### 2009-08-02 doc/latex/annotated.tex
documentation for serial, openmp, tbb

\section{Class List}
Here are the classes, structs, unions and interfaces with brief descriptions:\begin{CompactList}
\item\contentsline{section}{\hyperlink{class_cowichan}{Cowichan} (Base class for all C++ implementations )}{\pageref{class_cowichan}}{}
\item\contentsline{section}{\hyperlink{class_cowichan_open_m_p}{CowichanOpenMP} (Open Multi-Processing (OpenMP) implementation )}{\pageref{class_cowichan_open_m_p}}{}
\item\contentsline{section}{\hyperlink{class_cowichan_serial}{CowichanSerial} (Serial implementation )}{\pageref{class_cowichan_serial}}{}
\item\contentsline{section}{\hyperlink{class_cowichan_t_b_b}{CowichanTBB} (Thread Building Blocks (TBB) implementation )}{\pageref{class_cowichan_t_b_b}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_game_of_life}{cowichan\_\-tbb::GameOfLife} (Ping-pong solution to game of life )}{\pageref{classcowichan__tbb_1_1_game_of_life}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_histogram}{cowichan\_\-tbb::Histogram} (This class calculates the histogram of a matrix )}{\pageref{classcowichan__tbb_1_1_histogram}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_make_dominant}{cowichan\_\-tbb::MakeDominant} (Makes a given matrix diagonally dominant )}{\pageref{classcowichan__tbb_1_1_make_dominant}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_mandelbrot}{cowichan\_\-tbb::Mandelbrot} (Perform mandelbrot generation )}{\pageref{classcowichan__tbb_1_1_mandelbrot}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_maximum_distance}{cowichan\_\-tbb::MaximumDistance} (Compute maximum distance for hull )}{\pageref{classcowichan__tbb_1_1_maximum_distance}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_max_reducer}{cowichan\_\-tbb::MaxReducer} (Finds the maximum value in the image )}{\pageref{classcowichan__tbb_1_1_max_reducer}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_min_max_reducer}{cowichan\_\-tbb::MinMaxReducer} (Performs the minimum and maximum reductions )}{\pageref{classcowichan__tbb_1_1_min_max_reducer}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_normalizer}{cowichan\_\-tbb::Normalizer} (Normalizes points )}{\pageref{classcowichan__tbb_1_1_normalizer}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_norm_diff}{cowichan\_\-tbb::NormDiff} (This class takes the 1-norm of the difference between two vectors )}{\pageref{classcowichan__tbb_1_1_norm_diff}}{}
\item\contentsline{section}{\hyperlink{classcowichan__serial_1_1_perc_point}{cowichan\_\-serial::PercPoint} (Percolation point )}{\pageref{classcowichan__serial_1_1_perc_point}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_perc_point}{cowichan\_\-tbb::PercPoint} (Percolation point )}{\pageref{classcowichan__tbb_1_1_perc_point}}{}
\item\contentsline{section}{\hyperlink{classcowichan__openmp_1_1_perc_point}{cowichan\_\-openmp::PercPoint} (Percolation point )}{\pageref{classcowichan__openmp_1_1_perc_point}}{}
\item\contentsline{section}{\hyperlink{class_point}{Point} (Two-dimensional point )}{\pageref{class_point}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_point_count}{cowichan\_\-tbb::PointCount} (Calculates the number of points to sort for winnow )}{\pageref{classcowichan__tbb_1_1_point_count}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_point_distances}{cowichan\_\-tbb::PointDistances} (Fills in matrix and vector with distances )}{\pageref{classcowichan__tbb_1_1_point_distances}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_product}{cowichan\_\-tbb::Product} (Performs matrix vector product )}{\pageref{classcowichan__tbb_1_1_product}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_random_generator}{cowichan\_\-tbb::RandomGenerator} (Random generator )}{\pageref{classcowichan__tbb_1_1_random_generator}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_relaxer}{cowichan\_\-tbb::Relaxer} (Performs one iteration of relaxation )}{\pageref{classcowichan__tbb_1_1_relaxer}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_row_elimination}{cowichan\_\-tbb::RowElimination} (Performs row elimination, i.e. eliminate i-th column in j-th row )}{\pageref{classcowichan__tbb_1_1_row_elimination}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_shuffle}{cowichan\_\-tbb::Shuffle} (This class does a halving shuffle )}{\pageref{classcowichan__tbb_1_1_shuffle}}{}
\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_threshold}{cowichan\_\-tbb::Threshold} (Creates a boolean array based on a cut-off value )}{\pageref{classcowichan__tbb_1_1_threshold}}{}
\item\contentsline{section}{\hyperlink{class_weighted_point}{WeightedPoint} (Two-dimensional point with an integer weight )}{\pageref{class_weighted_point}}{}
\end{CompactList}

### 2009-08-02 doc/latex/class_cowichan.tex
documentation for serial, openmp, tbb

\hypertarget{class_cowichan}{
\section{Cowichan Class Reference}
\label{class_cowichan}\index{Cowichan@{Cowichan}}
}
Base class for all C++ implementations.

{\tt \#include $<$cowichan.hpp$>$}

Inheritance diagram for Cowichan::\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[height=2cm]{class_cowichan}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
{\footnotesize template$<$typename T $>$ }\\void \hyperlink{class_cowichan_14e706ef5a5f7661a623c595d7ed76f1}{print\_\-rect\_\-matrix} (T $\ast$matrix)
\item 
{\footnotesize template$<$typename T $>$ }\\void \hyperlink{class_cowichan_e8fdfb7dd3e8be0a7e9dbc531b1298d2}{print\_\-square\_\-matrix} (T $\ast$matrix)
\item 
{\footnotesize template$<$typename T $>$ }\\void \hyperlink{class_cowichan_90f1155f5c35308796b142b15d681a4b}{print\_\-vector} (T $\ast$vector)
\item 
void \hyperlink{class_cowichan_3b974ae1693ac661fb079b28981ca885}{print\_\-vector} (\hyperlink{class_point}{PointVector} points)
\item 
void \hyperlink{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]}, bool use\_\-randmat, bool use\_\-thresh)
\end{CompactItemize}
\subsection*{Protected Member Functions}
\begin{CompactItemize}
\item 
virtual void \hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{mandel} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)=0
\item 
virtual void \hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{randmat} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)=0
\item 
virtual void \hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{half} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixIn, \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixOut)=0
\item 
virtual void \hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{invperc} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)=0
\item 
virtual void \hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{thresh} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)=0
\item 
virtual void \hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{life} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixIn, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixOut)=0
\item 
virtual void \hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{winnow} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{class_point}{PointVector} points)=0
\item 
virtual void \hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{norm} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)=0
\item 
virtual void \hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{hull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)=0
\item 
virtual void \hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{outer} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector)=0
\item 
virtual void \hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{gauss} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)=0
\item 
virtual void \hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{sor} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)=0
\item 
virtual void \hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{product} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} candidate, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)=0
\item 
virtual \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{vecdiff} (\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} actual, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} computed)=0
\end{CompactItemize}
\subsection*{Protected Attributes}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_cowichan_44f139040042cc616e542be6faa28672}{nr}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_cowichan_81d2fc0da7b4ef388a1dbf08ac820eb1}{nc}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_cowichan_10e0c7c7bcccf11869821fa03cb7a6ee}{n}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_cowichan_edeb55a8b961f270ec35869dbba3afde}{lifeIterations}
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_f0bab1165abe53fea17e65764d247786}{mandelX0}
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_7b5e0e4e2026c78a6ec45fe306e5af93}{mandelY0}
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_edde2560c73a48dd44ea52c6a42c8649}{mandelDx}
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_7189d7127e740a2e6b14dcf0e757a500}{mandelDy}
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_a42c0ae639cc3ac124a013bc3e3162ad}{threshPercent}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_cowichan_c637b7380ab889ed42652790a3c542c6}{invpercNFill}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_cowichan_e9f8e9769f15e2648435064d4286b1a6}{seed}
\end{CompactItemize}
\subsection*{Static Protected Attributes}
\begin{CompactItemize}
\item 
static const char $\ast$ \hyperlink{class_cowichan_17937f5dfe24c46c00a76d48ca70ccaa}{CHAIN} = \char`\"{}chain\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_40d3f57ec83c7c1ec3af94c37d849049}{MANDEL} = \char`\"{}mandel\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_c9e66484bf3be07bfeac5511f514f102}{RANDMAT} = \char`\"{}randmat\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_6ef0b7dddca7331c9bfb2acd0ea60088}{HALF} = \char`\"{}half\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_0665b5094fb551e4f000502ec752a4e8}{INVPERC} = \char`\"{}invperc\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_e7adc012eb484f209b869db5062dacd0}{THRESH} = \char`\"{}thresh\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_6d8355328c9a2e36321d3d33930f56b2}{LIFE} = \char`\"{}life\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_a4d4acfeeb50c0552487a59a58b3f962}{WINNOW} = \char`\"{}winnow\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_4e2cd818307d1e806861489216e74a0f}{NORM} = \char`\"{}norm\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_c0ca862afefd5d6ab8fefdbf47428b21}{HULL} = \char`\"{}hull\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_b858724b8531751af4f611b978998be4}{OUTER} = \char`\"{}outer\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_b51b18cab3bd7cab0f3eb5c10f4cce0a}{GAUSS} = \char`\"{}gauss\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_eaac49e2ebb0c5149506f0544cd87794}{SOR} = \char`\"{}sor\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_57a1deea813a95ad7ed5a46dd000e576}{PRODUCT} = \char`\"{}product\char`\"{}
\item 
static const char $\ast$ \hyperlink{class_cowichan_d70f851efe3d20a89d7bcf5e7db70344}{VECDIFF} = \char`\"{}vecdiff\char`\"{}
\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{class_cowichan_08ee88eb612571e8faa20b2645dc906d}{chain} (bool use\_\-randmat, bool use\_\-thresh)
\end{CompactItemize}

\subsection{Detailed Description}
Base class for all C++ implementations.

The abstract class containing the inputs, definitions, and some debugging methods for the \hyperlink{class_cowichan}{Cowichan} problems.

\subsection{Member Function Documentation}
\hypertarget{class_cowichan_08ee88eb612571e8faa20b2645dc906d}{
\index{Cowichan@{Cowichan}!chain@{chain}}
\index{chain@{chain}!Cowichan@{Cowichan}}
\subsubsection[{chain}]{\setlength{\rightskip}{0pt plus 5cm}void Cowichan::chain (bool {\em use\_\-randmat}, \/  bool {\em use\_\-thresh})\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_cowichan_08ee88eb612571e8faa20b2645dc906d}

Runs the cowichan problem set, chained together. The order in the chain is: \begin{enumerate}
\item \hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{Cowichan::randmat} or \hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{Cowichan::mandel} \item \hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{Cowichan::half} \item \hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{Cowichan::invperc} or \hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{Cowichan::thresh} \item \hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{Cowichan::life} \item \hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{Cowichan::winnow} \item \hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{Cowichan::norm} \item \hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{Cowichan::hull} \item \hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{Cowichan::outer} \item \hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{Cowichan::gauss} \item \hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{Cowichan::sor} \item \hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan::product} (for gauss) \item \hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan::product} (for sor) \item \hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{Cowichan::vecdiff} \end{enumerate}
\begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em use\_\-randmat}]in step 1 use: randmat (if true) or mandel (if false). \item[{\em use\_\-thresh}]in step 3 use: thresh (if true) or invperc (if false). \end{description}
\end{Desc}
\hypertarget{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{
\index{Cowichan@{Cowichan}!gauss@{gauss}}
\index{gauss@{gauss}!Cowichan@{Cowichan}}
\subsubsection[{gauss}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::gauss ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}

For description see \hyperlink{index_gauss_sec}{11. Gaussian Elimination} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_b196d2384c2fe8b185e6269a050399eb}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_363bcd6f0c7b4d0fced94fe4cd59a267}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{CowichanTBB}.\hypertarget{class_cowichan_308603053675bccbe631f04af921f57c}{
\index{Cowichan@{Cowichan}!half@{half}}
\index{half@{half}!Cowichan@{Cowichan}}
\subsubsection[{half}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::half ({\bf IntMatrix} {\em matrixIn}, \/  {\bf IntMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_308603053675bccbe631f04af921f57c}

For description see \hyperlink{index_half_sec}{3. Two-Dimensional Shuffle} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]matrix to shuffle. \item[{\em matrixOut}]shuffled matrix. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_c08a6b3e23ce26959bac12af077f924f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_70989ffe182aebf590e39c56e146b0fb}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{CowichanTBB}.\hypertarget{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{
\index{Cowichan@{Cowichan}!hull@{hull}}
\index{hull@{hull}!Cowichan@{Cowichan}}
\subsubsection[{hull}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::hull ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}

For description see \hyperlink{index_hull_sec}{9. Convex Hull} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]points to run convex hull on. \item[{\em pointsOut}]reordered points from pointsIn. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{CowichanTBB}.\hypertarget{class_cowichan_ea126792a31e54a8722663b7ea768955}{
\index{Cowichan@{Cowichan}!invperc@{invperc}}
\index{invperc@{invperc}!Cowichan@{Cowichan}}
\subsubsection[{invperc}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::invperc ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_ea126792a31e54a8722663b7ea768955}

For description see \hyperlink{index_invperc_sec}{4. Invasion Percolation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]filling resistance matrix. \item[{\em mask}]filled cells. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4824c6b8509b5da835fbc5f64eb3e063}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{CowichanTBB}.\hypertarget{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{
\index{Cowichan@{Cowichan}!life@{life}}
\index{life@{life}!Cowichan@{Cowichan}}
\subsubsection[{life}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::life ({\bf BoolMatrix} {\em matrixIn}, \/  {\bf BoolMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}

For description see \hyperlink{index_life_sec}{6. Game of Life} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]initial world. \item[{\em matrixOut}]final world. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{CowichanTBB}.\hypertarget{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{
\index{Cowichan@{Cowichan}!main@{main}}
\index{main@{main}!Cowichan@{Cowichan}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}void Cowichan::main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]}, \/  bool {\em use\_\-randmat}, \/  bool {\em use\_\-thresh})}}
\label{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}

Runs cowichan problems based on command line input. Problem name can be specified on the command line. Otherwise, the chained version is run. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_08ee88eb612571e8faa20b2645dc906d}{Cowichan::chain} \end{Desc}
\begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em argc}]number of command line arguments. \item[{\em argv}]command line arguments. \item[{\em use\_\-randmat}]passed to chain if chained version is used. \item[{\em use\_\-thresh}]passed to chain if chained version is used. \end{description}
\end{Desc}
\hypertarget{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{
\index{Cowichan@{Cowichan}!mandel@{mandel}}
\index{mandel@{mandel}!Cowichan@{Cowichan}}
\subsubsection[{mandel}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::mandel ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}

For description see \hyperlink{index_mandel_sec}{1. Mandelbrot Set Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{CowichanTBB}.\hypertarget{class_cowichan_3df21e3c627958114e045c3559a29f30}{
\index{Cowichan@{Cowichan}!norm@{norm}}
\index{norm@{norm}!Cowichan@{Cowichan}}
\subsubsection[{norm}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::norm ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_3df21e3c627958114e045c3559a29f30}

For description see \hyperlink{index_norm_sec}{8. Point Location Normalization} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]points to normalize. \item[{\em pointsOut}]normalized points. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{CowichanTBB}.\hypertarget{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{
\index{Cowichan@{Cowichan}!outer@{outer}}
\index{outer@{outer}!Cowichan@{Cowichan}}
\subsubsection[{outer}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::outer ({\bf PointVector} {\em points}, \/  {\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em vector})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}

For description see \hyperlink{index_outer_sec}{10. Outer Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]vector of points. \item[{\em matrix}]resulting distance matrix. \item[{\em vector}]resulting distance vector. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_05f6899081a457d58978e4f6bda2db6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_99a0570f754ac82968d73093476de533}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{CowichanTBB}.\hypertarget{class_cowichan_14e706ef5a5f7661a623c595d7ed76f1}{
\index{Cowichan@{Cowichan}!print\_\-rect\_\-matrix@{print\_\-rect\_\-matrix}}
\index{print\_\-rect\_\-matrix@{print\_\-rect\_\-matrix}!Cowichan@{Cowichan}}
\subsubsection[{print\_\-rect\_\-matrix}]{\setlength{\rightskip}{0pt plus 5cm}template$<$typename T $>$ void Cowichan::print\_\-rect\_\-matrix (T $\ast$ {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_cowichan_14e706ef5a5f7661a623c595d7ed76f1}

DEBUGGING FUNCTION: Print a rectangular matrix on std::cout. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to print. \end{description}
\end{Desc}
\hypertarget{class_cowichan_e8fdfb7dd3e8be0a7e9dbc531b1298d2}{
\index{Cowichan@{Cowichan}!print\_\-square\_\-matrix@{print\_\-square\_\-matrix}}
\index{print\_\-square\_\-matrix@{print\_\-square\_\-matrix}!Cowichan@{Cowichan}}
\subsubsection[{print\_\-square\_\-matrix}]{\setlength{\rightskip}{0pt plus 5cm}template$<$typename T $>$ void Cowichan::print\_\-square\_\-matrix (T $\ast$ {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_cowichan_e8fdfb7dd3e8be0a7e9dbc531b1298d2}

DEBUGGING FUNCTION: Print a square matrix on std::cout. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to print. \end{description}
\end{Desc}
\hypertarget{class_cowichan_3b974ae1693ac661fb079b28981ca885}{
\index{Cowichan@{Cowichan}!print\_\-vector@{print\_\-vector}}
\index{print\_\-vector@{print\_\-vector}!Cowichan@{Cowichan}}
\subsubsection[{print\_\-vector}]{\setlength{\rightskip}{0pt plus 5cm}void Cowichan::print\_\-vector ({\bf PointVector} {\em points})}}
\label{class_cowichan_3b974ae1693ac661fb079b28981ca885}

DEBUGGING FUNCTION: Print a point vector. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]vector of points to print. \end{description}
\end{Desc}
\hypertarget{class_cowichan_90f1155f5c35308796b142b15d681a4b}{
\index{Cowichan@{Cowichan}!print\_\-vector@{print\_\-vector}}
\index{print\_\-vector@{print\_\-vector}!Cowichan@{Cowichan}}
\subsubsection[{print\_\-vector}]{\setlength{\rightskip}{0pt plus 5cm}template$<$typename T $>$ void Cowichan::print\_\-vector (T $\ast$ {\em vector})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_cowichan_90f1155f5c35308796b142b15d681a4b}

DEBUGGING FUNCTION: Print a vector on std::cout. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em vector}]vector to print. \end{description}
\end{Desc}
\hypertarget{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{
\index{Cowichan@{Cowichan}!product@{product}}
\index{product@{product}!Cowichan@{Cowichan}}
\subsubsection[{product}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::product ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em candidate}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}

For description see \hyperlink{index_product_sec}{13. Matrix-Vector Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em candidate}]vector X in AX = V. \item[{\em solution}]vector V in AX = V. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{CowichanTBB}.\hypertarget{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{
\index{Cowichan@{Cowichan}!randmat@{randmat}}
\index{randmat@{randmat}!Cowichan@{Cowichan}}
\subsubsection[{randmat}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::randmat ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}

For description see \hyperlink{index_randmat_sec}{2. Random Number Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{CowichanTBB}.\hypertarget{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{
\index{Cowichan@{Cowichan}!sor@{sor}}
\index{sor@{sor}!Cowichan@{Cowichan}}
\subsubsection[{sor}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::sor ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}

For description see \hyperlink{index_sor_sec}{12. Successive Over-Relaxation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_6e8b06711d976de1adc1e4dc81e560e5}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d6482d0369a26a51ef0e37ab238fc664}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{CowichanTBB}.\hypertarget{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{
\index{Cowichan@{Cowichan}!thresh@{thresh}}
\index{thresh@{thresh}!Cowichan@{Cowichan}}
\subsubsection[{thresh}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::thresh ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}

For description see \hyperlink{index_thresh_sec}{5. Histogram Thresholding} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]image. \item[{\em mask}]image after thresholding. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{CowichanTBB}.\hypertarget{class_cowichan_775d72b5e7d122f9f32555352278250e}{
\index{Cowichan@{Cowichan}!vecdiff@{vecdiff}}
\index{vecdiff@{vecdiff}!Cowichan@{Cowichan}}
\subsubsection[{vecdiff}]{\setlength{\rightskip}{0pt plus 5cm}virtual {\bf real} Cowichan::vecdiff ({\bf Vector} {\em actual}, \/  {\bf Vector} {\em computed})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_775d72b5e7d122f9f32555352278250e}

For description see \hyperlink{index_vecdiff_sec}{14. 1-Norm Vector Difference} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em actual}]first vector. \item[{\em computed}]second vector. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{CowichanTBB}.\hypertarget{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{
\index{Cowichan@{Cowichan}!winnow@{winnow}}
\index{winnow@{winnow}!Cowichan@{Cowichan}}
\subsubsection[{winnow}]{\setlength{\rightskip}{0pt plus 5cm}virtual void Cowichan::winnow ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask}, \/  {\bf PointVector} {\em points})\hspace{0.3cm}{\tt  \mbox{[}protected, pure virtual\mbox{]}}}}
\label{class_cowichan_13d60e06ced3b5da79d62c133ce82337}

For description see \hyperlink{index_winnow_sec}{7. Weighted Point Selection} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]integer matrix. \item[{\em mask}]boolean matrix. \item[{\em points}]evenly selected points. \end{description}
\end{Desc}

Implemented in \hyperlink{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{CowichanTBB}.

\subsection{Member Data Documentation}
\hypertarget{class_cowichan_17937f5dfe24c46c00a76d48ca70ccaa}{
\index{Cowichan@{Cowichan}!CHAIN@{CHAIN}}
\index{CHAIN@{CHAIN}!Cowichan@{Cowichan}}
\subsubsection[{CHAIN}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::CHAIN} = \char`\"{}chain\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_17937f5dfe24c46c00a76d48ca70ccaa}

Name for chain. \hypertarget{class_cowichan_b51b18cab3bd7cab0f3eb5c10f4cce0a}{
\index{Cowichan@{Cowichan}!GAUSS@{GAUSS}}
\index{GAUSS@{GAUSS}!Cowichan@{Cowichan}}
\subsubsection[{GAUSS}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::GAUSS} = \char`\"{}gauss\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_b51b18cab3bd7cab0f3eb5c10f4cce0a}

Name for gauss. \hypertarget{class_cowichan_6ef0b7dddca7331c9bfb2acd0ea60088}{
\index{Cowichan@{Cowichan}!HALF@{HALF}}
\index{HALF@{HALF}!Cowichan@{Cowichan}}
\subsubsection[{HALF}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::HALF} = \char`\"{}half\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_6ef0b7dddca7331c9bfb2acd0ea60088}

Name for half. \hypertarget{class_cowichan_c0ca862afefd5d6ab8fefdbf47428b21}{
\index{Cowichan@{Cowichan}!HULL@{HULL}}
\index{HULL@{HULL}!Cowichan@{Cowichan}}
\subsubsection[{HULL}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::HULL} = \char`\"{}hull\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_c0ca862afefd5d6ab8fefdbf47428b21}

Name for hull. \hypertarget{class_cowichan_0665b5094fb551e4f000502ec752a4e8}{
\index{Cowichan@{Cowichan}!INVPERC@{INVPERC}}
\index{INVPERC@{INVPERC}!Cowichan@{Cowichan}}
\subsubsection[{INVPERC}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::INVPERC} = \char`\"{}invperc\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_0665b5094fb551e4f000502ec752a4e8}

Name for invperc. \hypertarget{class_cowichan_c637b7380ab889ed42652790a3c542c6}{
\index{Cowichan@{Cowichan}!invpercNFill@{invpercNFill}}
\index{invpercNFill@{invpercNFill}!Cowichan@{Cowichan}}
\subsubsection[{invpercNFill}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf Cowichan::invpercNFill}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_c637b7380ab889ed42652790a3c542c6}

Number of cells to fill. \hypertarget{class_cowichan_6d8355328c9a2e36321d3d33930f56b2}{
\index{Cowichan@{Cowichan}!LIFE@{LIFE}}
\index{LIFE@{LIFE}!Cowichan@{Cowichan}}
\subsubsection[{LIFE}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::LIFE} = \char`\"{}life\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_6d8355328c9a2e36321d3d33930f56b2}

Name for life. \hypertarget{class_cowichan_edeb55a8b961f270ec35869dbba3afde}{
\index{Cowichan@{Cowichan}!lifeIterations@{lifeIterations}}
\index{lifeIterations@{lifeIterations}!Cowichan@{Cowichan}}
\subsubsection[{lifeIterations}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf Cowichan::lifeIterations}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_edeb55a8b961f270ec35869dbba3afde}

Number of iterations. \hypertarget{class_cowichan_40d3f57ec83c7c1ec3af94c37d849049}{
\index{Cowichan@{Cowichan}!MANDEL@{MANDEL}}
\index{MANDEL@{MANDEL}!Cowichan@{Cowichan}}
\subsubsection[{MANDEL}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::MANDEL} = \char`\"{}mandel\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_40d3f57ec83c7c1ec3af94c37d849049}

Name for mandel. \hypertarget{class_cowichan_edde2560c73a48dd44ea52c6a42c8649}{
\index{Cowichan@{Cowichan}!mandelDx@{mandelDx}}
\index{mandelDx@{mandelDx}!Cowichan@{Cowichan}}
\subsubsection[{mandelDx}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} {\bf Cowichan::mandelDx}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_edde2560c73a48dd44ea52c6a42c8649}

Extent of the region along the x axis. \hypertarget{class_cowichan_7189d7127e740a2e6b14dcf0e757a500}{
\index{Cowichan@{Cowichan}!mandelDy@{mandelDy}}
\index{mandelDy@{mandelDy}!Cowichan@{Cowichan}}
\subsubsection[{mandelDy}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} {\bf Cowichan::mandelDy}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_7189d7127e740a2e6b14dcf0e757a500}

Extent of the region along the y axis. \hypertarget{class_cowichan_f0bab1165abe53fea17e65764d247786}{
\index{Cowichan@{Cowichan}!mandelX0@{mandelX0}}
\index{mandelX0@{mandelX0}!Cowichan@{Cowichan}}
\subsubsection[{mandelX0}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} {\bf Cowichan::mandelX0}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_f0bab1165abe53fea17e65764d247786}

x-coordinate of the lower left corner. \hypertarget{class_cowichan_7b5e0e4e2026c78a6ec45fe306e5af93}{
\index{Cowichan@{Cowichan}!mandelY0@{mandelY0}}
\index{mandelY0@{mandelY0}!Cowichan@{Cowichan}}
\subsubsection[{mandelY0}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} {\bf Cowichan::mandelY0}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_7b5e0e4e2026c78a6ec45fe306e5af93}

y-coordinate of the lower left corner. \hypertarget{class_cowichan_10e0c7c7bcccf11869821fa03cb7a6ee}{
\index{Cowichan@{Cowichan}!n@{n}}
\index{n@{n}!Cowichan@{Cowichan}}
\subsubsection[{n}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf Cowichan::n}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_10e0c7c7bcccf11869821fa03cb7a6ee}

Number of rows/columns for square matrices. \hypertarget{class_cowichan_81d2fc0da7b4ef388a1dbf08ac820eb1}{
\index{Cowichan@{Cowichan}!nc@{nc}}
\index{nc@{nc}!Cowichan@{Cowichan}}
\subsubsection[{nc}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf Cowichan::nc}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_81d2fc0da7b4ef388a1dbf08ac820eb1}

Number of columns for rectangular matrices. \hypertarget{class_cowichan_4e2cd818307d1e806861489216e74a0f}{
\index{Cowichan@{Cowichan}!NORM@{NORM}}
\index{NORM@{NORM}!Cowichan@{Cowichan}}
\subsubsection[{NORM}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::NORM} = \char`\"{}norm\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_4e2cd818307d1e806861489216e74a0f}

Name for norm. \hypertarget{class_cowichan_44f139040042cc616e542be6faa28672}{
\index{Cowichan@{Cowichan}!nr@{nr}}
\index{nr@{nr}!Cowichan@{Cowichan}}
\subsubsection[{nr}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf Cowichan::nr}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_44f139040042cc616e542be6faa28672}

Number of rows for rectangular matrices. \hypertarget{class_cowichan_b858724b8531751af4f611b978998be4}{
\index{Cowichan@{Cowichan}!OUTER@{OUTER}}
\index{OUTER@{OUTER}!Cowichan@{Cowichan}}
\subsubsection[{OUTER}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::OUTER} = \char`\"{}outer\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_b858724b8531751af4f611b978998be4}

Name for outer. \hypertarget{class_cowichan_57a1deea813a95ad7ed5a46dd000e576}{
\index{Cowichan@{Cowichan}!PRODUCT@{PRODUCT}}
\index{PRODUCT@{PRODUCT}!Cowichan@{Cowichan}}
\subsubsection[{PRODUCT}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::PRODUCT} = \char`\"{}product\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_57a1deea813a95ad7ed5a46dd000e576}

Name for product. \hypertarget{class_cowichan_c9e66484bf3be07bfeac5511f514f102}{
\index{Cowichan@{Cowichan}!RANDMAT@{RANDMAT}}
\index{RANDMAT@{RANDMAT}!Cowichan@{Cowichan}}
\subsubsection[{RANDMAT}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::RANDMAT} = \char`\"{}randmat\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_c9e66484bf3be07bfeac5511f514f102}

Name for randmat. \hypertarget{class_cowichan_e9f8e9769f15e2648435064d4286b1a6}{
\index{Cowichan@{Cowichan}!seed@{seed}}
\index{seed@{seed}!Cowichan@{Cowichan}}
\subsubsection[{seed}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} {\bf Cowichan::seed}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_e9f8e9769f15e2648435064d4286b1a6}

Seed value for simple random number generator. \hypertarget{class_cowichan_eaac49e2ebb0c5149506f0544cd87794}{
\index{Cowichan@{Cowichan}!SOR@{SOR}}
\index{SOR@{SOR}!Cowichan@{Cowichan}}
\subsubsection[{SOR}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::SOR} = \char`\"{}sor\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_eaac49e2ebb0c5149506f0544cd87794}

Name for sor. \hypertarget{class_cowichan_e7adc012eb484f209b869db5062dacd0}{
\index{Cowichan@{Cowichan}!THRESH@{THRESH}}
\index{THRESH@{THRESH}!Cowichan@{Cowichan}}
\subsubsection[{THRESH}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::THRESH} = \char`\"{}thresh\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_e7adc012eb484f209b869db5062dacd0}

Name for thresh. \hypertarget{class_cowichan_a42c0ae639cc3ac124a013bc3e3162ad}{
\index{Cowichan@{Cowichan}!threshPercent@{threshPercent}}
\index{threshPercent@{threshPercent}!Cowichan@{Cowichan}}
\subsubsection[{threshPercent}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} {\bf Cowichan::threshPercent}\hspace{0.3cm}{\tt  \mbox{[}protected\mbox{]}}}}
\label{class_cowichan_a42c0ae639cc3ac124a013bc3e3162ad}

Thresholding percentage. \hypertarget{class_cowichan_d70f851efe3d20a89d7bcf5e7db70344}{
\index{Cowichan@{Cowichan}!VECDIFF@{VECDIFF}}
\index{VECDIFF@{VECDIFF}!Cowichan@{Cowichan}}
\subsubsection[{VECDIFF}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::VECDIFF} = \char`\"{}vecdiff\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_d70f851efe3d20a89d7bcf5e7db70344}

Name for vecdiff. \hypertarget{class_cowichan_a4d4acfeeb50c0552487a59a58b3f962}{
\index{Cowichan@{Cowichan}!WINNOW@{WINNOW}}
\index{WINNOW@{WINNOW}!Cowichan@{Cowichan}}
\subsubsection[{WINNOW}]{\setlength{\rightskip}{0pt plus 5cm}const char $\ast$ {\bf Cowichan::WINNOW} = \char`\"{}winnow\char`\"{}\hspace{0.3cm}{\tt  \mbox{[}static, protected\mbox{]}}}}
\label{class_cowichan_a4d4acfeeb50c0552487a59a58b3f962}

Name for winnow.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
cowichan/\hyperlink{cowichan_8hpp}{cowichan.hpp}\item 
cowichan/\hyperlink{cowichan_8cpp}{cowichan.cpp}\end{CompactItemize}

### 2009-08-02 doc/latex/class_cowichan_serial.tex
documentation for serial, openmp, tbb

\hypertarget{class_cowichan_serial}{
\section{CowichanSerial Class Reference}
\label{class_cowichan_serial}\index{CowichanSerial@{CowichanSerial}}
}
Serial implementation.

{\tt \#include $<$cowichan\_\-serial.hpp$>$}

Inheritance diagram for CowichanSerial::\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[height=2cm]{class_cowichan_serial}
\end{center}
\end{figure}
\subsection*{Protected Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{mandel} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{randmat} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_serial_c08a6b3e23ce26959bac12af077f924f}{half} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixIn, \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{invperc} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{thresh} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{life} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixIn, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{winnow} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{class_point}{PointVector} points)
\item 
void \hyperlink{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}{norm} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}{hull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_serial_05f6899081a457d58978e4f6bda2db6a}{outer} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector)
\item 
void \hyperlink{class_cowichan_serial_b196d2384c2fe8b185e6269a050399eb}{gauss} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_serial_6e8b06711d976de1adc1e4dc81e560e5}{sor} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{product} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} candidate, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{vecdiff} (\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} actual, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} computed)
\end{CompactItemize}

\subsection{Detailed Description}
Serial implementation.

This implementation can serve as a reference when comparing to parallel programming systems in terms of performance, code size, complexity, etc.

\subsection{Member Function Documentation}
\hypertarget{class_cowichan_serial_b196d2384c2fe8b185e6269a050399eb}{
\index{CowichanSerial@{CowichanSerial}!gauss@{gauss}}
\index{gauss@{gauss}!CowichanSerial@{CowichanSerial}}
\subsubsection[{gauss}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::gauss ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_b196d2384c2fe8b185e6269a050399eb}

For description see \hyperlink{index_gauss_sec}{11. Gaussian Elimination} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{Cowichan}.\hypertarget{class_cowichan_serial_c08a6b3e23ce26959bac12af077f924f}{
\index{CowichanSerial@{CowichanSerial}!half@{half}}
\index{half@{half}!CowichanSerial@{CowichanSerial}}
\subsubsection[{half}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::half ({\bf IntMatrix} {\em matrixIn}, \/  {\bf IntMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_c08a6b3e23ce26959bac12af077f924f}

For description see \hyperlink{index_half_sec}{3. Two-Dimensional Shuffle} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]matrix to shuffle. \item[{\em matrixOut}]shuffled matrix. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{Cowichan}.\hypertarget{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}{
\index{CowichanSerial@{CowichanSerial}!hull@{hull}}
\index{hull@{hull}!CowichanSerial@{CowichanSerial}}
\subsubsection[{hull}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::hull ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}

Runs quickhull algorithm until all points have been used up from the original vector. At each step the hull points are marked as used and a new convex hull is computed on the rest of points. The points that have been used up are in the range (n - hn, n), i.e. at the end of pointsIn vector. NOTE: pointsIn vector gets modified by the algorithm.

Implements \hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{Cowichan}.\hypertarget{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{
\index{CowichanSerial@{CowichanSerial}!invperc@{invperc}}
\index{invperc@{invperc}!CowichanSerial@{CowichanSerial}}
\subsubsection[{invperc}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::invperc ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}

For description see \hyperlink{index_invperc_sec}{4. Invasion Percolation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]filling resistance matrix. \item[{\em mask}]filled cells. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{Cowichan}.\hypertarget{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{
\index{CowichanSerial@{CowichanSerial}!life@{life}}
\index{life@{life}!CowichanSerial@{CowichanSerial}}
\subsubsection[{life}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::life ({\bf BoolMatrix} {\em matrixIn}, \/  {\bf BoolMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}

For description see \hyperlink{index_life_sec}{6. Game of Life} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]initial world. \item[{\em matrixOut}]final world. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{Cowichan}.\hypertarget{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{
\index{CowichanSerial@{CowichanSerial}!mandel@{mandel}}
\index{mandel@{mandel}!CowichanSerial@{CowichanSerial}}
\subsubsection[{mandel}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::mandel ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}

For description see \hyperlink{index_mandel_sec}{1. Mandelbrot Set Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{Cowichan}.\hypertarget{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}{
\index{CowichanSerial@{CowichanSerial}!norm@{norm}}
\index{norm@{norm}!CowichanSerial@{CowichanSerial}}
\subsubsection[{norm}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::norm ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}

For description see \hyperlink{index_norm_sec}{8. Point Location Normalization} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]points to normalize. \item[{\em pointsOut}]normalized points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{Cowichan}.\hypertarget{class_cowichan_serial_05f6899081a457d58978e4f6bda2db6a}{
\index{CowichanSerial@{CowichanSerial}!outer@{outer}}
\index{outer@{outer}!CowichanSerial@{CowichanSerial}}
\subsubsection[{outer}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::outer ({\bf PointVector} {\em points}, \/  {\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em vector})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_05f6899081a457d58978e4f6bda2db6a}

For description see \hyperlink{index_outer_sec}{10. Outer Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]vector of points. \item[{\em matrix}]resulting distance matrix. \item[{\em vector}]resulting distance vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{Cowichan}.\hypertarget{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{
\index{CowichanSerial@{CowichanSerial}!product@{product}}
\index{product@{product}!CowichanSerial@{CowichanSerial}}
\subsubsection[{product}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::product ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em candidate}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}

For description see \hyperlink{index_product_sec}{13. Matrix-Vector Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em candidate}]vector X in AX = V. \item[{\em solution}]vector V in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan}.\hypertarget{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{
\index{CowichanSerial@{CowichanSerial}!randmat@{randmat}}
\index{randmat@{randmat}!CowichanSerial@{CowichanSerial}}
\subsubsection[{randmat}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::randmat ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}

For description see \hyperlink{index_randmat_sec}{2. Random Number Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{Cowichan}.\hypertarget{class_cowichan_serial_6e8b06711d976de1adc1e4dc81e560e5}{
\index{CowichanSerial@{CowichanSerial}!sor@{sor}}
\index{sor@{sor}!CowichanSerial@{CowichanSerial}}
\subsubsection[{sor}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::sor ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_6e8b06711d976de1adc1e4dc81e560e5}

For description see \hyperlink{index_sor_sec}{12. Successive Over-Relaxation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{Cowichan}.\hypertarget{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{
\index{CowichanSerial@{CowichanSerial}!thresh@{thresh}}
\index{thresh@{thresh}!CowichanSerial@{CowichanSerial}}
\subsubsection[{thresh}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::thresh ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}

For description see \hyperlink{index_thresh_sec}{5. Histogram Thresholding} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]image. \item[{\em mask}]image after thresholding. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{Cowichan}.\hypertarget{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{
\index{CowichanSerial@{CowichanSerial}!vecdiff@{vecdiff}}
\index{vecdiff@{vecdiff}!CowichanSerial@{CowichanSerial}}
\subsubsection[{vecdiff}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} CowichanSerial::vecdiff ({\bf Vector} {\em actual}, \/  {\bf Vector} {\em computed})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}

For description see \hyperlink{index_vecdiff_sec}{14. 1-Norm Vector Difference} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em actual}]first vector. \item[{\em computed}]second vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{Cowichan}.\hypertarget{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{
\index{CowichanSerial@{CowichanSerial}!winnow@{winnow}}
\index{winnow@{winnow}!CowichanSerial@{CowichanSerial}}
\subsubsection[{winnow}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanSerial::winnow ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask}, \/  {\bf PointVector} {\em points})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}

For description see \hyperlink{index_winnow_sec}{7. Weighted Point Selection} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]integer matrix. \item[{\em mask}]boolean matrix. \item[{\em points}]evenly selected points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{Cowichan}.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_8hpp}{cowichan\_\-serial.hpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2gauss_8cpp}{gauss.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2half_8cpp}{half.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2hull_8cpp}{hull.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2invperc_8cpp}{invperc.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2life_8cpp}{life.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2mandel_8cpp}{mandel.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2norm_8cpp}{norm.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2outer_8cpp}{outer.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2product_8cpp}{product.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2randmat_8cpp}{randmat.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2sor_8cpp}{sor.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2thresh_8cpp}{thresh.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2vecdiff_8cpp}{vecdiff.cpp}\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2winnow_8cpp}{winnow.cpp}\end{CompactItemize}

### 2009-08-02 doc/latex/class_cowichan_t_b_b.tex
documentation for serial, openmp, tbb

\hypertarget{class_cowichan_t_b_b}{
\section{CowichanTBB Class Reference}
\label{class_cowichan_t_b_b}\index{CowichanTBB@{CowichanTBB}}
}
Thread Building Blocks (TBB) implementation.

{\tt \#include $<$cowichan\_\-tbb.hpp$>$}

Inheritance diagram for CowichanTBB::\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[height=2cm]{class_cowichan_t_b_b}
\end{center}
\end{figure}
\subsection*{Protected Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{mandel} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{randmat} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{half} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixIn, \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{invperc} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{thresh} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{life} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixIn, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{winnow} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{class_point}{PointVector} points)
\item 
void \hyperlink{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{norm} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{hull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{outer} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector)
\item 
void \hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{gauss} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{sor} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{product} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} candidate, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{vecdiff} (\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} actual, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} computed)
\end{CompactItemize}

\subsection{Detailed Description}
Thread Building Blocks (TBB) implementation.

Tags: shared memory, data parallel, task based.

\subsection{Member Function Documentation}
\hypertarget{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{
\index{CowichanTBB@{CowichanTBB}!gauss@{gauss}}
\index{gauss@{gauss}!CowichanTBB@{CowichanTBB}}
\subsubsection[{gauss}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::gauss ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}

For description see \hyperlink{index_gauss_sec}{11. Gaussian Elimination} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{Cowichan}.\hypertarget{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{
\index{CowichanTBB@{CowichanTBB}!half@{half}}
\index{half@{half}!CowichanTBB@{CowichanTBB}}
\subsubsection[{half}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::half ({\bf IntMatrix} {\em matrixIn}, \/  {\bf IntMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}

For description see \hyperlink{index_half_sec}{3. Two-Dimensional Shuffle} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]matrix to shuffle. \item[{\em matrixOut}]shuffled matrix. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{Cowichan}.\hypertarget{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{
\index{CowichanTBB@{CowichanTBB}!hull@{hull}}
\index{hull@{hull}!CowichanTBB@{CowichanTBB}}
\subsubsection[{hull}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::hull ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}

For description see \hyperlink{index_hull_sec}{9. Convex Hull}

Runs quickhull algorithm until all points have been used up from the original vector. At each step the hull points are marked as used and a new convex hull is computed on the rest of points. The points that have been used up are in the range (n - hn, n), i.e. at the end of pointsIn vector. NOTE: pointsIn vector gets modified by the algorithm.

Implements \hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{Cowichan}.\hypertarget{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{
\index{CowichanTBB@{CowichanTBB}!invperc@{invperc}}
\index{invperc@{invperc}!CowichanTBB@{CowichanTBB}}
\subsubsection[{invperc}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::invperc ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}

For description see \hyperlink{index_invperc_sec}{4. Invasion Percolation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]filling resistance matrix. \item[{\em mask}]filled cells. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{Cowichan}.\hypertarget{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{
\index{CowichanTBB@{CowichanTBB}!life@{life}}
\index{life@{life}!CowichanTBB@{CowichanTBB}}
\subsubsection[{life}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::life ({\bf BoolMatrix} {\em matrixIn}, \/  {\bf BoolMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}

For description see \hyperlink{index_life_sec}{6. Game of Life} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]initial world. \item[{\em matrixOut}]final world. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{Cowichan}.\hypertarget{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{
\index{CowichanTBB@{CowichanTBB}!mandel@{mandel}}
\index{mandel@{mandel}!CowichanTBB@{CowichanTBB}}
\subsubsection[{mandel}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::mandel ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}

For description see \hyperlink{index_mandel_sec}{1. Mandelbrot Set Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{Cowichan}.\hypertarget{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{
\index{CowichanTBB@{CowichanTBB}!norm@{norm}}
\index{norm@{norm}!CowichanTBB@{CowichanTBB}}
\subsubsection[{norm}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::norm ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}

For description see \hyperlink{index_norm_sec}{8. Point Location Normalization} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]points to normalize. \item[{\em pointsOut}]normalized points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{Cowichan}.\hypertarget{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{
\index{CowichanTBB@{CowichanTBB}!outer@{outer}}
\index{outer@{outer}!CowichanTBB@{CowichanTBB}}
\subsubsection[{outer}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::outer ({\bf PointVector} {\em points}, \/  {\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em vector})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}

For description see \hyperlink{index_outer_sec}{10. Outer Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]vector of points. \item[{\em matrix}]resulting distance matrix. \item[{\em vector}]resulting distance vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{Cowichan}.\hypertarget{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{
\index{CowichanTBB@{CowichanTBB}!product@{product}}
\index{product@{product}!CowichanTBB@{CowichanTBB}}
\subsubsection[{product}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::product ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em candidate}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}

For description see \hyperlink{index_product_sec}{13. Matrix-Vector Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em candidate}]vector X in AX = V. \item[{\em solution}]vector V in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan}.\hypertarget{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{
\index{CowichanTBB@{CowichanTBB}!randmat@{randmat}}
\index{randmat@{randmat}!CowichanTBB@{CowichanTBB}}
\subsubsection[{randmat}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::randmat ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}

For description see \hyperlink{index_randmat_sec}{2. Random Number Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{Cowichan}.\hypertarget{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{
\index{CowichanTBB@{CowichanTBB}!sor@{sor}}
\index{sor@{sor}!CowichanTBB@{CowichanTBB}}
\subsubsection[{sor}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::sor ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}

For description see \hyperlink{index_sor_sec}{12. Successive Over-Relaxation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{Cowichan}.\hypertarget{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{
\index{CowichanTBB@{CowichanTBB}!thresh@{thresh}}
\index{thresh@{thresh}!CowichanTBB@{CowichanTBB}}
\subsubsection[{thresh}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::thresh ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}

For description see \hyperlink{index_thresh_sec}{5. Histogram Thresholding} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]image. \item[{\em mask}]image after thresholding. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{Cowichan}.\hypertarget{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{
\index{CowichanTBB@{CowichanTBB}!vecdiff@{vecdiff}}
\index{vecdiff@{vecdiff}!CowichanTBB@{CowichanTBB}}
\subsubsection[{vecdiff}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} CowichanTBB::vecdiff ({\bf Vector} {\em actual}, \/  {\bf Vector} {\em computed})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}

For description see \hyperlink{index_vecdiff_sec}{14. 1-Norm Vector Difference} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em actual}]first vector. \item[{\em computed}]second vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{Cowichan}.\hypertarget{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{
\index{CowichanTBB@{CowichanTBB}!winnow@{winnow}}
\index{winnow@{winnow}!CowichanTBB@{CowichanTBB}}
\subsubsection[{winnow}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanTBB::winnow ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask}, \/  {\bf PointVector} {\em points})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}

For description see \hyperlink{index_winnow_sec}{7. Weighted Point Selection} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]integer matrix. \item[{\em mask}]boolean matrix. \item[{\em points}]evenly selected points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{Cowichan}.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_8hpp}{cowichan\_\-tbb.hpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2gauss_8cpp}{gauss.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2half_8cpp}{half.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2hull_8cpp}{hull.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2invperc_8cpp}{invperc.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2life_8cpp}{life.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2mandel_8cpp}{mandel.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2norm_8cpp}{norm.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2outer_8cpp}{outer.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2product_8cpp}{product.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2randmat_8cpp}{randmat.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2sor_8cpp}{sor.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2thresh_8cpp}{thresh.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2vecdiff_8cpp}{vecdiff.cpp}\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2winnow_8cpp}{winnow.cpp}\end{CompactItemize}

### 2009-08-02 doc/latex/class_game_of_life.tex
documentation for serial, openmp, tbb

\hypertarget{class_game_of_life}{
\section{GameOfLife Class Reference}
\label{class_game_of_life}\index{GameOfLife@{GameOfLife}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_game_of_life_7de73dfe2060ec6d0a5bab65cf771f63}{
\textbf{GameOfLife} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} first, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} second, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)}
\label{class_game_of_life_7de73dfe2060ec6d0a5bab65cf771f63}

\item 
\hypertarget{class_game_of_life_3027e63b5ad8a3263023f3590caf1d2e}{
void \textbf{swap} ()}
\label{class_game_of_life_3027e63b5ad8a3263023f3590caf1d2e}

\item 
void \hyperlink{class_game_of_life_dfa4429cd90a9a8a7539ecb5a4617ae8}{operator()} (const \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D} \&range) const 
\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_game_of_life_df162a0c6062460fa658ae9b9cd457b6}{sumNeighbours} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} x, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} y) const 
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_game_of_life_8bdab545960821d63692c05b875199db}{
\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} \textbf{\_\-first}}
\label{class_game_of_life_8bdab545960821d63692c05b875199db}

\item 
\hypertarget{class_game_of_life_d772ef5614e6b178fb8d653589f1f1f9}{
\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} \textbf{\_\-second}}
\label{class_game_of_life_d772ef5614e6b178fb8d653589f1f1f9}

\item 
\hypertarget{class_game_of_life_5a16e4e2b0d83c5e3f0e84cc3ff4db5f}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nr}}
\label{class_game_of_life_5a16e4e2b0d83c5e3f0e84cc3ff4db5f}

\item 
\hypertarget{class_game_of_life_eda1db7eabff4386e01714cf621b072d}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nc}}
\label{class_game_of_life_eda1db7eabff4386e01714cf621b072d}

\end{CompactItemize}

\subsection{Detailed Description}
This class does the game of life, and facilitates a ping-pong memory model.

\subsection{Member Function Documentation}
\hypertarget{class_game_of_life_dfa4429cd90a9a8a7539ecb5a4617ae8}{
\index{GameOfLife@{GameOfLife}!operator()@{operator()}}
\index{operator()@{operator()}!GameOfLife@{GameOfLife}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void GameOfLife::operator() (const {\bf Range2D} \& {\em range}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_game_of_life_dfa4429cd90a9a8a7539ecb5a4617ae8}

Performs the game of life operation over the given range. \hypertarget{class_game_of_life_df162a0c6062460fa658ae9b9cd457b6}{
\index{GameOfLife@{GameOfLife}!sumNeighbours@{sumNeighbours}}
\index{sumNeighbours@{sumNeighbours}!GameOfLife@{GameOfLife}}
\subsubsection[{sumNeighbours}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} GameOfLife::sumNeighbours ({\bf index\_\-t} {\em x}, \/  {\bf index\_\-t} {\em y}) const\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{class_game_of_life_df162a0c6062460fa658ae9b9cd457b6}

Calculate number of peers.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/life.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_make_dominant.tex
documentation for serial, openmp, tbb

\hypertarget{class_make_dominant}{
\section{MakeDominant Class Reference}
\label{class_make_dominant}\index{MakeDominant@{MakeDominant}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_make_dominant_f6d79453b946ddd294f9c4d4109fbb5c}{
\textbf{MakeDominant} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} value)}
\label{class_make_dominant_f6d79453b946ddd294f9c4d4109fbb5c}

\item 
void \hyperlink{class_make_dominant_79c1af1cd1e60a39f5385409209c6cef}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows) const 
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_make_dominant_73325f3005d8887ca20b10fd2cdbdb0b}{
\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} \textbf{\_\-matrix}}
\label{class_make_dominant_73325f3005d8887ca20b10fd2cdbdb0b}

\item 
\hypertarget{class_make_dominant_770833546d40345b9408226ffbcc6417}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{n}}
\label{class_make_dominant_770833546d40345b9408226ffbcc6417}

\item 
\hypertarget{class_make_dominant_b800551decfb81771f83ab869b459cd8}{
const \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{value}}
\label{class_make_dominant_b800551decfb81771f83ab869b459cd8}

\end{CompactItemize}

\subsection{Detailed Description}
Makes a given matrix diagonally dominant by modifying its diagonal elements.

\subsection{Member Function Documentation}
\hypertarget{class_make_dominant_79c1af1cd1e60a39f5385409209c6cef}{
\index{MakeDominant@{MakeDominant}!operator()@{operator()}}
\index{operator()@{operator()}!MakeDominant@{MakeDominant}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void MakeDominant::operator() (const {\bf Range} \& {\em rows}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_make_dominant_79c1af1cd1e60a39f5385409209c6cef}

Sets diagonal elements to a given constant.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/outer.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_max_reducer.tex
documentation for serial, openmp, tbb

\hypertarget{class_max_reducer}{
\section{MaxReducer Class Reference}
\label{class_max_reducer}\index{MaxReducer@{MaxReducer}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{class_max_reducer_15429c52ea4767f8c4ed60530814a924}{MaxReducer} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} image, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\item 
\hypertarget{class_max_reducer_de8cd5f464b4287fab14fdbe7b701452}{
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \textbf{getMaximum} () const }
\label{class_max_reducer_de8cd5f464b4287fab14fdbe7b701452}

\item 
void \hyperlink{class_max_reducer_80d53943773b8909921c53a2f094d078}{operator()} (const \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D} \&range)
\item 
\hyperlink{class_max_reducer_7c2c70197b7e05eb9ecacf0db0b8b3b4}{MaxReducer} (\hyperlink{class_max_reducer}{MaxReducer} \&other, split)
\item 
void \hyperlink{class_max_reducer_095078ad7a356586f27a79a97fc40ac9}{join} (const \hyperlink{class_max_reducer}{MaxReducer} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_max_reducer_90f7cb1f4a7460c19caf4419a0ae1c06}{
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \textbf{\_\-image}}
\label{class_max_reducer_90f7cb1f4a7460c19caf4419a0ae1c06}

\item 
\hypertarget{class_max_reducer_8635ce5b2d81812d4a7ddc051c961886}{
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \textbf{\_\-max}}
\label{class_max_reducer_8635ce5b2d81812d4a7ddc051c961886}

\item 
\hypertarget{class_max_reducer_c640e0798bc41ef929aa907cc94cec48}{
const \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nc}}
\label{class_max_reducer_c640e0798bc41ef929aa907cc94cec48}

\end{CompactItemize}

\subsection{Detailed Description}
Performs the maximum computation.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_max_reducer_15429c52ea4767f8c4ed60530814a924}{
\index{MaxReducer@{MaxReducer}!MaxReducer@{MaxReducer}}
\index{MaxReducer@{MaxReducer}!MaxReducer@{MaxReducer}}
\subsubsection[{MaxReducer}]{\setlength{\rightskip}{0pt plus 5cm}MaxReducer::MaxReducer ({\bf IntMatrix} {\em image}, \/  {\bf index\_\-t} {\em nc})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_max_reducer_15429c52ea4767f8c4ed60530814a924}

Initialise max with the lowest possible value. \hypertarget{class_max_reducer_7c2c70197b7e05eb9ecacf0db0b8b3b4}{
\index{MaxReducer@{MaxReducer}!MaxReducer@{MaxReducer}}
\index{MaxReducer@{MaxReducer}!MaxReducer@{MaxReducer}}
\subsubsection[{MaxReducer}]{\setlength{\rightskip}{0pt plus 5cm}MaxReducer::MaxReducer ({\bf MaxReducer} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_max_reducer_7c2c70197b7e05eb9ecacf0db0b8b3b4}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_max_reducer_095078ad7a356586f27a79a97fc40ac9}{
\index{MaxReducer@{MaxReducer}!join@{join}}
\index{join@{join}!MaxReducer@{MaxReducer}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void MaxReducer::join (const {\bf MaxReducer} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_max_reducer_095078ad7a356586f27a79a97fc40ac9}

Joiner (TBB). \hypertarget{class_max_reducer_80d53943773b8909921c53a2f094d078}{
\index{MaxReducer@{MaxReducer}!operator()@{operator()}}
\index{operator()@{operator()}!MaxReducer@{MaxReducer}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void MaxReducer::operator() (const {\bf Range2D} \& {\em range})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_max_reducer_80d53943773b8909921c53a2f094d078}

Calculates the maximum value over the given range.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/thresh.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_maximum_distance.tex
documentation for serial, openmp, tbb

\hypertarget{class_maximum_distance}{
\section{MaximumDistance Class Reference}
\label{class_maximum_distance}\index{MaximumDistance@{MaximumDistance}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_maximum_distance_53a46dbac3acecf2223f4966cd73cd69}{
\textbf{MaximumDistance} (const \hyperlink{class_point}{PointVector} \&points, const \hyperlink{class_point}{Point} \&p1, const \hyperlink{class_point}{Point} \&p2)}
\label{class_maximum_distance_53a46dbac3acecf2223f4966cd73cd69}

\item 
\hyperlink{class_point}{Point} $\ast$ \hyperlink{class_maximum_distance_8e1dae43c6ec3865ea79dd7caf27ad8f}{getPoint} () const 
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_maximum_distance_1a187dd3950b791434bb5223d6b16882}{getDistance} () const 
\item 
void \hyperlink{class_maximum_distance_69c66442a93a92b18266ba55909f6c15}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range)
\item 
\hyperlink{class_maximum_distance_1cfb6fc5edbc0495b2fe27304063b68e}{MaximumDistance} (\hyperlink{class_maximum_distance}{MaximumDistance} \&other, split)
\item 
void \hyperlink{class_maximum_distance_8e224531b8c5e9a9f9e506f3e674586b}{join} (const \hyperlink{class_maximum_distance}{MaximumDistance} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_maximum_distance_77b4e1cbd087d7f30f85884c1c9d02a4}{
const \hyperlink{class_point}{PointVector} \& \textbf{points}}
\label{class_maximum_distance_77b4e1cbd087d7f30f85884c1c9d02a4}

\item 
\hypertarget{class_maximum_distance_1ff10f979e3f1e837602e818ea0b8160}{
const \hyperlink{class_point}{Point} \& \textbf{p1}}
\label{class_maximum_distance_1ff10f979e3f1e837602e818ea0b8160}

\item 
\hypertarget{class_maximum_distance_1186a242927b7fc1a33fca0a6eb86006}{
const \hyperlink{class_point}{Point} \& \textbf{p2}}
\label{class_maximum_distance_1186a242927b7fc1a33fca0a6eb86006}

\item 
\hypertarget{class_maximum_distance_1f85a3af711d8c7608fb589d427c519f}{
\hyperlink{class_point}{Point} $\ast$ \textbf{maxPoint}}
\label{class_maximum_distance_1f85a3af711d8c7608fb589d427c519f}

\item 
\hypertarget{class_maximum_distance_b49089f3da83cc1a0e233ad6262ac6cc}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{maxCross}}
\label{class_maximum_distance_b49089f3da83cc1a0e233ad6262ac6cc}

\end{CompactItemize}

\subsection{Detailed Description}
Computes the maximum signed distance from a point to a line, given the line and list of points.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_maximum_distance_1cfb6fc5edbc0495b2fe27304063b68e}{
\index{MaximumDistance@{MaximumDistance}!MaximumDistance@{MaximumDistance}}
\index{MaximumDistance@{MaximumDistance}!MaximumDistance@{MaximumDistance}}
\subsubsection[{MaximumDistance}]{\setlength{\rightskip}{0pt plus 5cm}MaximumDistance::MaximumDistance ({\bf MaximumDistance} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_1cfb6fc5edbc0495b2fe27304063b68e}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_maximum_distance_1a187dd3950b791434bb5223d6b16882}{
\index{MaximumDistance@{MaximumDistance}!getDistance@{getDistance}}
\index{getDistance@{getDistance}!MaximumDistance@{MaximumDistance}}
\subsubsection[{getDistance}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} MaximumDistance::getDistance () const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_1a187dd3950b791434bb5223d6b16882}

Gets the signed distance to that point. \hypertarget{class_maximum_distance_8e1dae43c6ec3865ea79dd7caf27ad8f}{
\index{MaximumDistance@{MaximumDistance}!getPoint@{getPoint}}
\index{getPoint@{getPoint}!MaximumDistance@{MaximumDistance}}
\subsubsection[{getPoint}]{\setlength{\rightskip}{0pt plus 5cm}{\bf Point}$\ast$ MaximumDistance::getPoint () const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_8e1dae43c6ec3865ea79dd7caf27ad8f}

Gets the point with maximum signed distance (if it has already been calculated). \hypertarget{class_maximum_distance_8e224531b8c5e9a9f9e506f3e674586b}{
\index{MaximumDistance@{MaximumDistance}!join@{join}}
\index{join@{join}!MaximumDistance@{MaximumDistance}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void MaximumDistance::join (const {\bf MaximumDistance} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_8e224531b8c5e9a9f9e506f3e674586b}

Joiner (TBB). \hypertarget{class_maximum_distance_69c66442a93a92b18266ba55909f6c15}{
\index{MaximumDistance@{MaximumDistance}!operator()@{operator()}}
\index{operator()@{operator()}!MaximumDistance@{MaximumDistance}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void MaximumDistance::operator() (const {\bf Range} \& {\em range})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_maximum_distance_69c66442a93a92b18266ba55909f6c15}

Calculates the \hyperlink{class_point}{Point} with maximum signed distance.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/hull.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_min_max_reducer.tex
documentation for serial, openmp, tbb

\hypertarget{class_min_max_reducer}{
\section{MinMaxReducer Class Reference}
\label{class_min_max_reducer}\index{MinMaxReducer@{MinMaxReducer}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_min_max_reducer_21bb0b674efa16fe293f02e7ddac5baa}{
\textbf{MinMaxReducer} (\hyperlink{class_point}{PointVector} pointsIn)}
\label{class_min_max_reducer_21bb0b674efa16fe293f02e7ddac5baa}

\item 
\hypertarget{class_min_max_reducer_16c5224355e8fde2e50a8189fca1c2b5}{
\hyperlink{class_point}{Point} \textbf{getMinimum} () const }
\label{class_min_max_reducer_16c5224355e8fde2e50a8189fca1c2b5}

\item 
\hypertarget{class_min_max_reducer_af59d922da8dd11500e068f86e9e140a}{
\hyperlink{class_point}{Point} \textbf{getMaximum} () const }
\label{class_min_max_reducer_af59d922da8dd11500e068f86e9e140a}

\item 
void \hyperlink{class_min_max_reducer_966e6358c637bb3c663a3b6547170fd4}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range)
\item 
\hyperlink{class_min_max_reducer_282eda515485b150389c963f1024ce12}{MinMaxReducer} (\hyperlink{class_min_max_reducer}{MinMaxReducer} \&other, split)
\item 
void \hyperlink{class_min_max_reducer_d002be90fb5c5e273b9f0d5fcb1db988}{join} (const \hyperlink{class_min_max_reducer}{MinMaxReducer} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_min_max_reducer_850a77d5dc5ed32204e265cfc619abc2}{
\hyperlink{class_point}{PointVector} \textbf{pointsIn}}
\label{class_min_max_reducer_850a77d5dc5ed32204e265cfc619abc2}

\item 
\hypertarget{class_min_max_reducer_615c82d310905e78ae116be676357c46}{
\hyperlink{class_point}{Point} \textbf{minPoint}}
\label{class_min_max_reducer_615c82d310905e78ae116be676357c46}

\item 
\hypertarget{class_min_max_reducer_e096e17b34d0bc699a3ed349ed5798d8}{
\hyperlink{class_point}{Point} \textbf{maxPoint}}
\label{class_min_max_reducer_e096e17b34d0bc699a3ed349ed5798d8}

\end{CompactItemize}

\subsection{Detailed Description}
Performs the minimum and maximum computations.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_min_max_reducer_282eda515485b150389c963f1024ce12}{
\index{MinMaxReducer@{MinMaxReducer}!MinMaxReducer@{MinMaxReducer}}
\index{MinMaxReducer@{MinMaxReducer}!MinMaxReducer@{MinMaxReducer}}
\subsubsection[{MinMaxReducer}]{\setlength{\rightskip}{0pt plus 5cm}MinMaxReducer::MinMaxReducer ({\bf MinMaxReducer} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_min_max_reducer_282eda515485b150389c963f1024ce12}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_min_max_reducer_d002be90fb5c5e273b9f0d5fcb1db988}{
\index{MinMaxReducer@{MinMaxReducer}!join@{join}}
\index{join@{join}!MinMaxReducer@{MinMaxReducer}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void MinMaxReducer::join (const {\bf MinMaxReducer} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_min_max_reducer_d002be90fb5c5e273b9f0d5fcb1db988}

Joiner (TBB). \hypertarget{class_min_max_reducer_966e6358c637bb3c663a3b6547170fd4}{
\index{MinMaxReducer@{MinMaxReducer}!operator()@{operator()}}
\index{operator()@{operator()}!MinMaxReducer@{MinMaxReducer}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void MinMaxReducer::operator() (const {\bf Range} \& {\em range})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_min_max_reducer_966e6358c637bb3c663a3b6547170fd4}

Calculates the minimum and maximum co-ordinates over the given array range.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/norm.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_norm_diff.tex
documentation for serial, openmp, tbb

\hypertarget{class_norm_diff}{
\section{NormDiff Class Reference}
\label{class_norm_diff}\index{NormDiff@{NormDiff}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_norm_diff_47bd32e56a26272eaad47a4839448312}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{getNormDiff} () const }
\label{class_norm_diff_47bd32e56a26272eaad47a4839448312}

\item 
\hypertarget{class_norm_diff_2f6ff2bc440755f8ac2350b5185b5423}{
\textbf{NormDiff} (\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} actual, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} computed)}
\label{class_norm_diff_2f6ff2bc440755f8ac2350b5185b5423}

\item 
\hypertarget{class_norm_diff_44066147e706abbfce8e16fc92e0affc}{
void \textbf{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range)}
\label{class_norm_diff_44066147e706abbfce8e16fc92e0affc}

\item 
\hyperlink{class_norm_diff_edef38012c44577650ccb47a55177eee}{NormDiff} (\hyperlink{class_norm_diff}{NormDiff} \&other, split)
\item 
void \hyperlink{class_norm_diff_8a3ce83486dfa068fbc4d3601a6442b1}{join} (const \hyperlink{class_norm_diff}{NormDiff} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_norm_diff_f4dd514a2fb97d4e4742232b2d6f4930}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-actual}}
\label{class_norm_diff_f4dd514a2fb97d4e4742232b2d6f4930}

\item 
\hypertarget{class_norm_diff_c6815c7be114269afb88e6932a001496}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-computed}}
\label{class_norm_diff_c6815c7be114269afb88e6932a001496}

\item 
\hypertarget{class_norm_diff_99c21620f91d9f917bb0745a284275b3}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{normDiff}}
\label{class_norm_diff_99c21620f91d9f917bb0745a284275b3}

\end{CompactItemize}

\subsection{Detailed Description}
This class takes the 1-norm of the difference between two vectors.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_norm_diff_edef38012c44577650ccb47a55177eee}{
\index{NormDiff@{NormDiff}!NormDiff@{NormDiff}}
\index{NormDiff@{NormDiff}!NormDiff@{NormDiff}}
\subsubsection[{NormDiff}]{\setlength{\rightskip}{0pt plus 5cm}NormDiff::NormDiff ({\bf NormDiff} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_norm_diff_edef38012c44577650ccb47a55177eee}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_norm_diff_8a3ce83486dfa068fbc4d3601a6442b1}{
\index{NormDiff@{NormDiff}!join@{join}}
\index{join@{join}!NormDiff@{NormDiff}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void NormDiff::join (const {\bf NormDiff} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_norm_diff_8a3ce83486dfa068fbc4d3601a6442b1}

Joiner (TBB).

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/vecdiff.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_normalizer.tex
documentation for serial, openmp, tbb

\hypertarget{class_normalizer}{
\section{Normalizer Class Reference}
\label{class_normalizer}\index{Normalizer@{Normalizer}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_normalizer_56999f560190503e8f001bdc7ecc2b87}{
\textbf{Normalizer} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} minX, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} minY, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} xfactor, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} yfactor)}
\label{class_normalizer_56999f560190503e8f001bdc7ecc2b87}

\item 
\hypertarget{class_normalizer_f384e3a61919ce96f30b52248c384261}{
void \textbf{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range) const }
\label{class_normalizer_f384e3a61919ce96f30b52248c384261}

\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_normalizer_02f117411f07fae9fc034885e69f298b}{
\hyperlink{class_point}{PointVector} \textbf{pointsIn}}
\label{class_normalizer_02f117411f07fae9fc034885e69f298b}

\item 
\hypertarget{class_normalizer_5f731c7abc673530d5da15a5b73b03f3}{
\hyperlink{class_point}{PointVector} \textbf{pointsOut}}
\label{class_normalizer_5f731c7abc673530d5da15a5b73b03f3}

\item 
\hypertarget{class_normalizer_d928591add8f6af2fc4943b6a3800242}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{minX}}
\label{class_normalizer_d928591add8f6af2fc4943b6a3800242}

\item 
\hypertarget{class_normalizer_e154cea5c227b01fa743a18c20fc94bc}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{minY}}
\label{class_normalizer_e154cea5c227b01fa743a18c20fc94bc}

\item 
\hypertarget{class_normalizer_e25d5be008bd7b298b0e3c2a7969ed77}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{xfactor}}
\label{class_normalizer_e25d5be008bd7b298b0e3c2a7969ed77}

\item 
\hypertarget{class_normalizer_67a219d5d5e172dfca12b0a22bbe53a6}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{yfactor}}
\label{class_normalizer_67a219d5d5e172dfca12b0a22bbe53a6}

\end{CompactItemize}

\subsection{Detailed Description}
This class performs point normalization -- points are put onto the unit square.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/norm.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_perc_point.tex
documentation for serial, openmp, tbb

\hypertarget{class_perc_point}{
\section{PercPoint Class Reference}
\label{class_perc_point}\index{PercPoint@{PercPoint}}
}
Percolation point.

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}{PercPoint} (\hyperlink{class_point}{Point} \hyperlink{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{point})
\item 
bool \hyperlink{class_perc_point_f798aa389b770208bd172542267ec811}{operator$<$} (const \hyperlink{class_perc_point}{PercPoint} \&other) const 
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_perc_point_0099784a4e3728c1f28b546a5b94104a}{value} () const 
\item 
\hypertarget{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}{
\textbf{PercPoint} (\hyperlink{class_point}{Point} \hyperlink{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{point})}
\label{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}

\item 
\hypertarget{class_perc_point_f798aa389b770208bd172542267ec811}{
bool \textbf{operator$<$} (const \hyperlink{class_perc_point}{PercPoint} \&other) const }
\label{class_perc_point_f798aa389b770208bd172542267ec811}

\item 
\hypertarget{class_perc_point_0099784a4e3728c1f28b546a5b94104a}{
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \textbf{value} () const }
\label{class_perc_point_0099784a4e3728c1f28b546a5b94104a}

\item 
\hypertarget{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}{
\textbf{PercPoint} (\hyperlink{class_point}{Point} \hyperlink{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{point})}
\label{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}

\item 
\hypertarget{class_perc_point_f798aa389b770208bd172542267ec811}{
bool \textbf{operator$<$} (const \hyperlink{class_perc_point}{PercPoint} \&other) const }
\label{class_perc_point_f798aa389b770208bd172542267ec811}

\item 
\hypertarget{class_perc_point_0099784a4e3728c1f28b546a5b94104a}{
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \textbf{value} () const }
\label{class_perc_point_0099784a4e3728c1f28b546a5b94104a}

\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hyperlink{class_point}{Point} \hyperlink{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{point}
\item 
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \hyperlink{class_perc_point_730c461b6a04bcb883d60df9b297b5d5}{matrix}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_perc_point_dc46d54bf942be6e7774ae522e563f87}{nc}
\end{CompactItemize}

\subsection{Detailed Description}
Percolation point.

These are put into a heap to extract the one with lowest value.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}{
\index{PercPoint@{PercPoint}!PercPoint@{PercPoint}}
\index{PercPoint@{PercPoint}!PercPoint@{PercPoint}}
\subsubsection[{PercPoint}]{\setlength{\rightskip}{0pt plus 5cm}PercPoint::PercPoint ({\bf Point} {\em point})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_perc_point_52b0b83113727e5d43ac0d3db3125ad0}

Construct a percolation point. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em point}]point. \end{description}
\end{Desc}

\subsection{Member Function Documentation}
\hypertarget{class_perc_point_f798aa389b770208bd172542267ec811}{
\index{PercPoint@{PercPoint}!operator$<$@{operator$<$}}
\index{operator$<$@{operator$<$}!PercPoint@{PercPoint}}
\subsubsection[{operator$<$}]{\setlength{\rightskip}{0pt plus 5cm}bool PercPoint::operator$<$ (const {\bf PercPoint} \& {\em other}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_perc_point_f798aa389b770208bd172542267ec811}

Less than comparison - we want to extract lowest values. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em other}]right hand side percolation point. \end{description}
\end{Desc}
\hypertarget{class_perc_point_0099784a4e3728c1f28b546a5b94104a}{
\index{PercPoint@{PercPoint}!value@{value}}
\index{value@{value}!PercPoint@{PercPoint}}
\subsubsection[{value}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} PercPoint::value () const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_perc_point_0099784a4e3728c1f28b546a5b94104a}

Get the value at this point. \begin{Desc}
\item[Returns:]value from the matrix. \end{Desc}

\subsection{Member Data Documentation}
\hypertarget{class_perc_point_730c461b6a04bcb883d60df9b297b5d5}{
\index{PercPoint@{PercPoint}!matrix@{matrix}}
\index{matrix@{matrix}!PercPoint@{PercPoint}}
\subsubsection[{matrix}]{\setlength{\rightskip}{0pt plus 5cm}{\bf IntMatrix} {\bf PercPoint::matrix}}}
\label{class_perc_point_730c461b6a04bcb883d60df9b297b5d5}

Matrix. \hypertarget{class_perc_point_dc46d54bf942be6e7774ae522e563f87}{
\index{PercPoint@{PercPoint}!nc@{nc}}
\index{nc@{nc}!PercPoint@{PercPoint}}
\subsubsection[{nc}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf PercPoint::nc}}}
\label{class_perc_point_dc46d54bf942be6e7774ae522e563f87}

Number of columns. \hypertarget{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}{
\index{PercPoint@{PercPoint}!point@{point}}
\index{point@{point}!PercPoint@{PercPoint}}
\subsubsection[{point}]{\setlength{\rightskip}{0pt plus 5cm}{\bf Point} {\bf PercPoint::point}}}
\label{class_perc_point_6f5497e144144fa0f0e5e4783ab30761}

\hyperlink{class_point}{Point}.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
cowichan\_\-serial/\hyperlink{cowichan__serial_2invperc_8cpp}{invperc.cpp}\item 
cowichan\_\-openmp/invperc.cpp\item 
cowichan\_\-tbb/invperc.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_point_distances.tex
documentation for serial, openmp, tbb

\hypertarget{class_point_distances}{
\section{PointDistances Class Reference}
\label{class_point_distances}\index{PointDistances@{PointDistances}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_point_distances_12dba51e28ac2cf5ac0c9317d9ff242e}{
\textbf{PointDistances} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)}
\label{class_point_distances_12dba51e28ac2cf5ac0c9317d9ff242e}

\item 
\hypertarget{class_point_distances_ec6ace35911f7e26d573023f2381c673}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{getMaximum} () const }
\label{class_point_distances_ec6ace35911f7e26d573023f2381c673}

\item 
void \hyperlink{class_point_distances_ab3ec991122054f46c2d5b5f7f91e972}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows)
\item 
\hyperlink{class_point_distances_59396cd3119746888d0601cd78757da3}{PointDistances} (\hyperlink{class_point_distances}{PointDistances} \&other, split)
\item 
void \hyperlink{class_point_distances_609701c6fa66e3f0406db82bc70453d9}{join} (const \hyperlink{class_point_distances}{PointDistances} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_point_distances_a6640fdeda32580a760309f564f8ac9d}{
\hyperlink{class_point}{PointVector} \textbf{\_\-points}}
\label{class_point_distances_a6640fdeda32580a760309f564f8ac9d}

\item 
\hypertarget{class_point_distances_3044927709dad1b9a94784f077d83190}{
\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} \textbf{\_\-matrix}}
\label{class_point_distances_3044927709dad1b9a94784f077d83190}

\item 
\hypertarget{class_point_distances_3233cd7ba2cdb8a51ceadce064569c7d}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-vector}}
\label{class_point_distances_3233cd7ba2cdb8a51ceadce064569c7d}

\item 
\hypertarget{class_point_distances_633e1a7fb762471b9a4806d52e024fcc}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{n}}
\label{class_point_distances_633e1a7fb762471b9a4806d52e024fcc}

\item 
\hypertarget{class_point_distances_5aa37e9525054b926bb2ce5c6bb515a8}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{\_\-max}}
\label{class_point_distances_5aa37e9525054b926bb2ce5c6bb515a8}

\end{CompactItemize}

\subsection{Detailed Description}
This module turns a vector containing point positions into a dense, symmet- ric, diagonally dominant matrix by calculating the distances between each pair of points. It also constructs a real vector whose values are the distance of each point from the origin. Inputs are:

points: a vector of (x, y) points, where x and y are the point’s position. nelts: the number of points in the vector, and the size of the matrix along each axis.

Its outputs are:

matrix: a real matrix, whose values are filled with inter-point distances. vector: a real vector, whose values are filled with origin-to-point distances.

Each matrix element Mi,j such that i != j is given the value di,j, the Eu- clidean distance between point i and point j. The diagonal values Mi,i are then set to nelts times the maximum off-diagonal value to ensure that the matrix is diagonally dominant. The value of the vector element vi is set to the distance of point i from the origin, which is given by sqrt(xi$^\wedge$2 + yi$^\wedge$2).

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_point_distances_59396cd3119746888d0601cd78757da3}{
\index{PointDistances@{PointDistances}!PointDistances@{PointDistances}}
\index{PointDistances@{PointDistances}!PointDistances@{PointDistances}}
\subsubsection[{PointDistances}]{\setlength{\rightskip}{0pt plus 5cm}PointDistances::PointDistances ({\bf PointDistances} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_point_distances_59396cd3119746888d0601cd78757da3}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_point_distances_609701c6fa66e3f0406db82bc70453d9}{
\index{PointDistances@{PointDistances}!join@{join}}
\index{join@{join}!PointDistances@{PointDistances}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void PointDistances::join (const {\bf PointDistances} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_point_distances_609701c6fa66e3f0406db82bc70453d9}

Joiner (TBB). \hypertarget{class_point_distances_ab3ec991122054f46c2d5b5f7f91e972}{
\index{PointDistances@{PointDistances}!operator()@{operator()}}
\index{operator()@{operator()}!PointDistances@{PointDistances}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void PointDistances::operator() (const {\bf Range} \& {\em rows})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_point_distances_ab3ec991122054f46c2d5b5f7f91e972}

Calculates inter-point distances on the given range.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/outer.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_product.tex
documentation for serial, openmp, tbb

\hypertarget{class_product}{
\section{Product Class Reference}
\label{class_product}\index{Product@{Product}}
}
Performs matrix vector product.

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{class_product_516ecf8496b6684c7288dc0d6b002965}{Product} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} result, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_product_7216f072885841c00e8e4aa3372ea277}{n})
\item 
void \hyperlink{class_product_a88c9ebd1997feb3dbd98086d5dd7cff}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows) const 
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} \hyperlink{class_product_63011ee94f2a8c0aa49468c1a8029478}{\_\-matrix}
\item 
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \hyperlink{class_product_40c40247c21e89bce81a64001c52caab}{\_\-vector}
\item 
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \hyperlink{class_product_81890b0d05e574d9dc8dae60a9815e32}{\_\-result}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_product_7216f072885841c00e8e4aa3372ea277}{n}
\end{CompactItemize}

\subsection{Detailed Description}
Performs matrix vector product.

This class multiplies a matrix by a vector, producing a vector, a la: \mbox{[}x1 x2 x3\mbox{]} \mbox{[}y1\mbox{]} \mbox{[}(x1$\ast$y1 + x2$\ast$y2 + x3$\ast$y3)\mbox{]} \mbox{[}x4 x5 x6\mbox{]} $\ast$ \mbox{[}y2\mbox{]} = \mbox{[}(x4$\ast$y1 + x5$\ast$y2 + x6$\ast$y3)\mbox{]} \mbox{[}x7 x8 x9\mbox{]} \mbox{[}y3\mbox{]} \mbox{[}(x7$\ast$y1 + x8$\ast$y2 + x9$\ast$y3)\mbox{]}

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_product_516ecf8496b6684c7288dc0d6b002965}{
\index{Product@{Product}!Product@{Product}}
\index{Product@{Product}!Product@{Product}}
\subsubsection[{Product}]{\setlength{\rightskip}{0pt plus 5cm}Product::Product ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em vector}, \/  {\bf Vector} {\em result}, \/  {\bf index\_\-t} {\em n})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_product_516ecf8496b6684c7288dc0d6b002965}

Construct a product object. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]given matrix. \item[{\em vector}]given vector. \item[{\em result}]solution vector. \item[{\em n}]matrix size. \end{description}
\end{Desc}

\subsection{Member Function Documentation}
\hypertarget{class_product_a88c9ebd1997feb3dbd98086d5dd7cff}{
\index{Product@{Product}!operator()@{operator()}}
\index{operator()@{operator()}!Product@{Product}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void Product::operator() (const {\bf Range} \& {\em rows}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_product_a88c9ebd1997feb3dbd98086d5dd7cff}

Performs matrix-vector multiplication on the given row range. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em rows}]range of rows to use. \end{description}
\end{Desc}

\subsection{Member Data Documentation}
\hypertarget{class_product_63011ee94f2a8c0aa49468c1a8029478}{
\index{Product@{Product}!\_\-matrix@{\_\-matrix}}
\index{\_\-matrix@{\_\-matrix}!Product@{Product}}
\subsubsection[{\_\-matrix}]{\setlength{\rightskip}{0pt plus 5cm}{\bf Matrix} {\bf Product::\_\-matrix}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_product_63011ee94f2a8c0aa49468c1a8029478}

Given matrix. \hypertarget{class_product_81890b0d05e574d9dc8dae60a9815e32}{
\index{Product@{Product}!\_\-result@{\_\-result}}
\index{\_\-result@{\_\-result}!Product@{Product}}
\subsubsection[{\_\-result}]{\setlength{\rightskip}{0pt plus 5cm}{\bf Vector} {\bf Product::\_\-result}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_product_81890b0d05e574d9dc8dae60a9815e32}

Solution vector. \hypertarget{class_product_40c40247c21e89bce81a64001c52caab}{
\index{Product@{Product}!\_\-vector@{\_\-vector}}
\index{\_\-vector@{\_\-vector}!Product@{Product}}
\subsubsection[{\_\-vector}]{\setlength{\rightskip}{0pt plus 5cm}{\bf Vector} {\bf Product::\_\-vector}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_product_40c40247c21e89bce81a64001c52caab}

Given vector. \hypertarget{class_product_7216f072885841c00e8e4aa3372ea277}{
\index{Product@{Product}!n@{n}}
\index{n@{n}!Product@{Product}}
\subsubsection[{n}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf Product::n}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_product_7216f072885841c00e8e4aa3372ea277}

Matrix size.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2product_8cpp}{product.cpp}\end{CompactItemize}

### 2009-08-02 doc/latex/class_random_generator.tex
documentation for serial, openmp, tbb

\hypertarget{class_random_generator}{
\section{RandomGenerator Class Reference}
\label{class_random_generator}\index{RandomGenerator@{RandomGenerator}}
}
Random generator.

\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{class_random_generator_bfa47006c53f82e85cbc77ae5e667747}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows) const 
\item 
\hyperlink{class_random_generator_96274661438a80cfee1e7a33bc59f770}{RandomGenerator} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_random_generator_ccc3239cc9aa8571936ca78b0e82adde}{nr}, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_random_generator_b2273a84d69579c9e9e89d0073ced26c}{nc})
\item 
void \hyperlink{class_random_generator_4c5ae77662dfca92b1677e36b124e3c4}{execute} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\end{CompactItemize}
\subsection*{Private Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_random_generator_caea6ecf8e980861f1d370cb54eaeeda}{next} (\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \&current) const 
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_random_generator_d0ee441c774a78abd93a50c0b08eea43}{nextK} (\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \&current) const 
\item 
void \hyperlink{class_random_generator_8a8390ed7e511770795cf4204d1f1df0}{initialise} ()
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \hyperlink{class_random_generator_9fb2b8631b95d10abb2f013319d37f3f}{\_\-matrix}
\item 
\hyperlink{cowichan_8hpp_9bab229b7d95f858be62c35cca6ff294}{IntVector} \hyperlink{class_random_generator_4423c35148a29e6ddd8e108a144db2ac}{state}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_random_generator_3f733ddf90320aace1be173bdc7381da}{aPrime}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{class_random_generator_990ae8409e5290b6e911b609fb61d8c8}{cPrime}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_random_generator_ccc3239cc9aa8571936ca78b0e82adde}{nr}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_random_generator_b2273a84d69579c9e9e89d0073ced26c}{nc}
\end{CompactItemize}

\subsection{Detailed Description}
Random generator.

This class generates random integers in a matrix, using a simple linear congruential number generator with a given seed value, i.e. using the recurrence:

X\_\-i+1 = (a$\ast$X\_\-i + c) mod m

The class also provides a parallel implementation with parallel\_\-for.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_random_generator_96274661438a80cfee1e7a33bc59f770}{
\index{RandomGenerator@{RandomGenerator}!RandomGenerator@{RandomGenerator}}
\index{RandomGenerator@{RandomGenerator}!RandomGenerator@{RandomGenerator}}
\subsubsection[{RandomGenerator}]{\setlength{\rightskip}{0pt plus 5cm}RandomGenerator::RandomGenerator ({\bf index\_\-t} {\em nr}, \/  {\bf index\_\-t} {\em nc})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_random_generator_96274661438a80cfee1e7a33bc59f770}

Construct a random number generator. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em nr}]number of matrix rows. \item[{\em nc}]number of matrix columns. \end{description}
\end{Desc}

\subsection{Member Function Documentation}
\hypertarget{class_random_generator_4c5ae77662dfca92b1677e36b124e3c4}{
\index{RandomGenerator@{RandomGenerator}!execute@{execute}}
\index{execute@{execute}!RandomGenerator@{RandomGenerator}}
\subsubsection[{execute}]{\setlength{\rightskip}{0pt plus 5cm}void RandomGenerator::execute ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_random_generator_4c5ae77662dfca92b1677e36b124e3c4}

Run in parallel. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to use. \end{description}
\end{Desc}
\hypertarget{class_random_generator_8a8390ed7e511770795cf4204d1f1df0}{
\index{RandomGenerator@{RandomGenerator}!initialise@{initialise}}
\index{initialise@{initialise}!RandomGenerator@{RandomGenerator}}
\subsubsection[{initialise}]{\setlength{\rightskip}{0pt plus 5cm}void RandomGenerator::initialise ()\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{class_random_generator_8a8390ed7e511770795cf4204d1f1df0}

Initialises the random number generator to operate on NROWS different rows completely independently. \hypertarget{class_random_generator_caea6ecf8e980861f1d370cb54eaeeda}{
\index{RandomGenerator@{RandomGenerator}!next@{next}}
\index{next@{next}!RandomGenerator@{RandomGenerator}}
\subsubsection[{next}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} RandomGenerator::next ({\bf INT\_\-TYPE} \& {\em current}) const\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{class_random_generator_caea6ecf8e980861f1d370cb54eaeeda}

Generates the next random number. Convenience method for next(current, a, c); \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em current}]current element. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Next element. \end{Desc}
\hypertarget{class_random_generator_d0ee441c774a78abd93a50c0b08eea43}{
\index{RandomGenerator@{RandomGenerator}!nextK@{nextK}}
\index{nextK@{nextK}!RandomGenerator@{RandomGenerator}}
\subsubsection[{nextK}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} RandomGenerator::nextK ({\bf INT\_\-TYPE} \& {\em current}) const\hspace{0.3cm}{\tt  \mbox{[}inline, private\mbox{]}}}}
\label{class_random_generator_d0ee441c774a78abd93a50c0b08eea43}

Generates the k-th next random number, using an identity, where A = a$^\wedge$k mod m and C = c $\ast$ sum\mbox{[}j=0..k-1\mbox{]}(a$^\wedge$j mod m). \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em current}]element. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Next k-th element. \end{Desc}
\hypertarget{class_random_generator_bfa47006c53f82e85cbc77ae5e667747}{
\index{RandomGenerator@{RandomGenerator}!operator()@{operator()}}
\index{operator()@{operator()}!RandomGenerator@{RandomGenerator}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void RandomGenerator::operator() (const {\bf Range} \& {\em rows}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_random_generator_bfa47006c53f82e85cbc77ae5e667747}

Provides NROWS-seperated linear congruential RNG on the specified range of rows. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em rows}]range of rows to work on. \end{description}
\end{Desc}

\subsection{Member Data Documentation}
\hypertarget{class_random_generator_9fb2b8631b95d10abb2f013319d37f3f}{
\index{RandomGenerator@{RandomGenerator}!\_\-matrix@{\_\-matrix}}
\index{\_\-matrix@{\_\-matrix}!RandomGenerator@{RandomGenerator}}
\subsubsection[{\_\-matrix}]{\setlength{\rightskip}{0pt plus 5cm}{\bf IntMatrix} {\bf RandomGenerator::\_\-matrix}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_9fb2b8631b95d10abb2f013319d37f3f}

Matrix to fill. \hypertarget{class_random_generator_3f733ddf90320aace1be173bdc7381da}{
\index{RandomGenerator@{RandomGenerator}!aPrime@{aPrime}}
\index{aPrime@{aPrime}!RandomGenerator@{RandomGenerator}}
\subsubsection[{aPrime}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} {\bf RandomGenerator::aPrime}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_3f733ddf90320aace1be173bdc7381da}

Constant a prime. \hypertarget{class_random_generator_990ae8409e5290b6e911b609fb61d8c8}{
\index{RandomGenerator@{RandomGenerator}!cPrime@{cPrime}}
\index{cPrime@{cPrime}!RandomGenerator@{RandomGenerator}}
\subsubsection[{cPrime}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} {\bf RandomGenerator::cPrime}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_990ae8409e5290b6e911b609fb61d8c8}

Constant c prime. \hypertarget{class_random_generator_b2273a84d69579c9e9e89d0073ced26c}{
\index{RandomGenerator@{RandomGenerator}!nc@{nc}}
\index{nc@{nc}!RandomGenerator@{RandomGenerator}}
\subsubsection[{nc}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf RandomGenerator::nc}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_b2273a84d69579c9e9e89d0073ced26c}

Number of columns in the matrix. \hypertarget{class_random_generator_ccc3239cc9aa8571936ca78b0e82adde}{
\index{RandomGenerator@{RandomGenerator}!nr@{nr}}
\index{nr@{nr}!RandomGenerator@{RandomGenerator}}
\subsubsection[{nr}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} {\bf RandomGenerator::nr}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_ccc3239cc9aa8571936ca78b0e82adde}

Number of rows in the matrix. \hypertarget{class_random_generator_4423c35148a29e6ddd8e108a144db2ac}{
\index{RandomGenerator@{RandomGenerator}!state@{state}}
\index{state@{state}!RandomGenerator@{RandomGenerator}}
\subsubsection[{state}]{\setlength{\rightskip}{0pt plus 5cm}{\bf IntVector} {\bf RandomGenerator::state}\hspace{0.3cm}{\tt  \mbox{[}private\mbox{]}}}}
\label{class_random_generator_4423c35148a29e6ddd8e108a144db2ac}

First element vector.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/\hyperlink{cowichan__tbb_2randmat_8cpp}{randmat.cpp}\end{CompactItemize}

### 2009-08-02 doc/latex/class_relaxer.tex
documentation for serial, openmp, tbb

\hypertarget{class_relaxer}{
\section{Relaxer Class Reference}
\label{class_relaxer}\index{Relaxer@{Relaxer}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_relaxer_7867f2f26839f9faaa150b4d0ea180fa}{
\textbf{Relaxer} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)}
\label{class_relaxer_7867f2f26839f9faaa150b4d0ea180fa}

\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_relaxer_6b4b0383b11cfdc6fc48ec7e137e688d}{getMaxDiff} () const 
\item 
void \hyperlink{class_relaxer_bbcc97706ee506fb14f822fe8ae57ee9}{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&range)
\item 
\hyperlink{class_relaxer_34cd0a0220031f4ec6ab21dbcf9a9970}{Relaxer} (\hyperlink{class_relaxer}{Relaxer} \&other, split)
\item 
void \hyperlink{class_relaxer_521cb3709f2af24b307114d016b2cd44}{join} (const \hyperlink{class_relaxer}{Relaxer} \&other)
\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_relaxer_9055aba4c4a03387dd8809c3ab4a0629}{
\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} \textbf{\_\-matrix}}
\label{class_relaxer_9055aba4c4a03387dd8809c3ab4a0629}

\item 
\hypertarget{class_relaxer_352fb652b751f1df458ee2922fe6031f}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-target}}
\label{class_relaxer_352fb652b751f1df458ee2922fe6031f}

\item 
\hypertarget{class_relaxer_2c9cfda63081510022de66b90545382b}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-solution}}
\label{class_relaxer_2c9cfda63081510022de66b90545382b}

\item 
\hypertarget{class_relaxer_e5482dd547d0e30ab75a0d678ab0a6a7}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{n}}
\label{class_relaxer_e5482dd547d0e30ab75a0d678ab0a6a7}

\item 
\hypertarget{class_relaxer_8ea0071b5726e7a21f2cf33637ba8d00}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{maxDiff}}
\label{class_relaxer_8ea0071b5726e7a21f2cf33637ba8d00}

\end{CompactItemize}

\subsection{Detailed Description}
Performs one iteration of relaxation.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_relaxer_34cd0a0220031f4ec6ab21dbcf9a9970}{
\index{Relaxer@{Relaxer}!Relaxer@{Relaxer}}
\index{Relaxer@{Relaxer}!Relaxer@{Relaxer}}
\subsubsection[{Relaxer}]{\setlength{\rightskip}{0pt plus 5cm}Relaxer::Relaxer ({\bf Relaxer} \& {\em other}, \/  split)\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_relaxer_34cd0a0220031f4ec6ab21dbcf9a9970}

Splitting (TBB) constructor

\subsection{Member Function Documentation}
\hypertarget{class_relaxer_6b4b0383b11cfdc6fc48ec7e137e688d}{
\index{Relaxer@{Relaxer}!getMaxDiff@{getMaxDiff}}
\index{getMaxDiff@{getMaxDiff}!Relaxer@{Relaxer}}
\subsubsection[{getMaxDiff}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} Relaxer::getMaxDiff () const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_relaxer_6b4b0383b11cfdc6fc48ec7e137e688d}

Get maximum difference. \hypertarget{class_relaxer_521cb3709f2af24b307114d016b2cd44}{
\index{Relaxer@{Relaxer}!join@{join}}
\index{join@{join}!Relaxer@{Relaxer}}
\subsubsection[{join}]{\setlength{\rightskip}{0pt plus 5cm}void Relaxer::join (const {\bf Relaxer} \& {\em other})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_relaxer_521cb3709f2af24b307114d016b2cd44}

Joiner (TBB). \hypertarget{class_relaxer_bbcc97706ee506fb14f822fe8ae57ee9}{
\index{Relaxer@{Relaxer}!operator()@{operator()}}
\index{operator()@{operator()}!Relaxer@{Relaxer}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void Relaxer::operator() (const {\bf Range} \& {\em range})\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_relaxer_bbcc97706ee506fb14f822fe8ae57ee9}

Performs one iteration of relaxation.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/sor.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_row_elimination.tex
documentation for serial, openmp, tbb

\hypertarget{class_row_elimination}{
\section{RowElimination Class Reference}
\label{class_row_elimination}\index{RowElimination@{RowElimination}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_row_elimination_582416939554bbb8e462e53d6d86f8d6}{
\textbf{RowElimination} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)}
\label{class_row_elimination_582416939554bbb8e462e53d6d86f8d6}

\item 
\hypertarget{class_row_elimination_9037d2255c9363e4ed5315e4457e09c2}{
void \textbf{setI} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} i)}
\label{class_row_elimination_9037d2255c9363e4ed5315e4457e09c2}

\item 
\hypertarget{class_row_elimination_8fa4ceac659d237e6cf693321623e328}{
void \textbf{operator()} (const \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range} \&rows) const }
\label{class_row_elimination_8fa4ceac659d237e6cf693321623e328}

\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_row_elimination_49344262a09cdf006d4b249e8fddf3d4}{
\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} \textbf{\_\-matrix}}
\label{class_row_elimination_49344262a09cdf006d4b249e8fddf3d4}

\item 
\hypertarget{class_row_elimination_1f473bc02c05aa508b467e6097e13ddf}{
\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} \textbf{\_\-target}}
\label{class_row_elimination_1f473bc02c05aa508b467e6097e13ddf}

\item 
\hypertarget{class_row_elimination_593558f30cdc67d74b41e11197a56095}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{n}}
\label{class_row_elimination_593558f30cdc67d74b41e11197a56095}

\item 
\hypertarget{class_row_elimination_4ef655247f5284b1694b57f5687d2d40}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{i}}
\label{class_row_elimination_4ef655247f5284b1694b57f5687d2d40}

\item 
\hypertarget{class_row_elimination_5a96a12ab6503193733e9e4c2e545f07}{
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \textbf{column\_\-i}}
\label{class_row_elimination_5a96a12ab6503193733e9e4c2e545f07}

\end{CompactItemize}

\subsection{Detailed Description}
Performs row elimination, i.e. eliminate i-th column in j-th row.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/gauss.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_shuffle.tex
documentation for serial, openmp, tbb

\hypertarget{class_shuffle}{
\section{Shuffle Class Reference}
\label{class_shuffle}\index{Shuffle@{Shuffle}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_shuffle_b6e81ca9e45260ac098005caddd5a663}{
\textbf{Shuffle} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} input, \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} output, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)}
\label{class_shuffle_b6e81ca9e45260ac098005caddd5a663}

\item 
void \hyperlink{class_shuffle_7922aacb5aae207d2e51ded86b3c1c9d}{operator()} (const \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D} \&range) const 
\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_shuffle_b0fed7e167e632459538da3d6e810b2e}{
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \textbf{\_\-first}}
\label{class_shuffle_b0fed7e167e632459538da3d6e810b2e}

\item 
\hypertarget{class_shuffle_4b6d3c7369a12163f9ad5a080a5b208e}{
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \textbf{\_\-second}}
\label{class_shuffle_4b6d3c7369a12163f9ad5a080a5b208e}

\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_shuffle_d8f6038cc7fd4f722d7f00f2a6af9698}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{xBreak}}
\label{class_shuffle_d8f6038cc7fd4f722d7f00f2a6af9698}

\item 
\hypertarget{class_shuffle_a41b51804d77e262d65fc79474879e9c}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{yBreak}}
\label{class_shuffle_a41b51804d77e262d65fc79474879e9c}

\item 
\hypertarget{class_shuffle_d685d43331c6f9a9d0f0775d8e21ebff}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nr}}
\label{class_shuffle_d685d43331c6f9a9d0f0775d8e21ebff}

\item 
\hypertarget{class_shuffle_c7f2688b41a3bc24e7939164a28784e0}{
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nc}}
\label{class_shuffle_c7f2688b41a3bc24e7939164a28784e0}

\end{CompactItemize}

\subsection{Detailed Description}
This class does a halving shuffle.

\subsection{Member Function Documentation}
\hypertarget{class_shuffle_7922aacb5aae207d2e51ded86b3c1c9d}{
\index{Shuffle@{Shuffle}!operator()@{operator()}}
\index{operator()@{operator()}!Shuffle@{Shuffle}}
\subsubsection[{operator()}]{\setlength{\rightskip}{0pt plus 5cm}void Shuffle::operator() (const {\bf Range2D} \& {\em range}) const\hspace{0.3cm}{\tt  \mbox{[}inline\mbox{]}}}}
\label{class_shuffle_7922aacb5aae207d2e51ded86b3c1c9d}

Performs the halving shuffle over the given range.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/half.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/class_threshold.tex
documentation for serial, openmp, tbb

\hypertarget{class_threshold}{
\section{Threshold Class Reference}
\label{class_threshold}\index{Threshold@{Threshold}}
}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hypertarget{class_threshold_bdd06b7735ea6b46efab1685411e3d70}{
void \textbf{operator()} (const \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D} \&range) const }
\label{class_threshold_bdd06b7735ea6b46efab1685411e3d70}

\item 
\hypertarget{class_threshold_df53dd36b3ab7336627ab7361ab3eb97}{
\textbf{Threshold} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} image, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} retain, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} result, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)}
\label{class_threshold_df53dd36b3ab7336627ab7361ab3eb97}

\end{CompactItemize}
\subsection*{Private Attributes}
\begin{CompactItemize}
\item 
\hypertarget{class_threshold_53a4f784d8bd2d813a21aa1a52e5afbb}{
\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} \textbf{\_\-image}}
\label{class_threshold_53a4f784d8bd2d813a21aa1a52e5afbb}

\item 
\hypertarget{class_threshold_f5e82e77ffad10da1dc8c6501dddcdea}{
\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} \textbf{\_\-result}}
\label{class_threshold_f5e82e77ffad10da1dc8c6501dddcdea}

\item 
\hypertarget{class_threshold_cf3f8f39e3d71732c76525d62db39116}{
const \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{retain}}
\label{class_threshold_cf3f8f39e3d71732c76525d62db39116}

\item 
\hypertarget{class_threshold_0555939aa91492c2d58b8c650a992e77}{
const \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \textbf{nc}}
\label{class_threshold_0555939aa91492c2d58b8c650a992e77}

\end{CompactItemize}

\subsection{Detailed Description}
This class takes an Integer array to a boolean array based on a cut-off value.

The documentation for this class was generated from the following file:\begin{CompactItemize}
\item 
cowichan\_\-tbb/thresh.cpp\end{CompactItemize}

### 2009-08-02 doc/latex/cowichan__defaults_8hpp.tex
documentation for serial, openmp, tbb

\hypertarget{cowichan__defaults_8hpp}{
\section{cowichan/cowichan\_\-defaults.hpp File Reference}
\label{cowichan__defaults_8hpp}\index{cowichan/cowichan\_\-defaults.hpp@{cowichan/cowichan\_\-defaults.hpp}}
}
Defaults for \hyperlink{class_cowichan}{Cowichan} programs.

\subsection*{Defines}
\begin{CompactItemize}
\item 
\#define \hyperlink{cowichan__defaults_8hpp_30524c594b6941ca7822489d1794813b}{ALL\_\-NR}~15000
\item 
\#define \hyperlink{cowichan__defaults_8hpp_85076a3092aabd0bc4661e9184e2ddf3}{ALL\_\-NC}~15000
\item 
\#define \hyperlink{cowichan__defaults_8hpp_e36abbd26bf4ea979de0b63b1a37e349}{ALL\_\-N}~15000
\item 
\#define \hyperlink{cowichan__defaults_8hpp_0c6044d9326c14200929af66072b673a}{RAND\_\-MEAN}~0
\item 
\#define \hyperlink{cowichan__defaults_8hpp_40b63d443076ce61004bae3bf79a776b}{RAND\_\-RANGE}~20
\item 
\#define \hyperlink{cowichan__defaults_8hpp_32ee58f73d9de0d1d41ad46f7bff2719}{RAND\_\-SEED}~681304
\item 
\#define \hyperlink{cowichan__defaults_8hpp_d07476570307484eff173eedfbddd347}{RAND\_\-M}~56197
\item 
\#define \hyperlink{cowichan__defaults_8hpp_5f41a7e62c6e0138af5a65460416d6c7}{MANDEL\_\-NR}~ALL\_\-NR
\item 
\#define \hyperlink{cowichan__defaults_8hpp_d60d6cbcceef6bc7249b1480c829b7bb}{MANDEL\_\-NC}~ALL\_\-NC
\item 
\#define \hyperlink{cowichan__defaults_8hpp_5d8b0b1a78ed8a193734749178c248c1}{MANDEL\_\-X0}~0
\item 
\#define \hyperlink{cowichan__defaults_8hpp_47e550564bb994890b2ea6816155a0fe}{MANDEL\_\-Y0}~0
\item 
\#define \hyperlink{cowichan__defaults_8hpp_6fb2dcd18c6df1c377f613769d6314c3}{MANDEL\_\-DX}~1.5
\item 
\#define \hyperlink{cowichan__defaults_8hpp_b1759c852c10a8d03520a84beb5378fd}{MANDEL\_\-DY}~1.5
\item 
\#define \hyperlink{cowichan__defaults_8hpp_c911423783b8cba366d042acd8d062f3}{MANDEL\_\-INFINITY}~2.0
\item 
\#define \hyperlink{cowichan__defaults_8hpp_59c2aa73700269b6de563e60af48c6a3}{MANDEL\_\-MAX\_\-ITER}~150
\item 
\#define \hyperlink{cowichan__defaults_8hpp_6b2590531a8313fed76899a95ad3b625}{RANDMAT\_\-NR}~ALL\_\-NR
\item 
\#define \hyperlink{cowichan__defaults_8hpp_be3ee010372d99dfda2913d3d35a6a05}{RANDMAT\_\-NC}~ALL\_\-NC
\item 
\#define \hyperlink{cowichan__defaults_8hpp_7565d8f613d834da3b98134d4b564017}{RANDMAT\_\-A}~1291
\item 
\#define \hyperlink{cowichan__defaults_8hpp_cacc7f34c70c2ca998f0c1372ed1db83}{RANDMAT\_\-C}~917
\item 
\#define \hyperlink{cowichan__defaults_8hpp_3f8c101b36fa58d4d1d26cc4a77e1b51}{HALF\_\-NR}~ALL\_\-NR
\item 
\#define \hyperlink{cowichan__defaults_8hpp_d97fd174769938baa91d18ac273e1f93}{HALF\_\-NC}~ALL\_\-NC
\item 
\#define \hyperlink{cowichan__defaults_8hpp_966db6c2f5b3ed69daacad1d722b7071}{INVPERC\_\-NR}~ALL\_\-NR
\item 
\#define \hyperlink{cowichan__defaults_8hpp_3ae0f26f20ab30b6bbe513eff9a18291}{INVPERC\_\-NC}~ALL\_\-NC
\item 
\#define \hyperlink{cowichan__defaults_8hpp_cbaf3f96d8f75268024fb88255da23d9}{INVPERC\_\-NFILL}~1000000
\item 
\#define \hyperlink{cowichan__defaults_8hpp_4b80608f6c4617a66eba2f164afc4f3d}{THRESH\_\-NR}~ALL\_\-NR
\item 
\#define \hyperlink{cowichan__defaults_8hpp_7c356bfeab4913aa2e5dea26de44ebfc}{THRESH\_\-NC}~ALL\_\-NC
\item 
\#define \hyperlink{cowichan__defaults_8hpp_f85ccfff5fbed47416c64d4229d1234a}{THRESH\_\-PERCENT}~0.5
\item 
\#define \hyperlink{cowichan__defaults_8hpp_a597d047fbe34455bd7962e6e550adab}{LIFE\_\-NR}~ALL\_\-NR
\item 
\#define \hyperlink{cowichan__defaults_8hpp_9763036e0b951ebdb691f7336ecf0493}{LIFE\_\-NC}~ALL\_\-NC
\item 
\#define \hyperlink{cowichan__defaults_8hpp_02f3c90de37aa6314330a7e0cb68a1f5}{LIFE\_\-ITERATIONS}~10
\item 
\#define \hyperlink{cowichan__defaults_8hpp_cedd24970400253f48e19a285f8b8559}{WINNOW\_\-NR}~ALL\_\-NR
\item 
\#define \hyperlink{cowichan__defaults_8hpp_40f06800a12a97dde304a378d7d971cc}{WINNOW\_\-NC}~ALL\_\-NC
\item 
\#define \hyperlink{cowichan__defaults_8hpp_22c87347c9b1f8643a8b57e3d841303f}{WINNOW\_\-N}~ALL\_\-N
\item 
\#define \hyperlink{cowichan__defaults_8hpp_2ba93b0a30632568e28703c194c2345a}{NORM\_\-N}~ALL\_\-N
\item 
\#define \hyperlink{cowichan__defaults_8hpp_7f62220bbf93745fac5aa8e225f09126}{HULL\_\-N}~ALL\_\-N
\item 
\#define \hyperlink{cowichan__defaults_8hpp_113e7aa75d2b1403a5548913cd240cb4}{OUTER\_\-N}~ALL\_\-N
\item 
\#define \hyperlink{cowichan__defaults_8hpp_e1de0ce8df246861c84e405d66d37457}{GAUSS\_\-N}~ALL\_\-N
\item 
\#define \hyperlink{cowichan__defaults_8hpp_23487499c48e5b64d05fd2249be710c1}{SOR\_\-N}~ALL\_\-N
\item 
\#define \hyperlink{cowichan__defaults_8hpp_4b7aa2e2b41cee6b710c7ce48115f8ee}{SOR\_\-OMEGA}~0.9
\item 
\#define \hyperlink{cowichan__defaults_8hpp_e64dba51c415f6acbccbfd021a88769d}{SOR\_\-TOLERANCE}~10e-6
\item 
\#define \hyperlink{cowichan__defaults_8hpp_3673c55137aa755f4829b6b04329c379}{SOR\_\-MAX\_\-ITERS}~1000000
\item 
\#define \hyperlink{cowichan__defaults_8hpp_072a640a55e457c256faa59842e0368c}{PRODUCT\_\-N}~ALL\_\-N
\item 
\#define \hyperlink{cowichan__defaults_8hpp_d955f5a864096d87337347af906829b1}{VECDIFF\_\-N}~ALL\_\-N
\item 
\#define \hyperlink{cowichan__defaults_8hpp_5463dc5662d922fb1f8772a9e8e48672}{CHAIN\_\-NR}~ALL\_\-NR
\item 
\#define \hyperlink{cowichan__defaults_8hpp_fbc8b36ae001421db4f569c682ca7a45}{CHAIN\_\-NC}~ALL\_\-NC
\item 
\#define \hyperlink{cowichan__defaults_8hpp_40496e72e1bb9b2ca7ba854fa5679b8d}{CHAIN\_\-N}~ALL\_\-N
\end{CompactItemize}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
Defaults for \hyperlink{class_cowichan}{Cowichan} programs.

\subsection{Define Documentation}
\hypertarget{cowichan__defaults_8hpp_e36abbd26bf4ea979de0b63b1a37e349}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!ALL\_\-N@{ALL\_\-N}}
\index{ALL\_\-N@{ALL\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{ALL\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-N~15000}}
\label{cowichan__defaults_8hpp_e36abbd26bf4ea979de0b63b1a37e349}

Default matrix size for square matrices for all problems. \hypertarget{cowichan__defaults_8hpp_85076a3092aabd0bc4661e9184e2ddf3}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!ALL\_\-NC@{ALL\_\-NC}}
\index{ALL\_\-NC@{ALL\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{ALL\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-NC~15000}}
\label{cowichan__defaults_8hpp_85076a3092aabd0bc4661e9184e2ddf3}

Default number of columns for rectangular matrices for all problems. \hypertarget{cowichan__defaults_8hpp_30524c594b6941ca7822489d1794813b}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!ALL\_\-NR@{ALL\_\-NR}}
\index{ALL\_\-NR@{ALL\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{ALL\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-NR~15000}}
\label{cowichan__defaults_8hpp_30524c594b6941ca7822489d1794813b}

Default number of rows for rectangular matrices for all problems. \hypertarget{cowichan__defaults_8hpp_40496e72e1bb9b2ca7ba854fa5679b8d}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!CHAIN\_\-N@{CHAIN\_\-N}}
\index{CHAIN\_\-N@{CHAIN\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{CHAIN\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define CHAIN\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_40496e72e1bb9b2ca7ba854fa5679b8d}

Default matrix size for square matrices for chain. \hypertarget{cowichan__defaults_8hpp_fbc8b36ae001421db4f569c682ca7a45}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!CHAIN\_\-NC@{CHAIN\_\-NC}}
\index{CHAIN\_\-NC@{CHAIN\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{CHAIN\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define CHAIN\_\-NC~ALL\_\-NC}}
\label{cowichan__defaults_8hpp_fbc8b36ae001421db4f569c682ca7a45}

Default number of columns for rectangular matrices for chain. \hypertarget{cowichan__defaults_8hpp_5463dc5662d922fb1f8772a9e8e48672}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!CHAIN\_\-NR@{CHAIN\_\-NR}}
\index{CHAIN\_\-NR@{CHAIN\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{CHAIN\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define CHAIN\_\-NR~ALL\_\-NR}}
\label{cowichan__defaults_8hpp_5463dc5662d922fb1f8772a9e8e48672}

Default number of rows for rectangular matrices for chain. \hypertarget{cowichan__defaults_8hpp_e1de0ce8df246861c84e405d66d37457}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!GAUSS\_\-N@{GAUSS\_\-N}}
\index{GAUSS\_\-N@{GAUSS\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{GAUSS\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define GAUSS\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_e1de0ce8df246861c84e405d66d37457}

Default square matrix size for gauss. \hypertarget{cowichan__defaults_8hpp_d97fd174769938baa91d18ac273e1f93}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!HALF\_\-NC@{HALF\_\-NC}}
\index{HALF\_\-NC@{HALF\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{HALF\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define HALF\_\-NC~ALL\_\-NC}}
\label{cowichan__defaults_8hpp_d97fd174769938baa91d18ac273e1f93}

Default number of columns for half. \hypertarget{cowichan__defaults_8hpp_3f8c101b36fa58d4d1d26cc4a77e1b51}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!HALF\_\-NR@{HALF\_\-NR}}
\index{HALF\_\-NR@{HALF\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{HALF\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define HALF\_\-NR~ALL\_\-NR}}
\label{cowichan__defaults_8hpp_3f8c101b36fa58d4d1d26cc4a77e1b51}

Default number of rows for half. \hypertarget{cowichan__defaults_8hpp_7f62220bbf93745fac5aa8e225f09126}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!HULL\_\-N@{HULL\_\-N}}
\index{HULL\_\-N@{HULL\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{HULL\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define HULL\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_7f62220bbf93745fac5aa8e225f09126}

Default square matrix size for hull. \hypertarget{cowichan__defaults_8hpp_3ae0f26f20ab30b6bbe513eff9a18291}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!INVPERC\_\-NC@{INVPERC\_\-NC}}
\index{INVPERC\_\-NC@{INVPERC\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{INVPERC\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define INVPERC\_\-NC~ALL\_\-NC}}
\label{cowichan__defaults_8hpp_3ae0f26f20ab30b6bbe513eff9a18291}

Default number of columns for invperc. \hypertarget{cowichan__defaults_8hpp_cbaf3f96d8f75268024fb88255da23d9}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!INVPERC\_\-NFILL@{INVPERC\_\-NFILL}}
\index{INVPERC\_\-NFILL@{INVPERC\_\-NFILL}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{INVPERC\_\-NFILL}]{\setlength{\rightskip}{0pt plus 5cm}\#define INVPERC\_\-NFILL~1000000}}
\label{cowichan__defaults_8hpp_cbaf3f96d8f75268024fb88255da23d9}

Default number of cells to fill. \hypertarget{cowichan__defaults_8hpp_966db6c2f5b3ed69daacad1d722b7071}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!INVPERC\_\-NR@{INVPERC\_\-NR}}
\index{INVPERC\_\-NR@{INVPERC\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{INVPERC\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define INVPERC\_\-NR~ALL\_\-NR}}
\label{cowichan__defaults_8hpp_966db6c2f5b3ed69daacad1d722b7071}

Default number of rows for invperc. \hypertarget{cowichan__defaults_8hpp_02f3c90de37aa6314330a7e0cb68a1f5}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!LIFE\_\-ITERATIONS@{LIFE\_\-ITERATIONS}}
\index{LIFE\_\-ITERATIONS@{LIFE\_\-ITERATIONS}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{LIFE\_\-ITERATIONS}]{\setlength{\rightskip}{0pt plus 5cm}\#define LIFE\_\-ITERATIONS~10}}
\label{cowichan__defaults_8hpp_02f3c90de37aa6314330a7e0cb68a1f5}

Default number of life iterations. \hypertarget{cowichan__defaults_8hpp_9763036e0b951ebdb691f7336ecf0493}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!LIFE\_\-NC@{LIFE\_\-NC}}
\index{LIFE\_\-NC@{LIFE\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{LIFE\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define LIFE\_\-NC~ALL\_\-NC}}
\label{cowichan__defaults_8hpp_9763036e0b951ebdb691f7336ecf0493}

Default number of columns for life. \hypertarget{cowichan__defaults_8hpp_a597d047fbe34455bd7962e6e550adab}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!LIFE\_\-NR@{LIFE\_\-NR}}
\index{LIFE\_\-NR@{LIFE\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{LIFE\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define LIFE\_\-NR~ALL\_\-NR}}
\label{cowichan__defaults_8hpp_a597d047fbe34455bd7962e6e550adab}

Default number of rows for life. \hypertarget{cowichan__defaults_8hpp_6fb2dcd18c6df1c377f613769d6314c3}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!MANDEL\_\-DX@{MANDEL\_\-DX}}
\index{MANDEL\_\-DX@{MANDEL\_\-DX}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{MANDEL\_\-DX}]{\setlength{\rightskip}{0pt plus 5cm}\#define MANDEL\_\-DX~1.5}}
\label{cowichan__defaults_8hpp_6fb2dcd18c6df1c377f613769d6314c3}

Default extent of the region along the x axis. \hypertarget{cowichan__defaults_8hpp_b1759c852c10a8d03520a84beb5378fd}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!MANDEL\_\-DY@{MANDEL\_\-DY}}
\index{MANDEL\_\-DY@{MANDEL\_\-DY}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{MANDEL\_\-DY}]{\setlength{\rightskip}{0pt plus 5cm}\#define MANDEL\_\-DY~1.5}}
\label{cowichan__defaults_8hpp_b1759c852c10a8d03520a84beb5378fd}

Default extent of the region along the y axis. \hypertarget{cowichan__defaults_8hpp_c911423783b8cba366d042acd8d062f3}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!MANDEL\_\-INFINITY@{MANDEL\_\-INFINITY}}
\index{MANDEL\_\-INFINITY@{MANDEL\_\-INFINITY}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{MANDEL\_\-INFINITY}]{\setlength{\rightskip}{0pt plus 5cm}\#define MANDEL\_\-INFINITY~2.0}}
\label{cowichan__defaults_8hpp_c911423783b8cba366d042acd8d062f3}

Default mandel infinity. \hypertarget{cowichan__defaults_8hpp_59c2aa73700269b6de563e60af48c6a3}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!MANDEL\_\-MAX\_\-ITER@{MANDEL\_\-MAX\_\-ITER}}
\index{MANDEL\_\-MAX\_\-ITER@{MANDEL\_\-MAX\_\-ITER}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{MANDEL\_\-MAX\_\-ITER}]{\setlength{\rightskip}{0pt plus 5cm}\#define MANDEL\_\-MAX\_\-ITER~150}}
\label{cowichan__defaults_8hpp_59c2aa73700269b6de563e60af48c6a3}

Default mandel maximum number of iterations. \hypertarget{cowichan__defaults_8hpp_d60d6cbcceef6bc7249b1480c829b7bb}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!MANDEL\_\-NC@{MANDEL\_\-NC}}
\index{MANDEL\_\-NC@{MANDEL\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{MANDEL\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define MANDEL\_\-NC~ALL\_\-NC}}
\label{cowichan__defaults_8hpp_d60d6cbcceef6bc7249b1480c829b7bb}

Default number of columns for mandel. \hypertarget{cowichan__defaults_8hpp_5f41a7e62c6e0138af5a65460416d6c7}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!MANDEL\_\-NR@{MANDEL\_\-NR}}
\index{MANDEL\_\-NR@{MANDEL\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{MANDEL\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define MANDEL\_\-NR~ALL\_\-NR}}
\label{cowichan__defaults_8hpp_5f41a7e62c6e0138af5a65460416d6c7}

Default number of rows for mandel. \hypertarget{cowichan__defaults_8hpp_5d8b0b1a78ed8a193734749178c248c1}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!MANDEL\_\-X0@{MANDEL\_\-X0}}
\index{MANDEL\_\-X0@{MANDEL\_\-X0}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{MANDEL\_\-X0}]{\setlength{\rightskip}{0pt plus 5cm}\#define MANDEL\_\-X0~0}}
\label{cowichan__defaults_8hpp_5d8b0b1a78ed8a193734749178c248c1}

Default x-coordinate of the lower left corner. \hypertarget{cowichan__defaults_8hpp_47e550564bb994890b2ea6816155a0fe}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!MANDEL\_\-Y0@{MANDEL\_\-Y0}}
\index{MANDEL\_\-Y0@{MANDEL\_\-Y0}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{MANDEL\_\-Y0}]{\setlength{\rightskip}{0pt plus 5cm}\#define MANDEL\_\-Y0~0}}
\label{cowichan__defaults_8hpp_47e550564bb994890b2ea6816155a0fe}

Default y-coordinate of the lower left corner \hypertarget{cowichan__defaults_8hpp_2ba93b0a30632568e28703c194c2345a}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!NORM\_\-N@{NORM\_\-N}}
\index{NORM\_\-N@{NORM\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{NORM\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define NORM\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_2ba93b0a30632568e28703c194c2345a}

Default square matrix size for norm. \hypertarget{cowichan__defaults_8hpp_113e7aa75d2b1403a5548913cd240cb4}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!OUTER\_\-N@{OUTER\_\-N}}
\index{OUTER\_\-N@{OUTER\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{OUTER\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define OUTER\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_113e7aa75d2b1403a5548913cd240cb4}

Default square matrix size for outer. \hypertarget{cowichan__defaults_8hpp_072a640a55e457c256faa59842e0368c}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!PRODUCT\_\-N@{PRODUCT\_\-N}}
\index{PRODUCT\_\-N@{PRODUCT\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{PRODUCT\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define PRODUCT\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_072a640a55e457c256faa59842e0368c}

Default square matrix size for product. \hypertarget{cowichan__defaults_8hpp_d07476570307484eff173eedfbddd347}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!RAND\_\-M@{RAND\_\-M}}
\index{RAND\_\-M@{RAND\_\-M}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{RAND\_\-M}]{\setlength{\rightskip}{0pt plus 5cm}\#define RAND\_\-M~56197}}
\label{cowichan__defaults_8hpp_d07476570307484eff173eedfbddd347}

Default random number limit. \hypertarget{cowichan__defaults_8hpp_0c6044d9326c14200929af66072b673a}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!RAND\_\-MEAN@{RAND\_\-MEAN}}
\index{RAND\_\-MEAN@{RAND\_\-MEAN}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{RAND\_\-MEAN}]{\setlength{\rightskip}{0pt plus 5cm}\#define RAND\_\-MEAN~0}}
\label{cowichan__defaults_8hpp_0c6044d9326c14200929af66072b673a}

Default random number mean for inputs. \hypertarget{cowichan__defaults_8hpp_40b63d443076ce61004bae3bf79a776b}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!RAND\_\-RANGE@{RAND\_\-RANGE}}
\index{RAND\_\-RANGE@{RAND\_\-RANGE}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{RAND\_\-RANGE}]{\setlength{\rightskip}{0pt plus 5cm}\#define RAND\_\-RANGE~20}}
\label{cowichan__defaults_8hpp_40b63d443076ce61004bae3bf79a776b}

Default random number range for inputs. \hypertarget{cowichan__defaults_8hpp_32ee58f73d9de0d1d41ad46f7bff2719}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!RAND\_\-SEED@{RAND\_\-SEED}}
\index{RAND\_\-SEED@{RAND\_\-SEED}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{RAND\_\-SEED}]{\setlength{\rightskip}{0pt plus 5cm}\#define RAND\_\-SEED~681304}}
\label{cowichan__defaults_8hpp_32ee58f73d9de0d1d41ad46f7bff2719}

Default random number seed. \hypertarget{cowichan__defaults_8hpp_7565d8f613d834da3b98134d4b564017}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!RANDMAT\_\-A@{RANDMAT\_\-A}}
\index{RANDMAT\_\-A@{RANDMAT\_\-A}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{RANDMAT\_\-A}]{\setlength{\rightskip}{0pt plus 5cm}\#define RANDMAT\_\-A~1291}}
\label{cowichan__defaults_8hpp_7565d8f613d834da3b98134d4b564017}

Default constant A for randmat. \hypertarget{cowichan__defaults_8hpp_cacc7f34c70c2ca998f0c1372ed1db83}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!RANDMAT\_\-C@{RANDMAT\_\-C}}
\index{RANDMAT\_\-C@{RANDMAT\_\-C}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{RANDMAT\_\-C}]{\setlength{\rightskip}{0pt plus 5cm}\#define RANDMAT\_\-C~917}}
\label{cowichan__defaults_8hpp_cacc7f34c70c2ca998f0c1372ed1db83}

Default constant C for randmat. \hypertarget{cowichan__defaults_8hpp_be3ee010372d99dfda2913d3d35a6a05}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!RANDMAT\_\-NC@{RANDMAT\_\-NC}}
\index{RANDMAT\_\-NC@{RANDMAT\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{RANDMAT\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define RANDMAT\_\-NC~ALL\_\-NC}}
\label{cowichan__defaults_8hpp_be3ee010372d99dfda2913d3d35a6a05}

Default number of columns for randmat. \hypertarget{cowichan__defaults_8hpp_6b2590531a8313fed76899a95ad3b625}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!RANDMAT\_\-NR@{RANDMAT\_\-NR}}
\index{RANDMAT\_\-NR@{RANDMAT\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{RANDMAT\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define RANDMAT\_\-NR~ALL\_\-NR}}
\label{cowichan__defaults_8hpp_6b2590531a8313fed76899a95ad3b625}

Default number of rows for randmat. \hypertarget{cowichan__defaults_8hpp_3673c55137aa755f4829b6b04329c379}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!SOR\_\-MAX\_\-ITERS@{SOR\_\-MAX\_\-ITERS}}
\index{SOR\_\-MAX\_\-ITERS@{SOR\_\-MAX\_\-ITERS}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{SOR\_\-MAX\_\-ITERS}]{\setlength{\rightskip}{0pt plus 5cm}\#define SOR\_\-MAX\_\-ITERS~1000000}}
\label{cowichan__defaults_8hpp_3673c55137aa755f4829b6b04329c379}

Default maximum number of iterations for sor. \hypertarget{cowichan__defaults_8hpp_23487499c48e5b64d05fd2249be710c1}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!SOR\_\-N@{SOR\_\-N}}
\index{SOR\_\-N@{SOR\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{SOR\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define SOR\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_23487499c48e5b64d05fd2249be710c1}

Default square matrix size for sor. \hypertarget{cowichan__defaults_8hpp_4b7aa2e2b41cee6b710c7ce48115f8ee}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!SOR\_\-OMEGA@{SOR\_\-OMEGA}}
\index{SOR\_\-OMEGA@{SOR\_\-OMEGA}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{SOR\_\-OMEGA}]{\setlength{\rightskip}{0pt plus 5cm}\#define SOR\_\-OMEGA~0.9}}
\label{cowichan__defaults_8hpp_4b7aa2e2b41cee6b710c7ce48115f8ee}

Default omega for sor. \hypertarget{cowichan__defaults_8hpp_e64dba51c415f6acbccbfd021a88769d}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!SOR\_\-TOLERANCE@{SOR\_\-TOLERANCE}}
\index{SOR\_\-TOLERANCE@{SOR\_\-TOLERANCE}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{SOR\_\-TOLERANCE}]{\setlength{\rightskip}{0pt plus 5cm}\#define SOR\_\-TOLERANCE~10e-6}}
\label{cowichan__defaults_8hpp_e64dba51c415f6acbccbfd021a88769d}

Default floating point tolerance for sor. \hypertarget{cowichan__defaults_8hpp_7c356bfeab4913aa2e5dea26de44ebfc}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!THRESH\_\-NC@{THRESH\_\-NC}}
\index{THRESH\_\-NC@{THRESH\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{THRESH\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define THRESH\_\-NC~ALL\_\-NC}}
\label{cowichan__defaults_8hpp_7c356bfeab4913aa2e5dea26de44ebfc}

Default number of columns for thresh. \hypertarget{cowichan__defaults_8hpp_4b80608f6c4617a66eba2f164afc4f3d}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!THRESH\_\-NR@{THRESH\_\-NR}}
\index{THRESH\_\-NR@{THRESH\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{THRESH\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define THRESH\_\-NR~ALL\_\-NR}}
\label{cowichan__defaults_8hpp_4b80608f6c4617a66eba2f164afc4f3d}

Default number of rows for thresh. \hypertarget{cowichan__defaults_8hpp_f85ccfff5fbed47416c64d4229d1234a}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!THRESH\_\-PERCENT@{THRESH\_\-PERCENT}}
\index{THRESH\_\-PERCENT@{THRESH\_\-PERCENT}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{THRESH\_\-PERCENT}]{\setlength{\rightskip}{0pt plus 5cm}\#define THRESH\_\-PERCENT~0.5}}
\label{cowichan__defaults_8hpp_f85ccfff5fbed47416c64d4229d1234a}

Default threshold percentage. \hypertarget{cowichan__defaults_8hpp_d955f5a864096d87337347af906829b1}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!VECDIFF\_\-N@{VECDIFF\_\-N}}
\index{VECDIFF\_\-N@{VECDIFF\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{VECDIFF\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define VECDIFF\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_d955f5a864096d87337347af906829b1}

Default square matrix size for vecdiff. \hypertarget{cowichan__defaults_8hpp_22c87347c9b1f8643a8b57e3d841303f}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!WINNOW\_\-N@{WINNOW\_\-N}}
\index{WINNOW\_\-N@{WINNOW\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{WINNOW\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define WINNOW\_\-N~ALL\_\-N}}
\label{cowichan__defaults_8hpp_22c87347c9b1f8643a8b57e3d841303f}

Default square matrix size for winnow. \hypertarget{cowichan__defaults_8hpp_40f06800a12a97dde304a378d7d971cc}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!WINNOW\_\-NC@{WINNOW\_\-NC}}
\index{WINNOW\_\-NC@{WINNOW\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{WINNOW\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define WINNOW\_\-NC~ALL\_\-NC}}
\label{cowichan__defaults_8hpp_40f06800a12a97dde304a378d7d971cc}

Default rectangular matrix number of columns for winnow. \hypertarget{cowichan__defaults_8hpp_cedd24970400253f48e19a285f8b8559}{
\index{cowichan\_\-defaults.hpp@{cowichan\_\-defaults.hpp}!WINNOW\_\-NR@{WINNOW\_\-NR}}
\index{WINNOW\_\-NR@{WINNOW\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
\subsubsection[{WINNOW\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define WINNOW\_\-NR~ALL\_\-NR}}
\label{cowichan__defaults_8hpp_cedd24970400253f48e19a285f8b8559}

Default rectangular matrix number of rows for winnow.

### 2009-08-02 doc/latex/files.tex
documentation for serial, openmp, tbb

\section{File List}
Here is a list of all documented files with brief descriptions:\begin{CompactList}
\item\contentsline{section}{cowichan/\hyperlink{cowichan_8cpp}{cowichan.cpp} (Implementation for \hyperlink{class_cowichan}{Cowichan} class and common routines for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan_8cpp}}{}
\item\contentsline{section}{cowichan/\hyperlink{cowichan_8hpp}{cowichan.hpp} (Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan_8hpp}}{}
\item\contentsline{section}{cowichan/\hyperlink{cowichan__defaults_8hpp}{cowichan\_\-defaults.hpp} (Defaults for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan__defaults_8hpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_8cpp}{cowichan\_\-openmp.cpp} (This file contains main method that drives the OpenMP implementation )}{\pageref{cowichan__openmp_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_8hpp}{cowichan\_\-openmp.hpp} (Open Multi-Processing (OpenMP) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__openmp_8hpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2gauss_8cpp}{gauss.cpp} (OpenMP gauss implementation )}{\pageref{cowichan__openmp_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2half_8cpp}{half.cpp} (OpenMP half implementation )}{\pageref{cowichan__openmp_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2hull_8cpp}{hull.cpp} (OpenMP hull implementation )}{\pageref{cowichan__openmp_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2invperc_8cpp}{invperc.cpp} (OpenMP (NOT parallel) invperc implementation )}{\pageref{cowichan__openmp_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2life_8cpp}{life.cpp} (OpenMP life implementation )}{\pageref{cowichan__openmp_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2mandel_8cpp}{mandel.cpp} (OpenMP mandel implementation )}{\pageref{cowichan__openmp_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2norm_8cpp}{norm.cpp} (OpenMP norm implementation )}{\pageref{cowichan__openmp_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2outer_8cpp}{outer.cpp} (OpenMP outer implementation )}{\pageref{cowichan__openmp_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2product_8cpp}{product.cpp} (OpenMP product implementation )}{\pageref{cowichan__openmp_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2randmat_8cpp}{randmat.cpp} (OpenMP randmat implementation )}{\pageref{cowichan__openmp_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sor_8cpp}{sor.cpp} (OpenMP sor implementation )}{\pageref{cowichan__openmp_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{sort_8cpp}{sort.cpp} (Implementation of OpenMP sorting algorithms )}{\pageref{sort_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{sort_8hpp}{sort.hpp} (Sorting algorithms that use OpenMP )}{\pageref{sort_8hpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2thresh_8cpp}{thresh.cpp} (OpenMP thresh implementation )}{\pageref{cowichan__openmp_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2vecdiff_8cpp}{vecdiff.cpp} (OpenMP vecdiff implementation )}{\pageref{cowichan__openmp_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2winnow_8cpp}{winnow.cpp} (OpenMP winnow implementation )}{\pageref{cowichan__openmp_2winnow_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_8cpp}{cowichan\_\-serial.cpp} (This file contains main method that drives the serial implementation )}{\pageref{cowichan__serial_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_8hpp}{cowichan\_\-serial.hpp} (Serial implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__serial_8hpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2gauss_8cpp}{gauss.cpp} (Serial gauss implementation )}{\pageref{cowichan__serial_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2half_8cpp}{half.cpp} (Serial half implementation )}{\pageref{cowichan__serial_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2hull_8cpp}{hull.cpp} (Serial hull implementation )}{\pageref{cowichan__serial_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2invperc_8cpp}{invperc.cpp} (Serial invperc implementation )}{\pageref{cowichan__serial_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2life_8cpp}{life.cpp} (Serial life implementation )}{\pageref{cowichan__serial_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2mandel_8cpp}{mandel.cpp} (Serial mandel implementation )}{\pageref{cowichan__serial_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2norm_8cpp}{norm.cpp} (Serial norm implementation )}{\pageref{cowichan__serial_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2outer_8cpp}{outer.cpp} (Serial outer implementation )}{\pageref{cowichan__serial_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2product_8cpp}{product.cpp} (Serial product implementation )}{\pageref{cowichan__serial_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2randmat_8cpp}{randmat.cpp} (Serial randmat implementation )}{\pageref{cowichan__serial_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2sor_8cpp}{sor.cpp} (Serial sor implementation )}{\pageref{cowichan__serial_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2thresh_8cpp}{thresh.cpp} (Serial thresh implementation )}{\pageref{cowichan__serial_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2vecdiff_8cpp}{vecdiff.cpp} (Serial vecdiff implementation )}{\pageref{cowichan__serial_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-serial/\hyperlink{cowichan__serial_2winnow_8cpp}{winnow.cpp} (Serial winnow implementation )}{\pageref{cowichan__serial_2winnow_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_8cpp}{cowichan\_\-tbb.cpp} (This file contains main method that drives the TBB implementation )}{\pageref{cowichan__tbb_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_8hpp}{cowichan\_\-tbb.hpp} (Thread Building Blocks (TBB) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__tbb_8hpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2gauss_8cpp}{gauss.cpp} (TBB gauss implementation )}{\pageref{cowichan__tbb_2gauss_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2half_8cpp}{half.cpp} (TBB half implementation )}{\pageref{cowichan__tbb_2half_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2hull_8cpp}{hull.cpp} (TBB hull implementation )}{\pageref{cowichan__tbb_2hull_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2invperc_8cpp}{invperc.cpp} (TBB (NOT parallel) invperc implementation )}{\pageref{cowichan__tbb_2invperc_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2life_8cpp}{life.cpp} (TBB life implementation )}{\pageref{cowichan__tbb_2life_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2mandel_8cpp}{mandel.cpp} (TBB mandel implementation )}{\pageref{cowichan__tbb_2mandel_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2norm_8cpp}{norm.cpp} (TBB norm implementation )}{\pageref{cowichan__tbb_2norm_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2outer_8cpp}{outer.cpp} (TBB outer implementation )}{\pageref{cowichan__tbb_2outer_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2product_8cpp}{product.cpp} (TBB product implementation )}{\pageref{cowichan__tbb_2product_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2randmat_8cpp}{randmat.cpp} (TBB randmat implementation )}{\pageref{cowichan__tbb_2randmat_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2sor_8cpp}{sor.cpp} (TBB sor implementation )}{\pageref{cowichan__tbb_2sor_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2thresh_8cpp}{thresh.cpp} (TBB thresh implementation )}{\pageref{cowichan__tbb_2thresh_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2vecdiff_8cpp}{vecdiff.cpp} (TBB vecdiff implementation )}{\pageref{cowichan__tbb_2vecdiff_8cpp}}{}
\item\contentsline{section}{cowichan\_\-tbb/\hyperlink{cowichan__tbb_2winnow_8cpp}{winnow.cpp} (TBB winnow implementation )}{\pageref{cowichan__tbb_2winnow_8cpp}}{}
\end{CompactList}

### 2009-08-02 doc/latex/hierarchy.tex
documentation for serial, openmp, tbb

\section{Class Hierarchy}
This inheritance list is sorted roughly, but not completely, alphabetically:\begin{CompactList}
\item \contentsline{section}{Cowichan}{\pageref{class_cowichan}}{}
\begin{CompactList}
\item \contentsline{section}{CowichanOpenMP}{\pageref{class_cowichan_open_m_p}}{}
\item \contentsline{section}{CowichanSerial}{\pageref{class_cowichan_serial}}{}
\item \contentsline{section}{CowichanTBB}{\pageref{class_cowichan_t_b_b}}{}
\end{CompactList}
\item \contentsline{section}{cowichan\_\-tbb::GameOfLife}{\pageref{classcowichan__tbb_1_1_game_of_life}}{}
\item \contentsline{section}{cowichan\_\-tbb::Histogram}{\pageref{classcowichan__tbb_1_1_histogram}}{}
\item \contentsline{section}{cowichan\_\-tbb::MakeDominant}{\pageref{classcowichan__tbb_1_1_make_dominant}}{}
\item \contentsline{section}{cowichan\_\-tbb::Mandelbrot}{\pageref{classcowichan__tbb_1_1_mandelbrot}}{}
\item \contentsline{section}{cowichan\_\-tbb::MaximumDistance}{\pageref{classcowichan__tbb_1_1_maximum_distance}}{}
\item \contentsline{section}{cowichan\_\-tbb::MaxReducer}{\pageref{classcowichan__tbb_1_1_max_reducer}}{}
\item \contentsline{section}{cowichan\_\-tbb::MinMaxReducer}{\pageref{classcowichan__tbb_1_1_min_max_reducer}}{}
\item \contentsline{section}{cowichan\_\-tbb::Normalizer}{\pageref{classcowichan__tbb_1_1_normalizer}}{}
\item \contentsline{section}{cowichan\_\-tbb::NormDiff}{\pageref{classcowichan__tbb_1_1_norm_diff}}{}
\item \contentsline{section}{cowichan\_\-serial::PercPoint}{\pageref{classcowichan__serial_1_1_perc_point}}{}
\item \contentsline{section}{cowichan\_\-tbb::PercPoint}{\pageref{classcowichan__tbb_1_1_perc_point}}{}
\item \contentsline{section}{cowichan\_\-openmp::PercPoint}{\pageref{classcowichan__openmp_1_1_perc_point}}{}
\item \contentsline{section}{Point}{\pageref{class_point}}{}
\item \contentsline{section}{cowichan\_\-tbb::PointCount}{\pageref{classcowichan__tbb_1_1_point_count}}{}
\item \contentsline{section}{cowichan\_\-tbb::PointDistances}{\pageref{classcowichan__tbb_1_1_point_distances}}{}
\item \contentsline{section}{cowichan\_\-tbb::Product}{\pageref{classcowichan__tbb_1_1_product}}{}
\item \contentsline{section}{cowichan\_\-tbb::RandomGenerator}{\pageref{classcowichan__tbb_1_1_random_generator}}{}
\item \contentsline{section}{cowichan\_\-tbb::Relaxer}{\pageref{classcowichan__tbb_1_1_relaxer}}{}
\item \contentsline{section}{cowichan\_\-tbb::RowElimination}{\pageref{classcowichan__tbb_1_1_row_elimination}}{}
\item \contentsline{section}{cowichan\_\-tbb::Shuffle}{\pageref{classcowichan__tbb_1_1_shuffle}}{}
\item \contentsline{section}{cowichan\_\-tbb::Threshold}{\pageref{classcowichan__tbb_1_1_threshold}}{}
\item \contentsline{section}{WeightedPoint}{\pageref{class_weighted_point}}{}
\end{CompactList}

### 2009-08-02 doc/latex/index.tex
documentation for serial, openmp, tbb

\hypertarget{index_intro_sec}{}\section{Introduction}\label{index_intro_sec}
The \hyperlink{class_cowichan}{Cowichan} problems are implemented by various parallel programming systems. Serial C++ implementation is available for comparison. The \hyperlink{class_cowichan}{Cowichan} class is the base class for all C++ implementations.\hypertarget{index_problems_sec}{}\section{The Problems}\label{index_problems_sec}
Currently, there are 14 \hyperlink{class_cowichan}{Cowichan} problems. They are described below. \hyperlink{class_cowichan}{Cowichan} class. Problems can be run separately by passing the problem name as an argument on the command line. The inputs to the problems are defined in \hyperlink{cowichan__defaults_8hpp}{cowichan\_\-defaults.hpp}. There is currently no way to override the inputs without recompiling.\hypertarget{index_mandel_sec}{}\subsection{1. Mandelbrot Set Generation}\label{index_mandel_sec}
This module generates the Mandelbrot Set for a specified region of the complex plane. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{Cowichan::mandel}\end{Desc}
\hypertarget{index_randmat_sec}{}\subsection{2. Random Number Generation}\label{index_randmat_sec}
This module fills a matrix with pseudo-random integers. Note that, as in all problems, the output is required to be independent of the number of processors used. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{Cowichan::randmat}\end{Desc}
\hypertarget{index_half_sec}{}\subsection{3. Two-Dimensional Shuffle}\label{index_half_sec}
This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis, shuffles them, and then repeats this operation along the other axis. Values in odd-numbered locations are collected at the low end of each row or column, while values in even-numbered locations are moved to the high end. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{Cowichan::half}\end{Desc}
\hypertarget{index_invperc_sec}{}\subsection{4. Invasion Percolation}\label{index_invperc_sec}
Invasion percolation models the displacement of one fluid (such as oil) by another (such as water) in fractured rock. In two dimensions, this can be simulated by generating an NxN grid of random numbers in the range \mbox{[}1. . .R\mbox{]}, and then marking the center cell of the grid as filled. In each iteration, one examines the four orthogonal neighbors of all filled cells, chooses the one with the lowest value (i.e. the one with the least resistance to filling), and fills it in. The simulation continues until some fixed percentage of cells have been filled, or until some other condition (such as the presence of trapped regions) is achieved. The fractal structure of the filled and unfilled regions is then examined to determine how much oil could be recovered. The naive way to implement this is to repeatedly scan the array; a more sophisticated, and much faster, sequential technique is to maintain a priority queue of unfilled cells which are neighbors of filled cells. This latter technique is similar to the list-based methods used in some cellular automaton programs, and is very difficult to parallelize effectively. Filling begins at the central cell of the matrix (rounding down for even-sized axes). \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{Cowichan::invperc}\end{Desc}
\hypertarget{index_thresh_sec}{}\subsection{5. Histogram Thresholding}\label{index_thresh_sec}
This module performs histogram thresholding on an image. Given an integer image I and a target percentage p, it constructs a binary image B such that B\mbox{[}i,j\mbox{]} is set if no more than p percent of the pixels in I are brighter than I\mbox{[}i,j\mbox{]}. The general idea is that an image's histogram should have 2 peaks, one centered around the average foreground intensity, and one centered around the average background intensity. This program attempts to set a threshold between the two peaks in the histogram and select the pixels above the threshold. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{Cowichan::thresh}\end{Desc}
\hypertarget{index_life_sec}{}\subsection{6. Game of Life}\label{index_life_sec}
This module simulates the evolution of Conway's Game of Life, a two-dimensional cellular automaton. At each time step, this module must count the number of live (true) neighbors of each cell, using both orthogonal and diagonal connectivity and circular boundary conditions. The update rule is simple: if a cell has 3 live neighbors, or has 2 live neighbors and is already alive, it is alive in the next generation. In any other situation, the cell becomes, or stays, dead. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{Cowichan::life}\end{Desc}
\hypertarget{index_winnow_sec}{}\subsection{7. Weighted Point Selection}\label{index_winnow_sec}
This module converts a matrix of integer values to a vector of points, represented as x and y coordinates. Each location where mask is true becomes a candidate point, with a weight equal to the integer value in matrix at that location and x and y coordinates equal to its row and column indices. These candidate points are then {\bf sorted} into increasing order by weight, and n evenly-spaced points selected to create the result vector. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{Cowichan::winnow}\end{Desc}
\hypertarget{index_norm_sec}{}\subsection{8. Point Location Normalization}\label{index_norm_sec}
This module normalizes point coordinates so that all points lie within the unit square \mbox{[}0. . .1\mbox{]}x\mbox{[}0. . .1\mbox{]}. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{Cowichan::norm}\end{Desc}
\hypertarget{index_hull_sec}{}\subsection{9. Convex Hull}\label{index_hull_sec}
This module takes a list of two-dimensional points and reorders them by doing multiple convex hull computations. Convex hull is the boundary of the minimal convex set containing a given non-empty finite set of points in the plane. In other words, all points not in the convex hull are enclosed in the convex hull polygon. At each step the convex hull points are taken out of the input list and are put into the output list. The computation terminates when there are no more points left in the input list. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{Cowichan::hull}\end{Desc}
\hypertarget{index_outer_sec}{}\subsection{10. Outer Product}\label{index_outer_sec}
This module turns a vector containing point positions into a dense, symmetric, diagonally dominant matrix by calculating the distances between each pair of points. It also constructs a real vector whose values are the distance of each point from the origin. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{Cowichan::outer}\end{Desc}
\hypertarget{index_gauss_sec}{}\subsection{11. Gaussian Elimination}\label{index_gauss_sec}
This module solves a matrix equation AX = V for a dense, symmetric, diagonally dominant matrix A and an arbitrary vector non-zero V using explicit reduction (matrices are required to be symmetric and diagonally dominant in order to guarantee that there is a well-formed solution to the equation). \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{Cowichan::gauss}\end{Desc}
\hypertarget{index_sor_sec}{}\subsection{12. Successive Over-Relaxation}\label{index_sor_sec}
This module solves a matrix equation AX = V for a dense, symmetric, diagonally dominant matrix A and an arbitrary vector non-zero V using successive over-relaxation. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{Cowichan::sor}\end{Desc}
\hypertarget{index_product_sec}{}\subsection{13. Matrix-Vector Product}\label{index_product_sec}
This module calculates V in AX = V where A is a matrix and X is a vector. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan::product}\end{Desc}
\hypertarget{index_vecdiff_sec}{}\subsection{14. 1-Norm Vector Difference}\label{index_vecdiff_sec}
This module calculates the 1-norm of the difference between two vectors. In case the vectors are actual and computed values of some calculation the result represents the magnitude of the error. \begin{Desc}
\item[See also:]\hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{Cowichan::vecdiff} \end{Desc}

### 2009-08-02 doc/latex/namespacecowichan__openmp.tex
documentation for serial, openmp, tbb

\hypertarget{namespacecowichan__openmp}{
\section{cowichan\_\-openmp Namespace Reference}
\label{namespacecowichan__openmp}\index{cowichan\_\-openmp@{cowichan\_\-openmp}}
}
Additional classes and functions specific to openmp implementation.

\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__openmp_1_1_perc_point}{PercPoint}
\begin{CompactList}\small\item\em Percolation point. \item\end{CompactList}\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__openmp_f261011870b201d556308f43b5b25b89}{quickhull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn)
\item 
void \hyperlink{namespacecowichan__openmp_d31787f1b4c9d57119160411e96f769f}{split} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn, \hyperlink{class_point}{Point} $\ast$p1, \hyperlink{class_point}{Point} $\ast$p2)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__openmp_a4ce5b0ba78cd9480bf45e7c313f59d1}{sumNeighbours} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} first, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} r, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} c, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{namespacecowichan__openmp_2753e3856bc41c61099c4702d789823e}{mandel\_\-calc} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} x, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} y)
\item 
void \hyperlink{namespacecowichan__openmp_e96598e37e2b937c861c5bc93eda77b8}{findMinMax} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{Point} $\ast$minPoint, \hyperlink{class_point}{Point} $\ast$maxPoint)
\item 
void \hyperlink{namespacecowichan__openmp_9cbb3540b84f4e051f072a08c2c1d5db}{histogram\_\-sort} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len)
\item 
void \hyperlink{namespacecowichan__openmp_64208258cab6e46db879b46b40581134}{quick\_\-sort} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__openmp_2733c307de545e237e020a1f166850c5}{quick\_\-sort\_\-partition} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} pivotIndex)
\item 
void \hyperlink{namespacecowichan__openmp_9c68e88badede2262dd4ded923e804dc}{mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$buckets)
\end{CompactItemize}

\subsection{Detailed Description}
Additional classes and functions specific to openmp implementation.

\subsection{Function Documentation}
\hypertarget{namespacecowichan__openmp_e96598e37e2b937c861c5bc93eda77b8}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!findMinMax@{findMinMax}}
\index{findMinMax@{findMinMax}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{findMinMax}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::findMinMax ({\bf PointVector} {\em points}, \/  {\bf index\_\-t} {\em n}, \/  {\bf Point} $\ast$ {\em minPoint}, \/  {\bf Point} $\ast$ {\em maxPoint})}}
\label{namespacecowichan__openmp_e96598e37e2b937c861c5bc93eda77b8}

Find min/max x/y coordinates. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]point vector. \item[{\em n}]number of points. \item[{\em minPoint}]min x/y values. \item[{\em maxPoint}]max x/y values. \end{description}
\end{Desc}
\hypertarget{namespacecowichan__openmp_9cbb3540b84f4e051f072a08c2c1d5db}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!histogram\_\-sort@{histogram\_\-sort}}
\index{histogram\_\-sort@{histogram\_\-sort}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{histogram\_\-sort}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::histogram\_\-sort ({\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len})}}
\label{namespacecowichan__openmp_9cbb3540b84f4e051f072a08c2c1d5db}

This performs parallel histogram sort. \hypertarget{namespacecowichan__openmp_2753e3856bc41c61099c4702d789823e}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!mandel\_\-calc@{mandel\_\-calc}}
\index{mandel\_\-calc@{mandel\_\-calc}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{mandel\_\-calc}]{\setlength{\rightskip}{0pt plus 5cm}{\bf INT\_\-TYPE} cowichan\_\-openmp::mandel\_\-calc ({\bf real} {\em x}, \/  {\bf real} {\em y})}}
\label{namespacecowichan__openmp_2753e3856bc41c61099c4702d789823e}

Calculate mandelbrot value. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em x}]x-coordinate. \item[{\em y}]y-coordinate. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Mandelbrot value. \end{Desc}
\hypertarget{namespacecowichan__openmp_9c68e88badede2262dd4ded923e804dc}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!mask\_\-count@{mask\_\-count}}
\index{mask\_\-count@{mask\_\-count}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{mask\_\-count}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::mask\_\-count ({\bf BoolMatrix} {\em mask}, \/  {\bf index\_\-t} {\em nr}, \/  {\bf index\_\-t} {\em nc}, \/  {\bf index\_\-t} $\ast$ {\em buckets})}}
\label{namespacecowichan__openmp_9c68e88badede2262dd4ded923e804dc}

Count the number of set cells in each thread bucket. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em mask}]boolean mask. \item[{\em nr}]number of rows in mask. \item[{\em nc}]number of columns in mask. \item[{\em buckets}]number of cells in each bucket. \end{description}
\end{Desc}
\hypertarget{namespacecowichan__openmp_64208258cab6e46db879b46b40581134}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!quick\_\-sort@{quick\_\-sort}}
\index{quick\_\-sort@{quick\_\-sort}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{quick\_\-sort}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::quick\_\-sort ({\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len})}}
\label{namespacecowichan__openmp_64208258cab6e46db879b46b40581134}

This performs parallel quick sort. \hypertarget{namespacecowichan__openmp_2733c307de545e237e020a1f166850c5}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!quick\_\-sort\_\-partition@{quick\_\-sort\_\-partition}}
\index{quick\_\-sort\_\-partition@{quick\_\-sort\_\-partition}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{quick\_\-sort\_\-partition}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-openmp::quick\_\-sort\_\-partition ({\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len}, \/  {\bf index\_\-t} {\em pivotIndex})}}
\label{namespacecowichan__openmp_2733c307de545e237e020a1f166850c5}

Partitions vector into points less than and greater than the pivot. \hypertarget{namespacecowichan__openmp_f261011870b201d556308f43b5b25b89}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!quickhull@{quickhull}}
\index{quickhull@{quickhull}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{quickhull}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::quickhull ({\bf PointVector} {\em pointsIn}, \/  {\bf index\_\-t} {\em n}, \/  {\bf PointVector} {\em pointsOut}, \/  {\bf index\_\-t} $\ast$ {\em hn})}}
\label{namespacecowichan__openmp_f261011870b201d556308f43b5b25b89}

Runs quickhull algorithm. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]input points. \item[{\em n}]number of inputs points to use. \item[{\em pointsOut}]output points. \item[{\em hn}]number of output points generated so far. \end{description}
\end{Desc}
\hypertarget{namespacecowichan__openmp_d31787f1b4c9d57119160411e96f769f}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!split@{split}}
\index{split@{split}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{split}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-openmp::split ({\bf PointVector} {\em pointsIn}, \/  {\bf index\_\-t} {\em n}, \/  {\bf PointVector} {\em pointsOut}, \/  {\bf index\_\-t} $\ast$ {\em hn}, \/  {\bf Point} $\ast$ {\em p1}, \/  {\bf Point} $\ast$ {\em p2})}}
\label{namespacecowichan__openmp_d31787f1b4c9d57119160411e96f769f}

Recursive step of the quickhull algorithm - compute hull on one side of the splitting line. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]input points. \item[{\em n}]number of inputs points to use. \item[{\em pointsOut}]output points. \item[{\em hn}]number of output points generated so far. \item[{\em p1}]first point of the splitting line (p1,p2). \item[{\em p2}]second point of the splitting line (p1,p2). \end{description}
\end{Desc}
\hypertarget{namespacecowichan__openmp_a4ce5b0ba78cd9480bf45e7c313f59d1}{
\index{cowichan\_\-openmp@{cowichan\_\-openmp}!sumNeighbours@{sumNeighbours}}
\index{sumNeighbours@{sumNeighbours}!cowichan_openmp@{cowichan\_\-openmp}}
\subsubsection[{sumNeighbours}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-openmp::sumNeighbours ({\bf BoolMatrix} {\em first}, \/  {\bf index\_\-t} {\em r}, \/  {\bf index\_\-t} {\em c}, \/  {\bf index\_\-t} {\em nr}, \/  {\bf index\_\-t} {\em nc})}}
\label{namespacecowichan__openmp_a4ce5b0ba78cd9480bf45e7c313f59d1}

Calculate number of peers. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em first}]world matrix. \item[{\em r}]row. \item[{\em c}]column. \item[{\em nr}]number of rows in the matrix. \item[{\em nc}]number of columns in the matrix. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]The number of peers. \end{Desc}

### 2009-08-02 doc/latex/namespaces.tex
documentation for serial, openmp, tbb

\section{Namespace List}
Here is a list of all documented namespaces with brief descriptions:\begin{CompactList}
\item\contentsline{section}{\hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp} (Additional classes and functions specific to openmp implementation )}{\pageref{namespacecowichan__openmp}}{}
\item\contentsline{section}{\hyperlink{namespacecowichan__serial}{cowichan\_\-serial} (Additional classes and functions specific to serial implementation )}{\pageref{namespacecowichan__serial}}{}
\item\contentsline{section}{\hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb} (Additional classes and functions specific to tbb implementation )}{\pageref{namespacecowichan__tbb}}{}
\end{CompactList}

### 2009-08-02 doc/latex/sort_8hpp.tex
documentation for serial, openmp, tbb

\hypertarget{sort_8hpp}{
\section{cowichan\_\-openmp/sort.hpp File Reference}
\label{sort_8hpp}\index{cowichan\_\-openmp/sort.hpp@{cowichan\_\-openmp/sort.hpp}}
}
Sorting algorithms that use OpenMP.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Defines}
\begin{CompactItemize}
\item 
\#define \hyperlink{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{QUICK\_\-SORT\_\-CUTOFF}~100
\item 
\#define \hyperlink{sort_8hpp_66832e055934d011da279261bcfafdd1}{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}~1000
\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__openmp_9cbb3540b84f4e051f072a08c2c1d5db}{cowichan\_\-openmp::histogram\_\-sort} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len)
\item 
void \hyperlink{namespacecowichan__openmp_64208258cab6e46db879b46b40581134}{cowichan\_\-openmp::quick\_\-sort} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__openmp_2733c307de545e237e020a1f166850c5}{cowichan\_\-openmp::quick\_\-sort\_\-partition} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} pivotIndex)
\end{CompactItemize}

\label{_details}
\hypertarget{_details}{}
\subsection{Detailed Description}
Sorting algorithms that use OpenMP.

Two algorithms are included: \begin{enumerate}
\item Parallel Histogram Sort - similar to bucket sort, but does not require additional space for buckets (in-place).  \item Parallel Quick Sort - requires OpenMP 3.0 \end{enumerate}

\subsection{Define Documentation}
\hypertarget{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{
\index{sort.hpp@{sort.hpp}!QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}}
\index{QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}!sort.hpp@{sort.hpp}}
\subsubsection[{QUICK\_\-SORT\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}\#define QUICK\_\-SORT\_\-CUTOFF~100}}
\label{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}

Quick sort cutoff (becomes insertion sort) - must be $>$ 1 \hypertarget{sort_8hpp_66832e055934d011da279261bcfafdd1}{
\index{sort.hpp@{sort.hpp}!QUICK\_\-SORT\_\-TASK\_\-CUTOFF@{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}}
\index{QUICK\_\-SORT\_\-TASK\_\-CUTOFF@{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}!sort.hpp@{sort.hpp}}
\subsubsection[{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}\#define QUICK\_\-SORT\_\-TASK\_\-CUTOFF~1000}}
\label{sort_8hpp_66832e055934d011da279261bcfafdd1}

Parallel sort cutoff (no more tasks generated).

### 2009-08-02 cowichan_openmp/thresh.cpp

@@ -1,3 +1,9 @@
+/**
+ * \file cowichan_openmp/thresh.cpp
+ * \brief OpenMP thresh implementation.
+ * \see CowichanOpenMP::thresh
+ */
+
 #include "cowichan_openmp.hpp"
 
 void CowichanOpenMP::thresh(IntMatrix matrix, BoolMatrix mask) {

### 2009-08-02 cowichan_openmp/vecdiff.cpp

@@ -1,3 +1,9 @@
+/**
+ * \file cowichan_openmp/vecdiff.cpp
+ * \brief OpenMP vecdiff implementation.
+ * \see CowichanOpenMP::vecdiff
+ */
+
 #include "cowichan_openmp.hpp"
 
 real CowichanOpenMP::vecdiff (Vector actual, Vector computed)

### 2009-08-02 cowichan_openmp/winnow.cpp

@@ -1,9 +1,27 @@
+/**
+ * \file cowichan_openmp/winnow.cpp
+ * \brief OpenMP winnow implementation.
+ * \see CowichanOpenMP::winnow
+ */
+
 #include "cowichan_openmp.hpp"
 #include "sort.hpp"
 
+namespace cowichan_openmp
+{
+
+/**
+ * Count the number of set cells in each thread bucket.
+ * \param mask boolean mask.
+ * \param nr number of rows in mask.
+ * \param nc number of columns in mask.
+ * \param buckets number of cells in each bucket.
+ */
 void mask_count(BoolMatrix mask, index_t nr, index_t nc, index_t* buckets);
 
-void not_enough_points();
+}
+
+/*****************************************************************************/
 
 void CowichanOpenMP::winnow(IntMatrix matrix, BoolMatrix mask,
     PointVector points) {
@@ -104,12 +122,11 @@ void CowichanOpenMP::winnow(IntMatrix matrix, BoolMatrix mask,
   delete [] weightedPoints;
 }
 
-/**
- * Count the number of set cells in each thread bucket.
- * @param mask boolean mask.
- * @param nr number of rows.
- * @param nc number of columns.
- */
+/*****************************************************************************/
+
+namespace cowichan_openmp
+{
+
 void mask_count(BoolMatrix mask, index_t nr, index_t nc, index_t* buckets) {
 
   index_t r, c;
@@ -132,4 +149,5 @@ void mask_count(BoolMatrix mask, index_t nr, index_t nc, index_t* buckets) {
 
 }
 
+}
 

### 2009-08-02 cowichan_serial/thresh.cpp

@@ -1,10 +1,10 @@
-#include "cowichan_serial.hpp"
-
 /**
  * \file cowichan_serial/thresh.cpp
  * \brief Serial thresh implementation.
  * \see CowichanSerial::thresh
  */
+
+#include "cowichan_serial.hpp"
 
 void CowichanSerial::thresh(IntMatrix matrix, BoolMatrix mask) {
 

### 2009-08-02 cowichan_serial/vecdiff.cpp

@@ -1,10 +1,10 @@
-#include "cowichan_serial.hpp"
-
 /**
  * \file cowichan_serial/vecdiff.cpp
  * \brief Serial vecdiff implementation.
  * \see CowichanSerial::vecdiff
  */
+
+#include "cowichan_serial.hpp"
 
 real CowichanSerial::vecdiff (Vector actual, Vector computed)
 {

### 2009-08-02 cowichan_serial/winnow.cpp

@@ -1,10 +1,10 @@
-#include "cowichan_serial.hpp"
-
 /**
  * \file cowichan_serial/winnow.cpp
  * \brief Serial winnow implementation.
  * \see CowichanSerial::winnow
  */
+
+#include "cowichan_serial.hpp"
 
 namespace cowichan_serial
 {

### 2009-08-02 cowichan_tbb/life.cpp

@@ -96,6 +96,9 @@ public:
   GameOfLife(BoolMatrix first, BoolMatrix second, index_t nr, index_t nc):
     _first(first), _second(second), nr(nr), nc(nc), aliveCount(0) { }
 
+  /**
+   * Swap the matrices.
+   */
   void swap() {
     BoolMatrix temp = _first;
     _first = _second;

### 2009-08-02 cowichan_tbb/outer.cpp

@@ -94,7 +94,7 @@ public:
 
   /**
    * Joiner (TBB).
-   * \param object to join.
+   * \param other object to join.
    */
   void join(const PointDistances& other) {
     if (_max < other._max) {

### 2009-08-02 cowichan_tbb/thresh.cpp

@@ -1,29 +1,56 @@
+/**
+ * \file cowichan_tbb/thresh.cpp
+ * \brief TBB thresh implementation.
+ * \see CowichanTBB::thresh
+ */
+
 #include "cowichan_tbb.hpp"
 
+namespace cowichan_tbb
+{
+
 /**
- * Performs the maximum computation.
+ * \brief Finds the maximum value in the image.
  */
 class MaxReducer {
 private:
 
+  /**
+   * Image matrix.
+   */
   IntMatrix _image;
+
+  /**
+   * Maximum value.
+   */
   INT_TYPE _max;
+
+  /**
+   * Number of columns in the matrix.
+   */
   const index_t nc;
 
 public:
 
   /**
-   * Initialise max with the lowest possible value.
+   * Constructs a max reducer. Initialise max with the lowest possible value.
+   * \param image image matrix.
+   * \param nc number of matrix columns.
    */
   MaxReducer(IntMatrix image, index_t nc):
     _image(image), _max(MINIMUM_INT), nc(nc) { }
 
+  /**
+   * Get maximum value.
+   * \return Maximum value.
+   */
   INT_TYPE getMaximum() const {
     return _max;
   }
 
   /**
    * Calculates the maximum value over the given range.
+   * \param range row/column range.
    */
   void operator()(const Range2D& range) {
 
@@ -43,13 +70,15 @@ public:
   }
   
   /**
-   * Splitting (TBB) constructor
+   * Splitting (TBB) constructor.
+   * \param other object to split.
    */
   MaxReducer(MaxReducer& other, split) : _image(other._image),
       _max(MINIMUM_INT), nc(other.nc) { }
 
   /**
    * Joiner (TBB).
+   * \param other object to join.
    */
   void join(const MaxReducer& other) {
     if (_max < other._max) {
@@ -59,18 +88,45 @@ public:
   
 };
 
-/*
- * This class calculates the histogram of a matrix.
+/**
+ * \brief This class calculates the histogram of a matrix.
  */
 class Histogram {
   
+  /**
+   * Image matrix.
+   */
   IntMatrix _image;
+  
+  /**
+   * Histogram values.
+   */
   index_t* histogram;
+
+  /** 
+   * Number of bins in the histogram.
+   */
   const index_t bins;
-  const index_t nr, nc;
+  
+  /**
+   * Number of rows in the matrix.
+   */
+  const index_t nr;
+  
+  /**
[…]

### 2009-08-02 cowichan_tbb/vecdiff.cpp

@@ -1,23 +1,55 @@
+/**
+ * \file cowichan_tbb/vecdiff.cpp
+ * \brief TBB vecdiff implementation.
+ * \see CowichanTBB::vecdiff
+ */
+
 #include "cowichan_tbb.hpp"
 
+namespace cowichan_tbb
+{
+
 /**
- * This class takes the 1-norm of the difference between two vectors.
+ * \brief This class takes the 1-norm of the difference between two vectors.
  */
 class NormDiff {
 
+  /**
+   * First (actual) vector.
+   */
   Vector _actual;
+
+  /**
+   * Second (computed) vector.
+   */
   Vector _computed;
+
+  /**
+   * 1-norm difference.
+   */
   real normDiff;
 
 public:
   
+  /**
+   * Get 1-norm difference.
+   */
   real getNormDiff() const {
     return normDiff;
   }
   
+  /**
+   * Construct a norm diff object.
+   * \param actual first (actual) vector.
+   * \param computed second (computed) vector.
+   */
   NormDiff(Vector actual, Vector computed) : _actual(actual),
       _computed(computed), normDiff(0.0) { }
 
+  /**
+   * Perform the computation.
+   * \param range row range.
+   */
   void operator()(const Range& range) {
     Vector actual = _actual;
     Vector computed = _computed;
@@ -35,13 +67,15 @@ public:
   }
   
   /**
-   * Splitting (TBB) constructor
+   * Splitting (TBB) constructor.
+   * \param other object to split.
    */
   NormDiff(NormDiff& other, split):
     _actual(other._actual), _computed(other._computed), normDiff(0.0) { }
   
   /**
    * Joiner (TBB).
+   * \param other object to join.
    */
   void join(const NormDiff& other) {
     if (normDiff < other.normDiff) {
@@ -50,6 +84,8 @@ public:
   }
 
 };
+
+}
 
 /*****************************************************************************/
 

### 2009-08-02 cowichan_tbb/winnow.cpp

@@ -1,36 +1,44 @@
-/*
-  This module converts a matrix of integer values to a vector of points, rep-
-  resented as x and y coordinates. Its inputs are:
+/**
+ * \file cowichan_tbb/winnow.cpp
+ * \brief TBB winnow implementation.
+ * \see CowichanTBB::winnow
+ */
 
-  matrix: an integer matrix, whose values are used as masses.
-  mask: a Boolean matrix showing which points are eligible for consideration.
-  nr, nc: the number of rows and columns in the matrix.
-  n: the number of points to select.
-
-  Its output is:
-
-  points: a vector of (x, y) points.
-
-  Each location where mask is true becomes a candidate point, with a weight
-  equal to the integer value in matrix at that location and x and y coordinates
-  equal to its row and column indices. These candidate points are then sorted
-  into increasing order by weight, and nelts evenly-spaced points selected to
-  create the result vector.
-*/
 #include "cowichan_tbb.hpp"
 
-class ValueCount {
+namespace cowichan_tbb
+{
+
+/**
+ * \brief Calculates the number of points to sort for winnow.
+ */
+class PointCount {
 private:
 
+  /**
+   * Candidate points matrix.
+   */
   IntMatrix _candidates;
+
+  /**
+   * Mask.
+   */
   BoolMatrix _mask;
+
+  /**
+   * Number of columns in the matrix.
+   */
   index_t nc;
+
+  /**
+   * Number of points to sort.
+   */
   index_t count;
 
 public:
 
   /**
-   * Return value count.
+   * Return point count.
    */
   index_t getCount()
   {
@@ -38,13 +46,17 @@ public:
   }
 
   /**
-   * Standard constructor
+   * Construct a point count object.
+   * \param candidates candidate points matrix.
+   * \param mask mask.
+   * \param nc number of columns in the matrix.
    */
-  ValueCount(IntMatrix candidates, BoolMatrix mask, index_t nc):
+  PointCount(IntMatrix candidates, BoolMatrix mask, index_t nc):
     _candidates(candidates), _mask(mask), nc(nc), count(0) { }
 
   /**
-   * Add candidate values to the value list based on mask (TBB).
+   * Calculate number of points to sort (TBB).
+   * \param range row/column range to work on.
    */
   void operator()(const Range2D& range) {
     
@@ -68,30 +80,34 @@ public:
   }
   
   /**
-   * Splitting (TBB) constructor
[…]

### 2009-08-02 cowichan_mpi/gauss/parallel.cpp
removed old mpi implementation

@@ -1,68 +1,0 @@
-/**
- * Parallel implementation of generic gauss
- *
- * \file parallel.cpp
- * \author Andrew Borzenko
- * \date 02-03-09
- */
-
-#include "../include/main.h"
-#include "parallel.h"
-
-void gauss_mpi (mpi::communicator world,
-                real2D*	matrix,			/* to solve */
-                real1D*	vector,			/* target vector */
-                real1D*	answer,			/* solution found */
-                int		n)
-{
-  int		lo, hi;		/* work controls */
-  int		r, c, k;		/* indices */
-  int i;
-
-  // forward elimination
-  for (k=0; k<n; k++){
-    // do work
-    if (get_block_rows_mpi (world, k + 1, n, &lo, &hi)) {
-      // calculate pivots in k'th column
-      for (r = lo; r < hi; r++) {
-        matrix[r][k] = matrix[r][k] / matrix[k][k];
-      }
-      // update elements below k'th row
-      for (r = k + 1; r < n; r++) {
-        for (c = k + 1; c < n; c++) {
-          matrix[r][c] = matrix[r][c] - (matrix[r][k] * matrix[k][c]);
-        }
-      }
-      // update element of solution vector
-      for (r = k + 1; r < n; r++) {
-        vector[r] = vector[r] - (matrix[r][k] * vector[k]);
-      }
-    }
-    // broadcast matrix rows and vector
-    for (i = 0; i < world.size (); i++) {
-      if (get_block_rows_mpi (world, k + 1, n, &lo, &hi, i)) {
-        broadcast (world, matrix[lo], (hi - lo) * n, i);
-        broadcast (world, &vector[lo], hi - lo, i);
-      }
-    }
-  }
-
-  /* back substitution */
-  for (k=(n-1); k>=0; k--){
-    // do work
-    answer[k] = vector[k]/matrix[k][k];
-    if (get_block_rows_mpi (world, 0, k, &lo, &hi)) {
-      for (r = lo; r < hi; r++) {
-        vector[r] = vector[r] - (matrix[r][k] * answer[k]);
-      }
-    }
-    // broadcast vector
-    for (i = 0; i < world.size (); i++) {
-      if (get_block_rows_mpi (world, 0, k, &lo, &hi, i)) {
-        broadcast (world, &vector[lo], hi - lo, i);
-      }
-    }
-  }
-
-  /* return */
-}

### 2009-08-02 cowichan_mpi/half/parallel.cpp

@@ -1,86 +1,0 @@
-/**
- * Parallel implementation of halving shuffle
- *
- * \file parallel.cpp
- * \author Andrew Borzenko
- * \date 02-27-09
- */
-
-#include "../include/main.h"
-#include "parallel.h"
-
-/*
- * @ half : body of halving shuffle
- * > none
- * + shuffle rows and columns
- */
-
-void
-half_mpi (mpi::communicator world,
-  int2D*		matrix,			/* to shuffle */
-  int		nr,			/* row size */
-  int		nc			/* column size */
-){
-  bool work;
-  int		lo, hi;		/* work controls */
-  int		rlo, rhi;		/* work controls */
-  int		r, c, i;		/* loop indices */
-
-  int middle_r = (nr + 1) / 2;
-  int middle_c = (nc + 1) / 2;
-  int previous_r, previous_c;
-
-  int2D* tmp_matrix = new int2D[MAXEXT];
-
-  // work
-  work = get_block_rows_mpi (world, 0, nr, &lo, &hi);
-
-  // shuffle along rows
-  if (work) {
-    for (r = lo; r < hi; r++) {
-      for (c = 0; c < nc; c++) {
-        // get original column
-				if (c < middle_c) {
-					previous_c = c * 2;
-				} else {
-					previous_c = (c - middle_c) * 2 + 1;
-        }
-        tmp_matrix[r][c] = matrix[r][previous_c];
-      }
-    }
-  }
-  
-  // broadcast tmp_matrix rows
-  for (i = 0; i < world.size (); i++) {
-    if (get_block_rows_mpi (world, 0, nr, &rlo, &rhi, i)) {
-      broadcast (world, tmp_matrix[rlo], (rhi - rlo) * nc, i);
-    }
-  }
-
-  // shuffle along columns
-  if (work) {
-    for (r = lo; r < hi; r++) {
-      for (c = 0; c < nc; c++) {
-        // get original row
-				if (r < middle_r) {
-					previous_r = r * 2;
-				} else {
-					previous_r = (r - middle_r) * 2 + 1;
-        }
-        matrix[r][c] = tmp_matrix[previous_r][c];
-      }
-    }
-  }
-
-  // broadcast matrix rows
-  for (i = 0; i < world.size (); i++) {
-    if (get_block_rows_mpi (world, 0, nr, &rlo, &rhi, i)) {
[…]

### 2009-08-02 cowichan_mpi/hull/serial.cpp

@@ -1,84 +1,0 @@
-/**
- * Serial implementation of convex hull
- *
- * \file serial.cpp
- * \author Andrew Borzenko
- * \date 03-06-09
- */
-
-#include "../include/main.h"
-#include "serial.h"
-
-// public
-
-void hull (pt1D* points,     // list of points
-           int n,            // number of points
-           pt1D* hullPoints, // list of points in convex hull
-           int* hn)          // number of points in convex hull
-{
-  int i;
-  pt min_p;
-  pt max_p;
-
-  min_p = points[0];
-  max_p = points[0];
-  *hn = 0;
-
-  // figure out the points with minimum and maximum x values
-  for (i = 1; i < n; i++) {
-    if (min_p.x > points[i].x) {
-      min_p = points[i];
-    }
-    if (max_p.x < points[i].x) {
-      max_p = points[i];
-    }
-  }
-
-  // use these as initial pivots
-  split (points, n, hullPoints, hn, min_p, max_p);
-  split (points, n, hullPoints, hn, max_p, min_p);
-}
-
-// private
-
-void split (pt1D* points,     // list of points
-            int n,            // number of points
-            pt1D* hullPoints, // list of points in convex hull
-            int* hn,          // number of points in convex hull
-            pt p1,            // boundary point #1
-            pt p2)            // boundary point #2
-{
-  pt* maxPoint = NULL;
-  real maxCross = -std::numeric_limits<real>::infinity ();
-
-  // compute the signed distances from the line for each point
-  for (int i = 0; i < n; i++) {
-    real currentCross = cross (p1, p2, points[i]);
-    if (currentCross > maxCross) {
-      maxPoint = &(points[i]);
-      maxCross = currentCross;
-    }
-  }
-
-  // is there a point in the positive half-space?
-  // if so, it has maximal distance, and we must recurse based on that point.
-  if (maxCross > 0.0) {
-    // recurse on the new set with the given far point
-    split (points, n, hullPoints, hn, p1, *maxPoint);
-    split (points, n, hullPoints, hn, *maxPoint, p2);
-    return;
-  } 
-
-  // otherwise, it's not on the right side; we don't need to split anymore.
-  // this is because all points are inside the hull when we use this half-space.
-  // add the first point and return.
-  hullPoints[(*hn)++] = p1;
-}
-
-/**
[…]

### 2009-08-02 cowichan_mpi/include/util.h

@@ -1,62 +1,0 @@
-/**
- * Utility functions
- *
- * \file util.h
- * \author Andrew Borzenko
- * \date 01-26-09
- */
-
-void print_matrix (bool2D* matrix, int nr, int nc);
-void print_matrix (int2D* matrix, int nr, int nc);
-void print_matrix (real2D* matrix, int nr, int nc);
-void print_matrix (bool1DX* matrix, int nr, int nc);
-void print_matrix (int1DX* matrix, int nr, int nc);
-void print_matrix (real1DX* matrix, int nr, int nc);
-void print_vector (bool1D* vector, int nr);
-void print_vector (int1D* vector, int nr);
-void print_vector (real1D* vector, int nr);
-void print_vector (pt1D* vector, int nr);
-void print_points (pt1D* vector, int nr, int limit);
-
-bool get_block_rows_mpi (mpi::communicator world, int lo, int hi,
-                         int* start, int* end);
-bool get_block_rows_mpi (mpi::communicator world, int lo, int hi,
-                         int* start, int* end, int rank);
-int get_block_rank_mpi (mpi::communicator world, int lo, int hi,
-                        int row);
-bool get_cyclic_rows_mpi (mpi::communicator world, int lo, int hi,
-                         int* start, int* end, int* stride);
-int get_cyclic_rank_mpi (mpi::communicator world, int lo, int hi,
-                        int element);
-
-void randStateInit (unsigned int seed,       /* RNG seed */
-                    int width,      /* number of participants */
-                    unsigned int	* state,    /* per-thread state vector */
-                    unsigned int	* aPrime,   /* new multiplicative */
-                    unsigned int	* cPrime);  /* new additive value */
-
-void redPt1DPos (pt1D*		vec,			/* vector of points */
-                 int		n,			/* number of points */
-                 pt	      * ptMin,			/* minimum location */
-                 pt	      * ptMax);			/* maximum location */
-
-void ptSort(pt	      * ptVec,			/* points to sort */
-            int		len);			/* length of vectors */
-
-int ptCmp(pt	      * left,			/* left point */
-          pt	      * right);			/* right point */
-
-void intSort(int  * vec,			/* to sort */
-             int		len			/* length */
-);
-
-real ptDist(pt	      * left,			/* left point */
-            pt	      * right);			/* right point */
-
-real ptMag(pt	      * p);			/* point */
-
-INT64 get_ticks (); // tick count
[…]

### 2009-08-02 cowichan_mpi/invperc/invperc.cpp

@@ -1,84 +1,0 @@
-/**
- * Invasion percolation
- *
- * \file vecdiff.cpp
- * \author Andrew Borzenko
- * \date 02-28-09
- */
-
-#include "../include/main.h"
-#ifdef IS_PARALLEL
-  #include "parallel.h"
-#else
-  #include "serial.h"
-#endif
-
-int main(int argc, char* argv[])
-{
-#ifdef IS_PARALLEL
-  mpi::environment env(argc, argv);
-  mpi::communicator world;
-
-#ifdef TEST_OUTPUT
-  printf ("I am process %d\n", world.rank ());
-#endif
-#endif
-
-  int2D* matrix;
-  bool2D* mask;
-  int nr, nc;
-  int limit;
-  real fraction;
-  int i, j;
-
-  nr = MAXEXT;
-  nc = MAXEXT;
-  limit = 10;
-  fraction = 0.75;
-
-  srand (222);
-
-  matrix = new int2D[MAXEXT];
-  for (i = 0; i < nr; i++) {
-    for (j = 0; j < nc; j++) {
-      matrix[i][j] = rand () % limit;
-    }
-  }
-
-  mask = new bool2D[MAXEXT];
-  memset (mask, 0, sizeof(boolean) * nr * nc);
-
-#ifdef TEST_OUTPUT
-  printf ("Matrix:\n");
-  print_matrix (matrix, nr, nc);
-
-  printf ("Mask before:\n");
-  print_matrix (mask, nr, nc);
-#endif
-
-#ifdef TEST_TIME
-  INT64 start, end;
-  start = get_ticks ();
-#endif
-
-#ifdef IS_PARALLEL
-  invperc_mpi (world, matrix, mask, nr, nc, fraction);
-#else
-  invperc (matrix, mask, nr, nc, fraction);
-#endif
-
-#ifdef TEST_TIME
-  end = get_ticks ();
-  print_elapsed_time (start, end);
-#endif
-
-#ifdef TEST_OUTPUT
-  printf ("Mask after:\n");
-  print_matrix (mask, nr, nc);
-#endif
-
-  delete [] matrix;
-  delete [] mask;
-
-  return 0;
-}

### 2009-08-02 cowichan_mpi/invperc/parallel.h

@@ -1,65 +1,0 @@
-/**
- * Parallel implementation of invasion percolation
- *
- * \file parallel.h
- * \author Andrew Borzenko
- * \date 02-28-09
- */
-
-#pragma once
-#ifndef INVPERC_PARALLEL_H
-#define INVPERC_PARALLEL_H
-
-/*--------------------------------------------------------------*/
-/* private types						*/
-/*--------------------------------------------------------------*/
-
-typedef struct node_struct  node_t;
-typedef struct node_struct* node_p;
-
-struct node_struct {
-  int		val;			/* matrix value */
-  int		r, c;			/* location indices */
-};
-
-typedef std::multimap<int, node_p> mapNodesType;
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-void invperc_mpi (mpi::communicator world,
-                  int2D*		matrix,			/* matrix to invade */
-                  bool2D*	mask,			/* mask to fill */
-                  int		nr,			/* row size */
-                  int		nc,			/* column size */
-                  real		fraction);		/* how much to fill */
-
-/*--------------------------------------------------------------*/
-/* private functions 					*/
-/*--------------------------------------------------------------*/
-
-void inv_deq_mpi (mpi::communicator world,
-                  mapNodesType* nodes,
-                  int	      * r,			/* row index */
-                  int	      * c);			/* column index */
-
-void inv_enq_mpi (mpi::communicator world,
-                  mapNodesType* nodes,
-                  node_p	node);			/* what to enqueue */
-
-node_p inv_node_mpi (mpi::communicator world,
-                     int		val,			/* location value */
-                     int		r,			/* row index */
-                     int		c);			/* column index */
-
-void inv_enqPt_mpi (mpi::communicator world,
-                    mapNodesType* nodes,
-                    int2D*		matrix,			/* matrix of values */
-                    bool2D*	mask,			/* mask to be filled */
-                    int		nr,			/* number of rows */
-                    int		nc,			/* number of columns */
-                    int		r,			/* point row */
-                    int		c);			/* point column */
-
-#endif /* INVPERC_PARALLEL_H */

### 2009-08-02 cowichan_mpi/invperc/parallel_naive.cpp

@@ -1,71 +1,0 @@
-/**
- * Parallel implementation of invasion percolation
- *
- * \file parallel.cpp
- * \author Andrew Borzenko
- * \date 02-28-09
- */
-
-#include "../include/main.h"
-#include "parallel.h"
-
-// public
-
-void
-invperc_mpi (mpi::communicator world,
-  int2D*		matrix,			/* matrix to invade */
-  bool2D*	mask,			/* mask to fill */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  real		fraction		/* how much to fill */
-){
-  int		r, c;			/* row and column indices */
-  int		num, i;			/* filling index */
-  int min_r, min_c; // row and column of minimum value
-  int min_v; // minimum value
-
-  // init
-  num = (int)(fraction * nr * nc);
-
-  mask[nr / 2][nc / 2] = TRUE;
-
-  // fill
-  for (i = 0; i < num; i++) {
-    min_v = -1;
-    for (r = 0; r < nr; r++) {
-      for (c = 0; c < nc; c++) {
-        if ((mask[r][c] != TRUE) && (is_connected (mask, r, c, nr, nc))) {
-          if ((matrix[r][c] < min_v) || (min_v == -1)) {
-            min_r = r;
-            min_c = c;
-            min_v = matrix[r][c];
-          }
-        }
-      }
-    }
-    if (min_v != -1) {
-      mask[min_r][min_c] = TRUE;
-    }
-  }
-
-  /* return */
-}
-
-// private
-
-bool is_connected (bool2D* mask, int r, int c, int nr, int nc)
-{
-  if ((r - 1 >= 0) && (mask[r - 1][c])) {
-    return true;
-  }
-  else if ((r + 1 < nr) && (mask[r + 1][c])) {
-    return true;
-  }
-  else if ((c - 1 >= 0) && (mask[r][c - 1])) {
-    return true;
-  }
-  else if ((c + 1 < nc) && (mask[r][c + 1])) {
-    return true;
-  }
-  return false;
-}

### 2009-08-02 cowichan_mpi/invperc/parallel_naive.h

@@ -1,32 +1,0 @@
-/**
- * Parallel implementation of invasion percolation
- *
- * \file parallel.h
- * \author Andrew Borzenko
- * \date 02-28-09
- */
-
-#pragma once
-#ifndef PARALLEL_H
-#define PARALLEL_H
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-void
-invperc_mpi (mpi::communicator world,
-  int2D*		matrix,			/* matrix to invade */
-  bool2D*	mask,			/* mask to fill */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  real		fraction		/* how much to fill */
-);
-
-/*--------------------------------------------------------------*/
-/* private functions						*/
-/*--------------------------------------------------------------*/
-
-bool is_connected (bool2D* mask, int r, int c, int nr, int nc);
-
-#endif /* PARALLEL_H */

### 2009-08-02 cowichan_mpi/life/parallel.cpp

@@ -1,114 +1,0 @@
-/**
- * Parallel implementation of conway's game of life
- *
- * \file parallel.cpp
- * \author Andrew Borzenko
- * \date 01-26-09
- */
-
-#include "../include/main.h"
-#include "parallel.h"
-
-// public
-
-void
-life_mpi(
-  mpi::communicator world,
-  bool2D*	matrix,          // world to evolve
-  int nr,                  // row size
-  int nc,                  // column size
-  int iters                // number of iterations
-){
-  int i;                   // iteration index
-  int r;                   // row index
-  int alive;               // number alive
-  int lo, hi;              // work controls
-  int rlo, rhi;            // for broadcast
-  bool work;               // useful work to do?
-  int is_alive = 1;        // some cells still alive?
-  bool2D* matrix2;         // new matrix
-  bool2D* m_tmp;           // tmp pointer
-
-  // initialize
-  matrix2 = new bool2D[MAXEXT];
-
-  // work
-  work = get_block_rows_mpi (world, 0, nr, &lo, &hi);
-  for (i = 0; (i < iters) && (is_alive > 0); i++) {
-    // reset alive neighbour count
-    alive = 0;
-
-    // count neighbours and fill new matrix
-    if (work) {
-      for (r = lo; r < hi; r++) {
-        life_row_mpi(matrix, matrix2, &alive, nr, nc, r,
-          (nr + r - 1) % nr, (nr + r + 1) % nr);
-      }
-    }
-    // broadcast matrix
-    for (r = 0; r < world.size (); r++) {
-      if (get_block_rows_mpi (world, 0, nr, &rlo, &rhi, r)) {
-        broadcast (world, matrix2[rlo], (rhi - rlo) * nc, r);
-      }
-    }
-    // is_alive is maximum of local alive's
-    all_reduce (world, alive, is_alive, mpi::maximum<int>());
-    m_tmp = matrix;
-    matrix = matrix2;
-    matrix2 = m_tmp;
-  }
-
-  delete [] matrix2;
-
-  // check
-  ASSERT (is_alive > 0);
-}
-
-// private
-
-void
-life_one_mpi(
-  bool2D*       matrix,                 // old world
-  bool2D*       matrix2,                // new world
-  int*          alive,                  // alive count
-  int           r,                      // this row
-  int           r_lo,                   // lower row
-  int           r_hi,                   // higher row
-  int           c,                      // this column
[…]

### 2009-08-02 cowichan_mpi/life/parallel.h

@@ -1,61 +1,0 @@
-/**
- * Parallel implementation of conway's game of life
- *
- * \file parallel.h
- * \author Andrew Borzenko
- * \date 01-26-09
- */
-
-#pragma once
-#ifndef LIFE_PARALLEL_H
-#define LIFE_PARALLEL_H
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-/*
- * @ life : simulate Game of Life
- * > none
- * + evolve world
- */
-
-void
-life_mpi(
-  mpi::communicator world,
-  bool2D*	matrix,			/* world to evolve */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  int		iters			/* number of iterations */
-);
-
-#endif /* LIFE_PARALLEL_H */
-
-/*--------------------------------------------------------------*/
-/* private functions						*/
-/*--------------------------------------------------------------*/
-
-void
-life_one_mpi(
-  bool2D*       matrix,                 // old world
-  bool2D*       matrix2,                // new world
-  int*          alive,                  // alive count
-  int           r,                      // this row
-  int           r_lo,                   // lower row
-  int           r_hi,                   // higher row
-  int           c,                      // this column
-  int           c_lo,                   // lower column
-  int           c_hi                    // higher column
-);
-
-void
-life_row_mpi(
-  bool2D*       matrix,                 // old world
-  bool2D*       matrix2,                // new world
-  int*          alive,                  // alive count
-  int           nr,                     // row size
-  int           nc,                     // column size
-  int           r,                      // this row
-  int           r_lo,                   // lower row
-  int           r_hi                    // higher row
-);

### 2009-08-02 cowichan_mpi/life/serial.cpp

@@ -1,95 +1,0 @@
-/**
- * Serial implementation of conway's game of life
- *
- * \file serial.cpp
- * \author Andrew Borzenko
- * \date 01-26-09
- */
-
-#include "../include/main.h"
-#include "serial.h"
-
-// public
-
-void
-life(
-  bool2D*	world,			/* world to evolve */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  int		iters			/* number of iterations */
-){
-  int2D*		count;			/* neighborhood counts */
-  int		i;			/* iteration index */
-  int		r, c;			/* row/column indices */
-  int		alive = nr * nc;	/* number alive */
-
-  count = new int2D[MAXEXT];
-
-  for (i=0; (i<iters) && alive; i++){
-    /* fill neighborhood counts */
-    life_row(world, count, nr, nc, 0, nr-1, 1);
-    for (r=1; r<(nr-1); r++){
-      life_row(world, count, nr, nc, r, r-1, r+1);
-    }
-    life_row(world, count, nr, nc, nr-1, nr-2, 0);
-    /* update cells */
-    alive = 0;
-    for (r=0; r<nr; r++){
-      for (c=0; c<nc; c++){
-        if ((count[r][c] == 3) || ((count[r][c] == 2) && world[r][c])){
-          world[r][c] = TRUE;
-          alive += 1;
-        } else {
-          world[r][c] = FALSE;
-        }
-      }
-    }
-  }
-
-  delete [] count;
-
-  /* check */
-  ASSERT (alive != 0);
-
-  /* return */
-}
-
-// private
-
-void
-life_one(
-  bool2D*	world,			/* world to evolve */
-  int2D*		count,			/* neighborhood counts */
-  int		r,			/* this row */
-  int		r_lo,			/* lower row */
-  int		r_hi,			/* higher row */
-  int		c,			/* this column */
-  int		c_lo,			/* lower column */
-  int		c_hi			/* higher column */
-){
-  count[r][c] = world[r_lo][c_lo] + world[r_lo][c] + world[r_lo][c_hi]
-	      + world[r][c_lo]            +          world[r][c_hi]
-	      + world[r_hi][c_lo] + world[r_hi][c] + world[r_hi][c_hi];
-  /* return */
-}
-
-void
-life_row(
-  bool2D*	world,			/* world to evolve */
-  int2D*		count,			/* neighborhood counts */
[…]

### 2009-08-02 cowichan_mpi/mandel/mandel.cpp

@@ -1,65 +1,0 @@
-/**
- * Mandelbrot set implementation
- *
- * \file mandel.cpp
- * \author Andrew Borzenko
- * \date 02-09-09
- */
-
-#include "../include/main.h"
-#ifdef IS_PARALLEL
-  #include "parallel.h"
-#else
-  #include "serial.h"
-#endif
-
-int main(int argc, char* argv[])
-{
-#ifdef IS_PARALLEL
-  mpi::environment env(argc, argv);
-  mpi::communicator world;
-
-#ifdef TEST_OUTPUT
-  printf ("I am process %d\n", world.rank ());
-#endif
-#endif
-
-  int2D*	    matrix;			/* matrix to fill */
-  int		nr, nc;			/* matrix size */
-  real       base_x, base_y;
-  real       ext_x, ext_y;
-
-  nr = MAXEXT;
-  nc = MAXEXT;
-  base_x = 0;
-  base_y = 0;
-  ext_x = 1.5;
-  ext_y = 1.5;
-
-  matrix = new int2D[MAXEXT];
-
-#ifdef TEST_TIME
-  INT64 start, end;
-  start = get_ticks ();
-#endif
-
-#ifdef IS_PARALLEL
-  mandel_mpi (world, matrix, nr, nc, base_x, base_y, ext_x, ext_y);
-#else
-  mandel (matrix, nr, nc, base_x, base_y, ext_x, ext_y);
-#endif
-
-#ifdef TEST_TIME
-  end = get_ticks ();
-  print_elapsed_time (start, end);
-#endif
-
-#ifdef TEST_OUTPUT
-  printf ("Mandelbrot set:\n");
-  print_matrix (matrix, nr, nc);
-#endif
-
-  delete [] matrix;
-
-  return 0;
-}

### 2009-08-02 cowichan_mpi/norm/parallel.cpp

@@ -1,54 +1,0 @@
-/**
- * Parallel implementation of vector norm
- *
- * \file parallel.cpp
- * \author Andrew Borzenko
- * \date 02-10-09
- */
-
-#include "../include/main.h"
-#include "parallel.h"
-
-void norm_mpi (mpi::communicator world,
-               pt1D* vec,    // points to normalize
-               int  n)       // length of vector
-{
-  pt ptMin_local, ptMax_local; //pseudo-points
-  pt ptMin, ptMax; // pseudo-points
-  real sclX, sclY; // scaling factors
-  int i; // loop index
-  int lo, hi; // work controls
-  bool work; // useful work to do?
-
-  // initialize
-  ptMin_local = vec[0];
-  ptMax_local = vec[0];
-
-  work = get_block_rows_mpi (world, 0, n, &lo, &hi);
-  if (work) {
-    redPt1DPos(&vec[lo], hi - lo, &ptMin_local, &ptMax_local);
-  }
-
-  all_reduce (world, ptMin_local, ptMin, minimum_pt ());
-  all_reduce (world, ptMax_local, ptMax, maximum_pt ());
-
-  if (work) {
-    // scaling factors
-    sclX = (ptMax.x == ptMin.x) ? 0.0 : 1/(ptMax.x - ptMin.x);
-    sclY = (ptMax.y == ptMin.y) ? 0.0 : 1/(ptMax.y - ptMin.y);
-    // scale
-    for (i = lo; i < hi; i++) {
-      vec[i].x = sclX * (vec[i].x - ptMin.x);
-      vec[i].y = sclY * (vec[i].y - ptMin.y);
-    }
-  }
-
-  // broadcast normalized values
-  for (i = 0; i < world.size (); i++) {
-    if (get_block_rows_mpi (world, 0, n, &lo, &hi, i)) {
-      broadcast (world, &vec[lo], hi - lo, i);
-    }
-  }
-
-  /* return */
-}

### 2009-08-02 cowichan_mpi/outer/serial.cpp

@@ -1,50 +1,0 @@
-/**
- * Serial implementation of outer product matrix
- *
- * \file serial.cpp
- * \author Andrew Borzenko
- * \date 02-27-09
- */
-
-#include "../include/main.h"
-#include "serial.h"
-
-// public
-
-/*
- * @ outer : calculate outer product matrix
- * > none
- * + fill matrix
- */
-
-void
-outer(
-  pt1D*		ptVec,			/* vector of points */
-  real2D*	matrix,			/* matrix to fill */
-  real1D*	realVec,		/* vector to fill */
-  int		n			/* size */
-){
-  real		d;			/* distance between points */
-  real		dMax = -1.0;		/* maximum distance */
-  int		r, c;			/* loop indices */
-
-  /* all elements except matrix diagonal */
-  for (r=0; r<n; r++){
-    realVec[r] = ptMag(&(ptVec[r]));
-    for (c=0; c<r; c++){
-      d = ptDist(&(ptVec[r]), &(ptVec[c]));
-      if (d > dMax){
-        dMax = d;
-      }
-      matrix[r][c] = matrix[c][r] = d;
-    }
-  }
-
-  /* matrix diagonal */
-  dMax *= n;
-  for (r=0; r<n; r++){
-    matrix[r][r] = dMax;
-  }
-
-  /* return */
-}

### 2009-08-02 cowichan_mpi/sor/parallel.h

@@ -1,24 +1,0 @@
-/**
- * Parallel implementation of successive over-relaxation
- *
- * \file parallel.h
- * \author Andrew Borzenko
- * \date 03-02-09
- */
-
-#pragma once
-#ifndef SOR_PARALLEL_H
-#define SOR_PARALLEL_H
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-void sor_mpi (mpi::communicator world,
-              real2D*	matrix,			/* to solve */
-              real1D*	vector,			/* target vector */
-              real1D*	answer,			/* solution found */
-              int		n,			/* size */
-              real		tol);			/* tolerance on answer */
-
-#endif /* SOR_PARALLEL_H */

### 2009-08-02 cowichan_mpi/sor/serial.cpp

@@ -1,59 +1,0 @@
-/**
- * Serial implementation of successive over-relaxation
- *
- * \file serial.cpp
- * \author Andrew Borzenko
- * \date 03-02-09
- */
-
-#include "../include/main.h"
-#include "serial.h"
-
-// public
-
-/*
- * @ sor : do successive over-relaxation
- * > none
- * + relax matrix to find solution
- */
-
-void
-sor(
-  real2D*	matrix,			/* to solve */
-  real1D*	vector,			/* target vector */
-  real1D*	answer,			/* solution found */
-  int		n,			/* size */
-  real		tol			/* tolerance on answer */
-){
-  int		r, c, t;		/* indices */
-  real		sum, old, dMax, d;	/* temporaries */
-
-  /* initialize */
-  for (r=0; r<n; r++){
-    answer[r] = 1.0;
-  }
-  dMax = 2 * tol;			/* to forestall early exit */
-
-  for (t=0; (t<SOR_MAX_ITERS) && (dMax >= tol); t++){
-    dMax = 0.0;
-    for (r=0; r<n; r++){
-      /* compute sum */
-      sum = 0.0;
-      for (c=0; c<r; c++){
-        sum += matrix[r][c]*answer[c];
-      }
-      for (c=r+1; c<n; c++){
-        sum += matrix[r][c]*answer[c];
-      }
-    
-      /* compute difference */
-      old = answer[r];
-      answer[r] = (1.0-SOR_OMEGA)*old + SOR_OMEGA*(vector[r]-sum)/matrix[r][r];
-      d = (real)fabs((double)(old - answer[r]));
-      if (d > dMax){
-        dMax = d;
-      }
-    }
-  }
-  /* return */
-}

### 2009-08-02 cowichan_mpi/sor/sor.cpp

@@ -1,95 +1,0 @@
-/**
- * Successive over-relaxation
- *
- * \file sor.cpp
- * \author Andrew Borzenko
- * \date 03-02-09
- */
-
-#include "../include/main.h"
-#ifdef IS_PARALLEL
-  #include "parallel.h"
-#else
-  #include "serial.h"
-#endif
-
-int main(int argc, char* argv[])
-{
-#ifdef IS_PARALLEL
-  mpi::environment env(argc, argv);
-  mpi::communicator world;
-
-#ifdef TEST_OUTPUT
-  printf ("I am process %d\n", world.rank ());
-#endif
-#endif
-
-  real2D* matrix;
-  real1D* vector;
-  real1D* answer;
-  int n;
-  int limit;
-  int i, j;
-  real tolerance;
-  real max;
-
-  n = MAXEXT;
-  limit = 10;
-  tolerance = 10e-6;
-  max = -1.0;
-
-  srand (333);
-
-  matrix = new real2D[MAXEXT];
-  for (i = 0; i < n; i++) {
-    for (j = 0; j < n; j++) {
-      matrix[i][j] = rand () % limit;
-      if (matrix[i][j] > max) {
-        max = matrix[i][j];
-      }
-    }
-  }
-  // ensure diagonally dominant
-  max *= n;
-  for (i = 0; i < n; i++) {
-    matrix[i][i] = max;
-  }
-
-  vector = new real1D[MAXEXT];
-  for (i = 0; i < n; i++) {
-    vector[i] = rand () % limit;
-  }
-
-  answer = new real1D[MAXEXT];
-
-#ifdef TEST_OUTPUT
-  printf ("Matrix is:\n");
-  print_matrix (matrix, n, n);
-
-  printf ("Vector is:\n");
-  print_vector (vector, n);
-#endif
-
-#ifdef TEST_TIME
-  INT64 start, end;
-  start = get_ticks ();
-#endif
-
-#ifdef IS_PARALLEL
-  sor_mpi (world, matrix, vector, answer, n, tolerance);
-#else
-  sor (matrix, vector, answer, n, tolerance);
-#endif
-
-#ifdef TEST_TIME
-  end = get_ticks ();
-  print_elapsed_time (start, end);
-#endif
-
-#ifdef TEST_OUTPUT
-  printf ("Answer is:\n");
-  print_vector (answer, n);
-#endif
-
-  return 0;
-}

### 2009-08-02 cowichan_mpi/thresh/parallel.h

@@ -1,34 +1,0 @@
-/**
- * Parallel implementation of histogram thresholding
- *
- * \file parallel.h
- * \author Andrew Borzenko
- * \date 03-02-09
- */
-
-#pragma once
-#ifndef THRESH_PARALLEL_H
-#define THRESH_PARALLEL_H
-
-template<typename T>
-struct sum {
-  typedef T result_type;
-
-  T operator()(T a, T b)
-  {
-    return a + b;
-  }
-};
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-void thresh_mpi (mpi::communicator world,
-                 int2D*		matrix,			/* to threshold */
-                 bool2D*	mask,			/* threshold mask */
-                 int		nr,			/* row size */
-                 int		nc,			/* column size */
-                 real		fraction);		/* how much to keep */
-
-#endif /* THRESH_PARALLEL_H */

### 2009-08-02 cowichan_mpi/thresh/serial.h

@@ -1,26 +1,0 @@
-/**
- * Serial implementation of histogram thresholding
- *
- * \file serial.h
- * \author Andrew Borzenko
- * \date 03-02-09
- */
-
-#pragma once
-#ifndef THRESH_SERIAL_H
-#define THRESH_SERIAL_H
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-void
-thresh(
-  int2D*		matrix,			/* to threshold */
-  bool2D*	mask,			/* threshold mask */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  real		fraction		/* how much to keep */
-);
-
-#endif /* THRESH_SERIAL_H */

### 2009-08-02 cowichan_mpi/thresh/thresh.cpp

@@ -1,80 +1,0 @@
-/**
- * Histogram thresholding
- *
- * \file vecdiff.cpp
- * \author Andrew Borzenko
- * \date 03-02-09
- */
-
-#include "../include/main.h"
-#ifdef IS_PARALLEL
-  #include "parallel.h"
-#else
-  #include "serial.h"
-#endif
-
-int main(int argc, char* argv[])
-{
-#ifdef IS_PARALLEL
-  mpi::environment env(argc, argv);
-  mpi::communicator world;
-
-#ifdef TEST_OUTPUT
-  printf ("I am process %d\n", world.rank ());
-#endif
-#endif
-
-  int2D* matrix;
-  bool2D* mask;
-  int nr, nc;
-  real fraction;
-  int i, j;
-  int limit;
-
-  nr = MAXEXT;
-  nc = MAXEXT;
-  limit = 10;
-  fraction = 0.5;
-
-  srand (222);
-
-  matrix = new int2D[MAXEXT];
-  for (i = 0; i < nr; i++) {
-    for (j = 0; j < nc; j++) {
-      matrix[i][j] = rand () % limit;
-    }
-  }
-
-  mask = new bool2D[MAXEXT];
-
-#ifdef TEST_OUTPUT
-  printf ("Matrix is:\n");
-  print_matrix (matrix, nr, nc);
-#endif
-
-#ifdef TEST_TIME
-  INT64 start, end;
-  start = get_ticks ();
-#endif
-
-#ifdef IS_PARALLEL
-  thresh_mpi (world, matrix, mask, nr, nc, fraction);
-#else
-  thresh (matrix, mask, nr, nc, fraction);
-#endif
-
-#ifdef TEST_TIME
-  end = get_ticks ();
-  print_elapsed_time (start, end);
-#endif
-
-#ifdef TEST_OUTPUT
-  printf ("Mask is:\n");
-  print_matrix (mask, nr, nc);
-#endif
-
-  delete [] matrix;
-  delete [] mask;
-
-  return 0;
-}

### 2009-08-02 cowichan_mpi/vecdiff/parallel.h

@@ -1,25 +1,0 @@
-/**
- * Parallel implementation of vector difference
- *
- * \file parallel.h
- * \author Andrew Borzenko
- * \date 02-27-09
- */
-
-#pragma once
-#ifndef VECDIFF_PARALLEL_H
-#define VECDIFF_PARALLEL_H
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-void
-vecdiff_mpi(mpi::communicator world,
-  real1D*	left,			/* left vector */
-  real1D*	right,			/* right vector */
-  int		n,			/* vector lengths */
-  real	      * diff			/* norm-1 difference */
-);
-
-#endif /* VECDIFF_PARALLEL_H */

### 2009-08-02 cowichan_mpi/winnow/parallel.cpp

@@ -1,610 +1,0 @@
-/**
- * Parallel implementation of weighted point selection
- *
- * \file parallel.cpp
- * \author Andrew Borzenko
- * \date 02-23-09
- */
-
-#include "../include/main.h"
-#include "parallel.h"
-
-// private
-
-static int1D*	Totals;		/* row totals */
-static pt*	TmpPt;		/* temporary storage vector */
-static pt*	TmpPt2;		/* another temporary storage vector */
-static int*	Pivots;	/* pivot elements */
-static int*	Counts;	/* number of elements in each interval */
-static int*	Starts;	/* starts of sub-sections */
-static int	Len;		/* number of points */
-static int	NumSamples;	/* total number of samples */
-static int	SectionSize;	/* size of individual sections */
-static int	IntervalSize;	/* sampling interval */
-static int	ParWidth_1;	/* number of threads - 1 */ 
-
-static int*	SumProc;		/* partial sums */
-static int	SumShared;		/* final result */
-
-// public
-
-void
-winnow_mpi(mpi::communicator world,
-  int2D*		matrix,			/* point values */
-  bool2D*	mask,			/* suitable points */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  pt1D*		pts,			/* points to create */
-  int		npt			/* number of points */
-){
-  int		i, j;			/* loop indices */
-  int		sum, tmp;		/* for scanning */
-  int rank, size;
-
-  rank = world.rank ();
-  size = world.size ();
-  if (size > 1) {
-    ParWidth_1 = size - 1;
-
-    Totals = new int1D[MAXEXT];
-    TmpPt = new pt[MAXEXT * MAXEXT];
-    TmpPt2 = new pt[MAXEXT * MAXEXT];
-
-    Pivots = new int[size * size];
-    Counts = new int[size * size];
-    Starts = new int[size * size];
-
-    SumProc = new int[size];
-
-    /* pack points into temporary storage */
-
-    winnow_count(world, mask, nr, nc);
-    i = scanIntSum(world, Totals, nr);
-
-    Len = i;
-
-    /* set slice sizes */
-    NumSamples   = size * (ParWidth_1);
-    SectionSize  = INT_CEIL(Len, size);
[…]

### 2009-08-02 cowichan_mpi/winnow/parallel.h

@@ -1,92 +1,0 @@
-/**
- * Parallel implementation of weighted point selection
- *
- * \file parallel.h
- * \author Andrew Borzenko
- * \date 02-23-09
- */
-
-#pragma once
-#ifndef WINNOW_PARALLEL_H
-#define WINNOW_PARALLEL_H
-
-/*--------------------------------------------------------------*/
-/* private function prototypes					*/
-/*--------------------------------------------------------------*/
-
-void
-winnow_copy(mpi::communicator world,
-  int2D*		matrix,			/* matrix of values */
-  bool2D*	mask,			/* mask on values */
-  int		nr,			/* row size */
-  int		nc			/* column size */
-);
-void
-winnow_count(mpi::communicator world,
-  bool2D*	mask,			/* mask on points */
-  int		nr,			/* row size */
-  int		nc			/* column size */
-);
-void
-winnow_pack(
-  pt1D*		ptDst,			/* to pack into */
-  int		nDst,			/* number of points */
-  pt1D*		ptSrc,			/* to pull from */
-  int		nSrc,			/* number of tmps */
-  int		nt,			/* number of threads */
-  int		rank
-);
-void
-winnow_psrs_1(
-  mpi::communicator world
-);
-void
-winnow_psrs_2(
-  mpi::communicator world
-);
-void
-winnow_psrs_3(
-  mpi::communicator world
-);
-void
-winnow_psrs_4(
-  mpi::communicator world
-);
-void
-winnow_sched(
-  int		rank,			/* caller thread ID */
-  int	      * start,			/* start of own interval */
-  int	      * end			/* end of own interval */
-); 
-
-int
-scanIntSum(mpi::communicator world,
-  int	      * vec,			/* to sum */
-  int		len			/* vector length */
-);
-
-int winnow_redBool2DCount_mpi(bool2D* mask, // to reduce
-                              int nr,       // row size
-                              int nc);      // column size
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-/*
- * @ winnow : do point winnowing
- * > none
- * + create vector of points
- */
-
-void
-winnow_mpi(mpi::communicator world,
-  int2D*		matrix,			/* point values */
-  bool2D*	mask,			/* suitable points */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  pt1D*		pt,			/* points to create */
-  int		npt			/* number of points */
-);
-
-#endif /* WINNOW_PARALLEL_H */

### 2009-08-02 cowichan_mpi/winnow/serial.cpp

@@ -1,75 +1,0 @@
-/**
- * Serial implementation of weighted point selection
- *
- * \file serial.cpp
- * \author Andrew Borzenko
- * \date 02-23-09
- */
-
-#include "../include/main.h"
-#include "serial.h"
-
-// public
-
-void
-winnow(
-  int2D*		matrix,			/* point values */
-  bool2D*	mask,			/* suitable points */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  pt1D*		pt,			/* points to create */
-  int		npt			/* number of points */
-){
-  pt1DX*		ptVec;			/* temp point vector */
-  int		r, c, i, j, len;	/* indices and number of points */
-  int		stride;			/* selection stride */
-
-  ptVec = new pt1DX[MAXEXT * MAXEXT];
-
-  /* fill temporary vector */
-  len = winnow_redBool2DCount(mask, nr, nc);
-  ASSERT(len >= npt);
-  i = 0;
-  for (r=0; r<nr; r++){
-    for (c=0; c<nc; c++){
-      if (mask[r][c]){
-        ptVec[i].x = r;
-        ptVec[i].y = c;
-        ptVec[i].w = matrix[r][c];
-        i++;
-      }
-    }
-  }
-
-  /* sort */
-  ptSort(ptVec, len);
-
-  /* copy over points */
-  stride = len / npt;
-  for (i=npt-1, j=len-1; i>=0; i--, j-=stride){
-    pt[i] = ptVec[j];
-  }
-
-  delete [] ptVec;
-
-  /* return */
-}
-
-// private
-
-int
-winnow_redBool2DCount(
-  bool2D*	mask,			/* to reduce */
-  int		nr,			/* row size */
-  int		nc			/* column size */
-){
-  int		r, c, sum = 0;		/* indices and result */
-
-  for (r=0; r<nr; r++){
-    for (c=0; c<nc; c++){
-      if (mask[r][c]) sum++;
-    }
-  }
-
-  return sum;
-}

### 2009-08-02 cowichan_mpi/winnow/serial.h

@@ -1,44 +1,0 @@
-/**
- * Serial implementation of weighted point selection
- *
- * \file serial.h
- * \author Andrew Borzenko
- * \date 02-23-09
- */
-
-#pragma once
-#ifndef WINNOW_SERIAL_H
-#define WINNOW_SERIAL_H
-
-/*--------------------------------------------------------------*/
-/* private function prototypes					*/
-/*--------------------------------------------------------------*/
-
-int
-winnow_redBool2DCount(
-  bool2D*	mask,			/* to reduce */
-  int		nr,			/* row size */
-  int		nc			/* column size */
-);
-
-/*--------------------------------------------------------------*/
-/* public functions						*/
-/*--------------------------------------------------------------*/
-
-/*
- * @ winnow : do point winnowing
- * > none
- * + create vector of points
- */
-
-void
-winnow(
-  int2D*		matrix,			/* point values */
-  bool2D*	mask,			/* suitable points */
-  int		nr,			/* row size */
-  int		nc,			/* column size */
-  pt1D*		pt,			/* points to create */
-  int		npt			/* number of points */
-);
-
-#endif /* WINNOW_SERIAL_H */

### 2009-08-02 cowichan_mpi/winnow/winnow.cpp

@@ -1,97 +1,0 @@
-/**
- * Weighted point selection
- *
- * \file winnow.cpp
- * \author Andrew Borzenko
- * \date 02-23-09
- */
-
-#include "../include/main.h"
-#ifdef IS_PARALLEL
-  #include "parallel.h"
-#else
-  #include "serial.h"
-#endif
-
-int main(int argc, char* argv[])
-{
-#ifdef IS_PARALLEL
-  mpi::environment env(argc, argv);
-  mpi::communicator world;
-
-#ifdef TEST_OUTPUT
-  printf ("I am process %d\n", world.rank ());
-#endif
-#endif
-
-  int2D* matrix;
-  bool2D* mask;
-  pt1D* pt;
-  int    nr;    /* row size */
-  int    nc;    /* column size */
-  int limit;
-  int npt;
-  int i,j;
-
-  srand (222);
-
-  nr = MAXEXT;
-  nc = MAXEXT;
-  limit = 10;
-  npt = MAXEXT;
-  
-  matrix = new int2D[MAXEXT];
-  for (i = 0; i < nr; i++)
-  {
-    for (j = 0; j < nc; j++)
-    {
-      matrix[i][j] = rand () % limit;
-    }
-  }
-
-  mask = new bool2D[MAXEXT];
-  for (i = 0; i < nr; i++)
-  {
-    for (j = 0; j < nc; j++)
-    {
-      mask[i][j] = rand () % 2;
-    }
-  }
-
-  pt = new pt1D[npt];
-
-#ifdef TEST_OUTPUT
-  printf ("Matrix:\n");
-  print_matrix (matrix, nr, nc);
-
-  printf ("Mask:\n");
-  print_matrix (mask, nr, nc);
-#endif
-
-#ifdef TEST_TIME
-  INT64 start, end;
-  start = get_ticks ();
-#endif
-
-#ifdef IS_PARALLEL
-  winnow_mpi (world, matrix, mask, nr, nc, pt, npt); 
-#else
-  winnow (matrix, mask, nr, nc, pt, npt); 
-#endif
-
-#ifdef TEST_TIME
-  end = get_ticks ();
-  print_elapsed_time (start, end);
-#endif
-
-#ifdef TEST_OUTPUT
-  printf ("Points:\n");
-  print_vector (pt, nr);
-#endif
-
-  delete [] matrix;
-  delete [] mask;
-  delete [] pt;
-
-  return 0;
-}

### 2009-08-05 cowichan/cowichan.hpp
added a comment about randmat

@@ -23,7 +23,8 @@
  * \subsection randmat_sec 2. Random Number Generation
  * This module fills a matrix with pseudo-random integers.
  * Note that, as in all problems, the output is required to be independent of
- * the number of processors used.
+ * the number of processors used. However, the matrices generated may be
+ * different in each implementation.
  * \see Cowichan::randmat
  *
  * \subsection half_sec 3. Two-Dimensional Shuffle

### 2009-08-05 cowichan_serial/randmat.cpp

@@ -13,7 +13,7 @@ void CowichanSerial::randmat (IntMatrix matrix)
 
   for (c = 0; c < nc; c++) {
     for (r = 0; r < nr; r++) {
-      MATRIX_RECT(matrix, r, c) = v;
+      MATRIX_RECT(matrix, c, r) = v;
       v = (RANDMAT_A * v + RANDMAT_C) % RAND_M;
     }
   }

### 2009-08-05 cowichan/cowichan_defaults.hpp
changed life iterations default

@@ -160,7 +160,7 @@
 /**
  * Default number of life iterations.
  */
-#define LIFE_ITERATIONS 10
+#define LIFE_ITERATIONS 3
 
 // winnow
 /**

### 2009-08-05 doc/latex/class_cowichan.tex
doxygen documentation as of now

@@
 
-Implemented in \hyperlink{class_cowichan_serial_b196d2384c2fe8b185e6269a050399eb}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_363bcd6f0c7b4d0fced94fe4cd59a267}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{CowichanTBB}.\hypertarget{class_cowichan_308603053675bccbe631f04af921f57c}{
+Implemented in \hyperlink{class_cowichan_serial_b196d2384c2fe8b185e6269a050399eb}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_363bcd6f0c7b4d0fced94fe4cd59a267}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_975f8da6c166fe1db3cf9341eeaab000}{CowichanMPI}.\hypertarget{class_cowichan_308603053675bccbe631f04af921f57c}{
 \index{Cowichan@{Cowichan}!half@{half}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_c08a6b3e23ce26959bac12af077f924f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_70989ffe182aebf590e39c56e146b0fb}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{CowichanTBB}.\hypertarget{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{
+Implemented in \hyperlink{class_cowichan_serial_c08a6b3e23ce26959bac12af077f924f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_70989ffe182aebf590e39c56e146b0fb}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_154bbd1a400ee4571f761bff1cdf67cd}{CowichanMPI}.\hypertarget{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{
 \index{Cowichan@{Cowichan}!hull@{hull}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{CowichanTBB}.\hypertarget{class_cowichan_ea126792a31e54a8722663b7ea768955}{
+Implemented in \hyperlink{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_22b3dce35fd93635bd4c1596e7fb839c}{CowichanMPI}.\hypertarget{class_cowichan_ea126792a31e54a8722663b7ea768955}{
 \index{Cowichan@{Cowichan}!invperc@{invperc}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4824c6b8509b5da835fbc5f64eb3e063}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{CowichanTBB}.\hypertarget{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{
+Implemented in \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4824c6b8509b5da835fbc5f64eb3e063}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_bcd0f18fcccc8973ddaacf4dda190b53}{CowichanMPI}.\hypertarget{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{
 \index{Cowichan@{Cowichan}!life@{life}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{CowichanTBB}.\hypertarget{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{
+Implemented in \hyperlink{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_9c739951d036ef7905bda7cb896e0edf}{CowichanMPI}.\hypertarget{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{
 \index{Cowichan@{Cowichan}!main@{main}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{CowichanTBB}.\hypertarget{class_cowichan_3df21e3c627958114e045c3559a29f30}{
+Implemented in \hyperlink{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_58274fc55ed8a352a8924b130f3b5d2e}{CowichanMPI}.\hypertarget{class_cowichan_3df21e3c627958114e045c3559a29f30}{
 \index{Cowichan@{Cowichan}!norm@{norm}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{CowichanTBB}.\hypertarget{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{
+Implemented in \hyperlink{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_11df49e804427d4a0d085f4cff9302d0}{CowichanMPI}.\hypertarget{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{
 \index{Cowichan@{Cowichan}!outer@{outer}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_05f6899081a457d58978e4f6bda2db6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_99a0570f754ac82968d73093476de533}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{CowichanTBB}.\hypertarget{class_cowichan_14e706ef5a5f7661a623c595d7ed76f1}{
+Implemented in \hyperlink{class_cowichan_serial_05f6899081a457d58978e4f6bda2db6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_99a0570f754ac82968d73093476de533}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_39fcc8a331b0e26438fbf43fa01cf7c1}{CowichanMPI}.\hypertarget{class_cowichan_14e706ef5a5f7661a623c595d7ed76f1}{
 \index{Cowichan@{Cowichan}!print\_\-rect\_\-matrix@{print\_\-rect\_\-matrix}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{CowichanTBB}.\hypertarget{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{
+Implemented in \hyperlink{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_1b7dccf774caccb26839cd29fe0a5cb0}{CowichanMPI}.\hypertarget{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{
 \index{Cowichan@{Cowichan}!randmat@{randmat}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{CowichanTBB}.\hypertarget{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{
+Implemented in \hyperlink{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_6805f21144aeaeddc4549e1b2b42bca8}{CowichanMPI}.\hypertarget{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{
 \index{Cowichan@{Cowichan}!sor@{sor}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_6e8b06711d976de1adc1e4dc81e560e5}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d6482d0369a26a51ef0e37ab238fc664}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{CowichanTBB}.\hypertarget{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{
+Implemented in \hyperlink{class_cowichan_serial_6e8b06711d976de1adc1e4dc81e560e5}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d6482d0369a26a51ef0e37ab238fc664}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_7388c844e8aa73ab0923443a3a7ef069}{CowichanMPI}.\hypertarget{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{
 \index{Cowichan@{Cowichan}!thresh@{thresh}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{CowichanTBB}.\hypertarget{class_cowichan_775d72b5e7d122f9f32555352278250e}{
+Implemented in \hyperlink{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_49ff96b091a61e9cfd9aad8824e7fbbd}{CowichanMPI}.\hypertarget{class_cowichan_775d72b5e7d122f9f32555352278250e}{
 \index{Cowichan@{Cowichan}!vecdiff@{vecdiff}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{CowichanTBB}.\hypertarget{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{
+Implemented in \hyperlink{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_c5470a2876efecf843b19c37c21ecf19}{CowichanMPI}.\hypertarget{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{
 \index{Cowichan@{Cowichan}!winnow@{winnow}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{CowichanTBB}.
+Implemented in \hyperlink{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{CowichanTBB}, and \hyperlink{class_cowichan_m_p_i_9be48d86fc51ce13ef79d5fe2c8a16e0}{CowichanMPI}.
 

### 2009-08-05 doc/latex/class_cowichan_m_p_i.tex
doxygen documentation as of now

\hypertarget{class_cowichan_m_p_i}{
\section{CowichanMPI Class Reference}
\label{class_cowichan_m_p_i}\index{CowichanMPI@{CowichanMPI}}
}
Boost Message Passing Interface (MPI) implementation.

{\tt \#include $<$cowichan\_\-mpi.hpp$>$}

Inheritance diagram for CowichanMPI::\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[height=2cm]{class_cowichan_m_p_i}
\end{center}
\end{figure}
\subsection*{Public Member Functions}
\begin{CompactItemize}
\item 
\hyperlink{class_cowichan_m_p_i_ba4175b1aadbab75285b47819477dedf}{CowichanMPI} (const mpi::communicator \&\hyperlink{class_cowichan_m_p_i_cf8ae701dfbb7c17dcec165800c16969}{world})
\end{CompactItemize}
\subsection*{Public Attributes}
\begin{CompactItemize}
\item 
const mpi::communicator \& \hyperlink{class_cowichan_m_p_i_cf8ae701dfbb7c17dcec165800c16969}{world}
\end{CompactItemize}
\subsection*{Protected Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{class_cowichan_m_p_i_58274fc55ed8a352a8924b130f3b5d2e}{mandel} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_m_p_i_6805f21144aeaeddc4549e1b2b42bca8}{randmat} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_m_p_i_154bbd1a400ee4571f761bff1cdf67cd}{half} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixIn, \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_m_p_i_bcd0f18fcccc8973ddaacf4dda190b53}{invperc} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_m_p_i_49ff96b091a61e9cfd9aad8824e7fbbd}{thresh} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_m_p_i_9c739951d036ef7905bda7cb896e0edf}{life} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixIn, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_m_p_i_9be48d86fc51ce13ef79d5fe2c8a16e0}{winnow} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{class_point}{PointVector} points)
\item 
void \hyperlink{class_cowichan_m_p_i_11df49e804427d4a0d085f4cff9302d0}{norm} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_m_p_i_22b3dce35fd93635bd4c1596e7fb839c}{hull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_m_p_i_39fcc8a331b0e26438fbf43fa01cf7c1}{outer} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector)
\item 
void \hyperlink{class_cowichan_m_p_i_975f8da6c166fe1db3cf9341eeaab000}{gauss} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_m_p_i_7388c844e8aa73ab0923443a3a7ef069}{sor} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_m_p_i_1b7dccf774caccb26839cd29fe0a5cb0}{product} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} candidate, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_m_p_i_c5470a2876efecf843b19c37c21ecf19}{vecdiff} (\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} actual, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} computed)
\end{CompactItemize}

\subsection{Detailed Description}
Boost Message Passing Interface (MPI) implementation.

Boost MPI is a C++ interface that works with many MPI implementations. For example: MPICH1, MPICH2, OpenMPI.

Tags: distributed memory, message passing, mpi wrapper.

\subsection{Constructor \& Destructor Documentation}
\hypertarget{class_cowichan_m_p_i_ba4175b1aadbab75285b47819477dedf}{
\index{CowichanMPI@{CowichanMPI}!CowichanMPI@{CowichanMPI}}
\index{CowichanMPI@{CowichanMPI}!CowichanMPI@{CowichanMPI}}
\subsubsection[{CowichanMPI}]{\setlength{\rightskip}{0pt plus 5cm}CowichanMPI::CowichanMPI (const mpi::communicator \& {\em world})}}
\label{class_cowichan_m_p_i_ba4175b1aadbab75285b47819477dedf}

Constructs a cowichan mpi object. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]global communicator. \end{description}
\end{Desc}

\subsection{Member Function Documentation}
\hypertarget{class_cowichan_m_p_i_975f8da6c166fe1db3cf9341eeaab000}{
\index{CowichanMPI@{CowichanMPI}!gauss@{gauss}}
\index{gauss@{gauss}!CowichanMPI@{CowichanMPI}}
\subsubsection[{gauss}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::gauss ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_975f8da6c166fe1db3cf9341eeaab000}

For description see \hyperlink{index_gauss_sec}{11. Gaussian Elimination} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{Cowichan}.\hypertarget{class_cowichan_m_p_i_154bbd1a400ee4571f761bff1cdf67cd}{
\index{CowichanMPI@{CowichanMPI}!half@{half}}
\index{half@{half}!CowichanMPI@{CowichanMPI}}
\subsubsection[{half}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::half ({\bf IntMatrix} {\em matrixIn}, \/  {\bf IntMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_154bbd1a400ee4571f761bff1cdf67cd}

For description see \hyperlink{index_half_sec}{3. Two-Dimensional Shuffle} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]matrix to shuffle. \item[{\em matrixOut}]shuffled matrix. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{Cowichan}.\hypertarget{class_cowichan_m_p_i_22b3dce35fd93635bd4c1596e7fb839c}{
\index{CowichanMPI@{CowichanMPI}!hull@{hull}}
\index{hull@{hull}!CowichanMPI@{CowichanMPI}}
\subsubsection[{hull}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::hull ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_22b3dce35fd93635bd4c1596e7fb839c}

For description see \hyperlink{index_hull_sec}{9. Convex Hull} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]points to run convex hull on. \item[{\em pointsOut}]reordered points from pointsIn. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{Cowichan}.\hypertarget{class_cowichan_m_p_i_bcd0f18fcccc8973ddaacf4dda190b53}{
\index{CowichanMPI@{CowichanMPI}!invperc@{invperc}}
\index{invperc@{invperc}!CowichanMPI@{CowichanMPI}}
\subsubsection[{invperc}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::invperc ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_bcd0f18fcccc8973ddaacf4dda190b53}

For description see \hyperlink{index_invperc_sec}{4. Invasion Percolation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]filling resistance matrix. \item[{\em mask}]filled cells. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{Cowichan}.\hypertarget{class_cowichan_m_p_i_9c739951d036ef7905bda7cb896e0edf}{
\index{CowichanMPI@{CowichanMPI}!life@{life}}
\index{life@{life}!CowichanMPI@{CowichanMPI}}
\subsubsection[{life}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::life ({\bf BoolMatrix} {\em matrixIn}, \/  {\bf BoolMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_9c739951d036ef7905bda7cb896e0edf}

For description see \hyperlink{index_life_sec}{6. Game of Life} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]initial world. \item[{\em matrixOut}]final world. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{Cowichan}.\hypertarget{class_cowichan_m_p_i_58274fc55ed8a352a8924b130f3b5d2e}{
\index{CowichanMPI@{CowichanMPI}!mandel@{mandel}}
\index{mandel@{mandel}!CowichanMPI@{CowichanMPI}}
\subsubsection[{mandel}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::mandel ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_58274fc55ed8a352a8924b130f3b5d2e}

For description see \hyperlink{index_mandel_sec}{1. Mandelbrot Set Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{Cowichan}.\hypertarget{class_cowichan_m_p_i_11df49e804427d4a0d085f4cff9302d0}{
\index{CowichanMPI@{CowichanMPI}!norm@{norm}}
\index{norm@{norm}!CowichanMPI@{CowichanMPI}}
\subsubsection[{norm}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::norm ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_11df49e804427d4a0d085f4cff9302d0}

For description see \hyperlink{index_norm_sec}{8. Point Location Normalization} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]points to normalize. \item[{\em pointsOut}]normalized points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{Cowichan}.\hypertarget{class_cowichan_m_p_i_39fcc8a331b0e26438fbf43fa01cf7c1}{
\index{CowichanMPI@{CowichanMPI}!outer@{outer}}
\index{outer@{outer}!CowichanMPI@{CowichanMPI}}
\subsubsection[{outer}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::outer ({\bf PointVector} {\em points}, \/  {\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em vector})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_39fcc8a331b0e26438fbf43fa01cf7c1}

For description see \hyperlink{index_outer_sec}{10. Outer Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]vector of points. \item[{\em matrix}]resulting distance matrix. \item[{\em vector}]resulting distance vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{Cowichan}.\hypertarget{class_cowichan_m_p_i_1b7dccf774caccb26839cd29fe0a5cb0}{
\index{CowichanMPI@{CowichanMPI}!product@{product}}
\index{product@{product}!CowichanMPI@{CowichanMPI}}
\subsubsection[{product}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::product ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em candidate}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_1b7dccf774caccb26839cd29fe0a5cb0}

For description see \hyperlink{index_product_sec}{13. Matrix-Vector Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em candidate}]vector X in AX = V. \item[{\em solution}]vector V in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan}.\hypertarget{class_cowichan_m_p_i_6805f21144aeaeddc4549e1b2b42bca8}{
\index{CowichanMPI@{CowichanMPI}!randmat@{randmat}}
\index{randmat@{randmat}!CowichanMPI@{CowichanMPI}}
\subsubsection[{randmat}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::randmat ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_6805f21144aeaeddc4549e1b2b42bca8}

For description see \hyperlink{index_randmat_sec}{2. Random Number Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{Cowichan}.\hypertarget{class_cowichan_m_p_i_7388c844e8aa73ab0923443a3a7ef069}{
\index{CowichanMPI@{CowichanMPI}!sor@{sor}}
\index{sor@{sor}!CowichanMPI@{CowichanMPI}}
\subsubsection[{sor}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::sor ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_7388c844e8aa73ab0923443a3a7ef069}

For description see \hyperlink{index_sor_sec}{12. Successive Over-Relaxation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{Cowichan}.\hypertarget{class_cowichan_m_p_i_49ff96b091a61e9cfd9aad8824e7fbbd}{
\index{CowichanMPI@{CowichanMPI}!thresh@{thresh}}
\index{thresh@{thresh}!CowichanMPI@{CowichanMPI}}
\subsubsection[{thresh}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::thresh ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_49ff96b091a61e9cfd9aad8824e7fbbd}

For description see \hyperlink{index_thresh_sec}{5. Histogram Thresholding} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]image. \item[{\em mask}]image after thresholding. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{Cowichan}.\hypertarget{class_cowichan_m_p_i_c5470a2876efecf843b19c37c21ecf19}{
\index{CowichanMPI@{CowichanMPI}!vecdiff@{vecdiff}}
\index{vecdiff@{vecdiff}!CowichanMPI@{CowichanMPI}}
\subsubsection[{vecdiff}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} CowichanMPI::vecdiff ({\bf Vector} {\em actual}, \/  {\bf Vector} {\em computed})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_c5470a2876efecf843b19c37c21ecf19}

For description see \hyperlink{index_vecdiff_sec}{14. 1-Norm Vector Difference} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em actual}]first vector. \item[{\em computed}]second vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{Cowichan}.\hypertarget{class_cowichan_m_p_i_9be48d86fc51ce13ef79d5fe2c8a16e0}{
\index{CowichanMPI@{CowichanMPI}!winnow@{winnow}}
\index{winnow@{winnow}!CowichanMPI@{CowichanMPI}}
\subsubsection[{winnow}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanMPI::winnow ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask}, \/  {\bf PointVector} {\em points})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_m_p_i_9be48d86fc51ce13ef79d5fe2c8a16e0}

For description see \hyperlink{index_winnow_sec}{7. Weighted Point Selection} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]integer matrix. \item[{\em mask}]boolean matrix. \item[{\em points}]evenly selected points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{Cowichan}.

\subsection{Member Data Documentation}
\hypertarget{class_cowichan_m_p_i_cf8ae701dfbb7c17dcec165800c16969}{
\index{CowichanMPI@{CowichanMPI}!world@{world}}
\index{world@{world}!CowichanMPI@{CowichanMPI}}
\subsubsection[{world}]{\setlength{\rightskip}{0pt plus 5cm}const mpi::communicator\& {\bf CowichanMPI::world}}}
\label{class_cowichan_m_p_i_cf8ae701dfbb7c17dcec165800c16969}

Global communicator.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
cowichan\_\-mpi/\hyperlink{cowichan__mpi_8hpp}{cowichan\_\-mpi.hpp}\item 
cowichan\_\-mpi/\hyperlink{cowichan__mpi_8cpp}{cowichan\_\-mpi.cpp}\item 
cowichan\_\-mpi/\hyperlink{cowichan__mpi_2winnow_8cpp}{winnow.cpp}\end{CompactItemize}

### 2009-08-05 doc/latex/class_cowichan_open_m_p.tex
doxygen documentation as of now

\hypertarget{class_cowichan_open_m_p}{
\section{CowichanOpenMP Class Reference}
\label{class_cowichan_open_m_p}\index{CowichanOpenMP@{CowichanOpenMP}}
}
Open Multi-Processing (OpenMP) implementation.

{\tt \#include $<$cowichan\_\-openmp.hpp$>$}

Inheritance diagram for CowichanOpenMP::\begin{figure}[H]
\begin{center}
\leavevmode
\includegraphics[height=2cm]{class_cowichan_open_m_p}
\end{center}
\end{figure}
\subsection*{Static Public Attributes}
\begin{CompactItemize}
\item 
static const \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{class_cowichan_open_m_p_52ddfbbead2e1c115806cd07f5630bf5}{HULL\_\-CUTOFF} = 20000
\end{CompactItemize}
\subsection*{Protected Member Functions}
\begin{CompactItemize}
\item 
void \hyperlink{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{mandel} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{randmat} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix)
\item 
void \hyperlink{class_cowichan_open_m_p_70989ffe182aebf590e39c56e146b0fb}{half} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixIn, \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_open_m_p_4824c6b8509b5da835fbc5f64eb3e063}{invperc} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{thresh} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask)
\item 
void \hyperlink{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}{life} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixIn, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} matrixOut)
\item 
void \hyperlink{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{winnow} (\hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix} matrix, \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{class_point}{PointVector} points)
\item 
void \hyperlink{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{norm} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}{hull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{class_point}{PointVector} pointsOut)
\item 
void \hyperlink{class_cowichan_open_m_p_99a0570f754ac82968d73093476de533}{outer} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} vector)
\item 
void \hyperlink{class_cowichan_open_m_p_363bcd6f0c7b4d0fced94fe4cd59a267}{gauss} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_open_m_p_d6482d0369a26a51ef0e37ab238fc664}{sor} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} target, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
void \hyperlink{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{product} (\hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix} matrix, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} candidate, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} solution)
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{vecdiff} (\hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} actual, \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector} computed)
\end{CompactItemize}

\subsection{Detailed Description}
Open Multi-Processing (OpenMP) implementation.

Tags: shared memory, data parallel, shared variables, task based, compiler specific.

\subsection{Member Function Documentation}
\hypertarget{class_cowichan_open_m_p_363bcd6f0c7b4d0fced94fe4cd59a267}{
\index{CowichanOpenMP@{CowichanOpenMP}!gauss@{gauss}}
\index{gauss@{gauss}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{gauss}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::gauss ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_363bcd6f0c7b4d0fced94fe4cd59a267}

For description see \hyperlink{index_gauss_sec}{11. Gaussian Elimination} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{Cowichan}.\hypertarget{class_cowichan_open_m_p_70989ffe182aebf590e39c56e146b0fb}{
\index{CowichanOpenMP@{CowichanOpenMP}!half@{half}}
\index{half@{half}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{half}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::half ({\bf IntMatrix} {\em matrixIn}, \/  {\bf IntMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_70989ffe182aebf590e39c56e146b0fb}

For description see \hyperlink{index_half_sec}{3. Two-Dimensional Shuffle} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]matrix to shuffle. \item[{\em matrixOut}]shuffled matrix. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{Cowichan}.\hypertarget{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}{
\index{CowichanOpenMP@{CowichanOpenMP}!hull@{hull}}
\index{hull@{hull}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{hull}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::hull ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}

For description see \hyperlink{index_hull_sec}{9. Convex Hull}

Runs quickhull algorithm until all points have been used up from the original vector. At each step the hull points are marked as used and a new convex hull is computed on the rest of points. The points that have been used up are in the range (n - hn, n), i.e. at the end of pointsIn vector. NOTE: pointsIn vector gets modified by the algorithm.

Implements \hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{Cowichan}.\hypertarget{class_cowichan_open_m_p_4824c6b8509b5da835fbc5f64eb3e063}{
\index{CowichanOpenMP@{CowichanOpenMP}!invperc@{invperc}}
\index{invperc@{invperc}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{invperc}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::invperc ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_4824c6b8509b5da835fbc5f64eb3e063}

For description see \hyperlink{index_invperc_sec}{4. Invasion Percolation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]filling resistance matrix. \item[{\em mask}]filled cells. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{Cowichan}.\hypertarget{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}{
\index{CowichanOpenMP@{CowichanOpenMP}!life@{life}}
\index{life@{life}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{life}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::life ({\bf BoolMatrix} {\em matrixIn}, \/  {\bf BoolMatrix} {\em matrixOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}

For description see \hyperlink{index_life_sec}{6. Game of Life} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrixIn}]initial world. \item[{\em matrixOut}]final world. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{Cowichan}.\hypertarget{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{
\index{CowichanOpenMP@{CowichanOpenMP}!mandel@{mandel}}
\index{mandel@{mandel}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{mandel}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::mandel ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}

For description see \hyperlink{index_mandel_sec}{1. Mandelbrot Set Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{Cowichan}.\hypertarget{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{
\index{CowichanOpenMP@{CowichanOpenMP}!norm@{norm}}
\index{norm@{norm}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{norm}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::norm ({\bf PointVector} {\em pointsIn}, \/  {\bf PointVector} {\em pointsOut})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}

For description see \hyperlink{index_norm_sec}{8. Point Location Normalization} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em pointsIn}]points to normalize. \item[{\em pointsOut}]normalized points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{Cowichan}.\hypertarget{class_cowichan_open_m_p_99a0570f754ac82968d73093476de533}{
\index{CowichanOpenMP@{CowichanOpenMP}!outer@{outer}}
\index{outer@{outer}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{outer}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::outer ({\bf PointVector} {\em points}, \/  {\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em vector})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_99a0570f754ac82968d73093476de533}

For description see \hyperlink{index_outer_sec}{10. Outer Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em points}]vector of points. \item[{\em matrix}]resulting distance matrix. \item[{\em vector}]resulting distance vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{Cowichan}.\hypertarget{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{
\index{CowichanOpenMP@{CowichanOpenMP}!product@{product}}
\index{product@{product}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{product}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::product ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em candidate}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}

For description see \hyperlink{index_product_sec}{13. Matrix-Vector Product} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em candidate}]vector X in AX = V. \item[{\em solution}]vector V in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan}.\hypertarget{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{
\index{CowichanOpenMP@{CowichanOpenMP}!randmat@{randmat}}
\index{randmat@{randmat}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{randmat}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::randmat ({\bf IntMatrix} {\em matrix})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}

For description see \hyperlink{index_randmat_sec}{2. Random Number Generation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix to fill. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{Cowichan}.\hypertarget{class_cowichan_open_m_p_d6482d0369a26a51ef0e37ab238fc664}{
\index{CowichanOpenMP@{CowichanOpenMP}!sor@{sor}}
\index{sor@{sor}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{sor}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::sor ({\bf Matrix} {\em matrix}, \/  {\bf Vector} {\em target}, \/  {\bf Vector} {\em solution})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_d6482d0369a26a51ef0e37ab238fc664}

For description see \hyperlink{index_sor_sec}{12. Successive Over-Relaxation} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]matrix A in AX = V. \item[{\em target}]vector V in AX = V. \item[{\em solution}]vector X in AX = V. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{Cowichan}.\hypertarget{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{
\index{CowichanOpenMP@{CowichanOpenMP}!thresh@{thresh}}
\index{thresh@{thresh}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{thresh}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::thresh ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}

For description see \hyperlink{index_thresh_sec}{5. Histogram Thresholding} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]image. \item[{\em mask}]image after thresholding. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{Cowichan}.\hypertarget{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{
\index{CowichanOpenMP@{CowichanOpenMP}!vecdiff@{vecdiff}}
\index{vecdiff@{vecdiff}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{vecdiff}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} CowichanOpenMP::vecdiff ({\bf Vector} {\em actual}, \/  {\bf Vector} {\em computed})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}

For description see \hyperlink{index_vecdiff_sec}{14. 1-Norm Vector Difference} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em actual}]first vector. \item[{\em computed}]second vector. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{Cowichan}.\hypertarget{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{
\index{CowichanOpenMP@{CowichanOpenMP}!winnow@{winnow}}
\index{winnow@{winnow}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{winnow}]{\setlength{\rightskip}{0pt plus 5cm}void CowichanOpenMP::winnow ({\bf IntMatrix} {\em matrix}, \/  {\bf BoolMatrix} {\em mask}, \/  {\bf PointVector} {\em points})\hspace{0.3cm}{\tt  \mbox{[}protected, virtual\mbox{]}}}}
\label{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}

For description see \hyperlink{index_winnow_sec}{7. Weighted Point Selection} \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em matrix}]integer matrix. \item[{\em mask}]boolean matrix. \item[{\em points}]evenly selected points. \end{description}
\end{Desc}

Implements \hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{Cowichan}.

\subsection{Member Data Documentation}
\hypertarget{class_cowichan_open_m_p_52ddfbbead2e1c115806cd07f5630bf5}{
\index{CowichanOpenMP@{CowichanOpenMP}!HULL\_\-CUTOFF@{HULL\_\-CUTOFF}}
\index{HULL\_\-CUTOFF@{HULL\_\-CUTOFF}!CowichanOpenMP@{CowichanOpenMP}}
\subsubsection[{HULL\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}const {\bf index\_\-t} {\bf CowichanOpenMP::HULL\_\-CUTOFF} = 20000\hspace{0.3cm}{\tt  \mbox{[}static\mbox{]}}}}
\label{class_cowichan_open_m_p_52ddfbbead2e1c115806cd07f5630bf5}

Cutoff value n for hull.

The documentation for this class was generated from the following files:\begin{CompactItemize}
\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_8hpp}{cowichan\_\-openmp.hpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2gauss_8cpp}{gauss.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2half_8cpp}{half.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2hull_8cpp}{hull.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2invperc_8cpp}{invperc.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2life_8cpp}{life.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2mandel_8cpp}{mandel.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2norm_8cpp}{norm.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2outer_8cpp}{outer.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2product_8cpp}{product.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2randmat_8cpp}{randmat.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sor_8cpp}{sor.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2thresh_8cpp}{thresh.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2vecdiff_8cpp}{vecdiff.cpp}\item 
cowichan\_\-openmp/\hyperlink{cowichan__openmp_2winnow_8cpp}{winnow.cpp}\end{CompactItemize}

### 2009-08-05 doc/latex/files.tex
doxygen documentation as of now

@@
 \item\contentsline{section}{cowichan/\hyperlink{cowichan__defaults_8hpp}{cowichan\_\-defaults.hpp} (Defaults for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan__defaults_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_8cpp}{cowichan\_\-mpi.cpp} (This file contains main method that drives the MPI implementation )}{\pageref{cowichan__mpi_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_8hpp}{cowichan\_\-mpi.hpp} (Boost Message Passing Interface (MPI) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__mpi_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2sort_8cpp}{sort.cpp} (Implementation of MPI sorting algorithms )}{\pageref{cowichan__mpi_2sort_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2sort_8hpp}{sort.hpp} (Sorting algorithms that use MPI )}{\pageref{cowichan__mpi_2sort_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2winnow_8cpp}{winnow.cpp} (MPI winnow implementation )}{\pageref{cowichan__mpi_2winnow_8cpp}}{}
 \item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_8cpp}{cowichan\_\-openmp.cpp} (This file contains main method that drives the OpenMP implementation )}{\pageref{cowichan__openmp_8cpp}}{}
@@
 \item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sor_8cpp}{sor.cpp} (OpenMP sor implementation )}{\pageref{cowichan__openmp_2sor_8cpp}}{}
-\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{sort_8cpp}{sort.cpp} (Implementation of OpenMP sorting algorithms )}{\pageref{sort_8cpp}}{}
-\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{sort_8hpp}{sort.hpp} (Sorting algorithms that use OpenMP )}{\pageref{sort_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sort_8cpp}{sort.cpp} (Implementation of OpenMP sorting algorithms )}{\pageref{cowichan__openmp_2sort_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2sort_8hpp}{sort.hpp} (Sorting algorithms that use OpenMP )}{\pageref{cowichan__openmp_2sort_8hpp}}{}
 \item\contentsline{section}{cowichan\_\-openmp/\hyperlink{cowichan__openmp_2thresh_8cpp}{thresh.cpp} (OpenMP thresh implementation )}{\pageref{cowichan__openmp_2thresh_8cpp}}{}

### 2009-08-05 doc/latex/namespacecowichan__mpi.tex
doxygen documentation as of now

\hypertarget{namespacecowichan__mpi}{
\section{cowichan\_\-mpi Namespace Reference}
\label{namespacecowichan__mpi}\index{cowichan\_\-mpi@{cowichan\_\-mpi}}
}
Additional classes and functions specific to mpi implementation.

\subsection*{Functions}
\begin{CompactItemize}
\item 
bool \hyperlink{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}{get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end)
\item 
bool \hyperlink{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}{get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} rank)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}{getNumRemainingNodes} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}{power2} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)
\item 
void \hyperlink{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}{quick\_\-sort} (const mpi::communicator \&world, \hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} owner)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}{quick\_\-sort\_\-partition} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} pivotIndex)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_a54004a105356573d4bac92e1b234aa2}{mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\end{CompactItemize}

\subsection{Detailed Description}
Additional classes and functions specific to mpi implementation.

\subsection{Function Documentation}
\hypertarget{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!get\_\-block@{get\_\-block}}
\index{get\_\-block@{get\_\-block}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{get\_\-block}]{\setlength{\rightskip}{0pt plus 5cm}bool cowichan\_\-mpi::get\_\-block (const mpi::communicator \& {\em world}, \/  {\bf index\_\-t} {\em lo}, \/  {\bf index\_\-t} {\em hi}, \/  {\bf index\_\-t} $\ast$ {\em start}, \/  {\bf index\_\-t} $\ast$ {\em end}, \/  {\bf index\_\-t} {\em rank})}}
\label{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}

Get a block (start, end) to work on in the range (lo, hi) for process rank. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]communicator. \item[{\em lo}]lower end of the range. \item[{\em hi}]higher end of the range. \item[{\em start}]first index to work on. \item[{\em end}]last index to work on. \item[{\em rank}]process to get work for. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Whether any work is available for process rank. \end{Desc}
\hypertarget{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!get\_\-block@{get\_\-block}}
\index{get\_\-block@{get\_\-block}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{get\_\-block}]{\setlength{\rightskip}{0pt plus 5cm}bool cowichan\_\-mpi::get\_\-block (const mpi::communicator \& {\em world}, \/  {\bf index\_\-t} {\em lo}, \/  {\bf index\_\-t} {\em hi}, \/  {\bf index\_\-t} $\ast$ {\em start}, \/  {\bf index\_\-t} $\ast$ {\em end})}}
\label{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}

Get a block (start, end) to work on in the range (lo, hi) for the current process. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]communicator. \item[{\em lo}]lower end of the range. \item[{\em hi}]higher end of the range. \item[{\em start}]first index to work on. \item[{\em end}]last index to work on. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Whether any work is available for the current process. \end{Desc}
\hypertarget{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!getNumRemainingNodes@{getNumRemainingNodes}}
\index{getNumRemainingNodes@{getNumRemainingNodes}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{getNumRemainingNodes}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-mpi::getNumRemainingNodes (const mpi::communicator \& {\em world}, \/  {\bf index\_\-t} {\em level})}}
\label{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}

Get number of remaining nodes, i.e. not doing useful work. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]global communicator. \item[{\em level}]current recursion depth. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Number of remaining nodes. \end{Desc}
\hypertarget{namespacecowichan__mpi_a54004a105356573d4bac92e1b234aa2}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!mask\_\-count@{mask\_\-count}}
\index{mask\_\-count@{mask\_\-count}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{mask\_\-count}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-mpi::mask\_\-count ({\bf BoolMatrix} {\em mask}, \/  {\bf index\_\-t} {\em nr}, \/  {\bf index\_\-t} {\em nc})}}
\label{namespacecowichan__mpi_a54004a105356573d4bac92e1b234aa2}

Count the number of set cells in the mask. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em mask}]boolean mask. \item[{\em nr}]number of rows. \item[{\em nc}]number of columns. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]Number of set cells in the mask. \end{Desc}
\hypertarget{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!power2@{power2}}
\index{power2@{power2}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{power2}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-mpi::power2 ({\bf index\_\-t} {\em n})}}
\label{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}

Calculates a power of 2. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em n}]power is limited by sizeof(index\_\-t) $\ast$ 8 - 2 to prevent overflow. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]2$^\wedge$n. \end{Desc}
\hypertarget{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!quick\_\-sort@{quick\_\-sort}}
\index{quick\_\-sort@{quick\_\-sort}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{quick\_\-sort}]{\setlength{\rightskip}{0pt plus 5cm}void cowichan\_\-mpi::quick\_\-sort (const mpi::communicator \& {\em world}, \/  {\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len}, \/  {\bf index\_\-t} {\em level} = {\tt 0}, \/  {\bf index\_\-t} {\em owner} = {\tt 0})}}
\label{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}

This performs parallel quick sort. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em world}]global communicator. \item[{\em vector}]vector to sort. \item[{\em len}]vector length. \item[{\em level}]depth of recursion. \item[{\em owner}]rank of the owner of this sort. \end{description}
\end{Desc}
\hypertarget{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}{
\index{cowichan\_\-mpi@{cowichan\_\-mpi}!quick\_\-sort\_\-partition@{quick\_\-sort\_\-partition}}
\index{quick\_\-sort\_\-partition@{quick\_\-sort\_\-partition}!cowichan_mpi@{cowichan\_\-mpi}}
\subsubsection[{quick\_\-sort\_\-partition}]{\setlength{\rightskip}{0pt plus 5cm}{\bf index\_\-t} cowichan\_\-mpi::quick\_\-sort\_\-partition ({\bf WeightedPointVector} {\em vector}, \/  {\bf index\_\-t} {\em len}, \/  {\bf index\_\-t} {\em pivotIndex})}}
\label{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}

Partitions vector into points less than and greater than the pivot. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em vector}]vector to partition. \item[{\em len}]vector length. \item[{\em pivotIndex}]index to partition around. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]New pivot index. \end{Desc}

### 2009-08-05 doc/latex/sort_8cpp.tex
doxygen documentation as of now

--- 2009-08-02 doc/latex/sort_8hpp.tex
@@
-\hypertarget{sort_8hpp}{
-\section{cowichan\_\-openmp/sort.hpp File Reference}
-\label{sort_8hpp}\index{cowichan\_\-openmp/sort.hpp@{cowichan\_\-openmp/sort.hpp}}
+\hypertarget{sort_8cpp}{
+\section{cowichan\_\-openmp/sort.cpp File Reference}
+\label{sort_8cpp}\index{cowichan\_\-openmp/sort.cpp@{cowichan\_\-openmp/sort.cpp}}
 }
-Sorting algorithms that use OpenMP.
+Implementation of OpenMP sorting algorithms.
 
-{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
+{\tt \#include \char`\"{}sort.hpp\char`\"{}}\par
 \subsection*{Namespaces}
@@
 
-\end{CompactItemize}
-\subsection*{Defines}
-\begin{CompactItemize}
-\item 
-\#define \hyperlink{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{QUICK\_\-SORT\_\-CUTOFF}~100
-\item 
-\#define \hyperlink{sort_8hpp_66832e055934d011da279261bcfafdd1}{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}~1000
 \end{CompactItemize}
@@
 \subsection{Detailed Description}
-Sorting algorithms that use OpenMP.
-
-Two algorithms are included: \begin{enumerate}
-\item Parallel Histogram Sort - similar to bucket sort, but does not require additional space for buckets (in-place).  \item Parallel Quick Sort - requires OpenMP 3.0 \end{enumerate}
-
-\subsection{Define Documentation}
-\hypertarget{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{
-\index{sort.hpp@{sort.hpp}!QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}}
-\index{QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}!sort.hpp@{sort.hpp}}
-\subsubsection[{QUICK\_\-SORT\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}\#define QUICK\_\-SORT\_\-CUTOFF~100}}
-\label{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}
-
-Quick sort cutoff (becomes insertion sort) - must be $>$ 1 \hypertarget{sort_8hpp_66832e055934d011da279261bcfafdd1}{
-\index{sort.hpp@{sort.hpp}!QUICK\_\-SORT\_\-TASK\_\-CUTOFF@{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}}
-\index{QUICK\_\-SORT\_\-TASK\_\-CUTOFF@{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}!sort.hpp@{sort.hpp}}
-\subsubsection[{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}\#define QUICK\_\-SORT\_\-TASK\_\-CUTOFF~1000}}
-\label{sort_8hpp_66832e055934d011da279261bcfafdd1}
-
-Parallel sort cutoff (no more tasks generated).
+Implementation of OpenMP sorting algorithms.

### 2009-08-05 cowichan_lt/src/cowichan_lt.hpp
Round of fix-ups; all checked-in code working except hull, gauss, sor; winnow and outer still to be checked in as logic is not yet complete

@@ -6,9 +6,7 @@
 #define __cowichan_lt_hpp__
 
 #include "../../cowichan/cowichan.hpp"
-
-#define MAX(a,b) ((a) > (b) ? (a) : (b))
-#define MIN(a,b) ((a) < (b) ? (a) : (b))
+#include <algorithm> // for std::min, std::max
 
 // COWICHAN DEFINITIONS =====================================================//
 // aka. "inputs" to the toys, and chaining functions.

### 2009-08-05 cowichan_lt/src/hull.cpp

@@ -17,20 +17,20 @@ const char* MAX_POINT = "hull furthest point";
 
 const char* HULL_POINT = "hull point";
 const char* MASKED_POINT = "hull masked point";
-const char* NUM_POINTS = "hull # of points in convex hull"
+const char* NUM_POINTS = "hull # of points in convex hull";
 
 const char* FLAG_OUTPUT = "hull flag output";
 const char* FINISHED_MINMAX = "hull finshed min/max";
 
 
 void CowichanLinuxTuples::hull(PointVector pointsIn, PointVector pointsOut) {
-	real order = 0;
+	int order = 0;
 
 	// while not all points are used up then run quickhull on the rest of points
 	while (order < HULL_N) {
 
 		// Run quickhull on the decided points as a tuple-space problem.
-		LTHull app;
+		LTHull app(HULL_N - order);
 		app.addInput(0, pointsIn);
 		app.addOutput(0, pointsOut + order); // offset into the array
 		app.start(SERVER, PORT, NUM_WORKERS);
@@ -42,9 +42,17 @@ void CowichanLinuxTuples::hull(PointVector pointsIn, PointVector pointsOut) {
 
 	}
 
+	// cleanup: delete all mask tuples
+	tuple* maskTemplate = make_tuple("s?", MASKED_POINT);
+	while (true) {
+		if (get_tuple_nb(masked, &ctx) == NULL) break;
+	}
+
 }
 
 //===========================================================================//
+
+LTHull::LTHull(size_t left): numLeft(left) { }
 
 void LTHull::consumeInput() {
 
@@ -55,23 +63,30 @@ void LTHull::consumeInput() {
 	put_tuple(synchLock, &ctx);
 
 	// base case
-	if (n == 1) {
+	if (numLeft == 1) {
+
+		PointVector pointsIn = (PointVector) inputs[0];
+
+		// search for the only point
+		index_t pos;
+		for (pos = 0; pos < HULL_N; ++pos) {
+			if (!isMasked(pos)) break;
+		}
 
 		// emit the only point
-		tuple* hullPoint = make_tuple("sis", HULL_POINT, (*order)++);
-		hullPoint->elements[2].data.s.len = sizeof(Point);
-		hullPoint->elements[2].data.s.ptr = &p1;
+		tuple* hullPoint = make_tuple("sii", HULL_POINT, 0);
[…]

### 2009-08-05 cowichan_lt/src/hull.hpp

@@ -4,8 +4,17 @@
 	#include "tuple_common.hpp"
 	#include "cowichan_lt.hpp"
 
+	#define NO_POINT -1
+
 	class LTHull: public TupleApplication {
+	public:
+
+		int getNumPoints();
+		LTHull(size_t numLeft);
+
 	protected:
+
+		size_t numLeft;
 
 		void consumeInput();
 		void work();
@@ -30,12 +39,14 @@
 		static const char* FLAG_OUTPUT;
 		static const char* FINISHED_MINMAX;
 
-		void computeMinMax(size_t n, Point* minPoint, Point* maxPoint);
-		void split(PointVector pointsIn, index_t n, index_t* hn, const Point& p1, const Point& p2);
+		void split(const Point& p1, const Point& p2, index_t *order);
+
+		void computeCross(Point* p1, Point* p2, Point* maxPoint, real* maxCross);
+		void computeMinMax(Point* minPoint, Point* maxPoint);
+
 		void serviceMinMaxRequest(tuple* gotten);
 		void serviceCrossRequest(tuple* gotten);
 
-		int getNumPoints();
 		bool isMasked(index_t position);
 
 	};

### 2009-08-05 cowichan_lt/src/life.cpp

@@ -12,12 +12,12 @@ void CowichanLinuxTuples::life(BoolMatrix matrixIn, BoolMatrix matrixOut) {
 /**
  * Calculate number of peers.
  */
-index_t LTLife::sumNeighbours(index_t x, index_t y) const {
+index_t LTLife::sumNeighbours(index_t y, index_t x) {
 
 	index_t peers = 0;
 
 	// grab pointers locally.
-	BoolMatrix input = (BoolMatrix) inputs[0];
+	BoolMatrix first = (BoolMatrix) inputs[0];
 
 	// calculate possible neighbour positions
 	bool l = (x > 0);
@@ -26,14 +26,14 @@ index_t LTLife::sumNeighbours(index_t x, index_t y) const {
 	bool d = (y < (LIFE_NR - 1));
 
 	// calculate no. of neighbours
-	if (l &&       MATRIX_RECT_NC(_first, y    , x - 1, LIFE_NC)) ++peers;
-	if (l && u &&  MATRIX_RECT_NC(_first, y - 1, x - 1, LIFE_NC)) ++peers;
-	if (u &&       MATRIX_RECT_NC(_first, y - 1, x    , LIFE_NC)) ++peers;
-	if (r && u &&  MATRIX_RECT_NC(_first, y - 1, x + 1, LIFE_NC)) ++peers;
-	if (r &&       MATRIX_RECT_NC(_first, y    , x + 1, LIFE_NC)) ++peers;
-	if (r && d &&  MATRIX_RECT_NC(_first, y + 1, x + 1, LIFE_NC)) ++peers;
-	if (d &&       MATRIX_RECT_NC(_first, y + 1, x    , LIFE_NC)) ++peers;
-	if (l && d &&  MATRIX_RECT_NC(_first, y + 1, x - 1, LIFE_NC)) ++peers;
+	if (l &&       MATRIX_RECT_NC(first, y    , x - 1, LIFE_NC)) ++peers;
+	if (l && u &&  MATRIX_RECT_NC(first, y - 1, x - 1, LIFE_NC)) ++peers;
+	if (u &&       MATRIX_RECT_NC(first, y - 1, x    , LIFE_NC)) ++peers;
+	if (r && u &&  MATRIX_RECT_NC(first, y - 1, x + 1, LIFE_NC)) ++peers;
+	if (r &&       MATRIX_RECT_NC(first, y    , x + 1, LIFE_NC)) ++peers;
[…]

### 2009-08-05 cowichan_lt/src/life.hpp

@@ -10,7 +10,7 @@
 		void work();
 		void produceOutput();
 
-		index_t sumNeighbours();
+		index_t sumNeighbours(index_t y, index_t x);
 	};
 
 #endif

### 2009-08-05 cowichan_lt/src/norm.cpp

@@ -45,10 +45,10 @@ void LTBounds::consumeInput() {
 
 	// split points, based on a cluster size of the square-root of the
 	// number of points given.
-	size_t skip = (size_t) sqrt((real) NORM_N);
-	for (size_t pos = 0; pos < NORM_N; pos += skip) {
+	index_t skip = (size_t) sqrt((real) NORM_N);
+	for (index_t pos = 0; pos < NORM_N; pos += skip) {
 		send->elements[1].data.i = pos;
-		send->elements[2].data.i = min(pos + skip, NORM_N);
+		send->elements[2].data.i = std::min(pos + skip, NORM_N);
 		put_tuple(send, &ctx);
 	}
 
@@ -75,10 +75,10 @@ void LTBounds::work() {
 		Point maxPoint = input[start];
 		for (size_t pos = start + 1; pos < stop; ++pos) {
 			Point &cur = input[pos];
-			minPoint.x = min(minPoint.x, cur.x);
-			minPoint.y = min(minPoint.y, cur.y);
-			maxPoint.x = max(maxPoint.x, cur.x);
-			maxPoint.y = max(maxPoint.y, cur.y);
+			minPoint.x = std::min(minPoint.x, cur.x);
+			minPoint.y = std::min(minPoint.y, cur.y);
+			maxPoint.x = std::max(maxPoint.x, cur.x);
+			maxPoint.y = std::max(maxPoint.y, cur.y);
 		}
 
 		// purge local memory of the tuple we received
@@ -92,9 +92,9 @@ void LTBounds::work() {
 			tuple *tmpMin = make_tuple("s?", MIN_POINT);
 			tuple *tupleMin = get_nb_tuple(tmpMin, &ctx);
 			if (tupleMin != NULL) {
-				Point* worldMin = (Point*) tupleMin->elements[1].data.s;
-				minPoint.x = min(minPoint.x, worldMin->x);
-				minPoint.y = min(minPoint.y, worldMin->y);
+				Point* worldMin = (Point*) tupleMin->elements[1].data.s.ptr;
+				minPoint.x = std::min(minPoint.x, worldMin->x);
+				minPoint.y = std::min(minPoint.y, worldMin->y);
 				destroy_tuple(tupleMin);
 			}
 			tmpMin->elements[1].data.s.len = sizeof(Point);
@@ -106,9 +106,9 @@ void LTBounds::work() {
 			tuple *tmpMax = make_tuple("s?", MAX_POINT);
[…]

### 2009-08-05 cowichan_lt/src/product.cpp

@@ -16,7 +16,7 @@ void LTProduct::consumeInput() {
 	tuple *send = make_tuple("si", "product request");
 
 	// send off a request for each grid row.
-	for (size_t y = 0; y < PRODUCT_NR; ++y) {
+	for (size_t y = 0; y < PRODUCT_N; ++y) {
 		send->elements[1].data.i = y;
 		put_tuple(send, &ctx);
 	}
@@ -48,8 +48,8 @@ void LTProduct::work() {
 
 		// perform and store the actual computation for this row.
 		real result = 0.0;
-		for (int c = 0; c < PRODUCT_NC; ++c) {
-			result += MATRIX_RECT_NC(matrix, r, c, PRODUCT_NC) * VECTOR(candidate, c);
+		for (int c = 0; c < PRODUCT_N; ++c) {
+			result += MATRIX_SQUARE_N(matrix, r, c, PRODUCT_N) * VECTOR(candidate, c);
 		}
 		send->elements[2].data.d = result;
 	
@@ -75,7 +75,7 @@ void LTProduct::produceOutput() {
 
 	// grab all of the computations from the workers,
 	// in an unspecified order.
-	int computations = PRODUCT_NR;
+	int computations = PRODUCT_N;
 	while (computations > 0) {
 
 		// get the tuple and copy it into the matrix.

### 2009-08-05 cowichan_lt/src/sor.cpp

@@ -3,6 +3,12 @@
 #include <cmath>
 #include "sor.hpp"
 
+const char* LTSor::SYNCH_LOCK = "sor synch lock";
+const char* LTSor::ROWS_DONE = "sor rows reporting";
+const char* LTSor::SOLUTION_VECTOR = "sor solution row";
+const char* LTSor::SOLUTION_SUM = "sor inner sum";
+const char* LTSor::SOR_FLAG = "sor input consumed";
+
 void CowichanLinuxTuples::sor(Matrix matrix, Vector target, Vector solution) {
 
 	// calculate the 2D bounds of the point cloud
@@ -10,7 +16,7 @@ void CowichanLinuxTuples::sor(Matrix matrix, Vector target, Vector solution) {
 	app.addInput(0, matrix);
 	app.addInput(0, target);
 	app.addOutput(0, solution);
-	program.start(SERVER, PORT, NUM_WORKERS);
+	app.start(SERVER, PORT, NUM_WORKERS);
 
 }
 
@@ -22,35 +28,43 @@ void LTSor::consumeInput() {
 	tuple *synchLock = make_tuple("s", SYNCH_LOCK);
 	put_tuple(synchLock, &ctx);
 
-	// initialize
-	for (index_t r = 0; r < SOR_NR; ++r) {
+	// Initialise. note that as these can run in processes, the
+	// solution vector must live in tuple space to be communicated
+	// between the workers. Setting the solution variable is LOCAL ONLY
+	Matrix matrix = (Matrix) inputs[0];
+	Vector target = (Vector) inputs[1];
+	Vector solution = (Vector) outputs[0];
+	for (index_t r = 0; r < SOR_N; ++r) {
 		VECTOR(solution, r) = 1.0;
 	}
 
 	// put a tuple for the solution vector into tuple space
-	Vector solution = (Vector) outputs[0];
 	tuple *solutionTuple = make_tuple("ss", SOLUTION_VECTOR);
 	solutionTuple->elements[1].data.s.len = sizeof(solution);
-	solutionTuple->elements[1].data.s.ptr = solution;
+	solutionTuple->elements[1].data.s.ptr = (char*) solution;
 
 	// loop until we get the desired tolerance
 	real maxDiff = (real)(2 * SOR_TOLERANCE);
 	for (index_t t = 0; (t < SOR_MAX_ITERS) && (maxDiff >= SOR_TOLERANCE); t++) {
 
[…]

### 2009-08-05 cowichan_lt/src/sor.hpp

@@ -11,13 +11,13 @@
 		void work();
 		void produceOutput();
 
-		static const char* SYNCH_LOCK = "sor synch lock";
-		static const char* ROWS_DONE = "sor rows reporting";
+		static const char* SYNCH_LOCK;
+		static const char* ROWS_DONE;
 
-		static 	const char* SOLUTION_VECTOR = "sor solution row";
-		static const char* SOLUTION_SUM = "sor inner sum";
+		static const char* SOLUTION_VECTOR;
+		static const char* SOLUTION_SUM;
 
-		static const char* SOR_FLAG = "sor input consumed";
+		static const char* SOR_FLAG;
 
 		real solutionSum(index_t row);
 

### 2009-08-05 cowichan_lt/src/thresh.cpp

@@ -1,6 +1,10 @@
 #include <iostream>
 #include <cstdio>
+#include <map>
 #include "thresh.hpp"
+
+const char* LTFrequency::SYNCH_LOCK = "thresh synch lock";
+const char* LTFrequency::ROWS_DONE = "thresh rows reporting";
 
 void CowichanLinuxTuples::thresh(IntMatrix matrix, BoolMatrix outMatrix) {
 
@@ -36,6 +40,8 @@ void LTFrequency::consumeInput() {
 
 void LTFrequency::work() {
 
+	IntMatrix matrix = (IntMatrix) inputs[0];
+
 	// tuple templates
 	tuple *synchLock = make_tuple("s", SYNCH_LOCK);
 	tuple *recv = make_tuple("s?", "thresh request");
@@ -47,8 +53,9 @@ void LTFrequency::work() {
 		size_t y = gotten->elements[1].data.i;
 
 		// perform the actual computation for this row (counting)
-		for (int x = 0; x < THRESH_NC; ++x) {
-			// TODO create a map starting at all 0s and increment frequency
+		std::map<INT_TYPE, size_t> freq;
+		for (index_t x = 0; x < THRESH_NC; ++x) {
+			++freq[MATRIX_RECT_NC(matrix, y,x, THRESH_NC)];
 		}
 
 		// purge local memory of the tuple we received
@@ -59,15 +66,16 @@ void LTFrequency::work() {
 		get_tuple(synchLock, &ctx);
 
 			// combine results with master copy
-			for (std::iterator on map returning pairs) {
-				tuple *templateHist = make_tuple("sii", "thresh hist", n);
+			std::map<INT_TYPE, size_t>::const_iterator it;
+			for (it = freq.begin(); it != freq.end(); ++it) {
+				tuple *templateHist = make_tuple("sii", "thresh hist", it->first);
 				tuple *hist = get_nb_tuple(templateHist, &ctx);
 				if (hist == NULL) {
 					// use only this row's frequency
-					templateHist->elements[2].data.i = freq[n];
+					templateHist->elements[2].data.i = it->second;
 				} else {
 					// combine this row's frequency and existing frequency
-					templateHist->elements[2].data.i = freq[n] + hist->elements[2].data.i;
+					templateHist->elements[2].data.i = it->second + hist->elements[2].data.i;
 					destroy_tuple(hist);
 				}
 				put_tuple(templateHist, &ctx);
[…]

### 2009-08-05 cowichan_lt/src/thresh.hpp

@@ -11,8 +11,8 @@
 		void work();
 		void produceOutput();
 
-		static const char* SYNCH_LOCK = "thresh synch lock";
-		static const char* ROWS_DONE = "thresh rows reporting";
+		static const char* SYNCH_LOCK;
+		static const char* ROWS_DONE;
 
 	};
 

### 2009-08-05 cowichan_lt/src/vecdiff.cpp

@@ -3,16 +3,16 @@
 #include <cmath>
 #include "vecdiff.hpp"
 
-const char* LTVecDiff::SYNCH_LOCK = "vecdiff synch lock";
-const char* LTVecDiff::ELEMENTS_DONE = "vecdiff elements reporting";
-const char* LTVecDiff::MAX_DIFF = "vecdiff max difference";
+const char* LTVecdiff::SYNCH_LOCK = "vecdiff synch lock";
+const char* LTVecdiff::ELEMENTS_DONE = "vecdiff elements reporting";
+const char* LTVecdiff::MAX_DIFF = "vecdiff max difference";
 
 real CowichanLinuxTuples::vecdiff(Vector actual, Vector computed) {
 
 	real answer;
 
 	// calculate the 2D bounds of the point cloud
-	LTVecDiff program;
+	LTVecdiff program;
 	program.addInput(0, actual);
 	program.addInput(0, computed);
 	program.addOutput(0, &answer);
@@ -41,10 +41,10 @@ void LTVecdiff::consumeInput() {
 
 	// split points, based on a cluster size of the square-root of the
 	// number of elements in the given vectors.
-	size_t skip = (size_t) sqrt((real) VECDIFF_N);
-	for (size_t pos = 0; pos < VECDIFF_N; pos += skip) {
+	index_t skip = (size_t) sqrt((real) VECDIFF_N);
+	for (index_t pos = 0; pos < VECDIFF_N; pos += skip) {
 		send->elements[1].data.i = pos;
-		send->elements[2].data.i = min(pos + skip, VECDIFF_N);
+		send->elements[2].data.i = std::min(pos + skip, VECDIFF_N);
 		put_tuple(send, &ctx);
 	}
 
@@ -70,8 +70,8 @@ void LTVecdiff::work() {
 		// perform the actual computation for these elements (/max)
 		real maximum = 0.0;
 		for (size_t pos = start; pos < stop; ++pos) {
-			real thisMaximum = (real) fabs(input[pos] - actual[pos]);
-			maximum = max(thisMaximum, maximum);
+			real thisMaximum = (real) fabs(computed[pos] - actual[pos]);
+			maximum = std::max(thisMaximum, maximum);
 		}
 
 		// purge local memory of the tuple we received
@@ -85,7 +85,7 @@ void LTVecdiff::work() {
 			tuple *tmpMax = make_tuple("s?", MAX_DIFF);
[…]

### 2009-08-06 cowichan_lt/src/cowichan_lt.cpp
Things are starting to work a little bit better, but LinuxTuples has some serious issues. The console gets flooded with valgrind errors from the library!

@@ -10,7 +10,7 @@
  */
 int main(int argc, char* argv[]) {
 	Cowichan* linuxTuples = new CowichanLinuxTuples();
-	linuxTuples->main(argc, argv, false, true);
+	linuxTuples->main(argc, argv, true, true);
 	return 0;
 }
 

### 2009-08-06 cowichan_lt/src/half.cpp

@@ -21,24 +21,18 @@ void CowichanLinuxTuples::half(IntMatrix matrixIn, IntMatrix matrixOut) {
 
 void LTHalf::consumeInput() {
 
-	// tuple template
-	tuple *send = make_tuple("si", REQUEST);
-
 	// send off a request for each grid row.
 	for (size_t y = 0; y < HALF_NR; ++y) {
-		send->elements[1].data.i = y;
+		tuple *send = make_tuple("si", REQUEST, y);
 		put_tuple(send, &ctx);
+		destroy_tuple(send);
 	}
 	
-	// destroy the template tuple
-	destroy_tuple(send);
-
 }
 
 void LTHalf::work() {
 
 	tuple *recv = make_tuple("s?", REQUEST);
-	tuple *send = make_tuple("sis", DONE);
 	
 	// grab pointers locally.
 	IntMatrix input = (IntMatrix) inputs[0];
@@ -52,7 +46,7 @@ void LTHalf::work() {
 		// copy over row co-ordinate of the computation; create
 		// a buffer for the results of the computation.
 		size_t y = gotten->elements[1].data.i;
-		send->elements[1].data.i = y;
+		tuple *send = make_tuple("sis", DONE, y, "");
 		IntVector buffer = NEW_VECTOR_SZ(INT_TYPE, HALF_NC);
 		send->elements[2].data.s.len = sizeof(INT_TYPE) * HALF_NC;
 		send->elements[2].data.s.ptr = (char*) buffer;
@@ -66,12 +60,12 @@ void LTHalf::work() {
 		// send off the new tuple and purge local memory of the one we got
 		put_tuple(send, &ctx);
 		destroy_tuple(gotten);
+		destroy_tuple(send);
 		delete[] buffer;
 
 	}
 
 	// TODO destroy the template tuples; must send tuples for this
-//	destroy_tuple(send);
 //	destroy_tuple(recv);
 
 }

### 2009-08-06 cowichan_lt/src/life.cpp

@@ -9,10 +9,19 @@
 #include "life.hpp"
 
 void CowichanLinuxTuples::life(BoolMatrix matrixIn, BoolMatrix matrixOut) {
-	LTLife app;
-	app.addInput(0, matrixIn);
-	app.addOutput(0, matrixOut);
-	app.start(SERVER, PORT, NUM_WORKERS);
+
+	// overwrite the output matrix
+	memcpy(matrixOut, matrixIn,
+		sizeof(bool) * LIFE_NR * LIFE_NC);
+
+	// operate in-place on matrixOut the number of iterations we should.
+	for (size_t i = 0; i < LIFE_ITERATIONS; ++i) {
+		LTLife app;
+		app.addInput(0, matrixOut);
+		app.addOutput(0, matrixOut);
+		app.start(SERVER, PORT, NUM_WORKERS);
+	}
+
 }
 
 /**
@@ -64,7 +73,6 @@ void LTLife::consumeInput() {
 void LTLife::work() {
 
 	tuple *recv = make_tuple("s?", "life request");
-	tuple *send = make_tuple("sis", "life done");
 	
 	// grab pointers locally.
 	BoolMatrix input = (BoolMatrix) inputs[0];
@@ -78,8 +86,8 @@ void LTLife::work() {
 		// copy over row co-ordinate of the computation; create
 		// a buffer for the results of the computation.
 		size_t y = gotten->elements[1].data.i;
-		send->elements[1].data.i = y;
-		bool* buffer = (bool*) malloc(sizeof(bool) * LIFE_NC);
+		tuple *send = make_tuple("sis", "life done", y, "");
+		BoolVector buffer = (BoolVector) NEW_VECTOR_SZ(bool, LIFE_NC);
 		send->elements[2].data.s.len = sizeof(bool) * LIFE_NC;
 		send->elements[2].data.s.ptr = (char*) buffer;
 
@@ -100,19 +108,16 @@ void LTLife::work() {
 		// send off the new tuple and purge local memory of the one we got
 		put_tuple(send, &ctx);
 		destroy_tuple(gotten);
+		destroy_tuple(send);
 
 	}
 
 	// TODO destroy the template tuples; must send tuples for this
-//	destroy_tuple(send);
 //	destroy_tuple(recv);
 
 }
 
 void LTLife::produceOutput() {
-
-	// tuple template
-	tuple *recv = make_tuple("s??", "life done");
 
 	// grab output pointer locally.
[…]

### 2009-08-06 cowichan_lt/src/mandel.cpp

@@ -16,18 +16,14 @@ void CowichanLinuxTuples::mandel(IntMatrix matrix) {
 
 void LTMandel::consumeInput() {
 
-	// tuple template
-	tuple *send = make_tuple("si", "mandel request");
-
 	// send off a mandelbrot request for each grid row.
 	for (size_t y = 0; y < MANDEL_NR; ++y) {
+		tuple *send = make_tuple("si", "mandel request");
 		send->elements[1].data.i = y;
 		put_tuple(send, &ctx);
+		destroy_tuple(send);
 	}
 	
-	// destroy the template tuple
-	destroy_tuple(send);
-
 }
 
 int LTMandel::mandelCalc(real x, real y) {
@@ -59,7 +55,6 @@ int LTMandel::mandelCalc(real x, real y) {
 void LTMandel::work() {
 
 	tuple *recv = make_tuple("s?", "mandel request");
-	tuple *send = make_tuple("sis", "mandel done");
 	
 	// 2D delta between calculated points
 	real dX = MANDEL_DX / (MANDEL_NC - 1);
@@ -73,7 +68,7 @@ void LTMandel::work() {
 
 		// copy over row co-ordinate of the computation; create
 		// a buffer for the results of the computation.
-		send->elements[1].data.i = gotten->elements[1].data.i;
+		tuple *send = make_tuple("sis", "mandel done", gotten->elements[1].data.i, "");
 		int* buffer = (int*) malloc(sizeof(int) * MANDEL_NC);
 		send->elements[2].data.s.len = sizeof(int) * MANDEL_NC;
 		send->elements[2].data.s.ptr = (char*) buffer;
@@ -88,6 +83,7 @@ void LTMandel::work() {
 		// send off the new tuple and purge local memory of the one we gotten
 		put_tuple(send, &ctx);
 		destroy_tuple(gotten);
+		destroy_tuple(send);
 
 	}
 

### 2009-08-06 cowichan_lt/src/randmat.cpp

@@ -70,7 +70,6 @@ void LTRandmat::consumeInput() {
 void LTRandmat::work() {
 
 	tuple *recv = make_tuple("s?", REQUEST);
-	tuple *send = make_tuple("sis", "randmat done", 0, "");
 	tuple *tmpInit = make_tuple("s?", "randmat state");
 	tuple *init = NULL;
 	IntVector initVector = NULL;
@@ -90,8 +89,8 @@ void LTRandmat::work() {
 		// copy over row co-ordinate of the computation; create
 		// a buffer for the results of the computation.
 		size_t row = gotten->elements[1].data.i;
-		send->elements[1].data.i = row;
-		int* buffer = (int*) malloc(sizeof(INT_TYPE) * RANDMAT_NC);
+		tuple *send = make_tuple("sis", "randmat done", row, "");
+		IntVector buffer = (IntVector) NEW_VECTOR_SZ(INT_TYPE, RANDMAT_NC);
 		send->elements[2].data.s.len = sizeof(INT_TYPE) * RANDMAT_NC;
 		send->elements[2].data.s.ptr = (char*) buffer;
 
@@ -104,11 +103,12 @@ void LTRandmat::work() {
 		// send off the new tuple and purge local memory of the one we got
 		put_tuple(send, &ctx);
 		destroy_tuple(gotten);
+		destroy_tuple(send);
+		delete[] buffer;
 
 	}
 
 	// TODO destroy the template tuples; must send tuples for this
-//	destroy_tuple(send);
 //	destroy_tuple(recv);
 
 }

### 2009-08-06 cowichan_lt/src/thresh.cpp

@@ -40,19 +40,20 @@ void LTFrequency::consumeInput() {
 	// create a tuple synch lock
 	tuple *synchLock = make_tuple("s", SYNCH_LOCK);
 	put_tuple(synchLock, &ctx);
+	destroy_tuple(synchLock);
 
 	// create a "rows reporting" tuple, so that we
 	// know when the computation should end.
 	tuple *rowsReporting = make_tuple("si", ROWS_DONE, 0);
 	put_tuple(rowsReporting, &ctx);
+	destroy_tuple(rowsReporting);
 
 	// communicate all of the rows
-	tuple *send = make_tuple("s?", REQUEST);
 	for (size_t y = 0; y < THRESH_NR; ++y) {
-		send->elements[1].data.i = y;
+		tuple *send = make_tuple("si", REQUEST, y);
 		put_tuple(send, &ctx);
+		destroy_tuple(send);
 	}
-
 
 }
 
@@ -63,6 +64,7 @@ void LTFrequency::work() {
 	// tuple templates
 	tuple *synchLock = make_tuple("s", SYNCH_LOCK);
 	tuple *recv = make_tuple("s?", REQUEST);
+	tuple *templateRowsReporting = make_tuple("s?", ROWS_DONE);
 
 	while (1) {
 
@@ -101,10 +103,10 @@ void LTFrequency::work() {
 			}
 
 			// record the number of rows reporting
-			tuple *templateRowsReporting = make_tuple("s?", ROWS_DONE);
 			tuple *rowsReporting = get_tuple(templateRowsReporting, &ctx);
 			rowsReporting->elements[1].data.i += 1;
 			put_tuple(rowsReporting, &ctx);
+			destroy_tuple(rowsReporting);
 
 		// leave the critical section
 		put_tuple(synchLock, &ctx);
@@ -117,7 +119,8 @@ void LTFrequency::produceOutput() {
 
 	// wait for all rows to be done.
 	tuple *allRowsReporting = make_tuple("si", ROWS_DONE, THRESH_NR);
-	get_tuple(allRowsReporting, &ctx);
+	destroy_tuple(get_tuple(allRowsReporting, &ctx));
+	destroy_tuple(allRowsReporting);
 
 	// figure out when we have to stop.
 	index_t retain = (index_t)(THRESH_PERCENT * THRESH_NC * THRESH_NR);
@@ -129,22 +132,34 @@ void LTFrequency::produceOutput() {
 		if (hist != NULL) {
 			retain -= hist->elements[2].data.i;
 		}
[…]

### 2009-08-06 cowichan_lt/src/winnow.cpp

@@ -41,6 +41,7 @@ void LTWinnow::consumeInput() {
 	// create a tuple synch lock
 	tuple *synchLock = make_tuple("s", SYNCH_LOCK);
 	put_tuple(synchLock, &ctx);
+	destroy_tuple(synchLock);
 
 	// tuple template.
 	tuple *send = make_tuple("si", REQUEST, 0);
@@ -55,6 +56,7 @@ void LTWinnow::consumeInput() {
 		send->elements[1].data.i = row;
 		put_tuple(send, &ctx);
 	}
+	destroy_tuple(send);
 
 }
 
@@ -64,6 +66,7 @@ void LTWinnow::work() {
 	tuple *synchLock = make_tuple("s", SYNCH_LOCK);
 	tuple *recv = make_tuple("s?", REQUEST);
 	tuple *send = make_tuple("ssi", WEIGHTED_POINT, "", 0);
+	tuple *templateRowsReporting = make_tuple("s?", ROWS_DONE);
 
 	// grab pointers locally.
 	IntMatrix matrix = (IntMatrix) inputs[0];
@@ -113,10 +116,10 @@ void LTWinnow::work() {
 			destroy_tuple(tmpCount);
 
 			// record the number of rows reporting
-			tuple *templateRowsReporting = make_tuple("s?", ROWS_DONE);
 			tuple *rowsReporting = get_tuple(templateRowsReporting, &ctx);
 			rowsReporting->elements[1].data.i += 1;
 			put_tuple(rowsReporting, &ctx);
+			destroy_tuple(rowsReporting);
 
 		// leave the critical section
 		put_tuple(synchLock, &ctx);
@@ -145,6 +148,7 @@ Point LTWinnow::nextWeightedPoint(INT_TYPE* order) {
 	// strip data out of the received tuple and destroy it.
 	Point pt = *((Point*) gotten->elements[1].data.s.ptr);
 	destroy_tuple(gotten);
+	destroy_tuple(recv);
 
 	// return the point.
 	return pt;
@@ -156,6 +160,7 @@ void LTWinnow::produceOutput() {
 	// wait for all rows to be computed.
 	tuple *rowsReporting = make_tuple("si", ROWS_DONE, WINNOW_NR);
 	destroy_tuple(get_tuple(rowsReporting, &ctx));
+	destroy_tuple(rowsReporting);
 
 	// grab output pointers locally.
 	PointVector points = (PointVector) outputs[0];
[…]

### 2009-08-07 cowichan_lt/comp64.sh
Improved doxygen.

@@ -1,3 +1,3 @@
-#!/bin/sh
-g++ -Wall -m64 -fopenmp -O2 -c ../cowichan/cowichan.cpp
-g++ -Wall -m64 -fopenmp -O2 -o cowichan_openmp *.cpp cowichan.o
+#!/bin/bash
+g++ -Wall -m64 -Os -Iinclude -c ../cowichan/cowichan.cpp
+g++ -Wall -m64 -Os -llinuxtuples -Wno-write-strings -I../linuxtuples-1.03 -o cowichan_lt src/*.cpp cowichan.o

### 2009-08-07 cowichan_lt/src/cowichan_lt.hpp

@@ -9,6 +9,9 @@
 
 // COWICHAN DEFINITIONS =====================================================//
 // aka. "inputs" to the toys, and chaining functions.
+/**
+ * The LinuxTuples implementation of the Cowichan problem set.
+ */
 class CowichanLinuxTuples : public Cowichan {
 public:
 

### 2009-08-07 cowichan_lt/src/gauss.hpp

@@ -21,11 +21,26 @@
 		void work();
 		void produceOutput();
 
+		/**
+		 * Number of rows computed.
+		 */
 		static const char* ROWS_DONE;
+		/**
+		 * Is forward-elimination complete?
+		 */
 		static const char* FORWARD_DONE;
+		/**
+		 * A matrix row computation request.
+		 */
 		static const char* REQUEST;
 
+		/**
+		 * The target vector stored in the tuple space.
+		 */
 		static const char* TARGET;
+		/**
+		 * One of the matrix rows, stored in the tuple space.
+		 */
 		static const char* MATRIX_ROW;
 
 	};

### 2009-08-07 cowichan_lt/src/half.hpp

@@ -19,7 +19,16 @@
 		void consumeInput();
 		void work();
 		void produceOutput();
-		static const char *REQUEST, *DONE;
+
+		/**
+		 * A request to perform computation on a row of the matrix.
+		 */
+		static const char *REQUEST;
+
+		/**
+		 * Done with the halving shuffle.
+		 */
+		static const char *DONE;
 
 	};
 

### 2009-08-07 cowichan_lt/src/hull.hpp

@@ -15,6 +15,14 @@
 	 */
 	#define NO_POINT -1
 
+	/**
+	 * Performs an iterative convex hull in tuple space.
+	 * Iterative convex hull means that the output contains the convex hull of
+	 * all points, followed by the convex hull of those points not on the first
+	 * convex hull, followed by the convex hull of those points not on the first
+	 * or second, etc. etc. until no points remain. Thus the iterative convex hull
+	 * can be seen as a spiralling shape, or a permutation of the input points.
+	 */
 	class LTHull: public TupleApplication {
 	public:
 

### 2009-08-07 cowichan_lt/src/invperc.cpp

@@ -7,6 +7,9 @@
 #include "cowichan_lt.hpp"
 #include <queue>
 
+/**
+ * Point that also stores matrix value at its given row and column.
+ */
 class PercPoint {
 public:
 

### 2009-08-07 cowichan_lt/src/mandel.hpp

@@ -19,6 +19,13 @@
 		void consumeInput();
 		void work();
 		void produceOutput();
+
+		/**
+		 * Computes a part of the mandelbrot set.
+		 * \param x x co-ordinate to compute for.
+		 * \param y y co-ordinate to compute for.
+		 * \return the computed value at (x,y) on the mandelbrot set.
+		 */
 		int mandelCalc(real x, real y);
 	};
 

### 2009-08-07 cowichan_lt/src/norm.hpp

@@ -22,9 +22,21 @@
 		void work();
 		void produceOutput();
 
+		/**
+		 * Synchronization lock (critical section).
+		 */
 		static const char* SYNCH_LOCK;
+		/**
+		 * Number of points that have been processed.
+		 */
 		static const char* POINTS_DONE;
+		/**
+		 * The smallest point (in x and y) found
+		 */
 		static const char* MIN_POINT;
+		/**
+		 * The largest point (in x and y) found
+		 */
 		static const char* MAX_POINT;
 
 	};
@@ -40,7 +52,13 @@
 		void work();
 		void produceOutput();
 
+		/**
+		 * The smallest point (in x and y) found
+		 */
 		static const char* MIN_POINT;
+		/**
+		 * The largest point (in x and y) found
+		 */
 		static const char* MAX_POINT;
 
 	};

### 2009-08-07 cowichan_lt/src/outer.hpp

@@ -20,10 +20,22 @@
 		void work();
 		void produceOutput();
 
+		/**
+		 * Synchronization lock (critical section).
+		 */
 		static const char* SYNCH_LOCK;
+		/**
+		 * Request to compute for one vector.
+		 */
 		static const char* REQUEST;
 
+		/**
+		 * The maximum length of any vector.
+		 */
 		static const char* MAX_DISTANCE;
+		/**
+		 * An entry in the matrix, to be stored in the output.
+		 */
 		static const char* MATRIX_ENTRY;
 
 	};

### 2009-08-07 cowichan_lt/src/sor.hpp

@@ -10,6 +10,11 @@
 	#include "tuple_common.hpp"
 	#include "cowichan_lt.hpp"
 
+	/**
+	 * Performs a successive over-relaxation to solve a matrix
+	 * problem Ax = b. The solution is approximate, but it can be made
+	 * arbitrarily accurate. The computation is done in tuple space.
+	 */
 	class LTSor: public TupleApplication {
 	protected:
 

### 2009-08-07 cowichan_lt/src/thresh.hpp

@@ -22,9 +22,21 @@
 		void work();
 		void produceOutput();
 
+		/**
+		 * Synchronization lock (critical section).
+		 */
 		static const char* SYNCH_LOCK;
+		/**
+		 * Number of rows computed for.
+		 */
 		static const char* ROWS_DONE;
+		/**
+		 * A request to compute a row.
+		 */
 		static const char* REQUEST;
+		/**
+		 * The breaking-point for the thresholding process.
+		 */
 		static const char* POINT;
 
 	};
@@ -41,9 +53,18 @@
 		void work();
 		void produceOutput();
 
+		/**
+		 * A request to compute a row.
+		 */
 		static const char* REQUEST;
-		static const char* DONE;
+		/**
+		 * The breaking-point for the thresholding process.
+		 */
 		static const char* POINT;
+		/**
+		 * Flagged when finished.
+		 */
+		static const char* DONE;
 
 	};
 

### 2009-08-07 cowichan_lt/src/winnow.cpp

@@ -193,11 +193,11 @@ void LTWinnow::produceOutput() {
 
 		// skip over as many points as we need to
 		for (index_t i = 0; i < stride; ++i) {
-			//current = nextWeightedPoint(&order);
+			current = nextWeightedPoint(&order);
 		}
 
 		// put one in the list
-		points[pos++] = Point(0, 0); // current;
+		points[pos++] = current;
 
 	}
 

### 2009-08-07 cowichan_lt/src/winnow.hpp

@@ -28,10 +28,25 @@
 		 */
 		Point nextWeightedPoint(INT_TYPE* order);
 
+		/**
+		 * Synchronization lock (critical section).
+		 */
 		static const char* SYNCH_LOCK;
+		/**
+		 * Request to perform computation on a matrix row.
+		 */
 		static const char* REQUEST;
+		/**
+		 * A weighted point in the tuple space.
+		 */
 		static const char* WEIGHTED_POINT;
+		/**
+		 * The number of weighted points in the tuple space.
+		 */
 		static const char* COUNT;
+		/**
+		 * The number of rows that have had their computation finish.
+		 */
 		static const char* ROWS_DONE;
 
 	};

### 2009-08-07 cowichan_lt/src/hull.cpp
Improved doxygen.

@@ -120,11 +120,9 @@ void LTHull::consumeInput() {
 
 /**
  * Compute hull on one side of the splitting line.
- * @param pointsIn hull input points.
- * @param n number of input points.
- * @param pointsOut hull output points.
- * @param p1 boundary point #1.
- * @param p2 boundary point #2.
+ * \param p1 boundary point #1.
+ * \param p2 boundary point #2.
+ * \param order the ordering number of the next hull point to emit.
  */
 void LTHull::split(const index_t p1, const index_t p2, index_t *order) {
 

### 2009-08-07 cowichan_lt/src/hull.hpp

@@ -33,37 +33,82 @@
 
 		/**
 		 * Constructor.
-		 * \param numLeft the number of points left out of the original batch.
+		 * \param numLeft the number of points left unmasked in the input.
 		 */
 		LTHull(size_t numLeft);
 
-		// public, so we can clean-up outside of LTHull's tuple app.
+		/**
+		 * The index of a point that is masked off (i.e. already selected) in the input.
+		 * This is public so we can access it from CowichanLinuxTuples::hull(...).
+		 */
 		static const char* MASKED_POINT;
 
 	protected:
 
+		/**
+		 * The number of points left unmasked in the input.
+		 */
 		size_t numLeft;
 
 		void consumeInput();
 		void work();
 		void produceOutput();
 
+		/**
+		 * Generic request for work to be done.
+		 */
 		static const char* REQUEST;
+		/**
+		 * Specific work: min-max request over a length of points.
+		 */
 		static const char* REQUEST_MINMAX;
+		/**
+		 * Specific work: cross-product request over a length of points.
+		 */
 		static const char* REQUEST_CROSS;
 
+		/**
+		 * Synchronization lock (critical section).
+		 */
 		static const char* SYNCH_LOCK;
+		/**
+		 * The number of points that have been computed for.
+		 */
 		static const char* POINTS_DONE;
 
+		/**
+		 * The point with lowest x value, unmasked.
+		 */
 		static const char* MIN_X_POINT;
+		/**
+		 * The point with highest x value, unmasked.
+		 */
 		static const char* MAX_X_POINT;
+		/**
+		 * The cross product value of the "furthest" point.
+		 */
 		static const char* MAX_CROSS;
+		/**
+		 * The "furthest" point found.
+		 */
 		static const char* MAX_POINT;
 
+		/**
+		 * A point on the hull, ordered.
+		 */
 		static const char* HULL_POINT;
+		/**
+		 * The number of points in this convex hull.
+		 */
 		static const char* NUM_POINTS;
 
+		/**
+		 * Should the output producer run now?
+		 */
 		static const char* FLAG_OUTPUT;
+		/**
+		 * The min-max process mandated has been finished.
+		 */
 		static const char* FINISHED_MINMAX;
 
 		/**

### 2009-08-07 cowichan_lt/src/invperc.cpp

@@ -13,19 +13,38 @@
 class PercPoint {
 public:
 
+  /**
+   * Location on the matrix.
+   */
   Point point;
+  /**
+   * Matrix to draw values from.
+   */
   IntMatrix matrix;
+  /**
+   * Number of columns in the given matrix.
+   */
   index_t nc;
 
 public:
 
+  /**
+   * Construct with matrix location.
+   * \param point place on the matrix to "claim"
+   */
   PercPoint(Point point): point(point) { }
 
-  // we want to extract lowest values.
+  /**
+   * we want to extract lowest values.
+   */
   bool operator<(const PercPoint &other) const {
     return value() > other.value();
   }
 
+  /**
+   * Figure out matrix value at the point's position.
+   * \return the value of the given matrix at this position.
+   */
   INT_TYPE value() const {
     return MATRIX_RECT(matrix, (index_t)point.y, (index_t)point.x);
   }

### 2009-08-07 cowichan_lt/src/randmat.hpp

@@ -14,7 +14,11 @@
 	class LTRandmat: public TupleApplication {
 	protected:
 
-		INT_TYPE aPrime, cPrime;
+		/**
+		 * A and C values to use for the next-K parallel randmat computation.
+		 */
+		INT_TYPE aPrime;
+		INT_TYPE cPrime;
 
 		void consumeInput();
 		void work();
@@ -32,6 +36,9 @@
 		 */
 		INT_TYPE next(INT_TYPE& current) const;
 
+		/**
+		 * Request for a grid row to be generated.
+		 */
 		static const char *REQUEST;
 
 	};

### 2009-08-07 cowichan_lt/src/sor.hpp

@@ -22,12 +22,27 @@
 		void work();
 		void produceOutput();
 
+		/**
+		 * Synchronization lock (critical section).
+		 */
 		static const char* SYNCH_LOCK;
+		/**
+		 * The number of rows done in the middle SOR section.
+		 */
 		static const char* ROWS_DONE;
 
+		/**
+		 * The vector of the solution in tuple space
+		 */
 		static const char* SOLUTION_VECTOR;
+		/**
+		 * The on-going sum of the solution.
+		 */
 		static const char* SOLUTION_SUM;
 
+		/**
+		 * Should we flag the output producer?
+		 */
 		static const char* SOR_FLAG;
 
 		/**

### 2009-08-07 cowichan_lt/src/vecdiff.hpp

@@ -20,8 +20,17 @@
 		void work();
 		void produceOutput();
 
+		/**
+		 * Synchronization lock (critical section).
+		 */
 		static const char* SYNCH_LOCK;
+		/**
+		 * The number of differences already accomplished.
+		 */
 		static const char* ELEMENTS_DONE;
+		/**
+		 * The maximum difference found, so far.
+		 */
 		static const char* MAX_DIFF;
 
 	};

### 2009-08-07 cowichan_lt/src/randmat.hpp
Final doxygen fix.

@@ -15,9 +15,12 @@
 	protected:
 
 		/**
-		 * A and C values to use for the next-K parallel randmat computation.
+		 * A value to use for the next-K parallel randmat computation.
 		 */
 		INT_TYPE aPrime;
+		/**
+		 * C value to use for the next-K parallel randmat computation.
+		 */
 		INT_TYPE cPrime;
 
 		void consumeInput();

### 2009-08-07 doc/latex/annotated.tex
doxygen documentation as of now
Re-generated doxygen with linuxtuples additions.

@@
 \item\contentsline{section}{\hyperlink{class_cowichan}{Cowichan} (Base class for all C++ implementations )}{\pageref{class_cowichan}}{}
+\item\contentsline{section}{\hyperlink{class_cowichan_linux_tuples}{CowichanLinuxTuples} }{\pageref{class_cowichan_linux_tuples}}{}
+\item\contentsline{section}{\hyperlink{class_cowichan_m_p_i}{CowichanMPI} (Boost Message Passing Interface (MPI) implementation )}{\pageref{class_cowichan_m_p_i}}{}
 \item\contentsline{section}{\hyperlink{class_cowichan_open_m_p}{CowichanOpenMP} (Open Multi-Processing (OpenMP) implementation )}{\pageref{class_cowichan_open_m_p}}{}
@@
 \item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_histogram}{cowichan\_\-tbb::Histogram} (This class calculates the histogram of a matrix )}{\pageref{classcowichan__tbb_1_1_histogram}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_bounds}{LTBounds} }{\pageref{class_l_t_bounds}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_forward}{LTForward} }{\pageref{class_l_t_forward}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_frequency}{LTFrequency} }{\pageref{class_l_t_frequency}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_half}{LTHalf} }{\pageref{class_l_t_half}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_hull}{LTHull} }{\pageref{class_l_t_hull}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_life}{LTLife} }{\pageref{class_l_t_life}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_mandel}{LTMandel} }{\pageref{class_l_t_mandel}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_norm}{LTNorm} }{\pageref{class_l_t_norm}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_outer}{LTOuter} }{\pageref{class_l_t_outer}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_product}{LTProduct} }{\pageref{class_l_t_product}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_randmat}{LTRandmat} }{\pageref{class_l_t_randmat}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_sor}{LTSor} }{\pageref{class_l_t_sor}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_thresh}{LTThresh} }{\pageref{class_l_t_thresh}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_vecdiff}{LTVecdiff} }{\pageref{class_l_t_vecdiff}}{}
+\item\contentsline{section}{\hyperlink{class_l_t_winnow}{LTWinnow} }{\pageref{class_l_t_winnow}}{}
 \item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_make_dominant}{cowichan\_\-tbb::MakeDominant} (Makes a given matrix diagonally dominant )}{\pageref{classcowichan__tbb_1_1_make_dominant}}{}
@@
 \item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_norm_diff}{cowichan\_\-tbb::NormDiff} (This class takes the 1-norm of the difference between two vectors )}{\pageref{classcowichan__tbb_1_1_norm_diff}}{}
+\item\contentsline{section}{\hyperlink{class_perc_point}{PercPoint} }{\pageref{class_perc_point}}{}
+\item\contentsline{section}{\hyperlink{classcowichan__openmp_1_1_perc_point}{cowichan\_\-openmp::PercPoint} (Percolation point )}{\pageref{classcowichan__openmp_1_1_perc_point}}{}
+\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_perc_point}{cowichan\_\-tbb::PercPoint} (Percolation point )}{\pageref{classcowichan__tbb_1_1_perc_point}}{}
 \item\contentsline{section}{\hyperlink{classcowichan__serial_1_1_perc_point}{cowichan\_\-serial::PercPoint} (Percolation point )}{\pageref{classcowichan__serial_1_1_perc_point}}{}
-\item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_perc_point}{cowichan\_\-tbb::PercPoint} (Percolation point )}{\pageref{classcowichan__tbb_1_1_perc_point}}{}
-\item\contentsline{section}{\hyperlink{classcowichan__openmp_1_1_perc_point}{cowichan\_\-openmp::PercPoint} (Percolation point )}{\pageref{classcowichan__openmp_1_1_perc_point}}{}
 \item\contentsline{section}{\hyperlink{class_point}{Point} (Two-dimensional point )}{\pageref{class_point}}{}
@@
 \item\contentsline{section}{\hyperlink{classcowichan__tbb_1_1_threshold}{cowichan\_\-tbb::Threshold} (Creates a boolean array based on a cut-off value )}{\pageref{classcowichan__tbb_1_1_threshold}}{}
+\item\contentsline{section}{\hyperlink{class_tuple_application}{TupleApplication} }{\pageref{class_tuple_application}}{}
 \item\contentsline{section}{\hyperlink{class_weighted_point}{WeightedPoint} (Two-dimensional point with an integer weight )}{\pageref{class_weighted_point}}{}

### 2009-08-07 doc/latex/class_cowichan.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-02 doc/latex/class_cowichan.tex
@@
 \leavevmode
-\includegraphics[height=2cm]{class_cowichan}
+\includegraphics[height=1.56643cm]{class_cowichan}
 \end{center}
@@
 \end{Desc}
-\hypertarget{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{
+
+Reimplemented in \hyperlink{class_cowichan_linux_tuples_fd49b2a0af64bafb93a5fe1210c68acd}{CowichanLinuxTuples}.\hypertarget{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{
 \index{Cowichan@{Cowichan}!gauss@{gauss}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_b196d2384c2fe8b185e6269a050399eb}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_363bcd6f0c7b4d0fced94fe4cd59a267}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{CowichanTBB}.\hypertarget{class_cowichan_308603053675bccbe631f04af921f57c}{
+Implemented in \hyperlink{class_cowichan_serial_b196d2384c2fe8b185e6269a050399eb}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_363bcd6f0c7b4d0fced94fe4cd59a267}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_975f8da6c166fe1db3cf9341eeaab000}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_fd3e659afb38415f56fd508333b7bb1b}{CowichanLinuxTuples}.\hypertarget{class_cowichan_308603053675bccbe631f04af921f57c}{
 \index{Cowichan@{Cowichan}!half@{half}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_c08a6b3e23ce26959bac12af077f924f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_70989ffe182aebf590e39c56e146b0fb}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{CowichanTBB}.\hypertarget{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{
+Implemented in \hyperlink{class_cowichan_serial_c08a6b3e23ce26959bac12af077f924f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_70989ffe182aebf590e39c56e146b0fb}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_154bbd1a400ee4571f761bff1cdf67cd}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_f3bd58ad2653f06d242d473e7bffda0b}{CowichanLinuxTuples}.\hypertarget{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{
 \index{Cowichan@{Cowichan}!hull@{hull}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{CowichanTBB}.\hypertarget{class_cowichan_ea126792a31e54a8722663b7ea768955}{
+Implemented in \hyperlink{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_22b3dce35fd93635bd4c1596e7fb839c}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_501f87594d62af261ff0b1954a60ddc4}{CowichanLinuxTuples}.\hypertarget{class_cowichan_ea126792a31e54a8722663b7ea768955}{
 \index{Cowichan@{Cowichan}!invperc@{invperc}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4824c6b8509b5da835fbc5f64eb3e063}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{CowichanTBB}.\hypertarget{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{
+Implemented in \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4824c6b8509b5da835fbc5f64eb3e063}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_e4f9f8c31feea9b3a167fc2880a97610}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_bcd0f18fcccc8973ddaacf4dda190b53}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_5497a17a63cc5faad7afa54d013ea741}{CowichanLinuxTuples}.\hypertarget{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{
 \index{Cowichan@{Cowichan}!life@{life}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{CowichanTBB}.\hypertarget{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{
+Implemented in \hyperlink{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_9c739951d036ef7905bda7cb896e0edf}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_967138bbdead76eec50f31465dd9d5f1}{CowichanLinuxTuples}.\hypertarget{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{
 \index{Cowichan@{Cowichan}!main@{main}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{CowichanTBB}.\hypertarget{class_cowichan_3df21e3c627958114e045c3559a29f30}{
+Implemented in \hyperlink{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_58274fc55ed8a352a8924b130f3b5d2e}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_8774404842868381379ec627e8373cec}{CowichanLinuxTuples}.\hypertarget{class_cowichan_3df21e3c627958114e045c3559a29f30}{
 \index{Cowichan@{Cowichan}!norm@{norm}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{CowichanTBB}.\hypertarget{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{
+Implemented in \hyperlink{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_11df49e804427d4a0d085f4cff9302d0}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_922209c219bb5b9268263dae4000ea81}{CowichanLinuxTuples}.\hypertarget{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{
 \index{Cowichan@{Cowichan}!outer@{outer}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_05f6899081a457d58978e4f6bda2db6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_99a0570f754ac82968d73093476de533}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{CowichanTBB}.\hypertarget{class_cowichan_14e706ef5a5f7661a623c595d7ed76f1}{
+Implemented in \hyperlink{class_cowichan_serial_05f6899081a457d58978e4f6bda2db6a}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_99a0570f754ac82968d73093476de533}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_39fcc8a331b0e26438fbf43fa01cf7c1}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_b16e1c3b60f920c12460a74289558dcb}{CowichanLinuxTuples}.\hypertarget{class_cowichan_14e706ef5a5f7661a623c595d7ed76f1}{
 \index{Cowichan@{Cowichan}!print\_\-rect\_\-matrix@{print\_\-rect\_\-matrix}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{CowichanTBB}.\hypertarget{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{
+Implemented in \hyperlink{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_1b7dccf774caccb26839cd29fe0a5cb0}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_c4874f09c54056ac94d00f9e341a94f4}{CowichanLinuxTuples}.\hypertarget{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{
 \index{Cowichan@{Cowichan}!randmat@{randmat}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{CowichanTBB}.\hypertarget{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{
+Implemented in \hyperlink{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_6805f21144aeaeddc4549e1b2b42bca8}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_ff90c6a27db51366570b70d352de7fab}{CowichanLinuxTuples}.\hypertarget{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{
 \index{Cowichan@{Cowichan}!sor@{sor}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_6e8b06711d976de1adc1e4dc81e560e5}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d6482d0369a26a51ef0e37ab238fc664}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{CowichanTBB}.\hypertarget{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{
+Implemented in \hyperlink{class_cowichan_serial_6e8b06711d976de1adc1e4dc81e560e5}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_d6482d0369a26a51ef0e37ab238fc664}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_7388c844e8aa73ab0923443a3a7ef069}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_aa2c469c3f520a5a56c5ec1ac9a5a6c6}{CowichanLinuxTuples}.\hypertarget{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{
 \index{Cowichan@{Cowichan}!thresh@{thresh}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{CowichanTBB}.\hypertarget{class_cowichan_775d72b5e7d122f9f32555352278250e}{
+Implemented in \hyperlink{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_49ff96b091a61e9cfd9aad8824e7fbbd}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_b8cf055ec3d42a1ccde6d818c5a43a74}{CowichanLinuxTuples}.\hypertarget{class_cowichan_775d72b5e7d122f9f32555352278250e}{
 \index{Cowichan@{Cowichan}!vecdiff@{vecdiff}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{CowichanTBB}.\hypertarget{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{
+Implemented in \hyperlink{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_c5470a2876efecf843b19c37c21ecf19}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_213185666ea7dd0c9abc59899b454086}{CowichanLinuxTuples}.\hypertarget{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{
 \index{Cowichan@{Cowichan}!winnow@{winnow}}
@@
 
-Implemented in \hyperlink{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{CowichanOpenMP}, and \hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{CowichanTBB}.
+Implemented in \hyperlink{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{CowichanSerial}, \hyperlink{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{CowichanOpenMP}, \hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{CowichanTBB}, \hyperlink{class_cowichan_m_p_i_9be48d86fc51ce13ef79d5fe2c8a16e0}{CowichanMPI}, and \hyperlink{class_cowichan_linux_tuples_2b1a77ecbd01ec847cf63fe7ebb50e35}{CowichanLinuxTuples}.
 

### 2009-08-07 doc/latex/cowichan_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan_8cpp}{
\section{cowichan/cowichan.cpp File Reference}
\label{cowichan_8cpp}\index{cowichan/cowichan.cpp@{cowichan/cowichan.cpp}}
}
Implementation for \hyperlink{class_cowichan}{Cowichan} class and common routines for \hyperlink{class_cowichan}{Cowichan} programs.

{\tt \#include \char`\"{}cowichan.hpp\char`\"{}}\par
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{cowichan_8cpp_eb2386a0fab0a548605cd5f0898da5b4}{uniform} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} mean, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} range)
\item 
void \hyperlink{cowichan_8cpp_faa5a0a08b60255e5a59cbd3b1f5e45f}{out\_\-of\_\-memory} ()
\item 
void \hyperlink{cowichan_8cpp_501801b9dcd90a910432c7712ff763e7}{not\_\-enough\_\-points} ()
\item 
void \hyperlink{cowichan_8cpp_19e1551f7e954a984729d96c2c0bf5ba}{no\_\-cells\_\-alive} ()
\item 
INT64 \hyperlink{cowichan_8cpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}{get\_\-ticks} ()
\item 
INT64 \hyperlink{cowichan_8cpp_40356e6c8804a8f94e41c156ad368c83}{get\_\-freq} ()
\item 
void \hyperlink{cowichan_8cpp_560007e135455d06968b8a834d970cf9}{print\_\-elapsed\_\-time} (INT64 start, INT64 end)
\item 
void \hyperlink{cowichan_8cpp_1bf3d816eb607b5a06a97e05e9b95099}{timeInfo} (INT64 $\ast$start, INT64 $\ast$end, std::string message)
\end{CompactItemize}

\subsection{Detailed Description}
Implementation for \hyperlink{class_cowichan}{Cowichan} class and common routines for \hyperlink{class_cowichan}{Cowichan} programs.

\subsection{Function Documentation}
\hypertarget{cowichan_8cpp_40356e6c8804a8f94e41c156ad368c83}{
\index{cowichan.cpp@{cowichan.cpp}!get\_\-freq@{get\_\-freq}}
\index{get\_\-freq@{get\_\-freq}!cowichan.cpp@{cowichan.cpp}}
\subsubsection[{get\_\-freq}]{\setlength{\rightskip}{0pt plus 5cm}INT64 get\_\-freq ()}}
\label{cowichan_8cpp_40356e6c8804a8f94e41c156ad368c83}

Get tick frequency. \begin{Desc}
\item[Returns:]Tick frequency. \end{Desc}
\hypertarget{cowichan_8cpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}{
\index{cowichan.cpp@{cowichan.cpp}!get\_\-ticks@{get\_\-ticks}}
\index{get\_\-ticks@{get\_\-ticks}!cowichan.cpp@{cowichan.cpp}}
\subsubsection[{get\_\-ticks}]{\setlength{\rightskip}{0pt plus 5cm}INT64 get\_\-ticks ()}}
\label{cowichan_8cpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}

Get tick count. \begin{Desc}
\item[Returns:]Tick count. \end{Desc}
\hypertarget{cowichan_8cpp_19e1551f7e954a984729d96c2c0bf5ba}{
\index{cowichan.cpp@{cowichan.cpp}!no\_\-cells\_\-alive@{no\_\-cells\_\-alive}}
\index{no\_\-cells\_\-alive@{no\_\-cells\_\-alive}!cowichan.cpp@{cowichan.cpp}}
\subsubsection[{no\_\-cells\_\-alive}]{\setlength{\rightskip}{0pt plus 5cm}void no\_\-cells\_\-alive ()}}
\label{cowichan_8cpp_19e1551f7e954a984729d96c2c0bf5ba}

Prints no cells alive message and exits. \hypertarget{cowichan_8cpp_501801b9dcd90a910432c7712ff763e7}{
\index{cowichan.cpp@{cowichan.cpp}!not\_\-enough\_\-points@{not\_\-enough\_\-points}}
\index{not\_\-enough\_\-points@{not\_\-enough\_\-points}!cowichan.cpp@{cowichan.cpp}}
\subsubsection[{not\_\-enough\_\-points}]{\setlength{\rightskip}{0pt plus 5cm}void not\_\-enough\_\-points ()}}
\label{cowichan_8cpp_501801b9dcd90a910432c7712ff763e7}

Prints not enough points message and exits. \hypertarget{cowichan_8cpp_faa5a0a08b60255e5a59cbd3b1f5e45f}{
\index{cowichan.cpp@{cowichan.cpp}!out\_\-of\_\-memory@{out\_\-of\_\-memory}}
\index{out\_\-of\_\-memory@{out\_\-of\_\-memory}!cowichan.cpp@{cowichan.cpp}}
\subsubsection[{out\_\-of\_\-memory}]{\setlength{\rightskip}{0pt plus 5cm}void out\_\-of\_\-memory ()}}
\label{cowichan_8cpp_faa5a0a08b60255e5a59cbd3b1f5e45f}

Prints out of memory message and exits. \hypertarget{cowichan_8cpp_560007e135455d06968b8a834d970cf9}{
\index{cowichan.cpp@{cowichan.cpp}!print\_\-elapsed\_\-time@{print\_\-elapsed\_\-time}}
\index{print\_\-elapsed\_\-time@{print\_\-elapsed\_\-time}!cowichan.cpp@{cowichan.cpp}}
\subsubsection[{print\_\-elapsed\_\-time}]{\setlength{\rightskip}{0pt plus 5cm}void print\_\-elapsed\_\-time (INT64 {\em start}, \/  INT64 {\em end})}}
\label{cowichan_8cpp_560007e135455d06968b8a834d970cf9}

Print elapsed time to std::cout. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em start}]initial tick count. \item[{\em end}]final tick count. \end{description}
\end{Desc}
\hypertarget{cowichan_8cpp_1bf3d816eb607b5a06a97e05e9b95099}{
\index{cowichan.cpp@{cowichan.cpp}!timeInfo@{timeInfo}}
\index{timeInfo@{timeInfo}!cowichan.cpp@{cowichan.cpp}}
\subsubsection[{timeInfo}]{\setlength{\rightskip}{0pt plus 5cm}void timeInfo (INT64 $\ast$ {\em start}, \/  INT64 $\ast$ {\em end}, \/  std::string {\em message})}}
\label{cowichan_8cpp_1bf3d816eb607b5a06a97e05e9b95099}

Does a sort of swap-out, printing progress. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em start}]pointer to initial tick count. \item[{\em end}]pointeger to final tick count. \item[{\em message}]message to print before the elapsed time. \end{description}
\end{Desc}
\hypertarget{cowichan_8cpp_eb2386a0fab0a548605cd5f0898da5b4}{
\index{cowichan.cpp@{cowichan.cpp}!uniform@{uniform}}
\index{uniform@{uniform}!cowichan.cpp@{cowichan.cpp}}
\subsubsection[{uniform}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} uniform ({\bf real} {\em mean}, \/  {\bf real} {\em range})}}
\label{cowichan_8cpp_eb2386a0fab0a548605cd5f0898da5b4}

Returns a pseudorandom number $\sim$ U\mbox{[}mean - range, mean + range\mbox{]}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em mean}]mean. \item[{\em range}]range. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]The pseudorandom number. \end{Desc}

### 2009-08-07 doc/latex/cowichan_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan_8hpp}{
\section{cowichan/cowichan.hpp File Reference}
\label{cowichan_8hpp}\index{cowichan/cowichan.hpp@{cowichan/cowichan.hpp}}
}
Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs.

{\tt \#include \char`\"{}cowichan\_\-defaults.hpp\char`\"{}}\par
{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdlib$>$}\par
{\tt \#include $<$ctime$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include $<$algorithm$>$}\par
{\tt \#include $<$iomanip$>$}\par
{\tt \#include $<$climits$>$}\par
{\tt \#include $<$limits$>$}\par
{\tt \#include $<$string$>$}\par
{\tt \#include $<$cstring$>$}\par
{\tt \#include $<$windows.h$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_point}{Point}
\begin{CompactList}\small\item\em Two-dimensional point. \item\end{CompactList}\item 
class \hyperlink{class_weighted_point}{WeightedPoint}
\begin{CompactList}\small\item\em Two-dimensional point with an integer weight. \item\end{CompactList}\item 
class \hyperlink{class_cowichan}{Cowichan}
\begin{CompactList}\small\item\em Base class for all C++ implementations. \item\end{CompactList}\end{CompactItemize}
\subsection*{Defines}
\begin{CompactItemize}
\item 
\#define \hyperlink{cowichan_8hpp_9adcfa1dcb92dd0d8757264ae41d1710}{TEST\_\-TIME}
\item 
\#define \hyperlink{cowichan_8hpp_e0937999ad4f52ae055b86396280604a}{SORT\_\-TIME}
\item 
\#define \hyperlink{cowichan_8hpp_7db5e8ee7c27e6fb5c9caf9d674ea949}{REAL\_\-TYPE}~float
\item 
\#define \hyperlink{cowichan_8hpp_9f8bbf20bc9432260a35b2f3cdb009f8}{MAXIMUM\_\-INT}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE}$>$::max()
\item 
\#define \hyperlink{cowichan_8hpp_af258621cee3ac54ad546ab5297c2476}{MINIMUM\_\-INT}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE}$>$::min()
\item 
\#define \hyperlink{cowichan_8hpp_753345ec3932dd32057b98ab9386efa0}{MAXIMUM\_\-REAL}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real}$>$::min()
\item 
\#define \hyperlink{cowichan_8hpp_e618a02a4bb148a144f9b9f99aa0b581}{MINIMUM\_\-REAL}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real}$>$::min()
\item 
\#define \hyperlink{cowichan_8hpp_bb50b5f693a0b64cdf91bca21cb464de}{INFINITY\_\-REAL}~numeric\_\-limits$<$\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real}$>$::infinity()
\item 
\#define \hyperlink{cowichan_8hpp_e2f68e53d100109588bb54ccc1c3d85b}{MATRIX\_\-RECT}(mtrx, row, col)~(mtrx)\mbox{[}(row)$\ast$this $\rightarrow$ nc + col\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_2747f4d7d371c6ebdcaeb134876a2cf3}{MATRIX\_\-RECT\_\-NC}(mtrx, row, col, nc)~(mtrx)\mbox{[}(row)$\ast$(nc) + col\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_79012ec47cbea66cd6664a54ca8a754c}{MATRIX\_\-SQUARE}(mtrx, row, col)~(mtrx)\mbox{[}(row)$\ast$this $\rightarrow$ n + col\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_6772060ae1a2d5750934131acf95322a}{MATRIX\_\-SQUARE\_\-N}(mtrx, row, col, n)~(mtrx)\mbox{[}(row)$\ast$(n) + col\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_f1d98fb728b5c7300f80dd782702d1dd}{MATRIX}~MATRIX\_\-SQUARE
\item 
\#define \hyperlink{cowichan_8hpp_41a3c329b378fa03397ed8a8925cbc73}{VECTOR}(vect, row)~(vect)\mbox{[}row\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_bacaf90ccfec9927a6d97aa37f39917d}{DIAG}(mtrx, v)~(mtrx)\mbox{[}(v)$\ast$this $\rightarrow$ n + v\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_7b818ef6ccf7d45bd25b07a402ca500f}{DIAG\_\-SZ}(mtrx, v, n)~(mtrx)\mbox{[}(v)$\ast$(n) + v\mbox{]}
\item 
\#define \hyperlink{cowichan_8hpp_07e1668b7974a041604d27b62cbe76f5}{NEW\_\-MATRIX\_\-RECT}(\_\-\_\-type)~(new \_\-\_\-type\mbox{[}this $\rightarrow$ nr $\ast$ this $\rightarrow$ nc\mbox{]})
\item 
\#define \hyperlink{cowichan_8hpp_03c9abf69e2e60f4ed74eebad85b750f}{NEW\_\-MATRIX\_\-SQUARE}(\_\-\_\-type)~(new \_\-\_\-type\mbox{[}this $\rightarrow$ n $\ast$ this $\rightarrow$ n\mbox{]})
\item 
\#define \hyperlink{cowichan_8hpp_e37a7bc1a3de6387853e097f2a8b24b5}{NEW\_\-VECTOR\_\-SZ}(\_\-\_\-type, \_\-\_\-num)~(new \_\-\_\-type\mbox{[}\_\-\_\-num\mbox{]})
\item 
\#define \hyperlink{cowichan_8hpp_71e0e9927a2730d38d0335484f7554fe}{NEW\_\-VECTOR}(\_\-\_\-type)~NEW\_\-VECTOR\_\-SZ(\_\-\_\-type, this $\rightarrow$ n)
\end{CompactItemize}
\subsection*{Typedefs}
\begin{CompactItemize}
\item 
typedef REAL\_\-TYPE \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real}
\item 
typedef UINT32 \hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE}
\item 
typedef \hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} $\ast$ \hyperlink{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{IntMatrix}
\item 
typedef bool $\ast$ \hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix}
\item 
typedef \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} $\ast$ \hyperlink{cowichan_8hpp_7b68f24757f8f24846e7b556a6563030}{RealMatrix}
\item 
typedef \hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} $\ast$ \hyperlink{cowichan_8hpp_9bab229b7d95f858be62c35cca6ff294}{IntVector}
\item 
typedef bool $\ast$ \hyperlink{cowichan_8hpp_69facfadf10e8e6e1b3ea0d02af7fef4}{BoolVector}
\item 
typedef \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} $\ast$ \hyperlink{cowichan_8hpp_b40f3659328a6b06e3f83252e79e7b9a}{RealVector}
\item 
typedef \hyperlink{cowichan_8hpp_7b68f24757f8f24846e7b556a6563030}{RealMatrix} \hyperlink{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{Matrix}
\item 
typedef \hyperlink{cowichan_8hpp_b40f3659328a6b06e3f83252e79e7b9a}{RealVector} \hyperlink{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{Vector}
\item 
typedef ptrdiff\_\-t \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t}
\item 
typedef \hyperlink{class_point}{Point} $\ast$ \hyperlink{cowichan_8hpp_1551f7def82edc60d7e37053954b7a07}{PointVector}
\item 
typedef \hyperlink{class_weighted_point}{WeightedPoint} $\ast$ \hyperlink{cowichan_8hpp_bd6eb1c54f195e0982817565e2ca9c31}{WeightedPointVector}
\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
INT64 \hyperlink{cowichan_8hpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}{get\_\-ticks} ()
\item 
INT64 \hyperlink{cowichan_8hpp_40356e6c8804a8f94e41c156ad368c83}{get\_\-freq} ()
\item 
void \hyperlink{cowichan_8hpp_560007e135455d06968b8a834d970cf9}{print\_\-elapsed\_\-time} (INT64 start, INT64 end)
\item 
void \hyperlink{cowichan_8hpp_1bf3d816eb607b5a06a97e05e9b95099}{timeInfo} (INT64 $\ast$start, INT64 $\ast$end, std::string message)
\item 
void \hyperlink{cowichan_8hpp_faa5a0a08b60255e5a59cbd3b1f5e45f}{out\_\-of\_\-memory} ()
\item 
void \hyperlink{cowichan_8hpp_501801b9dcd90a910432c7712ff763e7}{not\_\-enough\_\-points} ()
\item 
void \hyperlink{cowichan_8hpp_19e1551f7e954a984729d96c2c0bf5ba}{no\_\-cells\_\-alive} ()
\item 
\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} \hyperlink{cowichan_8hpp_eb2386a0fab0a548605cd5f0898da5b4}{uniform} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} mean, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} range)
\end{CompactItemize}

\subsection{Detailed Description}
Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs.

\subsection{Define Documentation}
\hypertarget{cowichan_8hpp_bacaf90ccfec9927a6d97aa37f39917d}{
\index{cowichan.hpp@{cowichan.hpp}!DIAG@{DIAG}}
\index{DIAG@{DIAG}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{DIAG}]{\setlength{\rightskip}{0pt plus 5cm}\#define DIAG(mtrx, \/  v)~(mtrx)\mbox{[}(v)$\ast$this $\rightarrow$ n + v\mbox{]}}}
\label{cowichan_8hpp_bacaf90ccfec9927a6d97aa37f39917d}

Access a square matrix diagonal element v from a class using member variable n as the matrix size. \hypertarget{cowichan_8hpp_7b818ef6ccf7d45bd25b07a402ca500f}{
\index{cowichan.hpp@{cowichan.hpp}!DIAG\_\-SZ@{DIAG\_\-SZ}}
\index{DIAG\_\-SZ@{DIAG\_\-SZ}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{DIAG\_\-SZ}]{\setlength{\rightskip}{0pt plus 5cm}\#define DIAG\_\-SZ(mtrx, \/  v, \/  n)~(mtrx)\mbox{[}(v)$\ast$(n) + v\mbox{]}}}
\label{cowichan_8hpp_7b818ef6ccf7d45bd25b07a402ca500f}

Access a square matrix diagonal element v using n as the matrix size. \hypertarget{cowichan_8hpp_bb50b5f693a0b64cdf91bca21cb464de}{
\index{cowichan.hpp@{cowichan.hpp}!INFINITY\_\-REAL@{INFINITY\_\-REAL}}
\index{INFINITY\_\-REAL@{INFINITY\_\-REAL}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{INFINITY\_\-REAL}]{\setlength{\rightskip}{0pt plus 5cm}\#define INFINITY\_\-REAL~numeric\_\-limits$<${\bf real}$>$::infinity()}}
\label{cowichan_8hpp_bb50b5f693a0b64cdf91bca21cb464de}

Real infinity value. \hypertarget{cowichan_8hpp_f1d98fb728b5c7300f80dd782702d1dd}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX@{MATRIX}}
\index{MATRIX@{MATRIX}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX~MATRIX\_\-SQUARE}}
\label{cowichan_8hpp_f1d98fb728b5c7300f80dd782702d1dd}

Same as MATRIX\_\-SQUARE (shorter name) \hypertarget{cowichan_8hpp_e2f68e53d100109588bb54ccc1c3d85b}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX\_\-RECT@{MATRIX\_\-RECT}}
\index{MATRIX\_\-RECT@{MATRIX\_\-RECT}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX\_\-RECT}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX\_\-RECT(mtrx, \/  row, \/  col)~(mtrx)\mbox{[}(row)$\ast$this $\rightarrow$ nc + col\mbox{]}}}
\label{cowichan_8hpp_e2f68e53d100109588bb54ccc1c3d85b}

Access a rectangular matrix at row/col from a class using member variable nc as the number of columns. \hypertarget{cowichan_8hpp_2747f4d7d371c6ebdcaeb134876a2cf3}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX\_\-RECT\_\-NC@{MATRIX\_\-RECT\_\-NC}}
\index{MATRIX\_\-RECT\_\-NC@{MATRIX\_\-RECT\_\-NC}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX\_\-RECT\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX\_\-RECT\_\-NC(mtrx, \/  row, \/  col, \/  nc)~(mtrx)\mbox{[}(row)$\ast$(nc) + col\mbox{]}}}
\label{cowichan_8hpp_2747f4d7d371c6ebdcaeb134876a2cf3}

Access a rectangular matrix at row/col using nc as the number of columns. \hypertarget{cowichan_8hpp_79012ec47cbea66cd6664a54ca8a754c}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX\_\-SQUARE@{MATRIX\_\-SQUARE}}
\index{MATRIX\_\-SQUARE@{MATRIX\_\-SQUARE}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX\_\-SQUARE}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX\_\-SQUARE(mtrx, \/  row, \/  col)~(mtrx)\mbox{[}(row)$\ast$this $\rightarrow$ n + col\mbox{]}}}
\label{cowichan_8hpp_79012ec47cbea66cd6664a54ca8a754c}

Access a square matrix at row/col from a class using member variable n as the matrix size. \hypertarget{cowichan_8hpp_6772060ae1a2d5750934131acf95322a}{
\index{cowichan.hpp@{cowichan.hpp}!MATRIX\_\-SQUARE\_\-N@{MATRIX\_\-SQUARE\_\-N}}
\index{MATRIX\_\-SQUARE\_\-N@{MATRIX\_\-SQUARE\_\-N}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MATRIX\_\-SQUARE\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define MATRIX\_\-SQUARE\_\-N(mtrx, \/  row, \/  col, \/  n)~(mtrx)\mbox{[}(row)$\ast$(n) + col\mbox{]}}}
\label{cowichan_8hpp_6772060ae1a2d5750934131acf95322a}

Access a square matrix at row/col using n as the matrix size. \hypertarget{cowichan_8hpp_9f8bbf20bc9432260a35b2f3cdb009f8}{
\index{cowichan.hpp@{cowichan.hpp}!MAXIMUM\_\-INT@{MAXIMUM\_\-INT}}
\index{MAXIMUM\_\-INT@{MAXIMUM\_\-INT}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MAXIMUM\_\-INT}]{\setlength{\rightskip}{0pt plus 5cm}\#define MAXIMUM\_\-INT~numeric\_\-limits$<${\bf INT\_\-TYPE}$>$::max()}}
\label{cowichan_8hpp_9f8bbf20bc9432260a35b2f3cdb009f8}

Maximum integer value in matrices/vectors. \hypertarget{cowichan_8hpp_753345ec3932dd32057b98ab9386efa0}{
\index{cowichan.hpp@{cowichan.hpp}!MAXIMUM\_\-REAL@{MAXIMUM\_\-REAL}}
\index{MAXIMUM\_\-REAL@{MAXIMUM\_\-REAL}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MAXIMUM\_\-REAL}]{\setlength{\rightskip}{0pt plus 5cm}\#define MAXIMUM\_\-REAL~numeric\_\-limits$<${\bf real}$>$::min()}}
\label{cowichan_8hpp_753345ec3932dd32057b98ab9386efa0}

Maximum real value in matrices/vectors. \hypertarget{cowichan_8hpp_af258621cee3ac54ad546ab5297c2476}{
\index{cowichan.hpp@{cowichan.hpp}!MINIMUM\_\-INT@{MINIMUM\_\-INT}}
\index{MINIMUM\_\-INT@{MINIMUM\_\-INT}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MINIMUM\_\-INT}]{\setlength{\rightskip}{0pt plus 5cm}\#define MINIMUM\_\-INT~numeric\_\-limits$<${\bf INT\_\-TYPE}$>$::min()}}
\label{cowichan_8hpp_af258621cee3ac54ad546ab5297c2476}

Minimum integer value in matrices/vectors. \hypertarget{cowichan_8hpp_e618a02a4bb148a144f9b9f99aa0b581}{
\index{cowichan.hpp@{cowichan.hpp}!MINIMUM\_\-REAL@{MINIMUM\_\-REAL}}
\index{MINIMUM\_\-REAL@{MINIMUM\_\-REAL}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{MINIMUM\_\-REAL}]{\setlength{\rightskip}{0pt plus 5cm}\#define MINIMUM\_\-REAL~numeric\_\-limits$<${\bf real}$>$::min()}}
\label{cowichan_8hpp_e618a02a4bb148a144f9b9f99aa0b581}

Minimum real value in matrices/vectors. \hypertarget{cowichan_8hpp_07e1668b7974a041604d27b62cbe76f5}{
\index{cowichan.hpp@{cowichan.hpp}!NEW\_\-MATRIX\_\-RECT@{NEW\_\-MATRIX\_\-RECT}}
\index{NEW\_\-MATRIX\_\-RECT@{NEW\_\-MATRIX\_\-RECT}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{NEW\_\-MATRIX\_\-RECT}]{\setlength{\rightskip}{0pt plus 5cm}\#define NEW\_\-MATRIX\_\-RECT(\_\-\_\-type)~(new \_\-\_\-type\mbox{[}this $\rightarrow$ nr $\ast$ this $\rightarrow$ nc\mbox{]})}}
\label{cowichan_8hpp_07e1668b7974a041604d27b62cbe76f5}

Allocate a new rectangular matrix from a class using member variable nr as the number of rows and member variable nc as the number of columns. \hypertarget{cowichan_8hpp_03c9abf69e2e60f4ed74eebad85b750f}{
\index{cowichan.hpp@{cowichan.hpp}!NEW\_\-MATRIX\_\-SQUARE@{NEW\_\-MATRIX\_\-SQUARE}}
\index{NEW\_\-MATRIX\_\-SQUARE@{NEW\_\-MATRIX\_\-SQUARE}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{NEW\_\-MATRIX\_\-SQUARE}]{\setlength{\rightskip}{0pt plus 5cm}\#define NEW\_\-MATRIX\_\-SQUARE(\_\-\_\-type)~(new \_\-\_\-type\mbox{[}this $\rightarrow$ n $\ast$ this $\rightarrow$ n\mbox{]})}}
\label{cowichan_8hpp_03c9abf69e2e60f4ed74eebad85b750f}

Allocate a new square matrix from a class using member variable n as the matrix size. \hypertarget{cowichan_8hpp_71e0e9927a2730d38d0335484f7554fe}{
\index{cowichan.hpp@{cowichan.hpp}!NEW\_\-VECTOR@{NEW\_\-VECTOR}}
\index{NEW\_\-VECTOR@{NEW\_\-VECTOR}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{NEW\_\-VECTOR}]{\setlength{\rightskip}{0pt plus 5cm}\#define NEW\_\-VECTOR(\_\-\_\-type)~NEW\_\-VECTOR\_\-SZ(\_\-\_\-type, this $\rightarrow$ n)}}
\label{cowichan_8hpp_71e0e9927a2730d38d0335484f7554fe}

Allocate a new vector from a class using member variable n as the size of the vector. \hypertarget{cowichan_8hpp_e37a7bc1a3de6387853e097f2a8b24b5}{
\index{cowichan.hpp@{cowichan.hpp}!NEW\_\-VECTOR\_\-SZ@{NEW\_\-VECTOR\_\-SZ}}
\index{NEW\_\-VECTOR\_\-SZ@{NEW\_\-VECTOR\_\-SZ}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{NEW\_\-VECTOR\_\-SZ}]{\setlength{\rightskip}{0pt plus 5cm}\#define NEW\_\-VECTOR\_\-SZ(\_\-\_\-type, \/  \_\-\_\-num)~(new \_\-\_\-type\mbox{[}\_\-\_\-num\mbox{]})}}
\label{cowichan_8hpp_e37a7bc1a3de6387853e097f2a8b24b5}

Allocate a new vector of size \_\-\_\-num. \hypertarget{cowichan_8hpp_7db5e8ee7c27e6fb5c9caf9d674ea949}{
\index{cowichan.hpp@{cowichan.hpp}!REAL\_\-TYPE@{REAL\_\-TYPE}}
\index{REAL\_\-TYPE@{REAL\_\-TYPE}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{REAL\_\-TYPE}]{\setlength{\rightskip}{0pt plus 5cm}\#define REAL\_\-TYPE~float}}
\label{cowichan_8hpp_7db5e8ee7c27e6fb5c9caf9d674ea949}

Use IEEE single floating-point by default. \hypertarget{cowichan_8hpp_e0937999ad4f52ae055b86396280604a}{
\index{cowichan.hpp@{cowichan.hpp}!SORT\_\-TIME@{SORT\_\-TIME}}
\index{SORT\_\-TIME@{SORT\_\-TIME}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{SORT\_\-TIME}]{\setlength{\rightskip}{0pt plus 5cm}\#define SORT\_\-TIME}}
\label{cowichan_8hpp_e0937999ad4f52ae055b86396280604a}

Enables printing of sort time (in winnow). \hypertarget{cowichan_8hpp_9adcfa1dcb92dd0d8757264ae41d1710}{
\index{cowichan.hpp@{cowichan.hpp}!TEST\_\-TIME@{TEST\_\-TIME}}
\index{TEST\_\-TIME@{TEST\_\-TIME}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{TEST\_\-TIME}]{\setlength{\rightskip}{0pt plus 5cm}\#define TEST\_\-TIME}}
\label{cowichan_8hpp_9adcfa1dcb92dd0d8757264ae41d1710}

Enables printing of test time. \hypertarget{cowichan_8hpp_41a3c329b378fa03397ed8a8925cbc73}{
\index{cowichan.hpp@{cowichan.hpp}!VECTOR@{VECTOR}}
\index{VECTOR@{VECTOR}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{VECTOR}]{\setlength{\rightskip}{0pt plus 5cm}\#define VECTOR(vect, \/  row)~(vect)\mbox{[}row\mbox{]}}}
\label{cowichan_8hpp_41a3c329b378fa03397ed8a8925cbc73}

Access row of a vector.

\subsection{Typedef Documentation}
\hypertarget{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{
\index{cowichan.hpp@{cowichan.hpp}!BoolMatrix@{BoolMatrix}}
\index{BoolMatrix@{BoolMatrix}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{BoolMatrix}]{\setlength{\rightskip}{0pt plus 5cm}typedef bool$\ast$ {\bf BoolMatrix}}}
\label{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}

Boolean matrix type. \hypertarget{cowichan_8hpp_69facfadf10e8e6e1b3ea0d02af7fef4}{
\index{cowichan.hpp@{cowichan.hpp}!BoolVector@{BoolVector}}
\index{BoolVector@{BoolVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{BoolVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef bool$\ast$ {\bf BoolVector}}}
\label{cowichan_8hpp_69facfadf10e8e6e1b3ea0d02af7fef4}

Boolean vector type. \hypertarget{cowichan_8hpp_5b04577d5d21124855deaad298595371}{
\index{cowichan.hpp@{cowichan.hpp}!index\_\-t@{index\_\-t}}
\index{index\_\-t@{index\_\-t}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{index\_\-t}]{\setlength{\rightskip}{0pt plus 5cm}typedef ptrdiff\_\-t {\bf index\_\-t}}}
\label{cowichan_8hpp_5b04577d5d21124855deaad298595371}

Index type (signed version of size\_\-t). \hypertarget{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{
\index{cowichan.hpp@{cowichan.hpp}!INT\_\-TYPE@{INT\_\-TYPE}}
\index{INT\_\-TYPE@{INT\_\-TYPE}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{INT\_\-TYPE}]{\setlength{\rightskip}{0pt plus 5cm}typedef UINT32 {\bf INT\_\-TYPE}}}
\label{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}

Integer type (used in matrix/vector values). \hypertarget{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}{
\index{cowichan.hpp@{cowichan.hpp}!IntMatrix@{IntMatrix}}
\index{IntMatrix@{IntMatrix}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{IntMatrix}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf INT\_\-TYPE}$\ast$ {\bf IntMatrix}}}
\label{cowichan_8hpp_82321152ddeeefe9c61350a42ed9e7af}

Integer matrix type. \hypertarget{cowichan_8hpp_9bab229b7d95f858be62c35cca6ff294}{
\index{cowichan.hpp@{cowichan.hpp}!IntVector@{IntVector}}
\index{IntVector@{IntVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{IntVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf INT\_\-TYPE}$\ast$ {\bf IntVector}}}
\label{cowichan_8hpp_9bab229b7d95f858be62c35cca6ff294}

Integer vector type. \hypertarget{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}{
\index{cowichan.hpp@{cowichan.hpp}!Matrix@{Matrix}}
\index{Matrix@{Matrix}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{Matrix}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf RealMatrix} {\bf Matrix}}}
\label{cowichan_8hpp_3fb46f939e55c239fbc95656fc0f3399}

Real matrix type (shorter name). \hypertarget{cowichan_8hpp_1551f7def82edc60d7e37053954b7a07}{
\index{cowichan.hpp@{cowichan.hpp}!PointVector@{PointVector}}
\index{PointVector@{PointVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{PointVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf Point}$\ast$ {\bf PointVector}}}
\label{cowichan_8hpp_1551f7def82edc60d7e37053954b7a07}

Vector of points type. \hypertarget{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{
\index{cowichan.hpp@{cowichan.hpp}!real@{real}}
\index{real@{real}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{real}]{\setlength{\rightskip}{0pt plus 5cm}typedef REAL\_\-TYPE {\bf real}}}
\label{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}

Real type (used in matrix/vector values). \hypertarget{cowichan_8hpp_7b68f24757f8f24846e7b556a6563030}{
\index{cowichan.hpp@{cowichan.hpp}!RealMatrix@{RealMatrix}}
\index{RealMatrix@{RealMatrix}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{RealMatrix}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf real}$\ast$ {\bf RealMatrix}}}
\label{cowichan_8hpp_7b68f24757f8f24846e7b556a6563030}

Real matrix type. \hypertarget{cowichan_8hpp_b40f3659328a6b06e3f83252e79e7b9a}{
\index{cowichan.hpp@{cowichan.hpp}!RealVector@{RealVector}}
\index{RealVector@{RealVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{RealVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf real}$\ast$ {\bf RealVector}}}
\label{cowichan_8hpp_b40f3659328a6b06e3f83252e79e7b9a}

Real vector type. \hypertarget{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}{
\index{cowichan.hpp@{cowichan.hpp}!Vector@{Vector}}
\index{Vector@{Vector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{Vector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf RealVector} {\bf Vector}}}
\label{cowichan_8hpp_02bc1553e241b9b33408482658b3c355}

Real vector type (shorter name). \hypertarget{cowichan_8hpp_bd6eb1c54f195e0982817565e2ca9c31}{
\index{cowichan.hpp@{cowichan.hpp}!WeightedPointVector@{WeightedPointVector}}
\index{WeightedPointVector@{WeightedPointVector}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{WeightedPointVector}]{\setlength{\rightskip}{0pt plus 5cm}typedef {\bf WeightedPoint}$\ast$ {\bf WeightedPointVector}}}
\label{cowichan_8hpp_bd6eb1c54f195e0982817565e2ca9c31}

Vector of weighted points type.

\subsection{Function Documentation}
\hypertarget{cowichan_8hpp_40356e6c8804a8f94e41c156ad368c83}{
\index{cowichan.hpp@{cowichan.hpp}!get\_\-freq@{get\_\-freq}}
\index{get\_\-freq@{get\_\-freq}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{get\_\-freq}]{\setlength{\rightskip}{0pt plus 5cm}INT64 get\_\-freq ()}}
\label{cowichan_8hpp_40356e6c8804a8f94e41c156ad368c83}

Get tick frequency. \begin{Desc}
\item[Returns:]Tick frequency. \end{Desc}
\hypertarget{cowichan_8hpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}{
\index{cowichan.hpp@{cowichan.hpp}!get\_\-ticks@{get\_\-ticks}}
\index{get\_\-ticks@{get\_\-ticks}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{get\_\-ticks}]{\setlength{\rightskip}{0pt plus 5cm}INT64 get\_\-ticks ()}}
\label{cowichan_8hpp_fa824504b2ae4f2f1ebbf5e6f76fb4fa}

Get tick count. \begin{Desc}
\item[Returns:]Tick count. \end{Desc}
\hypertarget{cowichan_8hpp_19e1551f7e954a984729d96c2c0bf5ba}{
\index{cowichan.hpp@{cowichan.hpp}!no\_\-cells\_\-alive@{no\_\-cells\_\-alive}}
\index{no\_\-cells\_\-alive@{no\_\-cells\_\-alive}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{no\_\-cells\_\-alive}]{\setlength{\rightskip}{0pt plus 5cm}void no\_\-cells\_\-alive ()}}
\label{cowichan_8hpp_19e1551f7e954a984729d96c2c0bf5ba}

Prints no cells alive message and exits. \hypertarget{cowichan_8hpp_501801b9dcd90a910432c7712ff763e7}{
\index{cowichan.hpp@{cowichan.hpp}!not\_\-enough\_\-points@{not\_\-enough\_\-points}}
\index{not\_\-enough\_\-points@{not\_\-enough\_\-points}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{not\_\-enough\_\-points}]{\setlength{\rightskip}{0pt plus 5cm}void not\_\-enough\_\-points ()}}
\label{cowichan_8hpp_501801b9dcd90a910432c7712ff763e7}

Prints not enough points message and exits. \hypertarget{cowichan_8hpp_faa5a0a08b60255e5a59cbd3b1f5e45f}{
\index{cowichan.hpp@{cowichan.hpp}!out\_\-of\_\-memory@{out\_\-of\_\-memory}}
\index{out\_\-of\_\-memory@{out\_\-of\_\-memory}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{out\_\-of\_\-memory}]{\setlength{\rightskip}{0pt plus 5cm}void out\_\-of\_\-memory ()}}
\label{cowichan_8hpp_faa5a0a08b60255e5a59cbd3b1f5e45f}

Prints out of memory message and exits. \hypertarget{cowichan_8hpp_560007e135455d06968b8a834d970cf9}{
\index{cowichan.hpp@{cowichan.hpp}!print\_\-elapsed\_\-time@{print\_\-elapsed\_\-time}}
\index{print\_\-elapsed\_\-time@{print\_\-elapsed\_\-time}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{print\_\-elapsed\_\-time}]{\setlength{\rightskip}{0pt plus 5cm}void print\_\-elapsed\_\-time (INT64 {\em start}, \/  INT64 {\em end})}}
\label{cowichan_8hpp_560007e135455d06968b8a834d970cf9}

Print elapsed time to std::cout. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em start}]initial tick count. \item[{\em end}]final tick count. \end{description}
\end{Desc}
\hypertarget{cowichan_8hpp_1bf3d816eb607b5a06a97e05e9b95099}{
\index{cowichan.hpp@{cowichan.hpp}!timeInfo@{timeInfo}}
\index{timeInfo@{timeInfo}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{timeInfo}]{\setlength{\rightskip}{0pt plus 5cm}void timeInfo (INT64 $\ast$ {\em start}, \/  INT64 $\ast$ {\em end}, \/  std::string {\em message})}}
\label{cowichan_8hpp_1bf3d816eb607b5a06a97e05e9b95099}

Does a sort of swap-out, printing progress. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em start}]pointer to initial tick count. \item[{\em end}]pointeger to final tick count. \item[{\em message}]message to print before the elapsed time. \end{description}
\end{Desc}
\hypertarget{cowichan_8hpp_eb2386a0fab0a548605cd5f0898da5b4}{
\index{cowichan.hpp@{cowichan.hpp}!uniform@{uniform}}
\index{uniform@{uniform}!cowichan.hpp@{cowichan.hpp}}
\subsubsection[{uniform}]{\setlength{\rightskip}{0pt plus 5cm}{\bf real} uniform ({\bf real} {\em mean}, \/  {\bf real} {\em range})}}
\label{cowichan_8hpp_eb2386a0fab0a548605cd5f0898da5b4}

Returns a pseudorandom number $\sim$ U\mbox{[}mean - range, mean + range\mbox{]}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em mean}]mean. \item[{\em range}]range. \end{description}
\end{Desc}
\begin{Desc}
\item[Returns:]The pseudorandom number. \end{Desc}

### 2009-08-07 doc/latex/cowichan__defaults_8hpp.tex
Re-generated doxygen with linuxtuples additions.

@@
 \item 
-\#define \hyperlink{cowichan__defaults_8hpp_30524c594b6941ca7822489d1794813b}{ALL\_\-NR}~15000
+\#define \hyperlink{cowichan__defaults_8hpp_30524c594b6941ca7822489d1794813b}{ALL\_\-NR}~1000
 \item 
-\#define \hyperlink{cowichan__defaults_8hpp_85076a3092aabd0bc4661e9184e2ddf3}{ALL\_\-NC}~15000
+\#define \hyperlink{cowichan__defaults_8hpp_85076a3092aabd0bc4661e9184e2ddf3}{ALL\_\-NC}~1000
 \item 
-\#define \hyperlink{cowichan__defaults_8hpp_e36abbd26bf4ea979de0b63b1a37e349}{ALL\_\-N}~15000
+\#define \hyperlink{cowichan__defaults_8hpp_e36abbd26bf4ea979de0b63b1a37e349}{ALL\_\-N}~1000
 \item 
@@
 \item 
-\#define \hyperlink{cowichan__defaults_8hpp_02f3c90de37aa6314330a7e0cb68a1f5}{LIFE\_\-ITERATIONS}~10
+\#define \hyperlink{cowichan__defaults_8hpp_02f3c90de37aa6314330a7e0cb68a1f5}{LIFE\_\-ITERATIONS}~3
 \item 
@@
 
-\label{_details}
-\hypertarget{_details}{}
 \subsection{Detailed Description}
@@
 \index{ALL\_\-N@{ALL\_\-N}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
-\subsubsection[{ALL\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-N~15000}}
+\subsubsection[{ALL\_\-N}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-N~1000}}
 \label{cowichan__defaults_8hpp_e36abbd26bf4ea979de0b63b1a37e349}
@@
 \index{ALL\_\-NC@{ALL\_\-NC}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
-\subsubsection[{ALL\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-NC~15000}}
+\subsubsection[{ALL\_\-NC}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-NC~1000}}
 \label{cowichan__defaults_8hpp_85076a3092aabd0bc4661e9184e2ddf3}
@@
 \index{ALL\_\-NR@{ALL\_\-NR}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
-\subsubsection[{ALL\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-NR~15000}}
+\subsubsection[{ALL\_\-NR}]{\setlength{\rightskip}{0pt plus 5cm}\#define ALL\_\-NR~1000}}
 \label{cowichan__defaults_8hpp_30524c594b6941ca7822489d1794813b}
@@
 \index{LIFE\_\-ITERATIONS@{LIFE\_\-ITERATIONS}!cowichan_defaults.hpp@{cowichan\_\-defaults.hpp}}
-\subsubsection[{LIFE\_\-ITERATIONS}]{\setlength{\rightskip}{0pt plus 5cm}\#define LIFE\_\-ITERATIONS~10}}
+\subsubsection[{LIFE\_\-ITERATIONS}]{\setlength{\rightskip}{0pt plus 5cm}\#define LIFE\_\-ITERATIONS~3}}
 \label{cowichan__defaults_8hpp_02f3c90de37aa6314330a7e0cb68a1f5}

### 2009-08-07 doc/latex/cowichan__lt_2src_2gauss_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2gauss_8cpp}{
\section{cowichan\_\-lt/src/gauss.cpp File Reference}
\label{cowichan__lt_2src_2gauss_8cpp}\index{cowichan\_\-lt/src/gauss.cpp@{cowichan\_\-lt/src/gauss.cpp}}
}
LinuxTuples gauss implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}gauss.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples gauss implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_fd3e659afb38415f56fd508333b7bb1b}{CowichanLinuxTuples::gauss} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2half_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2half_8cpp}{
\section{cowichan\_\-lt/src/half.cpp File Reference}
\label{cowichan__lt_2src_2half_8cpp}\index{cowichan\_\-lt/src/half.cpp@{cowichan\_\-lt/src/half.cpp}}
}
LinuxTuples half implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include \char`\"{}half.hpp\char`\"{}}\par
\subsection*{Defines}
\begin{CompactItemize}
\item 
\hypertarget{cowichan__lt_2src_2half_8cpp_44133fb2267b03d2193d855394486d0f}{
\#define \textbf{FLIP}(item)~(item \% 2 == 0 ? (item + 1) : (item - 1))}
\label{cowichan__lt_2src_2half_8cpp_44133fb2267b03d2193d855394486d0f}

\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples half implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_f3bd58ad2653f06d242d473e7bffda0b}{CowichanLinuxTuples::half} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2hull_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2hull_8cpp}{
\section{cowichan\_\-lt/src/hull.cpp File Reference}
\label{cowichan__lt_2src_2hull_8cpp}\index{cowichan\_\-lt/src/hull.cpp@{cowichan\_\-lt/src/hull.cpp}}
}
LinuxTuples hull implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}hull.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples hull implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_501f87594d62af261ff0b1954a60ddc4}{CowichanLinuxTuples::hull} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2invperc_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2invperc_8cpp}{
\section{cowichan\_\-lt/src/invperc.cpp File Reference}
\label{cowichan__lt_2src_2invperc_8cpp}\index{cowichan\_\-lt/src/invperc.cpp@{cowichan\_\-lt/src/invperc.cpp}}
}
LinuxTuples (actually serial) invperc implementation.

{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
{\tt \#include $<$queue$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_perc_point}{PercPoint}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples (actually serial) invperc implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_5497a17a63cc5faad7afa54d013ea741}{CowichanLinuxTuples::invperc} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2life_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2life_8cpp}{
\section{cowichan\_\-lt/src/life.cpp File Reference}
\label{cowichan__lt_2src_2life_8cpp}\index{cowichan\_\-lt/src/life.cpp@{cowichan\_\-lt/src/life.cpp}}
}
LinuxTuples life implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include \char`\"{}life.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples life implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_967138bbdead76eec50f31465dd9d5f1}{CowichanLinuxTuples::life} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2mandel_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2mandel_8cpp}{
\section{cowichan\_\-lt/src/mandel.cpp File Reference}
\label{cowichan__lt_2src_2mandel_8cpp}\index{cowichan\_\-lt/src/mandel.cpp@{cowichan\_\-lt/src/mandel.cpp}}
}
LinuxTuples mandelbrot set implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include \char`\"{}mandel.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples mandelbrot set implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_8774404842868381379ec627e8373cec}{CowichanLinuxTuples::mandel} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2norm_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2norm_8cpp}{
\section{cowichan\_\-lt/src/norm.cpp File Reference}
\label{cowichan__lt_2src_2norm_8cpp}\index{cowichan\_\-lt/src/norm.cpp@{cowichan\_\-lt/src/norm.cpp}}
}
LinuxTuples norm implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}norm.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples norm implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_922209c219bb5b9268263dae4000ea81}{CowichanLinuxTuples::norm} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2outer_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2outer_8cpp}{
\section{cowichan\_\-lt/src/outer.cpp File Reference}
\label{cowichan__lt_2src_2outer_8cpp}\index{cowichan\_\-lt/src/outer.cpp@{cowichan\_\-lt/src/outer.cpp}}
}
LinuxTuples outer implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}outer.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples outer implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_b16e1c3b60f920c12460a74289558dcb}{CowichanLinuxTuples::outer} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2product_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2product_8cpp}{
\section{cowichan\_\-lt/src/product.cpp File Reference}
\label{cowichan__lt_2src_2product_8cpp}\index{cowichan\_\-lt/src/product.cpp@{cowichan\_\-lt/src/product.cpp}}
}
LinuxTuples product implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include \char`\"{}product.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples product implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_c4874f09c54056ac94d00f9e341a94f4}{CowichanLinuxTuples::product} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2randmat_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2randmat_8cpp}{
\section{cowichan\_\-lt/src/randmat.cpp File Reference}
\label{cowichan__lt_2src_2randmat_8cpp}\index{cowichan\_\-lt/src/randmat.cpp@{cowichan\_\-lt/src/randmat.cpp}}
}
LinuxTuples randmat implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include \char`\"{}randmat.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples randmat implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_ff90c6a27db51366570b70d352de7fab}{CowichanLinuxTuples::randmat} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2sor_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2sor_8cpp}{
\section{cowichan\_\-lt/src/sor.cpp File Reference}
\label{cowichan__lt_2src_2sor_8cpp}\index{cowichan\_\-lt/src/sor.cpp@{cowichan\_\-lt/src/sor.cpp}}
}
LinuxTuples successive over-relaxation implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}sor.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples successive over-relaxation implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_aa2c469c3f520a5a56c5ec1ac9a5a6c6}{CowichanLinuxTuples::sor} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2thresh_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2thresh_8cpp}{
\section{cowichan\_\-lt/src/thresh.cpp File Reference}
\label{cowichan__lt_2src_2thresh_8cpp}\index{cowichan\_\-lt/src/thresh.cpp@{cowichan\_\-lt/src/thresh.cpp}}
}
LinuxTuples thresh implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$map$>$}\par
{\tt \#include \char`\"{}thresh.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples thresh implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_b8cf055ec3d42a1ccde6d818c5a43a74}{CowichanLinuxTuples::thresh} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2vecdiff_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2vecdiff_8cpp}{
\section{cowichan\_\-lt/src/vecdiff.cpp File Reference}
\label{cowichan__lt_2src_2vecdiff_8cpp}\index{cowichan\_\-lt/src/vecdiff.cpp@{cowichan\_\-lt/src/vecdiff.cpp}}
}
LinuxTuples vecdiff implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}vecdiff.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples vecdiff implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_213185666ea7dd0c9abc59899b454086}{CowichanLinuxTuples::vecdiff} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_2src_2winnow_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_2src_2winnow_8cpp}{
\section{cowichan\_\-lt/src/winnow.cpp File Reference}
\label{cowichan__lt_2src_2winnow_8cpp}\index{cowichan\_\-lt/src/winnow.cpp@{cowichan\_\-lt/src/winnow.cpp}}
}
LinuxTuples winnow implementation.

{\tt \#include $<$iostream$>$}\par
{\tt \#include $<$cstdio$>$}\par
{\tt \#include $<$cmath$>$}\par
{\tt \#include \char`\"{}winnow.hpp\char`\"{}}\par

\subsection{Detailed Description}
LinuxTuples winnow implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_2b1a77ecbd01ec847cf63fe7ebb50e35}{CowichanLinuxTuples::winnow} \end{Desc}

### 2009-08-07 doc/latex/cowichan__lt_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_8cpp}{
\section{cowichan\_\-lt/src/cowichan\_\-lt.cpp File Reference}
\label{cowichan__lt_8cpp}\index{cowichan\_\-lt/src/cowichan\_\-lt.cpp@{cowichan\_\-lt/src/cowichan\_\-lt.cpp}}
}
Entry point for the cowichan linuxtuples application.

{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Functions}
\begin{CompactItemize}
\item 
int \hyperlink{cowichan__lt_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
\end{CompactItemize}

\subsection{Detailed Description}
Entry point for the cowichan linuxtuples application.

\subsection{Function Documentation}
\hypertarget{cowichan__lt_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
\index{cowichan\_\-lt.cpp@{cowichan\_\-lt.cpp}!main@{main}}
\index{main@{main}!cowichan_lt.cpp@{cowichan\_\-lt.cpp}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
\label{cowichan__lt_8cpp_0ddf1224851353fc92bfbff6f499fa97}

Entry point of the program.

### 2009-08-07 doc/latex/cowichan__lt_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__lt_8hpp}{
\section{cowichan\_\-lt/src/cowichan\_\-lt.hpp File Reference}
\label{cowichan__lt_8hpp}\index{cowichan\_\-lt/src/cowichan\_\-lt.hpp@{cowichan\_\-lt/src/cowichan\_\-lt.hpp}}
}
Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs; LinuxTuples implementation.

{\tt \#include \char`\"{}../../cowichan/cowichan.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_cowichan_linux_tuples}{CowichanLinuxTuples}
\end{CompactItemize}

\subsection{Detailed Description}
Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs; LinuxTuples implementation.

### 2009-08-07 doc/latex/cowichan__mpi_2sort_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__mpi_2sort_8cpp}{
\section{cowichan\_\-mpi/sort.cpp File Reference}
\label{cowichan__mpi_2sort_8cpp}\index{cowichan\_\-mpi/sort.cpp@{cowichan\_\-mpi/sort.cpp}}
}
Implementation of MPI sorting algorithms.

{\tt \#include \char`\"{}sort.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__mpi}{cowichan\_\-mpi}
\begin{CompactList}\small\item\em Additional classes and functions specific to mpi implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_7f82c459ad086060228b2bd8100bf9c7}{cowichan\_\-mpi::getNumRemainingNodes} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_a7c14e0bec9c9da1d56a348c9a0cf1f8}{cowichan\_\-mpi::power2} (\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n)
\item 
void \hyperlink{namespacecowichan__mpi_4ab537d0df99e9011b206c7259e82d35}{cowichan\_\-mpi::quick\_\-sort} (const mpi::communicator \&world, \hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} level, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} owner)
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_83904b95f448398ff15331182ee0c007}{cowichan\_\-mpi::quick\_\-sort\_\-partition} (\hyperlink{class_weighted_point}{WeightedPointVector} vector, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} len, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} pivotIndex)
\end{CompactItemize}

\subsection{Detailed Description}
Implementation of MPI sorting algorithms.

### 2009-08-07 doc/latex/cowichan__mpi_2sort_8hpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__mpi_2sort_8cpp.tex
@@
-\hypertarget{cowichan__mpi_2sort_8cpp}{
-\section{cowichan\_\-mpi/sort.cpp File Reference}
-\label{cowichan__mpi_2sort_8cpp}\index{cowichan\_\-mpi/sort.cpp@{cowichan\_\-mpi/sort.cpp}}
+\hypertarget{cowichan__mpi_2sort_8hpp}{
+\section{cowichan\_\-mpi/sort.hpp File Reference}
+\label{cowichan__mpi_2sort_8hpp}\index{cowichan\_\-mpi/sort.hpp@{cowichan\_\-mpi/sort.hpp}}
 }
-Implementation of MPI sorting algorithms.
+Sorting algorithms that use MPI.
 
-{\tt \#include \char`\"{}sort.hpp\char`\"{}}\par
+{\tt \#include \char`\"{}cowichan\_\-mpi.hpp\char`\"{}}\par
 \subsection*{Namespaces}
@@
 
+\end{CompactItemize}
+\subsection*{Defines}
+\begin{CompactItemize}
+\item 
+\#define \hyperlink{cowichan__mpi_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{QUICK\_\-SORT\_\-CUTOFF}~100
 \end{CompactItemize}
@@
 \subsection{Detailed Description}
-Implementation of MPI sorting algorithms.
+Sorting algorithms that use MPI.
+
+One algorithms is included: \begin{enumerate}
+\item Parallel Quick Sort \end{enumerate}
+
+\subsection{Define Documentation}
+\hypertarget{cowichan__mpi_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{
+\index{cowichan\_\-mpi/sort.hpp@{cowichan\_\-mpi/sort.hpp}!QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}}
+\index{QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}!cowichan_mpi/sort.hpp@{cowichan\_\-mpi/sort.hpp}}
+\subsubsection[{QUICK\_\-SORT\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}\#define QUICK\_\-SORT\_\-CUTOFF~100}}
+\label{cowichan__mpi_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}
+
+Quick sort cutoff (becomes insertion sort) - must be $>$ 1

### 2009-08-07 doc/latex/cowichan__mpi_2winnow_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__mpi_2winnow_8cpp}{
\section{cowichan\_\-mpi/winnow.cpp File Reference}
\label{cowichan__mpi_2winnow_8cpp}\index{cowichan\_\-mpi/winnow.cpp@{cowichan\_\-mpi/winnow.cpp}}
}
MPI winnow implementation.

{\tt \#include \char`\"{}cowichan\_\-mpi.hpp\char`\"{}}\par
{\tt \#include \char`\"{}sort.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__mpi}{cowichan\_\-mpi}
\begin{CompactList}\small\item\em Additional classes and functions specific to mpi implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__mpi_a54004a105356573d4bac92e1b234aa2}{cowichan\_\-mpi::mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\end{CompactItemize}

\subsection{Detailed Description}
MPI winnow implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_m_p_i_9be48d86fc51ce13ef79d5fe2c8a16e0}{CowichanMPI::winnow} \end{Desc}

### 2009-08-07 doc/latex/cowichan__mpi_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__mpi_8cpp}{
\section{cowichan\_\-mpi/cowichan\_\-mpi.cpp File Reference}
\label{cowichan__mpi_8cpp}\index{cowichan\_\-mpi/cowichan\_\-mpi.cpp@{cowichan\_\-mpi/cowichan\_\-mpi.cpp}}
}
This file contains main method that drives the MPI implementation.

{\tt \#include \char`\"{}cowichan\_\-mpi.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__mpi}{cowichan\_\-mpi}
\begin{CompactList}\small\item\em Additional classes and functions specific to mpi implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
int \hyperlink{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
\item 
bool \hyperlink{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}{cowichan\_\-mpi::get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end)
\item 
bool \hyperlink{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}{cowichan\_\-mpi::get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} rank)
\end{CompactItemize}

\subsection{Detailed Description}
This file contains main method that drives the MPI implementation.

\subsection{Function Documentation}
\hypertarget{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
\index{cowichan\_\-mpi.cpp@{cowichan\_\-mpi.cpp}!main@{main}}
\index{main@{main}!cowichan_mpi.cpp@{cowichan\_\-mpi.cpp}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
\label{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}

Main method - creates a \hyperlink{class_cowichan_m_p_i}{CowichanMPI} instance and executes \hyperlink{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{Cowichan::main}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em argc}]number of command line arguments. \item[{\em argv}]command line arguments. \end{description}
\end{Desc}

### 2009-08-07 doc/latex/cowichan__mpi_8hpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__mpi_8cpp.tex
@@
-\hypertarget{cowichan__mpi_8cpp}{
-\section{cowichan\_\-mpi/cowichan\_\-mpi.cpp File Reference}
-\label{cowichan__mpi_8cpp}\index{cowichan\_\-mpi/cowichan\_\-mpi.cpp@{cowichan\_\-mpi/cowichan\_\-mpi.cpp}}
+\hypertarget{cowichan__mpi_8hpp}{
+\section{cowichan\_\-mpi/cowichan\_\-mpi.hpp File Reference}
+\label{cowichan__mpi_8hpp}\index{cowichan\_\-mpi/cowichan\_\-mpi.hpp@{cowichan\_\-mpi/cowichan\_\-mpi.hpp}}
 }
-This file contains main method that drives the MPI implementation.
+Boost Message Passing Interface (MPI) implementation of \hyperlink{class_cowichan}{Cowichan} problems.
 
-{\tt \#include \char`\"{}cowichan\_\-mpi.hpp\char`\"{}}\par
+{\tt \#include \char`\"{}../cowichan/cowichan.hpp\char`\"{}}\par
+{\tt \#include $<$boost/mpi.hpp$>$}\par
+\subsection*{Classes}
+\begin{CompactItemize}
+\item 
+class \hyperlink{class_cowichan_m_p_i}{CowichanMPI}
+\begin{CompactList}\small\item\em Boost Message Passing Interface (MPI) implementation. \item\end{CompactList}\end{CompactItemize}
 \subsection*{Namespaces}
@@
 \item 
-int \hyperlink{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
-\item 
 bool \hyperlink{namespacecowichan__mpi_bb7c07d8b23660efe06c7ee12d60bbdc}{cowichan\_\-mpi::get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end)
@@
 bool \hyperlink{namespacecowichan__mpi_413bf41c8bf73b560d1c16a8e5206e5f}{cowichan\_\-mpi::get\_\-block} (const mpi::communicator \&world, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} lo, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} hi, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$start, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$end, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} rank)
+\item 
+{\footnotesize template$<$class Archive $>$ }\\void \hyperlink{namespaceboost_1_1serialization_214544bc307816738f05752310ed7336}{boost::serialization::serialize} (Archive \&ar, \hyperlink{class_weighted_point}{WeightedPoint} \&p, const unsigned int version)
+\item 
+{\footnotesize template$<$class Archive $>$ }\\void \hyperlink{namespaceboost_1_1serialization_899888015920ed0de50a4ec4786a4d5d}{boost::serialization::serialize} (Archive \&ar, \hyperlink{class_point}{Point} \&p, const unsigned int version)
 \end{CompactItemize}
@@
 \subsection{Detailed Description}
-This file contains main method that drives the MPI implementation.
-
-\subsection{Function Documentation}
-\hypertarget{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
-\index{cowichan\_\-mpi.cpp@{cowichan\_\-mpi.cpp}!main@{main}}
-\index{main@{main}!cowichan_mpi.cpp@{cowichan\_\-mpi.cpp}}
-\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
-\label{cowichan__mpi_8cpp_0ddf1224851353fc92bfbff6f499fa97}
-
-Main method - creates a \hyperlink{class_cowichan_m_p_i}{CowichanMPI} instance and executes \hyperlink{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{Cowichan::main}. \begin{Desc}
-\item[Parameters:]
-\begin{description}
-\item[{\em argc}]number of command line arguments. \item[{\em argv}]command line arguments. \end{description}
-\end{Desc}
+Boost Message Passing Interface (MPI) implementation of \hyperlink{class_cowichan}{Cowichan} problems.

### 2009-08-07 doc/latex/cowichan__openmp_2hull_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2hull_8cpp}{
\section{cowichan\_\-openmp/hull.cpp File Reference}
\label{cowichan__openmp_2hull_8cpp}\index{cowichan\_\-openmp/hull.cpp@{cowichan\_\-openmp/hull.cpp}}
}
OpenMP hull implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__openmp_f261011870b201d556308f43b5b25b89}{cowichan\_\-openmp::quickhull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn)
\item 
void \hyperlink{namespacecowichan__openmp_d31787f1b4c9d57119160411e96f769f}{cowichan\_\-openmp::split} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn, \hyperlink{class_point}{Point} $\ast$p1, \hyperlink{class_point}{Point} $\ast$p2)
\end{CompactItemize}

\subsection{Detailed Description}
OpenMP hull implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_cb444dd3e2c0f1f27f4135de0279d09e}{CowichanOpenMP::hull} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_2invperc_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2invperc_8cpp}{
\section{cowichan\_\-openmp/invperc.cpp File Reference}
\label{cowichan__openmp_2invperc_8cpp}\index{cowichan\_\-openmp/invperc.cpp@{cowichan\_\-openmp/invperc.cpp}}
}
OpenMP (NOT parallel) invperc implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
{\tt \#include $<$queue$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__openmp_1_1_perc_point}{cowichan\_\-openmp::PercPoint}
\begin{CompactList}\small\item\em Percolation point. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
OpenMP (NOT parallel) invperc implementation.

The same as \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial::invperc}

### 2009-08-07 doc/latex/cowichan__openmp_2life_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2life_8cpp}{
\section{cowichan\_\-openmp/life.cpp File Reference}
\label{cowichan__openmp_2life_8cpp}\index{cowichan\_\-openmp/life.cpp@{cowichan\_\-openmp/life.cpp}}
}
OpenMP life implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__openmp_a4ce5b0ba78cd9480bf45e7c313f59d1}{cowichan\_\-openmp::sumNeighbours} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} first, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} r, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} c, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\end{CompactItemize}

\subsection{Detailed Description}
OpenMP life implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_d24f3ef01289b8f1bd7b864f585daa62}{CowichanOpenMP::life} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_2mandel_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2mandel_8cpp}{
\section{cowichan\_\-openmp/mandel.cpp File Reference}
\label{cowichan__openmp_2mandel_8cpp}\index{cowichan\_\-openmp/mandel.cpp@{cowichan\_\-openmp/mandel.cpp}}
}
OpenMP mandel implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{namespacecowichan__openmp_2753e3856bc41c61099c4702d789823e}{cowichan\_\-openmp::mandel\_\-calc} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} x, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} y)
\end{CompactItemize}

\subsection{Detailed Description}
OpenMP mandel implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_6809c2738792de047ee59259636c1afd}{CowichanOpenMP::mandel} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_2norm_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2norm_8cpp}{
\section{cowichan\_\-openmp/norm.cpp File Reference}
\label{cowichan__openmp_2norm_8cpp}\index{cowichan\_\-openmp/norm.cpp@{cowichan\_\-openmp/norm.cpp}}
}
OpenMP norm implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__openmp_e96598e37e2b937c861c5bc93eda77b8}{cowichan\_\-openmp::findMinMax} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{Point} $\ast$minPoint, \hyperlink{class_point}{Point} $\ast$maxPoint)
\end{CompactItemize}

\subsection{Detailed Description}
OpenMP norm implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_4ffbe36816235bc6abec30eae2be2d78}{CowichanOpenMP::norm} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_2product_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2product_8cpp}{
\section{cowichan\_\-openmp/product.cpp File Reference}
\label{cowichan__openmp_2product_8cpp}\index{cowichan\_\-openmp/product.cpp@{cowichan\_\-openmp/product.cpp}}
}
OpenMP product implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par

\subsection{Detailed Description}
OpenMP product implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_41d0067382570d1e784f62f2c5963d49}{CowichanOpenMP::product} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_2randmat_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2randmat_8cpp}{
\section{cowichan\_\-openmp/randmat.cpp File Reference}
\label{cowichan__openmp_2randmat_8cpp}\index{cowichan\_\-openmp/randmat.cpp@{cowichan\_\-openmp/randmat.cpp}}
}
OpenMP randmat implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par

\subsection{Detailed Description}
OpenMP randmat implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_2c7c4e4dd96f82b7280a412c1fceed2c}{CowichanOpenMP::randmat} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_2sort_8hpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-02 doc/latex/sort_8hpp.tex
@@
-\hypertarget{sort_8hpp}{
+\hypertarget{cowichan__openmp_2sort_8hpp}{
 \section{cowichan\_\-openmp/sort.hpp File Reference}
-\label{sort_8hpp}\index{cowichan\_\-openmp/sort.hpp@{cowichan\_\-openmp/sort.hpp}}
+\label{cowichan__openmp_2sort_8hpp}\index{cowichan\_\-openmp/sort.hpp@{cowichan\_\-openmp/sort.hpp}}
 }
@@
 \item 
-\#define \hyperlink{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{QUICK\_\-SORT\_\-CUTOFF}~100
+\#define \hyperlink{cowichan__openmp_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{QUICK\_\-SORT\_\-CUTOFF}~100
 \item 
-\#define \hyperlink{sort_8hpp_66832e055934d011da279261bcfafdd1}{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}~1000
+\#define \hyperlink{cowichan__openmp_2sort_8hpp_66832e055934d011da279261bcfafdd1}{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}~1000
 \end{CompactItemize}
@@
 \subsection{Define Documentation}
-\hypertarget{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{
-\index{sort.hpp@{sort.hpp}!QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}}
-\index{QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}!sort.hpp@{sort.hpp}}
+\hypertarget{cowichan__openmp_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}{
+\index{cowichan\_\-openmp/sort.hpp@{cowichan\_\-openmp/sort.hpp}!QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}}
+\index{QUICK\_\-SORT\_\-CUTOFF@{QUICK\_\-SORT\_\-CUTOFF}!cowichan_openmp/sort.hpp@{cowichan\_\-openmp/sort.hpp}}
 \subsubsection[{QUICK\_\-SORT\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}\#define QUICK\_\-SORT\_\-CUTOFF~100}}
-\label{sort_8hpp_b3aee1394b5864d458c40b7878acbe83}
+\label{cowichan__openmp_2sort_8hpp_b3aee1394b5864d458c40b7878acbe83}
 
-Quick sort cutoff (becomes insertion sort) - must be $>$ 1 \hypertarget{sort_8hpp_66832e055934d011da279261bcfafdd1}{
-\index{sort.hpp@{sort.hpp}!QUICK\_\-SORT\_\-TASK\_\-CUTOFF@{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}}
-\index{QUICK\_\-SORT\_\-TASK\_\-CUTOFF@{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}!sort.hpp@{sort.hpp}}
+Quick sort cutoff (becomes insertion sort) - must be $>$ 1 \hypertarget{cowichan__openmp_2sort_8hpp_66832e055934d011da279261bcfafdd1}{
+\index{cowichan\_\-openmp/sort.hpp@{cowichan\_\-openmp/sort.hpp}!QUICK\_\-SORT\_\-TASK\_\-CUTOFF@{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}}
+\index{QUICK\_\-SORT\_\-TASK\_\-CUTOFF@{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}!cowichan_openmp/sort.hpp@{cowichan\_\-openmp/sort.hpp}}
 \subsubsection[{QUICK\_\-SORT\_\-TASK\_\-CUTOFF}]{\setlength{\rightskip}{0pt plus 5cm}\#define QUICK\_\-SORT\_\-TASK\_\-CUTOFF~1000}}
-\label{sort_8hpp_66832e055934d011da279261bcfafdd1}
+\label{cowichan__openmp_2sort_8hpp_66832e055934d011da279261bcfafdd1}
 

### 2009-08-07 doc/latex/cowichan__openmp_2thresh_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2thresh_8cpp}{
\section{cowichan\_\-openmp/thresh.cpp File Reference}
\label{cowichan__openmp_2thresh_8cpp}\index{cowichan\_\-openmp/thresh.cpp@{cowichan\_\-openmp/thresh.cpp}}
}
OpenMP thresh implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par

\subsection{Detailed Description}
OpenMP thresh implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_e72c4c0a162f30eac37333bd28db97bc}{CowichanOpenMP::thresh} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_2vecdiff_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2vecdiff_8cpp}{
\section{cowichan\_\-openmp/vecdiff.cpp File Reference}
\label{cowichan__openmp_2vecdiff_8cpp}\index{cowichan\_\-openmp/vecdiff.cpp@{cowichan\_\-openmp/vecdiff.cpp}}
}
OpenMP vecdiff implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par

\subsection{Detailed Description}
OpenMP vecdiff implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_92aa23ed47da0a5a3b43416ab08199b3}{CowichanOpenMP::vecdiff} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_2winnow_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_2winnow_8cpp}{
\section{cowichan\_\-openmp/winnow.cpp File Reference}
\label{cowichan__openmp_2winnow_8cpp}\index{cowichan\_\-openmp/winnow.cpp@{cowichan\_\-openmp/winnow.cpp}}
}
OpenMP winnow implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
{\tt \#include \char`\"{}sort.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__openmp_9c68e88badede2262dd4ded923e804dc}{cowichan\_\-openmp::mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$buckets)
\end{CompactItemize}

\subsection{Detailed Description}
OpenMP winnow implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_open_m_p_4a518f2b5590d4acd670f333471a380a}{CowichanOpenMP::winnow} \end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_8cpp}{
\section{cowichan\_\-openmp/cowichan\_\-openmp.cpp File Reference}
\label{cowichan__openmp_8cpp}\index{cowichan\_\-openmp/cowichan\_\-openmp.cpp@{cowichan\_\-openmp/cowichan\_\-openmp.cpp}}
}
This file contains main method that drives the OpenMP implementation.

{\tt \#include \char`\"{}cowichan\_\-openmp.hpp\char`\"{}}\par
\subsection*{Functions}
\begin{CompactItemize}
\item 
int \hyperlink{cowichan__openmp_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
\end{CompactItemize}

\subsection{Detailed Description}
This file contains main method that drives the OpenMP implementation.

\subsection{Function Documentation}
\hypertarget{cowichan__openmp_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
\index{cowichan\_\-openmp.cpp@{cowichan\_\-openmp.cpp}!main@{main}}
\index{main@{main}!cowichan_openmp.cpp@{cowichan\_\-openmp.cpp}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
\label{cowichan__openmp_8cpp_0ddf1224851353fc92bfbff6f499fa97}

Main method - creates a \hyperlink{class_cowichan_open_m_p}{CowichanOpenMP} instance and executes \hyperlink{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{Cowichan::main}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em argc}]number of command line arguments. \item[{\em argv}]command line arguments. \end{description}
\end{Desc}

### 2009-08-07 doc/latex/cowichan__openmp_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__openmp_8hpp}{
\section{cowichan\_\-openmp/cowichan\_\-openmp.hpp File Reference}
\label{cowichan__openmp_8hpp}\index{cowichan\_\-openmp/cowichan\_\-openmp.hpp@{cowichan\_\-openmp/cowichan\_\-openmp.hpp}}
}
Open Multi-Processing (OpenMP) implementation of \hyperlink{class_cowichan}{Cowichan} problems.

{\tt \#include \char`\"{}../cowichan/cowichan.hpp\char`\"{}}\par
{\tt \#include $<$omp.h$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_cowichan_open_m_p}{CowichanOpenMP}
\begin{CompactList}\small\item\em Open Multi-Processing (OpenMP) implementation. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__openmp}{cowichan\_\-openmp}
\begin{CompactList}\small\item\em Additional classes and functions specific to openmp implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
Open Multi-Processing (OpenMP) implementation of \hyperlink{class_cowichan}{Cowichan} problems.

### 2009-08-07 doc/latex/cowichan__serial_2hull_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2hull_8cpp}{
\section{cowichan\_\-serial/hull.cpp File Reference}
\label{cowichan__serial_2hull_8cpp}\index{cowichan\_\-serial/hull.cpp@{cowichan\_\-serial/hull.cpp}}
}
Serial hull implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__serial_b0aad3f081e04d4676fae7b6ae8d7d10}{cowichan\_\-serial::quickhull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn)
\item 
void \hyperlink{namespacecowichan__serial_e1e9f58ca5f603f8215abffb30b78d71}{cowichan\_\-serial::split} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn, \hyperlink{class_point}{Point} $\ast$p1, \hyperlink{class_point}{Point} $\ast$p2)
\end{CompactItemize}

\subsection{Detailed Description}
Serial hull implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_e75cedf0a296dd7ad3dc41cadc51776f}{CowichanSerial::hull} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2invperc_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2invperc_8cpp}{
\section{cowichan\_\-serial/invperc.cpp File Reference}
\label{cowichan__serial_2invperc_8cpp}\index{cowichan\_\-serial/invperc.cpp@{cowichan\_\-serial/invperc.cpp}}
}
Serial invperc implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
{\tt \#include $<$queue$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__serial_1_1_perc_point}{cowichan\_\-serial::PercPoint}
\begin{CompactList}\small\item\em Percolation point. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
Serial invperc implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial::invperc} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2life_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2life_8cpp}{
\section{cowichan\_\-serial/life.cpp File Reference}
\label{cowichan__serial_2life_8cpp}\index{cowichan\_\-serial/life.cpp@{cowichan\_\-serial/life.cpp}}
}
Serial life implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__serial_93fd16bd1e6cb78235571ce617773639}{cowichan\_\-serial::sumNeighbours} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} first, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} r, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} c, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\end{CompactItemize}

\subsection{Detailed Description}
Serial life implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_ffd5b022c1d24226e11924094d8af349}{CowichanSerial::life} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2mandel_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2mandel_8cpp}{
\section{cowichan\_\-serial/mandel.cpp File Reference}
\label{cowichan__serial_2mandel_8cpp}\index{cowichan\_\-serial/mandel.cpp@{cowichan\_\-serial/mandel.cpp}}
}
Serial mandel implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_c96945095fd0ce7186a1d00a89f77d2c}{INT\_\-TYPE} \hyperlink{namespacecowichan__serial_4cde20338a4976564f4ceaf7a8a2c68e}{cowichan\_\-serial::mandel\_\-calc} (\hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} x, \hyperlink{cowichan_8hpp_4d521b2c54a1f6312cc8fa04827eaf98}{real} y)
\end{CompactItemize}

\subsection{Detailed Description}
Serial mandel implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_97a58b7901d8a7680cc28d42cb94d532}{CowichanSerial::mandel} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2norm_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2norm_8cpp}{
\section{cowichan\_\-serial/norm.cpp File Reference}
\label{cowichan__serial_2norm_8cpp}\index{cowichan\_\-serial/norm.cpp@{cowichan\_\-serial/norm.cpp}}
}
Serial norm implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__serial_bd0760b2ea719ded039c711324fc1bd4}{cowichan\_\-serial::findMinMax} (\hyperlink{class_point}{PointVector} points, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{Point} $\ast$minPoint, \hyperlink{class_point}{Point} $\ast$maxPoint)
\end{CompactItemize}

\subsection{Detailed Description}
Serial norm implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_0eeb47447c6a6b94ff7c6999c96fda0e}{CowichanSerial::norm} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2product_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2product_8cpp}{
\section{cowichan\_\-serial/product.cpp File Reference}
\label{cowichan__serial_2product_8cpp}\index{cowichan\_\-serial/product.cpp@{cowichan\_\-serial/product.cpp}}
}
Serial product implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par

\subsection{Detailed Description}
Serial product implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_00411b35445d7d3038b96d53e43bdffa}{CowichanSerial::product} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2randmat_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2randmat_8cpp}{
\section{cowichan\_\-serial/randmat.cpp File Reference}
\label{cowichan__serial_2randmat_8cpp}\index{cowichan\_\-serial/randmat.cpp@{cowichan\_\-serial/randmat.cpp}}
}
Serial randmat implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par

\subsection{Detailed Description}
Serial randmat implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_2d24c0e562f7b109ec2ed916f38e5911}{CowichanSerial::randmat} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2thresh_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2thresh_8cpp}{
\section{cowichan\_\-serial/thresh.cpp File Reference}
\label{cowichan__serial_2thresh_8cpp}\index{cowichan\_\-serial/thresh.cpp@{cowichan\_\-serial/thresh.cpp}}
}
Serial thresh implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par

\subsection{Detailed Description}
Serial thresh implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_7c0f93b2099ce919f91b5d953ff76511}{CowichanSerial::thresh} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2vecdiff_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2vecdiff_8cpp}{
\section{cowichan\_\-serial/vecdiff.cpp File Reference}
\label{cowichan__serial_2vecdiff_8cpp}\index{cowichan\_\-serial/vecdiff.cpp@{cowichan\_\-serial/vecdiff.cpp}}
}
Serial vecdiff implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par

\subsection{Detailed Description}
Serial vecdiff implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_34b75a2084051b3677071bb3c334d1f4}{CowichanSerial::vecdiff} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_2winnow_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_2winnow_8cpp}{
\section{cowichan\_\-serial/winnow.cpp File Reference}
\label{cowichan__serial_2winnow_8cpp}\index{cowichan\_\-serial/winnow.cpp@{cowichan\_\-serial/winnow.cpp}}
}
Serial winnow implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
\hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} \hyperlink{namespacecowichan__serial_4a948387cf6c0310f0dc43d50fdef087}{cowichan\_\-serial::mask\_\-count} (\hyperlink{cowichan_8hpp_a64c8df2f1e9c8ea68a7bcc19aca683e}{BoolMatrix} mask, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nr, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} nc)
\end{CompactItemize}

\subsection{Detailed Description}
Serial winnow implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_serial_33daca65431f792c2f4f0e7f8d29fa01}{CowichanSerial::winnow} \end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_8cpp}{
\section{cowichan\_\-serial/cowichan\_\-serial.cpp File Reference}
\label{cowichan__serial_8cpp}\index{cowichan\_\-serial/cowichan\_\-serial.cpp@{cowichan\_\-serial/cowichan\_\-serial.cpp}}
}
This file contains main method that drives the serial implementation.

{\tt \#include \char`\"{}cowichan\_\-serial.hpp\char`\"{}}\par
\subsection*{Functions}
\begin{CompactItemize}
\item 
int \hyperlink{cowichan__serial_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
\end{CompactItemize}

\subsection{Detailed Description}
This file contains main method that drives the serial implementation.

\subsection{Function Documentation}
\hypertarget{cowichan__serial_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
\index{cowichan\_\-serial.cpp@{cowichan\_\-serial.cpp}!main@{main}}
\index{main@{main}!cowichan_serial.cpp@{cowichan\_\-serial.cpp}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
\label{cowichan__serial_8cpp_0ddf1224851353fc92bfbff6f499fa97}

Main method - creates a \hyperlink{class_cowichan_serial}{CowichanSerial} instance and executes \hyperlink{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{Cowichan::main}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em argc}]number of command line arguments. \item[{\em argv}]command line arguments. \end{description}
\end{Desc}

### 2009-08-07 doc/latex/cowichan__serial_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__serial_8hpp}{
\section{cowichan\_\-serial/cowichan\_\-serial.hpp File Reference}
\label{cowichan__serial_8hpp}\index{cowichan\_\-serial/cowichan\_\-serial.hpp@{cowichan\_\-serial/cowichan\_\-serial.hpp}}
}
Serial implementation of \hyperlink{class_cowichan}{Cowichan} problems.

{\tt \#include \char`\"{}../cowichan/cowichan.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_cowichan_serial}{CowichanSerial}
\begin{CompactList}\small\item\em Serial implementation. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__serial}{cowichan\_\-serial}
\begin{CompactList}\small\item\em Additional classes and functions specific to serial implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
Serial implementation of \hyperlink{class_cowichan}{Cowichan} problems.

### 2009-08-07 doc/latex/cowichan__tbb_2gauss_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_2gauss_8cpp}{
\section{cowichan\_\-tbb/gauss.cpp File Reference}
\label{cowichan__tbb_2gauss_8cpp}\index{cowichan\_\-tbb/gauss.cpp@{cowichan\_\-tbb/gauss.cpp}}
}
TBB gauss implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_row_elimination}{cowichan\_\-tbb::RowElimination}
\begin{CompactList}\small\item\em Performs row elimination, i.e. eliminate i-th column in j-th row. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB gauss implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{CowichanTBB::gauss} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2half_8cpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__tbb_2gauss_8cpp.tex
@@
-\hypertarget{cowichan__tbb_2gauss_8cpp}{
-\section{cowichan\_\-tbb/gauss.cpp File Reference}
-\label{cowichan__tbb_2gauss_8cpp}\index{cowichan\_\-tbb/gauss.cpp@{cowichan\_\-tbb/gauss.cpp}}
+\hypertarget{cowichan__tbb_2half_8cpp}{
+\section{cowichan\_\-tbb/half.cpp File Reference}
+\label{cowichan__tbb_2half_8cpp}\index{cowichan\_\-tbb/half.cpp@{cowichan\_\-tbb/half.cpp}}
 }
-TBB gauss implementation.
+TBB half implementation.
 
@@
 \item 
-class \hyperlink{classcowichan__tbb_1_1_row_elimination}{cowichan\_\-tbb::RowElimination}
-\begin{CompactList}\small\item\em Performs row elimination, i.e. eliminate i-th column in j-th row. \item\end{CompactList}\end{CompactItemize}
+class \hyperlink{classcowichan__tbb_1_1_shuffle}{cowichan\_\-tbb::Shuffle}
+\begin{CompactList}\small\item\em This class does a halving shuffle. \item\end{CompactList}\end{CompactItemize}
 \subsection*{Namespaces}
@@
 \subsection{Detailed Description}
-TBB gauss implementation.
+TBB half implementation.
 
 \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_t_b_b_d51868e0e65482f7c9e69d73b3904822}{CowichanTBB::gauss} \end{Desc}
+\item[See also:]\hyperlink{class_cowichan_t_b_b_bbf17e641657d54fc0c571c008ac8f7a}{CowichanTBB::half} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2hull_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_2hull_8cpp}{
\section{cowichan\_\-tbb/hull.cpp File Reference}
\label{cowichan__tbb_2hull_8cpp}\index{cowichan\_\-tbb/hull.cpp@{cowichan\_\-tbb/hull.cpp}}
}
TBB hull implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_maximum_distance}{cowichan\_\-tbb::MaximumDistance}
\begin{CompactList}\small\item\em Compute maximum distance for hull. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Functions}
\begin{CompactItemize}
\item 
void \hyperlink{namespacecowichan__tbb_08647459098335f4d3abda9054a99547}{cowichan\_\-tbb::quickhull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn)
\item 
void \hyperlink{namespacecowichan__tbb_689d689bc3d950d3dd7fb9df89d4156e}{cowichan\_\-tbb::hull\_\-split} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn, \hyperlink{class_point}{Point} $\ast$p1, \hyperlink{class_point}{Point} $\ast$p2)
\end{CompactItemize}

\subsection{Detailed Description}
TBB hull implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{CowichanTBB::hull} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2invperc_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_2invperc_8cpp}{
\section{cowichan\_\-tbb/invperc.cpp File Reference}
\label{cowichan__tbb_2invperc_8cpp}\index{cowichan\_\-tbb/invperc.cpp@{cowichan\_\-tbb/invperc.cpp}}
}
TBB (NOT parallel) invperc implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
{\tt \#include $<$queue$>$}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_perc_point}{cowichan\_\-tbb::PercPoint}
\begin{CompactList}\small\item\em Percolation point. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB (NOT parallel) invperc implementation.

The same as \hyperlink{class_cowichan_serial_9b1cf3fcbb40498609826433b8ea2f6a}{CowichanSerial::invperc}

### 2009-08-07 doc/latex/cowichan__tbb_2life_8cpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__tbb_2hull_8cpp.tex
@@
-\hypertarget{cowichan__tbb_2hull_8cpp}{
-\section{cowichan\_\-tbb/hull.cpp File Reference}
-\label{cowichan__tbb_2hull_8cpp}\index{cowichan\_\-tbb/hull.cpp@{cowichan\_\-tbb/hull.cpp}}
+\hypertarget{cowichan__tbb_2life_8cpp}{
+\section{cowichan\_\-tbb/life.cpp File Reference}
+\label{cowichan__tbb_2life_8cpp}\index{cowichan\_\-tbb/life.cpp@{cowichan\_\-tbb/life.cpp}}
 }
-TBB hull implementation.
+TBB life implementation.
 
@@
 \item 
-class \hyperlink{classcowichan__tbb_1_1_maximum_distance}{cowichan\_\-tbb::MaximumDistance}
-\begin{CompactList}\small\item\em Compute maximum distance for hull. \item\end{CompactList}\end{CompactItemize}
+class \hyperlink{classcowichan__tbb_1_1_game_of_life}{cowichan\_\-tbb::GameOfLife}
+\begin{CompactList}\small\item\em Ping-pong solution to game of life. \item\end{CompactList}\end{CompactItemize}
 \subsection*{Namespaces}
@@
 \end{CompactItemize}
-\subsection*{Functions}
-\begin{CompactItemize}
-\item 
-void \hyperlink{namespacecowichan__tbb_08647459098335f4d3abda9054a99547}{cowichan\_\-tbb::quickhull} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn)
-\item 
-void \hyperlink{namespacecowichan__tbb_689d689bc3d950d3dd7fb9df89d4156e}{cowichan\_\-tbb::hull\_\-split} (\hyperlink{class_point}{PointVector} pointsIn, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} n, \hyperlink{class_point}{PointVector} pointsOut, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $\ast$hn, \hyperlink{class_point}{Point} $\ast$p1, \hyperlink{class_point}{Point} $\ast$p2)
-\end{CompactItemize}
 
 \subsection{Detailed Description}
-TBB hull implementation.
+TBB life implementation.
 
 \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_t_b_b_0d23aa05dc21fd9e8b033769097b4b18}{CowichanTBB::hull} \end{Desc}
+\item[See also:]\hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{CowichanTBB::life} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2mandel_8cpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__tbb_2life_8cpp.tex
@@
-\hypertarget{cowichan__tbb_2life_8cpp}{
-\section{cowichan\_\-tbb/life.cpp File Reference}
-\label{cowichan__tbb_2life_8cpp}\index{cowichan\_\-tbb/life.cpp@{cowichan\_\-tbb/life.cpp}}
+\hypertarget{cowichan__tbb_2mandel_8cpp}{
+\section{cowichan\_\-tbb/mandel.cpp File Reference}
+\label{cowichan__tbb_2mandel_8cpp}\index{cowichan\_\-tbb/mandel.cpp@{cowichan\_\-tbb/mandel.cpp}}
 }
-TBB life implementation.
+TBB mandel implementation.
 
@@
 \item 
-class \hyperlink{classcowichan__tbb_1_1_game_of_life}{cowichan\_\-tbb::GameOfLife}
-\begin{CompactList}\small\item\em Ping-pong solution to game of life. \item\end{CompactList}\end{CompactItemize}
+class \hyperlink{classcowichan__tbb_1_1_mandelbrot}{cowichan\_\-tbb::Mandelbrot}
+\begin{CompactList}\small\item\em Perform mandelbrot generation. \item\end{CompactList}\end{CompactItemize}
 \subsection*{Namespaces}
@@
 \subsection{Detailed Description}
-TBB life implementation.
+TBB mandel implementation.
 
 \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_t_b_b_273e3e8a2e05108f59ec613472bfa363}{CowichanTBB::life} \end{Desc}
+\item[See also:]\hyperlink{class_cowichan_t_b_b_1e58b60ff22a58aba69567ecc6740878}{CowichanTBB::mandel} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2norm_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_2norm_8cpp}{
\section{cowichan\_\-tbb/norm.cpp File Reference}
\label{cowichan__tbb_2norm_8cpp}\index{cowichan\_\-tbb/norm.cpp@{cowichan\_\-tbb/norm.cpp}}
}
TBB norm implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_min_max_reducer}{cowichan\_\-tbb::MinMaxReducer}
\begin{CompactList}\small\item\em Performs the minimum and maximum reductions. \item\end{CompactList}\item 
class \hyperlink{classcowichan__tbb_1_1_normalizer}{cowichan\_\-tbb::Normalizer}
\begin{CompactList}\small\item\em Normalizes points. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB norm implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_ca08645b51a242317a115cd7ce7d81fe}{CowichanTBB::norm} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2outer_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_2outer_8cpp}{
\section{cowichan\_\-tbb/outer.cpp File Reference}
\label{cowichan__tbb_2outer_8cpp}\index{cowichan\_\-tbb/outer.cpp@{cowichan\_\-tbb/outer.cpp}}
}
TBB outer implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_point_distances}{cowichan\_\-tbb::PointDistances}
\begin{CompactList}\small\item\em Fills in matrix and vector with distances. \item\end{CompactList}\item 
class \hyperlink{classcowichan__tbb_1_1_make_dominant}{cowichan\_\-tbb::MakeDominant}
\begin{CompactList}\small\item\em Makes a given matrix diagonally dominant. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB outer implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{CowichanTBB::outer} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2product_8cpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__tbb_2outer_8cpp.tex
@@
-\hypertarget{cowichan__tbb_2outer_8cpp}{
-\section{cowichan\_\-tbb/outer.cpp File Reference}
-\label{cowichan__tbb_2outer_8cpp}\index{cowichan\_\-tbb/outer.cpp@{cowichan\_\-tbb/outer.cpp}}
+\hypertarget{cowichan__tbb_2product_8cpp}{
+\section{cowichan\_\-tbb/product.cpp File Reference}
+\label{cowichan__tbb_2product_8cpp}\index{cowichan\_\-tbb/product.cpp@{cowichan\_\-tbb/product.cpp}}
 }
-TBB outer implementation.
+TBB product implementation.
 
@@
 \item 
-class \hyperlink{classcowichan__tbb_1_1_point_distances}{cowichan\_\-tbb::PointDistances}
-\begin{CompactList}\small\item\em Fills in matrix and vector with distances. \item\end{CompactList}\item 
-class \hyperlink{classcowichan__tbb_1_1_make_dominant}{cowichan\_\-tbb::MakeDominant}
-\begin{CompactList}\small\item\em Makes a given matrix diagonally dominant. \item\end{CompactList}\end{CompactItemize}
+class \hyperlink{classcowichan__tbb_1_1_product}{cowichan\_\-tbb::Product}
+\begin{CompactList}\small\item\em Performs matrix vector product. \item\end{CompactList}\end{CompactItemize}
 \subsection*{Namespaces}
@@
 \subsection{Detailed Description}
-TBB outer implementation.
+TBB product implementation.
 
 \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_t_b_b_92a763b9733cdcceda13b7b7ea0bc229}{CowichanTBB::outer} \end{Desc}
+\item[See also:]\hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{CowichanTBB::product} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2randmat_8cpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__tbb_2product_8cpp.tex
@@
-\hypertarget{cowichan__tbb_2product_8cpp}{
-\section{cowichan\_\-tbb/product.cpp File Reference}
-\label{cowichan__tbb_2product_8cpp}\index{cowichan\_\-tbb/product.cpp@{cowichan\_\-tbb/product.cpp}}
+\hypertarget{cowichan__tbb_2randmat_8cpp}{
+\section{cowichan\_\-tbb/randmat.cpp File Reference}
+\label{cowichan__tbb_2randmat_8cpp}\index{cowichan\_\-tbb/randmat.cpp@{cowichan\_\-tbb/randmat.cpp}}
 }
-TBB product implementation.
+TBB randmat implementation.
 
@@
 \item 
-class \hyperlink{classcowichan__tbb_1_1_product}{cowichan\_\-tbb::Product}
-\begin{CompactList}\small\item\em Performs matrix vector product. \item\end{CompactList}\end{CompactItemize}
+class \hyperlink{classcowichan__tbb_1_1_random_generator}{cowichan\_\-tbb::RandomGenerator}
+\begin{CompactList}\small\item\em Random generator. \item\end{CompactList}\end{CompactItemize}
 \subsection*{Namespaces}
@@
 \subsection{Detailed Description}
-TBB product implementation.
+TBB randmat implementation.
 
 \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{CowichanTBB::product} \end{Desc}
+\item[See also:]\hyperlink{class_cowichan_t_b_b_b9b5cb4b4b5dc8907b2a01825cd4aaff}{CowichanTBB::randmat} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2sor_8cpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__tbb_2product_8cpp.tex
@@
-\hypertarget{cowichan__tbb_2product_8cpp}{
-\section{cowichan\_\-tbb/product.cpp File Reference}
-\label{cowichan__tbb_2product_8cpp}\index{cowichan\_\-tbb/product.cpp@{cowichan\_\-tbb/product.cpp}}
+\hypertarget{cowichan__tbb_2sor_8cpp}{
+\section{cowichan\_\-tbb/sor.cpp File Reference}
+\label{cowichan__tbb_2sor_8cpp}\index{cowichan\_\-tbb/sor.cpp@{cowichan\_\-tbb/sor.cpp}}
 }
-TBB product implementation.
+TBB sor implementation.
 
@@
 \item 
-class \hyperlink{classcowichan__tbb_1_1_product}{cowichan\_\-tbb::Product}
-\begin{CompactList}\small\item\em Performs matrix vector product. \item\end{CompactList}\end{CompactItemize}
+class \hyperlink{classcowichan__tbb_1_1_relaxer}{cowichan\_\-tbb::Relaxer}
+\begin{CompactList}\small\item\em Performs one iteration of relaxation. \item\end{CompactList}\end{CompactItemize}
 \subsection*{Namespaces}
@@
 \subsection{Detailed Description}
-TBB product implementation.
+TBB sor implementation.
 
 \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_t_b_b_f3144458520e2dff1f9fad5753f6fc3d}{CowichanTBB::product} \end{Desc}
+\item[See also:]\hyperlink{class_cowichan_t_b_b_dbb32ce457d0edca6815ab1cb2459276}{CowichanTBB::sor} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2thresh_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_2thresh_8cpp}{
\section{cowichan\_\-tbb/thresh.cpp File Reference}
\label{cowichan__tbb_2thresh_8cpp}\index{cowichan\_\-tbb/thresh.cpp@{cowichan\_\-tbb/thresh.cpp}}
}
TBB thresh implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_max_reducer}{cowichan\_\-tbb::MaxReducer}
\begin{CompactList}\small\item\em Finds the maximum value in the image. \item\end{CompactList}\item 
class \hyperlink{classcowichan__tbb_1_1_histogram}{cowichan\_\-tbb::Histogram}
\begin{CompactList}\small\item\em This class calculates the histogram of a matrix. \item\end{CompactList}\item 
class \hyperlink{classcowichan__tbb_1_1_threshold}{cowichan\_\-tbb::Threshold}
\begin{CompactList}\small\item\em Creates a boolean array based on a cut-off value. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB thresh implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_3306d21f0b3d12cc2e3b050b99812a27}{CowichanTBB::thresh} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2vecdiff_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_2vecdiff_8cpp}{
\section{cowichan\_\-tbb/vecdiff.cpp File Reference}
\label{cowichan__tbb_2vecdiff_8cpp}\index{cowichan\_\-tbb/vecdiff.cpp@{cowichan\_\-tbb/vecdiff.cpp}}
}
TBB vecdiff implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{classcowichan__tbb_1_1_norm_diff}{cowichan\_\-tbb::NormDiff}
\begin{CompactList}\small\item\em This class takes the 1-norm of the difference between two vectors. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}

\subsection{Detailed Description}
TBB vecdiff implementation.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{CowichanTBB::vecdiff} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_2winnow_8cpp.tex
Re-generated doxygen with linuxtuples additions.

--- 2009-08-07 doc/latex/cowichan__tbb_2vecdiff_8cpp.tex
@@
-\hypertarget{cowichan__tbb_2vecdiff_8cpp}{
-\section{cowichan\_\-tbb/vecdiff.cpp File Reference}
-\label{cowichan__tbb_2vecdiff_8cpp}\index{cowichan\_\-tbb/vecdiff.cpp@{cowichan\_\-tbb/vecdiff.cpp}}
+\hypertarget{cowichan__tbb_2winnow_8cpp}{
+\section{cowichan\_\-tbb/winnow.cpp File Reference}
+\label{cowichan__tbb_2winnow_8cpp}\index{cowichan\_\-tbb/winnow.cpp@{cowichan\_\-tbb/winnow.cpp}}
 }
-TBB vecdiff implementation.
+TBB winnow implementation.
 
@@
 \item 
-class \hyperlink{classcowichan__tbb_1_1_norm_diff}{cowichan\_\-tbb::NormDiff}
-\begin{CompactList}\small\item\em This class takes the 1-norm of the difference between two vectors. \item\end{CompactList}\end{CompactItemize}
+class \hyperlink{classcowichan__tbb_1_1_point_count}{cowichan\_\-tbb::PointCount}
+\begin{CompactList}\small\item\em Calculates the number of points to sort for winnow. \item\end{CompactList}\end{CompactItemize}
 \subsection*{Namespaces}
@@
 \subsection{Detailed Description}
-TBB vecdiff implementation.
+TBB winnow implementation.
 
 \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_t_b_b_28c976743df231fd183e4db9306050b1}{CowichanTBB::vecdiff} \end{Desc}
+\item[See also:]\hyperlink{class_cowichan_t_b_b_77178470ef780e0505137dc1d22a85a2}{CowichanTBB::winnow} \end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_8cpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_8cpp}{
\section{cowichan\_\-tbb/cowichan\_\-tbb.cpp File Reference}
\label{cowichan__tbb_8cpp}\index{cowichan\_\-tbb/cowichan\_\-tbb.cpp@{cowichan\_\-tbb/cowichan\_\-tbb.cpp}}
}
This file contains main method that drives the TBB implementation.

{\tt \#include \char`\"{}cowichan\_\-tbb.hpp\char`\"{}}\par
\subsection*{Functions}
\begin{CompactItemize}
\item 
int \hyperlink{cowichan__tbb_8cpp_0ddf1224851353fc92bfbff6f499fa97}{main} (int argc, char $\ast$argv\mbox{[}$\,$\mbox{]})
\end{CompactItemize}

\subsection{Detailed Description}
This file contains main method that drives the TBB implementation.

\subsection{Function Documentation}
\hypertarget{cowichan__tbb_8cpp_0ddf1224851353fc92bfbff6f499fa97}{
\index{cowichan\_\-tbb.cpp@{cowichan\_\-tbb.cpp}!main@{main}}
\index{main@{main}!cowichan_tbb.cpp@{cowichan\_\-tbb.cpp}}
\subsubsection[{main}]{\setlength{\rightskip}{0pt plus 5cm}int main (int {\em argc}, \/  char $\ast$ {\em argv}\mbox{[}$\,$\mbox{]})}}
\label{cowichan__tbb_8cpp_0ddf1224851353fc92bfbff6f499fa97}

Main method - creates a \hyperlink{class_cowichan_t_b_b}{CowichanTBB} instance and executes \hyperlink{class_cowichan_905f3eb45f21cdaa1c32a421d001fa4c}{Cowichan::main}. \begin{Desc}
\item[Parameters:]
\begin{description}
\item[{\em argc}]number of command line arguments. \item[{\em argv}]command line arguments. \end{description}
\end{Desc}

### 2009-08-07 doc/latex/cowichan__tbb_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{cowichan__tbb_8hpp}{
\section{cowichan\_\-tbb/cowichan\_\-tbb.hpp File Reference}
\label{cowichan__tbb_8hpp}\index{cowichan\_\-tbb/cowichan\_\-tbb.hpp@{cowichan\_\-tbb/cowichan\_\-tbb.hpp}}
}
Thread Building Blocks (TBB) implementation of \hyperlink{class_cowichan}{Cowichan} problems.

{\tt \#include \char`\"{}../cowichan/cowichan.hpp\char`\"{}}\par
{\tt \#include \char`\"{}tbb/task\_\-scheduler\_\-init.h\char`\"{}}\par
{\tt \#include \char`\"{}tbb/blocked\_\-range2d.h\char`\"{}}\par
{\tt \#include \char`\"{}tbb/parallel\_\-for.h\char`\"{}}\par
{\tt \#include \char`\"{}tbb/parallel\_\-reduce.h\char`\"{}}\par
{\tt \#include \char`\"{}tbb/parallel\_\-sort.h\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_cowichan_t_b_b}{CowichanTBB}
\begin{CompactList}\small\item\em Thread Building Blocks (TBB) implementation. \item\end{CompactList}\end{CompactItemize}
\subsection*{Namespaces}
\begin{CompactItemize}
\item 
namespace \hyperlink{namespacecowichan__tbb}{cowichan\_\-tbb}
\begin{CompactList}\small\item\em Additional classes and functions specific to tbb implementation. \item\end{CompactList}

\end{CompactItemize}
\subsection*{Typedefs}
\begin{CompactItemize}
\item 
typedef blocked\_\-range$<$ \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $>$ \hyperlink{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{Range}
\item 
typedef blocked\_\-range2d$<$ \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t}, \hyperlink{cowichan_8hpp_5b04577d5d21124855deaad298595371}{index\_\-t} $>$ \hyperlink{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{Range2D}
\end{CompactItemize}

\subsection{Detailed Description}
Thread Building Blocks (TBB) implementation of \hyperlink{class_cowichan}{Cowichan} problems.

\subsection{Typedef Documentation}
\hypertarget{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}{
\index{cowichan\_\-tbb.hpp@{cowichan\_\-tbb.hpp}!Range@{Range}}
\index{Range@{Range}!cowichan_tbb.hpp@{cowichan\_\-tbb.hpp}}
\subsubsection[{Range}]{\setlength{\rightskip}{0pt plus 5cm}typedef blocked\_\-range$<${\bf index\_\-t}$>$ {\bf Range}}}
\label{cowichan__tbb_8hpp_8e2057f725b08f3a15513c378a453a47}

One-dimensional tbb range. \hypertarget{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}{
\index{cowichan\_\-tbb.hpp@{cowichan\_\-tbb.hpp}!Range2D@{Range2D}}
\index{Range2D@{Range2D}!cowichan_tbb.hpp@{cowichan\_\-tbb.hpp}}
\subsubsection[{Range2D}]{\setlength{\rightskip}{0pt plus 5cm}typedef blocked\_\-range2d$<${\bf index\_\-t}, {\bf index\_\-t}$>$ {\bf Range2D}}}
\label{cowichan__tbb_8hpp_e591b8e6980ddc5982ee22655da2ab8e}

Two-dimensional tbb range.

### 2009-08-07 doc/latex/files.tex
Re-generated doxygen with linuxtuples additions.

@@
 \item\contentsline{section}{cowichan/\hyperlink{cowichan__defaults_8hpp}{cowichan\_\-defaults.hpp} (Defaults for \hyperlink{class_cowichan}{Cowichan} programs )}{\pageref{cowichan__defaults_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_8cpp}{cowichan\_\-lt.cpp} (Entry point for the cowichan linuxtuples application )}{\pageref{cowichan__lt_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_8hpp}{cowichan\_\-lt.hpp} (Datatypes and common routines for \hyperlink{class_cowichan}{Cowichan} programs; LinuxTuples implementation )}{\pageref{cowichan__lt_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2gauss_8cpp}{gauss.cpp} (LinuxTuples gauss implementation )}{\pageref{cowichan__lt_2src_2gauss_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{gauss_8hpp}{gauss.hpp} (LinuxTuples gauss header file )}{\pageref{gauss_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2half_8cpp}{half.cpp} (LinuxTuples half implementation )}{\pageref{cowichan__lt_2src_2half_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{half_8hpp}{half.hpp} (LinuxTuples halving shuffle header file )}{\pageref{half_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2hull_8cpp}{hull.cpp} (LinuxTuples hull implementation )}{\pageref{cowichan__lt_2src_2hull_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{hull_8hpp}{hull.hpp} (LinuxTuples iterative convex hull header file )}{\pageref{hull_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2invperc_8cpp}{invperc.cpp} (LinuxTuples (actually serial) invperc implementation )}{\pageref{cowichan__lt_2src_2invperc_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2life_8cpp}{life.cpp} (LinuxTuples life implementation )}{\pageref{cowichan__lt_2src_2life_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{life_8hpp}{life.hpp} (LinuxTuples conway's game of life header file )}{\pageref{life_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2mandel_8cpp}{mandel.cpp} (LinuxTuples mandelbrot set implementation )}{\pageref{cowichan__lt_2src_2mandel_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{mandel_8hpp}{mandel.hpp} (LinuxTuples mandelbrot set header file )}{\pageref{mandel_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2norm_8cpp}{norm.cpp} (LinuxTuples norm implementation )}{\pageref{cowichan__lt_2src_2norm_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{norm_8hpp}{norm.hpp} (LinuxTuples norm header file )}{\pageref{norm_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2outer_8cpp}{outer.cpp} (LinuxTuples outer implementation )}{\pageref{cowichan__lt_2src_2outer_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{outer_8hpp}{outer.hpp} (LinuxTuples outer product header file )}{\pageref{outer_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2product_8cpp}{product.cpp} (LinuxTuples product implementation )}{\pageref{cowichan__lt_2src_2product_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{product_8hpp}{product.hpp} (LinuxTuples product header file )}{\pageref{product_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2randmat_8cpp}{randmat.cpp} (LinuxTuples randmat implementation )}{\pageref{cowichan__lt_2src_2randmat_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{randmat_8hpp}{randmat.hpp} (LinuxTuples header file for the randmat cowichan problem )}{\pageref{randmat_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2sor_8cpp}{sor.cpp} (LinuxTuples successive over-relaxation implementation )}{\pageref{cowichan__lt_2src_2sor_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{sor_8hpp}{sor.hpp} (LinuxTuples successive over-relaxation header file )}{\pageref{sor_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2thresh_8cpp}{thresh.cpp} (LinuxTuples thresh implementation )}{\pageref{cowichan__lt_2src_2thresh_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{thresh_8hpp}{thresh.hpp} (LinuxTuples histogram thresholding header file )}{\pageref{thresh_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{tuple__common_8cpp}{tuple\_\-common.cpp} (Implementation of common tuple application logic )}{\pageref{tuple__common_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{tuple__common_8hpp}{tuple\_\-common.hpp} (Header file for tuple applications )}{\pageref{tuple__common_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2vecdiff_8cpp}{vecdiff.cpp} (LinuxTuples vecdiff implementation )}{\pageref{cowichan__lt_2src_2vecdiff_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{vecdiff_8hpp}{vecdiff.hpp} (LinuxTuples vector difference header file )}{\pageref{vecdiff_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{cowichan__lt_2src_2winnow_8cpp}{winnow.cpp} (LinuxTuples winnow implementation )}{\pageref{cowichan__lt_2src_2winnow_8cpp}}{}
+\item\contentsline{section}{cowichan\_\-lt/src/\hyperlink{winnow_8hpp}{winnow.hpp} (LinuxTuples winnow header file )}{\pageref{winnow_8hpp}}{}
 \item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_8cpp}{cowichan\_\-mpi.cpp} (This file contains main method that drives the MPI implementation )}{\pageref{cowichan__mpi_8cpp}}{}

### 2009-08-07 doc/latex/gauss_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{gauss_8hpp}{
\section{cowichan\_\-lt/src/gauss.hpp File Reference}
\label{gauss_8hpp}\index{cowichan\_\-lt/src/gauss.hpp@{cowichan\_\-lt/src/gauss.hpp}}
}
LinuxTuples gauss header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_forward}{LTForward}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples gauss header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_fd3e659afb38415f56fd508333b7bb1b}{CowichanLinuxTuples::gauss} \end{Desc}

### 2009-08-07 doc/latex/half_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{half_8hpp}{
\section{cowichan\_\-lt/src/half.hpp File Reference}
\label{half_8hpp}\index{cowichan\_\-lt/src/half.hpp@{cowichan\_\-lt/src/half.hpp}}
}
LinuxTuples halving shuffle header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_half}{LTHalf}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples halving shuffle header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_f3bd58ad2653f06d242d473e7bffda0b}{CowichanLinuxTuples::half} \end{Desc}

### 2009-08-07 doc/latex/hierarchy.tex
doxygen documentation as of now
Re-generated doxygen with linuxtuples additions.

@@
 \begin{CompactList}
+\item \contentsline{section}{CowichanLinuxTuples}{\pageref{class_cowichan_linux_tuples}}{}
+\item \contentsline{section}{CowichanMPI}{\pageref{class_cowichan_m_p_i}}{}
 \item \contentsline{section}{CowichanOpenMP}{\pageref{class_cowichan_open_m_p}}{}
@@
 \item \contentsline{section}{cowichan\_\-tbb::NormDiff}{\pageref{classcowichan__tbb_1_1_norm_diff}}{}
+\item \contentsline{section}{PercPoint}{\pageref{class_perc_point}}{}
+\item \contentsline{section}{cowichan\_\-openmp::PercPoint}{\pageref{classcowichan__openmp_1_1_perc_point}}{}
+\item \contentsline{section}{cowichan\_\-tbb::PercPoint}{\pageref{classcowichan__tbb_1_1_perc_point}}{}
 \item \contentsline{section}{cowichan\_\-serial::PercPoint}{\pageref{classcowichan__serial_1_1_perc_point}}{}
-\item \contentsline{section}{cowichan\_\-tbb::PercPoint}{\pageref{classcowichan__tbb_1_1_perc_point}}{}
-\item \contentsline{section}{cowichan\_\-openmp::PercPoint}{\pageref{classcowichan__openmp_1_1_perc_point}}{}
 \item \contentsline{section}{Point}{\pageref{class_point}}{}
@@
 \item \contentsline{section}{cowichan\_\-tbb::Threshold}{\pageref{classcowichan__tbb_1_1_threshold}}{}
+\item \contentsline{section}{TupleApplication}{\pageref{class_tuple_application}}{}
+\begin{CompactList}
+\item \contentsline{section}{LTBounds}{\pageref{class_l_t_bounds}}{}
+\item \contentsline{section}{LTForward}{\pageref{class_l_t_forward}}{}
+\item \contentsline{section}{LTFrequency}{\pageref{class_l_t_frequency}}{}
+\item \contentsline{section}{LTHalf}{\pageref{class_l_t_half}}{}
+\item \contentsline{section}{LTHull}{\pageref{class_l_t_hull}}{}
+\item \contentsline{section}{LTLife}{\pageref{class_l_t_life}}{}
+\item \contentsline{section}{LTMandel}{\pageref{class_l_t_mandel}}{}
+\item \contentsline{section}{LTNorm}{\pageref{class_l_t_norm}}{}
+\item \contentsline{section}{LTOuter}{\pageref{class_l_t_outer}}{}
+\item \contentsline{section}{LTProduct}{\pageref{class_l_t_product}}{}
+\item \contentsline{section}{LTRandmat}{\pageref{class_l_t_randmat}}{}
+\item \contentsline{section}{LTSor}{\pageref{class_l_t_sor}}{}
+\item \contentsline{section}{LTThresh}{\pageref{class_l_t_thresh}}{}
+\item \contentsline{section}{LTVecdiff}{\pageref{class_l_t_vecdiff}}{}
+\item \contentsline{section}{LTWinnow}{\pageref{class_l_t_winnow}}{}
+\end{CompactList}
 \item \contentsline{section}{WeightedPoint}{\pageref{class_weighted_point}}{}

### 2009-08-07 doc/latex/hull_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{hull_8hpp}{
\section{cowichan\_\-lt/src/hull.hpp File Reference}
\label{hull_8hpp}\index{cowichan\_\-lt/src/hull.hpp@{cowichan\_\-lt/src/hull.hpp}}
}
LinuxTuples iterative convex hull header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_hull}{LTHull}
\end{CompactItemize}
\subsection*{Defines}
\begin{CompactItemize}
\item 
\#define \hyperlink{hull_8hpp_ac10d34dd2d4faa7f96f7ed398c884f4}{NO\_\-POINT}~-1
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples iterative convex hull header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_501f87594d62af261ff0b1954a60ddc4}{CowichanLinuxTuples::hull} \end{Desc}

\subsection{Define Documentation}
\hypertarget{hull_8hpp_ac10d34dd2d4faa7f96f7ed398c884f4}{
\index{hull.hpp@{hull.hpp}!NO\_\-POINT@{NO\_\-POINT}}
\index{NO\_\-POINT@{NO\_\-POINT}!hull.hpp@{hull.hpp}}
\subsubsection[{NO\_\-POINT}]{\setlength{\rightskip}{0pt plus 5cm}\#define NO\_\-POINT~-1}}
\label{hull_8hpp_ac10d34dd2d4faa7f96f7ed398c884f4}

Indicator value to denote no valid point is stored.

### 2009-08-07 doc/latex/life_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{life_8hpp}{
\section{cowichan\_\-lt/src/life.hpp File Reference}
\label{life_8hpp}\index{cowichan\_\-lt/src/life.hpp@{cowichan\_\-lt/src/life.hpp}}
}
LinuxTuples conway's game of life header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_life}{LTLife}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples conway's game of life header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_967138bbdead76eec50f31465dd9d5f1}{CowichanLinuxTuples::life} \end{Desc}

### 2009-08-07 doc/latex/mandel_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{mandel_8hpp}{
\section{cowichan\_\-lt/src/mandel.hpp File Reference}
\label{mandel_8hpp}\index{cowichan\_\-lt/src/mandel.hpp@{cowichan\_\-lt/src/mandel.hpp}}
}
LinuxTuples mandelbrot set header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_mandel}{LTMandel}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples mandelbrot set header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_8774404842868381379ec627e8373cec}{CowichanLinuxTuples::mandel} \end{Desc}

### 2009-08-07 doc/latex/norm_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{norm_8hpp}{
\section{cowichan\_\-lt/src/norm.hpp File Reference}
\label{norm_8hpp}\index{cowichan\_\-lt/src/norm.hpp@{cowichan\_\-lt/src/norm.hpp}}
}
LinuxTuples norm header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_bounds}{LTBounds}
\item 
class \hyperlink{class_l_t_norm}{LTNorm}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples norm header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_922209c219bb5b9268263dae4000ea81}{CowichanLinuxTuples::norm} \end{Desc}

### 2009-08-07 doc/latex/outer_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{outer_8hpp}{
\section{cowichan\_\-lt/src/outer.hpp File Reference}
\label{outer_8hpp}\index{cowichan\_\-lt/src/outer.hpp@{cowichan\_\-lt/src/outer.hpp}}
}
LinuxTuples outer product header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_outer}{LTOuter}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples outer product header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_b16e1c3b60f920c12460a74289558dcb}{CowichanLinuxTuples::outer} \end{Desc}

### 2009-08-07 doc/latex/product_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{product_8hpp}{
\section{cowichan\_\-lt/src/product.hpp File Reference}
\label{product_8hpp}\index{cowichan\_\-lt/src/product.hpp@{cowichan\_\-lt/src/product.hpp}}
}
LinuxTuples product header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_product}{LTProduct}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples product header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_c4874f09c54056ac94d00f9e341a94f4}{CowichanLinuxTuples::product} \end{Desc}

### 2009-08-07 doc/latex/randmat_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{randmat_8hpp}{
\section{cowichan\_\-lt/src/randmat.hpp File Reference}
\label{randmat_8hpp}\index{cowichan\_\-lt/src/randmat.hpp@{cowichan\_\-lt/src/randmat.hpp}}
}
LinuxTuples header file for the randmat cowichan problem.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_randmat}{LTRandmat}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples header file for the randmat cowichan problem.

### 2009-08-07 doc/latex/sor_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{sor_8hpp}{
\section{cowichan\_\-lt/src/sor.hpp File Reference}
\label{sor_8hpp}\index{cowichan\_\-lt/src/sor.hpp@{cowichan\_\-lt/src/sor.hpp}}
}
LinuxTuples successive over-relaxation header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_sor}{LTSor}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples successive over-relaxation header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_aa2c469c3f520a5a56c5ec1ac9a5a6c6}{CowichanLinuxTuples::sor} \end{Desc}

### 2009-08-07 doc/latex/thresh_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{thresh_8hpp}{
\section{cowichan\_\-lt/src/thresh.hpp File Reference}
\label{thresh_8hpp}\index{cowichan\_\-lt/src/thresh.hpp@{cowichan\_\-lt/src/thresh.hpp}}
}
LinuxTuples histogram thresholding header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_frequency}{LTFrequency}
\item 
class \hyperlink{class_l_t_thresh}{LTThresh}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples histogram thresholding header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_b8cf055ec3d42a1ccde6d818c5a43a74}{CowichanLinuxTuples::thresh} \end{Desc}

### 2009-08-07 doc/latex/vecdiff_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{vecdiff_8hpp}{
\section{cowichan\_\-lt/src/vecdiff.hpp File Reference}
\label{vecdiff_8hpp}\index{cowichan\_\-lt/src/vecdiff.hpp@{cowichan\_\-lt/src/vecdiff.hpp}}
}
LinuxTuples vector difference header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_vecdiff}{LTVecdiff}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples vector difference header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_213185666ea7dd0c9abc59899b454086}{CowichanLinuxTuples::vecdiff} \end{Desc}

### 2009-08-07 doc/latex/winnow_8hpp.tex
Re-generated doxygen with linuxtuples additions.

\hypertarget{winnow_8hpp}{
\section{cowichan\_\-lt/src/winnow.hpp File Reference}
\label{winnow_8hpp}\index{cowichan\_\-lt/src/winnow.hpp@{cowichan\_\-lt/src/winnow.hpp}}
}
LinuxTuples winnow header file.

{\tt \#include \char`\"{}tuple\_\-common.hpp\char`\"{}}\par
{\tt \#include \char`\"{}cowichan\_\-lt.hpp\char`\"{}}\par
\subsection*{Classes}
\begin{CompactItemize}
\item 
class \hyperlink{class_l_t_winnow}{LTWinnow}
\end{CompactItemize}

\subsection{Detailed Description}
LinuxTuples winnow header file.

\begin{Desc}
\item[See also:]\hyperlink{class_cowichan_linux_tuples_2b1a77ecbd01ec847cf63fe7ebb50e35}{CowichanLinuxTuples::winnow} \end{Desc}

### 2009-08-11 doc/latex/class_cowichan_m_p_i.tex
add comment to thresh, invperc small fix

@@
 cowichan\_\-mpi/\hyperlink{cowichan__mpi_8cpp}{cowichan\_\-mpi.cpp}\item 
+cowichan\_\-mpi/\hyperlink{cowichan__mpi_2gauss_8cpp}{gauss.cpp}\item 
+cowichan\_\-mpi/\hyperlink{cowichan__mpi_2thresh_8cpp}{thresh.cpp}\item 
 cowichan\_\-mpi/\hyperlink{cowichan__mpi_2winnow_8cpp}{winnow.cpp}\end{CompactItemize}

### 2009-08-11 doc/latex/files.tex
add comment to thresh, invperc small fix

@@
 \item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_8hpp}{cowichan\_\-mpi.hpp} (Boost Message Passing Interface (MPI) implementation of \hyperlink{class_cowichan}{Cowichan} problems )}{\pageref{cowichan__mpi_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2gauss_8cpp}{gauss.cpp} (MPI gauss implementation )}{\pageref{cowichan__mpi_2gauss_8cpp}}{}
 \item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2sort_8cpp}{sort.cpp} (Implementation of MPI sorting algorithms )}{\pageref{cowichan__mpi_2sort_8cpp}}{}
 \item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2sort_8hpp}{sort.hpp} (Sorting algorithms that use MPI )}{\pageref{cowichan__mpi_2sort_8hpp}}{}
+\item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2thresh_8cpp}{thresh.cpp} (MPI thresh implementation )}{\pageref{cowichan__mpi_2thresh_8cpp}}{}
 \item\contentsline{section}{cowichan\_\-mpi/\hyperlink{cowichan__mpi_2winnow_8cpp}{winnow.cpp} (MPI winnow implementation )}{\pageref{cowichan__mpi_2winnow_8cpp}}{}

### 2009-08-11 cowichan/cowichan.cpp

@@ -33,12 +33,36 @@ void Cowichan::print_vector(PointVector points)
   index_t r;
 
   for (r = 0; r < n; r++) {
-    std::cout << "[" << points[r].x << ", " << points[r].y << "]\n";
+    std::cout << "[" << points[r].x << ", " << points[r].y << "]" << std::endl;
   }
-  std::cout << "\n";
+  std::cout << std::endl;
 }
 #else
 void Cowichan::print_vector(PointVector /* points */) { }
+#endif
+
+#ifdef OUTPUT_DATA
+  void Cowichan::print_bool_rect_matrix(BoolMatrix matrix)
+  {
+    index_t r, c;
+
+    for (r = 0; r < nr; r++) {
+      for (c = 0; c < nc; c++) {
+        if (MATRIX_RECT(matrix, r, c))
+        {
+          std::cout << "x";
+        }
+        else
+        {
+          std::cout << " ";
+        }
+      }
+      std::cout << std::endl;
+    }
+    std::cout << std::endl;
+  }
+#else
+  void Cowichan::print_bool_rect_matrix(BoolMatrix /* matrix */) { }
 #endif
 
 /*****************************************************************************/
@@ -227,7 +251,7 @@ void Cowichan::main (int argc, char* argv[], bool use_randmat, bool use_thresh)
       end = get_ticks ();
       invperc (matrix, mask);
       timeInfo(&start, &end, INVPERC);
-      print_rect_matrix<bool> (mask);
+      print_bool_rect_matrix (mask);
 
       // clean up
       delete [] matrix;
@@ -262,7 +286,7 @@ void Cowichan::main (int argc, char* argv[], bool use_randmat, bool use_thresh)
       end = get_ticks ();
       thresh (matrix, mask);
       timeInfo(&start, &end, THRESH);
-      print_rect_matrix<bool> (mask);
+      print_bool_rect_matrix (mask);
 
       // clean up
       delete [] matrix;
@@ -297,7 +321,7 @@ void Cowichan::main (int argc, char* argv[], bool use_randmat, bool use_thresh)
       end = get_ticks ();
       life (matrixIn, matrixOut);
       timeInfo(&start, &end, LIFE);
-      print_rect_matrix<bool> (matrixOut);
+      print_bool_rect_matrix (matrixOut);
 
       // clean up
       delete [] matrixIn;
[…]

### 2009-08-11 cowichan/cowichan.hpp

@@ -70,11 +70,10 @@
  * This module simulates the evolution of Conway's Game of Life, a
  * two-dimensional cellular automaton.
  * At each time step, this module must count the number of live (true)
- * neighbors of each cell, using both orthogonal and diagonal connectivity and
- * circular boundary conditions. The update rule is simple: if a cell has 3
- * live neighbors, or has 2 live neighbors and is already alive, it is alive in
- * the next generation. In any other situation, the cell becomes, or stays,
- * dead.
+ * neighbors of each cell, using both orthogonal and diagonal connectivity. The
+ * update rule is simple: if a cell has 3 live neighbors, or has 2 live
+ * neighbors and is already alive, it is alive in the next generation. In any
+ * other situation, the cell becomes, or stays, dead.
  * \see Cowichan::life
  *
  * \subsection winnow_sec 7. Weighted Point Selection
@@ -844,15 +843,20 @@ public:
       for (c = 0; c < nc; c++) {
         std::cout << MATRIX_RECT(matrix, r, c) << "\t";
       }
-      std::cout << "\n";
+      std::cout << std::endl;
     }
-    std::cout << "\n";
+    std::cout << std::endl;
   }
 #else
   template <typename T>
   void print_rect_matrix(T* /* matrix */) { }
 #endif
 
+  /**
+   * DEBUGGING FUNCTION: Print a rectangular boolean matrix on std::cout.
+   * \param matrix matrix to print.
+   */
+  void print_bool_rect_matrix(BoolMatrix matrix);
 
   /**
    * DEBUGGING FUNCTION: Print a square matrix on std::cout.
@@ -868,9 +872,9 @@ public:
       for (c = 0; c < n; c++) {
         std::cout << MATRIX_SQUARE(matrix, r, c) << "\t";
       }
-      std::cout << "\n";
+      std::cout << std::endl;
     }
-    std::cout << "\n";
+    std::cout << std::endl;
   }
 #else
   template <typename T>
@@ -888,9 +892,9 @@ public:
     index_t r;
 
     for (r = 0; r < n; r++) {
-      std::cout << VECTOR(vector, r) << "\n";
+      std::cout << VECTOR(vector, r) << std::endl;
     }
-    std::cout << "\n";
+    std::cout << std::endl;
   }
 #else
   template <typename T>

### 2009-08-11 cowichan_openmp/invperc.cpp

@@ -77,7 +77,7 @@ void CowichanOpenMP::invperc(IntMatrix matrix, BoolMatrix mask) {
   // "seed" with the middle value; start a priority queue.
   std::vector<PercPoint> points;
 
-  PercPoint initialPoint(Point((real) (nr / 2), (real) (nc / 2)));
+  PercPoint initialPoint(Point((real) (nc / 2), (real) (nr / 2)));
   initialPoint.matrix = matrix;
   initialPoint.nc = nc;
   points.push_back(initialPoint);

### 2009-08-11 cowichan_openmp/thresh.cpp

@@ -6,6 +6,9 @@
 
 #include "cowichan_openmp.hpp"
 
+/**
+ * Works only on positive input.
+ */
 void CowichanOpenMP::thresh(IntMatrix matrix, BoolMatrix mask) {
 
   index_t* hist = NULL; // histogram

### 2009-08-11 cowichan_serial/invperc.cpp

@@ -76,7 +76,7 @@ void CowichanSerial::invperc(IntMatrix matrix, BoolMatrix mask) {
   // "seed" with the middle value; start a priority queue.
   std::vector<PercPoint> points;
 
-  PercPoint initialPoint(Point((real) (nr / 2), (real) (nc / 2)));
+  PercPoint initialPoint(Point((real) (nc / 2), (real) (nr / 2)));
   initialPoint.matrix = matrix;
   initialPoint.nc = nc;
   points.push_back(initialPoint);

### 2009-08-11 cowichan_serial/thresh.cpp

@@ -6,6 +6,9 @@
 
 #include "cowichan_serial.hpp"
 
+/**
+ * Works only on positive input.
+ */
 void CowichanSerial::thresh(IntMatrix matrix, BoolMatrix mask) {
 
   index_t* hist = NULL; // histogram

### 2009-08-11 cowichan_tbb/invperc.cpp

@@ -77,7 +77,7 @@ void CowichanTBB::invperc(IntMatrix matrix, BoolMatrix mask) {
   // "seed" with the middle value; start a priority queue.
   std::vector<PercPoint> points;
 
-  PercPoint initialPoint(Point((real) (nr / 2), (real) (nc / 2)));
+  PercPoint initialPoint(Point((real) (nc / 2), (real) (nr / 2)));
   initialPoint.matrix = matrix;
   initialPoint.nc = nc;
   points.push_back(initialPoint);

### 2009-08-11 cowichan_tbb/thresh.cpp

@@ -260,6 +260,9 @@ public:
 
 /*****************************************************************************/
 
+/**
+ * Works only on positive input.
+ */
 void CowichanTBB::thresh(IntMatrix matrix, BoolMatrix mask) {
   
   // get the maximum value in the matrix (need 0-that number of bins)

### 2009-08-20 paper/2009-09/paper.tex
Starting a new version of the Cowichan paper

\documentstyle[epsf,11pt]{article}

\begin{document}

\begin{Large}\begin{center}
The Cowichan Problems: \\
Assessing the Usability of Parallel Programming Systems
\end{center}\end{Large}

\vspace{\baselineskip}

\begin{center}
\begin{tabular}{cccc}
Gregory V.~Wilson\footnote{\verb`http://www.third-bit.com`}
&
Andrew Borzenko
&
Cameron Gorrie
&
Zachary Kincaid
\end{tabular}
\end{center}

\vspace{\baselineskip}

\begin{abstract}
Most programmers believe that parallel programming is harder than sequential programming.
Many also believe that today's parallel programming systems make it even harder.
We have developed a suite of chained toy problems that can be used to assess
the usability of different scientific programming languages, parallel or otherwise.
This paper motivates and describes the suite,
and discusses what we have learned from porting it to four parallel C++ systems
(MPI, OpenMP, Intel Thread Building Blocks, and a Linux tuple space system),
and from translating it into sequential and parallel Haskell.
\end{abstract}

\section{Introduction\label{s:intro}}

Twenty years ago,
Bal {\em{et~al.}\/} listed more than 300 parallel programming systems \cite{b:par-lang-survey}.
A similar survey today could well uncover twice as many,
but a poll of application programmers would probably find that
only a handful are in general use.

The reasons why so many of these PPSs are only used by their developers (if at all) are murkier.
Part of the blame undoubtedly lies with the diversity of hardware,
and the attendant lack of portability,
alluded to above.
A second reason is that an academic researcher's principal interest is developing new languages,
not supporting existing ones.
By the time a system is 90\% complete,
its author is already thinking about its successor.
Few application programmers are willing to invest effort porting codes
to systems which are not properly supported,
and liable to change or evaporate without notice.
Third,
many research groups have chosen to explore the potential of parallelism
without consideration for compatibility with older languages.
Application programmers must therefore face the prospect of completely re-writing their programs,
rather than simply porting them.
As the next section argues,
this can obviate any performance that might be gained from parallelization.
Finally,
many PPSs simply cannot support the software engineering load of large applications.
Many provide only a single global namespace,
do not include I/O capabilities,
require users to marshal and unmarshal data explicitly,
or do not permit hardware to be timeshared.

The net result is that many PPSs are never used for anything more complicated than
generating the Mandelbrot Set,
simulating the Game of Life,
multiplying dense matrices,
or finding all solutions to the $N$-queens problem for some small $N$.
Lacking the feedback that comes from more demanding use,
few systems are improvements over their predecessors
in the way that successive versions of operating systems have improved on one another,
or the languages in the Algol$\rightarrow$Modula-$n$ family
have become both more powerful and more secure.
Most applications programmers therefore choose to play it safe
and stick to parallel programming's lowest common denominators:
data-parallel Fortran,
and message-passing libraries.

The historical reasons for tolerating this situation are becoming less valid with each passing year.
When only a few parallel computers existed,
and these were of widely different types,
software standardization simply didn't make sense.
Similarly,
when supercomputers were very expensive,
the cost of software development seemed relatively smaller,
and the requirement for maximum performance relatively greater,
than in sequential computing.
As \cite{b:social-limits-speed} points out,
however, these arguments are losing force.
Now,
as in sequential computing,
the first question is not ``How do I write programs for this machine?'',
but ``Will this machine run my programs?''
As microprocessor-based MIMD hardware comes to dominate massively-parallel computing,
we see increasing emphasis on standards such as High-Performance Fortran (HPF) \cite{b:hpf-overview},
and the Message-Passing Interface (MPI) \cite{b:mpi-overview}.

But herein lies a danger.
History shows that it is extremely difficult
to displace a programming language or system once it becomes established.
Cobol and MS-DOS are the most frequently cited examples,
but many systems which should have been superceded long ago have managed
not only to cling to life,
but to grow in influence.
(The fact that the mail signature
``The last good thing written in C was Schubert's Ninth Symphony''
can elicit smiles simply proves that even the truth can be funny.)
Established systems evolve slowly, if at all.
It would therefore seem that the best time to guarantee ourselves
a good parallel programming paradigm is right now,
before a large number of large applications are created and take on lives of their own.

We therefore propose to use a set of simple programming problems to compare
the usability of parallel programming systems.
Competent programmers, fluent in a specific system,
will implement solutions to these problems and report their experiences
in terms of development time, code size and clarity, and runtime efficiency.
These problems are on the scale of the toy applications described several paragraphs ago
both to keep implementation time low
and because we feel that any PPS produced today ought to be able to support such simple programs.
As the second author put it,
``We're all sick of seeing these,
but maybe if we all catch the disease at the same time we'll finally build up some immunity to it.''
One significant innovation in this work is that
we also require implementations of our suite to chain problems together,
so that the output of one problem is the input to another.
This will allow us to assess how well each PPS supports modularization and code re-use
(or, more realistically, how much extra effort programmers must invest
to achieve these goals when using a particular PPS).

Section~\ref{s:corollary} motivates this work by showing
how poor usability can limit the realizable parallelism in a program.
Section~\ref{s:previous} surveys previous work on the comparison of parallel programming systems,
while Section~\ref{s:method} outlines our choice of methods, and the constraints on our work.
Section~\ref{s:toys} then presents the suite of problems we intend to use.
In recognition of the original work done by Feo and others on
the Salishan Problems~\cite{b:salishan}, this suite is called
the Cowichan Problems\footnote{\noindent Pronounced {\em{Cow}\/}-i-chun.
	Like Salishan, the word is a Northwest Coast Indian place name.}.

\subsection*{History and Acknowledgments}

The first author originally developed this suite in the mid-1990s,
inspired both by \cite{b:salishan} and by
discussions with R.\ Bruce Irvin (then at the University of Wisconsin),
Henri Bal (Vrije Universiteit, Amsterdam),
and others.
Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.

\section{Quantifying the Importance of Usability\label{s:corollary}}

According to Amdahl's Law,
the time $t$ required to execute a program has two components:
$t_{seq}$, representing sequential operations that cannot be parallelized,
and $t_{par}$, representing those operations which can.
If we let $\sigma$ be the ``serial fraction'' of the program,
so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
then the speedup which can be achieved using $P$ processors is:
\begin{eqnarray*}
s(P)	& =	& \frac{t(1)}{t(P)}	\\
	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
\end{eqnarray*}
As $P{\rightarrow}{\infty}$,
the potential speedup is bounded by $1/{\sigma}$.

What is not usually acknowledged is
the influence of programming system usability on
the potential speedup of sequential programs.
As stated above, Amdahl's Law applies only to a single run of a program.
In the real world, programs are first developed, and then run many times.
If we let $T$ be the total running time of a program over its lifetime,
$D_{seq}$ be the time required to develop the sequential components of a program,
and $D_{par}$ be the time required to develop the parallel components
(or parallelize the sequential components),
then the lifetime speedup we can achieve is given by:
\begin{eqnarray*}
S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
\end{eqnarray*}
Assume that we are parallelizing a ``dusty deck'',
i.e.\ that the time $D_{seq}$ has already been invested.
If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
then achievable speedup is limited by:
\begin{eqnarray*}
S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
		& =	& \frac{1}{{\phi} + {\sigma}}
\end{eqnarray*}
The time required to parallelize a program can therefore be seen
as increasing that program's effective serial fraction.
Unless parallelization time can be substantially reduced,
massively-parallel computing is likely to remain attractive only in those cases in which:
\begin{itemize}
\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
\item	the expected total program runtime is very large
	(i.e.\ the program is a package used extensively in a field
	such as computational chemistry or vehicle crash simulation);
	or
\item	cost is not an obstacle (i.e.\ the customer is the NSA).
\end{itemize}

\section{Previous Work\label{s:previous}}

In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
\begin{quote}
As practiced by computer science,
the study of programming is an unholy mixture of mathematics,
literary criticism,
and folklore.
\end{quote}
Little has changed since.
Computer scientists pay lip service to the importance of usability,
but are strangely reluctant to try to quantify it,
particularly with respect to the tools they use themselves.
Religious debates about the ``right'' first-year teaching language
or the ``best'' GUI toolkit are common;
systematic use of the experimental method is not.

Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
such as readability and code-pattern recognition
\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
The little work that has been done on ``programming in the large'' has been anecdotal,
such as \cite{b:compare-ada-c-pascal}.
This book presented various authors' impressions of Pascal, C, and Ada
(before any complete Ada implementations existed).
Its editors formulated a list of questions to use in comparing languages,
but did not use it themselves.
A modified version of this questionnaire,
tailored for parallel programming languages,
is given in Appendix~\ref{s:questions},
and will be used as a guideline in our work.

Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
The first presented implementations of a simple numerical quadrature program
in more than a dozen different parallel languages
in use on mid-1980s hardware.
The second presented implementations of
the Salishan Problems---Hamming number generation,
isomer enumeration,
skyline matrix reduction,
and a simple discrete event simulator---in
C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
and a variety of dataflow and functional languages.
Both of these books convey the flavor of the languages they describe,
but neither made any effort to compare languages or problem implementations directly.

Inspired by the Salishan Problems,
the first author developed a suite of seven problems \cite{b:cowichan-1}
in order to test the limits of parallel programming systems.
Individual problems were badly load-balanced,
manipulated irregular pointer-based structures,
and so on.
Our experience implementing them was extremely valuable,
as it uncovered many bugs in the programming system used,
and revealed several shortcomings of the system's design.
However,
we also discovered that implementing each problem takes six to eight weeks,
which we feel is prohibitive at this exploratory stage.

\section{Methodology\label{s:method}}

The main aim of this work is to assess and compare
the ease with which programs can be developed and/or ported
using different parallel programming systems.
Here,
we discuss issues related to choosing the problems that make up the suite,
assessing the complexity of particular implementations,
and other purposes which implementations of these might serve.
Recognizing that these problems are extremely simple,
and easy to parallelize,
we refer to them as ``toys''.

\subsection{Criteria for Selection\label{s:method-criteria}}

Our criteria for including problems in this suite are as follows:
\begin{enumerate}
\item	Each toy should require no more than an afternoon
	to write and test in a well-supported sequential language.
	We feel that making individual problems more complicated will only discourage implementation.
\item	The correctness of each implementation must be relatively easy to verify.
	Toys whose output is easily visualized are therefore to be preferred,
	as are toys whose results are insensitive to floating-point discretization effects.
\item	Each toy application should be amenable to
	an analysis of speedup or isoefficiency \cite{b:isoefficiency-tutorial}
	to allow a comparison of theory and practice.
\item	At least some toys should not be ``infinitely scalable''.
	Many real-world applications are not,
	and this suite should reflect such limitations.
\item	At least some toys should require I/O,
	since this important aspect of real-world programming is often neglected by PPS designers.
\item	There should be some potential redundancy in the type of parallelization done,
	so that re-use of software can be demonstrated.
\item	Together,
	the toys in the suite must exercise a wide range of common parallel operations
	(Appendix~\ref{s:cliche-ops})
	and memory reference patterns
	(Appendix~\ref{s:cliche-memref}).
\item	The ``obvious'' implementations of these toys should span
	a wide range of parallel programming paradigms,
	such as task farming,
	geometric and functional decomposition,
	and speculative parallelism.
\item	Parallelizations for several (if not all) of
	the types of programming systems shown in Figure~\ref{f:taxonomy} should be known.
	This suite is not intended to be a collection of research problems;
	reasonably simple and efficient implementations in most environments
	should be straightforward to produce.
\end{enumerate}

\begin{figure}
\epsfysize=0.5\textheight
\begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
\caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
\end{figure}

\subsection{Specification\label{s:method-spec}}

Toys should be specified in terms of inputs and outputs,
rather than algorithmically.
For example,
``sort $N$ integers in the range $1{\ldots}R$,
where $R$ is not known in advance and there may be multiple instances of a given value''
is to be preferred over
``parallelize quicksort''.
As \cite{b:slalom} argued,
which parallel algorithm is best in a particular programming system
depends critically on the architectural assumptions made by that system.
Implementors should be free to employ those methods which they find natural and convenient.
The only exception we make to this rule is in the two matrix equation solvers,
where we require Gaussian elimination and successive over-relaxation.
Given the enormous number of matrix solution algorithms in use,
we felt we had to specify these particular algorithms in order to ensure that
different implementations of this suite would be comparable.

\subsection{Software Engineering Issues\label{s:method-softeng}}

These toys will exercise many different aspects of parallel systems.
However,
their ``single algorithm per program'' model is not representative of real applications.
These usually contain several discrete (and sometimes overlapping) phases,
each of which is qualitatively different.
A full implementation of this suite will therefore have two parts.
In the first,
each toy will be implemented as a stand-alone program.
In the second,
toys will be chained together to create the kind of phase behaviour seen in real applications
(Figure~\ref{f:chaining}).
This will test the ease with which heterogeneous parallelism can be mixed within a single program.
It will also show how well the system supports code re-use and information hiding,
which are crucial to the development of large programs.

\begin{figure}
\epsfxsize=12cm
\begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
\caption{Chaining\label{f:chaining}}
\end{figure}

A second reason for chaining toys is that
many groups are now using algorithmic skeletons to structure parallel programs
\cite{b:cole-skel,b:p3l-overview,b:enterprise}.
Such skeletons encapsulate the details of a particular style of parallelism,
and hide low-level or platform-dependent issues.
We hope that chaining will show the strengths and weaknesses of such systems.

Finally,
chaining should be designed so that some toys can be executed concurrently.
Many PPSs impose extraneous constraints on programs,
e.g.\ require all processes to participate in every barrier synchronization,
or require the same executable to be loaded onto each processor.
These constraints can limit the exploitation of potential concurrency.
Permitting, but not requiring, concurrent execution of several toys should uncover such limitations.

\subsection{Sizing\label{s:method-size}}

One crucial aspect of the specification of toys is
the way in which the sizes of problems are determined.
In a completely frozen model, the actual size of each problem would be compiled into each toy.
A fluid implementation, by contrast, would allow sizes to be specified at run-time,
and would only then allocate and partition data structures.
We have chosen to use an intermediate model (which might be called ``slushy''),
in which the maximum size of individual problems is specified during compilation,
but the actual size of a problem is only determined when the toy begins to execute.
We hope that the use of this model will force implementors to deal with
at least some of the software engineering issues
involved in building flexible large-scale applications,
without making implementation prohibitively difficult.

\subsection{Assessing Complexity\label{s:method-complex}}

There are at least two different ways in which usability might be compared.
One would be to measure the performance achieved by an ``average'' programmer
as a function of time on each of a range of problems.
This would lead to graphs of the kind shown in Figure~\ref{f:perf},
where the left graph shows a particular result
and the right graph a simplified abstraction of it.
By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
which is the length of vector on which a pipelined architecture achieves
half its theoretical peak performance,
we could in principle find the value of $p_{1/2}$---the programming time required to achieve
half of a machine's peak performance\footnote{Note that
	if performance was measured as a fraction of the figures quoted by manufacturers for their machines,
	it is unlikely that the halfway mark would ever be reached.}
for a particular combination of programming system and problem type.

\begin{figure}
\epsfxsize=12cm
\begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
\caption{Performance vs.\ Time\label{f:perf}}
\end{figure}

A second option would be to measure the complexities of
implementations of a single application using different programming systems.
With this approach,
comparative performance or speedup figures would be supplemented by complexity measures,
as shown in Figure~\ref{f:complex}.

\begin{figure}
\epsfxsize=6cm
\begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
\caption{Comparative Code Complexity\label{f:complex}}
\end{figure}

The drawback of the first approach is that there are no ``typical'' problems,
or ``average'' programmers.
Meaningful measurement of  $p_{1/2}$ would therefore require
measurement of a large number of programmers working on many problems.
Since this is too expensive to be practical,
the second approach is the one we have chosen to adopt.
It has drawbacks as well,
most notably the fact that no-one has ever demonstrated any better measure of program complexity
than a simple count of the number of lines in a program \cite{b:fenton-metrics}.
Since we anticipate a wide variation in the types of progamming systems that will be used in this work
(ranging from declarative dataflow languages
through functional languages with Lisp-like syntax
to parallel dialects of Fortran and C),
and in the indentation, commenting, and layout styles of particular programmers,
we may measure complexity by counting the number of tokens in the source code which are
keywords,
intrinsic operators,
user-defined functions,
type or variable declarations,
and miscellaneous punctuation.

\subsection{Other Uses for Implementations\label{s:method-uses}}

This suite is intended for assessing PPS usability,
but we envisage at least three other uses for implementations of it:
\begin{description}
\item[Related Research:]
	We hope that the existence of this suite will indicate
	to the developers of new parallel programming systems
	what a mature tool should be able to support.
	If sufficiently diverse,
	the suite could also be used to help test the correctness of new or improved systems.
	Finally,
	this suite could complement existing performance benchmarks,
	since performance figures for individual toys could be compared.
\item[Tool Development:]
	Requests from tool developers for existing parallel programs,
	or for event traces from them,
	are common on newsgroups such as {\tt{comp.parallel}}.
	We hope that traces from implementations of this suite
	will be used to compare different debugging and performance monitoring tools,
	that their source code will be used to test
	new parallelizing compilation techniques or mapping and scheduling algorithms,
	and so on.
\item[Teaching:]
	Implementations will be suitable for use as classroom examples,
	since they will be small enough to be understood quickly.
	The toys themselves should also be suitable as classroom exercises
	in a senior undergraduate course on parallel computing.
\end{description}

\subsection{Some Criticisms\label{s:method-criticism}}

\noindent {\em{Why not use an existing benchmark suite for this purpose?}\/}

Implementations of benchmark suites have usually been coded for absolute speed.
They are not representative of ``normal'' practice,
in which reasonable performance is the usual goal,
and so are inappropriate for this work.
In addition,
most existing benchmark suites are too large to be implemented or ported quickly
(e.g.\ SPEC~\cite{b:bench-over}),
over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
or concerned with a limited range of applications
(e.g.\ the NAS Parallel Benchmarks~\cite{b:nas-results}).
Finally,
most exist only in C or Fortran (or both);
many interesting parallel programming systems are built on top of other languages,
including Modula-2, Scheme, Prolog, and various dataflow languages.
We do not want the effort required to translate several large programs
to discourage researchers from participating in this exercise.

\vspace{\baselineskip}

\noindent {\em{Won't the results depend primarily on programmer ability?}\/}

The short answer to this question is yes,
but no more than the results from performance benchmarks.
A longer answer is that	since we intend to measure the combination of code complexity and performance achieved,
rather than performance as a function of development time,
programmers will be allowed to revise and improve their implementations when and as desired.

\section{The Cowichan Problems\label{s:toys}}

This section describes the problems making up the full problem suite.
As discussed above,
the implementation of each is to be used both as a stand-alone program
and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.

\begin{table}
\begin{center}
\begin{tabular}{cl|cl}
Code	& Toy				& Code	& Toy				\\
\hline
E	& elastic net			& O	& outer product			\\
G	& Gaussian elimination		& P	& matrix-vector product		\\
H	& halving shuffle		& R	& random matrix generation	\\
I	& invasion percolation		& S	& successive over-relaxation	\\
L	& Game of Life			& T	& image thresholding		\\
M	& Mandelbrot Set		& V	& vector difference		\\
N	& point normalization		& W	& point value winnowing
\end{tabular}
\caption{Single-Character Codes for Toys\label{t:toy-codes}}
\end{center}
\end{table}

\subsection{{\tt{mandel}}:
	Mandelbrot Set Generation
	\label{s:toys-mandel}}

\begin{figure}
\begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
\caption{The Mandelbrot Set\label{f:mandel}}
\end{figure}

This module generates the Mandelbrot Set
for a specified region of the complex plane
(Figure~\ref{f:mandel}).
The inputs to this module are:
\begin{description}
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{x0, y0}}:]
	the real coordinates of the lower-left corner
	of the region to be generated.
\item[{\tt{dx, dy}}:]
	the extent of the region to be generated.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix containing the iteration count at each point in the region.
\end{description}

Given initial coordinates $(x_0, y_0)$,
the Mandelbrot Set is generated by iterating the equation
\begin{eqnarray*}
x^{\prime}	& =	& x^2 - y^2 + y_0	\\
y^{\prime}	& =	& 2{x}{y} + x_0
\end{eqnarray*}
until either an iteration limit is reached,
or the values diverge.
The iteration limit used in this module is 150 steps;
divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
The integer value of each element of the matrix is
the number of iterations done.
Note that,
as in all problems,
the output is required to be
independent of the number of processors used.
This requires programmers to be careful about
floating-point roundoff effects.

\subsection{{\tt{randmat}}:
	Random Number Generation
	\label{s:toys-rng}}

This module fills a matrix with pseudo-random integers.
The inputs to this module are:
\begin{description}
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{s}}:]
	the random number generation seed.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix filled with random values.
\end{description}
Note that,
as in all problems,
the output is required to be independent of the number of processors used.
Generating new seed values from the given seed,
and running one copy of the random number generator
on each processor,
is therefore unlikely to qualify as a solution.

\subsection{{\tt{half}}:
	Two-Dimensional Shuffle
	\label{s:toys-half}}

This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
shuffles them,
and then repeats this operation along the other axis.
Values in odd-numbered locations are collected at the low end of each row or column,
while values in even-numbered locations are moved to the high end.
An example transformation is:

\begin{eqnarray*}
\begin{array}{cccc}
a & b & c & d \\
e & f & g & h \\
i & j & k & l
\end{array}
& = &
\begin{array}{cccc}
a & c & b & d \\
i & k & j & l \\
e & g & f & h
\end{array}
\end{eqnarray*}

Note that how an array element is moved depends only on whether its location index is odd or even,
not on whether its value is odd or even.

The inputs to this module are:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix containing shuffled values.
\end{description}

\subsection{{\tt{invperc}}:
	Invasion Percolation
	\label{s:toys-invperc}}

Invasion percolation models the displacement of one fluid
(such as oil)
by another
(such as water)
in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
In two dimensions,
this can be simulated by generating
an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
and then marking the center cell of the grid as filled.
In each iteration,
one examines the four orthogonal neighbors of all filled cells,
chooses the one with the lowest value
(i.e.\ the one with the least resistance to filling),
and fills it in.
Figure~\ref{f:invperc1} shows the first few steps in this process,
while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.

\begin{figure}
\begin{footnotesize}
\begin{math}
\begin{array}{ccc}
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & 29 & \circ & 29 & 77 \\
\hline
61 & 26 & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \star & \circ & 29 & 77 \\
\hline
61 & 26 & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & 29 & 77 \\
\hline
61 & \star & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
\\
 & & \\
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \star & 77 \\
\hline
61 & \circ & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & \star & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & \circ & \star \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
\\
 & & \\
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & \star & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
\star & \circ & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & \circ & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & \star & 72 & 45 & 38 \\
\hline
\circ & \circ & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & \circ & \circ \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
\end{array}
\end{math}
\end{footnotesize}
\caption{Invasion Percolation\label{f:invperc1}}
\end{figure}

\begin{figure}
\begin{center}
{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
\end{center}
\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
\end{figure}

The simulation continues until
some fixed percentage of cells have been filled,
or until some other condition
(such as the presence of trapped regions)
is achieved.
The fractal structure of the filled and unfilled regions
is then examined to determine how much oil could be recovered.
The na\"{\i}ve way to implement this is to repeatedly scan the array;
a more sophisticated, and much faster, sequential technique
is to maintain a priority queue of unfilled cells
which are neighbors of filled cells.
This latter technique is similar to the list-based methods used in some cellular automaton programs,
and is very difficult to parallelize effectively.

The inputs to this module are:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix and mask.
\item[{\tt{nfill}}:]
	the number of points to fill.
\end{description}
Its output is:
\begin{description}
\item[{\tt{mask}}:]
	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
\end{description}
Filling begins at the central cell of the matrix
(rounding down for even-sized axes).

\subsection{{\tt{thresh}}:
	Histogram Thresholding
	\label{s:toys-thresh}}

This module performs histogram thresholding on an image.
Given an integer image $I$ and a target percentage $p$,
it constructs a binary image $B$
such that $B_{i,j}$ is set
if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
The general idea is that an image's histogram should have 2 peaks,
one centered around the average foreground intensity,
and one centered around the average background intensity.
This program attempts to set
a threshold between the two peaks in the histogram
and select the pixels above the threshold.
This module's inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the integer matrix to be thresholded.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix and mask.
\item[{\tt{percent}}:]
	the minimum percentage of cells to retain.
\end{description}
Its output is:
\begin{description}
\item[{\tt{mask}}:]
	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
\end{description}

\subsection{{\tt{life}}:
	Game of Life
	\label{s:toys-life}}

\begin{figure}
\begin{center}
{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
\end{center}
\caption{Snapshots from the Game of Life\label{f:life}}
\end{figure}

This module simulates the evolution of Conway's Game of Life,
a two-dimensional cellular automaton (Figure~\ref{f:life}).
Its inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	a Boolean matrix representing the Life world.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{numgen}}:]
	the number of generations to simulate.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	a Boolean matrix representing the world after simulation.
\end{description}
At each time step,
this module must count the number of live ({\tt{true}}) neighbors
of each cell,
using both orthogonal and diagonal connectivity
and circular boundary conditions.
The update rule is simple:
if a cell has 3 live neighbors,
or has 2 live neighbors and is already alive,
it is alive in the next generation.
In any other situation,
the cell becomes, or stays, dead.

\subsection{{\tt{winnow}}:
	Weighted Point Selection
	\label{s:toys-winnow}}

This module converts a matrix of integer values
to a vector of points,
represented as $x$ and $y$ coordinates.
Its inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix, whose values are used as masses.
\item[{\tt{mask}}:]
	a Boolean matrix showing which points are eligible for consideration.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{nelts}}:]
	the number of points to select.
\end{description}
Its output is:
\begin{description}
\item[{\tt{points}}:]
	a vector of $(x,y)$ points.
\end{description}
Each location where {\tt{mask}} is {\tt{true}}
becomes a candidate point,
with a weight equal to the integer value in {\tt{matrix}} at that location
and $x$ and $y$ coordinates
equal to its row and column indices.
These candidate points are then sorted into increasing order by weight,
and {\tt{nelts}} evenly-spaced points selected to create the result vector.

\subsection{{\tt{norm}}:
	Point Location Normalization
	\label{s:toys-norm}}

This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
Its inputs are:
\begin{description}
\item[{\tt{points}}:]
	a vector of point locations.
\item[{\tt{nelts}}:]
	the number of elements in the point vector.
\end{description}
Its output is:
\begin{description}
\item[{\tt{points}}:]
	a vector of normalized point locations.
\end{description}
If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
then the normalization equation is:
\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
$y$ coordinates are normalized in the same fashion.

\subsection{{\tt{elastic}}:
	Elastic Net Simulation
	\label{s:toys-elastic}}

This module uses the elastic net algorithm \cite{b:elastic-net} to find an approximate solution to the travelling salesentity problem (TSP).
A closed, circular loop, modelled by a large number of doubly-connected points, is placed in the plane.
This path is slowly deformed until it connects the points representing the cities to be visited (Figure~\ref{f:elastic}).
The rate at which the loop is deformed determines the quality of the final path.

\begin{figure}
\begin{center}
\framebox{\epsffile{fig/elastic1.ps}}
\framebox{\epsffile{fig/elastic2.ps}}
\framebox{\epsffile{fig/elastic3.ps}}
\end{center}
\caption{Relaxation of an Elastic Net\label{f:elastic}}
\end{figure}

Let the (fixed) location of city $i$ be $c_i$, and the (changeable) location of point $j$ on the elastic net be $p_j$.
The change in the location of point $j$ during each iteration of the algorithm is given by:
\begin{equation}
{\delta}{p_j} = {\alpha}\sum_{i}{w_{ij}(c_i - p_j)} ~+~ K{\beta}(p_{j+1} - 2p_j + p_{j-1})
\end{equation}
$\alpha$ and $\beta$ are time-invariant factors relating the strengths of the city-to-point and point-to-point forces,
and $K$ is a time-dependent factor which controls the rate of annealing.
The strength of the force $w_{ij}$ between city $c_i$ and point $p_j$ is given by:
\begin{equation}
w_{ij} = {\phi}({\mid}c_i - p_j{\mid}, K) / \sum_{\ell}\phi({\mid}c_i - p_{\ell}{\mid}, K)
\end{equation}
where $\phi(d, K)$ is the Gaussian function $e^{-{d^2}/{K^2}}$.
Note the normalization of each $w_{ij}$; in order for this algorithm to work properly, the total force being exerted by each city must be the same.
This is accomplished by summing the forces exerted by each city, and then normalizing the city's forces by that total.

The inputs to this module are:
\begin{description}
\item[{\tt{cities}}:]
	a vector of points representing the locations of the cities to be toured.
	The $(x,y)$ coordinates of these cities must be scaled to lie in the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
\item[{\tt{n\_cities}}:]
	the number of cities in {\tt{cities}}.
\item[{\tt{n\_net}}:]
	the number of elastic net points to use.
\item[{\tt{n\_iters}}:]
	the total number of iterations to perform.
\item[{\tt{n\_relax}}:]
	the number of iterations to perform with each discrete value of $K$.
	After each {\tt{n\_relax}} iterations, $K$ is reduced by a (constant) pre-defined scaling factor.
\end{description}
Its output is:
\begin{description}
\item[{\tt{net}}:]
	a vector of {\tt{n\_net}} points representing the tour found.
\end{description}

\subsection{{\tt{outer}}:
	Outer Product
	\label{s:toys-outer}}

This module turns a vector containing point positions
into a dense, symmetric, diagonally dominant matrix
by calculating the distances between each pair of points.
It also constructs a real vector whose values are
the distance of each point from the origin.
Inputs are:
\begin{description}
\item[{\tt{points}}:]
	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
\item[{\tt{nelts}}:]
	the number of points in the vector, and the size of the matrix along each axis.
\end{description}
Its outputs are:
\begin{description}
\item[{\tt{matrix}}:]
	a real matrix, whose values are filled with inter-point distances.
\item[{\tt{vector}}:]
	a real vector, whose values are filled with origin-to-point distances.
\end{description}
Each matrix element $M_{i,j}$ such that $i \neq j$
is given the value $d_{i, j}$,
the Euclidean distance between point $i$ and point $j$.
The diagonal values $M_{i, i}$ are then set to
{\tt{nelts}} times the maximum off-diagonal value
to ensure that the matrix is diagonally dominant.
The value of the vector element $v_i$ is set to
the distance of point $i$ from the origin,
which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.

\subsection{{\tt{gauss}}:
	Gaussian Elimination
	\label{s:toys-gauss}}

This module solves a matrix equation $AX=V$ for
a dense, symmetric, diagonally dominant matrix $A$
and an arbitrary vector non-zero $V$
using explicit reduction.
(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
there is a well-formed solution to the equation.)
Inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the real matrix $A$.
\item[{\tt{target}}:]
	the real vector $V$.
\item[{\tt{nelts}}:]
	the number of values in each vector, and the size of the matrix along each axis.
\end{description}
Its output is:
\begin{description}
\item[{\tt{solution}}:]
	a real vector containing the solution $X$.
\end{description}

\subsection{{\tt{sor}}:
	Successive Over-Relaxation
	\label{s:toys-sor}}

This module solves a matrix equation $AX=V$ for
a dense, symmetric, diagonally dominant matrix $A$
and an arbitrary vector non-zero $V$
using successive over-relaxation.
Its inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the real matrix $A$.
\item[{\tt{target}}:]
	the real vector $V$.
\item[{\tt{nelts}}:]
	the number of values in each vector, and the size of the matrix along each axis.
\item[{\tt{tolerance}}:]
	the solution tolerance, e.g.\ $10^{-6}$.
\end{description}
Its output is:
\begin{description}
\item[{\tt{solution}}:]
	a real vector containing the solution $X$.
\end{description}

\subsection{{\tt{product}}:
	Matrix-Vector Product
	\label{s:toys-product}}

Given a matrix $A$,
a vector $V$,
and an assumed solution $X$ to the equation $AX=V$,
this module calculates the actual product $AX={V^{\prime}}$,
and then finds the magnitude of the error.
Inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the real matrix $A$.
\item[{\tt{actual}}:]
	the real vector $V$.
\item[{\tt{candidate}}:]
	a real vector containing the supposed solution.
\item[{\tt{nelts}}:]
	the number of values in each vector, and the size of the matrix along each axis.
\end{description}
The output of this function is:
\begin{description}
\item[{\tt{e}}:]
	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
\end{description}

\section{Other Issues\label{s:issues}}

As described in Section~\ref{s:method-softeng},
implementors are required to chain modules together in the manner shown in Figure~\ref{f:chaining}.
This section describes the flow of control and data between these chained problems,
and discusses several other important aspects of this suite,
including I/O and visualization.

\subsection{Chained Problem Sequences\label{s:issues-chain}}

The chained implementation of these problems
executes individual toys in the following order:
\begin{enumerate}
\item	An integer matrix $I$ with $r$ rows and $c$ columns is created
	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
	\label{e:chain-choice-1}
\item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
\item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
	to generate a Boolean mask $B$ from $I$ in which the (minimum) percentage of {\tt{true}} locations is $P$.
	Like $I$, $B$ has $r$ rows and $c$ columns.
	\label{e:chain-choice-2}
\item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
	This step overwrites the Boolean matrix $B$.
\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
	as described in Section~\ref{s:toys-outer}.
\item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
	In a parallel system, these two problems should if possible be executed concurrently.
\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
	In a parallel system, these two problems should if possible be executed concurrently.
\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
	This measures the agreement between the solutions found by the two methods.
\end{enumerate}
Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
there are four possible chained sequences.
A minimal implementation will provide one of these,
while a full implementation will be a single program which can execute any of the four sequences.
Intermediate implementations providing all four sequences as separate programs are also acceptable.

\subsection{Input and Output\label{s:issues-io}}

I/O is an important part of programming,
but is often treated as being of secondary importance by language designers.
This suite requires all stand-alone toys to read input values from files, and write results back;
the chained version must be able to checkpoint intermediate results between toys.
Finally,
implementors are strongly encouraged to include some means of visualizing the output or evolution of individual toys.

The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
A file containing a vector begins with 1 positive integer,
which specifies the number of elements in the vector;
a file containing a matrix begins with 2 positive integers,
which specify the number of rows and columns in the matrix respectively.
Elements of the vector or matrix then appear one per line,
from least-indexed to greatest-indexed
(and in row-major order for matrices).

The formats described above are required to be human-readable.
Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
using whatever file formats are convenient.
This will allow programmers to demonstrate the ``natural'' I/O capabilities of particular systems
which most probably be used for checkpointing intermediate results in real programs.

\begin{figure}
\epsfxsize=0.75\textwidth
\begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
\caption{Cowichan File Formats\label{f:io-formats}}
\end{figure}

\subsection{Visualization\label{s:issues-vis}}

Visualization is an increasingly important part of scientific and numerical computing.
Including it as an optional part of this problem set is also a good way to exercise both
the use of libraries within a parallel programming system,
and the ease with which a programming system can support variant versions of a program.

Our requirements for the graphics interface are:
\begin{enumerate}
\item	Visualization for individual toys should be as simple as possible,
	and should require as few changes to the basic toy code as possible.
\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
	The latter should create all images in a single window.
\item	A single source file should implement both the graphical and non-graphical versions of each toy.
\end{enumerate}
The first two points require the system to distinguish between the actual drawing code
(e.g.\ drawing a vector of points as circles)
and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
While the stand-alone toys use a whole window for drawing,
the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
At present,
this is implemented by using a call counter
in the interfaces for the individual toys and in the chained module.
The actual drawing routines called from within each toy could then be the same.

The present visualization routines were originally written by Anil Sukul,
a student at the Vrije Universiteit, Amsterdam.
The package uses the public domain graphics library VOGLE,
available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
as well as screen control routines (such as double buffering).
VOGLE was chosen because of its simplicity,
and because it includes Pascal and Fortran interfaces.

\subsection{Reproducibility\label{s:issues-reproduce}}

Reproducibility is an important issue for parallel programming systems.
While constraining the order of operations in a parallel system to guarantee reproducibility makes programs in that system easier to debug,
it can also reduce the expressiveness or performance of the system.

In this problem set, irreproducibility can appear in two forms: numerical and algorithmic.
The first arises in toys such as {\tt{gauss}}, {\tt{sor}}, and {\tt{elastic}}, which use floating-point numbers.
Precisely how round-off errors occur in these calculations can depend on the distribution of work among processors,
or the order in which those processors carry out particular operations.

Irreproducibility also arises in toys which only use exact numbers, such as {\tt{invperc}} and {\tt{randmat}}.
In the former, the percolation region is grown by repeatedly filling the lowest-valued cell on its perimeter.
If several cells have this value, implementations may choose one arbitrarily.
Thus, different implementations may produce very different fractal shapes.
In the case of random number generation, the simplest thing to do is to run the same generator independently on each processor,
although the values in the resulting matrix then depend on the number of processors used.

\section{Conclusion\label{s:conclude}}

A full ANSI~C version of this problem suite has been written, and is available by anonymous FTP from:
\begin{center}
{\tt{ftp.cs.toronto.edu :: pub/gvw/cowichan/src.tar.gz}}
\end{center}
This release includes two parallel versions,
both based on the POSIX threads package.
The first uses repeated fork-and-join parallelism,
while the second creates a fixed number of threads at the beginning of the program,
then synchronizes them repeatedly using barriers.

We would like to invite interested groups to participate in this project by parallelizing the problem suite using their favourite language or tool.
Our aim is to make both the implementations and their descriptions available as a hypertext document through the World-Wide Web,
and, if there is sufficient interest, to publish this material in book form.
If you would like to contribute to this effort, please contact either of the first two authors.

\bibliographystyle{plain}
\newcommand{\bibea}{{\em{et~al.}}}
\bibliography{paper}

\appendix

\newpage
\section{Parallel Clich\'{e}s\label{s:cliche}}

\subsection{Operations\label{s:cliche-ops}}

This list details some operations which are supported by many parallel programming systems.
The toys described in Section~\ref{s:toys} provide opportunities for exercising many of these,
and implementors are encouraged to phrase discussion of their work in terms of these operations where appropriate.

\begin{itemize}
\item	elementwise operations on arrays (unary, binary, and scalar promotion)
\item	cumulative operations on arrays (reduction and parallel prefix)
\item	array re-shaping and re-sizing (e.g.\ sub-array extraction)
\item	partial (conditionally masked) versions of all of the above
\item	regular data motion (e.g.\ circular and planar shifting)
\item	irregular data motion (e.g.\ 1-to-1, 1-to-many, and many-to-1 permutation)
\item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
\item	differential local addressing (i.e.\ subscripting an array with an array)
\item	full or partial replication of shared read-only values
\item	full or partial replication of shared read-mostly values with automatic consistency management
\item	structures with rotating ownership, suitable for migration
\item	producer-consumer structures
\item	partitionable structures
\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
\item	dynamic re-partitioning (e.g.\ for load balancing)
\item	committed mutual exclusion (e.g.\ waiting on a semaphore)
\item	uncommitted mutual exclusion (e.g.\ lock or fail)
\item	barrier synchronization
\item	multiple concurrent barriers used by non-overlapping groups of processes
\item	fetch-and-add, and other fetch-and-operate functions
\item	pre-scheduled receipt of messages of known composition
\item	variable-length message receipt
\item	message selection by type
\item	message selection by source
\item	message selection by contents
\item	guarded message selection (e.g.\ CSP's {\tt{alt}} construct)
\item	broadcast and partial broadcast
\item	split-phase (non-blocking) operations
\item	data marshalling and unmarshalling of ``flat'' structures (e.g.\ arrays of scalars)
\item	data marshalling and unmarshalling of nested or linked structures
\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
\item	pipelining
\item	distributed linked structures in systems without physically-shared memory
\item	indexing of distributed shared arrays
\item	collective I/O
\item	uniform-size interleaved I/O
\item	heterogeneous-size interleaved I/O
\item	independent I/O operations on a single shared file
\end{itemize}

\subsection{Memory Reference Patterns\label{s:cliche-memref}}

These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.

\begin{itemize}
\item	Write-once:
	variables which are assigned a value once, and only read thereafter.
	These can be supported through replication.
\item	Write-many:
	variables which are written and read many times.
	If several processes write to the variable concurrently, they typically write to different portions of it.
\item	Producer-consumer:
	variables which are written to by one or more objects, and read by one or more other objects.
	Entries in a shared queue, or a mailbox used for communication between two processes, are examples.
\item	Private:
	variables which are potentially shared, but actually private.
	The interior points of a mesh in a program which uses geometric decomposition fall into this category,
	while boundary points belong to the previous class.
\item	Migratory:
	variables which are read and written many times in succession by a single process before ownership passes to another process.
	The structures representing particles in an $N$-body simulation are the clearest example of this category.
\item	Result:
	Accumulators which are updated many times by different processes, but thereafter read.
\item	Read-mostly:
	Any variable which is read much more often than it is written to.
	The best score so far in a search problem is a typical instance of this class:
	while any process might update it, in practice processes read its value much more often than they change it.
\item	Synchronization:
	Variables used to force explicit synchronization in a program, such as semaphores and locks.
	These are typically ignored for long periods, and then the subject of intense bursts of access activity.
\item	General read-write:
	Any variable which cannot be put in one of the above categories.
\end{itemize}

\newpage
\section{Comparing Parallel Programming Systems\label{s:questions}}

This list of questions is a modified version of the one presented in \cite{b:compare-ada-c-pascal}.
It is presented here as an organizational aid;
when describing the PPS they have used,
implementors are encouraged to use the order and terminology of this list where appropriate.
We acknowledge that many of these questions will not apply to particular systems,
and encourage implementors to ask questions which we have left out.

\begin{enumerate}
\item History and Philosophy
  \begin{enumerate}
  \item For whom and for what purpose was the language designed?
  \end{enumerate}
\item Syntax
  \begin{enumerate}
  \item Model implemented (Figure~\ref{f:taxonomy})
  \item Class
    \begin{enumerate}
    \item library called from existing sequential language?
    \item syntactic extensions to existing sequential language?
    \item new language in Backus (Fortran/Algol/C) tradition?
    \item new (non-Backus) language?
    \end{enumerate}
  \item Concurrency mechanisms adhere to same philosophy as original
	language (where appropriate)?
  \item Readability?  Syntactic consistency?  Conciseness?
	(on an APL-to-Modula scale)
  \item Common errors (e.g.\ confusing {\tt{=}} and {\tt{==}} in C)?
  \end{enumerate}
\item Type Philosophy and Data Types
  \begin{enumerate}
  \item What primitive types does the language support?
  \item How are new types created?
  \item What intrinsic operations are defined on user-defined types?
  \item How strict is typing?
	How is type conversion done?
  \item Can type information be inspected at run-time?
  \item What concurrent operations on aggregates
	(such as lists and arrays) are intrinsic?
	What kinds of concurrent aggregate operations can users define?
  \item How are aggregates defined (e.g.\ by index set, by shape)?
	Can new kinds of aggregates (e.g.\ templates for arrays)
	be defined?
  \item How strict is typing during communication?
  \item What support is there for communication or manipulation
	of data of non-scalar or non-intrinsic types?
  \item Can structural information about aggregates
	(e.g.\ number of elements in a set, dimension of an array)
	be inspected at run-time?
  \item What built-in support is there for object classes and inheritance?
	For polymorphism?
	For class (as opposed to instance) data?
  \item Are functions first-class?
  \item Are continuations supported?  Are they first-class?
  \item Is dynamically-allocated memory supported?
	Is it done automatically?
	Is garbage-collection done automatically?
	Can a program ever create a dangling pointer?
	An alias?
  \end{enumerate}
\item Operators
  \begin{enumerate}
  \item What intrinsic operators are provided (brief list)?
	How strongly do these type-check their arguments?
  \item How are new operators defined?
  \item Are the precedence rules straightforward?
  \item Are side-effects/mutation allowed inside non-call expressions
	(e.g.\ C's side-effecting \verb`i++` construct)?
  \item Are their results platform-independent?
  \item Are the results of intrinsic parallel operations reproducible?
	Are they platform-independent?
  \item Are heterogeneous operations on aggregates allowed
	(e.g.\ every element of a list of polygons calculates its
	center of mass)?
	If so, how are these described, and how are they implemented?
  \item Which of the operators in Section~\ref{s:cliche-ops} are supported?
  \end{enumerate}
\item Control Flow
  \begin{enumerate}
  \item What synchronization is automatically imposed during execution
	(i.e.\ at what points can definite assertions about the state
	of the running program be made)?
  \item What control constructs does the language support?  Encourage?
  \item How is concurrency expressed?  How structured is this mechanism?
  \item Can the degree of concurrency be throttled (limited)?
	Is this done by the system, or by the user?
  \item How are exceptions handled/reported?
	Is it different for intrinsic operations and user-defined operations?
  \end{enumerate}
\item Subroutines, Scope, and Modularization
  \begin{enumerate}
  \item How efficient is subroutine call compared to in-line execution
	(i.e.\ how expensive are parameter passing and non-local control flow)?
  \item How much information about parameters is available (required)
	inside called subroutine
	(e.g.\ whether object is shared with other processes or private
	to executing process)?
  \item Are side-effecting procedures allowed?
  \item Can generic routines be written
	(e.g.\ to work on different types or array shapes)?
  \item Is there a restriction on return values (e.g.\ can a function return
	any object which could be declared)?
  \item Are there levels of scope?
	If multiple levels are permitted, how are references to stale scopes
	handled or prevented?
	Can information for declarations (such as size) be inherited (dynamically)
	from higher scopes?
  \item What synchronization requirements (if any) are imposed during calls
	(i.e.\ do all processes (in a group?) have to enter/exit the same subroutine
	at the same time)?
  \item What support is there for modularization?
	For hiding implementation details?
	For sharing such information?
	For detecting and resolving name clashes?
  \end{enumerate}
\item Concurrent Programming Facilities
  \begin{enumerate}
  \item At what level(s) is the system concurrent?
    \begin{enumerate}
    \item implicit (the user shouldn't ask such things)
    \item operational (e.g.\ data-parallel addition)
    \item program counter (single executable, multiple states)
    \item procedural (heterogeneous control parallelism)
      \begin{enumerate}
      \item one execution thread per address space
      \item multiple threads per address space
      \end{enumerate}
    \end{enumerate}
  \item What mutual exclusion and synchronization primitives are provided?
  \item At what level(s) is data sharing visible to the programmer?
    \begin{enumerate}
    \item sequential consistency
    \item sequential consistency within processor clusters
    \item object-level consistency
    \item there is no sharing
    \end{enumerate}
  \item What is the conceptual granularity of operations?
    \begin{enumerate}
    \item synchronization after atomic instruction
    \item homogeneous operations with large-grain synchronization
    \item heterogeneous operations with large-grain synchronization
    \item completely asynchronous execution
    \end{enumerate}
  \item How do processes communicate data values?
	How do they synchronize?
	Are non-blocking operations (ones in which buffers might be
	in a visibly volatile state) allowed?
	How are they checked for termination?
  \item If data are communicated explicitly,
	how does the receiver choose what to accept?
    \begin{enumerate}
    \item by address (e.g.\ channel or port)
    \item by pattern matching on contents
    \item by explicit message tagging
    \item by provision of a typed buffer
    \item by provision of an untyped (byte-block) buffer
    \item in FIFO order
    \end{enumerate}
  \item How do users specify the number of processors on which to run?
	Must this be built into the program,
	or can it be decided at run-time?
	Is there support for executing on heterogeneous processors?
  \item Can the user specify the mapping of processes to processors?
	If so, how?
	How tightly coupled are the mapping of processes to processors,
	and the mapping of distributed data structures (if supported)?
  \item Can the user specify the mapping of data structures to processors?
        Can data structures be decomposed or partitioned?
  \item What support is there for operations on groups of processes, such as:
    \begin{enumerate}
    \item broadcast
    \item multicast (partial broadcast)
    \item barrier synchronization
    \end{enumerate}
  \item Can arbitrary process groups be created,
	or must they be structured in some way
	(e.g.\ as logical mesh)?
	Can operations involving disjoint groups be executed concurrently
	(e.g.\ a broadcast within one group,
	and a barrier within another)?
  \end{enumerate}
\item Input/Output
  \begin{enumerate}
  \item Is simple sequential (single-process) I/O simple to write?
  \item Is concurrent I/O supported?  If so, which flavours:
    \begin{enumerate}
    \item single-reader-broadcast/collect-single-writer
    \item homogeneously-sized or -strided segments
    \item heterogeneously-sized or -strided segments
    \item arbitrary independent operations
    \end{enumerate}
  \item What types can be read/written intrinsically?
	E.g.\ can an entire record (or array, or list)
	be read or written in a single operation?
  \end{enumerate}
\item Access to Routines in Other Languages and to the Hardware
  \begin{enumerate}
  \item Can subroutines written in other languages be called?
        Can subroutines in this language be called from other languages?
  \item Are there ways to access the run-time system directly,
	or similar low-level facilities?
  \item Is there high-level access to physical addresses and devices?
  \end{enumerate}
\item Practical Considerations
  \begin{enumerate}
  \item Are both batch and interactive processing supported?
  \item How reliable and efficient are existing implementations?
	How fast is it (i.e.\ how quickly does code compile and load)?
	Is it supported (who fixes it if it breaks)?
  \item Is there a standard for the system (from some standards organization)?
  \item Is the documentation adequate? For whom (CS graduates, computer-literate
	users, the general population)? Is training available?	Consulting?
  \item How much code has been written (lines of code or weeks of effort)?
        By whom (computer science undergraduates, professional engineers, etc.)?
  \item How long does it take the compiler to generate a ``small'' executable
        (such as {\tt{"hello, world"}})?
  \item How large is a ``small'' executable?
        How long does it take to load such an executable?
  \end{enumerate}
\item Supporting Tools
  \begin{enumerate}
  \item What is the development environment like?
        Can code be developed on a workstation and then moved to a parallel platform?
  \item Are there debugging aids, performance analysers, and management tools?
  \item How much help do these really provide?
  \end{enumerate}
\end{enumerate}

\newpage
\section{A Critique of the C Implementation\label{s:critique}}

This section is a critique of the ANSI~C reference implementation of the problem suite.
It is intended to serve as a model for (self-)criticism of other programming systems.

\begin{itemize}
\item	C's support for multi-dimensional arrays (MDAs) is very weak.
	There is no way to dynamically allocate an MDA in a single go---one must either
	allocate a block of the same size as the desired array,
	and then do indexing calculations by hand,
	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
	MDAs do not carry dimension information with them,
	so it is impossible to determine the size of an array parameter within a function.
	Finally,
	C does not treat all axes of an array equally:
	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
	it is impossible to slice it along the other axis.
\item	C does not distinguish between Boolean and integer types.
	As a result,
	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
	can be passed in reverse order without a type error.
	Using {\tt{typedef}} to create a Boolean type name does not solve this
	(at least, not in {\tt{gcc}} V2.5.8).
\item	Union types cannot be safely initialized.
	This complicated the implementation of the graphics interface,
	where it would have been much more elegant to define a union type,
	each of whose variants held parameter specifications for a single toy.
	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
	initializing some and filling other with don't-care values.
	This is neither safe nor elegant.
\item	Parameter values cannot be used in the initialization of local variables inside functions.
	In particular,
	it is not possible to create a local vector with a length specified by an input parameter.
	Such a facility would be useful in the {\tt{winnow}} toy,
	where we have instead allocated local temporaries of the maximum possible size.
\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
	For example, the ``fail()'' error-handling routine was buggy because
	all \verb`real` arguments were being automatically converted to doubles,
	but were being taken off the stack as floats.
\item	no intrinsic notion of group ID/self ID in threads
\item	packaging parameters for threads
\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
\end{itemize}

\end{document}

### 2009-08-28 paper/2009-09/cliches.tex
Breaking paper into manageable chunks

--- 2009-08-20 paper/2009-09/paper.tex
@@
-\documentstyle[epsf,11pt]{article}
-
-\begin{document}
-
-\begin{Large}\begin{center}
-The Cowichan Problems: \\
-Assessing the Usability of Parallel Programming Systems
-\end{center}\end{Large}
-
-\vspace{\baselineskip}
-
-\begin{center}
-\begin{tabular}{cccc}
-Gregory V.~Wilson\footnote{\verb`http://www.third-bit.com`}
-&
-Andrew Borzenko
-&
-Cameron Gorrie
-&
-Zachary Kincaid
-\end{tabular}
-\end{center}
-
-\vspace{\baselineskip}
-
-\begin{abstract}
-Most programmers believe that parallel programming is harder than sequential programming.
-Many also believe that today's parallel programming systems make it even harder.
-We have developed a suite of chained toy problems that can be used to assess
-the usability of different scientific programming languages, parallel or otherwise.
-This paper motivates and describes the suite,
-and discusses what we have learned from porting it to four parallel C++ systems
-(MPI, OpenMP, Intel Thread Building Blocks, and a Linux tuple space system),
-and from translating it into sequential and parallel Haskell.
-\end{abstract}
-
-\section{Introduction\label{s:intro}}
-
-Twenty years ago,
-Bal {\em{et~al.}\/} listed more than 300 parallel programming systems \cite{b:par-lang-survey}.
-A similar survey today could well uncover twice as many,
-but a poll of application programmers would probably find that
-only a handful are in general use.
-
-The reasons why so many of these PPSs are only used by their developers (if at all) are murkier.
-Part of the blame undoubtedly lies with the diversity of hardware,
-and the attendant lack of portability,
-alluded to above.
-A second reason is that an academic researcher's principal interest is developing new languages,
-not supporting existing ones.
-By the time a system is 90\% complete,
-its author is already thinking about its successor.
-Few application programmers are willing to invest effort porting codes
-to systems which are not properly supported,
-and liable to change or evaporate without notice.
-Third,
-many research groups have chosen to explore the potential of parallelism
-without consideration for compatibility with older languages.
-Application programmers must therefore face the prospect of completely re-writing their programs,
-rather than simply porting them.
-As the next section argues,
-this can obviate any performance that might be gained from parallelization.
-Finally,
-many PPSs simply cannot support the software engineering load of large applications.
-Many provide only a single global namespace,
-do not include I/O capabilities,
-require users to marshal and unmarshal data explicitly,
-or do not permit hardware to be timeshared.
-
-The net result is that many PPSs are never used for anything more complicated than
-generating the Mandelbrot Set,
-simulating the Game of Life,
-multiplying dense matrices,
-or finding all solutions to the $N$-queens problem for some small $N$.
-Lacking the feedback that comes from more demanding use,
-few systems are improvements over their predecessors
-in the way that successive versions of operating systems have improved on one another,
-or the languages in the Algol$\rightarrow$Modula-$n$ family
-have become both more powerful and more secure.
-Most applications programmers therefore choose to play it safe
-and stick to parallel programming's lowest common denominators:
-data-parallel Fortran,
-and message-passing libraries.
-
-The historical reasons for tolerating this situation are becoming less valid with each passing year.
-When only a few parallel computers existed,
-and these were of widely different types,
-software standardization simply didn't make sense.
-Similarly,
-when supercomputers were very expensive,
-the cost of software development seemed relatively smaller,
-and the requirement for maximum performance relatively greater,
-than in sequential computing.
-As \cite{b:social-limits-speed} points out,
-however, these arguments are losing force.
-Now,
-as in sequential computing,
-the first question is not ``How do I write programs for this machine?'',
-but ``Will this machine run my programs?''
-As microprocessor-based MIMD hardware comes to dominate massively-parallel computing,
-we see increasing emphasis on standards such as High-Performance Fortran (HPF) \cite{b:hpf-overview},
-and the Message-Passing Interface (MPI) \cite{b:mpi-overview}.
-
-But herein lies a danger.
-History shows that it is extremely difficult
-to displace a programming language or system once it becomes established.
-Cobol and MS-DOS are the most frequently cited examples,
-but many systems which should have been superceded long ago have managed
-not only to cling to life,
-but to grow in influence.
-(The fact that the mail signature
-``The last good thing written in C was Schubert's Ninth Symphony''
-can elicit smiles simply proves that even the truth can be funny.)
-Established systems evolve slowly, if at all.
-It would therefore seem that the best time to guarantee ourselves
-a good parallel programming paradigm is right now,
-before a large number of large applications are created and take on lives of their own.
-
-We therefore propose to use a set of simple programming problems to compare
-the usability of parallel programming systems.
-Competent programmers, fluent in a specific system,
-will implement solutions to these problems and report their experiences
-in terms of development time, code size and clarity, and runtime efficiency.
-These problems are on the scale of the toy applications described several paragraphs ago
-both to keep implementation time low
-and because we feel that any PPS produced today ought to be able to support such simple programs.
-As the second author put it,
-``We're all sick of seeing these,
-but maybe if we all catch the disease at the same time we'll finally build up some immunity to it.''
-One significant innovation in this work is that
-we also require implementations of our suite to chain problems together,
-so that the output of one problem is the input to another.
-This will allow us to assess how well each PPS supports modularization and code re-use
-(or, more realistically, how much extra effort programmers must invest
-to achieve these goals when using a particular PPS).
-
-Section~\ref{s:corollary} motivates this work by showing
-how poor usability can limit the realizable parallelism in a program.
-Section~\ref{s:previous} surveys previous work on the comparison of parallel programming systems,
-while Section~\ref{s:method} outlines our choice of methods, and the constraints on our work.
-Section~\ref{s:toys} then presents the suite of problems we intend to use.
-In recognition of the original work done by Feo and others on
-the Salishan Problems~\cite{b:salishan}, this suite is called
-the Cowichan Problems\footnote{\noindent Pronounced {\em{Cow}\/}-i-chun.
-	Like Salishan, the word is a Northwest Coast Indian place name.}.
-
-\subsection*{History and Acknowledgments}
-
-The first author originally developed this suite in the mid-1990s,
-inspired both by \cite{b:salishan} and by
-discussions with R.\ Bruce Irvin (then at the University of Wisconsin),
-Henri Bal (Vrije Universiteit, Amsterdam),
-and others.
-Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.
-
-\section{Quantifying the Importance of Usability\label{s:corollary}}
-
-According to Amdahl's Law,
-the time $t$ required to execute a program has two components:
-$t_{seq}$, representing sequential operations that cannot be parallelized,
-and $t_{par}$, representing those operations which can.
-If we let $\sigma$ be the ``serial fraction'' of the program,
-so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
-then the speedup which can be achieved using $P$ processors is:
-\begin{eqnarray*}
-s(P)	& =	& \frac{t(1)}{t(P)}	\\
-	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
-\end{eqnarray*}
-As $P{\rightarrow}{\infty}$,
-the potential speedup is bounded by $1/{\sigma}$.
-
-What is not usually acknowledged is
-the influence of programming system usability on
-the potential speedup of sequential programs.
-As stated above, Amdahl's Law applies only to a single run of a program.
-In the real world, programs are first developed, and then run many times.
-If we let $T$ be the total running time of a program over its lifetime,
-$D_{seq}$ be the time required to develop the sequential components of a program,
-and $D_{par}$ be the time required to develop the parallel components
-(or parallelize the sequential components),
-then the lifetime speedup we can achieve is given by:
-\begin{eqnarray*}
-S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
-	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
-\end{eqnarray*}
-Assume that we are parallelizing a ``dusty deck'',
-i.e.\ that the time $D_{seq}$ has already been invested.
-If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
-then achievable speedup is limited by:
-\begin{eqnarray*}
-S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
-		& =	& \frac{1}{{\phi} + {\sigma}}
-\end{eqnarray*}
-The time required to parallelize a program can therefore be seen
-as increasing that program's effective serial fraction.
-Unless parallelization time can be substantially reduced,
-massively-parallel computing is likely to remain attractive only in those cases in which:
-\begin{itemize}
-\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
-\item	the expected total program runtime is very large
-	(i.e.\ the program is a package used extensively in a field
-	such as computational chemistry or vehicle crash simulation);
-	or
-\item	cost is not an obstacle (i.e.\ the customer is the NSA).
-\end{itemize}
-
-\section{Previous Work\label{s:previous}}
-
-In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
-\begin{quote}
-As practiced by computer science,
-the study of programming is an unholy mixture of mathematics,
-literary criticism,
-and folklore.
-\end{quote}
-Little has changed since.
-Computer scientists pay lip service to the importance of usability,
-but are strangely reluctant to try to quantify it,
-particularly with respect to the tools they use themselves.
-Religious debates about the ``right'' first-year teaching language
-or the ``best'' GUI toolkit are common;
-systematic use of the experimental method is not.
-
-Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
-such as readability and code-pattern recognition
-\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
-The little work that has been done on ``programming in the large'' has been anecdotal,
-such as \cite{b:compare-ada-c-pascal}.
-This book presented various authors' impressions of Pascal, C, and Ada
-(before any complete Ada implementations existed).
-Its editors formulated a list of questions to use in comparing languages,
-but did not use it themselves.
-A modified version of this questionnaire,
-tailored for parallel programming languages,
-is given in Appendix~\ref{s:questions},
-and will be used as a guideline in our work.
-
-Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
-The first presented implementations of a simple numerical quadrature program
-in more than a dozen different parallel languages
-in use on mid-1980s hardware.
-The second presented implementations of
-the Salishan Problems---Hamming number generation,
-isomer enumeration,
-skyline matrix reduction,
-and a simple discrete event simulator---in
-C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
-and a variety of dataflow and functional languages.
-Both of these books convey the flavor of the languages they describe,
-but neither made any effort to compare languages or problem implementations directly.
-
-Inspired by the Salishan Problems,
-the first author developed a suite of seven problems \cite{b:cowichan-1}
-in order to test the limits of parallel programming systems.
-Individual problems were badly load-balanced,
-manipulated irregular pointer-based structures,
-and so on.
-Our experience implementing them was extremely valuable,
-as it uncovered many bugs in the programming system used,
-and revealed several shortcomings of the system's design.
-However,
-we also discovered that implementing each problem takes six to eight weeks,
-which we feel is prohibitive at this exploratory stage.
-
-\section{Methodology\label{s:method}}
-
-The main aim of this work is to assess and compare
-the ease with which programs can be developed and/or ported
-using different parallel programming systems.
-Here,
-we discuss issues related to choosing the problems that make up the suite,
-assessing the complexity of particular implementations,
-and other purposes which implementations of these might serve.
-Recognizing that these problems are extremely simple,
-and easy to parallelize,
-we refer to them as ``toys''.
-
-\subsection{Criteria for Selection\label{s:method-criteria}}
-
-Our criteria for including problems in this suite are as follows:
-\begin{enumerate}
-\item	Each toy should require no more than an afternoon
-	to write and test in a well-supported sequential language.
-	We feel that making individual problems more complicated will only discourage implementation.
-\item	The correctness of each implementation must be relatively easy to verify.
-	Toys whose output is easily visualized are therefore to be preferred,
-	as are toys whose results are insensitive to floating-point discretization effects.
-\item	Each toy application should be amenable to
-	an analysis of speedup or isoefficiency \cite{b:isoefficiency-tutorial}
-	to allow a comparison of theory and practice.
-\item	At least some toys should not be ``infinitely scalable''.
-	Many real-world applications are not,
-	and this suite should reflect such limitations.
-\item	At least some toys should require I/O,
-	since this important aspect of real-world programming is often neglected by PPS designers.
-\item	There should be some potential redundancy in the type of parallelization done,
-	so that re-use of software can be demonstrated.
-\item	Together,
-	the toys in the suite must exercise a wide range of common parallel operations
-	(Appendix~\ref{s:cliche-ops})
-	and memory reference patterns
-	(Appendix~\ref{s:cliche-memref}).
-\item	The ``obvious'' implementations of these toys should span
-	a wide range of parallel programming paradigms,
-	such as task farming,
-	geometric and functional decomposition,
-	and speculative parallelism.
-\item	Parallelizations for several (if not all) of
-	the types of programming systems shown in Figure~\ref{f:taxonomy} should be known.
-	This suite is not intended to be a collection of research problems;
-	reasonably simple and efficient implementations in most environments
-	should be straightforward to produce.
-\end{enumerate}
-
-\begin{figure}
-\epsfysize=0.5\textheight
-\begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
-\caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
-\end{figure}
-
-\subsection{Specification\label{s:method-spec}}
-
-Toys should be specified in terms of inputs and outputs,
-rather than algorithmically.
-For example,
-``sort $N$ integers in the range $1{\ldots}R$,
-where $R$ is not known in advance and there may be multiple instances of a given value''
-is to be preferred over
-``parallelize quicksort''.
-As \cite{b:slalom} argued,
-which parallel algorithm is best in a particular programming system
-depends critically on the architectural assumptions made by that system.
-Implementors should be free to employ those methods which they find natural and convenient.
-The only exception we make to this rule is in the two matrix equation solvers,
-where we require Gaussian elimination and successive over-relaxation.
-Given the enormous number of matrix solution algorithms in use,
-we felt we had to specify these particular algorithms in order to ensure that
-different implementations of this suite would be comparable.
-
-\subsection{Software Engineering Issues\label{s:method-softeng}}
-
-These toys will exercise many different aspects of parallel systems.
-However,
-their ``single algorithm per program'' model is not representative of real applications.
-These usually contain several discrete (and sometimes overlapping) phases,
-each of which is qualitatively different.
-A full implementation of this suite will therefore have two parts.
-In the first,
-each toy will be implemented as a stand-alone program.
-In the second,
-toys will be chained together to create the kind of phase behaviour seen in real applications
-(Figure~\ref{f:chaining}).
-This will test the ease with which heterogeneous parallelism can be mixed within a single program.
-It will also show how well the system supports code re-use and information hiding,
-which are crucial to the development of large programs.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
-\caption{Chaining\label{f:chaining}}
-\end{figure}
-
-A second reason for chaining toys is that
-many groups are now using algorithmic skeletons to structure parallel programs
-\cite{b:cole-skel,b:p3l-overview,b:enterprise}.
-Such skeletons encapsulate the details of a particular style of parallelism,
-and hide low-level or platform-dependent issues.
-We hope that chaining will show the strengths and weaknesses of such systems.
-
-Finally,
-chaining should be designed so that some toys can be executed concurrently.
-Many PPSs impose extraneous constraints on programs,
-e.g.\ require all processes to participate in every barrier synchronization,
-or require the same executable to be loaded onto each processor.
-These constraints can limit the exploitation of potential concurrency.
-Permitting, but not requiring, concurrent execution of several toys should uncover such limitations.
-
-\subsection{Sizing\label{s:method-size}}
-
-One crucial aspect of the specification of toys is
-the way in which the sizes of problems are determined.
-In a completely frozen model, the actual size of each problem would be compiled into each toy.
-A fluid implementation, by contrast, would allow sizes to be specified at run-time,
-and would only then allocate and partition data structures.
-We have chosen to use an intermediate model (which might be called ``slushy''),
-in which the maximum size of individual problems is specified during compilation,
-but the actual size of a problem is only determined when the toy begins to execute.
-We hope that the use of this model will force implementors to deal with
-at least some of the software engineering issues
-involved in building flexible large-scale applications,
-without making implementation prohibitively difficult.
-
-\subsection{Assessing Complexity\label{s:method-complex}}
-
-There are at least two different ways in which usability might be compared.
-One would be to measure the performance achieved by an ``average'' programmer
-as a function of time on each of a range of problems.
-This would lead to graphs of the kind shown in Figure~\ref{f:perf},
-where the left graph shows a particular result
-and the right graph a simplified abstraction of it.
-By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
-which is the length of vector on which a pipelined architecture achieves
-half its theoretical peak performance,
-we could in principle find the value of $p_{1/2}$---the programming time required to achieve
-half of a machine's peak performance\footnote{Note that
-	if performance was measured as a fraction of the figures quoted by manufacturers for their machines,
-	it is unlikely that the halfway mark would ever be reached.}
-for a particular combination of programming system and problem type.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
-\caption{Performance vs.\ Time\label{f:perf}}
-\end{figure}
-
-A second option would be to measure the complexities of
-implementations of a single application using different programming systems.
-With this approach,
-comparative performance or speedup figures would be supplemented by complexity measures,
-as shown in Figure~\ref{f:complex}.
-
-\begin{figure}
-\epsfxsize=6cm
-\begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
-\caption{Comparative Code Complexity\label{f:complex}}
-\end{figure}
-
-The drawback of the first approach is that there are no ``typical'' problems,
-or ``average'' programmers.
-Meaningful measurement of  $p_{1/2}$ would therefore require
-measurement of a large number of programmers working on many problems.
-Since this is too expensive to be practical,
-the second approach is the one we have chosen to adopt.
-It has drawbacks as well,
-most notably the fact that no-one has ever demonstrated any better measure of program complexity
-than a simple count of the number of lines in a program \cite{b:fenton-metrics}.
-Since we anticipate a wide variation in the types of progamming systems that will be used in this work
-(ranging from declarative dataflow languages
-through functional languages with Lisp-like syntax
-to parallel dialects of Fortran and C),
-and in the indentation, commenting, and layout styles of particular programmers,
-we may measure complexity by counting the number of tokens in the source code which are
-keywords,
-intrinsic operators,
-user-defined functions,
-type or variable declarations,
-and miscellaneous punctuation.
-
-\subsection{Other Uses for Implementations\label{s:method-uses}}
-
-This suite is intended for assessing PPS usability,
-but we envisage at least three other uses for implementations of it:
-\begin{description}
-\item[Related Research:]
-	We hope that the existence of this suite will indicate
-	to the developers of new parallel programming systems
-	what a mature tool should be able to support.
-	If sufficiently diverse,
-	the suite could also be used to help test the correctness of new or improved systems.
-	Finally,
-	this suite could complement existing performance benchmarks,
-	since performance figures for individual toys could be compared.
-\item[Tool Development:]
-	Requests from tool developers for existing parallel programs,
-	or for event traces from them,
-	are common on newsgroups such as {\tt{comp.parallel}}.
-	We hope that traces from implementations of this suite
-	will be used to compare different debugging and performance monitoring tools,
-	that their source code will be used to test
-	new parallelizing compilation techniques or mapping and scheduling algorithms,
-	and so on.
-\item[Teaching:]
-	Implementations will be suitable for use as classroom examples,
-	since they will be small enough to be understood quickly.
-	The toys themselves should also be suitable as classroom exercises
-	in a senior undergraduate course on parallel computing.
-\end{description}
-
-\subsection{Some Criticisms\label{s:method-criticism}}
-
-\noindent {\em{Why not use an existing benchmark suite for this purpose?}\/}
-
-Implementations of benchmark suites have usually been coded for absolute speed.
-They are not representative of ``normal'' practice,
-in which reasonable performance is the usual goal,
-and so are inappropriate for this work.
-In addition,
-most existing benchmark suites are too large to be implemented or ported quickly
-(e.g.\ SPEC~\cite{b:bench-over}),
-over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
-or concerned with a limited range of applications
-(e.g.\ the NAS Parallel Benchmarks~\cite{b:nas-results}).
-Finally,
-most exist only in C or Fortran (or both);
-many interesting parallel programming systems are built on top of other languages,
-including Modula-2, Scheme, Prolog, and various dataflow languages.
-We do not want the effort required to translate several large programs
-to discourage researchers from participating in this exercise.
-
-\vspace{\baselineskip}
-
-\noindent {\em{Won't the results depend primarily on programmer ability?}\/}
-
-The short answer to this question is yes,
-but no more than the results from performance benchmarks.
-A longer answer is that	since we intend to measure the combination of code complexity and performance achieved,
-rather than performance as a function of development time,
-programmers will be allowed to revise and improve their implementations when and as desired.
-
-\section{The Cowichan Problems\label{s:toys}}
-
-This section describes the problems making up the full problem suite.
-As discussed above,
-the implementation of each is to be used both as a stand-alone program
-and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
-Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cl|cl}
-Code	& Toy				& Code	& Toy				\\
-\hline
-E	& elastic net			& O	& outer product			\\
-G	& Gaussian elimination		& P	& matrix-vector product		\\
-H	& halving shuffle		& R	& random matrix generation	\\
-I	& invasion percolation		& S	& successive over-relaxation	\\
-L	& Game of Life			& T	& image thresholding		\\
-M	& Mandelbrot Set		& V	& vector difference		\\
-N	& point normalization		& W	& point value winnowing
-\end{tabular}
-\caption{Single-Character Codes for Toys\label{t:toy-codes}}
-\end{center}
-\end{table}
-
-\subsection{{\tt{mandel}}:
-	Mandelbrot Set Generation
-	\label{s:toys-mandel}}
-
-\begin{figure}
-\begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
-\caption{The Mandelbrot Set\label{f:mandel}}
-\end{figure}
-
-This module generates the Mandelbrot Set
-for a specified region of the complex plane
-(Figure~\ref{f:mandel}).
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{x0, y0}}:]
-	the real coordinates of the lower-left corner
-	of the region to be generated.
-\item[{\tt{dx, dy}}:]
-	the extent of the region to be generated.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing the iteration count at each point in the region.
-\end{description}
-
-Given initial coordinates $(x_0, y_0)$,
-the Mandelbrot Set is generated by iterating the equation
-\begin{eqnarray*}
-x^{\prime}	& =	& x^2 - y^2 + y_0	\\
-y^{\prime}	& =	& 2{x}{y} + x_0
-\end{eqnarray*}
-until either an iteration limit is reached,
-or the values diverge.
-The iteration limit used in this module is 150 steps;
-divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
-The integer value of each element of the matrix is
-the number of iterations done.
-Note that,
-as in all problems,
-the output is required to be
-independent of the number of processors used.
-This requires programmers to be careful about
-floating-point roundoff effects.
-
-\subsection{{\tt{randmat}}:
-	Random Number Generation
-	\label{s:toys-rng}}
-
-This module fills a matrix with pseudo-random integers.
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{s}}:]
-	the random number generation seed.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix filled with random values.
-\end{description}
-Note that,
-as in all problems,
-the output is required to be independent of the number of processors used.
-Generating new seed values from the given seed,
-and running one copy of the random number generator
-on each processor,
-is therefore unlikely to qualify as a solution.
-
-\subsection{{\tt{half}}:
-	Two-Dimensional Shuffle
-	\label{s:toys-half}}
-
-This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
-shuffles them,
-and then repeats this operation along the other axis.
-Values in odd-numbered locations are collected at the low end of each row or column,
-while values in even-numbered locations are moved to the high end.
-An example transformation is:
-
-\begin{eqnarray*}
-\begin{array}{cccc}
-a & b & c & d \\
-e & f & g & h \\
-i & j & k & l
-\end{array}
-& = &
-\begin{array}{cccc}
-a & c & b & d \\
-i & k & j & l \\
-e & g & f & h
-\end{array}
-\end{eqnarray*}
-
-Note that how an array element is moved depends only on whether its location index is odd or even,
-not on whether its value is odd or even.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing shuffled values.
-\end{description}
-
-\subsection{{\tt{invperc}}:
-	Invasion Percolation
-	\label{s:toys-invperc}}
-
-Invasion percolation models the displacement of one fluid
-(such as oil)
-by another
-(such as water)
-in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
-In two dimensions,
-this can be simulated by generating
-an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
-and then marking the center cell of the grid as filled.
-In each iteration,
-one examines the four orthogonal neighbors of all filled cells,
-chooses the one with the lowest value
-(i.e.\ the one with the least resistance to filling),
-and fills it in.
-Figure~\ref{f:invperc1} shows the first few steps in this process,
-while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{math}
-\begin{array}{ccc}
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & 29 & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \star & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & 29 & 77 \\
-\hline
-61 & \star & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \star & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \star & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \star \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & \star & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-\star & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & \star & 72 & 45 & 38 \\
-\hline
-\circ & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \circ \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\end{array}
-\end{math}
-\end{footnotesize}
-\caption{Invasion Percolation\label{f:invperc1}}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
-\end{center}
-\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
-\end{figure}
-
-The simulation continues until
-some fixed percentage of cells have been filled,
-or until some other condition
-(such as the presence of trapped regions)
-is achieved.
-The fractal structure of the filled and unfilled regions
-is then examined to determine how much oil could be recovered.
-The na\"{\i}ve way to implement this is to repeatedly scan the array;
-a more sophisticated, and much faster, sequential technique
-is to maintain a priority queue of unfilled cells
-which are neighbors of filled cells.
-This latter technique is similar to the list-based methods used in some cellular automaton programs,
-and is very difficult to parallelize effectively.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{nfill}}:]
-	the number of points to fill.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
-\end{description}
-Filling begins at the central cell of the matrix
-(rounding down for even-sized axes).
-
-\subsection{{\tt{thresh}}:
-	Histogram Thresholding
-	\label{s:toys-thresh}}
-
-This module performs histogram thresholding on an image.
-Given an integer image $I$ and a target percentage $p$,
-it constructs a binary image $B$
-such that $B_{i,j}$ is set
-if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
-The general idea is that an image's histogram should have 2 peaks,
-one centered around the average foreground intensity,
-and one centered around the average background intensity.
-This program attempts to set
-a threshold between the two peaks in the histogram
-and select the pixels above the threshold.
-This module's inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the integer matrix to be thresholded.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{percent}}:]
-	the minimum percentage of cells to retain.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
-\end{description}
-
-\subsection{{\tt{life}}:
-	Game of Life
-	\label{s:toys-life}}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
-\end{center}
-\caption{Snapshots from the Game of Life\label{f:life}}
-\end{figure}
-
-This module simulates the evolution of Conway's Game of Life,
-a two-dimensional cellular automaton (Figure~\ref{f:life}).
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the Life world.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{numgen}}:]
-	the number of generations to simulate.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the world after simulation.
-\end{description}
-At each time step,
-this module must count the number of live ({\tt{true}}) neighbors
-of each cell,
-using both orthogonal and diagonal connectivity
-and circular boundary conditions.
-The update rule is simple:
-if a cell has 3 live neighbors,
-or has 2 live neighbors and is already alive,
-it is alive in the next generation.
-In any other situation,
-the cell becomes, or stays, dead.
-
-\subsection{{\tt{winnow}}:
-	Weighted Point Selection
-	\label{s:toys-winnow}}
-
-This module converts a matrix of integer values
-to a vector of points,
-represented as $x$ and $y$ coordinates.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix, whose values are used as masses.
-\item[{\tt{mask}}:]
-	a Boolean matrix showing which points are eligible for consideration.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{nelts}}:]
-	the number of points to select.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points.
-\end{description}
-Each location where {\tt{mask}} is {\tt{true}}
-becomes a candidate point,
-with a weight equal to the integer value in {\tt{matrix}} at that location
-and $x$ and $y$ coordinates
-equal to its row and column indices.
-These candidate points are then sorted into increasing order by weight,
-and {\tt{nelts}} evenly-spaced points selected to create the result vector.
-
-\subsection{{\tt{norm}}:
-	Point Location Normalization
-	\label{s:toys-norm}}
-
-This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-Its inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of point locations.
-\item[{\tt{nelts}}:]
-	the number of elements in the point vector.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of normalized point locations.
-\end{description}
-If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
-then the normalization equation is:
-\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
-$y$ coordinates are normalized in the same fashion.
-
-\subsection{{\tt{elastic}}:
-	Elastic Net Simulation
-	\label{s:toys-elastic}}
-
-This module uses the elastic net algorithm \cite{b:elastic-net} to find an approximate solution to the travelling salesentity problem (TSP).
-A closed, circular loop, modelled by a large number of doubly-connected points, is placed in the plane.
-This path is slowly deformed until it connects the points representing the cities to be visited (Figure~\ref{f:elastic}).
-The rate at which the loop is deformed determines the quality of the final path.
-
-\begin{figure}
-\begin{center}
-\framebox{\epsffile{fig/elastic1.ps}}
-\framebox{\epsffile{fig/elastic2.ps}}
-\framebox{\epsffile{fig/elastic3.ps}}
-\end{center}
-\caption{Relaxation of an Elastic Net\label{f:elastic}}
-\end{figure}
-
-Let the (fixed) location of city $i$ be $c_i$, and the (changeable) location of point $j$ on the elastic net be $p_j$.
-The change in the location of point $j$ during each iteration of the algorithm is given by:
-\begin{equation}
-{\delta}{p_j} = {\alpha}\sum_{i}{w_{ij}(c_i - p_j)} ~+~ K{\beta}(p_{j+1} - 2p_j + p_{j-1})
-\end{equation}
-$\alpha$ and $\beta$ are time-invariant factors relating the strengths of the city-to-point and point-to-point forces,
-and $K$ is a time-dependent factor which controls the rate of annealing.
-The strength of the force $w_{ij}$ between city $c_i$ and point $p_j$ is given by:
-\begin{equation}
-w_{ij} = {\phi}({\mid}c_i - p_j{\mid}, K) / \sum_{\ell}\phi({\mid}c_i - p_{\ell}{\mid}, K)
-\end{equation}
-where $\phi(d, K)$ is the Gaussian function $e^{-{d^2}/{K^2}}$.
-Note the normalization of each $w_{ij}$; in order for this algorithm to work properly, the total force being exerted by each city must be the same.
-This is accomplished by summing the forces exerted by each city, and then normalizing the city's forces by that total.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{cities}}:]
-	a vector of points representing the locations of the cities to be toured.
-	The $(x,y)$ coordinates of these cities must be scaled to lie in the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-\item[{\tt{n\_cities}}:]
-	the number of cities in {\tt{cities}}.
-\item[{\tt{n\_net}}:]
-	the number of elastic net points to use.
-\item[{\tt{n\_iters}}:]
-	the total number of iterations to perform.
-\item[{\tt{n\_relax}}:]
-	the number of iterations to perform with each discrete value of $K$.
-	After each {\tt{n\_relax}} iterations, $K$ is reduced by a (constant) pre-defined scaling factor.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{net}}:]
-	a vector of {\tt{n\_net}} points representing the tour found.
-\end{description}
-
-\subsection{{\tt{outer}}:
-	Outer Product
-	\label{s:toys-outer}}
-
-This module turns a vector containing point positions
-into a dense, symmetric, diagonally dominant matrix
-by calculating the distances between each pair of points.
-It also constructs a real vector whose values are
-the distance of each point from the origin.
-Inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
-\item[{\tt{nelts}}:]
-	the number of points in the vector, and the size of the matrix along each axis.
-\end{description}
-Its outputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a real matrix, whose values are filled with inter-point distances.
-\item[{\tt{vector}}:]
-	a real vector, whose values are filled with origin-to-point distances.
-\end{description}
-Each matrix element $M_{i,j}$ such that $i \neq j$
-is given the value $d_{i, j}$,
-the Euclidean distance between point $i$ and point $j$.
-The diagonal values $M_{i, i}$ are then set to
-{\tt{nelts}} times the maximum off-diagonal value
-to ensure that the matrix is diagonally dominant.
-The value of the vector element $v_i$ is set to
-the distance of point $i$ from the origin,
-which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.
-
-\subsection{{\tt{gauss}}:
-	Gaussian Elimination
-	\label{s:toys-gauss}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using explicit reduction.
-(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
-there is a well-formed solution to the equation.)
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{sor}}:
-	Successive Over-Relaxation
-	\label{s:toys-sor}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using successive over-relaxation.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\item[{\tt{tolerance}}:]
-	the solution tolerance, e.g.\ $10^{-6}$.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{product}}:
-	Matrix-Vector Product
-	\label{s:toys-product}}
-
-Given a matrix $A$,
-a vector $V$,
-and an assumed solution $X$ to the equation $AX=V$,
-this module calculates the actual product $AX={V^{\prime}}$,
-and then finds the magnitude of the error.
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{actual}}:]
-	the real vector $V$.
-\item[{\tt{candidate}}:]
-	a real vector containing the supposed solution.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-The output of this function is:
-\begin{description}
-\item[{\tt{e}}:]
-	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
-\end{description}
-
-\section{Other Issues\label{s:issues}}
-
-As described in Section~\ref{s:method-softeng},
-implementors are required to chain modules together in the manner shown in Figure~\ref{f:chaining}.
-This section describes the flow of control and data between these chained problems,
-and discusses several other important aspects of this suite,
-including I/O and visualization.
-
-\subsection{Chained Problem Sequences\label{s:issues-chain}}
-
-The chained implementation of these problems
-executes individual toys in the following order:
-\begin{enumerate}
-\item	An integer matrix $I$ with $r$ rows and $c$ columns is created
-	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
-	\label{e:chain-choice-1}
-\item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
-\item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
-	to generate a Boolean mask $B$ from $I$ in which the (minimum) percentage of {\tt{true}} locations is $P$.
-	Like $I$, $B$ has $r$ rows and $c$ columns.
-	\label{e:chain-choice-2}
-\item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
-	This step overwrites the Boolean matrix $B$.
-\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
-\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
-\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
-\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
-	as described in Section~\ref{s:toys-outer}.
-\item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
-	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
-	This measures the agreement between the solutions found by the two methods.
-\end{enumerate}
-Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
-there are four possible chained sequences.
-A minimal implementation will provide one of these,
-while a full implementation will be a single program which can execute any of the four sequences.
-Intermediate implementations providing all four sequences as separate programs are also acceptable.
-
-\subsection{Input and Output\label{s:issues-io}}
-
-I/O is an important part of programming,
-but is often treated as being of secondary importance by language designers.
-This suite requires all stand-alone toys to read input values from files, and write results back;
-the chained version must be able to checkpoint intermediate results between toys.
-Finally,
-implementors are strongly encouraged to include some means of visualizing the output or evolution of individual toys.
-
-The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
-A file containing a vector begins with 1 positive integer,
-which specifies the number of elements in the vector;
-a file containing a matrix begins with 2 positive integers,
-which specify the number of rows and columns in the matrix respectively.
-Elements of the vector or matrix then appear one per line,
-from least-indexed to greatest-indexed
-(and in row-major order for matrices).
-
-The formats described above are required to be human-readable.
-Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
-using whatever file formats are convenient.
-This will allow programmers to demonstrate the ``natural'' I/O capabilities of particular systems
-which most probably be used for checkpointing intermediate results in real programs.
-
-\begin{figure}
-\epsfxsize=0.75\textwidth
-\begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
-\caption{Cowichan File Formats\label{f:io-formats}}
-\end{figure}
-
-\subsection{Visualization\label{s:issues-vis}}
-
-Visualization is an increasingly important part of scientific and numerical computing.
-Including it as an optional part of this problem set is also a good way to exercise both
-the use of libraries within a parallel programming system,
-and the ease with which a programming system can support variant versions of a program.
-
-Our requirements for the graphics interface are:
-\begin{enumerate}
-\item	Visualization for individual toys should be as simple as possible,
-	and should require as few changes to the basic toy code as possible.
-\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
-	The latter should create all images in a single window.
-\item	A single source file should implement both the graphical and non-graphical versions of each toy.
-\end{enumerate}
-The first two points require the system to distinguish between the actual drawing code
-(e.g.\ drawing a vector of points as circles)
-and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
-While the stand-alone toys use a whole window for drawing,
-the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
-At present,
-this is implemented by using a call counter
-in the interfaces for the individual toys and in the chained module.
-The actual drawing routines called from within each toy could then be the same.
-
-The present visualization routines were originally written by Anil Sukul,
-a student at the Vrije Universiteit, Amsterdam.
-The package uses the public domain graphics library VOGLE,
-available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
-VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
-as well as screen control routines (such as double buffering).
-VOGLE was chosen because of its simplicity,
-and because it includes Pascal and Fortran interfaces.
-
-\subsection{Reproducibility\label{s:issues-reproduce}}
-
-Reproducibility is an important issue for parallel programming systems.
-While constraining the order of operations in a parallel system to guarantee reproducibility makes programs in that system easier to debug,
-it can also reduce the expressiveness or performance of the system.
-
-In this problem set, irreproducibility can appear in two forms: numerical and algorithmic.
-The first arises in toys such as {\tt{gauss}}, {\tt{sor}}, and {\tt{elastic}}, which use floating-point numbers.
-Precisely how round-off errors occur in these calculations can depend on the distribution of work among processors,
-or the order in which those processors carry out particular operations.
-
-Irreproducibility also arises in toys which only use exact numbers, such as {\tt{invperc}} and {\tt{randmat}}.
-In the former, the percolation region is grown by repeatedly filling the lowest-valued cell on its perimeter.
-If several cells have this value, implementations may choose one arbitrarily.
-Thus, different implementations may produce very different fractal shapes.
-In the case of random number generation, the simplest thing to do is to run the same generator independently on each processor,
-although the values in the resulting matrix then depend on the number of processors used.
-
-\section{Conclusion\label{s:conclude}}
-
-A full ANSI~C version of this problem suite has been written, and is available by anonymous FTP from:
-\begin{center}
-{\tt{ftp.cs.toronto.edu :: pub/gvw/cowichan/src.tar.gz}}
-\end{center}
-This release includes two parallel versions,
-both based on the POSIX threads package.
-The first uses repeated fork-and-join parallelism,
-while the second creates a fixed number of threads at the beginning of the program,
-then synchronizes them repeatedly using barriers.
-
-We would like to invite interested groups to participate in this project by parallelizing the problem suite using their favourite language or tool.
-Our aim is to make both the implementations and their descriptions available as a hypertext document through the World-Wide Web,
-and, if there is sufficient interest, to publish this material in book form.
-If you would like to contribute to this effort, please contact either of the first two authors.
-
-\bibliographystyle{plain}
-\newcommand{\bibea}{{\em{et~al.}}}
-\bibliography{paper}
-
-\appendix
-
-\newpage
 \section{Parallel Clich\'{e}s\label{s:cliche}}
@@
 \end{itemize}
-
-\newpage
-\section{Comparing Parallel Programming Systems\label{s:questions}}
-
-This list of questions is a modified version of the one presented in \cite{b:compare-ada-c-pascal}.
-It is presented here as an organizational aid;
-when describing the PPS they have used,
-implementors are encouraged to use the order and terminology of this list where appropriate.
-We acknowledge that many of these questions will not apply to particular systems,
-and encourage implementors to ask questions which we have left out.
-
-\begin{enumerate}
-\item History and Philosophy
-  \begin{enumerate}
-  \item For whom and for what purpose was the language designed?
-  \end{enumerate}
-\item Syntax
-  \begin{enumerate}
-  \item Model implemented (Figure~\ref{f:taxonomy})
-  \item Class
-    \begin{enumerate}
-    \item library called from existing sequential language?
-    \item syntactic extensions to existing sequential language?
-    \item new language in Backus (Fortran/Algol/C) tradition?
-    \item new (non-Backus) language?
-    \end{enumerate}
-  \item Concurrency mechanisms adhere to same philosophy as original
-	language (where appropriate)?
-  \item Readability?  Syntactic consistency?  Conciseness?
-	(on an APL-to-Modula scale)
-  \item Common errors (e.g.\ confusing {\tt{=}} and {\tt{==}} in C)?
-  \end{enumerate}
-\item Type Philosophy and Data Types
-  \begin{enumerate}
-  \item What primitive types does the language support?
-  \item How are new types created?
-  \item What intrinsic operations are defined on user-defined types?
-  \item How strict is typing?
-	How is type conversion done?
-  \item Can type information be inspected at run-time?
-  \item What concurrent operations on aggregates
-	(such as lists and arrays) are intrinsic?
-	What kinds of concurrent aggregate operations can users define?
-  \item How are aggregates defined (e.g.\ by index set, by shape)?
-	Can new kinds of aggregates (e.g.\ templates for arrays)
-	be defined?
-  \item How strict is typing during communication?
-  \item What support is there for communication or manipulation
-	of data of non-scalar or non-intrinsic types?
-  \item Can structural information about aggregates
-	(e.g.\ number of elements in a set, dimension of an array)
-	be inspected at run-time?
-  \item What built-in support is there for object classes and inheritance?
-	For polymorphism?
-	For class (as opposed to instance) data?
-  \item Are functions first-class?
-  \item Are continuations supported?  Are they first-class?
-  \item Is dynamically-allocated memory supported?
-	Is it done automatically?
-	Is garbage-collection done automatically?
-	Can a program ever create a dangling pointer?
-	An alias?
-  \end{enumerate}
-\item Operators
-  \begin{enumerate}
-  \item What intrinsic operators are provided (brief list)?
-	How strongly do these type-check their arguments?
-  \item How are new operators defined?
-  \item Are the precedence rules straightforward?
-  \item Are side-effects/mutation allowed inside non-call expressions
-	(e.g.\ C's side-effecting \verb`i++` construct)?
-  \item Are their results platform-independent?
-  \item Are the results of intrinsic parallel operations reproducible?
-	Are they platform-independent?
-  \item Are heterogeneous operations on aggregates allowed
-	(e.g.\ every element of a list of polygons calculates its
-	center of mass)?
-	If so, how are these described, and how are they implemented?
-  \item Which of the operators in Section~\ref{s:cliche-ops} are supported?
-  \end{enumerate}
-\item Control Flow
-  \begin{enumerate}
-  \item What synchronization is automatically imposed during execution
-	(i.e.\ at what points can definite assertions about the state
-	of the running program be made)?
-  \item What control constructs does the language support?  Encourage?
-  \item How is concurrency expressed?  How structured is this mechanism?
-  \item Can the degree of concurrency be throttled (limited)?
-	Is this done by the system, or by the user?
-  \item How are exceptions handled/reported?
-	Is it different for intrinsic operations and user-defined operations?
-  \end{enumerate}
-\item Subroutines, Scope, and Modularization
-  \begin{enumerate}
-  \item How efficient is subroutine call compared to in-line execution
-	(i.e.\ how expensive are parameter passing and non-local control flow)?
-  \item How much information about parameters is available (required)
-	inside called subroutine
-	(e.g.\ whether object is shared with other processes or private
-	to executing process)?
-  \item Are side-effecting procedures allowed?
-  \item Can generic routines be written
-	(e.g.\ to work on different types or array shapes)?
-  \item Is there a restriction on return values (e.g.\ can a function return
-	any object which could be declared)?
-  \item Are there levels of scope?
-	If multiple levels are permitted, how are references to stale scopes
-	handled or prevented?
-	Can information for declarations (such as size) be inherited (dynamically)
-	from higher scopes?
-  \item What synchronization requirements (if any) are imposed during calls
-	(i.e.\ do all processes (in a group?) have to enter/exit the same subroutine
-	at the same time)?
-  \item What support is there for modularization?
-	For hiding implementation details?
-	For sharing such information?
-	For detecting and resolving name clashes?
-  \end{enumerate}
-\item Concurrent Programming Facilities
-  \begin{enumerate}
-  \item At what level(s) is the system concurrent?
-    \begin{enumerate}
-    \item implicit (the user shouldn't ask such things)
-    \item operational (e.g.\ data-parallel addition)
-    \item program counter (single executable, multiple states)
-    \item procedural (heterogeneous control parallelism)
-      \begin{enumerate}
-      \item one execution thread per address space
-      \item multiple threads per address space
-      \end{enumerate}
-    \end{enumerate}
-  \item What mutual exclusion and synchronization primitives are provided?
-  \item At what level(s) is data sharing visible to the programmer?
-    \begin{enumerate}
-    \item sequential consistency
-    \item sequential consistency within processor clusters
-    \item object-level consistency
-    \item there is no sharing
-    \end{enumerate}
-  \item What is the conceptual granularity of operations?
-    \begin{enumerate}
-    \item synchronization after atomic instruction
-    \item homogeneous operations with large-grain synchronization
-    \item heterogeneous operations with large-grain synchronization
-    \item completely asynchronous execution
-    \end{enumerate}
-  \item How do processes communicate data values?
-	How do they synchronize?
-	Are non-blocking operations (ones in which buffers might be
-	in a visibly volatile state) allowed?
-	How are they checked for termination?
-  \item If data are communicated explicitly,
-	how does the receiver choose what to accept?
-    \begin{enumerate}
-    \item by address (e.g.\ channel or port)
-    \item by pattern matching on contents
-    \item by explicit message tagging
-    \item by provision of a typed buffer
-    \item by provision of an untyped (byte-block) buffer
-    \item in FIFO order
-    \end{enumerate}
-  \item How do users specify the number of processors on which to run?
-	Must this be built into the program,
-	or can it be decided at run-time?
-	Is there support for executing on heterogeneous processors?
-  \item Can the user specify the mapping of processes to processors?
-	If so, how?
-	How tightly coupled are the mapping of processes to processors,
-	and the mapping of distributed data structures (if supported)?
-  \item Can the user specify the mapping of data structures to processors?
-        Can data structures be decomposed or partitioned?
-  \item What support is there for operations on groups of processes, such as:
-    \begin{enumerate}
-    \item broadcast
-    \item multicast (partial broadcast)
-    \item barrier synchronization
-    \end{enumerate}
-  \item Can arbitrary process groups be created,
-	or must they be structured in some way
-	(e.g.\ as logical mesh)?
-	Can operations involving disjoint groups be executed concurrently
-	(e.g.\ a broadcast within one group,
-	and a barrier within another)?
-  \end{enumerate}
-\item Input/Output
-  \begin{enumerate}
-  \item Is simple sequential (single-process) I/O simple to write?
-  \item Is concurrent I/O supported?  If so, which flavours:
-    \begin{enumerate}
-    \item single-reader-broadcast/collect-single-writer
-    \item homogeneously-sized or -strided segments
-    \item heterogeneously-sized or -strided segments
-    \item arbitrary independent operations
-    \end{enumerate}
-  \item What types can be read/written intrinsically?
-	E.g.\ can an entire record (or array, or list)
-	be read or written in a single operation?
-  \end{enumerate}
-\item Access to Routines in Other Languages and to the Hardware
-  \begin{enumerate}
-  \item Can subroutines written in other languages be called?
-        Can subroutines in this language be called from other languages?
-  \item Are there ways to access the run-time system directly,
-	or similar low-level facilities?
-  \item Is there high-level access to physical addresses and devices?
-  \end{enumerate}
-\item Practical Considerations
-  \begin{enumerate}
-  \item Are both batch and interactive processing supported?
-  \item How reliable and efficient are existing implementations?
-	How fast is it (i.e.\ how quickly does code compile and load)?
-	Is it supported (who fixes it if it breaks)?
-  \item Is there a standard for the system (from some standards organization)?
-  \item Is the documentation adequate? For whom (CS graduates, computer-literate
-	users, the general population)? Is training available?	Consulting?
-  \item How much code has been written (lines of code or weeks of effort)?
-        By whom (computer science undergraduates, professional engineers, etc.)?
-  \item How long does it take the compiler to generate a ``small'' executable
-        (such as {\tt{"hello, world"}})?
-  \item How large is a ``small'' executable?
-        How long does it take to load such an executable?
-  \end{enumerate}
-\item Supporting Tools
-  \begin{enumerate}
-  \item What is the development environment like?
-        Can code be developed on a workstation and then moved to a parallel platform?
-  \item Are there debugging aids, performance analysers, and management tools?
-  \item How much help do these really provide?
-  \end{enumerate}
-\end{enumerate}
-
-\newpage
-\section{A Critique of the C Implementation\label{s:critique}}
-
-This section is a critique of the ANSI~C reference implementation of the problem suite.
-It is intended to serve as a model for (self-)criticism of other programming systems.
-
-\begin{itemize}
-\item	C's support for multi-dimensional arrays (MDAs) is very weak.
-	There is no way to dynamically allocate an MDA in a single go---one must either
-	allocate a block of the same size as the desired array,
-	and then do indexing calculations by hand,
-	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
-	MDAs do not carry dimension information with them,
-	so it is impossible to determine the size of an array parameter within a function.
-	Finally,
-	C does not treat all axes of an array equally:
-	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
-	it is impossible to slice it along the other axis.
-\item	C does not distinguish between Boolean and integer types.
-	As a result,
-	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
-	can be passed in reverse order without a type error.
-	Using {\tt{typedef}} to create a Boolean type name does not solve this
-	(at least, not in {\tt{gcc}} V2.5.8).
-\item	Union types cannot be safely initialized.
-	This complicated the implementation of the graphics interface,
-	where it would have been much more elegant to define a union type,
-	each of whose variants held parameter specifications for a single toy.
-	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
-	initializing some and filling other with don't-care values.
-	This is neither safe nor elegant.
-\item	Parameter values cannot be used in the initialization of local variables inside functions.
-	In particular,
-	it is not possible to create a local vector with a length specified by an input parameter.
-	Such a facility would be useful in the {\tt{winnow}} toy,
-	where we have instead allocated local temporaries of the maximum possible size.
-\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
-	For example, the ``fail()'' error-handling routine was buggy because
-	all \verb`real` arguments were being automatically converted to doubles,
-	but were being taken off the stack as floats.
-\item	no intrinsic notion of group ID/self ID in threads
-\item	packaging parameters for threads
-\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
-\end{itemize}
-
-\end{document}

### 2009-08-28 paper/2009-09/compare.tex
Breaking paper into manageable chunks

--- 2009-08-20 paper/2009-09/paper.tex
@@
-\documentstyle[epsf,11pt]{article}
-
-\begin{document}
-
-\begin{Large}\begin{center}
-The Cowichan Problems: \\
-Assessing the Usability of Parallel Programming Systems
-\end{center}\end{Large}
-
-\vspace{\baselineskip}
-
-\begin{center}
-\begin{tabular}{cccc}
-Gregory V.~Wilson\footnote{\verb`http://www.third-bit.com`}
-&
-Andrew Borzenko
-&
-Cameron Gorrie
-&
-Zachary Kincaid
-\end{tabular}
-\end{center}
-
-\vspace{\baselineskip}
-
-\begin{abstract}
-Most programmers believe that parallel programming is harder than sequential programming.
-Many also believe that today's parallel programming systems make it even harder.
-We have developed a suite of chained toy problems that can be used to assess
-the usability of different scientific programming languages, parallel or otherwise.
-This paper motivates and describes the suite,
-and discusses what we have learned from porting it to four parallel C++ systems
-(MPI, OpenMP, Intel Thread Building Blocks, and a Linux tuple space system),
-and from translating it into sequential and parallel Haskell.
-\end{abstract}
-
-\section{Introduction\label{s:intro}}
-
-Twenty years ago,
-Bal {\em{et~al.}\/} listed more than 300 parallel programming systems \cite{b:par-lang-survey}.
-A similar survey today could well uncover twice as many,
-but a poll of application programmers would probably find that
-only a handful are in general use.
-
-The reasons why so many of these PPSs are only used by their developers (if at all) are murkier.
-Part of the blame undoubtedly lies with the diversity of hardware,
-and the attendant lack of portability,
-alluded to above.
-A second reason is that an academic researcher's principal interest is developing new languages,
-not supporting existing ones.
-By the time a system is 90\% complete,
-its author is already thinking about its successor.
-Few application programmers are willing to invest effort porting codes
-to systems which are not properly supported,
-and liable to change or evaporate without notice.
-Third,
-many research groups have chosen to explore the potential of parallelism
-without consideration for compatibility with older languages.
-Application programmers must therefore face the prospect of completely re-writing their programs,
-rather than simply porting them.
-As the next section argues,
-this can obviate any performance that might be gained from parallelization.
-Finally,
-many PPSs simply cannot support the software engineering load of large applications.
-Many provide only a single global namespace,
-do not include I/O capabilities,
-require users to marshal and unmarshal data explicitly,
-or do not permit hardware to be timeshared.
-
-The net result is that many PPSs are never used for anything more complicated than
-generating the Mandelbrot Set,
-simulating the Game of Life,
-multiplying dense matrices,
-or finding all solutions to the $N$-queens problem for some small $N$.
-Lacking the feedback that comes from more demanding use,
-few systems are improvements over their predecessors
-in the way that successive versions of operating systems have improved on one another,
-or the languages in the Algol$\rightarrow$Modula-$n$ family
-have become both more powerful and more secure.
-Most applications programmers therefore choose to play it safe
-and stick to parallel programming's lowest common denominators:
-data-parallel Fortran,
-and message-passing libraries.
-
-The historical reasons for tolerating this situation are becoming less valid with each passing year.
-When only a few parallel computers existed,
-and these were of widely different types,
-software standardization simply didn't make sense.
-Similarly,
-when supercomputers were very expensive,
-the cost of software development seemed relatively smaller,
-and the requirement for maximum performance relatively greater,
-than in sequential computing.
-As \cite{b:social-limits-speed} points out,
-however, these arguments are losing force.
-Now,
-as in sequential computing,
-the first question is not ``How do I write programs for this machine?'',
-but ``Will this machine run my programs?''
-As microprocessor-based MIMD hardware comes to dominate massively-parallel computing,
-we see increasing emphasis on standards such as High-Performance Fortran (HPF) \cite{b:hpf-overview},
-and the Message-Passing Interface (MPI) \cite{b:mpi-overview}.
-
-But herein lies a danger.
-History shows that it is extremely difficult
-to displace a programming language or system once it becomes established.
-Cobol and MS-DOS are the most frequently cited examples,
-but many systems which should have been superceded long ago have managed
-not only to cling to life,
-but to grow in influence.
-(The fact that the mail signature
-``The last good thing written in C was Schubert's Ninth Symphony''
-can elicit smiles simply proves that even the truth can be funny.)
-Established systems evolve slowly, if at all.
-It would therefore seem that the best time to guarantee ourselves
-a good parallel programming paradigm is right now,
-before a large number of large applications are created and take on lives of their own.
-
-We therefore propose to use a set of simple programming problems to compare
-the usability of parallel programming systems.
-Competent programmers, fluent in a specific system,
-will implement solutions to these problems and report their experiences
-in terms of development time, code size and clarity, and runtime efficiency.
-These problems are on the scale of the toy applications described several paragraphs ago
-both to keep implementation time low
-and because we feel that any PPS produced today ought to be able to support such simple programs.
-As the second author put it,
-``We're all sick of seeing these,
-but maybe if we all catch the disease at the same time we'll finally build up some immunity to it.''
-One significant innovation in this work is that
-we also require implementations of our suite to chain problems together,
-so that the output of one problem is the input to another.
-This will allow us to assess how well each PPS supports modularization and code re-use
-(or, more realistically, how much extra effort programmers must invest
-to achieve these goals when using a particular PPS).
-
-Section~\ref{s:corollary} motivates this work by showing
-how poor usability can limit the realizable parallelism in a program.
-Section~\ref{s:previous} surveys previous work on the comparison of parallel programming systems,
-while Section~\ref{s:method} outlines our choice of methods, and the constraints on our work.
-Section~\ref{s:toys} then presents the suite of problems we intend to use.
-In recognition of the original work done by Feo and others on
-the Salishan Problems~\cite{b:salishan}, this suite is called
-the Cowichan Problems\footnote{\noindent Pronounced {\em{Cow}\/}-i-chun.
-	Like Salishan, the word is a Northwest Coast Indian place name.}.
-
-\subsection*{History and Acknowledgments}
-
-The first author originally developed this suite in the mid-1990s,
-inspired both by \cite{b:salishan} and by
-discussions with R.\ Bruce Irvin (then at the University of Wisconsin),
-Henri Bal (Vrije Universiteit, Amsterdam),
-and others.
-Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.
-
-\section{Quantifying the Importance of Usability\label{s:corollary}}
-
-According to Amdahl's Law,
-the time $t$ required to execute a program has two components:
-$t_{seq}$, representing sequential operations that cannot be parallelized,
-and $t_{par}$, representing those operations which can.
-If we let $\sigma$ be the ``serial fraction'' of the program,
-so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
-then the speedup which can be achieved using $P$ processors is:
-\begin{eqnarray*}
-s(P)	& =	& \frac{t(1)}{t(P)}	\\
-	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
-\end{eqnarray*}
-As $P{\rightarrow}{\infty}$,
-the potential speedup is bounded by $1/{\sigma}$.
-
-What is not usually acknowledged is
-the influence of programming system usability on
-the potential speedup of sequential programs.
-As stated above, Amdahl's Law applies only to a single run of a program.
-In the real world, programs are first developed, and then run many times.
-If we let $T$ be the total running time of a program over its lifetime,
-$D_{seq}$ be the time required to develop the sequential components of a program,
-and $D_{par}$ be the time required to develop the parallel components
-(or parallelize the sequential components),
-then the lifetime speedup we can achieve is given by:
-\begin{eqnarray*}
-S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
-	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
-\end{eqnarray*}
-Assume that we are parallelizing a ``dusty deck'',
-i.e.\ that the time $D_{seq}$ has already been invested.
-If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
-then achievable speedup is limited by:
-\begin{eqnarray*}
-S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
-		& =	& \frac{1}{{\phi} + {\sigma}}
-\end{eqnarray*}
-The time required to parallelize a program can therefore be seen
-as increasing that program's effective serial fraction.
-Unless parallelization time can be substantially reduced,
-massively-parallel computing is likely to remain attractive only in those cases in which:
-\begin{itemize}
-\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
-\item	the expected total program runtime is very large
-	(i.e.\ the program is a package used extensively in a field
-	such as computational chemistry or vehicle crash simulation);
-	or
-\item	cost is not an obstacle (i.e.\ the customer is the NSA).
-\end{itemize}
-
-\section{Previous Work\label{s:previous}}
-
-In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
-\begin{quote}
-As practiced by computer science,
-the study of programming is an unholy mixture of mathematics,
-literary criticism,
-and folklore.
-\end{quote}
-Little has changed since.
-Computer scientists pay lip service to the importance of usability,
-but are strangely reluctant to try to quantify it,
-particularly with respect to the tools they use themselves.
-Religious debates about the ``right'' first-year teaching language
-or the ``best'' GUI toolkit are common;
-systematic use of the experimental method is not.
-
-Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
-such as readability and code-pattern recognition
-\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
-The little work that has been done on ``programming in the large'' has been anecdotal,
-such as \cite{b:compare-ada-c-pascal}.
-This book presented various authors' impressions of Pascal, C, and Ada
-(before any complete Ada implementations existed).
-Its editors formulated a list of questions to use in comparing languages,
-but did not use it themselves.
-A modified version of this questionnaire,
-tailored for parallel programming languages,
-is given in Appendix~\ref{s:questions},
-and will be used as a guideline in our work.
-
-Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
-The first presented implementations of a simple numerical quadrature program
-in more than a dozen different parallel languages
-in use on mid-1980s hardware.
-The second presented implementations of
-the Salishan Problems---Hamming number generation,
-isomer enumeration,
-skyline matrix reduction,
-and a simple discrete event simulator---in
-C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
-and a variety of dataflow and functional languages.
-Both of these books convey the flavor of the languages they describe,
-but neither made any effort to compare languages or problem implementations directly.
-
-Inspired by the Salishan Problems,
-the first author developed a suite of seven problems \cite{b:cowichan-1}
-in order to test the limits of parallel programming systems.
-Individual problems were badly load-balanced,
-manipulated irregular pointer-based structures,
-and so on.
-Our experience implementing them was extremely valuable,
-as it uncovered many bugs in the programming system used,
-and revealed several shortcomings of the system's design.
-However,
-we also discovered that implementing each problem takes six to eight weeks,
-which we feel is prohibitive at this exploratory stage.
-
-\section{Methodology\label{s:method}}
-
-The main aim of this work is to assess and compare
-the ease with which programs can be developed and/or ported
-using different parallel programming systems.
-Here,
-we discuss issues related to choosing the problems that make up the suite,
-assessing the complexity of particular implementations,
-and other purposes which implementations of these might serve.
-Recognizing that these problems are extremely simple,
-and easy to parallelize,
-we refer to them as ``toys''.
-
-\subsection{Criteria for Selection\label{s:method-criteria}}
-
-Our criteria for including problems in this suite are as follows:
-\begin{enumerate}
-\item	Each toy should require no more than an afternoon
-	to write and test in a well-supported sequential language.
-	We feel that making individual problems more complicated will only discourage implementation.
-\item	The correctness of each implementation must be relatively easy to verify.
-	Toys whose output is easily visualized are therefore to be preferred,
-	as are toys whose results are insensitive to floating-point discretization effects.
-\item	Each toy application should be amenable to
-	an analysis of speedup or isoefficiency \cite{b:isoefficiency-tutorial}
-	to allow a comparison of theory and practice.
-\item	At least some toys should not be ``infinitely scalable''.
-	Many real-world applications are not,
-	and this suite should reflect such limitations.
-\item	At least some toys should require I/O,
-	since this important aspect of real-world programming is often neglected by PPS designers.
-\item	There should be some potential redundancy in the type of parallelization done,
-	so that re-use of software can be demonstrated.
-\item	Together,
-	the toys in the suite must exercise a wide range of common parallel operations
-	(Appendix~\ref{s:cliche-ops})
-	and memory reference patterns
-	(Appendix~\ref{s:cliche-memref}).
-\item	The ``obvious'' implementations of these toys should span
-	a wide range of parallel programming paradigms,
-	such as task farming,
-	geometric and functional decomposition,
-	and speculative parallelism.
-\item	Parallelizations for several (if not all) of
-	the types of programming systems shown in Figure~\ref{f:taxonomy} should be known.
-	This suite is not intended to be a collection of research problems;
-	reasonably simple and efficient implementations in most environments
-	should be straightforward to produce.
-\end{enumerate}
-
-\begin{figure}
-\epsfysize=0.5\textheight
-\begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
-\caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
-\end{figure}
-
-\subsection{Specification\label{s:method-spec}}
-
-Toys should be specified in terms of inputs and outputs,
-rather than algorithmically.
-For example,
-``sort $N$ integers in the range $1{\ldots}R$,
-where $R$ is not known in advance and there may be multiple instances of a given value''
-is to be preferred over
-``parallelize quicksort''.
-As \cite{b:slalom} argued,
-which parallel algorithm is best in a particular programming system
-depends critically on the architectural assumptions made by that system.
-Implementors should be free to employ those methods which they find natural and convenient.
-The only exception we make to this rule is in the two matrix equation solvers,
-where we require Gaussian elimination and successive over-relaxation.
-Given the enormous number of matrix solution algorithms in use,
-we felt we had to specify these particular algorithms in order to ensure that
-different implementations of this suite would be comparable.
-
-\subsection{Software Engineering Issues\label{s:method-softeng}}
-
-These toys will exercise many different aspects of parallel systems.
-However,
-their ``single algorithm per program'' model is not representative of real applications.
-These usually contain several discrete (and sometimes overlapping) phases,
-each of which is qualitatively different.
-A full implementation of this suite will therefore have two parts.
-In the first,
-each toy will be implemented as a stand-alone program.
-In the second,
-toys will be chained together to create the kind of phase behaviour seen in real applications
-(Figure~\ref{f:chaining}).
-This will test the ease with which heterogeneous parallelism can be mixed within a single program.
-It will also show how well the system supports code re-use and information hiding,
-which are crucial to the development of large programs.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
-\caption{Chaining\label{f:chaining}}
-\end{figure}
-
-A second reason for chaining toys is that
-many groups are now using algorithmic skeletons to structure parallel programs
-\cite{b:cole-skel,b:p3l-overview,b:enterprise}.
-Such skeletons encapsulate the details of a particular style of parallelism,
-and hide low-level or platform-dependent issues.
-We hope that chaining will show the strengths and weaknesses of such systems.
-
-Finally,
-chaining should be designed so that some toys can be executed concurrently.
-Many PPSs impose extraneous constraints on programs,
-e.g.\ require all processes to participate in every barrier synchronization,
-or require the same executable to be loaded onto each processor.
-These constraints can limit the exploitation of potential concurrency.
-Permitting, but not requiring, concurrent execution of several toys should uncover such limitations.
-
-\subsection{Sizing\label{s:method-size}}
-
-One crucial aspect of the specification of toys is
-the way in which the sizes of problems are determined.
-In a completely frozen model, the actual size of each problem would be compiled into each toy.
-A fluid implementation, by contrast, would allow sizes to be specified at run-time,
-and would only then allocate and partition data structures.
-We have chosen to use an intermediate model (which might be called ``slushy''),
-in which the maximum size of individual problems is specified during compilation,
-but the actual size of a problem is only determined when the toy begins to execute.
-We hope that the use of this model will force implementors to deal with
-at least some of the software engineering issues
-involved in building flexible large-scale applications,
-without making implementation prohibitively difficult.
-
-\subsection{Assessing Complexity\label{s:method-complex}}
-
-There are at least two different ways in which usability might be compared.
-One would be to measure the performance achieved by an ``average'' programmer
-as a function of time on each of a range of problems.
-This would lead to graphs of the kind shown in Figure~\ref{f:perf},
-where the left graph shows a particular result
-and the right graph a simplified abstraction of it.
-By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
-which is the length of vector on which a pipelined architecture achieves
-half its theoretical peak performance,
-we could in principle find the value of $p_{1/2}$---the programming time required to achieve
-half of a machine's peak performance\footnote{Note that
-	if performance was measured as a fraction of the figures quoted by manufacturers for their machines,
-	it is unlikely that the halfway mark would ever be reached.}
-for a particular combination of programming system and problem type.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
-\caption{Performance vs.\ Time\label{f:perf}}
-\end{figure}
-
-A second option would be to measure the complexities of
-implementations of a single application using different programming systems.
-With this approach,
-comparative performance or speedup figures would be supplemented by complexity measures,
-as shown in Figure~\ref{f:complex}.
-
-\begin{figure}
-\epsfxsize=6cm
-\begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
-\caption{Comparative Code Complexity\label{f:complex}}
-\end{figure}
-
-The drawback of the first approach is that there are no ``typical'' problems,
-or ``average'' programmers.
-Meaningful measurement of  $p_{1/2}$ would therefore require
-measurement of a large number of programmers working on many problems.
-Since this is too expensive to be practical,
-the second approach is the one we have chosen to adopt.
-It has drawbacks as well,
-most notably the fact that no-one has ever demonstrated any better measure of program complexity
-than a simple count of the number of lines in a program \cite{b:fenton-metrics}.
-Since we anticipate a wide variation in the types of progamming systems that will be used in this work
-(ranging from declarative dataflow languages
-through functional languages with Lisp-like syntax
-to parallel dialects of Fortran and C),
-and in the indentation, commenting, and layout styles of particular programmers,
-we may measure complexity by counting the number of tokens in the source code which are
-keywords,
-intrinsic operators,
-user-defined functions,
-type or variable declarations,
-and miscellaneous punctuation.
-
-\subsection{Other Uses for Implementations\label{s:method-uses}}
-
-This suite is intended for assessing PPS usability,
-but we envisage at least three other uses for implementations of it:
-\begin{description}
-\item[Related Research:]
-	We hope that the existence of this suite will indicate
-	to the developers of new parallel programming systems
-	what a mature tool should be able to support.
-	If sufficiently diverse,
-	the suite could also be used to help test the correctness of new or improved systems.
-	Finally,
-	this suite could complement existing performance benchmarks,
-	since performance figures for individual toys could be compared.
-\item[Tool Development:]
-	Requests from tool developers for existing parallel programs,
-	or for event traces from them,
-	are common on newsgroups such as {\tt{comp.parallel}}.
-	We hope that traces from implementations of this suite
-	will be used to compare different debugging and performance monitoring tools,
-	that their source code will be used to test
-	new parallelizing compilation techniques or mapping and scheduling algorithms,
-	and so on.
-\item[Teaching:]
-	Implementations will be suitable for use as classroom examples,
-	since they will be small enough to be understood quickly.
-	The toys themselves should also be suitable as classroom exercises
-	in a senior undergraduate course on parallel computing.
-\end{description}
-
-\subsection{Some Criticisms\label{s:method-criticism}}
-
-\noindent {\em{Why not use an existing benchmark suite for this purpose?}\/}
-
-Implementations of benchmark suites have usually been coded for absolute speed.
-They are not representative of ``normal'' practice,
-in which reasonable performance is the usual goal,
-and so are inappropriate for this work.
-In addition,
-most existing benchmark suites are too large to be implemented or ported quickly
-(e.g.\ SPEC~\cite{b:bench-over}),
-over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
-or concerned with a limited range of applications
-(e.g.\ the NAS Parallel Benchmarks~\cite{b:nas-results}).
-Finally,
-most exist only in C or Fortran (or both);
-many interesting parallel programming systems are built on top of other languages,
-including Modula-2, Scheme, Prolog, and various dataflow languages.
-We do not want the effort required to translate several large programs
-to discourage researchers from participating in this exercise.
-
-\vspace{\baselineskip}
-
-\noindent {\em{Won't the results depend primarily on programmer ability?}\/}
-
-The short answer to this question is yes,
-but no more than the results from performance benchmarks.
-A longer answer is that	since we intend to measure the combination of code complexity and performance achieved,
-rather than performance as a function of development time,
-programmers will be allowed to revise and improve their implementations when and as desired.
-
-\section{The Cowichan Problems\label{s:toys}}
-
-This section describes the problems making up the full problem suite.
-As discussed above,
-the implementation of each is to be used both as a stand-alone program
-and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
-Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cl|cl}
-Code	& Toy				& Code	& Toy				\\
-\hline
-E	& elastic net			& O	& outer product			\\
-G	& Gaussian elimination		& P	& matrix-vector product		\\
-H	& halving shuffle		& R	& random matrix generation	\\
-I	& invasion percolation		& S	& successive over-relaxation	\\
-L	& Game of Life			& T	& image thresholding		\\
-M	& Mandelbrot Set		& V	& vector difference		\\
-N	& point normalization		& W	& point value winnowing
-\end{tabular}
-\caption{Single-Character Codes for Toys\label{t:toy-codes}}
-\end{center}
-\end{table}
-
-\subsection{{\tt{mandel}}:
-	Mandelbrot Set Generation
-	\label{s:toys-mandel}}
-
-\begin{figure}
-\begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
-\caption{The Mandelbrot Set\label{f:mandel}}
-\end{figure}
-
-This module generates the Mandelbrot Set
-for a specified region of the complex plane
-(Figure~\ref{f:mandel}).
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{x0, y0}}:]
-	the real coordinates of the lower-left corner
-	of the region to be generated.
-\item[{\tt{dx, dy}}:]
-	the extent of the region to be generated.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing the iteration count at each point in the region.
-\end{description}
-
-Given initial coordinates $(x_0, y_0)$,
-the Mandelbrot Set is generated by iterating the equation
-\begin{eqnarray*}
-x^{\prime}	& =	& x^2 - y^2 + y_0	\\
-y^{\prime}	& =	& 2{x}{y} + x_0
-\end{eqnarray*}
-until either an iteration limit is reached,
-or the values diverge.
-The iteration limit used in this module is 150 steps;
-divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
-The integer value of each element of the matrix is
-the number of iterations done.
-Note that,
-as in all problems,
-the output is required to be
-independent of the number of processors used.
-This requires programmers to be careful about
-floating-point roundoff effects.
-
-\subsection{{\tt{randmat}}:
-	Random Number Generation
-	\label{s:toys-rng}}
-
-This module fills a matrix with pseudo-random integers.
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{s}}:]
-	the random number generation seed.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix filled with random values.
-\end{description}
-Note that,
-as in all problems,
-the output is required to be independent of the number of processors used.
-Generating new seed values from the given seed,
-and running one copy of the random number generator
-on each processor,
-is therefore unlikely to qualify as a solution.
-
-\subsection{{\tt{half}}:
-	Two-Dimensional Shuffle
-	\label{s:toys-half}}
-
-This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
-shuffles them,
-and then repeats this operation along the other axis.
-Values in odd-numbered locations are collected at the low end of each row or column,
-while values in even-numbered locations are moved to the high end.
-An example transformation is:
-
-\begin{eqnarray*}
-\begin{array}{cccc}
-a & b & c & d \\
-e & f & g & h \\
-i & j & k & l
-\end{array}
-& = &
-\begin{array}{cccc}
-a & c & b & d \\
-i & k & j & l \\
-e & g & f & h
-\end{array}
-\end{eqnarray*}
-
-Note that how an array element is moved depends only on whether its location index is odd or even,
-not on whether its value is odd or even.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing shuffled values.
-\end{description}
-
-\subsection{{\tt{invperc}}:
-	Invasion Percolation
-	\label{s:toys-invperc}}
-
-Invasion percolation models the displacement of one fluid
-(such as oil)
-by another
-(such as water)
-in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
-In two dimensions,
-this can be simulated by generating
-an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
-and then marking the center cell of the grid as filled.
-In each iteration,
-one examines the four orthogonal neighbors of all filled cells,
-chooses the one with the lowest value
-(i.e.\ the one with the least resistance to filling),
-and fills it in.
-Figure~\ref{f:invperc1} shows the first few steps in this process,
-while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{math}
-\begin{array}{ccc}
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & 29 & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \star & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & 29 & 77 \\
-\hline
-61 & \star & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \star & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \star & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \star \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & \star & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-\star & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & \star & 72 & 45 & 38 \\
-\hline
-\circ & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \circ \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\end{array}
-\end{math}
-\end{footnotesize}
-\caption{Invasion Percolation\label{f:invperc1}}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
-\end{center}
-\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
-\end{figure}
-
-The simulation continues until
-some fixed percentage of cells have been filled,
-or until some other condition
-(such as the presence of trapped regions)
-is achieved.
-The fractal structure of the filled and unfilled regions
-is then examined to determine how much oil could be recovered.
-The na\"{\i}ve way to implement this is to repeatedly scan the array;
-a more sophisticated, and much faster, sequential technique
-is to maintain a priority queue of unfilled cells
-which are neighbors of filled cells.
-This latter technique is similar to the list-based methods used in some cellular automaton programs,
-and is very difficult to parallelize effectively.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{nfill}}:]
-	the number of points to fill.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
-\end{description}
-Filling begins at the central cell of the matrix
-(rounding down for even-sized axes).
-
-\subsection{{\tt{thresh}}:
-	Histogram Thresholding
-	\label{s:toys-thresh}}
-
-This module performs histogram thresholding on an image.
-Given an integer image $I$ and a target percentage $p$,
-it constructs a binary image $B$
-such that $B_{i,j}$ is set
-if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
-The general idea is that an image's histogram should have 2 peaks,
-one centered around the average foreground intensity,
-and one centered around the average background intensity.
-This program attempts to set
-a threshold between the two peaks in the histogram
-and select the pixels above the threshold.
-This module's inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the integer matrix to be thresholded.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{percent}}:]
-	the minimum percentage of cells to retain.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
-\end{description}
-
-\subsection{{\tt{life}}:
-	Game of Life
-	\label{s:toys-life}}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
-\end{center}
-\caption{Snapshots from the Game of Life\label{f:life}}
-\end{figure}
-
-This module simulates the evolution of Conway's Game of Life,
-a two-dimensional cellular automaton (Figure~\ref{f:life}).
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the Life world.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{numgen}}:]
-	the number of generations to simulate.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the world after simulation.
-\end{description}
-At each time step,
-this module must count the number of live ({\tt{true}}) neighbors
-of each cell,
-using both orthogonal and diagonal connectivity
-and circular boundary conditions.
-The update rule is simple:
-if a cell has 3 live neighbors,
-or has 2 live neighbors and is already alive,
-it is alive in the next generation.
-In any other situation,
-the cell becomes, or stays, dead.
-
-\subsection{{\tt{winnow}}:
-	Weighted Point Selection
-	\label{s:toys-winnow}}
-
-This module converts a matrix of integer values
-to a vector of points,
-represented as $x$ and $y$ coordinates.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix, whose values are used as masses.
-\item[{\tt{mask}}:]
-	a Boolean matrix showing which points are eligible for consideration.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{nelts}}:]
-	the number of points to select.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points.
-\end{description}
-Each location where {\tt{mask}} is {\tt{true}}
-becomes a candidate point,
-with a weight equal to the integer value in {\tt{matrix}} at that location
-and $x$ and $y$ coordinates
-equal to its row and column indices.
-These candidate points are then sorted into increasing order by weight,
-and {\tt{nelts}} evenly-spaced points selected to create the result vector.
-
-\subsection{{\tt{norm}}:
-	Point Location Normalization
-	\label{s:toys-norm}}
-
-This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-Its inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of point locations.
-\item[{\tt{nelts}}:]
-	the number of elements in the point vector.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of normalized point locations.
-\end{description}
-If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
-then the normalization equation is:
-\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
-$y$ coordinates are normalized in the same fashion.
-
-\subsection{{\tt{elastic}}:
-	Elastic Net Simulation
-	\label{s:toys-elastic}}
-
-This module uses the elastic net algorithm \cite{b:elastic-net} to find an approximate solution to the travelling salesentity problem (TSP).
-A closed, circular loop, modelled by a large number of doubly-connected points, is placed in the plane.
-This path is slowly deformed until it connects the points representing the cities to be visited (Figure~\ref{f:elastic}).
-The rate at which the loop is deformed determines the quality of the final path.
-
-\begin{figure}
-\begin{center}
-\framebox{\epsffile{fig/elastic1.ps}}
-\framebox{\epsffile{fig/elastic2.ps}}
-\framebox{\epsffile{fig/elastic3.ps}}
-\end{center}
-\caption{Relaxation of an Elastic Net\label{f:elastic}}
-\end{figure}
-
-Let the (fixed) location of city $i$ be $c_i$, and the (changeable) location of point $j$ on the elastic net be $p_j$.
-The change in the location of point $j$ during each iteration of the algorithm is given by:
-\begin{equation}
-{\delta}{p_j} = {\alpha}\sum_{i}{w_{ij}(c_i - p_j)} ~+~ K{\beta}(p_{j+1} - 2p_j + p_{j-1})
-\end{equation}
-$\alpha$ and $\beta$ are time-invariant factors relating the strengths of the city-to-point and point-to-point forces,
-and $K$ is a time-dependent factor which controls the rate of annealing.
-The strength of the force $w_{ij}$ between city $c_i$ and point $p_j$ is given by:
-\begin{equation}
-w_{ij} = {\phi}({\mid}c_i - p_j{\mid}, K) / \sum_{\ell}\phi({\mid}c_i - p_{\ell}{\mid}, K)
-\end{equation}
-where $\phi(d, K)$ is the Gaussian function $e^{-{d^2}/{K^2}}$.
-Note the normalization of each $w_{ij}$; in order for this algorithm to work properly, the total force being exerted by each city must be the same.
-This is accomplished by summing the forces exerted by each city, and then normalizing the city's forces by that total.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{cities}}:]
-	a vector of points representing the locations of the cities to be toured.
-	The $(x,y)$ coordinates of these cities must be scaled to lie in the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-\item[{\tt{n\_cities}}:]
-	the number of cities in {\tt{cities}}.
-\item[{\tt{n\_net}}:]
-	the number of elastic net points to use.
-\item[{\tt{n\_iters}}:]
-	the total number of iterations to perform.
-\item[{\tt{n\_relax}}:]
-	the number of iterations to perform with each discrete value of $K$.
-	After each {\tt{n\_relax}} iterations, $K$ is reduced by a (constant) pre-defined scaling factor.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{net}}:]
-	a vector of {\tt{n\_net}} points representing the tour found.
-\end{description}
-
-\subsection{{\tt{outer}}:
-	Outer Product
-	\label{s:toys-outer}}
-
-This module turns a vector containing point positions
-into a dense, symmetric, diagonally dominant matrix
-by calculating the distances between each pair of points.
-It also constructs a real vector whose values are
-the distance of each point from the origin.
-Inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
-\item[{\tt{nelts}}:]
-	the number of points in the vector, and the size of the matrix along each axis.
-\end{description}
-Its outputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a real matrix, whose values are filled with inter-point distances.
-\item[{\tt{vector}}:]
-	a real vector, whose values are filled with origin-to-point distances.
-\end{description}
-Each matrix element $M_{i,j}$ such that $i \neq j$
-is given the value $d_{i, j}$,
-the Euclidean distance between point $i$ and point $j$.
-The diagonal values $M_{i, i}$ are then set to
-{\tt{nelts}} times the maximum off-diagonal value
-to ensure that the matrix is diagonally dominant.
-The value of the vector element $v_i$ is set to
-the distance of point $i$ from the origin,
-which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.
-
-\subsection{{\tt{gauss}}:
-	Gaussian Elimination
-	\label{s:toys-gauss}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using explicit reduction.
-(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
-there is a well-formed solution to the equation.)
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{sor}}:
-	Successive Over-Relaxation
-	\label{s:toys-sor}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using successive over-relaxation.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\item[{\tt{tolerance}}:]
-	the solution tolerance, e.g.\ $10^{-6}$.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{product}}:
-	Matrix-Vector Product
-	\label{s:toys-product}}
-
-Given a matrix $A$,
-a vector $V$,
-and an assumed solution $X$ to the equation $AX=V$,
-this module calculates the actual product $AX={V^{\prime}}$,
-and then finds the magnitude of the error.
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{actual}}:]
-	the real vector $V$.
-\item[{\tt{candidate}}:]
-	a real vector containing the supposed solution.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-The output of this function is:
-\begin{description}
-\item[{\tt{e}}:]
-	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
-\end{description}
-
-\section{Other Issues\label{s:issues}}
-
-As described in Section~\ref{s:method-softeng},
-implementors are required to chain modules together in the manner shown in Figure~\ref{f:chaining}.
-This section describes the flow of control and data between these chained problems,
-and discusses several other important aspects of this suite,
-including I/O and visualization.
-
-\subsection{Chained Problem Sequences\label{s:issues-chain}}
-
-The chained implementation of these problems
-executes individual toys in the following order:
-\begin{enumerate}
-\item	An integer matrix $I$ with $r$ rows and $c$ columns is created
-	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
-	\label{e:chain-choice-1}
-\item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
-\item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
-	to generate a Boolean mask $B$ from $I$ in which the (minimum) percentage of {\tt{true}} locations is $P$.
-	Like $I$, $B$ has $r$ rows and $c$ columns.
-	\label{e:chain-choice-2}
-\item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
-	This step overwrites the Boolean matrix $B$.
-\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
-\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
-\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
-\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
-	as described in Section~\ref{s:toys-outer}.
-\item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
-	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
-	This measures the agreement between the solutions found by the two methods.
-\end{enumerate}
-Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
-there are four possible chained sequences.
-A minimal implementation will provide one of these,
-while a full implementation will be a single program which can execute any of the four sequences.
-Intermediate implementations providing all four sequences as separate programs are also acceptable.
-
-\subsection{Input and Output\label{s:issues-io}}
-
-I/O is an important part of programming,
-but is often treated as being of secondary importance by language designers.
-This suite requires all stand-alone toys to read input values from files, and write results back;
-the chained version must be able to checkpoint intermediate results between toys.
-Finally,
-implementors are strongly encouraged to include some means of visualizing the output or evolution of individual toys.
-
-The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
-A file containing a vector begins with 1 positive integer,
-which specifies the number of elements in the vector;
-a file containing a matrix begins with 2 positive integers,
-which specify the number of rows and columns in the matrix respectively.
-Elements of the vector or matrix then appear one per line,
-from least-indexed to greatest-indexed
-(and in row-major order for matrices).
-
-The formats described above are required to be human-readable.
-Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
-using whatever file formats are convenient.
-This will allow programmers to demonstrate the ``natural'' I/O capabilities of particular systems
-which most probably be used for checkpointing intermediate results in real programs.
-
-\begin{figure}
-\epsfxsize=0.75\textwidth
-\begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
-\caption{Cowichan File Formats\label{f:io-formats}}
-\end{figure}
-
-\subsection{Visualization\label{s:issues-vis}}
-
-Visualization is an increasingly important part of scientific and numerical computing.
-Including it as an optional part of this problem set is also a good way to exercise both
-the use of libraries within a parallel programming system,
-and the ease with which a programming system can support variant versions of a program.
-
-Our requirements for the graphics interface are:
-\begin{enumerate}
-\item	Visualization for individual toys should be as simple as possible,
-	and should require as few changes to the basic toy code as possible.
-\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
-	The latter should create all images in a single window.
-\item	A single source file should implement both the graphical and non-graphical versions of each toy.
-\end{enumerate}
-The first two points require the system to distinguish between the actual drawing code
-(e.g.\ drawing a vector of points as circles)
-and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
-While the stand-alone toys use a whole window for drawing,
-the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
-At present,
-this is implemented by using a call counter
-in the interfaces for the individual toys and in the chained module.
-The actual drawing routines called from within each toy could then be the same.
-
-The present visualization routines were originally written by Anil Sukul,
-a student at the Vrije Universiteit, Amsterdam.
-The package uses the public domain graphics library VOGLE,
-available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
-VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
-as well as screen control routines (such as double buffering).
-VOGLE was chosen because of its simplicity,
-and because it includes Pascal and Fortran interfaces.
-
-\subsection{Reproducibility\label{s:issues-reproduce}}
-
-Reproducibility is an important issue for parallel programming systems.
-While constraining the order of operations in a parallel system to guarantee reproducibility makes programs in that system easier to debug,
-it can also reduce the expressiveness or performance of the system.
-
-In this problem set, irreproducibility can appear in two forms: numerical and algorithmic.
-The first arises in toys such as {\tt{gauss}}, {\tt{sor}}, and {\tt{elastic}}, which use floating-point numbers.
-Precisely how round-off errors occur in these calculations can depend on the distribution of work among processors,
-or the order in which those processors carry out particular operations.
-
-Irreproducibility also arises in toys which only use exact numbers, such as {\tt{invperc}} and {\tt{randmat}}.
-In the former, the percolation region is grown by repeatedly filling the lowest-valued cell on its perimeter.
-If several cells have this value, implementations may choose one arbitrarily.
-Thus, different implementations may produce very different fractal shapes.
-In the case of random number generation, the simplest thing to do is to run the same generator independently on each processor,
-although the values in the resulting matrix then depend on the number of processors used.
-
-\section{Conclusion\label{s:conclude}}
-
-A full ANSI~C version of this problem suite has been written, and is available by anonymous FTP from:
-\begin{center}
-{\tt{ftp.cs.toronto.edu :: pub/gvw/cowichan/src.tar.gz}}
-\end{center}
-This release includes two parallel versions,
-both based on the POSIX threads package.
-The first uses repeated fork-and-join parallelism,
-while the second creates a fixed number of threads at the beginning of the program,
-then synchronizes them repeatedly using barriers.
-
-We would like to invite interested groups to participate in this project by parallelizing the problem suite using their favourite language or tool.
-Our aim is to make both the implementations and their descriptions available as a hypertext document through the World-Wide Web,
-and, if there is sufficient interest, to publish this material in book form.
-If you would like to contribute to this effort, please contact either of the first two authors.
-
-\bibliographystyle{plain}
-\newcommand{\bibea}{{\em{et~al.}}}
-\bibliography{paper}
-
-\appendix
-
-\newpage
-\section{Parallel Clich\'{e}s\label{s:cliche}}
-
-\subsection{Operations\label{s:cliche-ops}}
-
-This list details some operations which are supported by many parallel programming systems.
-The toys described in Section~\ref{s:toys} provide opportunities for exercising many of these,
-and implementors are encouraged to phrase discussion of their work in terms of these operations where appropriate.
-
-\begin{itemize}
-\item	elementwise operations on arrays (unary, binary, and scalar promotion)
-\item	cumulative operations on arrays (reduction and parallel prefix)
-\item	array re-shaping and re-sizing (e.g.\ sub-array extraction)
-\item	partial (conditionally masked) versions of all of the above
-\item	regular data motion (e.g.\ circular and planar shifting)
-\item	irregular data motion (e.g.\ 1-to-1, 1-to-many, and many-to-1 permutation)
-\item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
-\item	differential local addressing (i.e.\ subscripting an array with an array)
-\item	full or partial replication of shared read-only values
-\item	full or partial replication of shared read-mostly values with automatic consistency management
-\item	structures with rotating ownership, suitable for migration
-\item	producer-consumer structures
-\item	partitionable structures
-\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
-\item	dynamic re-partitioning (e.g.\ for load balancing)
-\item	committed mutual exclusion (e.g.\ waiting on a semaphore)
-\item	uncommitted mutual exclusion (e.g.\ lock or fail)
-\item	barrier synchronization
-\item	multiple concurrent barriers used by non-overlapping groups of processes
-\item	fetch-and-add, and other fetch-and-operate functions
-\item	pre-scheduled receipt of messages of known composition
-\item	variable-length message receipt
-\item	message selection by type
-\item	message selection by source
-\item	message selection by contents
-\item	guarded message selection (e.g.\ CSP's {\tt{alt}} construct)
-\item	broadcast and partial broadcast
-\item	split-phase (non-blocking) operations
-\item	data marshalling and unmarshalling of ``flat'' structures (e.g.\ arrays of scalars)
-\item	data marshalling and unmarshalling of nested or linked structures
-\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
-\item	pipelining
-\item	distributed linked structures in systems without physically-shared memory
-\item	indexing of distributed shared arrays
-\item	collective I/O
-\item	uniform-size interleaved I/O
-\item	heterogeneous-size interleaved I/O
-\item	independent I/O operations on a single shared file
-\end{itemize}
-
-\subsection{Memory Reference Patterns\label{s:cliche-memref}}
-
-These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
-Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.
-
-\begin{itemize}
-\item	Write-once:
-	variables which are assigned a value once, and only read thereafter.
-	These can be supported through replication.
-\item	Write-many:
-	variables which are written and read many times.
-	If several processes write to the variable concurrently, they typically write to different portions of it.
-\item	Producer-consumer:
-	variables which are written to by one or more objects, and read by one or more other objects.
-	Entries in a shared queue, or a mailbox used for communication between two processes, are examples.
-\item	Private:
-	variables which are potentially shared, but actually private.
-	The interior points of a mesh in a program which uses geometric decomposition fall into this category,
-	while boundary points belong to the previous class.
-\item	Migratory:
-	variables which are read and written many times in succession by a single process before ownership passes to another process.
-	The structures representing particles in an $N$-body simulation are the clearest example of this category.
-\item	Result:
-	Accumulators which are updated many times by different processes, but thereafter read.
-\item	Read-mostly:
-	Any variable which is read much more often than it is written to.
-	The best score so far in a search problem is a typical instance of this class:
-	while any process might update it, in practice processes read its value much more often than they change it.
-\item	Synchronization:
-	Variables used to force explicit synchronization in a program, such as semaphores and locks.
-	These are typically ignored for long periods, and then the subject of intense bursts of access activity.
-\item	General read-write:
-	Any variable which cannot be put in one of the above categories.
-\end{itemize}
-
-\newpage
 \section{Comparing Parallel Programming Systems\label{s:questions}}
@@
 \end{enumerate}
-
-\newpage
-\section{A Critique of the C Implementation\label{s:critique}}
-
-This section is a critique of the ANSI~C reference implementation of the problem suite.
-It is intended to serve as a model for (self-)criticism of other programming systems.
-
-\begin{itemize}
-\item	C's support for multi-dimensional arrays (MDAs) is very weak.
-	There is no way to dynamically allocate an MDA in a single go---one must either
-	allocate a block of the same size as the desired array,
-	and then do indexing calculations by hand,
-	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
-	MDAs do not carry dimension information with them,
-	so it is impossible to determine the size of an array parameter within a function.
-	Finally,
-	C does not treat all axes of an array equally:
-	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
-	it is impossible to slice it along the other axis.
-\item	C does not distinguish between Boolean and integer types.
-	As a result,
-	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
-	can be passed in reverse order without a type error.
-	Using {\tt{typedef}} to create a Boolean type name does not solve this
-	(at least, not in {\tt{gcc}} V2.5.8).
-\item	Union types cannot be safely initialized.
-	This complicated the implementation of the graphics interface,
-	where it would have been much more elegant to define a union type,
-	each of whose variants held parameter specifications for a single toy.
-	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
-	initializing some and filling other with don't-care values.
-	This is neither safe nor elegant.
-\item	Parameter values cannot be used in the initialization of local variables inside functions.
-	In particular,
-	it is not possible to create a local vector with a length specified by an input parameter.
-	Such a facility would be useful in the {\tt{winnow}} toy,
-	where we have instead allocated local temporaries of the maximum possible size.
-\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
-	For example, the ``fail()'' error-handling routine was buggy because
-	all \verb`real` arguments were being automatically converted to doubles,
-	but were being taken off the stack as floats.
-\item	no intrinsic notion of group ID/self ID in threads
-\item	packaging parameters for threads
-\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
-\end{itemize}
-
-\end{document}

### 2009-08-28 paper/2009-09/concl.tex
Breaking paper into manageable chunks

--- 2009-08-20 paper/2009-09/paper.tex
@@
-\documentstyle[epsf,11pt]{article}
-
-\begin{document}
-
-\begin{Large}\begin{center}
-The Cowichan Problems: \\
-Assessing the Usability of Parallel Programming Systems
-\end{center}\end{Large}
-
-\vspace{\baselineskip}
-
-\begin{center}
-\begin{tabular}{cccc}
-Gregory V.~Wilson\footnote{\verb`http://www.third-bit.com`}
-&
-Andrew Borzenko
-&
-Cameron Gorrie
-&
-Zachary Kincaid
-\end{tabular}
-\end{center}
-
-\vspace{\baselineskip}
-
-\begin{abstract}
-Most programmers believe that parallel programming is harder than sequential programming.
-Many also believe that today's parallel programming systems make it even harder.
-We have developed a suite of chained toy problems that can be used to assess
-the usability of different scientific programming languages, parallel or otherwise.
-This paper motivates and describes the suite,
-and discusses what we have learned from porting it to four parallel C++ systems
-(MPI, OpenMP, Intel Thread Building Blocks, and a Linux tuple space system),
-and from translating it into sequential and parallel Haskell.
-\end{abstract}
-
-\section{Introduction\label{s:intro}}
-
-Twenty years ago,
-Bal {\em{et~al.}\/} listed more than 300 parallel programming systems \cite{b:par-lang-survey}.
-A similar survey today could well uncover twice as many,
-but a poll of application programmers would probably find that
-only a handful are in general use.
-
-The reasons why so many of these PPSs are only used by their developers (if at all) are murkier.
-Part of the blame undoubtedly lies with the diversity of hardware,
-and the attendant lack of portability,
-alluded to above.
-A second reason is that an academic researcher's principal interest is developing new languages,
-not supporting existing ones.
-By the time a system is 90\% complete,
-its author is already thinking about its successor.
-Few application programmers are willing to invest effort porting codes
-to systems which are not properly supported,
-and liable to change or evaporate without notice.
-Third,
-many research groups have chosen to explore the potential of parallelism
-without consideration for compatibility with older languages.
-Application programmers must therefore face the prospect of completely re-writing their programs,
-rather than simply porting them.
-As the next section argues,
-this can obviate any performance that might be gained from parallelization.
-Finally,
-many PPSs simply cannot support the software engineering load of large applications.
-Many provide only a single global namespace,
-do not include I/O capabilities,
-require users to marshal and unmarshal data explicitly,
-or do not permit hardware to be timeshared.
-
-The net result is that many PPSs are never used for anything more complicated than
-generating the Mandelbrot Set,
-simulating the Game of Life,
-multiplying dense matrices,
-or finding all solutions to the $N$-queens problem for some small $N$.
-Lacking the feedback that comes from more demanding use,
-few systems are improvements over their predecessors
-in the way that successive versions of operating systems have improved on one another,
-or the languages in the Algol$\rightarrow$Modula-$n$ family
-have become both more powerful and more secure.
-Most applications programmers therefore choose to play it safe
-and stick to parallel programming's lowest common denominators:
-data-parallel Fortran,
-and message-passing libraries.
-
-The historical reasons for tolerating this situation are becoming less valid with each passing year.
-When only a few parallel computers existed,
-and these were of widely different types,
-software standardization simply didn't make sense.
-Similarly,
-when supercomputers were very expensive,
-the cost of software development seemed relatively smaller,
-and the requirement for maximum performance relatively greater,
-than in sequential computing.
-As \cite{b:social-limits-speed} points out,
-however, these arguments are losing force.
-Now,
-as in sequential computing,
-the first question is not ``How do I write programs for this machine?'',
-but ``Will this machine run my programs?''
-As microprocessor-based MIMD hardware comes to dominate massively-parallel computing,
-we see increasing emphasis on standards such as High-Performance Fortran (HPF) \cite{b:hpf-overview},
-and the Message-Passing Interface (MPI) \cite{b:mpi-overview}.
-
-But herein lies a danger.
-History shows that it is extremely difficult
-to displace a programming language or system once it becomes established.
-Cobol and MS-DOS are the most frequently cited examples,
-but many systems which should have been superceded long ago have managed
-not only to cling to life,
-but to grow in influence.
-(The fact that the mail signature
-``The last good thing written in C was Schubert's Ninth Symphony''
-can elicit smiles simply proves that even the truth can be funny.)
-Established systems evolve slowly, if at all.
-It would therefore seem that the best time to guarantee ourselves
-a good parallel programming paradigm is right now,
-before a large number of large applications are created and take on lives of their own.
-
-We therefore propose to use a set of simple programming problems to compare
-the usability of parallel programming systems.
-Competent programmers, fluent in a specific system,
-will implement solutions to these problems and report their experiences
-in terms of development time, code size and clarity, and runtime efficiency.
-These problems are on the scale of the toy applications described several paragraphs ago
-both to keep implementation time low
-and because we feel that any PPS produced today ought to be able to support such simple programs.
-As the second author put it,
-``We're all sick of seeing these,
-but maybe if we all catch the disease at the same time we'll finally build up some immunity to it.''
-One significant innovation in this work is that
-we also require implementations of our suite to chain problems together,
-so that the output of one problem is the input to another.
-This will allow us to assess how well each PPS supports modularization and code re-use
-(or, more realistically, how much extra effort programmers must invest
-to achieve these goals when using a particular PPS).
-
-Section~\ref{s:corollary} motivates this work by showing
-how poor usability can limit the realizable parallelism in a program.
-Section~\ref{s:previous} surveys previous work on the comparison of parallel programming systems,
-while Section~\ref{s:method} outlines our choice of methods, and the constraints on our work.
-Section~\ref{s:toys} then presents the suite of problems we intend to use.
-In recognition of the original work done by Feo and others on
-the Salishan Problems~\cite{b:salishan}, this suite is called
-the Cowichan Problems\footnote{\noindent Pronounced {\em{Cow}\/}-i-chun.
-	Like Salishan, the word is a Northwest Coast Indian place name.}.
-
-\subsection*{History and Acknowledgments}
-
-The first author originally developed this suite in the mid-1990s,
-inspired both by \cite{b:salishan} and by
-discussions with R.\ Bruce Irvin (then at the University of Wisconsin),
-Henri Bal (Vrije Universiteit, Amsterdam),
-and others.
-Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.
-
-\section{Quantifying the Importance of Usability\label{s:corollary}}
-
-According to Amdahl's Law,
-the time $t$ required to execute a program has two components:
-$t_{seq}$, representing sequential operations that cannot be parallelized,
-and $t_{par}$, representing those operations which can.
-If we let $\sigma$ be the ``serial fraction'' of the program,
-so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
-then the speedup which can be achieved using $P$ processors is:
-\begin{eqnarray*}
-s(P)	& =	& \frac{t(1)}{t(P)}	\\
-	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
-\end{eqnarray*}
-As $P{\rightarrow}{\infty}$,
-the potential speedup is bounded by $1/{\sigma}$.
-
-What is not usually acknowledged is
-the influence of programming system usability on
-the potential speedup of sequential programs.
-As stated above, Amdahl's Law applies only to a single run of a program.
-In the real world, programs are first developed, and then run many times.
-If we let $T$ be the total running time of a program over its lifetime,
-$D_{seq}$ be the time required to develop the sequential components of a program,
-and $D_{par}$ be the time required to develop the parallel components
-(or parallelize the sequential components),
-then the lifetime speedup we can achieve is given by:
-\begin{eqnarray*}
-S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
-	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
-\end{eqnarray*}
-Assume that we are parallelizing a ``dusty deck'',
-i.e.\ that the time $D_{seq}$ has already been invested.
-If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
-then achievable speedup is limited by:
-\begin{eqnarray*}
-S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
-		& =	& \frac{1}{{\phi} + {\sigma}}
-\end{eqnarray*}
-The time required to parallelize a program can therefore be seen
-as increasing that program's effective serial fraction.
-Unless parallelization time can be substantially reduced,
-massively-parallel computing is likely to remain attractive only in those cases in which:
-\begin{itemize}
-\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
-\item	the expected total program runtime is very large
-	(i.e.\ the program is a package used extensively in a field
-	such as computational chemistry or vehicle crash simulation);
-	or
-\item	cost is not an obstacle (i.e.\ the customer is the NSA).
-\end{itemize}
-
-\section{Previous Work\label{s:previous}}
-
-In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
-\begin{quote}
-As practiced by computer science,
-the study of programming is an unholy mixture of mathematics,
-literary criticism,
-and folklore.
-\end{quote}
-Little has changed since.
-Computer scientists pay lip service to the importance of usability,
-but are strangely reluctant to try to quantify it,
-particularly with respect to the tools they use themselves.
-Religious debates about the ``right'' first-year teaching language
-or the ``best'' GUI toolkit are common;
-systematic use of the experimental method is not.
-
-Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
-such as readability and code-pattern recognition
-\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
-The little work that has been done on ``programming in the large'' has been anecdotal,
-such as \cite{b:compare-ada-c-pascal}.
-This book presented various authors' impressions of Pascal, C, and Ada
-(before any complete Ada implementations existed).
-Its editors formulated a list of questions to use in comparing languages,
-but did not use it themselves.
-A modified version of this questionnaire,
-tailored for parallel programming languages,
-is given in Appendix~\ref{s:questions},
-and will be used as a guideline in our work.
-
-Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
-The first presented implementations of a simple numerical quadrature program
-in more than a dozen different parallel languages
-in use on mid-1980s hardware.
-The second presented implementations of
-the Salishan Problems---Hamming number generation,
-isomer enumeration,
-skyline matrix reduction,
-and a simple discrete event simulator---in
-C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
-and a variety of dataflow and functional languages.
-Both of these books convey the flavor of the languages they describe,
-but neither made any effort to compare languages or problem implementations directly.
-
-Inspired by the Salishan Problems,
-the first author developed a suite of seven problems \cite{b:cowichan-1}
-in order to test the limits of parallel programming systems.
-Individual problems were badly load-balanced,
-manipulated irregular pointer-based structures,
-and so on.
-Our experience implementing them was extremely valuable,
-as it uncovered many bugs in the programming system used,
-and revealed several shortcomings of the system's design.
-However,
-we also discovered that implementing each problem takes six to eight weeks,
-which we feel is prohibitive at this exploratory stage.
-
-\section{Methodology\label{s:method}}
-
-The main aim of this work is to assess and compare
-the ease with which programs can be developed and/or ported
-using different parallel programming systems.
-Here,
-we discuss issues related to choosing the problems that make up the suite,
-assessing the complexity of particular implementations,
-and other purposes which implementations of these might serve.
-Recognizing that these problems are extremely simple,
-and easy to parallelize,
-we refer to them as ``toys''.
-
-\subsection{Criteria for Selection\label{s:method-criteria}}
-
-Our criteria for including problems in this suite are as follows:
-\begin{enumerate}
-\item	Each toy should require no more than an afternoon
-	to write and test in a well-supported sequential language.
-	We feel that making individual problems more complicated will only discourage implementation.
-\item	The correctness of each implementation must be relatively easy to verify.
-	Toys whose output is easily visualized are therefore to be preferred,
-	as are toys whose results are insensitive to floating-point discretization effects.
-\item	Each toy application should be amenable to
-	an analysis of speedup or isoefficiency \cite{b:isoefficiency-tutorial}
-	to allow a comparison of theory and practice.
-\item	At least some toys should not be ``infinitely scalable''.
-	Many real-world applications are not,
-	and this suite should reflect such limitations.
-\item	At least some toys should require I/O,
-	since this important aspect of real-world programming is often neglected by PPS designers.
-\item	There should be some potential redundancy in the type of parallelization done,
-	so that re-use of software can be demonstrated.
-\item	Together,
-	the toys in the suite must exercise a wide range of common parallel operations
-	(Appendix~\ref{s:cliche-ops})
-	and memory reference patterns
-	(Appendix~\ref{s:cliche-memref}).
-\item	The ``obvious'' implementations of these toys should span
-	a wide range of parallel programming paradigms,
-	such as task farming,
-	geometric and functional decomposition,
-	and speculative parallelism.
-\item	Parallelizations for several (if not all) of
-	the types of programming systems shown in Figure~\ref{f:taxonomy} should be known.
-	This suite is not intended to be a collection of research problems;
-	reasonably simple and efficient implementations in most environments
-	should be straightforward to produce.
-\end{enumerate}
-
-\begin{figure}
-\epsfysize=0.5\textheight
-\begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
-\caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
-\end{figure}
-
-\subsection{Specification\label{s:method-spec}}
-
-Toys should be specified in terms of inputs and outputs,
-rather than algorithmically.
-For example,
-``sort $N$ integers in the range $1{\ldots}R$,
-where $R$ is not known in advance and there may be multiple instances of a given value''
-is to be preferred over
-``parallelize quicksort''.
-As \cite{b:slalom} argued,
-which parallel algorithm is best in a particular programming system
-depends critically on the architectural assumptions made by that system.
-Implementors should be free to employ those methods which they find natural and convenient.
-The only exception we make to this rule is in the two matrix equation solvers,
-where we require Gaussian elimination and successive over-relaxation.
-Given the enormous number of matrix solution algorithms in use,
-we felt we had to specify these particular algorithms in order to ensure that
-different implementations of this suite would be comparable.
-
-\subsection{Software Engineering Issues\label{s:method-softeng}}
-
-These toys will exercise many different aspects of parallel systems.
-However,
-their ``single algorithm per program'' model is not representative of real applications.
-These usually contain several discrete (and sometimes overlapping) phases,
-each of which is qualitatively different.
-A full implementation of this suite will therefore have two parts.
-In the first,
-each toy will be implemented as a stand-alone program.
-In the second,
-toys will be chained together to create the kind of phase behaviour seen in real applications
-(Figure~\ref{f:chaining}).
-This will test the ease with which heterogeneous parallelism can be mixed within a single program.
-It will also show how well the system supports code re-use and information hiding,
-which are crucial to the development of large programs.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
-\caption{Chaining\label{f:chaining}}
-\end{figure}
-
-A second reason for chaining toys is that
-many groups are now using algorithmic skeletons to structure parallel programs
-\cite{b:cole-skel,b:p3l-overview,b:enterprise}.
-Such skeletons encapsulate the details of a particular style of parallelism,
-and hide low-level or platform-dependent issues.
-We hope that chaining will show the strengths and weaknesses of such systems.
-
-Finally,
-chaining should be designed so that some toys can be executed concurrently.
-Many PPSs impose extraneous constraints on programs,
-e.g.\ require all processes to participate in every barrier synchronization,
-or require the same executable to be loaded onto each processor.
-These constraints can limit the exploitation of potential concurrency.
-Permitting, but not requiring, concurrent execution of several toys should uncover such limitations.
-
-\subsection{Sizing\label{s:method-size}}
-
-One crucial aspect of the specification of toys is
-the way in which the sizes of problems are determined.
-In a completely frozen model, the actual size of each problem would be compiled into each toy.
-A fluid implementation, by contrast, would allow sizes to be specified at run-time,
-and would only then allocate and partition data structures.
-We have chosen to use an intermediate model (which might be called ``slushy''),
-in which the maximum size of individual problems is specified during compilation,
-but the actual size of a problem is only determined when the toy begins to execute.
-We hope that the use of this model will force implementors to deal with
-at least some of the software engineering issues
-involved in building flexible large-scale applications,
-without making implementation prohibitively difficult.
-
-\subsection{Assessing Complexity\label{s:method-complex}}
-
-There are at least two different ways in which usability might be compared.
-One would be to measure the performance achieved by an ``average'' programmer
-as a function of time on each of a range of problems.
-This would lead to graphs of the kind shown in Figure~\ref{f:perf},
-where the left graph shows a particular result
-and the right graph a simplified abstraction of it.
-By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
-which is the length of vector on which a pipelined architecture achieves
-half its theoretical peak performance,
-we could in principle find the value of $p_{1/2}$---the programming time required to achieve
-half of a machine's peak performance\footnote{Note that
-	if performance was measured as a fraction of the figures quoted by manufacturers for their machines,
-	it is unlikely that the halfway mark would ever be reached.}
-for a particular combination of programming system and problem type.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
-\caption{Performance vs.\ Time\label{f:perf}}
-\end{figure}
-
-A second option would be to measure the complexities of
-implementations of a single application using different programming systems.
-With this approach,
-comparative performance or speedup figures would be supplemented by complexity measures,
-as shown in Figure~\ref{f:complex}.
-
-\begin{figure}
-\epsfxsize=6cm
-\begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
-\caption{Comparative Code Complexity\label{f:complex}}
-\end{figure}
-
-The drawback of the first approach is that there are no ``typical'' problems,
-or ``average'' programmers.
-Meaningful measurement of  $p_{1/2}$ would therefore require
-measurement of a large number of programmers working on many problems.
-Since this is too expensive to be practical,
-the second approach is the one we have chosen to adopt.
-It has drawbacks as well,
-most notably the fact that no-one has ever demonstrated any better measure of program complexity
-than a simple count of the number of lines in a program \cite{b:fenton-metrics}.
-Since we anticipate a wide variation in the types of progamming systems that will be used in this work
-(ranging from declarative dataflow languages
-through functional languages with Lisp-like syntax
-to parallel dialects of Fortran and C),
-and in the indentation, commenting, and layout styles of particular programmers,
-we may measure complexity by counting the number of tokens in the source code which are
-keywords,
-intrinsic operators,
-user-defined functions,
-type or variable declarations,
-and miscellaneous punctuation.
-
-\subsection{Other Uses for Implementations\label{s:method-uses}}
-
-This suite is intended for assessing PPS usability,
-but we envisage at least three other uses for implementations of it:
-\begin{description}
-\item[Related Research:]
-	We hope that the existence of this suite will indicate
-	to the developers of new parallel programming systems
-	what a mature tool should be able to support.
-	If sufficiently diverse,
-	the suite could also be used to help test the correctness of new or improved systems.
-	Finally,
-	this suite could complement existing performance benchmarks,
-	since performance figures for individual toys could be compared.
-\item[Tool Development:]
-	Requests from tool developers for existing parallel programs,
-	or for event traces from them,
-	are common on newsgroups such as {\tt{comp.parallel}}.
-	We hope that traces from implementations of this suite
-	will be used to compare different debugging and performance monitoring tools,
-	that their source code will be used to test
-	new parallelizing compilation techniques or mapping and scheduling algorithms,
-	and so on.
-\item[Teaching:]
-	Implementations will be suitable for use as classroom examples,
-	since they will be small enough to be understood quickly.
-	The toys themselves should also be suitable as classroom exercises
-	in a senior undergraduate course on parallel computing.
-\end{description}
-
-\subsection{Some Criticisms\label{s:method-criticism}}
-
-\noindent {\em{Why not use an existing benchmark suite for this purpose?}\/}
-
-Implementations of benchmark suites have usually been coded for absolute speed.
-They are not representative of ``normal'' practice,
-in which reasonable performance is the usual goal,
-and so are inappropriate for this work.
-In addition,
-most existing benchmark suites are too large to be implemented or ported quickly
-(e.g.\ SPEC~\cite{b:bench-over}),
-over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
-or concerned with a limited range of applications
-(e.g.\ the NAS Parallel Benchmarks~\cite{b:nas-results}).
-Finally,
-most exist only in C or Fortran (or both);
-many interesting parallel programming systems are built on top of other languages,
-including Modula-2, Scheme, Prolog, and various dataflow languages.
-We do not want the effort required to translate several large programs
-to discourage researchers from participating in this exercise.
-
-\vspace{\baselineskip}
-
-\noindent {\em{Won't the results depend primarily on programmer ability?}\/}
-
-The short answer to this question is yes,
-but no more than the results from performance benchmarks.
-A longer answer is that	since we intend to measure the combination of code complexity and performance achieved,
-rather than performance as a function of development time,
-programmers will be allowed to revise and improve their implementations when and as desired.
-
-\section{The Cowichan Problems\label{s:toys}}
-
-This section describes the problems making up the full problem suite.
-As discussed above,
-the implementation of each is to be used both as a stand-alone program
-and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
-Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cl|cl}
-Code	& Toy				& Code	& Toy				\\
-\hline
-E	& elastic net			& O	& outer product			\\
-G	& Gaussian elimination		& P	& matrix-vector product		\\
-H	& halving shuffle		& R	& random matrix generation	\\
-I	& invasion percolation		& S	& successive over-relaxation	\\
-L	& Game of Life			& T	& image thresholding		\\
-M	& Mandelbrot Set		& V	& vector difference		\\
-N	& point normalization		& W	& point value winnowing
-\end{tabular}
-\caption{Single-Character Codes for Toys\label{t:toy-codes}}
-\end{center}
-\end{table}
-
-\subsection{{\tt{mandel}}:
-	Mandelbrot Set Generation
-	\label{s:toys-mandel}}
-
-\begin{figure}
-\begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
-\caption{The Mandelbrot Set\label{f:mandel}}
-\end{figure}
-
-This module generates the Mandelbrot Set
-for a specified region of the complex plane
-(Figure~\ref{f:mandel}).
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{x0, y0}}:]
-	the real coordinates of the lower-left corner
-	of the region to be generated.
-\item[{\tt{dx, dy}}:]
-	the extent of the region to be generated.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing the iteration count at each point in the region.
-\end{description}
-
-Given initial coordinates $(x_0, y_0)$,
-the Mandelbrot Set is generated by iterating the equation
-\begin{eqnarray*}
-x^{\prime}	& =	& x^2 - y^2 + y_0	\\
-y^{\prime}	& =	& 2{x}{y} + x_0
-\end{eqnarray*}
-until either an iteration limit is reached,
-or the values diverge.
-The iteration limit used in this module is 150 steps;
-divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
-The integer value of each element of the matrix is
-the number of iterations done.
-Note that,
-as in all problems,
-the output is required to be
-independent of the number of processors used.
-This requires programmers to be careful about
-floating-point roundoff effects.
-
-\subsection{{\tt{randmat}}:
-	Random Number Generation
-	\label{s:toys-rng}}
-
-This module fills a matrix with pseudo-random integers.
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{s}}:]
-	the random number generation seed.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix filled with random values.
-\end{description}
-Note that,
-as in all problems,
-the output is required to be independent of the number of processors used.
-Generating new seed values from the given seed,
-and running one copy of the random number generator
-on each processor,
-is therefore unlikely to qualify as a solution.
-
-\subsection{{\tt{half}}:
-	Two-Dimensional Shuffle
-	\label{s:toys-half}}
-
-This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
-shuffles them,
-and then repeats this operation along the other axis.
-Values in odd-numbered locations are collected at the low end of each row or column,
-while values in even-numbered locations are moved to the high end.
-An example transformation is:
-
-\begin{eqnarray*}
-\begin{array}{cccc}
-a & b & c & d \\
-e & f & g & h \\
-i & j & k & l
-\end{array}
-& = &
-\begin{array}{cccc}
-a & c & b & d \\
-i & k & j & l \\
-e & g & f & h
-\end{array}
-\end{eqnarray*}
-
-Note that how an array element is moved depends only on whether its location index is odd or even,
-not on whether its value is odd or even.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing shuffled values.
-\end{description}
-
-\subsection{{\tt{invperc}}:
-	Invasion Percolation
-	\label{s:toys-invperc}}
-
-Invasion percolation models the displacement of one fluid
-(such as oil)
-by another
-(such as water)
-in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
-In two dimensions,
-this can be simulated by generating
-an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
-and then marking the center cell of the grid as filled.
-In each iteration,
-one examines the four orthogonal neighbors of all filled cells,
-chooses the one with the lowest value
-(i.e.\ the one with the least resistance to filling),
-and fills it in.
-Figure~\ref{f:invperc1} shows the first few steps in this process,
-while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{math}
-\begin{array}{ccc}
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & 29 & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \star & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & 29 & 77 \\
-\hline
-61 & \star & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \star & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \star & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \star \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & \star & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-\star & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & \star & 72 & 45 & 38 \\
-\hline
-\circ & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \circ \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\end{array}
-\end{math}
-\end{footnotesize}
-\caption{Invasion Percolation\label{f:invperc1}}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
-\end{center}
-\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
-\end{figure}
-
-The simulation continues until
-some fixed percentage of cells have been filled,
-or until some other condition
-(such as the presence of trapped regions)
-is achieved.
-The fractal structure of the filled and unfilled regions
-is then examined to determine how much oil could be recovered.
-The na\"{\i}ve way to implement this is to repeatedly scan the array;
-a more sophisticated, and much faster, sequential technique
-is to maintain a priority queue of unfilled cells
-which are neighbors of filled cells.
-This latter technique is similar to the list-based methods used in some cellular automaton programs,
-and is very difficult to parallelize effectively.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{nfill}}:]
-	the number of points to fill.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
-\end{description}
-Filling begins at the central cell of the matrix
-(rounding down for even-sized axes).
-
-\subsection{{\tt{thresh}}:
-	Histogram Thresholding
-	\label{s:toys-thresh}}
-
-This module performs histogram thresholding on an image.
-Given an integer image $I$ and a target percentage $p$,
-it constructs a binary image $B$
-such that $B_{i,j}$ is set
-if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
-The general idea is that an image's histogram should have 2 peaks,
-one centered around the average foreground intensity,
-and one centered around the average background intensity.
-This program attempts to set
-a threshold between the two peaks in the histogram
-and select the pixels above the threshold.
-This module's inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the integer matrix to be thresholded.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{percent}}:]
-	the minimum percentage of cells to retain.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
-\end{description}
-
-\subsection{{\tt{life}}:
-	Game of Life
-	\label{s:toys-life}}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
-\end{center}
-\caption{Snapshots from the Game of Life\label{f:life}}
-\end{figure}
-
-This module simulates the evolution of Conway's Game of Life,
-a two-dimensional cellular automaton (Figure~\ref{f:life}).
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the Life world.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{numgen}}:]
-	the number of generations to simulate.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the world after simulation.
-\end{description}
-At each time step,
-this module must count the number of live ({\tt{true}}) neighbors
-of each cell,
-using both orthogonal and diagonal connectivity
-and circular boundary conditions.
-The update rule is simple:
-if a cell has 3 live neighbors,
-or has 2 live neighbors and is already alive,
-it is alive in the next generation.
-In any other situation,
-the cell becomes, or stays, dead.
-
-\subsection{{\tt{winnow}}:
-	Weighted Point Selection
-	\label{s:toys-winnow}}
-
-This module converts a matrix of integer values
-to a vector of points,
-represented as $x$ and $y$ coordinates.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix, whose values are used as masses.
-\item[{\tt{mask}}:]
-	a Boolean matrix showing which points are eligible for consideration.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{nelts}}:]
-	the number of points to select.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points.
-\end{description}
-Each location where {\tt{mask}} is {\tt{true}}
-becomes a candidate point,
-with a weight equal to the integer value in {\tt{matrix}} at that location
-and $x$ and $y$ coordinates
-equal to its row and column indices.
-These candidate points are then sorted into increasing order by weight,
-and {\tt{nelts}} evenly-spaced points selected to create the result vector.
-
-\subsection{{\tt{norm}}:
-	Point Location Normalization
-	\label{s:toys-norm}}
-
-This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-Its inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of point locations.
-\item[{\tt{nelts}}:]
-	the number of elements in the point vector.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of normalized point locations.
-\end{description}
-If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
-then the normalization equation is:
-\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
-$y$ coordinates are normalized in the same fashion.
-
-\subsection{{\tt{elastic}}:
-	Elastic Net Simulation
-	\label{s:toys-elastic}}
-
-This module uses the elastic net algorithm \cite{b:elastic-net} to find an approximate solution to the travelling salesentity problem (TSP).
-A closed, circular loop, modelled by a large number of doubly-connected points, is placed in the plane.
-This path is slowly deformed until it connects the points representing the cities to be visited (Figure~\ref{f:elastic}).
-The rate at which the loop is deformed determines the quality of the final path.
-
-\begin{figure}
-\begin{center}
-\framebox{\epsffile{fig/elastic1.ps}}
-\framebox{\epsffile{fig/elastic2.ps}}
-\framebox{\epsffile{fig/elastic3.ps}}
-\end{center}
-\caption{Relaxation of an Elastic Net\label{f:elastic}}
-\end{figure}
-
-Let the (fixed) location of city $i$ be $c_i$, and the (changeable) location of point $j$ on the elastic net be $p_j$.
-The change in the location of point $j$ during each iteration of the algorithm is given by:
-\begin{equation}
-{\delta}{p_j} = {\alpha}\sum_{i}{w_{ij}(c_i - p_j)} ~+~ K{\beta}(p_{j+1} - 2p_j + p_{j-1})
-\end{equation}
-$\alpha$ and $\beta$ are time-invariant factors relating the strengths of the city-to-point and point-to-point forces,
-and $K$ is a time-dependent factor which controls the rate of annealing.
-The strength of the force $w_{ij}$ between city $c_i$ and point $p_j$ is given by:
-\begin{equation}
-w_{ij} = {\phi}({\mid}c_i - p_j{\mid}, K) / \sum_{\ell}\phi({\mid}c_i - p_{\ell}{\mid}, K)
-\end{equation}
-where $\phi(d, K)$ is the Gaussian function $e^{-{d^2}/{K^2}}$.
-Note the normalization of each $w_{ij}$; in order for this algorithm to work properly, the total force being exerted by each city must be the same.
-This is accomplished by summing the forces exerted by each city, and then normalizing the city's forces by that total.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{cities}}:]
-	a vector of points representing the locations of the cities to be toured.
-	The $(x,y)$ coordinates of these cities must be scaled to lie in the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-\item[{\tt{n\_cities}}:]
-	the number of cities in {\tt{cities}}.
-\item[{\tt{n\_net}}:]
-	the number of elastic net points to use.
-\item[{\tt{n\_iters}}:]
-	the total number of iterations to perform.
-\item[{\tt{n\_relax}}:]
-	the number of iterations to perform with each discrete value of $K$.
-	After each {\tt{n\_relax}} iterations, $K$ is reduced by a (constant) pre-defined scaling factor.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{net}}:]
-	a vector of {\tt{n\_net}} points representing the tour found.
-\end{description}
-
-\subsection{{\tt{outer}}:
-	Outer Product
-	\label{s:toys-outer}}
-
-This module turns a vector containing point positions
-into a dense, symmetric, diagonally dominant matrix
-by calculating the distances between each pair of points.
-It also constructs a real vector whose values are
-the distance of each point from the origin.
-Inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
-\item[{\tt{nelts}}:]
-	the number of points in the vector, and the size of the matrix along each axis.
-\end{description}
-Its outputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a real matrix, whose values are filled with inter-point distances.
-\item[{\tt{vector}}:]
-	a real vector, whose values are filled with origin-to-point distances.
-\end{description}
-Each matrix element $M_{i,j}$ such that $i \neq j$
-is given the value $d_{i, j}$,
-the Euclidean distance between point $i$ and point $j$.
-The diagonal values $M_{i, i}$ are then set to
-{\tt{nelts}} times the maximum off-diagonal value
-to ensure that the matrix is diagonally dominant.
-The value of the vector element $v_i$ is set to
-the distance of point $i$ from the origin,
-which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.
-
-\subsection{{\tt{gauss}}:
-	Gaussian Elimination
-	\label{s:toys-gauss}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using explicit reduction.
-(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
-there is a well-formed solution to the equation.)
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{sor}}:
-	Successive Over-Relaxation
-	\label{s:toys-sor}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using successive over-relaxation.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\item[{\tt{tolerance}}:]
-	the solution tolerance, e.g.\ $10^{-6}$.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{product}}:
-	Matrix-Vector Product
-	\label{s:toys-product}}
-
-Given a matrix $A$,
-a vector $V$,
-and an assumed solution $X$ to the equation $AX=V$,
-this module calculates the actual product $AX={V^{\prime}}$,
-and then finds the magnitude of the error.
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{actual}}:]
-	the real vector $V$.
-\item[{\tt{candidate}}:]
-	a real vector containing the supposed solution.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-The output of this function is:
-\begin{description}
-\item[{\tt{e}}:]
-	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
-\end{description}
-
-\section{Other Issues\label{s:issues}}
-
-As described in Section~\ref{s:method-softeng},
-implementors are required to chain modules together in the manner shown in Figure~\ref{f:chaining}.
-This section describes the flow of control and data between these chained problems,
-and discusses several other important aspects of this suite,
-including I/O and visualization.
-
-\subsection{Chained Problem Sequences\label{s:issues-chain}}
-
-The chained implementation of these problems
-executes individual toys in the following order:
-\begin{enumerate}
-\item	An integer matrix $I$ with $r$ rows and $c$ columns is created
-	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
-	\label{e:chain-choice-1}
-\item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
-\item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
-	to generate a Boolean mask $B$ from $I$ in which the (minimum) percentage of {\tt{true}} locations is $P$.
-	Like $I$, $B$ has $r$ rows and $c$ columns.
-	\label{e:chain-choice-2}
-\item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
-	This step overwrites the Boolean matrix $B$.
-\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
-\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
-\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
-\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
-	as described in Section~\ref{s:toys-outer}.
-\item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
-	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
-	This measures the agreement between the solutions found by the two methods.
-\end{enumerate}
-Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
-there are four possible chained sequences.
-A minimal implementation will provide one of these,
-while a full implementation will be a single program which can execute any of the four sequences.
-Intermediate implementations providing all four sequences as separate programs are also acceptable.
-
-\subsection{Input and Output\label{s:issues-io}}
-
-I/O is an important part of programming,
-but is often treated as being of secondary importance by language designers.
-This suite requires all stand-alone toys to read input values from files, and write results back;
-the chained version must be able to checkpoint intermediate results between toys.
-Finally,
-implementors are strongly encouraged to include some means of visualizing the output or evolution of individual toys.
-
-The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
-A file containing a vector begins with 1 positive integer,
-which specifies the number of elements in the vector;
-a file containing a matrix begins with 2 positive integers,
-which specify the number of rows and columns in the matrix respectively.
-Elements of the vector or matrix then appear one per line,
-from least-indexed to greatest-indexed
-(and in row-major order for matrices).
-
-The formats described above are required to be human-readable.
-Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
-using whatever file formats are convenient.
-This will allow programmers to demonstrate the ``natural'' I/O capabilities of particular systems
-which most probably be used for checkpointing intermediate results in real programs.
-
-\begin{figure}
-\epsfxsize=0.75\textwidth
-\begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
-\caption{Cowichan File Formats\label{f:io-formats}}
-\end{figure}
-
-\subsection{Visualization\label{s:issues-vis}}
-
-Visualization is an increasingly important part of scientific and numerical computing.
-Including it as an optional part of this problem set is also a good way to exercise both
-the use of libraries within a parallel programming system,
-and the ease with which a programming system can support variant versions of a program.
-
-Our requirements for the graphics interface are:
-\begin{enumerate}
-\item	Visualization for individual toys should be as simple as possible,
-	and should require as few changes to the basic toy code as possible.
-\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
-	The latter should create all images in a single window.
-\item	A single source file should implement both the graphical and non-graphical versions of each toy.
-\end{enumerate}
-The first two points require the system to distinguish between the actual drawing code
-(e.g.\ drawing a vector of points as circles)
-and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
-While the stand-alone toys use a whole window for drawing,
-the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
-At present,
-this is implemented by using a call counter
-in the interfaces for the individual toys and in the chained module.
-The actual drawing routines called from within each toy could then be the same.
-
-The present visualization routines were originally written by Anil Sukul,
-a student at the Vrije Universiteit, Amsterdam.
-The package uses the public domain graphics library VOGLE,
-available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
-VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
-as well as screen control routines (such as double buffering).
-VOGLE was chosen because of its simplicity,
-and because it includes Pascal and Fortran interfaces.
-
-\subsection{Reproducibility\label{s:issues-reproduce}}
-
-Reproducibility is an important issue for parallel programming systems.
-While constraining the order of operations in a parallel system to guarantee reproducibility makes programs in that system easier to debug,
-it can also reduce the expressiveness or performance of the system.
-
-In this problem set, irreproducibility can appear in two forms: numerical and algorithmic.
-The first arises in toys such as {\tt{gauss}}, {\tt{sor}}, and {\tt{elastic}}, which use floating-point numbers.
-Precisely how round-off errors occur in these calculations can depend on the distribution of work among processors,
-or the order in which those processors carry out particular operations.
-
-Irreproducibility also arises in toys which only use exact numbers, such as {\tt{invperc}} and {\tt{randmat}}.
-In the former, the percolation region is grown by repeatedly filling the lowest-valued cell on its perimeter.
-If several cells have this value, implementations may choose one arbitrarily.
-Thus, different implementations may produce very different fractal shapes.
-In the case of random number generation, the simplest thing to do is to run the same generator independently on each processor,
-although the values in the resulting matrix then depend on the number of processors used.
-
 \section{Conclusion\label{s:conclude}}
@@
 If you would like to contribute to this effort, please contact either of the first two authors.
-
-\bibliographystyle{plain}
-\newcommand{\bibea}{{\em{et~al.}}}
-\bibliography{paper}
-
-\appendix
-
-\newpage
-\section{Parallel Clich\'{e}s\label{s:cliche}}
-
-\subsection{Operations\label{s:cliche-ops}}
-
-This list details some operations which are supported by many parallel programming systems.
-The toys described in Section~\ref{s:toys} provide opportunities for exercising many of these,
-and implementors are encouraged to phrase discussion of their work in terms of these operations where appropriate.
-
-\begin{itemize}
-\item	elementwise operations on arrays (unary, binary, and scalar promotion)
-\item	cumulative operations on arrays (reduction and parallel prefix)
-\item	array re-shaping and re-sizing (e.g.\ sub-array extraction)
-\item	partial (conditionally masked) versions of all of the above
-\item	regular data motion (e.g.\ circular and planar shifting)
-\item	irregular data motion (e.g.\ 1-to-1, 1-to-many, and many-to-1 permutation)
-\item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
-\item	differential local addressing (i.e.\ subscripting an array with an array)
-\item	full or partial replication of shared read-only values
-\item	full or partial replication of shared read-mostly values with automatic consistency management
-\item	structures with rotating ownership, suitable for migration
-\item	producer-consumer structures
-\item	partitionable structures
-\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
-\item	dynamic re-partitioning (e.g.\ for load balancing)
-\item	committed mutual exclusion (e.g.\ waiting on a semaphore)
-\item	uncommitted mutual exclusion (e.g.\ lock or fail)
-\item	barrier synchronization
-\item	multiple concurrent barriers used by non-overlapping groups of processes
-\item	fetch-and-add, and other fetch-and-operate functions
-\item	pre-scheduled receipt of messages of known composition
-\item	variable-length message receipt
-\item	message selection by type
-\item	message selection by source
-\item	message selection by contents
-\item	guarded message selection (e.g.\ CSP's {\tt{alt}} construct)
-\item	broadcast and partial broadcast
-\item	split-phase (non-blocking) operations
-\item	data marshalling and unmarshalling of ``flat'' structures (e.g.\ arrays of scalars)
-\item	data marshalling and unmarshalling of nested or linked structures
-\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
-\item	pipelining
-\item	distributed linked structures in systems without physically-shared memory
-\item	indexing of distributed shared arrays
-\item	collective I/O
-\item	uniform-size interleaved I/O
-\item	heterogeneous-size interleaved I/O
-\item	independent I/O operations on a single shared file
-\end{itemize}
-
-\subsection{Memory Reference Patterns\label{s:cliche-memref}}
-
-These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
-Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.
-
-\begin{itemize}
-\item	Write-once:
-	variables which are assigned a value once, and only read thereafter.
-	These can be supported through replication.
-\item	Write-many:
-	variables which are written and read many times.
-	If several processes write to the variable concurrently, they typically write to different portions of it.
-\item	Producer-consumer:
-	variables which are written to by one or more objects, and read by one or more other objects.
-	Entries in a shared queue, or a mailbox used for communication between two processes, are examples.
-\item	Private:
-	variables which are potentially shared, but actually private.
-	The interior points of a mesh in a program which uses geometric decomposition fall into this category,
-	while boundary points belong to the previous class.
-\item	Migratory:
-	variables which are read and written many times in succession by a single process before ownership passes to another process.
-	The structures representing particles in an $N$-body simulation are the clearest example of this category.
-\item	Result:
-	Accumulators which are updated many times by different processes, but thereafter read.
-\item	Read-mostly:
-	Any variable which is read much more often than it is written to.
-	The best score so far in a search problem is a typical instance of this class:
-	while any process might update it, in practice processes read its value much more often than they change it.
-\item	Synchronization:
-	Variables used to force explicit synchronization in a program, such as semaphores and locks.
-	These are typically ignored for long periods, and then the subject of intense bursts of access activity.
-\item	General read-write:
-	Any variable which cannot be put in one of the above categories.
-\end{itemize}
-
-\newpage
-\section{Comparing Parallel Programming Systems\label{s:questions}}
-
-This list of questions is a modified version of the one presented in \cite{b:compare-ada-c-pascal}.
-It is presented here as an organizational aid;
-when describing the PPS they have used,
-implementors are encouraged to use the order and terminology of this list where appropriate.
-We acknowledge that many of these questions will not apply to particular systems,
-and encourage implementors to ask questions which we have left out.
-
-\begin{enumerate}
-\item History and Philosophy
-  \begin{enumerate}
-  \item For whom and for what purpose was the language designed?
-  \end{enumerate}
-\item Syntax
-  \begin{enumerate}
-  \item Model implemented (Figure~\ref{f:taxonomy})
-  \item Class
-    \begin{enumerate}
-    \item library called from existing sequential language?
-    \item syntactic extensions to existing sequential language?
-    \item new language in Backus (Fortran/Algol/C) tradition?
-    \item new (non-Backus) language?
-    \end{enumerate}
-  \item Concurrency mechanisms adhere to same philosophy as original
-	language (where appropriate)?
-  \item Readability?  Syntactic consistency?  Conciseness?
-	(on an APL-to-Modula scale)
-  \item Common errors (e.g.\ confusing {\tt{=}} and {\tt{==}} in C)?
-  \end{enumerate}
-\item Type Philosophy and Data Types
-  \begin{enumerate}
-  \item What primitive types does the language support?
-  \item How are new types created?
-  \item What intrinsic operations are defined on user-defined types?
-  \item How strict is typing?
-	How is type conversion done?
-  \item Can type information be inspected at run-time?
-  \item What concurrent operations on aggregates
-	(such as lists and arrays) are intrinsic?
-	What kinds of concurrent aggregate operations can users define?
-  \item How are aggregates defined (e.g.\ by index set, by shape)?
-	Can new kinds of aggregates (e.g.\ templates for arrays)
-	be defined?
-  \item How strict is typing during communication?
-  \item What support is there for communication or manipulation
-	of data of non-scalar or non-intrinsic types?
-  \item Can structural information about aggregates
-	(e.g.\ number of elements in a set, dimension of an array)
-	be inspected at run-time?
-  \item What built-in support is there for object classes and inheritance?
-	For polymorphism?
-	For class (as opposed to instance) data?
-  \item Are functions first-class?
-  \item Are continuations supported?  Are they first-class?
-  \item Is dynamically-allocated memory supported?
-	Is it done automatically?
-	Is garbage-collection done automatically?
-	Can a program ever create a dangling pointer?
-	An alias?
-  \end{enumerate}
-\item Operators
-  \begin{enumerate}
-  \item What intrinsic operators are provided (brief list)?
-	How strongly do these type-check their arguments?
-  \item How are new operators defined?
-  \item Are the precedence rules straightforward?
-  \item Are side-effects/mutation allowed inside non-call expressions
-	(e.g.\ C's side-effecting \verb`i++` construct)?
-  \item Are their results platform-independent?
-  \item Are the results of intrinsic parallel operations reproducible?
-	Are they platform-independent?
-  \item Are heterogeneous operations on aggregates allowed
-	(e.g.\ every element of a list of polygons calculates its
-	center of mass)?
-	If so, how are these described, and how are they implemented?
-  \item Which of the operators in Section~\ref{s:cliche-ops} are supported?
-  \end{enumerate}
-\item Control Flow
-  \begin{enumerate}
-  \item What synchronization is automatically imposed during execution
-	(i.e.\ at what points can definite assertions about the state
-	of the running program be made)?
-  \item What control constructs does the language support?  Encourage?
-  \item How is concurrency expressed?  How structured is this mechanism?
-  \item Can the degree of concurrency be throttled (limited)?
-	Is this done by the system, or by the user?
-  \item How are exceptions handled/reported?
-	Is it different for intrinsic operations and user-defined operations?
-  \end{enumerate}
-\item Subroutines, Scope, and Modularization
-  \begin{enumerate}
-  \item How efficient is subroutine call compared to in-line execution
-	(i.e.\ how expensive are parameter passing and non-local control flow)?
-  \item How much information about parameters is available (required)
-	inside called subroutine
-	(e.g.\ whether object is shared with other processes or private
-	to executing process)?
-  \item Are side-effecting procedures allowed?
-  \item Can generic routines be written
-	(e.g.\ to work on different types or array shapes)?
-  \item Is there a restriction on return values (e.g.\ can a function return
-	any object which could be declared)?
-  \item Are there levels of scope?
-	If multiple levels are permitted, how are references to stale scopes
-	handled or prevented?
-	Can information for declarations (such as size) be inherited (dynamically)
-	from higher scopes?
-  \item What synchronization requirements (if any) are imposed during calls
-	(i.e.\ do all processes (in a group?) have to enter/exit the same subroutine
-	at the same time)?
-  \item What support is there for modularization?
-	For hiding implementation details?
-	For sharing such information?
-	For detecting and resolving name clashes?
-  \end{enumerate}
-\item Concurrent Programming Facilities
-  \begin{enumerate}
-  \item At what level(s) is the system concurrent?
-    \begin{enumerate}
-    \item implicit (the user shouldn't ask such things)
-    \item operational (e.g.\ data-parallel addition)
-    \item program counter (single executable, multiple states)
-    \item procedural (heterogeneous control parallelism)
-      \begin{enumerate}
-      \item one execution thread per address space
-      \item multiple threads per address space
-      \end{enumerate}
-    \end{enumerate}
-  \item What mutual exclusion and synchronization primitives are provided?
-  \item At what level(s) is data sharing visible to the programmer?
-    \begin{enumerate}
-    \item sequential consistency
-    \item sequential consistency within processor clusters
-    \item object-level consistency
-    \item there is no sharing
-    \end{enumerate}
-  \item What is the conceptual granularity of operations?
-    \begin{enumerate}
-    \item synchronization after atomic instruction
-    \item homogeneous operations with large-grain synchronization
-    \item heterogeneous operations with large-grain synchronization
-    \item completely asynchronous execution
-    \end{enumerate}
-  \item How do processes communicate data values?
-	How do they synchronize?
-	Are non-blocking operations (ones in which buffers might be
-	in a visibly volatile state) allowed?
-	How are they checked for termination?
-  \item If data are communicated explicitly,
-	how does the receiver choose what to accept?
-    \begin{enumerate}
-    \item by address (e.g.\ channel or port)
-    \item by pattern matching on contents
-    \item by explicit message tagging
-    \item by provision of a typed buffer
-    \item by provision of an untyped (byte-block) buffer
-    \item in FIFO order
-    \end{enumerate}
-  \item How do users specify the number of processors on which to run?
-	Must this be built into the program,
-	or can it be decided at run-time?
-	Is there support for executing on heterogeneous processors?
-  \item Can the user specify the mapping of processes to processors?
-	If so, how?
-	How tightly coupled are the mapping of processes to processors,
-	and the mapping of distributed data structures (if supported)?
-  \item Can the user specify the mapping of data structures to processors?
-        Can data structures be decomposed or partitioned?
-  \item What support is there for operations on groups of processes, such as:
-    \begin{enumerate}
-    \item broadcast
-    \item multicast (partial broadcast)
-    \item barrier synchronization
-    \end{enumerate}
-  \item Can arbitrary process groups be created,
-	or must they be structured in some way
-	(e.g.\ as logical mesh)?
-	Can operations involving disjoint groups be executed concurrently
-	(e.g.\ a broadcast within one group,
-	and a barrier within another)?
-  \end{enumerate}
-\item Input/Output
-  \begin{enumerate}
-  \item Is simple sequential (single-process) I/O simple to write?
-  \item Is concurrent I/O supported?  If so, which flavours:
-    \begin{enumerate}
-    \item single-reader-broadcast/collect-single-writer
-    \item homogeneously-sized or -strided segments
-    \item heterogeneously-sized or -strided segments
-    \item arbitrary independent operations
-    \end{enumerate}
-  \item What types can be read/written intrinsically?
-	E.g.\ can an entire record (or array, or list)
-	be read or written in a single operation?
-  \end{enumerate}
-\item Access to Routines in Other Languages and to the Hardware
-  \begin{enumerate}
-  \item Can subroutines written in other languages be called?
-        Can subroutines in this language be called from other languages?
-  \item Are there ways to access the run-time system directly,
-	or similar low-level facilities?
-  \item Is there high-level access to physical addresses and devices?
-  \end{enumerate}
-\item Practical Considerations
-  \begin{enumerate}
-  \item Are both batch and interactive processing supported?
-  \item How reliable and efficient are existing implementations?
-	How fast is it (i.e.\ how quickly does code compile and load)?
-	Is it supported (who fixes it if it breaks)?
-  \item Is there a standard for the system (from some standards organization)?
-  \item Is the documentation adequate? For whom (CS graduates, computer-literate
-	users, the general population)? Is training available?	Consulting?
-  \item How much code has been written (lines of code or weeks of effort)?
-        By whom (computer science undergraduates, professional engineers, etc.)?
-  \item How long does it take the compiler to generate a ``small'' executable
-        (such as {\tt{"hello, world"}})?
-  \item How large is a ``small'' executable?
-        How long does it take to load such an executable?
-  \end{enumerate}
-\item Supporting Tools
-  \begin{enumerate}
-  \item What is the development environment like?
-        Can code be developed on a workstation and then moved to a parallel platform?
-  \item Are there debugging aids, performance analysers, and management tools?
-  \item How much help do these really provide?
-  \end{enumerate}
-\end{enumerate}
-
-\newpage
-\section{A Critique of the C Implementation\label{s:critique}}
-
-This section is a critique of the ANSI~C reference implementation of the problem suite.
-It is intended to serve as a model for (self-)criticism of other programming systems.
-
-\begin{itemize}
-\item	C's support for multi-dimensional arrays (MDAs) is very weak.
-	There is no way to dynamically allocate an MDA in a single go---one must either
-	allocate a block of the same size as the desired array,
-	and then do indexing calculations by hand,
-	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
-	MDAs do not carry dimension information with them,
-	so it is impossible to determine the size of an array parameter within a function.
-	Finally,
-	C does not treat all axes of an array equally:
-	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
-	it is impossible to slice it along the other axis.
-\item	C does not distinguish between Boolean and integer types.
-	As a result,
-	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
-	can be passed in reverse order without a type error.
-	Using {\tt{typedef}} to create a Boolean type name does not solve this
-	(at least, not in {\tt{gcc}} V2.5.8).
-\item	Union types cannot be safely initialized.
-	This complicated the implementation of the graphics interface,
-	where it would have been much more elegant to define a union type,
-	each of whose variants held parameter specifications for a single toy.
-	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
-	initializing some and filling other with don't-care values.
-	This is neither safe nor elegant.
-\item	Parameter values cannot be used in the initialization of local variables inside functions.
-	In particular,
-	it is not possible to create a local vector with a length specified by an input parameter.
-	Such a facility would be useful in the {\tt{winnow}} toy,
-	where we have instead allocated local temporaries of the maximum possible size.
-\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
-	For example, the ``fail()'' error-handling routine was buggy because
-	all \verb`real` arguments were being automatically converted to doubles,
-	but were being taken off the stack as floats.
-\item	no intrinsic notion of group ID/self ID in threads
-\item	packaging parameters for threads
-\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
-\end{itemize}
-
-\end{document}

### 2009-08-28 paper/2009-09/intro.tex
Breaking paper into manageable chunks

--- 2009-08-20 paper/2009-09/paper.tex
@@
-\documentstyle[epsf,11pt]{article}
-
-\begin{document}
-
-\begin{Large}\begin{center}
-The Cowichan Problems: \\
-Assessing the Usability of Parallel Programming Systems
-\end{center}\end{Large}
-
-\vspace{\baselineskip}
-
-\begin{center}
-\begin{tabular}{cccc}
-Gregory V.~Wilson\footnote{\verb`http://www.third-bit.com`}
-&
-Andrew Borzenko
-&
-Cameron Gorrie
-&
-Zachary Kincaid
-\end{tabular}
-\end{center}
-
-\vspace{\baselineskip}
-
-\begin{abstract}
-Most programmers believe that parallel programming is harder than sequential programming.
-Many also believe that today's parallel programming systems make it even harder.
-We have developed a suite of chained toy problems that can be used to assess
-the usability of different scientific programming languages, parallel or otherwise.
-This paper motivates and describes the suite,
-and discusses what we have learned from porting it to four parallel C++ systems
-(MPI, OpenMP, Intel Thread Building Blocks, and a Linux tuple space system),
-and from translating it into sequential and parallel Haskell.
-\end{abstract}
-
 \section{Introduction\label{s:intro}}
@@
 Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.
-
-\section{Quantifying the Importance of Usability\label{s:corollary}}
-
-According to Amdahl's Law,
-the time $t$ required to execute a program has two components:
-$t_{seq}$, representing sequential operations that cannot be parallelized,
-and $t_{par}$, representing those operations which can.
-If we let $\sigma$ be the ``serial fraction'' of the program,
-so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
-then the speedup which can be achieved using $P$ processors is:
-\begin{eqnarray*}
-s(P)	& =	& \frac{t(1)}{t(P)}	\\
-	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
-\end{eqnarray*}
-As $P{\rightarrow}{\infty}$,
-the potential speedup is bounded by $1/{\sigma}$.
-
-What is not usually acknowledged is
-the influence of programming system usability on
-the potential speedup of sequential programs.
-As stated above, Amdahl's Law applies only to a single run of a program.
-In the real world, programs are first developed, and then run many times.
-If we let $T$ be the total running time of a program over its lifetime,
-$D_{seq}$ be the time required to develop the sequential components of a program,
-and $D_{par}$ be the time required to develop the parallel components
-(or parallelize the sequential components),
-then the lifetime speedup we can achieve is given by:
-\begin{eqnarray*}
-S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
-	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
-\end{eqnarray*}
-Assume that we are parallelizing a ``dusty deck'',
-i.e.\ that the time $D_{seq}$ has already been invested.
-If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
-then achievable speedup is limited by:
-\begin{eqnarray*}
-S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
-		& =	& \frac{1}{{\phi} + {\sigma}}
-\end{eqnarray*}
-The time required to parallelize a program can therefore be seen
-as increasing that program's effective serial fraction.
-Unless parallelization time can be substantially reduced,
-massively-parallel computing is likely to remain attractive only in those cases in which:
-\begin{itemize}
-\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
-\item	the expected total program runtime is very large
-	(i.e.\ the program is a package used extensively in a field
-	such as computational chemistry or vehicle crash simulation);
-	or
-\item	cost is not an obstacle (i.e.\ the customer is the NSA).
-\end{itemize}
-
-\section{Previous Work\label{s:previous}}
-
-In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
-\begin{quote}
-As practiced by computer science,
-the study of programming is an unholy mixture of mathematics,
-literary criticism,
-and folklore.
-\end{quote}
-Little has changed since.
-Computer scientists pay lip service to the importance of usability,
-but are strangely reluctant to try to quantify it,
-particularly with respect to the tools they use themselves.
-Religious debates about the ``right'' first-year teaching language
-or the ``best'' GUI toolkit are common;
-systematic use of the experimental method is not.
-
-Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
-such as readability and code-pattern recognition
-\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
-The little work that has been done on ``programming in the large'' has been anecdotal,
-such as \cite{b:compare-ada-c-pascal}.
-This book presented various authors' impressions of Pascal, C, and Ada
-(before any complete Ada implementations existed).
-Its editors formulated a list of questions to use in comparing languages,
-but did not use it themselves.
-A modified version of this questionnaire,
-tailored for parallel programming languages,
-is given in Appendix~\ref{s:questions},
-and will be used as a guideline in our work.
-
-Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
-The first presented implementations of a simple numerical quadrature program
-in more than a dozen different parallel languages
-in use on mid-1980s hardware.
-The second presented implementations of
-the Salishan Problems---Hamming number generation,
-isomer enumeration,
-skyline matrix reduction,
-and a simple discrete event simulator---in
-C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
-and a variety of dataflow and functional languages.
-Both of these books convey the flavor of the languages they describe,
-but neither made any effort to compare languages or problem implementations directly.
-
-Inspired by the Salishan Problems,
-the first author developed a suite of seven problems \cite{b:cowichan-1}
-in order to test the limits of parallel programming systems.
-Individual problems were badly load-balanced,
-manipulated irregular pointer-based structures,
-and so on.
-Our experience implementing them was extremely valuable,
-as it uncovered many bugs in the programming system used,
-and revealed several shortcomings of the system's design.
-However,
-we also discovered that implementing each problem takes six to eight weeks,
-which we feel is prohibitive at this exploratory stage.
-
-\section{Methodology\label{s:method}}
-
-The main aim of this work is to assess and compare
-the ease with which programs can be developed and/or ported
-using different parallel programming systems.
-Here,
-we discuss issues related to choosing the problems that make up the suite,
-assessing the complexity of particular implementations,
-and other purposes which implementations of these might serve.
-Recognizing that these problems are extremely simple,
-and easy to parallelize,
-we refer to them as ``toys''.
-
-\subsection{Criteria for Selection\label{s:method-criteria}}
-
-Our criteria for including problems in this suite are as follows:
-\begin{enumerate}
-\item	Each toy should require no more than an afternoon
-	to write and test in a well-supported sequential language.
-	We feel that making individual problems more complicated will only discourage implementation.
-\item	The correctness of each implementation must be relatively easy to verify.
-	Toys whose output is easily visualized are therefore to be preferred,
-	as are toys whose results are insensitive to floating-point discretization effects.
-\item	Each toy application should be amenable to
-	an analysis of speedup or isoefficiency \cite{b:isoefficiency-tutorial}
-	to allow a comparison of theory and practice.
-\item	At least some toys should not be ``infinitely scalable''.
-	Many real-world applications are not,
-	and this suite should reflect such limitations.
-\item	At least some toys should require I/O,
-	since this important aspect of real-world programming is often neglected by PPS designers.
-\item	There should be some potential redundancy in the type of parallelization done,
-	so that re-use of software can be demonstrated.
-\item	Together,
-	the toys in the suite must exercise a wide range of common parallel operations
-	(Appendix~\ref{s:cliche-ops})
-	and memory reference patterns
-	(Appendix~\ref{s:cliche-memref}).
-\item	The ``obvious'' implementations of these toys should span
-	a wide range of parallel programming paradigms,
-	such as task farming,
-	geometric and functional decomposition,
-	and speculative parallelism.
-\item	Parallelizations for several (if not all) of
-	the types of programming systems shown in Figure~\ref{f:taxonomy} should be known.
-	This suite is not intended to be a collection of research problems;
-	reasonably simple and efficient implementations in most environments
-	should be straightforward to produce.
-\end{enumerate}
-
-\begin{figure}
-\epsfysize=0.5\textheight
-\begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
-\caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
-\end{figure}
-
-\subsection{Specification\label{s:method-spec}}
-
-Toys should be specified in terms of inputs and outputs,
-rather than algorithmically.
-For example,
-``sort $N$ integers in the range $1{\ldots}R$,
-where $R$ is not known in advance and there may be multiple instances of a given value''
-is to be preferred over
-``parallelize quicksort''.
-As \cite{b:slalom} argued,
-which parallel algorithm is best in a particular programming system
-depends critically on the architectural assumptions made by that system.
-Implementors should be free to employ those methods which they find natural and convenient.
-The only exception we make to this rule is in the two matrix equation solvers,
-where we require Gaussian elimination and successive over-relaxation.
-Given the enormous number of matrix solution algorithms in use,
-we felt we had to specify these particular algorithms in order to ensure that
-different implementations of this suite would be comparable.
-
-\subsection{Software Engineering Issues\label{s:method-softeng}}
-
-These toys will exercise many different aspects of parallel systems.
-However,
-their ``single algorithm per program'' model is not representative of real applications.
-These usually contain several discrete (and sometimes overlapping) phases,
-each of which is qualitatively different.
-A full implementation of this suite will therefore have two parts.
-In the first,
-each toy will be implemented as a stand-alone program.
-In the second,
-toys will be chained together to create the kind of phase behaviour seen in real applications
-(Figure~\ref{f:chaining}).
-This will test the ease with which heterogeneous parallelism can be mixed within a single program.
-It will also show how well the system supports code re-use and information hiding,
-which are crucial to the development of large programs.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
-\caption{Chaining\label{f:chaining}}
-\end{figure}
-
-A second reason for chaining toys is that
-many groups are now using algorithmic skeletons to structure parallel programs
-\cite{b:cole-skel,b:p3l-overview,b:enterprise}.
-Such skeletons encapsulate the details of a particular style of parallelism,
-and hide low-level or platform-dependent issues.
-We hope that chaining will show the strengths and weaknesses of such systems.
-
-Finally,
-chaining should be designed so that some toys can be executed concurrently.
-Many PPSs impose extraneous constraints on programs,
-e.g.\ require all processes to participate in every barrier synchronization,
-or require the same executable to be loaded onto each processor.
-These constraints can limit the exploitation of potential concurrency.
-Permitting, but not requiring, concurrent execution of several toys should uncover such limitations.
-
-\subsection{Sizing\label{s:method-size}}
-
-One crucial aspect of the specification of toys is
-the way in which the sizes of problems are determined.
-In a completely frozen model, the actual size of each problem would be compiled into each toy.
-A fluid implementation, by contrast, would allow sizes to be specified at run-time,
-and would only then allocate and partition data structures.
-We have chosen to use an intermediate model (which might be called ``slushy''),
-in which the maximum size of individual problems is specified during compilation,
-but the actual size of a problem is only determined when the toy begins to execute.
-We hope that the use of this model will force implementors to deal with
-at least some of the software engineering issues
-involved in building flexible large-scale applications,
-without making implementation prohibitively difficult.
-
-\subsection{Assessing Complexity\label{s:method-complex}}
-
-There are at least two different ways in which usability might be compared.
-One would be to measure the performance achieved by an ``average'' programmer
-as a function of time on each of a range of problems.
-This would lead to graphs of the kind shown in Figure~\ref{f:perf},
-where the left graph shows a particular result
-and the right graph a simplified abstraction of it.
-By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
-which is the length of vector on which a pipelined architecture achieves
-half its theoretical peak performance,
-we could in principle find the value of $p_{1/2}$---the programming time required to achieve
-half of a machine's peak performance\footnote{Note that
-	if performance was measured as a fraction of the figures quoted by manufacturers for their machines,
-	it is unlikely that the halfway mark would ever be reached.}
-for a particular combination of programming system and problem type.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
-\caption{Performance vs.\ Time\label{f:perf}}
-\end{figure}
-
-A second option would be to measure the complexities of
-implementations of a single application using different programming systems.
-With this approach,
-comparative performance or speedup figures would be supplemented by complexity measures,
-as shown in Figure~\ref{f:complex}.
-
-\begin{figure}
-\epsfxsize=6cm
-\begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
-\caption{Comparative Code Complexity\label{f:complex}}
-\end{figure}
-
-The drawback of the first approach is that there are no ``typical'' problems,
-or ``average'' programmers.
-Meaningful measurement of  $p_{1/2}$ would therefore require
-measurement of a large number of programmers working on many problems.
-Since this is too expensive to be practical,
-the second approach is the one we have chosen to adopt.
-It has drawbacks as well,
-most notably the fact that no-one has ever demonstrated any better measure of program complexity
-than a simple count of the number of lines in a program \cite{b:fenton-metrics}.
-Since we anticipate a wide variation in the types of progamming systems that will be used in this work
-(ranging from declarative dataflow languages
-through functional languages with Lisp-like syntax
-to parallel dialects of Fortran and C),
-and in the indentation, commenting, and layout styles of particular programmers,
-we may measure complexity by counting the number of tokens in the source code which are
-keywords,
-intrinsic operators,
-user-defined functions,
-type or variable declarations,
-and miscellaneous punctuation.
-
-\subsection{Other Uses for Implementations\label{s:method-uses}}
-
-This suite is intended for assessing PPS usability,
-but we envisage at least three other uses for implementations of it:
-\begin{description}
-\item[Related Research:]
-	We hope that the existence of this suite will indicate
-	to the developers of new parallel programming systems
-	what a mature tool should be able to support.
-	If sufficiently diverse,
-	the suite could also be used to help test the correctness of new or improved systems.
-	Finally,
-	this suite could complement existing performance benchmarks,
-	since performance figures for individual toys could be compared.
-\item[Tool Development:]
-	Requests from tool developers for existing parallel programs,
-	or for event traces from them,
-	are common on newsgroups such as {\tt{comp.parallel}}.
-	We hope that traces from implementations of this suite
-	will be used to compare different debugging and performance monitoring tools,
-	that their source code will be used to test
-	new parallelizing compilation techniques or mapping and scheduling algorithms,
-	and so on.
-\item[Teaching:]
-	Implementations will be suitable for use as classroom examples,
-	since they will be small enough to be understood quickly.
-	The toys themselves should also be suitable as classroom exercises
-	in a senior undergraduate course on parallel computing.
-\end{description}
-
-\subsection{Some Criticisms\label{s:method-criticism}}
-
-\noindent {\em{Why not use an existing benchmark suite for this purpose?}\/}
-
-Implementations of benchmark suites have usually been coded for absolute speed.
-They are not representative of ``normal'' practice,
-in which reasonable performance is the usual goal,
-and so are inappropriate for this work.
-In addition,
-most existing benchmark suites are too large to be implemented or ported quickly
-(e.g.\ SPEC~\cite{b:bench-over}),
-over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
-or concerned with a limited range of applications
-(e.g.\ the NAS Parallel Benchmarks~\cite{b:nas-results}).
-Finally,
-most exist only in C or Fortran (or both);
-many interesting parallel programming systems are built on top of other languages,
-including Modula-2, Scheme, Prolog, and various dataflow languages.
-We do not want the effort required to translate several large programs
-to discourage researchers from participating in this exercise.
-
-\vspace{\baselineskip}
-
-\noindent {\em{Won't the results depend primarily on programmer ability?}\/}
-
-The short answer to this question is yes,
-but no more than the results from performance benchmarks.
-A longer answer is that	since we intend to measure the combination of code complexity and performance achieved,
-rather than performance as a function of development time,
-programmers will be allowed to revise and improve their implementations when and as desired.
-
-\section{The Cowichan Problems\label{s:toys}}
-
-This section describes the problems making up the full problem suite.
-As discussed above,
-the implementation of each is to be used both as a stand-alone program
-and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
-Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cl|cl}
-Code	& Toy				& Code	& Toy				\\
-\hline
-E	& elastic net			& O	& outer product			\\
-G	& Gaussian elimination		& P	& matrix-vector product		\\
-H	& halving shuffle		& R	& random matrix generation	\\
-I	& invasion percolation		& S	& successive over-relaxation	\\
-L	& Game of Life			& T	& image thresholding		\\
-M	& Mandelbrot Set		& V	& vector difference		\\
-N	& point normalization		& W	& point value winnowing
-\end{tabular}
-\caption{Single-Character Codes for Toys\label{t:toy-codes}}
-\end{center}
-\end{table}
-
-\subsection{{\tt{mandel}}:
-	Mandelbrot Set Generation
-	\label{s:toys-mandel}}
-
-\begin{figure}
-\begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
-\caption{The Mandelbrot Set\label{f:mandel}}
-\end{figure}
-
-This module generates the Mandelbrot Set
-for a specified region of the complex plane
-(Figure~\ref{f:mandel}).
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{x0, y0}}:]
-	the real coordinates of the lower-left corner
-	of the region to be generated.
-\item[{\tt{dx, dy}}:]
-	the extent of the region to be generated.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing the iteration count at each point in the region.
-\end{description}
-
-Given initial coordinates $(x_0, y_0)$,
-the Mandelbrot Set is generated by iterating the equation
-\begin{eqnarray*}
-x^{\prime}	& =	& x^2 - y^2 + y_0	\\
-y^{\prime}	& =	& 2{x}{y} + x_0
-\end{eqnarray*}
-until either an iteration limit is reached,
-or the values diverge.
-The iteration limit used in this module is 150 steps;
-divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
-The integer value of each element of the matrix is
-the number of iterations done.
-Note that,
-as in all problems,
-the output is required to be
-independent of the number of processors used.
-This requires programmers to be careful about
-floating-point roundoff effects.
-
-\subsection{{\tt{randmat}}:
-	Random Number Generation
-	\label{s:toys-rng}}
-
-This module fills a matrix with pseudo-random integers.
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{s}}:]
-	the random number generation seed.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix filled with random values.
-\end{description}
-Note that,
-as in all problems,
-the output is required to be independent of the number of processors used.
-Generating new seed values from the given seed,
-and running one copy of the random number generator
-on each processor,
-is therefore unlikely to qualify as a solution.
-
-\subsection{{\tt{half}}:
-	Two-Dimensional Shuffle
-	\label{s:toys-half}}
-
-This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
-shuffles them,
-and then repeats this operation along the other axis.
-Values in odd-numbered locations are collected at the low end of each row or column,
-while values in even-numbered locations are moved to the high end.
-An example transformation is:
-
-\begin{eqnarray*}
-\begin{array}{cccc}
-a & b & c & d \\
-e & f & g & h \\
-i & j & k & l
-\end{array}
-& = &
-\begin{array}{cccc}
-a & c & b & d \\
-i & k & j & l \\
-e & g & f & h
-\end{array}
-\end{eqnarray*}
-
-Note that how an array element is moved depends only on whether its location index is odd or even,
-not on whether its value is odd or even.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing shuffled values.
-\end{description}
-
-\subsection{{\tt{invperc}}:
-	Invasion Percolation
-	\label{s:toys-invperc}}
-
-Invasion percolation models the displacement of one fluid
-(such as oil)
-by another
-(such as water)
-in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
-In two dimensions,
-this can be simulated by generating
-an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
-and then marking the center cell of the grid as filled.
-In each iteration,
-one examines the four orthogonal neighbors of all filled cells,
-chooses the one with the lowest value
-(i.e.\ the one with the least resistance to filling),
-and fills it in.
-Figure~\ref{f:invperc1} shows the first few steps in this process,
-while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{math}
-\begin{array}{ccc}
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & 29 & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \star & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & 29 & 77 \\
-\hline
-61 & \star & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \star & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \star & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \star \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & \star & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-\star & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & \star & 72 & 45 & 38 \\
-\hline
-\circ & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \circ \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\end{array}
-\end{math}
-\end{footnotesize}
-\caption{Invasion Percolation\label{f:invperc1}}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
-\end{center}
-\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
-\end{figure}
-
-The simulation continues until
-some fixed percentage of cells have been filled,
-or until some other condition
-(such as the presence of trapped regions)
-is achieved.
-The fractal structure of the filled and unfilled regions
-is then examined to determine how much oil could be recovered.
-The na\"{\i}ve way to implement this is to repeatedly scan the array;
-a more sophisticated, and much faster, sequential technique
-is to maintain a priority queue of unfilled cells
-which are neighbors of filled cells.
-This latter technique is similar to the list-based methods used in some cellular automaton programs,
-and is very difficult to parallelize effectively.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{nfill}}:]
-	the number of points to fill.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
-\end{description}
-Filling begins at the central cell of the matrix
-(rounding down for even-sized axes).
-
-\subsection{{\tt{thresh}}:
-	Histogram Thresholding
-	\label{s:toys-thresh}}
-
-This module performs histogram thresholding on an image.
-Given an integer image $I$ and a target percentage $p$,
-it constructs a binary image $B$
-such that $B_{i,j}$ is set
-if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
-The general idea is that an image's histogram should have 2 peaks,
-one centered around the average foreground intensity,
-and one centered around the average background intensity.
-This program attempts to set
-a threshold between the two peaks in the histogram
-and select the pixels above the threshold.
-This module's inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the integer matrix to be thresholded.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{percent}}:]
-	the minimum percentage of cells to retain.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
-\end{description}
-
-\subsection{{\tt{life}}:
-	Game of Life
-	\label{s:toys-life}}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
-\end{center}
-\caption{Snapshots from the Game of Life\label{f:life}}
-\end{figure}
-
-This module simulates the evolution of Conway's Game of Life,
-a two-dimensional cellular automaton (Figure~\ref{f:life}).
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the Life world.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{numgen}}:]
-	the number of generations to simulate.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the world after simulation.
-\end{description}
-At each time step,
-this module must count the number of live ({\tt{true}}) neighbors
-of each cell,
-using both orthogonal and diagonal connectivity
-and circular boundary conditions.
-The update rule is simple:
-if a cell has 3 live neighbors,
-or has 2 live neighbors and is already alive,
-it is alive in the next generation.
-In any other situation,
-the cell becomes, or stays, dead.
-
-\subsection{{\tt{winnow}}:
-	Weighted Point Selection
-	\label{s:toys-winnow}}
-
-This module converts a matrix of integer values
-to a vector of points,
-represented as $x$ and $y$ coordinates.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix, whose values are used as masses.
-\item[{\tt{mask}}:]
-	a Boolean matrix showing which points are eligible for consideration.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{nelts}}:]
-	the number of points to select.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points.
-\end{description}
-Each location where {\tt{mask}} is {\tt{true}}
-becomes a candidate point,
-with a weight equal to the integer value in {\tt{matrix}} at that location
-and $x$ and $y$ coordinates
-equal to its row and column indices.
-These candidate points are then sorted into increasing order by weight,
-and {\tt{nelts}} evenly-spaced points selected to create the result vector.
-
-\subsection{{\tt{norm}}:
-	Point Location Normalization
-	\label{s:toys-norm}}
-
-This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-Its inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of point locations.
-\item[{\tt{nelts}}:]
-	the number of elements in the point vector.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of normalized point locations.
-\end{description}
-If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
-then the normalization equation is:
-\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
-$y$ coordinates are normalized in the same fashion.
-
-\subsection{{\tt{elastic}}:
-	Elastic Net Simulation
-	\label{s:toys-elastic}}
-
-This module uses the elastic net algorithm \cite{b:elastic-net} to find an approximate solution to the travelling salesentity problem (TSP).
-A closed, circular loop, modelled by a large number of doubly-connected points, is placed in the plane.
-This path is slowly deformed until it connects the points representing the cities to be visited (Figure~\ref{f:elastic}).
-The rate at which the loop is deformed determines the quality of the final path.
-
-\begin{figure}
-\begin{center}
-\framebox{\epsffile{fig/elastic1.ps}}
-\framebox{\epsffile{fig/elastic2.ps}}
-\framebox{\epsffile{fig/elastic3.ps}}
-\end{center}
-\caption{Relaxation of an Elastic Net\label{f:elastic}}
-\end{figure}
-
-Let the (fixed) location of city $i$ be $c_i$, and the (changeable) location of point $j$ on the elastic net be $p_j$.
-The change in the location of point $j$ during each iteration of the algorithm is given by:
-\begin{equation}
-{\delta}{p_j} = {\alpha}\sum_{i}{w_{ij}(c_i - p_j)} ~+~ K{\beta}(p_{j+1} - 2p_j + p_{j-1})
-\end{equation}
-$\alpha$ and $\beta$ are time-invariant factors relating the strengths of the city-to-point and point-to-point forces,
-and $K$ is a time-dependent factor which controls the rate of annealing.
-The strength of the force $w_{ij}$ between city $c_i$ and point $p_j$ is given by:
-\begin{equation}
-w_{ij} = {\phi}({\mid}c_i - p_j{\mid}, K) / \sum_{\ell}\phi({\mid}c_i - p_{\ell}{\mid}, K)
-\end{equation}
-where $\phi(d, K)$ is the Gaussian function $e^{-{d^2}/{K^2}}$.
-Note the normalization of each $w_{ij}$; in order for this algorithm to work properly, the total force being exerted by each city must be the same.
-This is accomplished by summing the forces exerted by each city, and then normalizing the city's forces by that total.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{cities}}:]
-	a vector of points representing the locations of the cities to be toured.
-	The $(x,y)$ coordinates of these cities must be scaled to lie in the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-\item[{\tt{n\_cities}}:]
-	the number of cities in {\tt{cities}}.
-\item[{\tt{n\_net}}:]
-	the number of elastic net points to use.
-\item[{\tt{n\_iters}}:]
-	the total number of iterations to perform.
-\item[{\tt{n\_relax}}:]
-	the number of iterations to perform with each discrete value of $K$.
-	After each {\tt{n\_relax}} iterations, $K$ is reduced by a (constant) pre-defined scaling factor.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{net}}:]
-	a vector of {\tt{n\_net}} points representing the tour found.
-\end{description}
-
-\subsection{{\tt{outer}}:
-	Outer Product
-	\label{s:toys-outer}}
-
-This module turns a vector containing point positions
-into a dense, symmetric, diagonally dominant matrix
-by calculating the distances between each pair of points.
-It also constructs a real vector whose values are
-the distance of each point from the origin.
-Inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
-\item[{\tt{nelts}}:]
-	the number of points in the vector, and the size of the matrix along each axis.
-\end{description}
-Its outputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a real matrix, whose values are filled with inter-point distances.
-\item[{\tt{vector}}:]
-	a real vector, whose values are filled with origin-to-point distances.
-\end{description}
-Each matrix element $M_{i,j}$ such that $i \neq j$
-is given the value $d_{i, j}$,
-the Euclidean distance between point $i$ and point $j$.
-The diagonal values $M_{i, i}$ are then set to
-{\tt{nelts}} times the maximum off-diagonal value
-to ensure that the matrix is diagonally dominant.
-The value of the vector element $v_i$ is set to
-the distance of point $i$ from the origin,
-which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.
-
-\subsection{{\tt{gauss}}:
-	Gaussian Elimination
-	\label{s:toys-gauss}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using explicit reduction.
-(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
-there is a well-formed solution to the equation.)
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{sor}}:
-	Successive Over-Relaxation
-	\label{s:toys-sor}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using successive over-relaxation.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\item[{\tt{tolerance}}:]
-	the solution tolerance, e.g.\ $10^{-6}$.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{product}}:
-	Matrix-Vector Product
-	\label{s:toys-product}}
-
-Given a matrix $A$,
-a vector $V$,
-and an assumed solution $X$ to the equation $AX=V$,
-this module calculates the actual product $AX={V^{\prime}}$,
-and then finds the magnitude of the error.
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{actual}}:]
-	the real vector $V$.
-\item[{\tt{candidate}}:]
-	a real vector containing the supposed solution.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-The output of this function is:
-\begin{description}
-\item[{\tt{e}}:]
-	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
-\end{description}
-
-\section{Other Issues\label{s:issues}}
-
-As described in Section~\ref{s:method-softeng},
-implementors are required to chain modules together in the manner shown in Figure~\ref{f:chaining}.
-This section describes the flow of control and data between these chained problems,
-and discusses several other important aspects of this suite,
-including I/O and visualization.
-
-\subsection{Chained Problem Sequences\label{s:issues-chain}}
-
-The chained implementation of these problems
-executes individual toys in the following order:
-\begin{enumerate}
-\item	An integer matrix $I$ with $r$ rows and $c$ columns is created
-	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
-	\label{e:chain-choice-1}
-\item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
-\item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
-	to generate a Boolean mask $B$ from $I$ in which the (minimum) percentage of {\tt{true}} locations is $P$.
-	Like $I$, $B$ has $r$ rows and $c$ columns.
-	\label{e:chain-choice-2}
-\item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
-	This step overwrites the Boolean matrix $B$.
-\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
-\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
-\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
-\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
-	as described in Section~\ref{s:toys-outer}.
-\item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
-	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
-	This measures the agreement between the solutions found by the two methods.
-\end{enumerate}
-Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
-there are four possible chained sequences.
-A minimal implementation will provide one of these,
-while a full implementation will be a single program which can execute any of the four sequences.
-Intermediate implementations providing all four sequences as separate programs are also acceptable.
-
-\subsection{Input and Output\label{s:issues-io}}
-
-I/O is an important part of programming,
-but is often treated as being of secondary importance by language designers.
-This suite requires all stand-alone toys to read input values from files, and write results back;
-the chained version must be able to checkpoint intermediate results between toys.
-Finally,
-implementors are strongly encouraged to include some means of visualizing the output or evolution of individual toys.
-
-The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
-A file containing a vector begins with 1 positive integer,
-which specifies the number of elements in the vector;
-a file containing a matrix begins with 2 positive integers,
-which specify the number of rows and columns in the matrix respectively.
-Elements of the vector or matrix then appear one per line,
-from least-indexed to greatest-indexed
-(and in row-major order for matrices).
-
-The formats described above are required to be human-readable.
-Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
-using whatever file formats are convenient.
-This will allow programmers to demonstrate the ``natural'' I/O capabilities of particular systems
-which most probably be used for checkpointing intermediate results in real programs.
-
-\begin{figure}
-\epsfxsize=0.75\textwidth
-\begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
-\caption{Cowichan File Formats\label{f:io-formats}}
-\end{figure}
-
-\subsection{Visualization\label{s:issues-vis}}
-
-Visualization is an increasingly important part of scientific and numerical computing.
-Including it as an optional part of this problem set is also a good way to exercise both
-the use of libraries within a parallel programming system,
-and the ease with which a programming system can support variant versions of a program.
-
-Our requirements for the graphics interface are:
-\begin{enumerate}
-\item	Visualization for individual toys should be as simple as possible,
-	and should require as few changes to the basic toy code as possible.
-\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
-	The latter should create all images in a single window.
-\item	A single source file should implement both the graphical and non-graphical versions of each toy.
-\end{enumerate}
-The first two points require the system to distinguish between the actual drawing code
-(e.g.\ drawing a vector of points as circles)
-and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
-While the stand-alone toys use a whole window for drawing,
-the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
-At present,
-this is implemented by using a call counter
-in the interfaces for the individual toys and in the chained module.
-The actual drawing routines called from within each toy could then be the same.
-
-The present visualization routines were originally written by Anil Sukul,
-a student at the Vrije Universiteit, Amsterdam.
-The package uses the public domain graphics library VOGLE,
-available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
-VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
-as well as screen control routines (such as double buffering).
-VOGLE was chosen because of its simplicity,
-and because it includes Pascal and Fortran interfaces.
-
-\subsection{Reproducibility\label{s:issues-reproduce}}
-
-Reproducibility is an important issue for parallel programming systems.
-While constraining the order of operations in a parallel system to guarantee reproducibility makes programs in that system easier to debug,
-it can also reduce the expressiveness or performance of the system.
-
-In this problem set, irreproducibility can appear in two forms: numerical and algorithmic.
-The first arises in toys such as {\tt{gauss}}, {\tt{sor}}, and {\tt{elastic}}, which use floating-point numbers.
-Precisely how round-off errors occur in these calculations can depend on the distribution of work among processors,
-or the order in which those processors carry out particular operations.
-
-Irreproducibility also arises in toys which only use exact numbers, such as {\tt{invperc}} and {\tt{randmat}}.
-In the former, the percolation region is grown by repeatedly filling the lowest-valued cell on its perimeter.
-If several cells have this value, implementations may choose one arbitrarily.
-Thus, different implementations may produce very different fractal shapes.
-In the case of random number generation, the simplest thing to do is to run the same generator independently on each processor,
-although the values in the resulting matrix then depend on the number of processors used.
-
-\section{Conclusion\label{s:conclude}}
-
-A full ANSI~C version of this problem suite has been written, and is available by anonymous FTP from:
-\begin{center}
-{\tt{ftp.cs.toronto.edu :: pub/gvw/cowichan/src.tar.gz}}
-\end{center}
-This release includes two parallel versions,
-both based on the POSIX threads package.
-The first uses repeated fork-and-join parallelism,
-while the second creates a fixed number of threads at the beginning of the program,
-then synchronizes them repeatedly using barriers.
-
-We would like to invite interested groups to participate in this project by parallelizing the problem suite using their favourite language or tool.
-Our aim is to make both the implementations and their descriptions available as a hypertext document through the World-Wide Web,
-and, if there is sufficient interest, to publish this material in book form.
-If you would like to contribute to this effort, please contact either of the first two authors.
-
-\bibliographystyle{plain}
-\newcommand{\bibea}{{\em{et~al.}}}
-\bibliography{paper}
-
-\appendix
-
-\newpage
-\section{Parallel Clich\'{e}s\label{s:cliche}}
-
-\subsection{Operations\label{s:cliche-ops}}
-
-This list details some operations which are supported by many parallel programming systems.
-The toys described in Section~\ref{s:toys} provide opportunities for exercising many of these,
-and implementors are encouraged to phrase discussion of their work in terms of these operations where appropriate.
-
-\begin{itemize}
-\item	elementwise operations on arrays (unary, binary, and scalar promotion)
-\item	cumulative operations on arrays (reduction and parallel prefix)
-\item	array re-shaping and re-sizing (e.g.\ sub-array extraction)
-\item	partial (conditionally masked) versions of all of the above
-\item	regular data motion (e.g.\ circular and planar shifting)
-\item	irregular data motion (e.g.\ 1-to-1, 1-to-many, and many-to-1 permutation)
-\item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
-\item	differential local addressing (i.e.\ subscripting an array with an array)
-\item	full or partial replication of shared read-only values
-\item	full or partial replication of shared read-mostly values with automatic consistency management
-\item	structures with rotating ownership, suitable for migration
-\item	producer-consumer structures
-\item	partitionable structures
-\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
-\item	dynamic re-partitioning (e.g.\ for load balancing)
-\item	committed mutual exclusion (e.g.\ waiting on a semaphore)
-\item	uncommitted mutual exclusion (e.g.\ lock or fail)
-\item	barrier synchronization
-\item	multiple concurrent barriers used by non-overlapping groups of processes
-\item	fetch-and-add, and other fetch-and-operate functions
-\item	pre-scheduled receipt of messages of known composition
-\item	variable-length message receipt
-\item	message selection by type
-\item	message selection by source
-\item	message selection by contents
-\item	guarded message selection (e.g.\ CSP's {\tt{alt}} construct)
-\item	broadcast and partial broadcast
-\item	split-phase (non-blocking) operations
-\item	data marshalling and unmarshalling of ``flat'' structures (e.g.\ arrays of scalars)
-\item	data marshalling and unmarshalling of nested or linked structures
-\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
-\item	pipelining
-\item	distributed linked structures in systems without physically-shared memory
-\item	indexing of distributed shared arrays
-\item	collective I/O
-\item	uniform-size interleaved I/O
-\item	heterogeneous-size interleaved I/O
-\item	independent I/O operations on a single shared file
-\end{itemize}
-
-\subsection{Memory Reference Patterns\label{s:cliche-memref}}
-
-These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
-Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.
-
-\begin{itemize}
-\item	Write-once:
-	variables which are assigned a value once, and only read thereafter.
-	These can be supported through replication.
-\item	Write-many:
-	variables which are written and read many times.
-	If several processes write to the variable concurrently, they typically write to different portions of it.
-\item	Producer-consumer:
-	variables which are written to by one or more objects, and read by one or more other objects.
-	Entries in a shared queue, or a mailbox used for communication between two processes, are examples.
-\item	Private:
-	variables which are potentially shared, but actually private.
-	The interior points of a mesh in a program which uses geometric decomposition fall into this category,
-	while boundary points belong to the previous class.
-\item	Migratory:
-	variables which are read and written many times in succession by a single process before ownership passes to another process.
-	The structures representing particles in an $N$-body simulation are the clearest example of this category.
-\item	Result:
-	Accumulators which are updated many times by different processes, but thereafter read.
-\item	Read-mostly:
-	Any variable which is read much more often than it is written to.
-	The best score so far in a search problem is a typical instance of this class:
-	while any process might update it, in practice processes read its value much more often than they change it.
-\item	Synchronization:
-	Variables used to force explicit synchronization in a program, such as semaphores and locks.
-	These are typically ignored for long periods, and then the subject of intense bursts of access activity.
-\item	General read-write:
-	Any variable which cannot be put in one of the above categories.
-\end{itemize}
-
-\newpage
-\section{Comparing Parallel Programming Systems\label{s:questions}}
-
-This list of questions is a modified version of the one presented in \cite{b:compare-ada-c-pascal}.
-It is presented here as an organizational aid;
-when describing the PPS they have used,
-implementors are encouraged to use the order and terminology of this list where appropriate.
-We acknowledge that many of these questions will not apply to particular systems,
-and encourage implementors to ask questions which we have left out.
-
-\begin{enumerate}
-\item History and Philosophy
-  \begin{enumerate}
-  \item For whom and for what purpose was the language designed?
-  \end{enumerate}
-\item Syntax
-  \begin{enumerate}
-  \item Model implemented (Figure~\ref{f:taxonomy})
-  \item Class
-    \begin{enumerate}
-    \item library called from existing sequential language?
-    \item syntactic extensions to existing sequential language?
-    \item new language in Backus (Fortran/Algol/C) tradition?
-    \item new (non-Backus) language?
-    \end{enumerate}
-  \item Concurrency mechanisms adhere to same philosophy as original
-	language (where appropriate)?
-  \item Readability?  Syntactic consistency?  Conciseness?
-	(on an APL-to-Modula scale)
-  \item Common errors (e.g.\ confusing {\tt{=}} and {\tt{==}} in C)?
-  \end{enumerate}
-\item Type Philosophy and Data Types
-  \begin{enumerate}
-  \item What primitive types does the language support?
-  \item How are new types created?
-  \item What intrinsic operations are defined on user-defined types?
-  \item How strict is typing?
-	How is type conversion done?
-  \item Can type information be inspected at run-time?
-  \item What concurrent operations on aggregates
-	(such as lists and arrays) are intrinsic?
-	What kinds of concurrent aggregate operations can users define?
-  \item How are aggregates defined (e.g.\ by index set, by shape)?
-	Can new kinds of aggregates (e.g.\ templates for arrays)
-	be defined?
-  \item How strict is typing during communication?
-  \item What support is there for communication or manipulation
-	of data of non-scalar or non-intrinsic types?
-  \item Can structural information about aggregates
-	(e.g.\ number of elements in a set, dimension of an array)
-	be inspected at run-time?
-  \item What built-in support is there for object classes and inheritance?
-	For polymorphism?
-	For class (as opposed to instance) data?
-  \item Are functions first-class?
-  \item Are continuations supported?  Are they first-class?
-  \item Is dynamically-allocated memory supported?
-	Is it done automatically?
-	Is garbage-collection done automatically?
-	Can a program ever create a dangling pointer?
-	An alias?
-  \end{enumerate}
-\item Operators
-  \begin{enumerate}
-  \item What intrinsic operators are provided (brief list)?
-	How strongly do these type-check their arguments?
-  \item How are new operators defined?
-  \item Are the precedence rules straightforward?
-  \item Are side-effects/mutation allowed inside non-call expressions
-	(e.g.\ C's side-effecting \verb`i++` construct)?
-  \item Are their results platform-independent?
-  \item Are the results of intrinsic parallel operations reproducible?
-	Are they platform-independent?
-  \item Are heterogeneous operations on aggregates allowed
-	(e.g.\ every element of a list of polygons calculates its
-	center of mass)?
-	If so, how are these described, and how are they implemented?
-  \item Which of the operators in Section~\ref{s:cliche-ops} are supported?
-  \end{enumerate}
-\item Control Flow
-  \begin{enumerate}
-  \item What synchronization is automatically imposed during execution
-	(i.e.\ at what points can definite assertions about the state
-	of the running program be made)?
-  \item What control constructs does the language support?  Encourage?
-  \item How is concurrency expressed?  How structured is this mechanism?
-  \item Can the degree of concurrency be throttled (limited)?
-	Is this done by the system, or by the user?
-  \item How are exceptions handled/reported?
-	Is it different for intrinsic operations and user-defined operations?
-  \end{enumerate}
-\item Subroutines, Scope, and Modularization
-  \begin{enumerate}
-  \item How efficient is subroutine call compared to in-line execution
-	(i.e.\ how expensive are parameter passing and non-local control flow)?
-  \item How much information about parameters is available (required)
-	inside called subroutine
-	(e.g.\ whether object is shared with other processes or private
-	to executing process)?
-  \item Are side-effecting procedures allowed?
-  \item Can generic routines be written
-	(e.g.\ to work on different types or array shapes)?
-  \item Is there a restriction on return values (e.g.\ can a function return
-	any object which could be declared)?
-  \item Are there levels of scope?
-	If multiple levels are permitted, how are references to stale scopes
-	handled or prevented?
-	Can information for declarations (such as size) be inherited (dynamically)
-	from higher scopes?
-  \item What synchronization requirements (if any) are imposed during calls
-	(i.e.\ do all processes (in a group?) have to enter/exit the same subroutine
-	at the same time)?
-  \item What support is there for modularization?
-	For hiding implementation details?
-	For sharing such information?
-	For detecting and resolving name clashes?
-  \end{enumerate}
-\item Concurrent Programming Facilities
-  \begin{enumerate}
-  \item At what level(s) is the system concurrent?
-    \begin{enumerate}
-    \item implicit (the user shouldn't ask such things)
-    \item operational (e.g.\ data-parallel addition)
-    \item program counter (single executable, multiple states)
-    \item procedural (heterogeneous control parallelism)
-      \begin{enumerate}
-      \item one execution thread per address space
-      \item multiple threads per address space
-      \end{enumerate}
-    \end{enumerate}
-  \item What mutual exclusion and synchronization primitives are provided?
-  \item At what level(s) is data sharing visible to the programmer?
-    \begin{enumerate}
-    \item sequential consistency
-    \item sequential consistency within processor clusters
-    \item object-level consistency
-    \item there is no sharing
-    \end{enumerate}
-  \item What is the conceptual granularity of operations?
-    \begin{enumerate}
-    \item synchronization after atomic instruction
-    \item homogeneous operations with large-grain synchronization
-    \item heterogeneous operations with large-grain synchronization
-    \item completely asynchronous execution
-    \end{enumerate}
-  \item How do processes communicate data values?
-	How do they synchronize?
-	Are non-blocking operations (ones in which buffers might be
-	in a visibly volatile state) allowed?
-	How are they checked for termination?
-  \item If data are communicated explicitly,
-	how does the receiver choose what to accept?
-    \begin{enumerate}
-    \item by address (e.g.\ channel or port)
-    \item by pattern matching on contents
-    \item by explicit message tagging
-    \item by provision of a typed buffer
-    \item by provision of an untyped (byte-block) buffer
-    \item in FIFO order
-    \end{enumerate}
-  \item How do users specify the number of processors on which to run?
-	Must this be built into the program,
-	or can it be decided at run-time?
-	Is there support for executing on heterogeneous processors?
-  \item Can the user specify the mapping of processes to processors?
-	If so, how?
-	How tightly coupled are the mapping of processes to processors,
-	and the mapping of distributed data structures (if supported)?
-  \item Can the user specify the mapping of data structures to processors?
-        Can data structures be decomposed or partitioned?
-  \item What support is there for operations on groups of processes, such as:
-    \begin{enumerate}
-    \item broadcast
-    \item multicast (partial broadcast)
-    \item barrier synchronization
-    \end{enumerate}
-  \item Can arbitrary process groups be created,
-	or must they be structured in some way
-	(e.g.\ as logical mesh)?
-	Can operations involving disjoint groups be executed concurrently
-	(e.g.\ a broadcast within one group,
-	and a barrier within another)?
-  \end{enumerate}
-\item Input/Output
-  \begin{enumerate}
-  \item Is simple sequential (single-process) I/O simple to write?
-  \item Is concurrent I/O supported?  If so, which flavours:
-    \begin{enumerate}
-    \item single-reader-broadcast/collect-single-writer
-    \item homogeneously-sized or -strided segments
-    \item heterogeneously-sized or -strided segments
-    \item arbitrary independent operations
-    \end{enumerate}
-  \item What types can be read/written intrinsically?
-	E.g.\ can an entire record (or array, or list)
-	be read or written in a single operation?
-  \end{enumerate}
-\item Access to Routines in Other Languages and to the Hardware
-  \begin{enumerate}
-  \item Can subroutines written in other languages be called?
-        Can subroutines in this language be called from other languages?
-  \item Are there ways to access the run-time system directly,
-	or similar low-level facilities?
-  \item Is there high-level access to physical addresses and devices?
-  \end{enumerate}
-\item Practical Considerations
-  \begin{enumerate}
-  \item Are both batch and interactive processing supported?
-  \item How reliable and efficient are existing implementations?
-	How fast is it (i.e.\ how quickly does code compile and load)?
-	Is it supported (who fixes it if it breaks)?
-  \item Is there a standard for the system (from some standards organization)?
-  \item Is the documentation adequate? For whom (CS graduates, computer-literate
-	users, the general population)? Is training available?	Consulting?
-  \item How much code has been written (lines of code or weeks of effort)?
-        By whom (computer science undergraduates, professional engineers, etc.)?
-  \item How long does it take the compiler to generate a ``small'' executable
-        (such as {\tt{"hello, world"}})?
-  \item How large is a ``small'' executable?
-        How long does it take to load such an executable?
-  \end{enumerate}
-\item Supporting Tools
-  \begin{enumerate}
-  \item What is the development environment like?
-        Can code be developed on a workstation and then moved to a parallel platform?
-  \item Are there debugging aids, performance analysers, and management tools?
-  \item How much help do these really provide?
-  \end{enumerate}
-\end{enumerate}
-
-\newpage
-\section{A Critique of the C Implementation\label{s:critique}}
-
-This section is a critique of the ANSI~C reference implementation of the problem suite.
-It is intended to serve as a model for (self-)criticism of other programming systems.
-
-\begin{itemize}
-\item	C's support for multi-dimensional arrays (MDAs) is very weak.
-	There is no way to dynamically allocate an MDA in a single go---one must either
-	allocate a block of the same size as the desired array,
-	and then do indexing calculations by hand,
-	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
-	MDAs do not carry dimension information with them,
-	so it is impossible to determine the size of an array parameter within a function.
-	Finally,
-	C does not treat all axes of an array equally:
-	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
-	it is impossible to slice it along the other axis.
-\item	C does not distinguish between Boolean and integer types.
-	As a result,
-	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
-	can be passed in reverse order without a type error.
-	Using {\tt{typedef}} to create a Boolean type name does not solve this
-	(at least, not in {\tt{gcc}} V2.5.8).
-\item	Union types cannot be safely initialized.
-	This complicated the implementation of the graphics interface,
-	where it would have been much more elegant to define a union type,
-	each of whose variants held parameter specifications for a single toy.
-	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
-	initializing some and filling other with don't-care values.
-	This is neither safe nor elegant.
-\item	Parameter values cannot be used in the initialization of local variables inside functions.
-	In particular,
-	it is not possible to create a local vector with a length specified by an input parameter.
-	Such a facility would be useful in the {\tt{winnow}} toy,
-	where we have instead allocated local temporaries of the maximum possible size.
-\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
-	For example, the ``fail()'' error-handling routine was buggy because
-	all \verb`real` arguments were being automatically converted to doubles,
-	but were being taken off the stack as floats.
-\item	no intrinsic notion of group ID/self ID in threads
-\item	packaging parameters for threads
-\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
-\end{itemize}
-
-\end{document}

### 2009-08-28 paper/2009-09/issues.tex
Breaking paper into manageable chunks

--- 2009-08-20 paper/2009-09/paper.tex
@@
-\documentstyle[epsf,11pt]{article}
-
-\begin{document}
-
-\begin{Large}\begin{center}
-The Cowichan Problems: \\
-Assessing the Usability of Parallel Programming Systems
-\end{center}\end{Large}
-
-\vspace{\baselineskip}
-
-\begin{center}
-\begin{tabular}{cccc}
-Gregory V.~Wilson\footnote{\verb`http://www.third-bit.com`}
-&
-Andrew Borzenko
-&
-Cameron Gorrie
-&
-Zachary Kincaid
-\end{tabular}
-\end{center}
-
-\vspace{\baselineskip}
-
-\begin{abstract}
-Most programmers believe that parallel programming is harder than sequential programming.
-Many also believe that today's parallel programming systems make it even harder.
-We have developed a suite of chained toy problems that can be used to assess
-the usability of different scientific programming languages, parallel or otherwise.
-This paper motivates and describes the suite,
-and discusses what we have learned from porting it to four parallel C++ systems
-(MPI, OpenMP, Intel Thread Building Blocks, and a Linux tuple space system),
-and from translating it into sequential and parallel Haskell.
-\end{abstract}
-
-\section{Introduction\label{s:intro}}
-
-Twenty years ago,
-Bal {\em{et~al.}\/} listed more than 300 parallel programming systems \cite{b:par-lang-survey}.
-A similar survey today could well uncover twice as many,
-but a poll of application programmers would probably find that
-only a handful are in general use.
-
-The reasons why so many of these PPSs are only used by their developers (if at all) are murkier.
-Part of the blame undoubtedly lies with the diversity of hardware,
-and the attendant lack of portability,
-alluded to above.
-A second reason is that an academic researcher's principal interest is developing new languages,
-not supporting existing ones.
-By the time a system is 90\% complete,
-its author is already thinking about its successor.
-Few application programmers are willing to invest effort porting codes
-to systems which are not properly supported,
-and liable to change or evaporate without notice.
-Third,
-many research groups have chosen to explore the potential of parallelism
-without consideration for compatibility with older languages.
-Application programmers must therefore face the prospect of completely re-writing their programs,
-rather than simply porting them.
-As the next section argues,
-this can obviate any performance that might be gained from parallelization.
-Finally,
-many PPSs simply cannot support the software engineering load of large applications.
-Many provide only a single global namespace,
-do not include I/O capabilities,
-require users to marshal and unmarshal data explicitly,
-or do not permit hardware to be timeshared.
-
-The net result is that many PPSs are never used for anything more complicated than
-generating the Mandelbrot Set,
-simulating the Game of Life,
-multiplying dense matrices,
-or finding all solutions to the $N$-queens problem for some small $N$.
-Lacking the feedback that comes from more demanding use,
-few systems are improvements over their predecessors
-in the way that successive versions of operating systems have improved on one another,
-or the languages in the Algol$\rightarrow$Modula-$n$ family
-have become both more powerful and more secure.
-Most applications programmers therefore choose to play it safe
-and stick to parallel programming's lowest common denominators:
-data-parallel Fortran,
-and message-passing libraries.
-
-The historical reasons for tolerating this situation are becoming less valid with each passing year.
-When only a few parallel computers existed,
-and these were of widely different types,
-software standardization simply didn't make sense.
-Similarly,
-when supercomputers were very expensive,
-the cost of software development seemed relatively smaller,
-and the requirement for maximum performance relatively greater,
-than in sequential computing.
-As \cite{b:social-limits-speed} points out,
-however, these arguments are losing force.
-Now,
-as in sequential computing,
-the first question is not ``How do I write programs for this machine?'',
-but ``Will this machine run my programs?''
-As microprocessor-based MIMD hardware comes to dominate massively-parallel computing,
-we see increasing emphasis on standards such as High-Performance Fortran (HPF) \cite{b:hpf-overview},
-and the Message-Passing Interface (MPI) \cite{b:mpi-overview}.
-
-But herein lies a danger.
-History shows that it is extremely difficult
-to displace a programming language or system once it becomes established.
-Cobol and MS-DOS are the most frequently cited examples,
-but many systems which should have been superceded long ago have managed
-not only to cling to life,
-but to grow in influence.
-(The fact that the mail signature
-``The last good thing written in C was Schubert's Ninth Symphony''
-can elicit smiles simply proves that even the truth can be funny.)
-Established systems evolve slowly, if at all.
-It would therefore seem that the best time to guarantee ourselves
-a good parallel programming paradigm is right now,
-before a large number of large applications are created and take on lives of their own.
-
-We therefore propose to use a set of simple programming problems to compare
-the usability of parallel programming systems.
-Competent programmers, fluent in a specific system,
-will implement solutions to these problems and report their experiences
-in terms of development time, code size and clarity, and runtime efficiency.
-These problems are on the scale of the toy applications described several paragraphs ago
-both to keep implementation time low
-and because we feel that any PPS produced today ought to be able to support such simple programs.
-As the second author put it,
-``We're all sick of seeing these,
-but maybe if we all catch the disease at the same time we'll finally build up some immunity to it.''
-One significant innovation in this work is that
-we also require implementations of our suite to chain problems together,
-so that the output of one problem is the input to another.
-This will allow us to assess how well each PPS supports modularization and code re-use
-(or, more realistically, how much extra effort programmers must invest
-to achieve these goals when using a particular PPS).
-
-Section~\ref{s:corollary} motivates this work by showing
-how poor usability can limit the realizable parallelism in a program.
-Section~\ref{s:previous} surveys previous work on the comparison of parallel programming systems,
-while Section~\ref{s:method} outlines our choice of methods, and the constraints on our work.
-Section~\ref{s:toys} then presents the suite of problems we intend to use.
-In recognition of the original work done by Feo and others on
-the Salishan Problems~\cite{b:salishan}, this suite is called
-the Cowichan Problems\footnote{\noindent Pronounced {\em{Cow}\/}-i-chun.
-	Like Salishan, the word is a Northwest Coast Indian place name.}.
-
-\subsection*{History and Acknowledgments}
-
-The first author originally developed this suite in the mid-1990s,
-inspired both by \cite{b:salishan} and by
-discussions with R.\ Bruce Irvin (then at the University of Wisconsin),
-Henri Bal (Vrije Universiteit, Amsterdam),
-and others.
-Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.
-
-\section{Quantifying the Importance of Usability\label{s:corollary}}
-
-According to Amdahl's Law,
-the time $t$ required to execute a program has two components:
-$t_{seq}$, representing sequential operations that cannot be parallelized,
-and $t_{par}$, representing those operations which can.
-If we let $\sigma$ be the ``serial fraction'' of the program,
-so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
-then the speedup which can be achieved using $P$ processors is:
-\begin{eqnarray*}
-s(P)	& =	& \frac{t(1)}{t(P)}	\\
-	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
-\end{eqnarray*}
-As $P{\rightarrow}{\infty}$,
-the potential speedup is bounded by $1/{\sigma}$.
-
-What is not usually acknowledged is
-the influence of programming system usability on
-the potential speedup of sequential programs.
-As stated above, Amdahl's Law applies only to a single run of a program.
-In the real world, programs are first developed, and then run many times.
-If we let $T$ be the total running time of a program over its lifetime,
-$D_{seq}$ be the time required to develop the sequential components of a program,
-and $D_{par}$ be the time required to develop the parallel components
-(or parallelize the sequential components),
-then the lifetime speedup we can achieve is given by:
-\begin{eqnarray*}
-S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
-	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
-\end{eqnarray*}
-Assume that we are parallelizing a ``dusty deck'',
-i.e.\ that the time $D_{seq}$ has already been invested.
-If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
-then achievable speedup is limited by:
-\begin{eqnarray*}
-S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
-		& =	& \frac{1}{{\phi} + {\sigma}}
-\end{eqnarray*}
-The time required to parallelize a program can therefore be seen
-as increasing that program's effective serial fraction.
-Unless parallelization time can be substantially reduced,
-massively-parallel computing is likely to remain attractive only in those cases in which:
-\begin{itemize}
-\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
-\item	the expected total program runtime is very large
-	(i.e.\ the program is a package used extensively in a field
-	such as computational chemistry or vehicle crash simulation);
-	or
-\item	cost is not an obstacle (i.e.\ the customer is the NSA).
-\end{itemize}
-
-\section{Previous Work\label{s:previous}}
-
-In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
-\begin{quote}
-As practiced by computer science,
-the study of programming is an unholy mixture of mathematics,
-literary criticism,
-and folklore.
-\end{quote}
-Little has changed since.
-Computer scientists pay lip service to the importance of usability,
-but are strangely reluctant to try to quantify it,
-particularly with respect to the tools they use themselves.
-Religious debates about the ``right'' first-year teaching language
-or the ``best'' GUI toolkit are common;
-systematic use of the experimental method is not.
-
-Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
-such as readability and code-pattern recognition
-\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
-The little work that has been done on ``programming in the large'' has been anecdotal,
-such as \cite{b:compare-ada-c-pascal}.
-This book presented various authors' impressions of Pascal, C, and Ada
-(before any complete Ada implementations existed).
-Its editors formulated a list of questions to use in comparing languages,
-but did not use it themselves.
-A modified version of this questionnaire,
-tailored for parallel programming languages,
-is given in Appendix~\ref{s:questions},
-and will be used as a guideline in our work.
-
-Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
-The first presented implementations of a simple numerical quadrature program
-in more than a dozen different parallel languages
-in use on mid-1980s hardware.
-The second presented implementations of
-the Salishan Problems---Hamming number generation,
-isomer enumeration,
-skyline matrix reduction,
-and a simple discrete event simulator---in
-C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
-and a variety of dataflow and functional languages.
-Both of these books convey the flavor of the languages they describe,
-but neither made any effort to compare languages or problem implementations directly.
-
-Inspired by the Salishan Problems,
-the first author developed a suite of seven problems \cite{b:cowichan-1}
-in order to test the limits of parallel programming systems.
-Individual problems were badly load-balanced,
-manipulated irregular pointer-based structures,
-and so on.
-Our experience implementing them was extremely valuable,
-as it uncovered many bugs in the programming system used,
-and revealed several shortcomings of the system's design.
-However,
-we also discovered that implementing each problem takes six to eight weeks,
-which we feel is prohibitive at this exploratory stage.
-
-\section{Methodology\label{s:method}}
-
-The main aim of this work is to assess and compare
-the ease with which programs can be developed and/or ported
-using different parallel programming systems.
-Here,
-we discuss issues related to choosing the problems that make up the suite,
-assessing the complexity of particular implementations,
-and other purposes which implementations of these might serve.
-Recognizing that these problems are extremely simple,
-and easy to parallelize,
-we refer to them as ``toys''.
-
-\subsection{Criteria for Selection\label{s:method-criteria}}
-
-Our criteria for including problems in this suite are as follows:
-\begin{enumerate}
-\item	Each toy should require no more than an afternoon
-	to write and test in a well-supported sequential language.
-	We feel that making individual problems more complicated will only discourage implementation.
-\item	The correctness of each implementation must be relatively easy to verify.
-	Toys whose output is easily visualized are therefore to be preferred,
-	as are toys whose results are insensitive to floating-point discretization effects.
-\item	Each toy application should be amenable to
-	an analysis of speedup or isoefficiency \cite{b:isoefficiency-tutorial}
-	to allow a comparison of theory and practice.
-\item	At least some toys should not be ``infinitely scalable''.
-	Many real-world applications are not,
-	and this suite should reflect such limitations.
-\item	At least some toys should require I/O,
-	since this important aspect of real-world programming is often neglected by PPS designers.
-\item	There should be some potential redundancy in the type of parallelization done,
-	so that re-use of software can be demonstrated.
-\item	Together,
-	the toys in the suite must exercise a wide range of common parallel operations
-	(Appendix~\ref{s:cliche-ops})
-	and memory reference patterns
-	(Appendix~\ref{s:cliche-memref}).
-\item	The ``obvious'' implementations of these toys should span
-	a wide range of parallel programming paradigms,
-	such as task farming,
-	geometric and functional decomposition,
-	and speculative parallelism.
-\item	Parallelizations for several (if not all) of
-	the types of programming systems shown in Figure~\ref{f:taxonomy} should be known.
-	This suite is not intended to be a collection of research problems;
-	reasonably simple and efficient implementations in most environments
-	should be straightforward to produce.
-\end{enumerate}
-
-\begin{figure}
-\epsfysize=0.5\textheight
-\begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
-\caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
-\end{figure}
-
-\subsection{Specification\label{s:method-spec}}
-
-Toys should be specified in terms of inputs and outputs,
-rather than algorithmically.
-For example,
-``sort $N$ integers in the range $1{\ldots}R$,
-where $R$ is not known in advance and there may be multiple instances of a given value''
-is to be preferred over
-``parallelize quicksort''.
-As \cite{b:slalom} argued,
-which parallel algorithm is best in a particular programming system
-depends critically on the architectural assumptions made by that system.
-Implementors should be free to employ those methods which they find natural and convenient.
-The only exception we make to this rule is in the two matrix equation solvers,
-where we require Gaussian elimination and successive over-relaxation.
-Given the enormous number of matrix solution algorithms in use,
-we felt we had to specify these particular algorithms in order to ensure that
-different implementations of this suite would be comparable.
-
-\subsection{Software Engineering Issues\label{s:method-softeng}}
-
-These toys will exercise many different aspects of parallel systems.
-However,
-their ``single algorithm per program'' model is not representative of real applications.
-These usually contain several discrete (and sometimes overlapping) phases,
-each of which is qualitatively different.
-A full implementation of this suite will therefore have two parts.
-In the first,
-each toy will be implemented as a stand-alone program.
-In the second,
-toys will be chained together to create the kind of phase behaviour seen in real applications
-(Figure~\ref{f:chaining}).
-This will test the ease with which heterogeneous parallelism can be mixed within a single program.
-It will also show how well the system supports code re-use and information hiding,
-which are crucial to the development of large programs.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
-\caption{Chaining\label{f:chaining}}
-\end{figure}
-
-A second reason for chaining toys is that
-many groups are now using algorithmic skeletons to structure parallel programs
-\cite{b:cole-skel,b:p3l-overview,b:enterprise}.
-Such skeletons encapsulate the details of a particular style of parallelism,
-and hide low-level or platform-dependent issues.
-We hope that chaining will show the strengths and weaknesses of such systems.
-
-Finally,
-chaining should be designed so that some toys can be executed concurrently.
-Many PPSs impose extraneous constraints on programs,
-e.g.\ require all processes to participate in every barrier synchronization,
-or require the same executable to be loaded onto each processor.
-These constraints can limit the exploitation of potential concurrency.
-Permitting, but not requiring, concurrent execution of several toys should uncover such limitations.
-
-\subsection{Sizing\label{s:method-size}}
-
-One crucial aspect of the specification of toys is
-the way in which the sizes of problems are determined.
-In a completely frozen model, the actual size of each problem would be compiled into each toy.
-A fluid implementation, by contrast, would allow sizes to be specified at run-time,
-and would only then allocate and partition data structures.
-We have chosen to use an intermediate model (which might be called ``slushy''),
-in which the maximum size of individual problems is specified during compilation,
-but the actual size of a problem is only determined when the toy begins to execute.
-We hope that the use of this model will force implementors to deal with
-at least some of the software engineering issues
-involved in building flexible large-scale applications,
-without making implementation prohibitively difficult.
-
-\subsection{Assessing Complexity\label{s:method-complex}}
-
-There are at least two different ways in which usability might be compared.
-One would be to measure the performance achieved by an ``average'' programmer
-as a function of time on each of a range of problems.
-This would lead to graphs of the kind shown in Figure~\ref{f:perf},
-where the left graph shows a particular result
-and the right graph a simplified abstraction of it.
-By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
-which is the length of vector on which a pipelined architecture achieves
-half its theoretical peak performance,
-we could in principle find the value of $p_{1/2}$---the programming time required to achieve
-half of a machine's peak performance\footnote{Note that
-	if performance was measured as a fraction of the figures quoted by manufacturers for their machines,
-	it is unlikely that the halfway mark would ever be reached.}
-for a particular combination of programming system and problem type.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
-\caption{Performance vs.\ Time\label{f:perf}}
-\end{figure}
-
-A second option would be to measure the complexities of
-implementations of a single application using different programming systems.
-With this approach,
-comparative performance or speedup figures would be supplemented by complexity measures,
-as shown in Figure~\ref{f:complex}.
-
-\begin{figure}
-\epsfxsize=6cm
-\begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
-\caption{Comparative Code Complexity\label{f:complex}}
-\end{figure}
-
-The drawback of the first approach is that there are no ``typical'' problems,
-or ``average'' programmers.
-Meaningful measurement of  $p_{1/2}$ would therefore require
-measurement of a large number of programmers working on many problems.
-Since this is too expensive to be practical,
-the second approach is the one we have chosen to adopt.
-It has drawbacks as well,
-most notably the fact that no-one has ever demonstrated any better measure of program complexity
-than a simple count of the number of lines in a program \cite{b:fenton-metrics}.
-Since we anticipate a wide variation in the types of progamming systems that will be used in this work
-(ranging from declarative dataflow languages
-through functional languages with Lisp-like syntax
-to parallel dialects of Fortran and C),
-and in the indentation, commenting, and layout styles of particular programmers,
-we may measure complexity by counting the number of tokens in the source code which are
-keywords,
-intrinsic operators,
-user-defined functions,
-type or variable declarations,
-and miscellaneous punctuation.
-
-\subsection{Other Uses for Implementations\label{s:method-uses}}
-
-This suite is intended for assessing PPS usability,
-but we envisage at least three other uses for implementations of it:
-\begin{description}
-\item[Related Research:]
-	We hope that the existence of this suite will indicate
-	to the developers of new parallel programming systems
-	what a mature tool should be able to support.
-	If sufficiently diverse,
-	the suite could also be used to help test the correctness of new or improved systems.
-	Finally,
-	this suite could complement existing performance benchmarks,
-	since performance figures for individual toys could be compared.
-\item[Tool Development:]
-	Requests from tool developers for existing parallel programs,
-	or for event traces from them,
-	are common on newsgroups such as {\tt{comp.parallel}}.
-	We hope that traces from implementations of this suite
-	will be used to compare different debugging and performance monitoring tools,
-	that their source code will be used to test
-	new parallelizing compilation techniques or mapping and scheduling algorithms,
-	and so on.
-\item[Teaching:]
-	Implementations will be suitable for use as classroom examples,
-	since they will be small enough to be understood quickly.
-	The toys themselves should also be suitable as classroom exercises
-	in a senior undergraduate course on parallel computing.
-\end{description}
-
-\subsection{Some Criticisms\label{s:method-criticism}}
-
-\noindent {\em{Why not use an existing benchmark suite for this purpose?}\/}
-
-Implementations of benchmark suites have usually been coded for absolute speed.
-They are not representative of ``normal'' practice,
-in which reasonable performance is the usual goal,
-and so are inappropriate for this work.
-In addition,
-most existing benchmark suites are too large to be implemented or ported quickly
-(e.g.\ SPEC~\cite{b:bench-over}),
-over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
-or concerned with a limited range of applications
-(e.g.\ the NAS Parallel Benchmarks~\cite{b:nas-results}).
-Finally,
-most exist only in C or Fortran (or both);
-many interesting parallel programming systems are built on top of other languages,
-including Modula-2, Scheme, Prolog, and various dataflow languages.
-We do not want the effort required to translate several large programs
-to discourage researchers from participating in this exercise.
-
-\vspace{\baselineskip}
-
-\noindent {\em{Won't the results depend primarily on programmer ability?}\/}
-
-The short answer to this question is yes,
-but no more than the results from performance benchmarks.
-A longer answer is that	since we intend to measure the combination of code complexity and performance achieved,
-rather than performance as a function of development time,
-programmers will be allowed to revise and improve their implementations when and as desired.
-
-\section{The Cowichan Problems\label{s:toys}}
-
-This section describes the problems making up the full problem suite.
-As discussed above,
-the implementation of each is to be used both as a stand-alone program
-and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
-Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cl|cl}
-Code	& Toy				& Code	& Toy				\\
-\hline
-E	& elastic net			& O	& outer product			\\
-G	& Gaussian elimination		& P	& matrix-vector product		\\
-H	& halving shuffle		& R	& random matrix generation	\\
-I	& invasion percolation		& S	& successive over-relaxation	\\
-L	& Game of Life			& T	& image thresholding		\\
-M	& Mandelbrot Set		& V	& vector difference		\\
-N	& point normalization		& W	& point value winnowing
-\end{tabular}
-\caption{Single-Character Codes for Toys\label{t:toy-codes}}
-\end{center}
-\end{table}
-
-\subsection{{\tt{mandel}}:
-	Mandelbrot Set Generation
-	\label{s:toys-mandel}}
-
-\begin{figure}
-\begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
-\caption{The Mandelbrot Set\label{f:mandel}}
-\end{figure}
-
-This module generates the Mandelbrot Set
-for a specified region of the complex plane
-(Figure~\ref{f:mandel}).
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{x0, y0}}:]
-	the real coordinates of the lower-left corner
-	of the region to be generated.
-\item[{\tt{dx, dy}}:]
-	the extent of the region to be generated.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing the iteration count at each point in the region.
-\end{description}
-
-Given initial coordinates $(x_0, y_0)$,
-the Mandelbrot Set is generated by iterating the equation
-\begin{eqnarray*}
-x^{\prime}	& =	& x^2 - y^2 + y_0	\\
-y^{\prime}	& =	& 2{x}{y} + x_0
-\end{eqnarray*}
-until either an iteration limit is reached,
-or the values diverge.
-The iteration limit used in this module is 150 steps;
-divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
-The integer value of each element of the matrix is
-the number of iterations done.
-Note that,
-as in all problems,
-the output is required to be
-independent of the number of processors used.
-This requires programmers to be careful about
-floating-point roundoff effects.
-
-\subsection{{\tt{randmat}}:
-	Random Number Generation
-	\label{s:toys-rng}}
-
-This module fills a matrix with pseudo-random integers.
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{s}}:]
-	the random number generation seed.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix filled with random values.
-\end{description}
-Note that,
-as in all problems,
-the output is required to be independent of the number of processors used.
-Generating new seed values from the given seed,
-and running one copy of the random number generator
-on each processor,
-is therefore unlikely to qualify as a solution.
-
-\subsection{{\tt{half}}:
-	Two-Dimensional Shuffle
-	\label{s:toys-half}}
-
-This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
-shuffles them,
-and then repeats this operation along the other axis.
-Values in odd-numbered locations are collected at the low end of each row or column,
-while values in even-numbered locations are moved to the high end.
-An example transformation is:
-
-\begin{eqnarray*}
-\begin{array}{cccc}
-a & b & c & d \\
-e & f & g & h \\
-i & j & k & l
-\end{array}
-& = &
-\begin{array}{cccc}
-a & c & b & d \\
-i & k & j & l \\
-e & g & f & h
-\end{array}
-\end{eqnarray*}
-
-Note that how an array element is moved depends only on whether its location index is odd or even,
-not on whether its value is odd or even.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing shuffled values.
-\end{description}
-
-\subsection{{\tt{invperc}}:
-	Invasion Percolation
-	\label{s:toys-invperc}}
-
-Invasion percolation models the displacement of one fluid
-(such as oil)
-by another
-(such as water)
-in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
-In two dimensions,
-this can be simulated by generating
-an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
-and then marking the center cell of the grid as filled.
-In each iteration,
-one examines the four orthogonal neighbors of all filled cells,
-chooses the one with the lowest value
-(i.e.\ the one with the least resistance to filling),
-and fills it in.
-Figure~\ref{f:invperc1} shows the first few steps in this process,
-while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{math}
-\begin{array}{ccc}
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & 29 & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \star & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & 29 & 77 \\
-\hline
-61 & \star & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \star & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \star & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \star \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & \star & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-\star & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & \star & 72 & 45 & 38 \\
-\hline
-\circ & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \circ \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\end{array}
-\end{math}
-\end{footnotesize}
-\caption{Invasion Percolation\label{f:invperc1}}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
-\end{center}
-\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
-\end{figure}
-
-The simulation continues until
-some fixed percentage of cells have been filled,
-or until some other condition
-(such as the presence of trapped regions)
-is achieved.
-The fractal structure of the filled and unfilled regions
-is then examined to determine how much oil could be recovered.
-The na\"{\i}ve way to implement this is to repeatedly scan the array;
-a more sophisticated, and much faster, sequential technique
-is to maintain a priority queue of unfilled cells
-which are neighbors of filled cells.
-This latter technique is similar to the list-based methods used in some cellular automaton programs,
-and is very difficult to parallelize effectively.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{nfill}}:]
-	the number of points to fill.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
-\end{description}
-Filling begins at the central cell of the matrix
-(rounding down for even-sized axes).
-
-\subsection{{\tt{thresh}}:
-	Histogram Thresholding
-	\label{s:toys-thresh}}
-
-This module performs histogram thresholding on an image.
-Given an integer image $I$ and a target percentage $p$,
-it constructs a binary image $B$
-such that $B_{i,j}$ is set
-if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
-The general idea is that an image's histogram should have 2 peaks,
-one centered around the average foreground intensity,
-and one centered around the average background intensity.
-This program attempts to set
-a threshold between the two peaks in the histogram
-and select the pixels above the threshold.
-This module's inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the integer matrix to be thresholded.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{percent}}:]
-	the minimum percentage of cells to retain.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
-\end{description}
-
-\subsection{{\tt{life}}:
-	Game of Life
-	\label{s:toys-life}}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
-\end{center}
-\caption{Snapshots from the Game of Life\label{f:life}}
-\end{figure}
-
-This module simulates the evolution of Conway's Game of Life,
-a two-dimensional cellular automaton (Figure~\ref{f:life}).
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the Life world.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{numgen}}:]
-	the number of generations to simulate.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the world after simulation.
-\end{description}
-At each time step,
-this module must count the number of live ({\tt{true}}) neighbors
-of each cell,
-using both orthogonal and diagonal connectivity
-and circular boundary conditions.
-The update rule is simple:
-if a cell has 3 live neighbors,
-or has 2 live neighbors and is already alive,
-it is alive in the next generation.
-In any other situation,
-the cell becomes, or stays, dead.
-
-\subsection{{\tt{winnow}}:
-	Weighted Point Selection
-	\label{s:toys-winnow}}
-
-This module converts a matrix of integer values
-to a vector of points,
-represented as $x$ and $y$ coordinates.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix, whose values are used as masses.
-\item[{\tt{mask}}:]
-	a Boolean matrix showing which points are eligible for consideration.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{nelts}}:]
-	the number of points to select.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points.
-\end{description}
-Each location where {\tt{mask}} is {\tt{true}}
-becomes a candidate point,
-with a weight equal to the integer value in {\tt{matrix}} at that location
-and $x$ and $y$ coordinates
-equal to its row and column indices.
-These candidate points are then sorted into increasing order by weight,
-and {\tt{nelts}} evenly-spaced points selected to create the result vector.
-
-\subsection{{\tt{norm}}:
-	Point Location Normalization
-	\label{s:toys-norm}}
-
-This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-Its inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of point locations.
-\item[{\tt{nelts}}:]
-	the number of elements in the point vector.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of normalized point locations.
-\end{description}
-If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
-then the normalization equation is:
-\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
-$y$ coordinates are normalized in the same fashion.
-
-\subsection{{\tt{elastic}}:
-	Elastic Net Simulation
-	\label{s:toys-elastic}}
-
-This module uses the elastic net algorithm \cite{b:elastic-net} to find an approximate solution to the travelling salesentity problem (TSP).
-A closed, circular loop, modelled by a large number of doubly-connected points, is placed in the plane.
-This path is slowly deformed until it connects the points representing the cities to be visited (Figure~\ref{f:elastic}).
-The rate at which the loop is deformed determines the quality of the final path.
-
-\begin{figure}
-\begin{center}
-\framebox{\epsffile{fig/elastic1.ps}}
-\framebox{\epsffile{fig/elastic2.ps}}
-\framebox{\epsffile{fig/elastic3.ps}}
-\end{center}
-\caption{Relaxation of an Elastic Net\label{f:elastic}}
-\end{figure}
-
-Let the (fixed) location of city $i$ be $c_i$, and the (changeable) location of point $j$ on the elastic net be $p_j$.
-The change in the location of point $j$ during each iteration of the algorithm is given by:
-\begin{equation}
-{\delta}{p_j} = {\alpha}\sum_{i}{w_{ij}(c_i - p_j)} ~+~ K{\beta}(p_{j+1} - 2p_j + p_{j-1})
-\end{equation}
-$\alpha$ and $\beta$ are time-invariant factors relating the strengths of the city-to-point and point-to-point forces,
-and $K$ is a time-dependent factor which controls the rate of annealing.
-The strength of the force $w_{ij}$ between city $c_i$ and point $p_j$ is given by:
-\begin{equation}
-w_{ij} = {\phi}({\mid}c_i - p_j{\mid}, K) / \sum_{\ell}\phi({\mid}c_i - p_{\ell}{\mid}, K)
-\end{equation}
-where $\phi(d, K)$ is the Gaussian function $e^{-{d^2}/{K^2}}$.
-Note the normalization of each $w_{ij}$; in order for this algorithm to work properly, the total force being exerted by each city must be the same.
-This is accomplished by summing the forces exerted by each city, and then normalizing the city's forces by that total.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{cities}}:]
-	a vector of points representing the locations of the cities to be toured.
-	The $(x,y)$ coordinates of these cities must be scaled to lie in the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-\item[{\tt{n\_cities}}:]
-	the number of cities in {\tt{cities}}.
-\item[{\tt{n\_net}}:]
-	the number of elastic net points to use.
-\item[{\tt{n\_iters}}:]
-	the total number of iterations to perform.
-\item[{\tt{n\_relax}}:]
-	the number of iterations to perform with each discrete value of $K$.
-	After each {\tt{n\_relax}} iterations, $K$ is reduced by a (constant) pre-defined scaling factor.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{net}}:]
-	a vector of {\tt{n\_net}} points representing the tour found.
-\end{description}
-
-\subsection{{\tt{outer}}:
-	Outer Product
-	\label{s:toys-outer}}
-
-This module turns a vector containing point positions
-into a dense, symmetric, diagonally dominant matrix
-by calculating the distances between each pair of points.
-It also constructs a real vector whose values are
-the distance of each point from the origin.
-Inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
-\item[{\tt{nelts}}:]
-	the number of points in the vector, and the size of the matrix along each axis.
-\end{description}
-Its outputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a real matrix, whose values are filled with inter-point distances.
-\item[{\tt{vector}}:]
-	a real vector, whose values are filled with origin-to-point distances.
-\end{description}
-Each matrix element $M_{i,j}$ such that $i \neq j$
-is given the value $d_{i, j}$,
-the Euclidean distance between point $i$ and point $j$.
-The diagonal values $M_{i, i}$ are then set to
-{\tt{nelts}} times the maximum off-diagonal value
-to ensure that the matrix is diagonally dominant.
-The value of the vector element $v_i$ is set to
-the distance of point $i$ from the origin,
-which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.
-
-\subsection{{\tt{gauss}}:
-	Gaussian Elimination
-	\label{s:toys-gauss}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using explicit reduction.
-(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
-there is a well-formed solution to the equation.)
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{sor}}:
-	Successive Over-Relaxation
-	\label{s:toys-sor}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using successive over-relaxation.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\item[{\tt{tolerance}}:]
-	the solution tolerance, e.g.\ $10^{-6}$.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{product}}:
-	Matrix-Vector Product
-	\label{s:toys-product}}
-
-Given a matrix $A$,
-a vector $V$,
-and an assumed solution $X$ to the equation $AX=V$,
-this module calculates the actual product $AX={V^{\prime}}$,
-and then finds the magnitude of the error.
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{actual}}:]
-	the real vector $V$.
-\item[{\tt{candidate}}:]
-	a real vector containing the supposed solution.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-The output of this function is:
-\begin{description}
-\item[{\tt{e}}:]
-	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
-\end{description}
-
 \section{Other Issues\label{s:issues}}
@@
 \begin{figure}
-\epsfxsize=0.75\textwidth
-\begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
+% \epsfxsize=0.75\textwidth
+% \begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
 \caption{Cowichan File Formats\label{f:io-formats}}
@@
 although the values in the resulting matrix then depend on the number of processors used.
-
-\section{Conclusion\label{s:conclude}}
-
-A full ANSI~C version of this problem suite has been written, and is available by anonymous FTP from:
-\begin{center}
-{\tt{ftp.cs.toronto.edu :: pub/gvw/cowichan/src.tar.gz}}
-\end{center}
-This release includes two parallel versions,
-both based on the POSIX threads package.
-The first uses repeated fork-and-join parallelism,
-while the second creates a fixed number of threads at the beginning of the program,
-then synchronizes them repeatedly using barriers.
-
-We would like to invite interested groups to participate in this project by parallelizing the problem suite using their favourite language or tool.
-Our aim is to make both the implementations and their descriptions available as a hypertext document through the World-Wide Web,
-and, if there is sufficient interest, to publish this material in book form.
-If you would like to contribute to this effort, please contact either of the first two authors.
-
-\bibliographystyle{plain}
-\newcommand{\bibea}{{\em{et~al.}}}
-\bibliography{paper}
-
-\appendix
-
-\newpage
-\section{Parallel Clich\'{e}s\label{s:cliche}}
-
-\subsection{Operations\label{s:cliche-ops}}
-
-This list details some operations which are supported by many parallel programming systems.
-The toys described in Section~\ref{s:toys} provide opportunities for exercising many of these,
-and implementors are encouraged to phrase discussion of their work in terms of these operations where appropriate.
-
-\begin{itemize}
-\item	elementwise operations on arrays (unary, binary, and scalar promotion)
-\item	cumulative operations on arrays (reduction and parallel prefix)
-\item	array re-shaping and re-sizing (e.g.\ sub-array extraction)
-\item	partial (conditionally masked) versions of all of the above
-\item	regular data motion (e.g.\ circular and planar shifting)
-\item	irregular data motion (e.g.\ 1-to-1, 1-to-many, and many-to-1 permutation)
-\item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
-\item	differential local addressing (i.e.\ subscripting an array with an array)
-\item	full or partial replication of shared read-only values
-\item	full or partial replication of shared read-mostly values with automatic consistency management
-\item	structures with rotating ownership, suitable for migration
-\item	producer-consumer structures
-\item	partitionable structures
-\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
-\item	dynamic re-partitioning (e.g.\ for load balancing)
-\item	committed mutual exclusion (e.g.\ waiting on a semaphore)
-\item	uncommitted mutual exclusion (e.g.\ lock or fail)
-\item	barrier synchronization
-\item	multiple concurrent barriers used by non-overlapping groups of processes
-\item	fetch-and-add, and other fetch-and-operate functions
-\item	pre-scheduled receipt of messages of known composition
-\item	variable-length message receipt
-\item	message selection by type
-\item	message selection by source
-\item	message selection by contents
-\item	guarded message selection (e.g.\ CSP's {\tt{alt}} construct)
-\item	broadcast and partial broadcast
-\item	split-phase (non-blocking) operations
-\item	data marshalling and unmarshalling of ``flat'' structures (e.g.\ arrays of scalars)
-\item	data marshalling and unmarshalling of nested or linked structures
-\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
-\item	pipelining
-\item	distributed linked structures in systems without physically-shared memory
-\item	indexing of distributed shared arrays
-\item	collective I/O
-\item	uniform-size interleaved I/O
-\item	heterogeneous-size interleaved I/O
-\item	independent I/O operations on a single shared file
-\end{itemize}
-
-\subsection{Memory Reference Patterns\label{s:cliche-memref}}
-
-These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
-Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.
-
-\begin{itemize}
-\item	Write-once:
-	variables which are assigned a value once, and only read thereafter.
-	These can be supported through replication.
-\item	Write-many:
-	variables which are written and read many times.
-	If several processes write to the variable concurrently, they typically write to different portions of it.
-\item	Producer-consumer:
-	variables which are written to by one or more objects, and read by one or more other objects.
-	Entries in a shared queue, or a mailbox used for communication between two processes, are examples.
-\item	Private:
-	variables which are potentially shared, but actually private.
-	The interior points of a mesh in a program which uses geometric decomposition fall into this category,
-	while boundary points belong to the previous class.
-\item	Migratory:
-	variables which are read and written many times in succession by a single process before ownership passes to another process.
-	The structures representing particles in an $N$-body simulation are the clearest example of this category.
-\item	Result:
-	Accumulators which are updated many times by different processes, but thereafter read.
-\item	Read-mostly:
-	Any variable which is read much more often than it is written to.
-	The best score so far in a search problem is a typical instance of this class:
-	while any process might update it, in practice processes read its value much more often than they change it.
-\item	Synchronization:
-	Variables used to force explicit synchronization in a program, such as semaphores and locks.
-	These are typically ignored for long periods, and then the subject of intense bursts of access activity.
-\item	General read-write:
-	Any variable which cannot be put in one of the above categories.
-\end{itemize}
-
-\newpage
-\section{Comparing Parallel Programming Systems\label{s:questions}}
-
-This list of questions is a modified version of the one presented in \cite{b:compare-ada-c-pascal}.
-It is presented here as an organizational aid;
-when describing the PPS they have used,
-implementors are encouraged to use the order and terminology of this list where appropriate.
-We acknowledge that many of these questions will not apply to particular systems,
-and encourage implementors to ask questions which we have left out.
-
-\begin{enumerate}
-\item History and Philosophy
-  \begin{enumerate}
-  \item For whom and for what purpose was the language designed?
-  \end{enumerate}
-\item Syntax
-  \begin{enumerate}
-  \item Model implemented (Figure~\ref{f:taxonomy})
-  \item Class
-    \begin{enumerate}
-    \item library called from existing sequential language?
-    \item syntactic extensions to existing sequential language?
-    \item new language in Backus (Fortran/Algol/C) tradition?
-    \item new (non-Backus) language?
-    \end{enumerate}
-  \item Concurrency mechanisms adhere to same philosophy as original
-	language (where appropriate)?
-  \item Readability?  Syntactic consistency?  Conciseness?
-	(on an APL-to-Modula scale)
-  \item Common errors (e.g.\ confusing {\tt{=}} and {\tt{==}} in C)?
-  \end{enumerate}
-\item Type Philosophy and Data Types
-  \begin{enumerate}
-  \item What primitive types does the language support?
-  \item How are new types created?
-  \item What intrinsic operations are defined on user-defined types?
-  \item How strict is typing?
-	How is type conversion done?
-  \item Can type information be inspected at run-time?
-  \item What concurrent operations on aggregates
-	(such as lists and arrays) are intrinsic?
-	What kinds of concurrent aggregate operations can users define?
-  \item How are aggregates defined (e.g.\ by index set, by shape)?
-	Can new kinds of aggregates (e.g.\ templates for arrays)
-	be defined?
-  \item How strict is typing during communication?
-  \item What support is there for communication or manipulation
-	of data of non-scalar or non-intrinsic types?
-  \item Can structural information about aggregates
-	(e.g.\ number of elements in a set, dimension of an array)
-	be inspected at run-time?
-  \item What built-in support is there for object classes and inheritance?
-	For polymorphism?
-	For class (as opposed to instance) data?
-  \item Are functions first-class?
-  \item Are continuations supported?  Are they first-class?
-  \item Is dynamically-allocated memory supported?
-	Is it done automatically?
-	Is garbage-collection done automatically?
-	Can a program ever create a dangling pointer?
-	An alias?
-  \end{enumerate}
-\item Operators
-  \begin{enumerate}
-  \item What intrinsic operators are provided (brief list)?
-	How strongly do these type-check their arguments?
-  \item How are new operators defined?
-  \item Are the precedence rules straightforward?
-  \item Are side-effects/mutation allowed inside non-call expressions
-	(e.g.\ C's side-effecting \verb`i++` construct)?
-  \item Are their results platform-independent?
-  \item Are the results of intrinsic parallel operations reproducible?
-	Are they platform-independent?
-  \item Are heterogeneous operations on aggregates allowed
-	(e.g.\ every element of a list of polygons calculates its
-	center of mass)?
-	If so, how are these described, and how are they implemented?
-  \item Which of the operators in Section~\ref{s:cliche-ops} are supported?
-  \end{enumerate}
-\item Control Flow
-  \begin{enumerate}
-  \item What synchronization is automatically imposed during execution
-	(i.e.\ at what points can definite assertions about the state
-	of the running program be made)?
-  \item What control constructs does the language support?  Encourage?
-  \item How is concurrency expressed?  How structured is this mechanism?
-  \item Can the degree of concurrency be throttled (limited)?
-	Is this done by the system, or by the user?
-  \item How are exceptions handled/reported?
-	Is it different for intrinsic operations and user-defined operations?
-  \end{enumerate}
-\item Subroutines, Scope, and Modularization
-  \begin{enumerate}
-  \item How efficient is subroutine call compared to in-line execution
-	(i.e.\ how expensive are parameter passing and non-local control flow)?
-  \item How much information about parameters is available (required)
-	inside called subroutine
-	(e.g.\ whether object is shared with other processes or private
-	to executing process)?
-  \item Are side-effecting procedures allowed?
-  \item Can generic routines be written
-	(e.g.\ to work on different types or array shapes)?
-  \item Is there a restriction on return values (e.g.\ can a function return
-	any object which could be declared)?
-  \item Are there levels of scope?
-	If multiple levels are permitted, how are references to stale scopes
-	handled or prevented?
-	Can information for declarations (such as size) be inherited (dynamically)
-	from higher scopes?
-  \item What synchronization requirements (if any) are imposed during calls
-	(i.e.\ do all processes (in a group?) have to enter/exit the same subroutine
-	at the same time)?
-  \item What support is there for modularization?
-	For hiding implementation details?
-	For sharing such information?
-	For detecting and resolving name clashes?
-  \end{enumerate}
-\item Concurrent Programming Facilities
-  \begin{enumerate}
-  \item At what level(s) is the system concurrent?
-    \begin{enumerate}
-    \item implicit (the user shouldn't ask such things)
-    \item operational (e.g.\ data-parallel addition)
-    \item program counter (single executable, multiple states)
-    \item procedural (heterogeneous control parallelism)
-      \begin{enumerate}
-      \item one execution thread per address space
-      \item multiple threads per address space
-      \end{enumerate}
-    \end{enumerate}
-  \item What mutual exclusion and synchronization primitives are provided?
-  \item At what level(s) is data sharing visible to the programmer?
-    \begin{enumerate}
-    \item sequential consistency
-    \item sequential consistency within processor clusters
-    \item object-level consistency
-    \item there is no sharing
-    \end{enumerate}
-  \item What is the conceptual granularity of operations?
-    \begin{enumerate}
-    \item synchronization after atomic instruction
-    \item homogeneous operations with large-grain synchronization
-    \item heterogeneous operations with large-grain synchronization
-    \item completely asynchronous execution
-    \end{enumerate}
-  \item How do processes communicate data values?
-	How do they synchronize?
-	Are non-blocking operations (ones in which buffers might be
-	in a visibly volatile state) allowed?
-	How are they checked for termination?
-  \item If data are communicated explicitly,
-	how does the receiver choose what to accept?
-    \begin{enumerate}
-    \item by address (e.g.\ channel or port)
-    \item by pattern matching on contents
-    \item by explicit message tagging
-    \item by provision of a typed buffer
-    \item by provision of an untyped (byte-block) buffer
-    \item in FIFO order
-    \end{enumerate}
-  \item How do users specify the number of processors on which to run?
-	Must this be built into the program,
-	or can it be decided at run-time?
-	Is there support for executing on heterogeneous processors?
-  \item Can the user specify the mapping of processes to processors?
-	If so, how?
-	How tightly coupled are the mapping of processes to processors,
-	and the mapping of distributed data structures (if supported)?
-  \item Can the user specify the mapping of data structures to processors?
-        Can data structures be decomposed or partitioned?
-  \item What support is there for operations on groups of processes, such as:
-    \begin{enumerate}
-    \item broadcast
-    \item multicast (partial broadcast)
-    \item barrier synchronization
-    \end{enumerate}
-  \item Can arbitrary process groups be created,
-	or must they be structured in some way
-	(e.g.\ as logical mesh)?
-	Can operations involving disjoint groups be executed concurrently
-	(e.g.\ a broadcast within one group,
-	and a barrier within another)?
-  \end{enumerate}
-\item Input/Output
-  \begin{enumerate}
-  \item Is simple sequential (single-process) I/O simple to write?
-  \item Is concurrent I/O supported?  If so, which flavours:
-    \begin{enumerate}
-    \item single-reader-broadcast/collect-single-writer
-    \item homogeneously-sized or -strided segments
-    \item heterogeneously-sized or -strided segments
-    \item arbitrary independent operations
-    \end{enumerate}
-  \item What types can be read/written intrinsically?
-	E.g.\ can an entire record (or array, or list)
-	be read or written in a single operation?
-  \end{enumerate}
-\item Access to Routines in Other Languages and to the Hardware
-  \begin{enumerate}
-  \item Can subroutines written in other languages be called?
-        Can subroutines in this language be called from other languages?
-  \item Are there ways to access the run-time system directly,
-	or similar low-level facilities?
-  \item Is there high-level access to physical addresses and devices?
-  \end{enumerate}
-\item Practical Considerations
-  \begin{enumerate}
-  \item Are both batch and interactive processing supported?
-  \item How reliable and efficient are existing implementations?
-	How fast is it (i.e.\ how quickly does code compile and load)?
-	Is it supported (who fixes it if it breaks)?
-  \item Is there a standard for the system (from some standards organization)?
-  \item Is the documentation adequate? For whom (CS graduates, computer-literate
-	users, the general population)? Is training available?	Consulting?
-  \item How much code has been written (lines of code or weeks of effort)?
-        By whom (computer science undergraduates, professional engineers, etc.)?
-  \item How long does it take the compiler to generate a ``small'' executable
-        (such as {\tt{"hello, world"}})?
-  \item How large is a ``small'' executable?
-        How long does it take to load such an executable?
-  \end{enumerate}
-\item Supporting Tools
-  \begin{enumerate}
-  \item What is the development environment like?
-        Can code be developed on a workstation and then moved to a parallel platform?
-  \item Are there debugging aids, performance analysers, and management tools?
-  \item How much help do these really provide?
-  \end{enumerate}
-\end{enumerate}
-
-\newpage
-\section{A Critique of the C Implementation\label{s:critique}}
-
-This section is a critique of the ANSI~C reference implementation of the problem suite.
-It is intended to serve as a model for (self-)criticism of other programming systems.
-
-\begin{itemize}
-\item	C's support for multi-dimensional arrays (MDAs) is very weak.
-	There is no way to dynamically allocate an MDA in a single go---one must either
-	allocate a block of the same size as the desired array,
-	and then do indexing calculations by hand,
-	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
-	MDAs do not carry dimension information with them,
-	so it is impossible to determine the size of an array parameter within a function.
-	Finally,
-	C does not treat all axes of an array equally:
-	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
-	it is impossible to slice it along the other axis.
-\item	C does not distinguish between Boolean and integer types.
-	As a result,
-	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
-	can be passed in reverse order without a type error.
-	Using {\tt{typedef}} to create a Boolean type name does not solve this
-	(at least, not in {\tt{gcc}} V2.5.8).
-\item	Union types cannot be safely initialized.
-	This complicated the implementation of the graphics interface,
-	where it would have been much more elegant to define a union type,
-	each of whose variants held parameter specifications for a single toy.
-	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
-	initializing some and filling other with don't-care values.
-	This is neither safe nor elegant.
-\item	Parameter values cannot be used in the initialization of local variables inside functions.
-	In particular,
-	it is not possible to create a local vector with a length specified by an input parameter.
-	Such a facility would be useful in the {\tt{winnow}} toy,
-	where we have instead allocated local temporaries of the maximum possible size.
-\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
-	For example, the ``fail()'' error-handling routine was buggy because
-	all \verb`real` arguments were being automatically converted to doubles,
-	but were being taken off the stack as floats.
-\item	no intrinsic notion of group ID/self ID in threads
-\item	packaging parameters for threads
-\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
-\end{itemize}
-
-\end{document}

### 2009-08-28 paper/2009-09/method.tex
Breaking paper into manageable chunks

--- 2009-08-20 paper/2009-09/paper.tex
@@
-\documentstyle[epsf,11pt]{article}
-
-\begin{document}
-
-\begin{Large}\begin{center}
-The Cowichan Problems: \\
-Assessing the Usability of Parallel Programming Systems
-\end{center}\end{Large}
-
-\vspace{\baselineskip}
-
-\begin{center}
-\begin{tabular}{cccc}
-Gregory V.~Wilson\footnote{\verb`http://www.third-bit.com`}
-&
-Andrew Borzenko
-&
-Cameron Gorrie
-&
-Zachary Kincaid
-\end{tabular}
-\end{center}
-
-\vspace{\baselineskip}
-
-\begin{abstract}
-Most programmers believe that parallel programming is harder than sequential programming.
-Many also believe that today's parallel programming systems make it even harder.
-We have developed a suite of chained toy problems that can be used to assess
-the usability of different scientific programming languages, parallel or otherwise.
-This paper motivates and describes the suite,
-and discusses what we have learned from porting it to four parallel C++ systems
-(MPI, OpenMP, Intel Thread Building Blocks, and a Linux tuple space system),
-and from translating it into sequential and parallel Haskell.
-\end{abstract}
-
-\section{Introduction\label{s:intro}}
-
-Twenty years ago,
-Bal {\em{et~al.}\/} listed more than 300 parallel programming systems \cite{b:par-lang-survey}.
-A similar survey today could well uncover twice as many,
-but a poll of application programmers would probably find that
-only a handful are in general use.
-
-The reasons why so many of these PPSs are only used by their developers (if at all) are murkier.
-Part of the blame undoubtedly lies with the diversity of hardware,
-and the attendant lack of portability,
-alluded to above.
-A second reason is that an academic researcher's principal interest is developing new languages,
-not supporting existing ones.
-By the time a system is 90\% complete,
-its author is already thinking about its successor.
-Few application programmers are willing to invest effort porting codes
-to systems which are not properly supported,
-and liable to change or evaporate without notice.
-Third,
-many research groups have chosen to explore the potential of parallelism
-without consideration for compatibility with older languages.
-Application programmers must therefore face the prospect of completely re-writing their programs,
-rather than simply porting them.
-As the next section argues,
-this can obviate any performance that might be gained from parallelization.
-Finally,
-many PPSs simply cannot support the software engineering load of large applications.
-Many provide only a single global namespace,
-do not include I/O capabilities,
-require users to marshal and unmarshal data explicitly,
-or do not permit hardware to be timeshared.
-
-The net result is that many PPSs are never used for anything more complicated than
-generating the Mandelbrot Set,
-simulating the Game of Life,
-multiplying dense matrices,
-or finding all solutions to the $N$-queens problem for some small $N$.
-Lacking the feedback that comes from more demanding use,
-few systems are improvements over their predecessors
-in the way that successive versions of operating systems have improved on one another,
-or the languages in the Algol$\rightarrow$Modula-$n$ family
-have become both more powerful and more secure.
-Most applications programmers therefore choose to play it safe
-and stick to parallel programming's lowest common denominators:
-data-parallel Fortran,
-and message-passing libraries.
-
-The historical reasons for tolerating this situation are becoming less valid with each passing year.
-When only a few parallel computers existed,
-and these were of widely different types,
-software standardization simply didn't make sense.
-Similarly,
-when supercomputers were very expensive,
-the cost of software development seemed relatively smaller,
-and the requirement for maximum performance relatively greater,
-than in sequential computing.
-As \cite{b:social-limits-speed} points out,
-however, these arguments are losing force.
-Now,
-as in sequential computing,
-the first question is not ``How do I write programs for this machine?'',
-but ``Will this machine run my programs?''
-As microprocessor-based MIMD hardware comes to dominate massively-parallel computing,
-we see increasing emphasis on standards such as High-Performance Fortran (HPF) \cite{b:hpf-overview},
-and the Message-Passing Interface (MPI) \cite{b:mpi-overview}.
-
-But herein lies a danger.
-History shows that it is extremely difficult
-to displace a programming language or system once it becomes established.
-Cobol and MS-DOS are the most frequently cited examples,
-but many systems which should have been superceded long ago have managed
-not only to cling to life,
-but to grow in influence.
-(The fact that the mail signature
-``The last good thing written in C was Schubert's Ninth Symphony''
-can elicit smiles simply proves that even the truth can be funny.)
-Established systems evolve slowly, if at all.
-It would therefore seem that the best time to guarantee ourselves
-a good parallel programming paradigm is right now,
-before a large number of large applications are created and take on lives of their own.
-
-We therefore propose to use a set of simple programming problems to compare
-the usability of parallel programming systems.
-Competent programmers, fluent in a specific system,
-will implement solutions to these problems and report their experiences
-in terms of development time, code size and clarity, and runtime efficiency.
-These problems are on the scale of the toy applications described several paragraphs ago
-both to keep implementation time low
-and because we feel that any PPS produced today ought to be able to support such simple programs.
-As the second author put it,
-``We're all sick of seeing these,
-but maybe if we all catch the disease at the same time we'll finally build up some immunity to it.''
-One significant innovation in this work is that
-we also require implementations of our suite to chain problems together,
-so that the output of one problem is the input to another.
-This will allow us to assess how well each PPS supports modularization and code re-use
-(or, more realistically, how much extra effort programmers must invest
-to achieve these goals when using a particular PPS).
-
-Section~\ref{s:corollary} motivates this work by showing
-how poor usability can limit the realizable parallelism in a program.
-Section~\ref{s:previous} surveys previous work on the comparison of parallel programming systems,
-while Section~\ref{s:method} outlines our choice of methods, and the constraints on our work.
-Section~\ref{s:toys} then presents the suite of problems we intend to use.
-In recognition of the original work done by Feo and others on
-the Salishan Problems~\cite{b:salishan}, this suite is called
-the Cowichan Problems\footnote{\noindent Pronounced {\em{Cow}\/}-i-chun.
-	Like Salishan, the word is a Northwest Coast Indian place name.}.
-
-\subsection*{History and Acknowledgments}
-
-The first author originally developed this suite in the mid-1990s,
-inspired both by \cite{b:salishan} and by
-discussions with R.\ Bruce Irvin (then at the University of Wisconsin),
-Henri Bal (Vrije Universiteit, Amsterdam),
-and others.
-Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.
-
-\section{Quantifying the Importance of Usability\label{s:corollary}}
-
-According to Amdahl's Law,
-the time $t$ required to execute a program has two components:
-$t_{seq}$, representing sequential operations that cannot be parallelized,
-and $t_{par}$, representing those operations which can.
-If we let $\sigma$ be the ``serial fraction'' of the program,
-so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
-then the speedup which can be achieved using $P$ processors is:
-\begin{eqnarray*}
-s(P)	& =	& \frac{t(1)}{t(P)}	\\
-	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
-\end{eqnarray*}
-As $P{\rightarrow}{\infty}$,
-the potential speedup is bounded by $1/{\sigma}$.
-
-What is not usually acknowledged is
-the influence of programming system usability on
-the potential speedup of sequential programs.
-As stated above, Amdahl's Law applies only to a single run of a program.
-In the real world, programs are first developed, and then run many times.
-If we let $T$ be the total running time of a program over its lifetime,
-$D_{seq}$ be the time required to develop the sequential components of a program,
-and $D_{par}$ be the time required to develop the parallel components
-(or parallelize the sequential components),
-then the lifetime speedup we can achieve is given by:
-\begin{eqnarray*}
-S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
-	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
-\end{eqnarray*}
-Assume that we are parallelizing a ``dusty deck'',
-i.e.\ that the time $D_{seq}$ has already been invested.
-If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
-then achievable speedup is limited by:
-\begin{eqnarray*}
-S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
-		& =	& \frac{1}{{\phi} + {\sigma}}
-\end{eqnarray*}
-The time required to parallelize a program can therefore be seen
-as increasing that program's effective serial fraction.
-Unless parallelization time can be substantially reduced,
-massively-parallel computing is likely to remain attractive only in those cases in which:
-\begin{itemize}
-\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
-\item	the expected total program runtime is very large
-	(i.e.\ the program is a package used extensively in a field
-	such as computational chemistry or vehicle crash simulation);
-	or
-\item	cost is not an obstacle (i.e.\ the customer is the NSA).
-\end{itemize}
-
-\section{Previous Work\label{s:previous}}
-
-In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
-\begin{quote}
-As practiced by computer science,
-the study of programming is an unholy mixture of mathematics,
-literary criticism,
-and folklore.
-\end{quote}
-Little has changed since.
-Computer scientists pay lip service to the importance of usability,
-but are strangely reluctant to try to quantify it,
-particularly with respect to the tools they use themselves.
-Religious debates about the ``right'' first-year teaching language
-or the ``best'' GUI toolkit are common;
-systematic use of the experimental method is not.
-
-Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
-such as readability and code-pattern recognition
-\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
-The little work that has been done on ``programming in the large'' has been anecdotal,
-such as \cite{b:compare-ada-c-pascal}.
-This book presented various authors' impressions of Pascal, C, and Ada
-(before any complete Ada implementations existed).
-Its editors formulated a list of questions to use in comparing languages,
-but did not use it themselves.
-A modified version of this questionnaire,
-tailored for parallel programming languages,
-is given in Appendix~\ref{s:questions},
-and will be used as a guideline in our work.
-
-Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
-The first presented implementations of a simple numerical quadrature program
-in more than a dozen different parallel languages
-in use on mid-1980s hardware.
-The second presented implementations of
-the Salishan Problems---Hamming number generation,
-isomer enumeration,
-skyline matrix reduction,
-and a simple discrete event simulator---in
-C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
-and a variety of dataflow and functional languages.
-Both of these books convey the flavor of the languages they describe,
-but neither made any effort to compare languages or problem implementations directly.
-
-Inspired by the Salishan Problems,
-the first author developed a suite of seven problems \cite{b:cowichan-1}
-in order to test the limits of parallel programming systems.
-Individual problems were badly load-balanced,
-manipulated irregular pointer-based structures,
-and so on.
-Our experience implementing them was extremely valuable,
-as it uncovered many bugs in the programming system used,
-and revealed several shortcomings of the system's design.
-However,
-we also discovered that implementing each problem takes six to eight weeks,
-which we feel is prohibitive at this exploratory stage.
-
 \section{Methodology\label{s:method}}
@@
 \begin{figure}
-\epsfysize=0.5\textheight
-\begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
+% \epsfysize=0.5\textheight
+% \begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
 \caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
@@
 \begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
+% \epsfxsize=12cm
+% \begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
 \caption{Chaining\label{f:chaining}}
@@
 \begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
+% \epsfxsize=12cm
+% \begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
 \caption{Performance vs.\ Time\label{f:perf}}
@@
 \begin{figure}
-\epsfxsize=6cm
-\begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
+% \epsfxsize=6cm
+% \begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
 \caption{Comparative Code Complexity\label{f:complex}}
@@
 programmers will be allowed to revise and improve their implementations when and as desired.
-
-\section{The Cowichan Problems\label{s:toys}}
-
-This section describes the problems making up the full problem suite.
-As discussed above,
-the implementation of each is to be used both as a stand-alone program
-and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
-Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cl|cl}
-Code	& Toy				& Code	& Toy				\\
-\hline
-E	& elastic net			& O	& outer product			\\
-G	& Gaussian elimination		& P	& matrix-vector product		\\
-H	& halving shuffle		& R	& random matrix generation	\\
-I	& invasion percolation		& S	& successive over-relaxation	\\
-L	& Game of Life			& T	& image thresholding		\\
-M	& Mandelbrot Set		& V	& vector difference		\\
-N	& point normalization		& W	& point value winnowing
-\end{tabular}
-\caption{Single-Character Codes for Toys\label{t:toy-codes}}
-\end{center}
-\end{table}
-
-\subsection{{\tt{mandel}}:
-	Mandelbrot Set Generation
-	\label{s:toys-mandel}}
-
-\begin{figure}
-\begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
-\caption{The Mandelbrot Set\label{f:mandel}}
-\end{figure}
-
-This module generates the Mandelbrot Set
-for a specified region of the complex plane
-(Figure~\ref{f:mandel}).
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{x0, y0}}:]
-	the real coordinates of the lower-left corner
-	of the region to be generated.
-\item[{\tt{dx, dy}}:]
-	the extent of the region to be generated.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing the iteration count at each point in the region.
-\end{description}
-
-Given initial coordinates $(x_0, y_0)$,
-the Mandelbrot Set is generated by iterating the equation
-\begin{eqnarray*}
-x^{\prime}	& =	& x^2 - y^2 + y_0	\\
-y^{\prime}	& =	& 2{x}{y} + x_0
-\end{eqnarray*}
-until either an iteration limit is reached,
-or the values diverge.
-The iteration limit used in this module is 150 steps;
-divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
-The integer value of each element of the matrix is
-the number of iterations done.
-Note that,
-as in all problems,
-the output is required to be
-independent of the number of processors used.
-This requires programmers to be careful about
-floating-point roundoff effects.
-
-\subsection{{\tt{randmat}}:
-	Random Number Generation
-	\label{s:toys-rng}}
-
-This module fills a matrix with pseudo-random integers.
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{s}}:]
-	the random number generation seed.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix filled with random values.
-\end{description}
-Note that,
-as in all problems,
-the output is required to be independent of the number of processors used.
-Generating new seed values from the given seed,
-and running one copy of the random number generator
-on each processor,
-is therefore unlikely to qualify as a solution.
-
-\subsection{{\tt{half}}:
-	Two-Dimensional Shuffle
-	\label{s:toys-half}}
-
-This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
-shuffles them,
-and then repeats this operation along the other axis.
-Values in odd-numbered locations are collected at the low end of each row or column,
-while values in even-numbered locations are moved to the high end.
-An example transformation is:
-
-\begin{eqnarray*}
-\begin{array}{cccc}
-a & b & c & d \\
-e & f & g & h \\
-i & j & k & l
-\end{array}
-& = &
-\begin{array}{cccc}
-a & c & b & d \\
-i & k & j & l \\
-e & g & f & h
-\end{array}
-\end{eqnarray*}
-
-Note that how an array element is moved depends only on whether its location index is odd or even,
-not on whether its value is odd or even.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing shuffled values.
-\end{description}
-
-\subsection{{\tt{invperc}}:
-	Invasion Percolation
-	\label{s:toys-invperc}}
-
-Invasion percolation models the displacement of one fluid
-(such as oil)
-by another
-(such as water)
-in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
-In two dimensions,
-this can be simulated by generating
-an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
-and then marking the center cell of the grid as filled.
-In each iteration,
-one examines the four orthogonal neighbors of all filled cells,
-chooses the one with the lowest value
-(i.e.\ the one with the least resistance to filling),
-and fills it in.
-Figure~\ref{f:invperc1} shows the first few steps in this process,
-while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{math}
-\begin{array}{ccc}
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & 29 & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \star & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & 29 & 77 \\
-\hline
-61 & \star & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \star & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \star & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \star \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & \star & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-\star & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & \star & 72 & 45 & 38 \\
-\hline
-\circ & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \circ \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\end{array}
-\end{math}
-\end{footnotesize}
-\caption{Invasion Percolation\label{f:invperc1}}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
-\end{center}
-\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
-\end{figure}
-
-The simulation continues until
-some fixed percentage of cells have been filled,
-or until some other condition
-(such as the presence of trapped regions)
-is achieved.
-The fractal structure of the filled and unfilled regions
-is then examined to determine how much oil could be recovered.
-The na\"{\i}ve way to implement this is to repeatedly scan the array;
-a more sophisticated, and much faster, sequential technique
-is to maintain a priority queue of unfilled cells
-which are neighbors of filled cells.
-This latter technique is similar to the list-based methods used in some cellular automaton programs,
-and is very difficult to parallelize effectively.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{nfill}}:]
-	the number of points to fill.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
-\end{description}
-Filling begins at the central cell of the matrix
-(rounding down for even-sized axes).
-
-\subsection{{\tt{thresh}}:
-	Histogram Thresholding
-	\label{s:toys-thresh}}
-
-This module performs histogram thresholding on an image.
-Given an integer image $I$ and a target percentage $p$,
-it constructs a binary image $B$
-such that $B_{i,j}$ is set
-if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
-The general idea is that an image's histogram should have 2 peaks,
-one centered around the average foreground intensity,
-and one centered around the average background intensity.
-This program attempts to set
-a threshold between the two peaks in the histogram
-and select the pixels above the threshold.
-This module's inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the integer matrix to be thresholded.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{percent}}:]
-	the minimum percentage of cells to retain.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
-\end{description}
-
-\subsection{{\tt{life}}:
-	Game of Life
-	\label{s:toys-life}}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
-\end{center}
-\caption{Snapshots from the Game of Life\label{f:life}}
-\end{figure}
-
-This module simulates the evolution of Conway's Game of Life,
-a two-dimensional cellular automaton (Figure~\ref{f:life}).
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the Life world.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{numgen}}:]
-	the number of generations to simulate.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the world after simulation.
-\end{description}
-At each time step,
-this module must count the number of live ({\tt{true}}) neighbors
-of each cell,
-using both orthogonal and diagonal connectivity
-and circular boundary conditions.
-The update rule is simple:
-if a cell has 3 live neighbors,
-or has 2 live neighbors and is already alive,
-it is alive in the next generation.
-In any other situation,
-the cell becomes, or stays, dead.
-
-\subsection{{\tt{winnow}}:
-	Weighted Point Selection
-	\label{s:toys-winnow}}
-
-This module converts a matrix of integer values
-to a vector of points,
-represented as $x$ and $y$ coordinates.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix, whose values are used as masses.
-\item[{\tt{mask}}:]
-	a Boolean matrix showing which points are eligible for consideration.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{nelts}}:]
-	the number of points to select.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points.
-\end{description}
-Each location where {\tt{mask}} is {\tt{true}}
-becomes a candidate point,
-with a weight equal to the integer value in {\tt{matrix}} at that location
-and $x$ and $y$ coordinates
-equal to its row and column indices.
-These candidate points are then sorted into increasing order by weight,
-and {\tt{nelts}} evenly-spaced points selected to create the result vector.
-
-\subsection{{\tt{norm}}:
-	Point Location Normalization
-	\label{s:toys-norm}}
-
-This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-Its inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of point locations.
-\item[{\tt{nelts}}:]
-	the number of elements in the point vector.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of normalized point locations.
-\end{description}
-If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
-then the normalization equation is:
-\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
-$y$ coordinates are normalized in the same fashion.
-
-\subsection{{\tt{elastic}}:
-	Elastic Net Simulation
-	\label{s:toys-elastic}}
-
-This module uses the elastic net algorithm \cite{b:elastic-net} to find an approximate solution to the travelling salesentity problem (TSP).
-A closed, circular loop, modelled by a large number of doubly-connected points, is placed in the plane.
-This path is slowly deformed until it connects the points representing the cities to be visited (Figure~\ref{f:elastic}).
-The rate at which the loop is deformed determines the quality of the final path.
-
-\begin{figure}
-\begin{center}
-\framebox{\epsffile{fig/elastic1.ps}}
-\framebox{\epsffile{fig/elastic2.ps}}
-\framebox{\epsffile{fig/elastic3.ps}}
-\end{center}
-\caption{Relaxation of an Elastic Net\label{f:elastic}}
-\end{figure}
-
-Let the (fixed) location of city $i$ be $c_i$, and the (changeable) location of point $j$ on the elastic net be $p_j$.
-The change in the location of point $j$ during each iteration of the algorithm is given by:
-\begin{equation}
-{\delta}{p_j} = {\alpha}\sum_{i}{w_{ij}(c_i - p_j)} ~+~ K{\beta}(p_{j+1} - 2p_j + p_{j-1})
-\end{equation}
-$\alpha$ and $\beta$ are time-invariant factors relating the strengths of the city-to-point and point-to-point forces,
-and $K$ is a time-dependent factor which controls the rate of annealing.
-The strength of the force $w_{ij}$ between city $c_i$ and point $p_j$ is given by:
-\begin{equation}
-w_{ij} = {\phi}({\mid}c_i - p_j{\mid}, K) / \sum_{\ell}\phi({\mid}c_i - p_{\ell}{\mid}, K)
-\end{equation}
-where $\phi(d, K)$ is the Gaussian function $e^{-{d^2}/{K^2}}$.
-Note the normalization of each $w_{ij}$; in order for this algorithm to work properly, the total force being exerted by each city must be the same.
-This is accomplished by summing the forces exerted by each city, and then normalizing the city's forces by that total.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{cities}}:]
-	a vector of points representing the locations of the cities to be toured.
-	The $(x,y)$ coordinates of these cities must be scaled to lie in the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-\item[{\tt{n\_cities}}:]
-	the number of cities in {\tt{cities}}.
-\item[{\tt{n\_net}}:]
-	the number of elastic net points to use.
-\item[{\tt{n\_iters}}:]
-	the total number of iterations to perform.
-\item[{\tt{n\_relax}}:]
-	the number of iterations to perform with each discrete value of $K$.
-	After each {\tt{n\_relax}} iterations, $K$ is reduced by a (constant) pre-defined scaling factor.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{net}}:]
-	a vector of {\tt{n\_net}} points representing the tour found.
-\end{description}
-
-\subsection{{\tt{outer}}:
-	Outer Product
-	\label{s:toys-outer}}
-
-This module turns a vector containing point positions
-into a dense, symmetric, diagonally dominant matrix
-by calculating the distances between each pair of points.
-It also constructs a real vector whose values are
-the distance of each point from the origin.
-Inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
-\item[{\tt{nelts}}:]
-	the number of points in the vector, and the size of the matrix along each axis.
-\end{description}
-Its outputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a real matrix, whose values are filled with inter-point distances.
-\item[{\tt{vector}}:]
-	a real vector, whose values are filled with origin-to-point distances.
-\end{description}
-Each matrix element $M_{i,j}$ such that $i \neq j$
-is given the value $d_{i, j}$,
-the Euclidean distance between point $i$ and point $j$.
-The diagonal values $M_{i, i}$ are then set to
-{\tt{nelts}} times the maximum off-diagonal value
-to ensure that the matrix is diagonally dominant.
-The value of the vector element $v_i$ is set to
-the distance of point $i$ from the origin,
-which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.
-
-\subsection{{\tt{gauss}}:
-	Gaussian Elimination
-	\label{s:toys-gauss}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using explicit reduction.
-(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
-there is a well-formed solution to the equation.)
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{sor}}:
-	Successive Over-Relaxation
-	\label{s:toys-sor}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using successive over-relaxation.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\item[{\tt{tolerance}}:]
-	the solution tolerance, e.g.\ $10^{-6}$.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{product}}:
-	Matrix-Vector Product
-	\label{s:toys-product}}
-
-Given a matrix $A$,
-a vector $V$,
-and an assumed solution $X$ to the equation $AX=V$,
-this module calculates the actual product $AX={V^{\prime}}$,
-and then finds the magnitude of the error.
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{actual}}:]
-	the real vector $V$.
-\item[{\tt{candidate}}:]
-	a real vector containing the supposed solution.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-The output of this function is:
-\begin{description}
-\item[{\tt{e}}:]
-	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
-\end{description}
-
-\section{Other Issues\label{s:issues}}
-
-As described in Section~\ref{s:method-softeng},
-implementors are required to chain modules together in the manner shown in Figure~\ref{f:chaining}.
-This section describes the flow of control and data between these chained problems,
-and discusses several other important aspects of this suite,
-including I/O and visualization.
-
-\subsection{Chained Problem Sequences\label{s:issues-chain}}
-
-The chained implementation of these problems
-executes individual toys in the following order:
-\begin{enumerate}
-\item	An integer matrix $I$ with $r$ rows and $c$ columns is created
-	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
-	\label{e:chain-choice-1}
-\item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
-\item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
-	to generate a Boolean mask $B$ from $I$ in which the (minimum) percentage of {\tt{true}} locations is $P$.
-	Like $I$, $B$ has $r$ rows and $c$ columns.
-	\label{e:chain-choice-2}
-\item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
-	This step overwrites the Boolean matrix $B$.
-\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
-\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
-\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
-\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
-	as described in Section~\ref{s:toys-outer}.
-\item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
-	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
-	This measures the agreement between the solutions found by the two methods.
-\end{enumerate}
-Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
-there are four possible chained sequences.
-A minimal implementation will provide one of these,
-while a full implementation will be a single program which can execute any of the four sequences.
-Intermediate implementations providing all four sequences as separate programs are also acceptable.
-
-\subsection{Input and Output\label{s:issues-io}}
-
-I/O is an important part of programming,
-but is often treated as being of secondary importance by language designers.
-This suite requires all stand-alone toys to read input values from files, and write results back;
-the chained version must be able to checkpoint intermediate results between toys.
-Finally,
-implementors are strongly encouraged to include some means of visualizing the output or evolution of individual toys.
-
-The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
-A file containing a vector begins with 1 positive integer,
-which specifies the number of elements in the vector;
-a file containing a matrix begins with 2 positive integers,
-which specify the number of rows and columns in the matrix respectively.
-Elements of the vector or matrix then appear one per line,
-from least-indexed to greatest-indexed
-(and in row-major order for matrices).
-
-The formats described above are required to be human-readable.
-Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
-using whatever file formats are convenient.
-This will allow programmers to demonstrate the ``natural'' I/O capabilities of particular systems
-which most probably be used for checkpointing intermediate results in real programs.
-
-\begin{figure}
-\epsfxsize=0.75\textwidth
-\begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
-\caption{Cowichan File Formats\label{f:io-formats}}
-\end{figure}
-
-\subsection{Visualization\label{s:issues-vis}}
-
-Visualization is an increasingly important part of scientific and numerical computing.
-Including it as an optional part of this problem set is also a good way to exercise both
-the use of libraries within a parallel programming system,
-and the ease with which a programming system can support variant versions of a program.
-
-Our requirements for the graphics interface are:
-\begin{enumerate}
-\item	Visualization for individual toys should be as simple as possible,
-	and should require as few changes to the basic toy code as possible.
-\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
-	The latter should create all images in a single window.
-\item	A single source file should implement both the graphical and non-graphical versions of each toy.
-\end{enumerate}
-The first two points require the system to distinguish between the actual drawing code
-(e.g.\ drawing a vector of points as circles)
-and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
-While the stand-alone toys use a whole window for drawing,
-the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
-At present,
-this is implemented by using a call counter
-in the interfaces for the individual toys and in the chained module.
-The actual drawing routines called from within each toy could then be the same.
-
-The present visualization routines were originally written by Anil Sukul,
-a student at the Vrije Universiteit, Amsterdam.
-The package uses the public domain graphics library VOGLE,
-available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
-VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
-as well as screen control routines (such as double buffering).
-VOGLE was chosen because of its simplicity,
-and because it includes Pascal and Fortran interfaces.
-
-\subsection{Reproducibility\label{s:issues-reproduce}}
-
-Reproducibility is an important issue for parallel programming systems.
-While constraining the order of operations in a parallel system to guarantee reproducibility makes programs in that system easier to debug,
-it can also reduce the expressiveness or performance of the system.
-
-In this problem set, irreproducibility can appear in two forms: numerical and algorithmic.
-The first arises in toys such as {\tt{gauss}}, {\tt{sor}}, and {\tt{elastic}}, which use floating-point numbers.
-Precisely how round-off errors occur in these calculations can depend on the distribution of work among processors,
-or the order in which those processors carry out particular operations.
-
-Irreproducibility also arises in toys which only use exact numbers, such as {\tt{invperc}} and {\tt{randmat}}.
-In the former, the percolation region is grown by repeatedly filling the lowest-valued cell on its perimeter.
-If several cells have this value, implementations may choose one arbitrarily.
-Thus, different implementations may produce very different fractal shapes.
-In the case of random number generation, the simplest thing to do is to run the same generator independently on each processor,
-although the values in the resulting matrix then depend on the number of processors used.
-
-\section{Conclusion\label{s:conclude}}
-
-A full ANSI~C version of this problem suite has been written, and is available by anonymous FTP from:
-\begin{center}
-{\tt{ftp.cs.toronto.edu :: pub/gvw/cowichan/src.tar.gz}}
-\end{center}
-This release includes two parallel versions,
-both based on the POSIX threads package.
-The first uses repeated fork-and-join parallelism,
-while the second creates a fixed number of threads at the beginning of the program,
-then synchronizes them repeatedly using barriers.
-
-We would like to invite interested groups to participate in this project by parallelizing the problem suite using their favourite language or tool.
-Our aim is to make both the implementations and their descriptions available as a hypertext document through the World-Wide Web,
-and, if there is sufficient interest, to publish this material in book form.
-If you would like to contribute to this effort, please contact either of the first two authors.
-
-\bibliographystyle{plain}
-\newcommand{\bibea}{{\em{et~al.}}}
-\bibliography{paper}
-
-\appendix
-
-\newpage
-\section{Parallel Clich\'{e}s\label{s:cliche}}
-
-\subsection{Operations\label{s:cliche-ops}}
-
-This list details some operations which are supported by many parallel programming systems.
-The toys described in Section~\ref{s:toys} provide opportunities for exercising many of these,
-and implementors are encouraged to phrase discussion of their work in terms of these operations where appropriate.
-
-\begin{itemize}
-\item	elementwise operations on arrays (unary, binary, and scalar promotion)
-\item	cumulative operations on arrays (reduction and parallel prefix)
-\item	array re-shaping and re-sizing (e.g.\ sub-array extraction)
-\item	partial (conditionally masked) versions of all of the above
-\item	regular data motion (e.g.\ circular and planar shifting)
-\item	irregular data motion (e.g.\ 1-to-1, 1-to-many, and many-to-1 permutation)
-\item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
-\item	differential local addressing (i.e.\ subscripting an array with an array)
-\item	full or partial replication of shared read-only values
-\item	full or partial replication of shared read-mostly values with automatic consistency management
-\item	structures with rotating ownership, suitable for migration
-\item	producer-consumer structures
-\item	partitionable structures
-\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
-\item	dynamic re-partitioning (e.g.\ for load balancing)
-\item	committed mutual exclusion (e.g.\ waiting on a semaphore)
-\item	uncommitted mutual exclusion (e.g.\ lock or fail)
-\item	barrier synchronization
-\item	multiple concurrent barriers used by non-overlapping groups of processes
-\item	fetch-and-add, and other fetch-and-operate functions
-\item	pre-scheduled receipt of messages of known composition
-\item	variable-length message receipt
-\item	message selection by type
-\item	message selection by source
-\item	message selection by contents
-\item	guarded message selection (e.g.\ CSP's {\tt{alt}} construct)
-\item	broadcast and partial broadcast
-\item	split-phase (non-blocking) operations
-\item	data marshalling and unmarshalling of ``flat'' structures (e.g.\ arrays of scalars)
-\item	data marshalling and unmarshalling of nested or linked structures
-\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
-\item	pipelining
-\item	distributed linked structures in systems without physically-shared memory
-\item	indexing of distributed shared arrays
-\item	collective I/O
-\item	uniform-size interleaved I/O
-\item	heterogeneous-size interleaved I/O
-\item	independent I/O operations on a single shared file
-\end{itemize}
-
-\subsection{Memory Reference Patterns\label{s:cliche-memref}}
-
-These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
-Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.
-
-\begin{itemize}
-\item	Write-once:
-	variables which are assigned a value once, and only read thereafter.
-	These can be supported through replication.
-\item	Write-many:
-	variables which are written and read many times.
-	If several processes write to the variable concurrently, they typically write to different portions of it.
-\item	Producer-consumer:
-	variables which are written to by one or more objects, and read by one or more other objects.
-	Entries in a shared queue, or a mailbox used for communication between two processes, are examples.
-\item	Private:
-	variables which are potentially shared, but actually private.
-	The interior points of a mesh in a program which uses geometric decomposition fall into this category,
-	while boundary points belong to the previous class.
-\item	Migratory:
-	variables which are read and written many times in succession by a single process before ownership passes to another process.
-	The structures representing particles in an $N$-body simulation are the clearest example of this category.
-\item	Result:
-	Accumulators which are updated many times by different processes, but thereafter read.
-\item	Read-mostly:
-	Any variable which is read much more often than it is written to.
-	The best score so far in a search problem is a typical instance of this class:
-	while any process might update it, in practice processes read its value much more often than they change it.
-\item	Synchronization:
-	Variables used to force explicit synchronization in a program, such as semaphores and locks.
-	These are typically ignored for long periods, and then the subject of intense bursts of access activity.
-\item	General read-write:
-	Any variable which cannot be put in one of the above categories.
-\end{itemize}
-
-\newpage
-\section{Comparing Parallel Programming Systems\label{s:questions}}
-
-This list of questions is a modified version of the one presented in \cite{b:compare-ada-c-pascal}.
-It is presented here as an organizational aid;
-when describing the PPS they have used,
-implementors are encouraged to use the order and terminology of this list where appropriate.
-We acknowledge that many of these questions will not apply to particular systems,
-and encourage implementors to ask questions which we have left out.
-
-\begin{enumerate}
-\item History and Philosophy
-  \begin{enumerate}
-  \item For whom and for what purpose was the language designed?
-  \end{enumerate}
-\item Syntax
-  \begin{enumerate}
-  \item Model implemented (Figure~\ref{f:taxonomy})
-  \item Class
-    \begin{enumerate}
-    \item library called from existing sequential language?
-    \item syntactic extensions to existing sequential language?
-    \item new language in Backus (Fortran/Algol/C) tradition?
-    \item new (non-Backus) language?
-    \end{enumerate}
-  \item Concurrency mechanisms adhere to same philosophy as original
-	language (where appropriate)?
-  \item Readability?  Syntactic consistency?  Conciseness?
-	(on an APL-to-Modula scale)
-  \item Common errors (e.g.\ confusing {\tt{=}} and {\tt{==}} in C)?
-  \end{enumerate}
-\item Type Philosophy and Data Types
-  \begin{enumerate}
-  \item What primitive types does the language support?
-  \item How are new types created?
-  \item What intrinsic operations are defined on user-defined types?
-  \item How strict is typing?
-	How is type conversion done?
-  \item Can type information be inspected at run-time?
-  \item What concurrent operations on aggregates
-	(such as lists and arrays) are intrinsic?
-	What kinds of concurrent aggregate operations can users define?
-  \item How are aggregates defined (e.g.\ by index set, by shape)?
-	Can new kinds of aggregates (e.g.\ templates for arrays)
-	be defined?
-  \item How strict is typing during communication?
-  \item What support is there for communication or manipulation
-	of data of non-scalar or non-intrinsic types?
-  \item Can structural information about aggregates
-	(e.g.\ number of elements in a set, dimension of an array)
-	be inspected at run-time?
-  \item What built-in support is there for object classes and inheritance?
-	For polymorphism?
-	For class (as opposed to instance) data?
-  \item Are functions first-class?
-  \item Are continuations supported?  Are they first-class?
-  \item Is dynamically-allocated memory supported?
-	Is it done automatically?
-	Is garbage-collection done automatically?
-	Can a program ever create a dangling pointer?
-	An alias?
-  \end{enumerate}
-\item Operators
-  \begin{enumerate}
-  \item What intrinsic operators are provided (brief list)?
-	How strongly do these type-check their arguments?
-  \item How are new operators defined?
-  \item Are the precedence rules straightforward?
-  \item Are side-effects/mutation allowed inside non-call expressions
-	(e.g.\ C's side-effecting \verb`i++` construct)?
-  \item Are their results platform-independent?
-  \item Are the results of intrinsic parallel operations reproducible?
-	Are they platform-independent?
-  \item Are heterogeneous operations on aggregates allowed
-	(e.g.\ every element of a list of polygons calculates its
-	center of mass)?
-	If so, how are these described, and how are they implemented?
-  \item Which of the operators in Section~\ref{s:cliche-ops} are supported?
-  \end{enumerate}
-\item Control Flow
-  \begin{enumerate}
-  \item What synchronization is automatically imposed during execution
-	(i.e.\ at what points can definite assertions about the state
-	of the running program be made)?
-  \item What control constructs does the language support?  Encourage?
-  \item How is concurrency expressed?  How structured is this mechanism?
-  \item Can the degree of concurrency be throttled (limited)?
-	Is this done by the system, or by the user?
-  \item How are exceptions handled/reported?
-	Is it different for intrinsic operations and user-defined operations?
-  \end{enumerate}
-\item Subroutines, Scope, and Modularization
-  \begin{enumerate}
-  \item How efficient is subroutine call compared to in-line execution
-	(i.e.\ how expensive are parameter passing and non-local control flow)?
-  \item How much information about parameters is available (required)
-	inside called subroutine
-	(e.g.\ whether object is shared with other processes or private
-	to executing process)?
-  \item Are side-effecting procedures allowed?
-  \item Can generic routines be written
-	(e.g.\ to work on different types or array shapes)?
-  \item Is there a restriction on return values (e.g.\ can a function return
-	any object which could be declared)?
-  \item Are there levels of scope?
-	If multiple levels are permitted, how are references to stale scopes
-	handled or prevented?
-	Can information for declarations (such as size) be inherited (dynamically)
-	from higher scopes?
-  \item What synchronization requirements (if any) are imposed during calls
-	(i.e.\ do all processes (in a group?) have to enter/exit the same subroutine
-	at the same time)?
-  \item What support is there for modularization?
-	For hiding implementation details?
-	For sharing such information?
-	For detecting and resolving name clashes?
-  \end{enumerate}
-\item Concurrent Programming Facilities
-  \begin{enumerate}
-  \item At what level(s) is the system concurrent?
-    \begin{enumerate}
-    \item implicit (the user shouldn't ask such things)
-    \item operational (e.g.\ data-parallel addition)
-    \item program counter (single executable, multiple states)
-    \item procedural (heterogeneous control parallelism)
-      \begin{enumerate}
-      \item one execution thread per address space
-      \item multiple threads per address space
-      \end{enumerate}
-    \end{enumerate}
-  \item What mutual exclusion and synchronization primitives are provided?
-  \item At what level(s) is data sharing visible to the programmer?
-    \begin{enumerate}
-    \item sequential consistency
-    \item sequential consistency within processor clusters
-    \item object-level consistency
-    \item there is no sharing
-    \end{enumerate}
-  \item What is the conceptual granularity of operations?
-    \begin{enumerate}
-    \item synchronization after atomic instruction
-    \item homogeneous operations with large-grain synchronization
-    \item heterogeneous operations with large-grain synchronization
-    \item completely asynchronous execution
-    \end{enumerate}
-  \item How do processes communicate data values?
-	How do they synchronize?
-	Are non-blocking operations (ones in which buffers might be
-	in a visibly volatile state) allowed?
-	How are they checked for termination?
-  \item If data are communicated explicitly,
-	how does the receiver choose what to accept?
-    \begin{enumerate}
-    \item by address (e.g.\ channel or port)
-    \item by pattern matching on contents
-    \item by explicit message tagging
-    \item by provision of a typed buffer
-    \item by provision of an untyped (byte-block) buffer
-    \item in FIFO order
-    \end{enumerate}
-  \item How do users specify the number of processors on which to run?
-	Must this be built into the program,
-	or can it be decided at run-time?
-	Is there support for executing on heterogeneous processors?
-  \item Can the user specify the mapping of processes to processors?
-	If so, how?
-	How tightly coupled are the mapping of processes to processors,
-	and the mapping of distributed data structures (if supported)?
-  \item Can the user specify the mapping of data structures to processors?
-        Can data structures be decomposed or partitioned?
-  \item What support is there for operations on groups of processes, such as:
-    \begin{enumerate}
-    \item broadcast
-    \item multicast (partial broadcast)
-    \item barrier synchronization
-    \end{enumerate}
-  \item Can arbitrary process groups be created,
-	or must they be structured in some way
-	(e.g.\ as logical mesh)?
-	Can operations involving disjoint groups be executed concurrently
-	(e.g.\ a broadcast within one group,
-	and a barrier within another)?
-  \end{enumerate}
-\item Input/Output
-  \begin{enumerate}
-  \item Is simple sequential (single-process) I/O simple to write?
-  \item Is concurrent I/O supported?  If so, which flavours:
-    \begin{enumerate}
-    \item single-reader-broadcast/collect-single-writer
-    \item homogeneously-sized or -strided segments
-    \item heterogeneously-sized or -strided segments
-    \item arbitrary independent operations
-    \end{enumerate}
-  \item What types can be read/written intrinsically?
-	E.g.\ can an entire record (or array, or list)
-	be read or written in a single operation?
-  \end{enumerate}
-\item Access to Routines in Other Languages and to the Hardware
-  \begin{enumerate}
-  \item Can subroutines written in other languages be called?
-        Can subroutines in this language be called from other languages?
-  \item Are there ways to access the run-time system directly,
-	or similar low-level facilities?
-  \item Is there high-level access to physical addresses and devices?
-  \end{enumerate}
-\item Practical Considerations
-  \begin{enumerate}
-  \item Are both batch and interactive processing supported?
-  \item How reliable and efficient are existing implementations?
-	How fast is it (i.e.\ how quickly does code compile and load)?
-	Is it supported (who fixes it if it breaks)?
-  \item Is there a standard for the system (from some standards organization)?
-  \item Is the documentation adequate? For whom (CS graduates, computer-literate
-	users, the general population)? Is training available?	Consulting?
-  \item How much code has been written (lines of code or weeks of effort)?
-        By whom (computer science undergraduates, professional engineers, etc.)?
-  \item How long does it take the compiler to generate a ``small'' executable
-        (such as {\tt{"hello, world"}})?
-  \item How large is a ``small'' executable?
-        How long does it take to load such an executable?
-  \end{enumerate}
-\item Supporting Tools
-  \begin{enumerate}
-  \item What is the development environment like?
-        Can code be developed on a workstation and then moved to a parallel platform?
-  \item Are there debugging aids, performance analysers, and management tools?
-  \item How much help do these really provide?
-  \end{enumerate}
-\end{enumerate}
-
-\newpage
-\section{A Critique of the C Implementation\label{s:critique}}
-
-This section is a critique of the ANSI~C reference implementation of the problem suite.
-It is intended to serve as a model for (self-)criticism of other programming systems.
-
-\begin{itemize}
-\item	C's support for multi-dimensional arrays (MDAs) is very weak.
-	There is no way to dynamically allocate an MDA in a single go---one must either
-	allocate a block of the same size as the desired array,
-	and then do indexing calculations by hand,
-	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
-	MDAs do not carry dimension information with them,
-	so it is impossible to determine the size of an array parameter within a function.
-	Finally,
-	C does not treat all axes of an array equally:
-	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
-	it is impossible to slice it along the other axis.
-\item	C does not distinguish between Boolean and integer types.
-	As a result,
-	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
-	can be passed in reverse order without a type error.
-	Using {\tt{typedef}} to create a Boolean type name does not solve this
-	(at least, not in {\tt{gcc}} V2.5.8).
-\item	Union types cannot be safely initialized.
-	This complicated the implementation of the graphics interface,
-	where it would have been much more elegant to define a union type,
-	each of whose variants held parameter specifications for a single toy.
-	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
-	initializing some and filling other with don't-care values.
-	This is neither safe nor elegant.
-\item	Parameter values cannot be used in the initialization of local variables inside functions.
-	In particular,
-	it is not possible to create a local vector with a length specified by an input parameter.
-	Such a facility would be useful in the {\tt{winnow}} toy,
-	where we have instead allocated local temporaries of the maximum possible size.
-\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
-	For example, the ``fail()'' error-handling routine was buggy because
-	all \verb`real` arguments were being automatically converted to doubles,
-	but were being taken off the stack as floats.
-\item	no intrinsic notion of group ID/self ID in threads
-\item	packaging parameters for threads
-\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
-\end{itemize}
-
-\end{document}

### 2009-08-28 paper/2009-09/paper.tex
Convex hull description
Breaking paper into manageable chunks

@@
 
-\begin{Large}\begin{center}
-The Cowichan Problems: \\
-Assessing the Usability of Parallel Programming Systems
-\end{center}\end{Large}
+\input{titlepage}
 
@@
 
-\begin{center}
-\begin{tabular}{cccc}
-Gregory V.~Wilson\footnote{\verb`http://www.third-bit.com`}
-&
-Andrew Borzenko
-&
-Cameron Gorrie
-&
-Zachary Kincaid
-\end{tabular}
-\end{center}
-
-\vspace{\baselineskip}
-
-\begin{abstract}
-Most programmers believe that parallel programming is harder than sequential programming.
-Many also believe that today's parallel programming systems make it even harder.
-We have developed a suite of chained toy problems that can be used to assess
-the usability of different scientific programming languages, parallel or otherwise.
-This paper motivates and describes the suite,
-and discusses what we have learned from porting it to four parallel C++ systems
-(MPI, OpenMP, Intel Thread Building Blocks, and a Linux tuple space system),
-and from translating it into sequential and parallel Haskell.
-\end{abstract}
-
-\section{Introduction\label{s:intro}}
-
-Twenty years ago,
-Bal {\em{et~al.}\/} listed more than 300 parallel programming systems \cite{b:par-lang-survey}.
-A similar survey today could well uncover twice as many,
-but a poll of application programmers would probably find that
-only a handful are in general use.
-
-The reasons why so many of these PPSs are only used by their developers (if at all) are murkier.
-Part of the blame undoubtedly lies with the diversity of hardware,
-and the attendant lack of portability,
-alluded to above.
-A second reason is that an academic researcher's principal interest is developing new languages,
-not supporting existing ones.
-By the time a system is 90\% complete,
-its author is already thinking about its successor.
-Few application programmers are willing to invest effort porting codes
-to systems which are not properly supported,
-and liable to change or evaporate without notice.
-Third,
-many research groups have chosen to explore the potential of parallelism
-without consideration for compatibility with older languages.
-Application programmers must therefore face the prospect of completely re-writing their programs,
-rather than simply porting them.
-As the next section argues,
-this can obviate any performance that might be gained from parallelization.
-Finally,
-many PPSs simply cannot support the software engineering load of large applications.
-Many provide only a single global namespace,
-do not include I/O capabilities,
-require users to marshal and unmarshal data explicitly,
-or do not permit hardware to be timeshared.
-
-The net result is that many PPSs are never used for anything more complicated than
-generating the Mandelbrot Set,
-simulating the Game of Life,
-multiplying dense matrices,
-or finding all solutions to the $N$-queens problem for some small $N$.
-Lacking the feedback that comes from more demanding use,
-few systems are improvements over their predecessors
-in the way that successive versions of operating systems have improved on one another,
-or the languages in the Algol$\rightarrow$Modula-$n$ family
-have become both more powerful and more secure.
-Most applications programmers therefore choose to play it safe
-and stick to parallel programming's lowest common denominators:
-data-parallel Fortran,
-and message-passing libraries.
-
-The historical reasons for tolerating this situation are becoming less valid with each passing year.
-When only a few parallel computers existed,
-and these were of widely different types,
-software standardization simply didn't make sense.
-Similarly,
-when supercomputers were very expensive,
-the cost of software development seemed relatively smaller,
-and the requirement for maximum performance relatively greater,
-than in sequential computing.
-As \cite{b:social-limits-speed} points out,
-however, these arguments are losing force.
-Now,
-as in sequential computing,
-the first question is not ``How do I write programs for this machine?'',
-but ``Will this machine run my programs?''
-As microprocessor-based MIMD hardware comes to dominate massively-parallel computing,
-we see increasing emphasis on standards such as High-Performance Fortran (HPF) \cite{b:hpf-overview},
-and the Message-Passing Interface (MPI) \cite{b:mpi-overview}.
-
-But herein lies a danger.
-History shows that it is extremely difficult
-to displace a programming language or system once it becomes established.
-Cobol and MS-DOS are the most frequently cited examples,
-but many systems which should have been superceded long ago have managed
-not only to cling to life,
-but to grow in influence.
-(The fact that the mail signature
-``The last good thing written in C was Schubert's Ninth Symphony''
-can elicit smiles simply proves that even the truth can be funny.)
-Established systems evolve slowly, if at all.
-It would therefore seem that the best time to guarantee ourselves
-a good parallel programming paradigm is right now,
-before a large number of large applications are created and take on lives of their own.
-
-We therefore propose to use a set of simple programming problems to compare
-the usability of parallel programming systems.
-Competent programmers, fluent in a specific system,
-will implement solutions to these problems and report their experiences
-in terms of development time, code size and clarity, and runtime efficiency.
-These problems are on the scale of the toy applications described several paragraphs ago
-both to keep implementation time low
-and because we feel that any PPS produced today ought to be able to support such simple programs.
-As the second author put it,
-``We're all sick of seeing these,
-but maybe if we all catch the disease at the same time we'll finally build up some immunity to it.''
-One significant innovation in this work is that
-we also require implementations of our suite to chain problems together,
-so that the output of one problem is the input to another.
-This will allow us to assess how well each PPS supports modularization and code re-use
-(or, more realistically, how much extra effort programmers must invest
-to achieve these goals when using a particular PPS).
-
-Section~\ref{s:corollary} motivates this work by showing
-how poor usability can limit the realizable parallelism in a program.
-Section~\ref{s:previous} surveys previous work on the comparison of parallel programming systems,
-while Section~\ref{s:method} outlines our choice of methods, and the constraints on our work.
-Section~\ref{s:toys} then presents the suite of problems we intend to use.
-In recognition of the original work done by Feo and others on
-the Salishan Problems~\cite{b:salishan}, this suite is called
-the Cowichan Problems\footnote{\noindent Pronounced {\em{Cow}\/}-i-chun.
-	Like Salishan, the word is a Northwest Coast Indian place name.}.
-
-\subsection*{History and Acknowledgments}
-
-The first author originally developed this suite in the mid-1990s,
-inspired both by \cite{b:salishan} and by
-discussions with R.\ Bruce Irvin (then at the University of Wisconsin),
-Henri Bal (Vrije Universiteit, Amsterdam),
-and others.
-Early experiences were described in \cite{b:cowichan-ifip} and \cite{b:cowichan-orca}.
-
-\section{Quantifying the Importance of Usability\label{s:corollary}}
-
-According to Amdahl's Law,
-the time $t$ required to execute a program has two components:
-$t_{seq}$, representing sequential operations that cannot be parallelized,
-and $t_{par}$, representing those operations which can.
-If we let $\sigma$ be the ``serial fraction'' of the program,
-so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
-then the speedup which can be achieved using $P$ processors is:
-\begin{eqnarray*}
-s(P)	& =	& \frac{t(1)}{t(P)}	\\
-	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
-\end{eqnarray*}
-As $P{\rightarrow}{\infty}$,
-the potential speedup is bounded by $1/{\sigma}$.
-
-What is not usually acknowledged is
-the influence of programming system usability on
-the potential speedup of sequential programs.
-As stated above, Amdahl's Law applies only to a single run of a program.
-In the real world, programs are first developed, and then run many times.
-If we let $T$ be the total running time of a program over its lifetime,
-$D_{seq}$ be the time required to develop the sequential components of a program,
-and $D_{par}$ be the time required to develop the parallel components
-(or parallelize the sequential components),
-then the lifetime speedup we can achieve is given by:
-\begin{eqnarray*}
-S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
-	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
-\end{eqnarray*}
-Assume that we are parallelizing a ``dusty deck'',
-i.e.\ that the time $D_{seq}$ has already been invested.
-If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
-then achievable speedup is limited by:
-\begin{eqnarray*}
-S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
-		& =	& \frac{1}{{\phi} + {\sigma}}
-\end{eqnarray*}
-The time required to parallelize a program can therefore be seen
-as increasing that program's effective serial fraction.
-Unless parallelization time can be substantially reduced,
-massively-parallel computing is likely to remain attractive only in those cases in which:
-\begin{itemize}
-\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
-\item	the expected total program runtime is very large
-	(i.e.\ the program is a package used extensively in a field
-	such as computational chemistry or vehicle crash simulation);
-	or
-\item	cost is not an obstacle (i.e.\ the customer is the NSA).
-\end{itemize}
-
-\section{Previous Work\label{s:previous}}
-
-In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
-\begin{quote}
-As practiced by computer science,
-the study of programming is an unholy mixture of mathematics,
-literary criticism,
-and folklore.
-\end{quote}
-Little has changed since.
-Computer scientists pay lip service to the importance of usability,
-but are strangely reluctant to try to quantify it,
-particularly with respect to the tools they use themselves.
-Religious debates about the ``right'' first-year teaching language
-or the ``best'' GUI toolkit are common;
-systematic use of the experimental method is not.
-
-Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
-such as readability and code-pattern recognition
-\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
-The little work that has been done on ``programming in the large'' has been anecdotal,
-such as \cite{b:compare-ada-c-pascal}.
-This book presented various authors' impressions of Pascal, C, and Ada
-(before any complete Ada implementations existed).
-Its editors formulated a list of questions to use in comparing languages,
-but did not use it themselves.
-A modified version of this questionnaire,
-tailored for parallel programming languages,
-is given in Appendix~\ref{s:questions},
-and will be used as a guideline in our work.
-
-Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
-The first presented implementations of a simple numerical quadrature program
-in more than a dozen different parallel languages
-in use on mid-1980s hardware.
-The second presented implementations of
-the Salishan Problems---Hamming number generation,
-isomer enumeration,
-skyline matrix reduction,
-and a simple discrete event simulator---in
-C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
-and a variety of dataflow and functional languages.
-Both of these books convey the flavor of the languages they describe,
-but neither made any effort to compare languages or problem implementations directly.
-
-Inspired by the Salishan Problems,
-the first author developed a suite of seven problems \cite{b:cowichan-1}
-in order to test the limits of parallel programming systems.
-Individual problems were badly load-balanced,
-manipulated irregular pointer-based structures,
-and so on.
-Our experience implementing them was extremely valuable,
-as it uncovered many bugs in the programming system used,
-and revealed several shortcomings of the system's design.
-However,
-we also discovered that implementing each problem takes six to eight weeks,
-which we feel is prohibitive at this exploratory stage.
-
-\section{Methodology\label{s:method}}
-
-The main aim of this work is to assess and compare
-the ease with which programs can be developed and/or ported
-using different parallel programming systems.
-Here,
-we discuss issues related to choosing the problems that make up the suite,
-assessing the complexity of particular implementations,
-and other purposes which implementations of these might serve.
-Recognizing that these problems are extremely simple,
-and easy to parallelize,
-we refer to them as ``toys''.
-
-\subsection{Criteria for Selection\label{s:method-criteria}}
-
-Our criteria for including problems in this suite are as follows:
-\begin{enumerate}
-\item	Each toy should require no more than an afternoon
-	to write and test in a well-supported sequential language.
-	We feel that making individual problems more complicated will only discourage implementation.
-\item	The correctness of each implementation must be relatively easy to verify.
-	Toys whose output is easily visualized are therefore to be preferred,
-	as are toys whose results are insensitive to floating-point discretization effects.
-\item	Each toy application should be amenable to
-	an analysis of speedup or isoefficiency \cite{b:isoefficiency-tutorial}
-	to allow a comparison of theory and practice.
-\item	At least some toys should not be ``infinitely scalable''.
-	Many real-world applications are not,
-	and this suite should reflect such limitations.
-\item	At least some toys should require I/O,
-	since this important aspect of real-world programming is often neglected by PPS designers.
-\item	There should be some potential redundancy in the type of parallelization done,
-	so that re-use of software can be demonstrated.
-\item	Together,
-	the toys in the suite must exercise a wide range of common parallel operations
-	(Appendix~\ref{s:cliche-ops})
-	and memory reference patterns
-	(Appendix~\ref{s:cliche-memref}).
-\item	The ``obvious'' implementations of these toys should span
-	a wide range of parallel programming paradigms,
-	such as task farming,
-	geometric and functional decomposition,
-	and speculative parallelism.
-\item	Parallelizations for several (if not all) of
-	the types of programming systems shown in Figure~\ref{f:taxonomy} should be known.
-	This suite is not intended to be a collection of research problems;
-	reasonably simple and efficient implementations in most environments
-	should be straightforward to produce.
-\end{enumerate}
-
-\begin{figure}
-\epsfysize=0.5\textheight
-\begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
-\caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
-\end{figure}
-
-\subsection{Specification\label{s:method-spec}}
-
-Toys should be specified in terms of inputs and outputs,
-rather than algorithmically.
-For example,
-``sort $N$ integers in the range $1{\ldots}R$,
-where $R$ is not known in advance and there may be multiple instances of a given value''
-is to be preferred over
-``parallelize quicksort''.
-As \cite{b:slalom} argued,
-which parallel algorithm is best in a particular programming system
-depends critically on the architectural assumptions made by that system.
-Implementors should be free to employ those methods which they find natural and convenient.
-The only exception we make to this rule is in the two matrix equation solvers,
-where we require Gaussian elimination and successive over-relaxation.
-Given the enormous number of matrix solution algorithms in use,
-we felt we had to specify these particular algorithms in order to ensure that
-different implementations of this suite would be comparable.
-
-\subsection{Software Engineering Issues\label{s:method-softeng}}
-
-These toys will exercise many different aspects of parallel systems.
-However,
-their ``single algorithm per program'' model is not representative of real applications.
-These usually contain several discrete (and sometimes overlapping) phases,
-each of which is qualitatively different.
-A full implementation of this suite will therefore have two parts.
-In the first,
-each toy will be implemented as a stand-alone program.
-In the second,
-toys will be chained together to create the kind of phase behaviour seen in real applications
-(Figure~\ref{f:chaining}).
-This will test the ease with which heterogeneous parallelism can be mixed within a single program.
-It will also show how well the system supports code re-use and information hiding,
-which are crucial to the development of large programs.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/chain.eps}}\end{center}
-\caption{Chaining\label{f:chaining}}
-\end{figure}
-
-A second reason for chaining toys is that
-many groups are now using algorithmic skeletons to structure parallel programs
-\cite{b:cole-skel,b:p3l-overview,b:enterprise}.
-Such skeletons encapsulate the details of a particular style of parallelism,
-and hide low-level or platform-dependent issues.
-We hope that chaining will show the strengths and weaknesses of such systems.
-
-Finally,
-chaining should be designed so that some toys can be executed concurrently.
-Many PPSs impose extraneous constraints on programs,
-e.g.\ require all processes to participate in every barrier synchronization,
-or require the same executable to be loaded onto each processor.
-These constraints can limit the exploitation of potential concurrency.
-Permitting, but not requiring, concurrent execution of several toys should uncover such limitations.
-
-\subsection{Sizing\label{s:method-size}}
-
-One crucial aspect of the specification of toys is
-the way in which the sizes of problems are determined.
-In a completely frozen model, the actual size of each problem would be compiled into each toy.
-A fluid implementation, by contrast, would allow sizes to be specified at run-time,
-and would only then allocate and partition data structures.
-We have chosen to use an intermediate model (which might be called ``slushy''),
-in which the maximum size of individual problems is specified during compilation,
-but the actual size of a problem is only determined when the toy begins to execute.
-We hope that the use of this model will force implementors to deal with
-at least some of the software engineering issues
-involved in building flexible large-scale applications,
-without making implementation prohibitively difficult.
-
-\subsection{Assessing Complexity\label{s:method-complex}}
-
-There are at least two different ways in which usability might be compared.
-One would be to measure the performance achieved by an ``average'' programmer
-as a function of time on each of a range of problems.
-This would lead to graphs of the kind shown in Figure~\ref{f:perf},
-where the left graph shows a particular result
-and the right graph a simplified abstraction of it.
-By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
-which is the length of vector on which a pipelined architecture achieves
-half its theoretical peak performance,
-we could in principle find the value of $p_{1/2}$---the programming time required to achieve
-half of a machine's peak performance\footnote{Note that
-	if performance was measured as a fraction of the figures quoted by manufacturers for their machines,
-	it is unlikely that the halfway mark would ever be reached.}
-for a particular combination of programming system and problem type.
-
-\begin{figure}
-\epsfxsize=12cm
-\begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
-\caption{Performance vs.\ Time\label{f:perf}}
-\end{figure}
-
-A second option would be to measure the complexities of
-implementations of a single application using different programming systems.
-With this approach,
-comparative performance or speedup figures would be supplemented by complexity measures,
-as shown in Figure~\ref{f:complex}.
-
-\begin{figure}
-\epsfxsize=6cm
-\begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
-\caption{Comparative Code Complexity\label{f:complex}}
-\end{figure}
-
-The drawback of the first approach is that there are no ``typical'' problems,
-or ``average'' programmers.
-Meaningful measurement of  $p_{1/2}$ would therefore require
-measurement of a large number of programmers working on many problems.
-Since this is too expensive to be practical,
-the second approach is the one we have chosen to adopt.
-It has drawbacks as well,
-most notably the fact that no-one has ever demonstrated any better measure of program complexity
-than a simple count of the number of lines in a program \cite{b:fenton-metrics}.
-Since we anticipate a wide variation in the types of progamming systems that will be used in this work
-(ranging from declarative dataflow languages
-through functional languages with Lisp-like syntax
-to parallel dialects of Fortran and C),
-and in the indentation, commenting, and layout styles of particular programmers,
-we may measure complexity by counting the number of tokens in the source code which are
-keywords,
-intrinsic operators,
-user-defined functions,
-type or variable declarations,
-and miscellaneous punctuation.
-
-\subsection{Other Uses for Implementations\label{s:method-uses}}
-
-This suite is intended for assessing PPS usability,
-but we envisage at least three other uses for implementations of it:
-\begin{description}
-\item[Related Research:]
-	We hope that the existence of this suite will indicate
-	to the developers of new parallel programming systems
-	what a mature tool should be able to support.
-	If sufficiently diverse,
-	the suite could also be used to help test the correctness of new or improved systems.
-	Finally,
-	this suite could complement existing performance benchmarks,
-	since performance figures for individual toys could be compared.
-\item[Tool Development:]
-	Requests from tool developers for existing parallel programs,
-	or for event traces from them,
-	are common on newsgroups such as {\tt{comp.parallel}}.
-	We hope that traces from implementations of this suite
-	will be used to compare different debugging and performance monitoring tools,
-	that their source code will be used to test
-	new parallelizing compilation techniques or mapping and scheduling algorithms,
-	and so on.
-\item[Teaching:]
-	Implementations will be suitable for use as classroom examples,
-	since they will be small enough to be understood quickly.
-	The toys themselves should also be suitable as classroom exercises
-	in a senior undergraduate course on parallel computing.
-\end{description}
-
-\subsection{Some Criticisms\label{s:method-criticism}}
-
-\noindent {\em{Why not use an existing benchmark suite for this purpose?}\/}
-
-Implementations of benchmark suites have usually been coded for absolute speed.
-They are not representative of ``normal'' practice,
-in which reasonable performance is the usual goal,
-and so are inappropriate for this work.
-In addition,
-most existing benchmark suites are too large to be implemented or ported quickly
-(e.g.\ SPEC~\cite{b:bench-over}),
-over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
-or concerned with a limited range of applications
-(e.g.\ the NAS Parallel Benchmarks~\cite{b:nas-results}).
-Finally,
-most exist only in C or Fortran (or both);
-many interesting parallel programming systems are built on top of other languages,
-including Modula-2, Scheme, Prolog, and various dataflow languages.
-We do not want the effort required to translate several large programs
-to discourage researchers from participating in this exercise.
-
-\vspace{\baselineskip}
-
-\noindent {\em{Won't the results depend primarily on programmer ability?}\/}
-
-The short answer to this question is yes,
-but no more than the results from performance benchmarks.
-A longer answer is that	since we intend to measure the combination of code complexity and performance achieved,
-rather than performance as a function of development time,
-programmers will be allowed to revise and improve their implementations when and as desired.
-
-\section{The Cowichan Problems\label{s:toys}}
-
-This section describes the problems making up the full problem suite.
-As discussed above,
-the implementation of each is to be used both as a stand-alone program
-and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
-Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cl|cl}
-Code	& Toy				& Code	& Toy				\\
-\hline
-E	& elastic net			& O	& outer product			\\
-G	& Gaussian elimination		& P	& matrix-vector product		\\
-H	& halving shuffle		& R	& random matrix generation	\\
-I	& invasion percolation		& S	& successive over-relaxation	\\
-L	& Game of Life			& T	& image thresholding		\\
-M	& Mandelbrot Set		& V	& vector difference		\\
-N	& point normalization		& W	& point value winnowing
-\end{tabular}
-\caption{Single-Character Codes for Toys\label{t:toy-codes}}
-\end{center}
-\end{table}
-
-\subsection{{\tt{mandel}}:
-	Mandelbrot Set Generation
-	\label{s:toys-mandel}}
-
-\begin{figure}
-\begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
-\caption{The Mandelbrot Set\label{f:mandel}}
-\end{figure}
-
-This module generates the Mandelbrot Set
-for a specified region of the complex plane
-(Figure~\ref{f:mandel}).
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{x0, y0}}:]
-	the real coordinates of the lower-left corner
-	of the region to be generated.
-\item[{\tt{dx, dy}}:]
-	the extent of the region to be generated.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing the iteration count at each point in the region.
-\end{description}
-
-Given initial coordinates $(x_0, y_0)$,
-the Mandelbrot Set is generated by iterating the equation
-\begin{eqnarray*}
-x^{\prime}	& =	& x^2 - y^2 + y_0	\\
-y^{\prime}	& =	& 2{x}{y} + x_0
-\end{eqnarray*}
-until either an iteration limit is reached,
-or the values diverge.
-The iteration limit used in this module is 150 steps;
-divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
-The integer value of each element of the matrix is
-the number of iterations done.
-Note that,
-as in all problems,
-the output is required to be
-independent of the number of processors used.
-This requires programmers to be careful about
-floating-point roundoff effects.
-
-\subsection{{\tt{randmat}}:
-	Random Number Generation
-	\label{s:toys-rng}}
-
-This module fills a matrix with pseudo-random integers.
-The inputs to this module are:
-\begin{description}
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{s}}:]
-	the random number generation seed.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix filled with random values.
-\end{description}
-Note that,
-as in all problems,
-the output is required to be independent of the number of processors used.
-Generating new seed values from the given seed,
-and running one copy of the random number generator
-on each processor,
-is therefore unlikely to qualify as a solution.
-
-\subsection{{\tt{half}}:
-	Two-Dimensional Shuffle
-	\label{s:toys-half}}
-
-This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
-shuffles them,
-and then repeats this operation along the other axis.
-Values in odd-numbered locations are collected at the low end of each row or column,
-while values in even-numbered locations are moved to the high end.
-An example transformation is:
-
-\begin{eqnarray*}
-\begin{array}{cccc}
-a & b & c & d \\
-e & f & g & h \\
-i & j & k & l
-\end{array}
-& = &
-\begin{array}{cccc}
-a & c & b & d \\
-i & k & j & l \\
-e & g & f & h
-\end{array}
-\end{eqnarray*}
-
-Note that how an array element is moved depends only on whether its location index is odd or even,
-not on whether its value is odd or even.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix containing shuffled values.
-\end{description}
-
-\subsection{{\tt{invperc}}:
-	Invasion Percolation
-	\label{s:toys-invperc}}
-
-Invasion percolation models the displacement of one fluid
-(such as oil)
-by another
-(such as water)
-in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
-In two dimensions,
-this can be simulated by generating
-an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
-and then marking the center cell of the grid as filled.
-In each iteration,
-one examines the four orthogonal neighbors of all filled cells,
-chooses the one with the lowest value
-(i.e.\ the one with the least resistance to filling),
-and fills it in.
-Figure~\ref{f:invperc1} shows the first few steps in this process,
-while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{math}
-\begin{array}{ccc}
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & 29 & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \star & \circ & 29 & 77 \\
-\hline
-61 & 26 & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & 29 & 77 \\
-\hline
-61 & \star & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \star & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \star & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & 38 & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \star \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\\
- & & \\
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-10 & \star & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & 35 & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & 12 & 72 & 45 & 38 \\
-\hline
-\star & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & 11 \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-&
-\begin{array}{|c|c|c|c|c|}
-\hline
-26 & \star & 72 & 45 & 38 \\
-\hline
-\circ & \circ & 39 & 92 & 38 \\
-\hline
-44 & \circ & \circ & \circ & 77 \\
-\hline
-61 & \circ & 90 & \circ & \circ \\
-\hline
-83 & 84 & 18 & 56 & 52 \\
-\hline
-\end{array}
-\end{array}
-\end{math}
-\end{footnotesize}
-\caption{Invasion Percolation\label{f:invperc1}}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
-\end{center}
-\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
-\end{figure}
-
-The simulation continues until
-some fixed percentage of cells have been filled,
-or until some other condition
-(such as the presence of trapped regions)
-is achieved.
-The fractal structure of the filled and unfilled regions
-is then examined to determine how much oil could be recovered.
-The na\"{\i}ve way to implement this is to repeatedly scan the array;
-a more sophisticated, and much faster, sequential technique
-is to maintain a priority queue of unfilled cells
-which are neighbors of filled cells.
-This latter technique is similar to the list-based methods used in some cellular automaton programs,
-and is very difficult to parallelize effectively.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{nfill}}:]
-	the number of points to fill.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
-\end{description}
-Filling begins at the central cell of the matrix
-(rounding down for even-sized axes).
-
-\subsection{{\tt{thresh}}:
-	Histogram Thresholding
-	\label{s:toys-thresh}}
-
-This module performs histogram thresholding on an image.
-Given an integer image $I$ and a target percentage $p$,
-it constructs a binary image $B$
-such that $B_{i,j}$ is set
-if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
-The general idea is that an image's histogram should have 2 peaks,
-one centered around the average foreground intensity,
-and one centered around the average background intensity.
-This program attempts to set
-a threshold between the two peaks in the histogram
-and select the pixels above the threshold.
-This module's inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the integer matrix to be thresholded.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix and mask.
-\item[{\tt{percent}}:]
-	the minimum percentage of cells to retain.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{mask}}:]
-	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
-\end{description}
-
-\subsection{{\tt{life}}:
-	Game of Life
-	\label{s:toys-life}}
-
-\begin{figure}
-\begin{center}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
-{\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
-\end{center}
-\caption{Snapshots from the Game of Life\label{f:life}}
-\end{figure}
-
-This module simulates the evolution of Conway's Game of Life,
-a two-dimensional cellular automaton (Figure~\ref{f:life}).
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the Life world.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{numgen}}:]
-	the number of generations to simulate.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a Boolean matrix representing the world after simulation.
-\end{description}
-At each time step,
-this module must count the number of live ({\tt{true}}) neighbors
-of each cell,
-using both orthogonal and diagonal connectivity
-and circular boundary conditions.
-The update rule is simple:
-if a cell has 3 live neighbors,
-or has 2 live neighbors and is already alive,
-it is alive in the next generation.
-In any other situation,
-the cell becomes, or stays, dead.
-
-\subsection{{\tt{winnow}}:
-	Weighted Point Selection
-	\label{s:toys-winnow}}
-
-This module converts a matrix of integer values
-to a vector of points,
-represented as $x$ and $y$ coordinates.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	an integer matrix, whose values are used as masses.
-\item[{\tt{mask}}:]
-	a Boolean matrix showing which points are eligible for consideration.
-\item[{\tt{nrows, ncols}}:]
-	the number of rows and columns in the matrix.
-\item[{\tt{nelts}}:]
-	the number of points to select.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points.
-\end{description}
-Each location where {\tt{mask}} is {\tt{true}}
-becomes a candidate point,
-with a weight equal to the integer value in {\tt{matrix}} at that location
-and $x$ and $y$ coordinates
-equal to its row and column indices.
-These candidate points are then sorted into increasing order by weight,
-and {\tt{nelts}} evenly-spaced points selected to create the result vector.
-
-\subsection{{\tt{norm}}:
-	Point Location Normalization
-	\label{s:toys-norm}}
-
-This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-Its inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of point locations.
-\item[{\tt{nelts}}:]
-	the number of elements in the point vector.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of normalized point locations.
-\end{description}
-If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
-then the normalization equation is:
-\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
-$y$ coordinates are normalized in the same fashion.
-
-\subsection{{\tt{elastic}}:
-	Elastic Net Simulation
-	\label{s:toys-elastic}}
-
-This module uses the elastic net algorithm \cite{b:elastic-net} to find an approximate solution to the travelling salesentity problem (TSP).
-A closed, circular loop, modelled by a large number of doubly-connected points, is placed in the plane.
-This path is slowly deformed until it connects the points representing the cities to be visited (Figure~\ref{f:elastic}).
-The rate at which the loop is deformed determines the quality of the final path.
-
-\begin{figure}
-\begin{center}
-\framebox{\epsffile{fig/elastic1.ps}}
-\framebox{\epsffile{fig/elastic2.ps}}
-\framebox{\epsffile{fig/elastic3.ps}}
-\end{center}
-\caption{Relaxation of an Elastic Net\label{f:elastic}}
-\end{figure}
-
-Let the (fixed) location of city $i$ be $c_i$, and the (changeable) location of point $j$ on the elastic net be $p_j$.
-The change in the location of point $j$ during each iteration of the algorithm is given by:
-\begin{equation}
-{\delta}{p_j} = {\alpha}\sum_{i}{w_{ij}(c_i - p_j)} ~+~ K{\beta}(p_{j+1} - 2p_j + p_{j-1})
-\end{equation}
-$\alpha$ and $\beta$ are time-invariant factors relating the strengths of the city-to-point and point-to-point forces,
-and $K$ is a time-dependent factor which controls the rate of annealing.
-The strength of the force $w_{ij}$ between city $c_i$ and point $p_j$ is given by:
-\begin{equation}
-w_{ij} = {\phi}({\mid}c_i - p_j{\mid}, K) / \sum_{\ell}\phi({\mid}c_i - p_{\ell}{\mid}, K)
-\end{equation}
-where $\phi(d, K)$ is the Gaussian function $e^{-{d^2}/{K^2}}$.
-Note the normalization of each $w_{ij}$; in order for this algorithm to work properly, the total force being exerted by each city must be the same.
-This is accomplished by summing the forces exerted by each city, and then normalizing the city's forces by that total.
-
-The inputs to this module are:
-\begin{description}
-\item[{\tt{cities}}:]
-	a vector of points representing the locations of the cities to be toured.
-	The $(x,y)$ coordinates of these cities must be scaled to lie in the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
-\item[{\tt{n\_cities}}:]
-	the number of cities in {\tt{cities}}.
-\item[{\tt{n\_net}}:]
-	the number of elastic net points to use.
-\item[{\tt{n\_iters}}:]
-	the total number of iterations to perform.
-\item[{\tt{n\_relax}}:]
-	the number of iterations to perform with each discrete value of $K$.
-	After each {\tt{n\_relax}} iterations, $K$ is reduced by a (constant) pre-defined scaling factor.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{net}}:]
-	a vector of {\tt{n\_net}} points representing the tour found.
-\end{description}
-
-\subsection{{\tt{outer}}:
-	Outer Product
-	\label{s:toys-outer}}
-
-This module turns a vector containing point positions
-into a dense, symmetric, diagonally dominant matrix
-by calculating the distances between each pair of points.
-It also constructs a real vector whose values are
-the distance of each point from the origin.
-Inputs are:
-\begin{description}
-\item[{\tt{points}}:]
-	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
-\item[{\tt{nelts}}:]
-	the number of points in the vector, and the size of the matrix along each axis.
-\end{description}
-Its outputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	a real matrix, whose values are filled with inter-point distances.
-\item[{\tt{vector}}:]
-	a real vector, whose values are filled with origin-to-point distances.
-\end{description}
-Each matrix element $M_{i,j}$ such that $i \neq j$
-is given the value $d_{i, j}$,
-the Euclidean distance between point $i$ and point $j$.
-The diagonal values $M_{i, i}$ are then set to
-{\tt{nelts}} times the maximum off-diagonal value
-to ensure that the matrix is diagonally dominant.
-The value of the vector element $v_i$ is set to
-the distance of point $i$ from the origin,
-which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.
-
-\subsection{{\tt{gauss}}:
-	Gaussian Elimination
-	\label{s:toys-gauss}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using explicit reduction.
-(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
-there is a well-formed solution to the equation.)
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{sor}}:
-	Successive Over-Relaxation
-	\label{s:toys-sor}}
-
-This module solves a matrix equation $AX=V$ for
-a dense, symmetric, diagonally dominant matrix $A$
-and an arbitrary vector non-zero $V$
-using successive over-relaxation.
-Its inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{target}}:]
-	the real vector $V$.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\item[{\tt{tolerance}}:]
-	the solution tolerance, e.g.\ $10^{-6}$.
-\end{description}
-Its output is:
-\begin{description}
-\item[{\tt{solution}}:]
-	a real vector containing the solution $X$.
-\end{description}
-
-\subsection{{\tt{product}}:
-	Matrix-Vector Product
-	\label{s:toys-product}}
-
-Given a matrix $A$,
-a vector $V$,
-and an assumed solution $X$ to the equation $AX=V$,
-this module calculates the actual product $AX={V^{\prime}}$,
-and then finds the magnitude of the error.
-Inputs are:
-\begin{description}
-\item[{\tt{matrix}}:]
-	the real matrix $A$.
-\item[{\tt{actual}}:]
-	the real vector $V$.
-\item[{\tt{candidate}}:]
-	a real vector containing the supposed solution.
-\item[{\tt{nelts}}:]
-	the number of values in each vector, and the size of the matrix along each axis.
-\end{description}
-The output of this function is:
-\begin{description}
-\item[{\tt{e}}:]
-	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
-\end{description}
-
-\section{Other Issues\label{s:issues}}
-
-As described in Section~\ref{s:method-softeng},
-implementors are required to chain modules together in the manner shown in Figure~\ref{f:chaining}.
-This section describes the flow of control and data between these chained problems,
-and discusses several other important aspects of this suite,
-including I/O and visualization.
-
-\subsection{Chained Problem Sequences\label{s:issues-chain}}
-
-The chained implementation of these problems
-executes individual toys in the following order:
-\begin{enumerate}
-\item	An integer matrix $I$ with $r$ rows and $c$ columns is created
-	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
-	\label{e:chain-choice-1}
-\item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
-\item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
-	to generate a Boolean mask $B$ from $I$ in which the (minimum) percentage of {\tt{true}} locations is $P$.
-	Like $I$, $B$ has $r$ rows and $c$ columns.
-	\label{e:chain-choice-2}
-\item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
-	This step overwrites the Boolean matrix $B$.
-\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
-\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
-\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
-\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
-	as described in Section~\ref{s:toys-outer}.
-\item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
-	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
-	This measures the agreement between the solutions found by the two methods.
-\end{enumerate}
-Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
-there are four possible chained sequences.
-A minimal implementation will provide one of these,
-while a full implementation will be a single program which can execute any of the four sequences.
-Intermediate implementations providing all four sequences as separate programs are also acceptable.
-
-\subsection{Input and Output\label{s:issues-io}}
-
-I/O is an important part of programming,
-but is often treated as being of secondary importance by language designers.
-This suite requires all stand-alone toys to read input values from files, and write results back;
-the chained version must be able to checkpoint intermediate results between toys.
-Finally,
-implementors are strongly encouraged to include some means of visualizing the output or evolution of individual toys.
-
-The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
-A file containing a vector begins with 1 positive integer,
-which specifies the number of elements in the vector;
-a file containing a matrix begins with 2 positive integers,
-which specify the number of rows and columns in the matrix respectively.
-Elements of the vector or matrix then appear one per line,
-from least-indexed to greatest-indexed
-(and in row-major order for matrices).
-
-The formats described above are required to be human-readable.
-Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
-using whatever file formats are convenient.
-This will allow programmers to demonstrate the ``natural'' I/O capabilities of particular systems
-which most probably be used for checkpointing intermediate results in real programs.
-
-\begin{figure}
-\epsfxsize=0.75\textwidth
-\begin{center}\mbox{\epsffile{fig/file-formats.eps}}\end{center}
-\caption{Cowichan File Formats\label{f:io-formats}}
-\end{figure}
-
-\subsection{Visualization\label{s:issues-vis}}
-
-Visualization is an increasingly important part of scientific and numerical computing.
-Including it as an optional part of this problem set is also a good way to exercise both
-the use of libraries within a parallel programming system,
-and the ease with which a programming system can support variant versions of a program.
-
-Our requirements for the graphics interface are:
-\begin{enumerate}
-\item	Visualization for individual toys should be as simple as possible,
-	and should require as few changes to the basic toy code as possible.
-\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
-	The latter should create all images in a single window.
-\item	A single source file should implement both the graphical and non-graphical versions of each toy.
-\end{enumerate}
-The first two points require the system to distinguish between the actual drawing code
-(e.g.\ drawing a vector of points as circles)
-and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
-While the stand-alone toys use a whole window for drawing,
-the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
-At present,
-this is implemented by using a call counter
-in the interfaces for the individual toys and in the chained module.
-The actual drawing routines called from within each toy could then be the same.
-
-The present visualization routines were originally written by Anil Sukul,
-a student at the Vrije Universiteit, Amsterdam.
-The package uses the public domain graphics library VOGLE,
-available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
-VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
-as well as screen control routines (such as double buffering).
-VOGLE was chosen because of its simplicity,
-and because it includes Pascal and Fortran interfaces.
-
-\subsection{Reproducibility\label{s:issues-reproduce}}
-
-Reproducibility is an important issue for parallel programming systems.
-While constraining the order of operations in a parallel system to guarantee reproducibility makes programs in that system easier to debug,
-it can also reduce the expressiveness or performance of the system.
-
-In this problem set, irreproducibility can appear in two forms: numerical and algorithmic.
-The first arises in toys such as {\tt{gauss}}, {\tt{sor}}, and {\tt{elastic}}, which use floating-point numbers.
-Precisely how round-off errors occur in these calculations can depend on the distribution of work among processors,
-or the order in which those processors carry out particular operations.
-
-Irreproducibility also arises in toys which only use exact numbers, such as {\tt{invperc}} and {\tt{randmat}}.
-In the former, the percolation region is grown by repeatedly filling the lowest-valued cell on its perimeter.
-If several cells have this value, implementations may choose one arbitrarily.
-Thus, different implementations may produce very different fractal shapes.
-In the case of random number generation, the simplest thing to do is to run the same generator independently on each processor,
-although the values in the resulting matrix then depend on the number of processors used.
-
-\section{Conclusion\label{s:conclude}}
-
-A full ANSI~C version of this problem suite has been written, and is available by anonymous FTP from:
-\begin{center}
-{\tt{ftp.cs.toronto.edu :: pub/gvw/cowichan/src.tar.gz}}
-\end{center}
-This release includes two parallel versions,
-both based on the POSIX threads package.
-The first uses repeated fork-and-join parallelism,
-while the second creates a fixed number of threads at the beginning of the program,
-then synchronizes them repeatedly using barriers.
-
-We would like to invite interested groups to participate in this project by parallelizing the problem suite using their favourite language or tool.
-Our aim is to make both the implementations and their descriptions available as a hypertext document through the World-Wide Web,
-and, if there is sufficient interest, to publish this material in book form.
-If you would like to contribute to this effort, please contact either of the first two authors.
+\input{abstract}
+\input{intro}
+\input{quant}
+\input{method}
+\input{problems}
+\input{issues}
+\input{concl}
 
@@
 
-\newpage
-\section{Parallel Clich\'{e}s\label{s:cliche}}
-
-\subsection{Operations\label{s:cliche-ops}}
-
-This list details some operations which are supported by many parallel programming systems.
-The toys described in Section~\ref{s:toys} provide opportunities for exercising many of these,
-and implementors are encouraged to phrase discussion of their work in terms of these operations where appropriate.
-
-\begin{itemize}
-\item	elementwise operations on arrays (unary, binary, and scalar promotion)
-\item	cumulative operations on arrays (reduction and parallel prefix)
-\item	array re-shaping and re-sizing (e.g.\ sub-array extraction)
-\item	partial (conditionally masked) versions of all of the above
-\item	regular data motion (e.g.\ circular and planar shifting)
-\item	irregular data motion (e.g.\ 1-to-1, 1-to-many, and many-to-1 permutation)
-\item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
-\item	differential local addressing (i.e.\ subscripting an array with an array)
-\item	full or partial replication of shared read-only values
-\item	full or partial replication of shared read-mostly values with automatic consistency management
-\item	structures with rotating ownership, suitable for migration
-\item	producer-consumer structures
-\item	partitionable structures
-\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
-\item	dynamic re-partitioning (e.g.\ for load balancing)
-\item	committed mutual exclusion (e.g.\ waiting on a semaphore)
-\item	uncommitted mutual exclusion (e.g.\ lock or fail)
-\item	barrier synchronization
-\item	multiple concurrent barriers used by non-overlapping groups of processes
-\item	fetch-and-add, and other fetch-and-operate functions
-\item	pre-scheduled receipt of messages of known composition
-\item	variable-length message receipt
-\item	message selection by type
-\item	message selection by source
-\item	message selection by contents
-\item	guarded message selection (e.g.\ CSP's {\tt{alt}} construct)
-\item	broadcast and partial broadcast
-\item	split-phase (non-blocking) operations
-\item	data marshalling and unmarshalling of ``flat'' structures (e.g.\ arrays of scalars)
-\item	data marshalling and unmarshalling of nested or linked structures
-\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
-\item	pipelining
-\item	distributed linked structures in systems without physically-shared memory
-\item	indexing of distributed shared arrays
-\item	collective I/O
-\item	uniform-size interleaved I/O
-\item	heterogeneous-size interleaved I/O
-\item	independent I/O operations on a single shared file
-\end{itemize}
-
-\subsection{Memory Reference Patterns\label{s:cliche-memref}}
-
-These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
-Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.
-
-\begin{itemize}
-\item	Write-once:
-	variables which are assigned a value once, and only read thereafter.
-	These can be supported through replication.
-\item	Write-many:
-	variables which are written and read many times.
-	If several processes write to the variable concurrently, they typically write to different portions of it.
-\item	Producer-consumer:
-	variables which are written to by one or more objects, and read by one or more other objects.
-	Entries in a shared queue, or a mailbox used for communication between two processes, are examples.
-\item	Private:
-	variables which are potentially shared, but actually private.
-	The interior points of a mesh in a program which uses geometric decomposition fall into this category,
-	while boundary points belong to the previous class.
-\item	Migratory:
-	variables which are read and written many times in succession by a single process before ownership passes to another process.
-	The structures representing particles in an $N$-body simulation are the clearest example of this category.
-\item	Result:
-	Accumulators which are updated many times by different processes, but thereafter read.
-\item	Read-mostly:
-	Any variable which is read much more often than it is written to.
-	The best score so far in a search problem is a typical instance of this class:
-	while any process might update it, in practice processes read its value much more often than they change it.
-\item	Synchronization:
-	Variables used to force explicit synchronization in a program, such as semaphores and locks.
-	These are typically ignored for long periods, and then the subject of intense bursts of access activity.
-\item	General read-write:
-	Any variable which cannot be put in one of the above categories.
-\end{itemize}
-
-\newpage
-\section{Comparing Parallel Programming Systems\label{s:questions}}
-
-This list of questions is a modified version of the one presented in \cite{b:compare-ada-c-pascal}.
-It is presented here as an organizational aid;
-when describing the PPS they have used,
-implementors are encouraged to use the order and terminology of this list where appropriate.
-We acknowledge that many of these questions will not apply to particular systems,
-and encourage implementors to ask questions which we have left out.
-
-\begin{enumerate}
-\item History and Philosophy
-  \begin{enumerate}
-  \item For whom and for what purpose was the language designed?
-  \end{enumerate}
-\item Syntax
-  \begin{enumerate}
-  \item Model implemented (Figure~\ref{f:taxonomy})
-  \item Class
-    \begin{enumerate}
-    \item library called from existing sequential language?
-    \item syntactic extensions to existing sequential language?
-    \item new language in Backus (Fortran/Algol/C) tradition?
-    \item new (non-Backus) language?
-    \end{enumerate}
-  \item Concurrency mechanisms adhere to same philosophy as original
-	language (where appropriate)?
-  \item Readability?  Syntactic consistency?  Conciseness?
-	(on an APL-to-Modula scale)
-  \item Common errors (e.g.\ confusing {\tt{=}} and {\tt{==}} in C)?
-  \end{enumerate}
-\item Type Philosophy and Data Types
-  \begin{enumerate}
-  \item What primitive types does the language support?
-  \item How are new types created?
-  \item What intrinsic operations are defined on user-defined types?
-  \item How strict is typing?
-	How is type conversion done?
-  \item Can type information be inspected at run-time?
-  \item What concurrent operations on aggregates
-	(such as lists and arrays) are intrinsic?
-	What kinds of concurrent aggregate operations can users define?
-  \item How are aggregates defined (e.g.\ by index set, by shape)?
-	Can new kinds of aggregates (e.g.\ templates for arrays)
-	be defined?
-  \item How strict is typing during communication?
-  \item What support is there for communication or manipulation
-	of data of non-scalar or non-intrinsic types?
-  \item Can structural information about aggregates
-	(e.g.\ number of elements in a set, dimension of an array)
-	be inspected at run-time?
-  \item What built-in support is there for object classes and inheritance?
-	For polymorphism?
-	For class (as opposed to instance) data?
-  \item Are functions first-class?
-  \item Are continuations supported?  Are they first-class?
-  \item Is dynamically-allocated memory supported?
-	Is it done automatically?
-	Is garbage-collection done automatically?
-	Can a program ever create a dangling pointer?
-	An alias?
-  \end{enumerate}
-\item Operators
-  \begin{enumerate}
-  \item What intrinsic operators are provided (brief list)?
-	How strongly do these type-check their arguments?
-  \item How are new operators defined?
-  \item Are the precedence rules straightforward?
-  \item Are side-effects/mutation allowed inside non-call expressions
-	(e.g.\ C's side-effecting \verb`i++` construct)?
-  \item Are their results platform-independent?
-  \item Are the results of intrinsic parallel operations reproducible?
-	Are they platform-independent?
-  \item Are heterogeneous operations on aggregates allowed
-	(e.g.\ every element of a list of polygons calculates its
-	center of mass)?
-	If so, how are these described, and how are they implemented?
-  \item Which of the operators in Section~\ref{s:cliche-ops} are supported?
-  \end{enumerate}
-\item Control Flow
-  \begin{enumerate}
-  \item What synchronization is automatically imposed during execution
-	(i.e.\ at what points can definite assertions about the state
-	of the running program be made)?
-  \item What control constructs does the language support?  Encourage?
-  \item How is concurrency expressed?  How structured is this mechanism?
-  \item Can the degree of concurrency be throttled (limited)?
-	Is this done by the system, or by the user?
-  \item How are exceptions handled/reported?
-	Is it different for intrinsic operations and user-defined operations?
-  \end{enumerate}
-\item Subroutines, Scope, and Modularization
-  \begin{enumerate}
-  \item How efficient is subroutine call compared to in-line execution
-	(i.e.\ how expensive are parameter passing and non-local control flow)?
-  \item How much information about parameters is available (required)
-	inside called subroutine
-	(e.g.\ whether object is shared with other processes or private
-	to executing process)?
-  \item Are side-effecting procedures allowed?
-  \item Can generic routines be written
-	(e.g.\ to work on different types or array shapes)?
-  \item Is there a restriction on return values (e.g.\ can a function return
-	any object which could be declared)?
-  \item Are there levels of scope?
-	If multiple levels are permitted, how are references to stale scopes
-	handled or prevented?
-	Can information for declarations (such as size) be inherited (dynamically)
-	from higher scopes?
-  \item What synchronization requirements (if any) are imposed during calls
-	(i.e.\ do all processes (in a group?) have to enter/exit the same subroutine
-	at the same time)?
-  \item What support is there for modularization?
-	For hiding implementation details?
-	For sharing such information?
-	For detecting and resolving name clashes?
-  \end{enumerate}
-\item Concurrent Programming Facilities
-  \begin{enumerate}
-  \item At what level(s) is the system concurrent?
-    \begin{enumerate}
-    \item implicit (the user shouldn't ask such things)
-    \item operational (e.g.\ data-parallel addition)
-    \item program counter (single executable, multiple states)
-    \item procedural (heterogeneous control parallelism)
-      \begin{enumerate}
-      \item one execution thread per address space
-      \item multiple threads per address space
-      \end{enumerate}
-    \end{enumerate}
-  \item What mutual exclusion and synchronization primitives are provided?
-  \item At what level(s) is data sharing visible to the programmer?
-    \begin{enumerate}
-    \item sequential consistency
-    \item sequential consistency within processor clusters
-    \item object-level consistency
-    \item there is no sharing
-    \end{enumerate}
-  \item What is the conceptual granularity of operations?
-    \begin{enumerate}
-    \item synchronization after atomic instruction
-    \item homogeneous operations with large-grain synchronization
-    \item heterogeneous operations with large-grain synchronization
-    \item completely asynchronous execution
-    \end{enumerate}
-  \item How do processes communicate data values?
-	How do they synchronize?
-	Are non-blocking operations (ones in which buffers might be
-	in a visibly volatile state) allowed?
-	How are they checked for termination?
-  \item If data are communicated explicitly,
-	how does the receiver choose what to accept?
-    \begin{enumerate}
-    \item by address (e.g.\ channel or port)
-    \item by pattern matching on contents
-    \item by explicit message tagging
-    \item by provision of a typed buffer
-    \item by provision of an untyped (byte-block) buffer
-    \item in FIFO order
-    \end{enumerate}
-  \item How do users specify the number of processors on which to run?
-	Must this be built into the program,
-	or can it be decided at run-time?
-	Is there support for executing on heterogeneous processors?
-  \item Can the user specify the mapping of processes to processors?
-	If so, how?
-	How tightly coupled are the mapping of processes to processors,
-	and the mapping of distributed data structures (if supported)?
-  \item Can the user specify the mapping of data structures to processors?
-        Can data structures be decomposed or partitioned?
-  \item What support is there for operations on groups of processes, such as:
-    \begin{enumerate}
-    \item broadcast
-    \item multicast (partial broadcast)
-    \item barrier synchronization
-    \end{enumerate}
-  \item Can arbitrary process groups be created,
-	or must they be structured in some way
-	(e.g.\ as logical mesh)?
-	Can operations involving disjoint groups be executed concurrently
-	(e.g.\ a broadcast within one group,
-	and a barrier within another)?
-  \end{enumerate}
-\item Input/Output
-  \begin{enumerate}
-  \item Is simple sequential (single-process) I/O simple to write?
-  \item Is concurrent I/O supported?  If so, which flavours:
-    \begin{enumerate}
-    \item single-reader-broadcast/collect-single-writer
-    \item homogeneously-sized or -strided segments
-    \item heterogeneously-sized or -strided segments
-    \item arbitrary independent operations
-    \end{enumerate}
-  \item What types can be read/written intrinsically?
-	E.g.\ can an entire record (or array, or list)
-	be read or written in a single operation?
-  \end{enumerate}
-\item Access to Routines in Other Languages and to the Hardware
-  \begin{enumerate}
-  \item Can subroutines written in other languages be called?
-        Can subroutines in this language be called from other languages?
-  \item Are there ways to access the run-time system directly,
-	or similar low-level facilities?
-  \item Is there high-level access to physical addresses and devices?
-  \end{enumerate}
-\item Practical Considerations
-  \begin{enumerate}
-  \item Are both batch and interactive processing supported?
-  \item How reliable and efficient are existing implementations?
-	How fast is it (i.e.\ how quickly does code compile and load)?
-	Is it supported (who fixes it if it breaks)?
-  \item Is there a standard for the system (from some standards organization)?
-  \item Is the documentation adequate? For whom (CS graduates, computer-literate
-	users, the general population)? Is training available?	Consulting?
-  \item How much code has been written (lines of code or weeks of effort)?
-        By whom (computer science undergraduates, professional engineers, etc.)?
-  \item How long does it take the compiler to generate a ``small'' executable
-        (such as {\tt{"hello, world"}})?
-  \item How large is a ``small'' executable?
-        How long does it take to load such an executable?
-  \end{enumerate}
-\item Supporting Tools
-  \begin{enumerate}
-  \item What is the development environment like?
-        Can code be developed on a workstation and then moved to a parallel platform?
-  \item Are there debugging aids, performance analysers, and management tools?
-  \item How much help do these really provide?
-  \end{enumerate}
-\end{enumerate}
-
-\newpage
-\section{A Critique of the C Implementation\label{s:critique}}
-
-This section is a critique of the ANSI~C reference implementation of the problem suite.
-It is intended to serve as a model for (self-)criticism of other programming systems.
-
-\begin{itemize}
-\item	C's support for multi-dimensional arrays (MDAs) is very weak.
-	There is no way to dynamically allocate an MDA in a single go---one must either
-	allocate a block of the same size as the desired array,
-	and then do indexing calculations by hand,
-	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
-	MDAs do not carry dimension information with them,
-	so it is impossible to determine the size of an array parameter within a function.
-	Finally,
-	C does not treat all axes of an array equally:
-	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
-	it is impossible to slice it along the other axis.
-\item	C does not distinguish between Boolean and integer types.
-	As a result,
-	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
-	can be passed in reverse order without a type error.
-	Using {\tt{typedef}} to create a Boolean type name does not solve this
-	(at least, not in {\tt{gcc}} V2.5.8).
-\item	Union types cannot be safely initialized.
-	This complicated the implementation of the graphics interface,
-	where it would have been much more elegant to define a union type,
-	each of whose variants held parameter specifications for a single toy.
-	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
-	initializing some and filling other with don't-care values.
-	This is neither safe nor elegant.
-\item	Parameter values cannot be used in the initialization of local variables inside functions.
-	In particular,
-	it is not possible to create a local vector with a length specified by an input parameter.
-	Such a facility would be useful in the {\tt{winnow}} toy,
-	where we have instead allocated local temporaries of the maximum possible size.
-\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
-	For example, the ``fail()'' error-handling routine was buggy because
-	all \verb`real` arguments were being automatically converted to doubles,
-	but were being taken off the stack as floats.
-\item	no intrinsic notion of group ID/self ID in threads
-\item	packaging parameters for threads
-\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
-\end{itemize}
+\input{cliches}
+\input{compare}
 

### 2009-08-28 paper/2009-09/prob-convex.tex
Breaking paper into manageable chunks

--- 2009-08-02 doc/latex/index.tex
@@
-\hypertarget{index_intro_sec}{}\section{Introduction}\label{index_intro_sec}
-The \hyperlink{class_cowichan}{Cowichan} problems are implemented by various parallel programming systems.
-Serial C++ implementation is available for comparison.
-The \hyperlink{class_cowichan}{Cowichan} class is the base class for all C++ implementations.\hypertarget{index_problems_sec}{}\section{The Problems}\label{index_problems_sec}
-Currently, there are 14 \hyperlink{class_cowichan}{Cowichan} problems.
-They are described below.
-\hyperlink{class_cowichan}{Cowichan} class.
-Problems can be run separately by passing the problem name as an argument on the command line.
-The inputs to the problems are defined in \hyperlink{cowichan__defaults_8hpp}{cowichan\_\-defaults.hpp}.
-There is currently no way to override the inputs without recompiling.\hypertarget{index_mandel_sec}{}\subsection{1.
-Mandelbrot Set Generation}\label{index_mandel_sec}
-This module generates the Mandelbrot Set for a specified region of the complex plane. \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_ec6cc4eb2ad444474b923532167e98a2}{Cowichan::mandel}\end{Desc}
-\hypertarget{index_randmat_sec}{}\subsection{2. Random Number Generation}\label{index_randmat_sec}
-This module fills a matrix with pseudo-random integers. Note that, as in all problems, the output is required to be independent of the number of processors used. \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_c44cacf9d9e363a5b076bcee8b9a7a73}{Cowichan::randmat}\end{Desc}
-\hypertarget{index_half_sec}{}\subsection{3. Two-Dimensional Shuffle}\label{index_half_sec}
-This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis, shuffles them, and then repeats this operation along the other axis.
-Values in odd-numbered locations are collected at the low end of each row or column, while values in even-numbered locations are moved to the high end.
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_308603053675bccbe631f04af921f57c}{Cowichan::half}\end{Desc}
-\hypertarget{index_invperc_sec}{}\subsection{4. Invasion Percolation}\label{index_invperc_sec}
-Invasion percolation models the displacement of one fluid (such as oil) by another (such as water) in fractured rock.
-In two dimensions, this can be simulated by generating an NxN grid of random numbers in the range \mbox{[}1.
-.
-.R\mbox{]}, and then marking the center cell of the grid as filled.
-In each iteration, one examines the four orthogonal neighbors of all filled cells, chooses the one with the lowest value (i.e.
-the one with the least resistance to filling), and fills it in.
-The simulation continues until some fixed percentage of cells have been filled, or until some other condition (such as the presence of trapped regions) is achieved.
-The fractal structure of the filled and unfilled regions is then examined to determine how much oil could be recovered.
-The naive way to implement this is to repeatedly scan the array; a more sophisticated, and much faster, sequential technique is to maintain a priority queue of unfilled cells which are neighbors of filled cells.
-This latter technique is similar to the list-based methods used in some cellular automaton programs, and is very difficult to parallelize effectively.
-Filling begins at the central cell of the matrix (rounding down for even-sized axes).
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_ea126792a31e54a8722663b7ea768955}{Cowichan::invperc}\end{Desc}
-\hypertarget{index_thresh_sec}{}\subsection{5. Histogram Thresholding}\label{index_thresh_sec}
-This module performs histogram thresholding on an image.
-Given an integer image I and a target percentage p, it constructs a binary image B such that B\mbox{[}i,j\mbox{]} is set if no more than p percent of the pixels in I are brighter than I\mbox{[}i,j\mbox{]}.
-The general idea is that an image's histogram should have 2 peaks, one centered around the average foreground intensity, and one centered around the average background intensity.
-This program attempts to set a threshold between the two peaks in the histogram and select the pixels above the threshold.
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_a0b633b8c1f21884e0998a9c7020c08c}{Cowichan::thresh}\end{Desc}
-\hypertarget{index_life_sec}{}\subsection{6. Game of Life}\label{index_life_sec}
-This module simulates the evolution of Conway's Game of Life, a two-dimensional cellular automaton.
-At each time step, this module must count the number of live (true) neighbors of each cell, using both orthogonal and diagonal connectivity and circular boundary conditions.
-The update rule is simple: if a cell has 3 live neighbors, or has 2 live neighbors and is already alive, it is alive in the next generation.
-In any other situation, the cell becomes, or stays, dead.
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_d449595ef2fe934bdd128ac8b1f51d07}{Cowichan::life}\end{Desc}
-\hypertarget{index_winnow_sec}{}\subsection{7. Weighted Point Selection}\label{index_winnow_sec}
-This module converts a matrix of integer values to a vector of points, represented as x and y coordinates.
-Each location where mask is true becomes a candidate point, with a weight equal to the integer value in matrix at that location and x and y coordinates equal to its row and column indices.
-These candidate points are then {\bf sorted} into increasing order by weight, and n evenly-spaced points selected to create the result vector.
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_13d60e06ced3b5da79d62c133ce82337}{Cowichan::winnow}\end{Desc}
-\hypertarget{index_norm_sec}{}\subsection{8. Point Location Normalization}\label{index_norm_sec}
-This module normalizes point coordinates so that all points lie within the unit square \mbox{[}0. . .1\mbox{]}x\mbox{[}0. . .1\mbox{]}. \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_3df21e3c627958114e045c3559a29f30}{Cowichan::norm}\end{Desc}
-\hypertarget{index_hull_sec}{}\subsection{9. Convex Hull}\label{index_hull_sec}
-This module takes a list of two-dimensional points and reorders them by doing multiple convex hull computations.
-Convex hull is the boundary of the minimal convex set containing a given non-empty finite set of points in the plane.
-In other words, all points not in the convex hull are enclosed in the convex hull polygon.
-At each step the convex hull points are taken out of the input list and are put into the output list.
-The computation terminates when there are no more points left in the input list.
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_0c6b68ae3c059b66893405f8530a2e0a}{Cowichan::hull}\end{Desc}
-\hypertarget{index_outer_sec}{}\subsection{10. Outer Product}\label{index_outer_sec}
-This module turns a vector containing point positions into a dense, symmetric, diagonally dominant matrix by calculating the distances between each pair of points.
-It also constructs a real vector whose values are the distance of each point from the origin.
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_52f17221019290b88334b0ca7f3bcdb9}{Cowichan::outer}\end{Desc}
-\hypertarget{index_gauss_sec}{}\subsection{11. Gaussian Elimination}\label{index_gauss_sec}
-This module solves a matrix equation AX = V for a dense, symmetric, diagonally dominant matrix A and an arbitrary vector non-zero V using explicit reduction (matrices are required to be symmetric and diagonally dominant in order to guarantee that there is a well-formed solution to the equation).
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_aa9aac74b96dc5ed33e821d94649d1b2}{Cowichan::gauss}\end{Desc}
-\hypertarget{index_sor_sec}{}\subsection{12. Successive Over-Relaxation}\label{index_sor_sec}
-This module solves a matrix equation AX = V for a dense, symmetric, diagonally dominant matrix A and an arbitrary vector non-zero V using successive over-relaxation. \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_92d8d9ae77208115fdfe69e1174f601c}{Cowichan::sor}\end{Desc}
-\hypertarget{index_product_sec}{}\subsection{13. Matrix-Vector Product}\label{index_product_sec}
-This module calculates V in AX = V where A is a matrix and X is a vector. \begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_3d7d4b581a1d6f0392dc452830fb3b03}{Cowichan::product}\end{Desc}
-\hypertarget{index_vecdiff_sec}{}\subsection{14. 1-Norm Vector Difference}\label{index_vecdiff_sec}
-This module calculates the 1-norm of the difference between two vectors.
-In case the vectors are actual and computed values of some calculation the result represents the magnitude of the error.
-\begin{Desc}
-\item[See also:]\hyperlink{class_cowichan_775d72b5e7d122f9f32555352278250e}{Cowichan::vecdiff} \end{Desc}
+\subsection{{\tt{convex}}:
+	Convex Hull
+	\label{s:toys-convex}}
+
+This module takes a list of two-dimensional points and reorders them by
+doing multiple convex hull computations. The convex hull is the boundary of
+the minimal convex set containing a given non-empty finite set of points
+in the plane. In other words, all points not in the convex hull are
+enclosed in the convex hull polygon. At each step the convex hull points
+are taken out of the input list and are put into the output list. The
+computation terminates when there are no more points left in the input
+list.  Inputs are:
+
+\begin{description}
+\item[{\tt{pointsIn}}:]
+	the vector of input points.
+\end{description}
+Its output is:
+\begin{description}
+\item[{\tt{pointsOut}}:]
+	the vector of output points (a permutation of the input).
+\end{description}

### 2009-08-28 paper/2009-09/prob-gauss.tex
Breaking paper into manageable chunks

\subsection{{\tt{gauss}}:
	Gaussian Elimination
	\label{s:toys-gauss}}

This module solves a matrix equation $AX=V$ for
a dense, symmetric, diagonally dominant matrix $A$
and an arbitrary vector non-zero $V$
using explicit reduction.
(Matrices are required to be symmetric and diagonally dominant in order to guarantee that
there is a well-formed solution to the equation.)
Inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the real matrix $A$.
\item[{\tt{target}}:]
	the real vector $V$.
\item[{\tt{nelts}}:]
	the number of values in each vector, and the size of the matrix along each axis.
\end{description}
Its output is:
\begin{description}
\item[{\tt{solution}}:]
	a real vector containing the solution $X$.
\end{description}

### 2009-08-28 paper/2009-09/prob-half.tex
Breaking paper into manageable chunks

\subsection{{\tt{half}}:
	Two-Dimensional Shuffle
	\label{s:toys-half}}

This module divides the values in a rectangular two-dimensional integer matrix into two halves along one axis,
shuffles them,
and then repeats this operation along the other axis.
Values in odd-numbered locations are collected at the low end of each row or column,
while values in even-numbered locations are moved to the high end.
An example transformation is:

\begin{eqnarray*}
\begin{array}{cccc}
a & b & c & d \\
e & f & g & h \\
i & j & k & l
\end{array}
& = &
\begin{array}{cccc}
a & c & b & d \\
i & k & j & l \\
e & g & f & h
\end{array}
\end{eqnarray*}

Note that how an array element is moved depends only on whether its location index is odd or even,
not on whether its value is odd or even.

The inputs to this module are:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix containing shuffled values.
\end{description}

### 2009-08-28 paper/2009-09/prob-invperc.tex
Breaking paper into manageable chunks

\subsection{{\tt{invperc}}:
	Invasion Percolation
	\label{s:toys-invperc}}

Invasion percolation models the displacement of one fluid
(such as oil)
by another
(such as water)
in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
In two dimensions,
this can be simulated by generating
an $N{\times}N$ grid of random numbers in the range $[1{\ldots}R]$,
and then marking the center cell of the grid as filled.
In each iteration,
one examines the four orthogonal neighbors of all filled cells,
chooses the one with the lowest value
(i.e.\ the one with the least resistance to filling),
and fills it in.
Figure~\ref{f:invperc1} shows the first few steps in this process,
while Figure~\ref{f:invperc2} shows the evolution of the final fractal shape.

\begin{figure}
\begin{footnotesize}
\begin{math}
\begin{array}{ccc}
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & 29 & \circ & 29 & 77 \\
\hline
61 & 26 & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \star & \circ & 29 & 77 \\
\hline
61 & 26 & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & 29 & 77 \\
\hline
61 & \star & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
\\
 & & \\
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \star & 77 \\
\hline
61 & \circ & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & \star & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & 38 & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & \circ & \star \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
\\
 & & \\
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
10 & \star & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & 35 & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & 12 & 72 & 45 & 38 \\
\hline
\star & \circ & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & \circ & 11 \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
&
\begin{array}{|c|c|c|c|c|}
\hline
26 & \star & 72 & 45 & 38 \\
\hline
\circ & \circ & 39 & 92 & 38 \\
\hline
44 & \circ & \circ & \circ & 77 \\
\hline
61 & \circ & 90 & \circ & \circ \\
\hline
83 & 84 & 18 & 56 & 52 \\
\hline
\end{array}
\end{array}
\end{math}
\end{footnotesize}
\caption{Invasion Percolation\label{f:invperc1}}
\end{figure}

\begin{figure}
\begin{center}
% {\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc1.ps}}}
% {\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc2.ps}}}
% {\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc3.ps}}}
% {\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/invperc4.ps}}}
\end{center}
\caption{Fractal Generated by Invasion Percolation\label{f:invperc2}}
\end{figure}

The simulation continues until
some fixed percentage of cells have been filled,
or until some other condition
(such as the presence of trapped regions)
is achieved.
The fractal structure of the filled and unfilled regions
is then examined to determine how much oil could be recovered.
The na\"{\i}ve way to implement this is to repeatedly scan the array;
a more sophisticated, and much faster, sequential technique
is to maintain a priority queue of unfilled cells
which are neighbors of filled cells.
This latter technique is similar to the list-based methods used in some cellular automaton programs,
and is very difficult to parallelize effectively.

The inputs to this module are:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix and mask.
\item[{\tt{nfill}}:]
	the number of points to fill.
\end{description}
Its output is:
\begin{description}
\item[{\tt{mask}}:]
	a Boolean matrix filled with {\tt{true}} (showing a filled cell) or {\tt{false}} (showing an unfilled cell).
\end{description}
Filling begins at the central cell of the matrix
(rounding down for even-sized axes).

### 2009-08-28 paper/2009-09/prob-life.tex
Breaking paper into manageable chunks

\subsection{{\tt{life}}:
	Game of Life
	\label{s:toys-life}}

\begin{figure}
\begin{center}
% {\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life1.ps}}}
% {\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life2.ps}}}
% {\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life3.ps}}}
% {\epsfxsize=0.2\textwidth\framebox{\epsffile{fig/life4.ps}}}
\end{center}
\caption{Snapshots from the Game of Life\label{f:life}}
\end{figure}

This module simulates the evolution of Conway's Game of Life,
a two-dimensional cellular automaton (Figure~\ref{f:life}).
Its inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	a Boolean matrix representing the Life world.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{numgen}}:]
	the number of generations to simulate.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	a Boolean matrix representing the world after simulation.
\end{description}
At each time step,
this module must count the number of live ({\tt{true}}) neighbors
of each cell,
using both orthogonal and diagonal connectivity
and circular boundary conditions.
The update rule is simple:
if a cell has 3 live neighbors,
or has 2 live neighbors and is already alive,
it is alive in the next generation.
In any other situation,
the cell becomes, or stays, dead.

### 2009-08-28 paper/2009-09/prob-mandel.tex
Breaking paper into manageable chunks

\subsection{{\tt{mandel}}:
	Mandelbrot Set Generation
	\label{s:toys-mandel}}

\begin{figure}
% \begin{center}\framebox{\epsffile{fig/mandel.ps}}\end{center}
\caption{The Mandelbrot Set\label{f:mandel}}
\end{figure}

This module generates the Mandelbrot Set
for a specified region of the complex plane
(Figure~\ref{f:mandel}).
The inputs to this module are:
\begin{description}
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{x0, y0}}:]
	the real coordinates of the lower-left corner
	of the region to be generated.
\item[{\tt{dx, dy}}:]
	the extent of the region to be generated.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix containing the iteration count at each point in the region.
\end{description}

Given initial coordinates $(x_0, y_0)$,
the Mandelbrot Set is generated by iterating the equation
\begin{eqnarray*}
x^{\prime}	& =	& x^2 - y^2 + y_0	\\
y^{\prime}	& =	& 2{x}{y} + x_0
\end{eqnarray*}
until either an iteration limit is reached,
or the values diverge.
The iteration limit used in this module is 150 steps;
divergence occurs when $x^2 + y^2$ becomes 2.0 or greater.
The integer value of each element of the matrix is
the number of iterations done.
Note that,
as in all problems,
the output is required to be
independent of the number of processors used.
This requires programmers to be careful about
floating-point roundoff effects.

### 2009-08-28 paper/2009-09/prob-norm.tex
Breaking paper into manageable chunks

\subsection{{\tt{norm}}:
	Point Location Normalization
	\label{s:toys-norm}}

This module normalizes point coordinates so that all points lie within the unit square $[0{\ldots}1]{\times}[0{\ldots}1]$.
Its inputs are:
\begin{description}
\item[{\tt{points}}:]
	a vector of point locations.
\item[{\tt{nelts}}:]
	the number of elements in the point vector.
\end{description}
Its output is:
\begin{description}
\item[{\tt{points}}:]
	a vector of normalized point locations.
\end{description}
If $x_{min}$ and $x_{max}$ are the minimum and maximum $x$ coordinate values in the input vector,
then the normalization equation is:
\[	x_{i}^{\prime} = \frac{x_i - x_{min}}{x_{max} - x_{min}}	\]
$y$ coordinates are normalized in the same fashion.

### 2009-08-28 paper/2009-09/prob-outer.tex
Breaking paper into manageable chunks

\subsection{{\tt{outer}}:
	Outer Product
	\label{s:toys-outer}}

This module turns a vector containing point positions
into a dense, symmetric, diagonally dominant matrix
by calculating the distances between each pair of points.
It also constructs a real vector whose values are
the distance of each point from the origin.
Inputs are:
\begin{description}
\item[{\tt{points}}:]
	a vector of $(x,y)$ points, where $x$ and $y$ are the point's position.
\item[{\tt{nelts}}:]
	the number of points in the vector, and the size of the matrix along each axis.
\end{description}
Its outputs are:
\begin{description}
\item[{\tt{matrix}}:]
	a real matrix, whose values are filled with inter-point distances.
\item[{\tt{vector}}:]
	a real vector, whose values are filled with origin-to-point distances.
\end{description}
Each matrix element $M_{i,j}$ such that $i \neq j$
is given the value $d_{i, j}$,
the Euclidean distance between point $i$ and point $j$.
The diagonal values $M_{i, i}$ are then set to
{\tt{nelts}} times the maximum off-diagonal value
to ensure that the matrix is diagonally dominant.
The value of the vector element $v_i$ is set to
the distance of point $i$ from the origin,
which is given by $\sqrt{{x_i}^2 + {y_i}^2}$.

### 2009-08-28 paper/2009-09/prob-product.tex
Breaking paper into manageable chunks

\subsection{{\tt{product}}:
	Matrix-Vector Product
	\label{s:toys-product}}

Given a matrix $A$,
a vector $V$,
and an assumed solution $X$ to the equation $AX=V$,
this module calculates the actual product $AX={V^{\prime}}$,
and then finds the magnitude of the error.
Inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the real matrix $A$.
\item[{\tt{actual}}:]
	the real vector $V$.
\item[{\tt{candidate}}:]
	a real vector containing the supposed solution.
\item[{\tt{nelts}}:]
	the number of values in each vector, and the size of the matrix along each axis.
\end{description}
The output of this function is:
\begin{description}
\item[{\tt{e}}:]
	the largest absolute value in the element-wise difference of $V$ and $V^{\prime}$.
\end{description}

### 2009-08-28 paper/2009-09/prob-randmat.tex
Breaking paper into manageable chunks

\subsection{{\tt{randmat}}:
	Random Number Generation
	\label{s:toys-rng}}

This module fills a matrix with pseudo-random integers.
The inputs to this module are:
\begin{description}
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{s}}:]
	the random number generation seed.
\end{description}
Its output is:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix filled with random values.
\end{description}
Note that,
as in all problems,
the output is required to be independent of the number of processors used.
Generating new seed values from the given seed,
and running one copy of the random number generator
on each processor,
is therefore unlikely to qualify as a solution.

### 2009-08-28 paper/2009-09/prob-thresh.tex
Breaking paper into manageable chunks

\subsection{{\tt{thresh}}:
	Histogram Thresholding
	\label{s:toys-thresh}}

This module performs histogram thresholding on an image.
Given an integer image $I$ and a target percentage $p$,
it constructs a binary image $B$
such that $B_{i,j}$ is set
if no more than $p$ percent of the pixels in $I$ are brighter than $I_{i,j}$.
The general idea is that an image's histogram should have 2 peaks,
one centered around the average foreground intensity,
and one centered around the average background intensity.
This program attempts to set
a threshold between the two peaks in the histogram
and select the pixels above the threshold.
This module's inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	the integer matrix to be thresholded.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix and mask.
\item[{\tt{percent}}:]
	the minimum percentage of cells to retain.
\end{description}
Its output is:
\begin{description}
\item[{\tt{mask}}:]
	a Boolean matrix filled with {\tt{true}} (showing a cell that is kept) or {\tt{false}} (showing a cell that is discarded).
\end{description}

### 2009-08-28 paper/2009-09/prob-winnow.tex
Breaking paper into manageable chunks

\subsection{{\tt{winnow}}:
	Weighted Point Selection
	\label{s:toys-winnow}}

This module converts a matrix of integer values
to a vector of points,
represented as $x$ and $y$ coordinates.
Its inputs are:
\begin{description}
\item[{\tt{matrix}}:]
	an integer matrix, whose values are used as masses.
\item[{\tt{mask}}:]
	a Boolean matrix showing which points are eligible for consideration.
\item[{\tt{nrows, ncols}}:]
	the number of rows and columns in the matrix.
\item[{\tt{nelts}}:]
	the number of points to select.
\end{description}
Its output is:
\begin{description}
\item[{\tt{points}}:]
	a vector of $(x,y)$ points.
\end{description}
Each location where {\tt{mask}} is {\tt{true}}
becomes a candidate point,
with a weight equal to the integer value in {\tt{matrix}} at that location
and $x$ and $y$ coordinates
equal to its row and column indices.
These candidate points are then sorted into increasing order by weight,
and {\tt{nelts}} evenly-spaced points selected to create the result vector.

### 2009-08-28 paper/2009-09/problems.tex
Breaking paper into manageable chunks

\section{The Cowichan Problems\label{s:toys}}

This section describes the problems making up the full problem suite.
As discussed above,
the implementation of each is to be used both as a stand-alone program
and as a module in the chained problem sequence described in Section~\ref{s:issues-chain}.
Table~\ref{t:toy-codes} lists the one-character codes sometimes used to identify toys.

\begin{table}
\begin{center}
\begin{tabular}{cl|cl}
Code	& Toy				& Code	& Toy				\\
\hline
C	& convex hull			& O	& outer product			\\
G	& Gaussian elimination		& P	& matrix-vector product		\\
H	& halving shuffle		& R	& random matrix generation	\\
I	& invasion percolation		& S	& successive over-relaxation	\\
L	& Game of Life			& T	& image thresholding		\\
M	& Mandelbrot Set		& V	& vector difference		\\
N	& point normalization		& W	& point value winnowing
\end{tabular}
\caption{Single-Character Codes for Toys\label{t:toy-codes}}
\end{center}
\end{table}

\input{prob-convex}
\input{prob-gauss}
\input{prob-half}
\input{prob-invperc}
\input{prob-life}
\input{prob-mandel}
\input{prob-norm}
\input{prob-outer}
\input{prob-product}
\input{prob-randmat}
\input{prob-sor}
\input{prob-thresh}
\input{prob-vecdiff}
\input{prob-winnow}

### 2009-08-28 paper/2009-09/quant.tex
Breaking paper into manageable chunks

\section{Quantifying the Importance of Usability\label{s:corollary}}

According to Amdahl's Law,
the time $t$ required to execute a program has two components:
$t_{seq}$, representing sequential operations that cannot be parallelized,
and $t_{par}$, representing those operations which can.
If we let $\sigma$ be the ``serial fraction'' of the program,
so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
then the speedup which can be achieved using $P$ processors is:
\begin{eqnarray*}
s(P)	& =	& \frac{t(1)}{t(P)}	\\
	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
\end{eqnarray*}
As $P{\rightarrow}{\infty}$,
the potential speedup is bounded by $1/{\sigma}$.

What is not usually acknowledged is
the influence of programming system usability on
the potential speedup of sequential programs.
As stated above, Amdahl's Law applies only to a single run of a program.
In the real world, programs are first developed, and then run many times.
If we let $T$ be the total running time of a program over its lifetime,
$D_{seq}$ be the time required to develop the sequential components of a program,
and $D_{par}$ be the time required to develop the parallel components
(or parallelize the sequential components),
then the lifetime speedup we can achieve is given by:
\begin{eqnarray*}
S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
\end{eqnarray*}
Assume that we are parallelizing a ``dusty deck'',
i.e.\ that the time $D_{seq}$ has already been invested.
If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
then achievable speedup is limited by:
\begin{eqnarray*}
S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
		& =	& \frac{1}{{\phi} + {\sigma}}
\end{eqnarray*}
The time required to parallelize a program can therefore be seen
as increasing that program's effective serial fraction.
Unless parallelization time can be substantially reduced,
massively-parallel computing is likely to remain attractive only in those cases in which:
\begin{itemize}
\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
\item	the expected total program runtime is very large
	(i.e.\ the program is a package used extensively in a field
	such as computational chemistry or vehicle crash simulation);
	or
\item	cost is not an obstacle (i.e.\ the customer is the NSA).
\end{itemize}

\subsection{Previous Work\label{s:previous}}

In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
\begin{quote}
As practiced by computer science,
the study of programming is an unholy mixture of mathematics,
literary criticism,
and folklore.
\end{quote}
Little has changed since.
Computer scientists pay lip service to the importance of usability,
but are strangely reluctant to try to quantify it,
particularly with respect to the tools they use themselves.
Religious debates about the ``right'' first-year teaching language
or the ``best'' GUI toolkit are common;
systematic use of the experimental method is not.

Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
such as readability and code-pattern recognition
\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
The little work that has been done on ``programming in the large'' has been anecdotal,
such as \cite{b:compare-ada-c-pascal}.
This book presented various authors' impressions of Pascal, C, and Ada
(before any complete Ada implementations existed).
Its editors formulated a list of questions to use in comparing languages,
but did not use it themselves.
A modified version of this questionnaire,
tailored for parallel programming languages,
is given in Appendix~\ref{s:questions},
and will be used as a guideline in our work.

Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
The first presented implementations of a simple numerical quadrature program
in more than a dozen different parallel languages
in use on mid-1980s hardware.
The second presented implementations of
the Salishan Problems---Hamming number generation,
isomer enumeration,
skyline matrix reduction,
and a simple discrete event simulator---in
C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
and a variety of dataflow and functional languages.
Both of these books convey the flavor of the languages they describe,
but neither made any effort to compare languages or problem implementations directly.

Inspired by the Salishan Problems,
the first author developed a suite of seven problems \cite{b:cowichan-1}
in order to test the limits of parallel programming systems.
Individual problems were badly load-balanced,
manipulated irregular pointer-based structures,
and so on.
Our experience implementing them was extremely valuable,
as it uncovered many bugs in the programming system used,
and revealed several shortcomings of the system's design.
However,
we also discovered that implementing each problem takes six to eight weeks,
which we feel is prohibitive at this exploratory stage.

### 2009-08-28 paper/2009-09/unused.tex
Breaking paper into manageable chunks

\section{A Critique of the C Implementation\label{s:critique}}

This section is a critique of the ANSI~C reference implementation of the problem suite.
It is intended to serve as a model for (self-)criticism of other programming systems.

\begin{itemize}
\item	C's support for multi-dimensional arrays (MDAs) is very weak.
	There is no way to dynamically allocate an MDA in a single go---one must either
	allocate a block of the same size as the desired array,
	and then do indexing calculations by hand,
	or allocate a vector of pointers to allocated vectors of pointers to{\ldots}to vectors holding data.
	MDAs do not carry dimension information with them,
	so it is impossible to determine the size of an array parameter within a function.
	Finally,
	C does not treat all axes of an array equally:
	while it is trivial to take a slice out of a 2-dimensional array along the most-significant axis,
	it is impossible to slice it along the other axis.
\item	C does not distinguish between Boolean and integer types.
	As a result,
	the {\tt{matrix}} and {\tt{mask}} arguments to the invasion percolation problem
	can be passed in reverse order without a type error.
	Using {\tt{typedef}} to create a Boolean type name does not solve this
	(at least, not in {\tt{gcc}} V2.5.8).
\item	Union types cannot be safely initialized.
	This complicated the implementation of the graphics interface,
	where it would have been much more elegant to define a union type,
	each of whose variants held parameter specifications for a single toy.
	The code in the graphics module {\tt{gfx.c}} relies instead on arrays of {\tt{int}}s and {\tt{float}}s,
	initializing some and filling other with don't-care values.
	This is neither safe nor elegant.
\item	Parameter values cannot be used in the initialization of local variables inside functions.
	In particular,
	it is not possible to create a local vector with a length specified by an input parameter.
	Such a facility would be useful in the {\tt{winnow}} toy,
	where we have instead allocated local temporaries of the maximum possible size.
\item	The automatic conversion of floats to doubles across function calls can quite often be a nuisance.
	For example, the ``fail()'' error-handling routine was buggy because
	all \verb`real` arguments were being automatically converted to doubles,
	but were being taken off the stack as floats.
\item	no intrinsic notion of group ID/self ID in threads
\item	packaging parameters for threads
\item	pointers to array vs.\ arrays themselves (semantics of definition depends upon context)
\end{itemize}

### 2009-08-28 paper/2009-09/compare.tex
Firming up quantification section

@@
 It is presented here as an organizational aid;
-when describing the PPS they have used,
+when describing the programming system they have used,
 implementors are encouraged to use the order and terminology of this list where appropriate.

### 2009-08-28 paper/2009-09/quant.tex
Rewriting the introduction
Firming up quantification section

@@
-\section{Quantifying the Importance of Usability\label{s:corollary}}
+\section{Quantifying the Importance of Usability\label{s:quant}}
 
 According to Amdahl's Law,
-the time $t$ required to execute a program has two components:
-$t_{seq}$, representing sequential operations that cannot be parallelized,
-and $t_{par}$, representing those operations which can.
-If we let $\sigma$ be the ``serial fraction'' of the program,
-so that $t_{seq} = {\sigma}t$ and $t_{par} = (1-{\sigma})t$,
-then the speedup which can be achieved using $P$ processors is:
+if $t(n)$ is an algorithm's running time on $n$ processors
+and $\sigma$ is the algorithm's serial fraction
+(i.e., the fraction of the algorithm that cannot be parallelized),
+then:
 \begin{eqnarray*}
-s(P)	& =	& \frac{t(1)}{t(P)}	\\
-	& =	& \frac{t}{{{\sigma}t} + {{(1-{\sigma})t}}/{P}}
+t(p) & = & {\sigma}t(1) + (1 - {\sigma})t(1)/p
 \end{eqnarray*}
-As $P{\rightarrow}{\infty}$,
-the potential speedup is bounded by $1/{\sigma}$.
+The speedup on $p$ processors is therefore:
+\begin{eqnarray*}
+s(p) & = & t(1)/t(p) \\
+     & = & t(1)/({\sigma}t(1) + (1 - {\sigma})t(1)/p) \\
+     & = & 1/({\sigma} + (1 - {\sigma})/p)
+\end{eqnarray*}
+As $p{\leftarrow}{\infty}$, the speedup is bounded by $1/{\sigma}$.
+This makes intuitive sense:
+if 10\% of an algorithm can't be parallelized,
+then even with infinite hardware,
+the program can only be made to run 10 times faster.
 
-What is not usually acknowledged is
-the influence of programming system usability on
-the potential speedup of sequential programs.
-As stated above, Amdahl's Law applies only to a single run of a program.
-In the real world, programs are first developed, and then run many times.
-If we let $T$ be the total running time of a program over its lifetime,
-$D_{seq}$ be the time required to develop the sequential components of a program,
-and $D_{par}$ be the time required to develop the parallel components
-(or parallelize the sequential components),
-then the lifetime speedup we can achieve is given by:
+But software doesn't just happen:
+it has to be written and debugged,
+and these activities usually aren't made faster by parallel hardware\footnote{In fact,
+most programmers believe that doing parallel programming is harder and slower than doing it sequentially.}.
+If we let $T$ and $S$ be the equivalents of $t$ and $s$ over the program's whole lifetime,
+and $D$ be its total development time,
+then:
 \begin{eqnarray*}
-S(P)	& =	& \frac{D_{seq} + T(1)}{D_{seq} + D_{par} + T(P)}	\\
-	& =	& \frac{D_{seq} + T}{D_{seq} + D_{par} + {{\sigma}T} + {{(1-{\sigma})T}}/{P}}
+T(p) & = & {\sigma}T(1) + (1 - {\sigma})t(1)/p + D
 \end{eqnarray*}
-Assume that we are parallelizing a ``dusty deck'',
-i.e.\ that the time $D_{seq}$ has already been invested.
-If we let $\phi = D_{par}/T$ be the ratio of parallelization time to total program runtime,
-then achievable speedup is limited by:
+The achievable \emph{lifetime} speedup is then:
 \begin{eqnarray*}
-S_{legacy}(P)	& =	& \frac{T}{D_{par} + {\sigma}T}		\\
-		& =	& \frac{1}{{\phi} + {\sigma}}
+S(\infty) & = & 1/({\sigma} + D/T(1))
 \end{eqnarray*}
-The time required to parallelize a program can therefore be seen
-as increasing that program's effective serial fraction.
+
+The ratio of development time to (sequential) running time is therefore as important to speedup
+as the serial fraction of the algorithm.
+Put another way,
+the time required to parallelize a program effectively increases the program's effective serial fraction.
 Unless parallelization time can be substantially reduced,
-massively-parallel computing is likely to remain attractive only in those cases in which:
+parallel programming will only be cost-effective when:
 \begin{itemize}
-\item	the application is trivially parallel (e.g.\ task-farming a rendering calculation);
+\item	the application is trivially parallel (e.g., task-farming a rendering calculation);
 \item	the expected total program runtime is very large
-	(i.e.\ the program is a package used extensively in a field
+	(i.e., the program is a package used extensively in a field
 	such as computational chemistry or vehicle crash simulation);
 	or
-\item	cost is not an obstacle (i.e.\ the customer is the NSA).
+\item	cost is not an obstacle (i.e., the customer is Disney or the NSA).
 \end{itemize}
+
+The aim of this problem suite is to capture enough information about $D$
+to allow competing parallel programming systems to be ranked and compared.
 
@@
 
-In a 1981 survey article \cite{b:psych-prog-survey}, B.~A.\ Sheil observed that:
-\begin{quote}
-As practiced by computer science,
-the study of programming is an unholy mixture of mathematics,
-literary criticism,
-and folklore.
-\end{quote}
-Little has changed since.
-Computer scientists pay lip service to the importance of usability,
-but are strangely reluctant to try to quantify it,
-particularly with respect to the tools they use themselves.
-Religious debates about the ``right'' first-year teaching language
-or the ``best'' GUI toolkit are common;
-systematic use of the experimental method is not.
-
-Most of the experiments which have been done have concentrated on ``programming in the small'' issues,
-such as readability and code-pattern recognition
-\cite{b:reverse-parsing-2,b:cloze-program-1,b:prog-comprehension-1}.
-The little work that has been done on ``programming in the large'' has been anecdotal,
-such as \cite{b:compare-ada-c-pascal}.
-This book presented various authors' impressions of Pascal, C, and Ada
-(before any complete Ada implementations existed).
-Its editors formulated a list of questions to use in comparing languages,
-but did not use it themselves.
-A modified version of this questionnaire,
-tailored for parallel programming languages,
-is given in Appendix~\ref{s:questions},
-and will be used as a guideline in our work.
-
 Two comparative efforts in parallel programming are \cite{b:babb-cases} and \cite{b:salishan}.
 The first presented implementations of a simple numerical quadrature program
-in more than a dozen different parallel languages
-in use on mid-1980s hardware.
+in more than a dozen different parallel languages used on mid-1980s hardware.
 The second presented implementations of
@@
 and a simple discrete event simulator---in
-C* \cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
-and a variety of dataflow and functional languages.
-Both of these books convey the flavor of the languages they describe,
-but neither made any effort to compare languages or problem implementations directly.
+C*~\cite{b:dataparallel-c}, Occam~\cite{b:occam}, Ada~\cite{b:ada},
+and several dataflow and functional languages.
+Both books conveyed the flavor of the languages they describe,
+but neither compared languages or problem implementations directly.
 
-Inspired by the Salishan Problems,
-the first author developed a suite of seven problems \cite{b:cowichan-1}
-in order to test the limits of parallel programming systems.
-Individual problems were badly load-balanced,
-manipulated irregular pointer-based structures,
-and so on.
-Our experience implementing them was extremely valuable,
-as it uncovered many bugs in the programming system used,
-and revealed several shortcomings of the system's design.
-However,
-we also discovered that implementing each problem takes six to eight weeks,
-which we feel is prohibitive at this exploratory stage.
+More recently,
+Lutz has explored quantitative comparison of the relative productivity
+of different programming languages \cite{b:prechelt-languages,b:prechelt-platforms}.
+Our long-term aim is to replicate his work in the domain of parallel programming.

### 2009-08-31 paper/2009-09/memory.tex
More work reformatting and reorganizing paper

--- 2009-08-28 paper/2009-09/cliches.tex
@@
-\section{Parallel Clich\'{e}s\label{s:cliche}}
+\section{Memory Reference Patterns\label{s:memory}}
 
-\subsection{Operations\label{s:cliche-ops}}
-
-This list details some operations which are supported by many parallel programming systems.
-The toys described in Section~\ref{s:toys} provide opportunities for exercising many of these,
-and implementors are encouraged to phrase discussion of their work in terms of these operations where appropriate.
-
-\begin{itemize}
-\item	elementwise operations on arrays (unary, binary, and scalar promotion)
-\item	cumulative operations on arrays (reduction and parallel prefix)
-\item	array re-shaping and re-sizing (e.g.\ sub-array extraction)
-\item	partial (conditionally masked) versions of all of the above
-\item	regular data motion (e.g.\ circular and planar shifting)
-\item	irregular data motion (e.g.\ 1-to-1, 1-to-many, and many-to-1 permutation)
-\item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
-\item	differential local addressing (i.e.\ subscripting an array with an array)
-\item	full or partial replication of shared read-only values
-\item	full or partial replication of shared read-mostly values with automatic consistency management
-\item	structures with rotating ownership, suitable for migration
-\item	producer-consumer structures
-\item	partitionable structures
-\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
-\item	dynamic re-partitioning (e.g.\ for load balancing)
-\item	committed mutual exclusion (e.g.\ waiting on a semaphore)
-\item	uncommitted mutual exclusion (e.g.\ lock or fail)
-\item	barrier synchronization
-\item	multiple concurrent barriers used by non-overlapping groups of processes
-\item	fetch-and-add, and other fetch-and-operate functions
-\item	pre-scheduled receipt of messages of known composition
-\item	variable-length message receipt
-\item	message selection by type
-\item	message selection by source
-\item	message selection by contents
-\item	guarded message selection (e.g.\ CSP's {\tt{alt}} construct)
-\item	broadcast and partial broadcast
-\item	split-phase (non-blocking) operations
-\item	data marshalling and unmarshalling of ``flat'' structures (e.g.\ arrays of scalars)
-\item	data marshalling and unmarshalling of nested or linked structures
-\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
-\item	pipelining
-\item	distributed linked structures in systems without physically-shared memory
-\item	indexing of distributed shared arrays
-\item	collective I/O
-\item	uniform-size interleaved I/O
-\item	heterogeneous-size interleaved I/O
-\item	independent I/O operations on a single shared file
-\end{itemize}
-
-\subsection{Memory Reference Patterns\label{s:cliche-memref}}
-
-These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
+These are inspired by the categorization originally used in the Munin system \cite{b:munin-overview}.
 Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.

### 2009-08-31 paper/2009-09/method.tex
More work reformatting and reorganizing paper

@@
 
-The main aim of this work is to assess and compare
-the ease with which programs can be developed and/or ported
-using different parallel programming systems.
+Our aim is to assess and compare the usability of parallel programming systems.
 Here,
-we discuss issues related to choosing the problems that make up the suite,
-assessing the complexity of particular implementations,
+we discuss issues related to choosing the problems that make up our benchmark suite,
+how to assess particular implementations,
 and other purposes which implementations of these might serve.
-Recognizing that these problems are extremely simple,
-and easy to parallelize,
+Recognizing that our problems are extremely simple,
 we refer to them as ``toys''.
@@
 
-Our criteria for including problems in this suite are as follows:
+Our criteria for including toys are:
+
 \begin{enumerate}
-\item	Each toy should require no more than an afternoon
+
+\item	Each toy should require no more than an hour
 	to write and test in a well-supported sequential language.
-	We feel that making individual problems more complicated will only discourage implementation.
-\item	The correctness of each implementation must be relatively easy to verify.
-	Toys whose output is easily visualized are therefore to be preferred,
-	as are toys whose results are insensitive to floating-point discretization effects.
-\item	Each toy application should be amenable to
-	an analysis of speedup or isoefficiency \cite{b:isoefficiency-tutorial}
-	to allow a comparison of theory and practice.
+	We feel that making individual problems more complicated will only discourage uptake.
+
+\item	Correctness must be easy to verify.
+	Toys whose output is easily visualized are therefore preferred,
+	as are toys whose results are insensitive to floating-point effects.
+
+\item	Speedup must similarly be easy to measure---while machine performance is not our focus,
+	we are generally uninterested in parallel implementations that are slower than
+	the sequential originals.
+
 \item	At least some toys should not be ``infinitely scalable''.
@@
 	and this suite should reflect such limitations.
+
 \item	At least some toys should require I/O,
 	since this important aspect of real-world programming is often neglected by PPS designers.
-\item	There should be some potential redundancy in the type of parallelization done,
-	so that re-use of software can be demonstrated.
+
+\item	There should be some overlap in the toys' implementations,
+	so that implementors can demonstrate how well
+	their chosen systems take advantage of opportunities for code re-use.
+
 \item	Together,
-	the toys in the suite must exercise a wide range of common parallel operations
-	(Appendix~\ref{s:cliche-ops})
-	and memory reference patterns
-	(Appendix~\ref{s:cliche-memref}).
-\item	The ``obvious'' implementations of these toys should span
-	a wide range of parallel programming paradigms,
-	such as task farming,
-	geometric and functional decomposition,
-	and speculative parallelism.
-\item	Parallelizations for several (if not all) of
-	the types of programming systems shown in Figure~\ref{f:taxonomy} should be known.
-	This suite is not intended to be a collection of research problems;
-	reasonably simple and efficient implementations in most environments
-	should be straightforward to produce.
+	the toys in the suite must exercise a wide range of parallel clich\'{e}s
+	(Appendix~\ref{s:ops} and ~\ref{s:memory}).
+	In particular,
+
+\item	Toys should be specified by inputs and outputs rather than algorithmically,
+	i.e.,
+	``sort $N$ integers'' rather than ``parallelize quicksort'',
+	so that implementors can choose algorithms that are ``natural'' for their systems.
+	Implementations that parallelize a grossly inefficient sequential algorithm
+	should expect to be criticized for doing so;
+	see Section~\ref{s:toys-invperc} for more discussion of this.
+
 \end{enumerate}
-
-\begin{figure}
-% \epsfysize=0.5\textheight
-% \begin{center}\mbox{\epsffile{fig/taxonomy.eps}}\end{center}
-\caption{A Taxonomy of Parallel Programming Systems\label{f:taxonomy}}
-\end{figure}
-
-\subsection{Specification\label{s:method-spec}}
-
-Toys should be specified in terms of inputs and outputs,
-rather than algorithmically.
-For example,
-``sort $N$ integers in the range $1{\ldots}R$,
-where $R$ is not known in advance and there may be multiple instances of a given value''
-is to be preferred over
-``parallelize quicksort''.
-As \cite{b:slalom} argued,
-which parallel algorithm is best in a particular programming system
-depends critically on the architectural assumptions made by that system.
-Implementors should be free to employ those methods which they find natural and convenient.
-The only exception we make to this rule is in the two matrix equation solvers,
-where we require Gaussian elimination and successive over-relaxation.
-Given the enormous number of matrix solution algorithms in use,
-we felt we had to specify these particular algorithms in order to ensure that
-different implementations of this suite would be comparable.
 
@@
 
-These toys will exercise many different aspects of parallel systems.
-However,
-their ``single algorithm per program'' model is not representative of real applications.
-These usually contain several discrete (and sometimes overlapping) phases,
-each of which is qualitatively different.
+The ``single algorithm per program'' model of many benchmarks is not representative of real programs,
+which often contain several qualitatively different phases or modules.
 A full implementation of this suite will therefore have two parts.
@@
 In the second,
-toys will be chained together to create the kind of phase behaviour seen in real applications
-(Figure~\ref{f:chaining}).
+toys will be chained together as shown in Figure~\ref{f:chaining}.
 This will test the ease with which heterogeneous parallelism can be mixed within a single program.
@@
 
-A second reason for chaining toys is that
-many groups are now using algorithmic skeletons to structure parallel programs
-\cite{b:cole-skel,b:p3l-overview,b:enterprise}.
-Such skeletons encapsulate the details of a particular style of parallelism,
-and hide low-level or platform-dependent issues.
-We hope that chaining will show the strengths and weaknesses of such systems.
-
-Finally,
-chaining should be designed so that some toys can be executed concurrently.
-Many PPSs impose extraneous constraints on programs,
-e.g.\ require all processes to participate in every barrier synchronization,
+Where possible,
+chaining should execute toys concurrently.
+Some parallel programming systems impose extraneous constraints on programs,
+e.g., require all processes to participate in every barrier synchronization,
 or require the same executable to be loaded onto each processor.
@@
 the way in which the sizes of problems are determined.
-In a completely frozen model, the actual size of each problem would be compiled into each toy.
-A fluid implementation, by contrast, would allow sizes to be specified at run-time,
-and would only then allocate and partition data structures.
-We have chosen to use an intermediate model (which might be called ``slushy''),
+In a \emph{frozen} model,
+the actual size of each problem and/or the number of processors available would be compiled into each toy.
+A fully \emph{fluid} implementation, by contrast, would allow these sizes to be specified at run-time.
+If a system does not support fully-fluid implementations,
+it must at least use an intermediate model (which might be called \emph{slushy}),
 in which the maximum size of individual problems is specified during compilation,
 but the actual size of a problem is only determined when the toy begins to execute.
-We hope that the use of this model will force implementors to deal with
-at least some of the software engineering issues
-involved in building flexible large-scale applications,
-without making implementation prohibitively difficult.
 
@@
 
-There are at least two different ways in which usability might be compared.
+There are at least two different ways in which usability could be compared.
 One would be to measure the performance achieved by an ``average'' programmer
-as a function of time on each of a range of problems.
-This would lead to graphs of the kind shown in Figure~\ref{f:perf},
-where the left graph shows a particular result
-and the right graph a simplified abstraction of it.
+as a function of time on each toy.
 By analogy with Hockney's $n_{1/2}$ measure \cite{b:hockney-perfparam},
-which is the length of vector on which a pipelined architecture achieves
-half its theoretical peak performance,
 we could in principle find the value of $p_{1/2}$---the programming time required to achieve
@@
 for a particular combination of programming system and problem type.
+However,
+the number of implementors is likely to be too small for averages to be meaningful.
 
-\begin{figure}
-% \epsfxsize=12cm
-% \begin{center}\mbox{\epsffile{fig/perf-time.eps}}\end{center}
-\caption{Performance vs.\ Time\label{f:perf}}
-\end{figure}
+Alternatively,
+we could use code metrics
+to compare implementations of various toys.
+This is also dubious,
+as \cite{b:el-emam-metrics} has shown that code metrics do little more than measure lines of code.
 
-A second option would be to measure the complexities of
-implementations of a single application using different programming systems.
-With this approach,
-comparative performance or speedup figures would be supplemented by complexity measures,
-as shown in Figure~\ref{f:complex}.
-
-\begin{figure}
-% \epsfxsize=6cm
-% \begin{center}\mbox{\epsffile{fig/perf-complex.eps}}\end{center}
-\caption{Comparative Code Complexity\label{f:complex}}
-\end{figure}
-
-The drawback of the first approach is that there are no ``typical'' problems,
-or ``average'' programmers.
-Meaningful measurement of  $p_{1/2}$ would therefore require
-measurement of a large number of programmers working on many problems.
-Since this is too expensive to be practical,
-the second approach is the one we have chosen to adopt.
-It has drawbacks as well,
-most notably the fact that no-one has ever demonstrated any better measure of program complexity
-than a simple count of the number of lines in a program \cite{b:fenton-metrics}.
-Since we anticipate a wide variation in the types of progamming systems that will be used in this work
-(ranging from declarative dataflow languages
-through functional languages with Lisp-like syntax
-to parallel dialects of Fortran and C),
-and in the indentation, commenting, and layout styles of particular programmers,
-we may measure complexity by counting the number of tokens in the source code which are
-keywords,
-intrinsic operators,
-user-defined functions,
-type or variable declarations,
-and miscellaneous punctuation.
+We therefore intend to compare implementations through group consensus,
+e.g.,
+through interviews with implementors
+and by asking other programmers who are familiar with the base languages used
+to read and comment on the parallel versions of the toys.
+We will also compare the length of code in parallel and sequential implementations,
+though we realize that this can exaggerate the impact of parallelization.
 
@@
 
-This suite is intended for assessing PPS usability,
-but we envisage at least three other uses for implementations of it:
-\begin{description}
-\item[Related Research:]
-	We hope that the existence of this suite will indicate
-	to the developers of new parallel programming systems
-	what a mature tool should be able to support.
-	If sufficiently diverse,
-	the suite could also be used to help test the correctness of new or improved systems.
-	Finally,
-	this suite could complement existing performance benchmarks,
-	since performance figures for individual toys could be compared.
-\item[Tool Development:]
-	Requests from tool developers for existing parallel programs,
-	or for event traces from them,
-	are common on newsgroups such as {\tt{comp.parallel}}.
-	We hope that traces from implementations of this suite
-	will be used to compare different debugging and performance monitoring tools,
-	that their source code will be used to test
-	new parallelizing compilation techniques or mapping and scheduling algorithms,
-	and so on.
-\item[Teaching:]
-	Implementations will be suitable for use as classroom examples,
-	since they will be small enough to be understood quickly.
-	The toys themselves should also be suitable as classroom exercises
+This suite is intended for assessing the usability of parallel programming systems,
+but we envisage other uses.
+First, this suite will indicate
+what a mature parallel programming system should be able to support.
+In particular,
+we will ask implementors to describe how they debugged correctness and performance problems,
+and in particular what tools they used.
+(The lack of useful debugging tools is a chronic complaint among parallel programmers.)
+
+This suite should also be useful for education,
+since toys will be small enough to be understood quickly,
+and parallelizing them should be a suitable classroom exercise
 in a senior undergraduate course on parallel computing.
-\end{description}
 
@@
 over-specified (e.g.\ the Livermore Loops~\cite{b:livermore-loops}),
-or concerned with a limited range of applications
-(e.g.\ the NAS Parallel Benchmarks~\cite{b:nas-results}).
+or concerned with a limited range of applications.
 Finally,
 most exist only in C or Fortran (or both);
-many interesting parallel programming systems are built on top of other languages,
-including Modula-2, Scheme, Prolog, and various dataflow languages.
+many interesting parallel programming systems are built on top of other languages.
 We do not want the effort required to translate several large programs
@@
 
-\vspace{\baselineskip}
-
 \noindent {\em{Won't the results depend primarily on programmer ability?}\/}
 
-The short answer to this question is yes,
+Yes,
 but no more than the results from performance benchmarks.
-A longer answer is that	since we intend to measure the combination of code complexity and performance achieved,
+A longer answer is that
+since we intend to measure the combination of code complexity and performance achieved,
 rather than performance as a function of development time,
 programmers will be allowed to revise and improve their implementations when and as desired.
+We hope this will encourage the evolutionary improvement seen in some online programming contests,
+e.g., \cite{b:gulley-contest}.

### 2009-08-31 paper/2009-09/ops.tex
More work reformatting and reorganizing paper

--- 2009-08-28 paper/2009-09/cliches.tex
@@
-\section{Parallel Clich\'{e}s\label{s:cliche}}
-
-\subsection{Operations\label{s:cliche-ops}}
+\section{Parallel Operations\label{s:ops}}
 
@@
 \end{itemize}
-
-\subsection{Memory Reference Patterns\label{s:cliche-memref}}
-
-These are inspired by the categorization used in the Munin system \cite{b:munin-overview}.
-Again, implementors are encouraged to phrase discussion of their work in these terms where appropriate.
-
-\begin{itemize}
-\item	Write-once:
-	variables which are assigned a value once, and only read thereafter.
-	These can be supported through replication.
-\item	Write-many:
-	variables which are written and read many times.
-	If several processes write to the variable concurrently, they typically write to different portions of it.
-\item	Producer-consumer:
-	variables which are written to by one or more objects, and read by one or more other objects.
-	Entries in a shared queue, or a mailbox used for communication between two processes, are examples.
-\item	Private:
-	variables which are potentially shared, but actually private.
-	The interior points of a mesh in a program which uses geometric decomposition fall into this category,
-	while boundary points belong to the previous class.
-\item	Migratory:
-	variables which are read and written many times in succession by a single process before ownership passes to another process.
-	The structures representing particles in an $N$-body simulation are the clearest example of this category.
-\item	Result:
-	Accumulators which are updated many times by different processes, but thereafter read.
-\item	Read-mostly:
-	Any variable which is read much more often than it is written to.
-	The best score so far in a search problem is a typical instance of this class:
-	while any process might update it, in practice processes read its value much more often than they change it.
-\item	Synchronization:
-	Variables used to force explicit synchronization in a program, such as semaphores and locks.
-	These are typically ignored for long periods, and then the subject of intense bursts of access activity.
-\item	General read-write:
-	Any variable which cannot be put in one of the above categories.
-\end{itemize}

### 2009-08-31 paper/2009-09/prob-invperc.tex
More work reformatting and reorganizing paper

@@
 (such as water)
-in fractured rock \cite{b:percolation-theory,b:invasion-percolation}.
+in fractured rock \cite{b:percolation-theory}.
 In two dimensions,

### 2009-09-01 paper/2009-09/issues.tex
More tidying up on paper

@@
 
-The chained implementation of these problems
-executes individual toys in the following order:
+The chained implementation of these toys executes in the order shown below.
+Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
+there are four possible chained sequences.
+
 \begin{enumerate}
+
 \item	An integer matrix $I$ with $r$ rows and $c$ columns is created
-	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel}) or by filling locations with random values (Section~\ref{s:toys-rng}).
+	either by using the Mandelbrot Set algorithm (Section~\ref{s:toys-mandel})
+	or by filling locations with random values (Section~\ref{s:toys-rng}).
 	\label{e:chain-choice-1}
+
 \item	The integer matrix $I$ is shuffled in both dimensions (Section~\ref{s:toys-half}).
+
 \item	Either invasion percolation (Section~\ref{s:toys-invperc}) or histogram thresholding (Section~\ref{s:toys-thresh}) is used
@@
 	\label{e:chain-choice-2}
+
 \item	The Game of Life (Section~\ref{s:toys-life}) is simulated for $G$ generations, using $B$ as an initial configuration.
 	This step overwrites the Boolean matrix $B$.
-\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ as described in Section~\ref{s:toys-winnow}.
-\item	The elastic net algorithm is simulated using the points created in the previous step as cities.
-\item	The final locations of the net points from the prevoius step are normalized, as described in Section~\ref{s:toys-norm}.
-\item	An $L{\times}L$ matrix $A$ and an $L$ vector $V$ are created using the normalized point locations from the previous step,
-	as described in Section~\ref{s:toys-outer}.
+
+\item	A vector of $L$ $(m,x,y)$ points is created using the integer matrix $I$ and the Boolean matrix $B$ (Section~\ref{s:toys-winnow}).
+
+\item	A series of convex hulls are obtained from the points produced above (Section~\ref{s:toys-convex}).
+
+\item	The final locations of the net points from the previous step are normalized (Section~\ref{s:toys-norm}).
+
+\item	An $L{\times}L$ matrix $A$ and an $L$-vector $V$ are created using the normalized point locations from the previous step
+	(Section~\ref{s:toys-outer}).
+
 \item	The matrix equation $AX=V$ is solved using Gaussian elimination and using successive over-relaxation
 	to generate two solution vectors $X_{gauss}$ and $X_{sor}$.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The checking vectors $V_{gauss} = A X_{gauss}$ and $V_{sor} = A X_{sor}$ are calculated.
-	In a parallel system, these two problems should if possible be executed concurrently.
-\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$, i.e.\ the greatest element-wise absolute difference, is calculated.
+	These two toys should execute concurrently if possible.
+
+\item	The checking vectors $V_{gauss}=AX_{gauss}$ and $V_{sor}=AX_{sor}$ are calculated.
+	These two toys should execute concurrently if possible.
+
+\item	The norm-1 distance between $V_{gauss}$ and $V_{sor}$ (i.e., the greatest element-wise absolute difference) is calculated.
 	This measures the agreement between the solutions found by the two methods.
+
 \end{enumerate}
-Because of choices in steps~\ref{e:chain-choice-1} and~\ref{e:chain-choice-2},
-there are four possible chained sequences.
-A minimal implementation will provide one of these,
-while a full implementation will be a single program which can execute any of the four sequences.
-Intermediate implementations providing all four sequences as separate programs are also acceptable.
 
@@
 
-The file formats used in the Cowichan problems are shown in Figure~\ref{f:io-formats}.
-A file containing a vector begins with 1 positive integer,
+The file formats used in the Cowichan Problems are shown in Figure~\ref{f:io-formats}.
+A file containing a vector begins with a single positive integer,
 which specifies the number of elements in the vector;
-a file containing a matrix begins with 2 positive integers,
-which specify the number of rows and columns in the matrix respectively.
+a file containing a matrix begins with a pair of positive integers,
+which specify the number of rows and columns in the matrix respectively\footnote{Note
+that this means the first number is the Y extent, and the second number is the X extent.}.
 Elements of the vector or matrix then appear one per line,
@@
 The formats described above are required to be human-readable.
-Implementations may also include I/O using ``raw'' files, i.e.\ binary data dumps rather than formatted ASCII,
+Implementations may also include I/O using ``raw'' files,
+i.e., binary data dumps rather than formatted ASCII,
 using whatever file formats are convenient.
@@
 \end{figure}
-
-\subsection{Visualization\label{s:issues-vis}}
-
-Visualization is an increasingly important part of scientific and numerical computing.
-Including it as an optional part of this problem set is also a good way to exercise both
-the use of libraries within a parallel programming system,
-and the ease with which a programming system can support variant versions of a program.
-
-Our requirements for the graphics interface are:
-\begin{enumerate}
-\item	Visualization for individual toys should be as simple as possible,
-	and should require as few changes to the basic toy code as possible.
-\item	The same code should be used for both the stand-alone versions of the toys and for the chained versions.
-	The latter should create all images in a single window.
-\item	A single source file should implement both the graphical and non-graphical versions of each toy.
-\end{enumerate}
-The first two points require the system to distinguish between the actual drawing code
-(e.g.\ drawing a vector of points as circles)
-and the place where the drawing has to be done in the window (called the {\em{viewport}\/}).
-While the stand-alone toys use a whole window for drawing,
-the chained versions may only draw in some portion of the window, e.g.\ the upper left corner.
-At present,
-this is implemented by using a call counter
-in the interfaces for the individual toys and in the chained module.
-The actual drawing routines called from within each toy could then be the same.
-
-The present visualization routines were originally written by Anil Sukul,
-a student at the Vrije Universiteit, Amsterdam.
-The package uses the public domain graphics library VOGLE,
-available by anonymous FTP from {\tt{gondwana.ecr.mu.oz.au::/pub}}.
-VOGLE includes functions for drawing points, lines, polygons, and text in a device-independent fashion, 
-as well as screen control routines (such as double buffering).
-VOGLE was chosen because of its simplicity,
-and because it includes Pascal and Fortran interfaces.
 

### 2009-09-01 paper/2009-09/method.tex
More tidying up on paper

@@
 
-\subsection{Assessing Complexity\label{s:method-complex}}
+\subsection{Assessing Usability\label{s:method-assess}}
 
-There are at least two different ways in which usability could be compared.
-One would be to measure the performance achieved by an ``average'' programmer
+One way to compare the usability of parallel programming systems would be
+to measure the performance achieved by an ``average'' programmer
 as a function of time on each toy.
@@
 However,
-the number of implementors is likely to be too small for averages to be meaningful.
+the number of implementors required is much greater than the number we can reasonably expect.
 
@@
 This is also dubious,
-as \cite{b:el-emam-metrics} has shown that code metrics do little more than measure lines of code.
+both because \cite{b:el-emam-metrics} has shown that most metrics do little more than measure lines of code,
+and because of the difficulty of comparing metric values between languages.
 
-We therefore intend to compare implementations through group consensus,
+The best we can hope for now is therefore qualitative consensus,
 e.g.,
-through interviews with implementors
-and by asking other programmers who are familiar with the base languages used
+to interview implementors about their experiences
+and ask other programmers who are familiar with the base languages used
 to read and comment on the parallel versions of the toys.

### 2009-09-01 paper/2009-09/ops.tex
More tidying up on paper

@@
 \item	scalar and non-scalar value location (e.g.\ finding a scalar or record value in an array)
-\item	differential local addressing (i.e.\ subscripting an array with an array)
+\item	differential local addressing (i.e., subscripting an array with an array)
 \item	full or partial replication of shared read-only values
@@
 \item	partitionable structures
-\item	pre-determined run-time re-partitioning (i.e.\ re-distributing an array)
+\item	pre-determined run-time re-partitioning (i.e., re-distributing an array)
 \item	dynamic re-partitioning (e.g.\ for load balancing)
@@
 \item	data marshalling and unmarshalling of nested or linked structures
-\item	heterogeneous parallelism (i.e.\ running different applications concurrently in one machine)
+\item	heterogeneous parallelism (i.e., running different applications concurrently in one machine)
 \item	pipelining

### 2009-09-01 paper/2009-09/prob-invperc.tex
More tidying up on paper

@@
 chooses the one with the lowest value
-(i.e.\ the one with the least resistance to filling),
+(i.e., the one with the least resistance to filling),
 and fills it in.
